### 2016-09-06 paper.tex
Structure document

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[10pt, conference, compsocconf]{IEEEtran}
% Add the compsocconf option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Everthing you wanted to know about PanDA but were truly truly afraid to ask}

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations

\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
line 2: name of organization, acronyms acceptable\\
line 3: City, Country\\
line 4: Email: name@xyz.com}
\and
\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
line 2: name of organization, acronyms acceptable\\
line 3: City, Country\\
line 4: Email: name@xyz.com}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
The abstract goes here. DO NOT USE SPECIAL CHARACTERS, SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.
\end{abstract}

\begin{IEEEkeywords}
TBD.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

% ----------------------------------------------------------------------------
% I
% ----------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}

\begin{itemize}
  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
  \item Why are WMS needed?
  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
  \item \ldots
  \item Overview/Summary of paper.
\end{itemize}

% ----------------------------------------------------------------------------
% II
% ----------------------------------------------------------------------------
\section{Background}
\label{sec:background}

\begin{itemize}
  \item Brief overview of distributed CI.
  \item Workload vs Workflow
  \item Distinguish between WFMS and WLMS!!!!
  \begin{itemize}
    \item Role of Workload MS: help “hide” and federate heterogeneous resources.
  \end{itemize}
\end{itemize}

% ----------------------------------------------------------------------------
% III
% ----------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}

\begin{itemize}
  \item Other workload management systems (WMS) (not exclusive)
  \begin{itemize}
    \item glidein-WMS
    \item Dirac
    \item ALICE-AlieN
  \end{itemize}
  \item Comparison and contrast
  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
\end{itemize}

% ----------------------------------------------------------------------------
% IV
% ----------------------------------------------------------------------------
\section{PanDA Overview}
\label{sec:panda_overview}

\begin{itemize}
  \item PanDA WMS
  \begin{itemize}
    \item State Diagrams
    \item Control Flow Diagrams
    \item Software Module and Architecture diagrams.
  \end{itemize}
  \item PANDA Pilot
\end{itemize}

% ----------------------------------------------------------------------------
% V
% ----------------------------------------------------------------------------
\section{Deploying PanDA on a Leadership-scale system}
\label{sec:panda_deployment}

\begin{itemize}
  \item Challenges
  \item Opportunities
  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
  \item Take-aways, lessons learned
\end{itemize}

% ----------------------------------------------------------------------------
% VI
% ----------------------------------------------------------------------------
\section{Performance Characterization}
\label{sec:panda_performance}

\begin{itemize}
  \item Internal PanDA performance
  \item External PanDA performance (workload specific on underlying system/platform)
\end{itemize}

% ----------------------------------------------------------------------------
% VII
% ----------------------------------------------------------------------------
\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
\label{sec:panda_titan}

\begin{itemize}
  \item Metadata performance issue, and how it’s resolved
  \item File I/O performance and impact
  \item ?
\end{itemize}

% ----------------------------------------------------------------------------
% VIII
% ----------------------------------------------------------------------------
\section{PANDA: Future/RoadMap}
\label{sec:panda_roadmap}

% section panda_future_roadmap (end)

% \subsubsection{Subsubsection Heading Here}
% Subsubsection text here.

% \section{Type style and Fonts}
% Wherever Times is specified, Times Roman or Times New Roman may be used. If
% neither is available on your system, please use the font closest in
% appearance to Times. Avoid using bit-mapped fonts if possible. True-Type 1
% or Open Type fonts are preferred. Please embed symbol fonts, as well, for
% math, etc.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex,
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

% \section{Conclusion}
% The conclusion goes here. this is more of the conclusion

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank... More thanks here...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% that's all folks
\end{document}

### 2016-09-13 WMSLiterature.tex
First draft - WMS Literature

\subsection{Alien}
Alien is a Workload and Data Management system composed of a 	set of middleware tools and services entirely based on web-services and standard protocols. Alien was originally developed for the ALICE experiment \cite{} but subsequently used by several virtual organizations \cite{}. 
%The system has been deployed in 2001 for distributed producti	on of Monte Carlo data, detector simulation and reconstruction.
Alien is composed of two type of services:
\begin{itemize}
\item
\emph{Central services}, these services are unique for each virtual organisation, therefore there is only one  configuration point for the management;
\item \emph{Site services}, they provide the interfacing to local resources and Grid services running on a VO-box;
\end{itemize}

The most important central services are:
\begin{itemize}
\item \emph{Job Manager}, a database that keeps track of all the jobs submitted to the system and their current execution status;
\item \emph{Brokers}, they are the core of task executions and data transfers; they receive tasks in form of JDL,  keep them ordered by priority and send them to the CE for execution;
\item \emph{Optimizers}, they are used to minimize the work of the Broker by scanning periodially the task queue and re-arranging the tasks in such a way that fairness and priority policies are guaranted;
\item \emph{Data Catalogue}, it keeps track of the scripts and files uploaded on Storage Elements.
\end{itemize}

%The Computing Agents are instead site services that monitor the local Computing Element, advertise site's capabilities and are responsible for submitting the JobAgents.

%Information about the status of the sites and central services, full job statistics and monitoring information are kept in a MonALISA repository.
%% CLUSTER MONITOR SHOULD BE EQUAL TO COMPUTING AGENT
%% Job Manager should be equal to TaskQueue
%% Process Monitor == PIlot????

The job execution in Alien is usually distributed over several sites. Each of these sites has at least one service called ClusterMonitor. On one side the Cluster Monitor is used to communicate with Central services (Job Manager and Broker), on the other side it can manage Computing Elements (CE) by starting and stopping them whenever it receives the signal.
The CE is the resource in charge of the execution of the jobs. A CE usually is associated with a batch queue and can send the jobs to the worker nodes controlled by the queue.

The CE asks the Broker for jobs to execute by sending its JDL. Once received the JDL, the Broker will try to match it with the JDL of the jobs in queue. If a match exists then the Broker sends the jobs JDL to the CE.
Immediately after receving a job JDL, the CE create a new service on the worker node called ProcessMonitor. This service allows the CE (and the rest of Alien services through the CE) to interact with the job while is running. 
This execution strategy is called ``pull mode'' due to the fact that CE asks for jobs.
It is worth to mention that more recent version of Alien implement exploit gLite and CE-CREAM. The latter allows the system to by-pass the broker during the job submission \cite{}.

Alien uses the Job Description Language (JDL) to allow users to describe their workload task by task;  i.e.,  users can specify features such as task priorities, the level of parallelism (one core, multi-core, MPI etc..) and  also the DCR that should be targeted for the execution.

%Job submission is implemented by following the so-called ``pull mode'' which is composed of the following steps: 
%\begin{enumerate}
% \item the VO-Box monitors the status of the site queues through polls to the resource running on the CE; 
%\item  the Job Broker receives a report everytime slots become available; 
%\item  if the Task Queue is not empty, the Job Broker asks the VO-box to submit a number of Agents;
%\item finally, the JobAgents are submitted  to the site Computing Element either by way of that sends them back to the site Computing Element or, wherever available, directly through the CREAM interface on the CE itself.
%\end{enumerate}

\subsection{DIRAC} 
DIRAC (Distributed Infrastructure with Remote Agent Control) Workload and Data Management System is a software product, developed within the CERN LHCb project, to manage the processing of detector data, Monte Carlo simulations, and end-user analyses. 
DIRAC  architecture relies on four entities:
\begin{itemize}
\item \emph{Clients}: consist in a set of APIs that allows users to submit job requests. Clients interact directly with DIRAC central services.
\item \emph{Services}: serve Clients and Agents by performing crucial operations such as Job Management, Configuration, Bookkeping and Accounting.
\item \emph{Agents}: perform repetitive tasks like querying file catalogs,  monitoring of jobs on resources.
\item \emph{Resources}: they can be PC's, site cluters and Grids. Agents interact with them without distinction.
\end{itemize}
In the same way of Alien, DIRAC implements a pull scheduling. Furthermore DIRAC was the first WMS to exploit the concept of Pilot Agent on the Grid. 
Pilots Agents are gLite jobs that are submitted to the grid when jobs arrive into the WMS. 
DIRAC pilot system has four main logical components:
\begin{itemize}
\item a set of TaskQueues that collect tasks submitted by users, multiple TaskQueue being created depending on the requirements and ownership of the tasks;
\item a set of JobWrappers that are executed on the DCR to bind compute resources and execute tasks submitted by the users;
\item a set of TaskQueueDirectors that submits JobWrappers to target DCRs;
\item a MatchMaker that matches requests from JobWrappers to suitable tasks into TaskQeues.
\end{itemize}
The DIRAC execution model can be summarized in five
steps: 1. a user submits its workload in form of tasks to the WMS Job Manager; 2. submitted tasks are validated and added to a new or an existing TaskQueue, depending on the task properties; 3. TaskQueueDirector evaluates TaskQueues and a suitable number of JobWrappers are submitted to available
DCRs; 4. JobWrappers get instantiated on the DCRs and, then,  ask for tasks to the MatchMaker; 5. JobWrappers execute tasks while JobWrapper’s Watchdog monitor them.
TaskQueueDirectors deploy Pilots by getting a list of TaskQueues and calculating the number of pilot to submit  according to user priorities.
Once deployed on the compute resource, Pilots, a.k.a. JobWrappers, hold the resource in the form of single or multiple cores, spanning portions, whole, or multiple compute nodes. Pilots do not expose data capabilities although the system allows the user to perform both data staging and data replication. 
TaskQueues, TaskQueueDirectors, and the MatchMaker are implemented as services whereas the JobWrapper is implemented within the Agents together with the WatchDog.

\subsection{HTCondor Glidein and GlideinWMS}
The HTCondor Glidein system  as part of
the HTCondor [153] software ecosystem. The HTCondor
Glidein is a pilot-based system to aggregate DCRs with
heterogeneous middleware into HTCondor resource pools.
The logical components of HTCondor relevant to the
Glidein system are: 
\begin{itemize}
\item \emph{Schedd}, a daemon that implement a queuing system that holds workload tasks;
\item \emph{Startd}, a  daemon to control the DCR resources. 
\item \emph{Collector}, holds references to all the active
Schedd/Startd daemons; 
\item \emph{Negotiator} matches tasks queued in a Schedd to resources handled by a Startd.
\end{itemize}
In
HTCondor Glidein has been complemented by Glidein-WMS, a Glidein-based workload management system thatautomates deployment and management of Glideins on multiple types of DCR middleware. GlideinWMS builds upon the
HTCondor Glidein system by adding the following logical components: 
\begin{itemize}
\item \emph{Glidein Factories} that submit tasks to the DCRs middleware;
\item a set of \emph{Virtual Organizations (VO) Frontend} daemons that match the tasks on one or more Schedd to the resource attributes;
\item a \emph{Collector} that holds references to all the active Glidein Factories and VO Frontend daemons. 
\end{itemize}

 The execution model of the HTCondor Glidein system can be summarized in nine steps: 1. the user submits a Glidein (i.e., a job) to a DCR batch scheduler; 2. once executed, this Glidein bootstraps a Startd daemon; 3. the Startd daemon advertises itself with the Collector; 4. the user submits the tasks of the workload to the Schedd daemon; 5. the Schedd advertises these tasks to the Collector; 6. the Negotiator matches the requirements of the tasks to the properties of one of the available Startd daemon (i.e., a Glidein); 7. the Negotiator communicates the match to the Schedd; 8. the Schedd submits the tasks to the Startd daemon indicated by
the Negotiator; 9. the task is executed.

By using GlideinWMS, the user does not have to submit Glidein directly but only tasks to Schedd. From there: 1. every Schedd advertises its tasks with the VO Frontend; 2. the VO Frontend matches the tasks’ requirements to the resource properties advertised by the WMS Connector; 3. the VO Frontend places requests for Glideins instantiation to the WMS Collector; 4. the WMS Collector contacts the appropriate Glidein Factory to execute the requested Glideins; 5. the requested Glideins become active on the DCRs; and 6. the Glideins advertise their availability to the (HTCondor) Collector. From there on the execution model is the same as described for the HTCondor Glidein Service.

The resources managed by a single Glidein (i.e., pilot) are limited to compute resources. Glideins may bind one or more cores, depending on the target DCRs. For example, heterogeneous HTCondor pools with resources for desktops, workstations, small campus clusters, and some larger clusters will run mostly single core Glideins. More specialized pools that hold, for example, only DCRs with HTC, Grid, or Cloud middleware may instantiate Glideins with a larger number of cores. Both HTCondor Glidein and GlideinWMS provide abstractions for file staging but pilots are not used to hold data or network resources.
The process of pilot deployment is the main difference between HTCondor Glidein and GlideinWMS. While the
HTCondor Glidein system requires users to submit the pilots to the DCRs, GlideinWMS automates and optimizes pilot provisioning. GlideinWMS attempts to maximize the throughput of task execution by continuously instantiating Glideins until the queues of the available Schedd are emptied. Once all the tasks have been executed, the remaining Glideins are terminated.
HTCondor Glidein and GlideWMS expose the interfaces of HTCondor to the application layer and no theoretical
limitations are posed on the type and complexity of the workloads that can be executed. For example, DAGMan
(Directed Acyclic Graph Manager) has been designed to execute workflows by submitting tasks to Schedd, and a tool is available to design applications based on the master-worker coordination pattern.

Both HTCondor Glidein and GlideWMS rely on one or more HTCondor Collectors to match task requirements and resource properties, represented as ClassAds. This
matching can be evaluated right before the scheduling of the task. In this way, late binding is achieved but early binding remains unsupported.

### 2016-09-13 paper.tex
Add usual comments and headers
Itemize -> Enumerate
adding olcf boilerplate to acks
Second draft, plus integration of Sergey's part

@@
 % \documentclass[conference]{../sty/IEEEtran}
+
+\input{head}
+\input{include}
 
@@
 
+\newcommand{\sarp}[1]{{\color{green}\textit{sarp: {#1}}}}
+
 \begin{document}
@@
 \begin{abstract}
-The abstract goes here. DO NOT USE SPECIAL CHARACTERS, SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.
+ Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.
+Heterogeneous  re- sources are distributed worldwide, thousands of physicists analyz- ing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year.
+The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
+In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model.
+The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.
+PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
+2013.
+We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
 \end{abstract}
@@
 % ----------------------------------------------------------------------------
-\section{Introduction}
-\label{sec:intro}
+\section{Introduction}\label{sec:intro}
+The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades.
+Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers.
+Thousands of physicists analyze peta-scale  data volumes daily.
+One of  the LHC experiments  the ATLAS  [1]  utilizes the PanDA  workload management system [4] (WMS) for distributed data processing and analysis.
+The ATLAS Computing model [2] is based on a Grid paradigm [3], with multilevel, hierarchically distributed computing  and storage resources.
+Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
+PanDA has a highly scalable architecture.
+Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade.
+PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middle- ware).
+This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
+PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user anal- ysis, while actively evolving to meet rapidly changing science needs.
+Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figure/DailyJobs.png}
+\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
+\end{center}
+\end{figure}
 
-\begin{itemize}
+-------------------------------------------------------------------------------------------
+\begin{enumerate}
   \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
@@
   \item Overview/Summary of paper.
-\end{itemize}
+\end{enumerate}
 
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item Brief overview of distributed CI.
@@
   \item Distinguish between WFMS and WLMS!!!!
-  \begin{itemize}
+  \begin{enumerate}
     \item Role of Workload MS: help “hide” and federate heterogeneous resources.
-  \end{itemize}
-\end{itemize}
+  \end{enumerate}
+\end{enumerate}
 
@@
 \label{sec:related}
-
-\begin{itemize}
+\input{WMSLiterature}
+--------------------------------------------------------------------------------------------------
+\begin{enumerate}
   \item Other workload management systems (WMS) (not exclusive)
-  \begin{itemize}
+  \begin{enumerate}
     \item glidein-WMS
@@
     \item ALICE-AlieN
-  \end{itemize}
+  \end{enumerate}
   \item Comparison and contrast
   \item Core features of a general WMS, i.e., minimal complete model of a WMS.
-\end{itemize}
+\end{enumerate}
 
@@
 % ----------------------------------------------------------------------------
-\section{PanDA Overview}
-\label{sec:panda_overview}
+\section{PanDA Overview}\label{sec:panda_overview}
+PanDA consists of several interconnected subsystems, most of them are built from off-the-shelf and Open Source components. Figure \ref{fig:architecture} shows a schematic view of the PanDA system.
+In the following we will briefly describe system’s architecture and components.
 
-\begin{itemize}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figure/PandaArch.jpg}
+\caption{Schematic view of the PanDA system\label{fig:architecture}}
+\end{center}
+\end{figure}
+\subsection{PanDA Server}
+The PanDA server is the main component of the system, It provides  a task queue managing all job information centrally.
+The PanDA  server receives jobs through the client interface into the task queue, upon which a brokerage module  operates to prioritize  and assign work on the basis of job type, priority, input data and its locality, and available CPU resources.
+The PanDA  server operates as a web service.
+It runs on Apache web server, interacting with the back-end database running  on separate servers.
+It uses the Apache worker  model, with many independent   processes  handling client requests  in  parallel.
+Since all  state is maintained  in  the central database,  the PanDA Server application instance itself is stateless.
+Currently production version of PanDA utilizes Oracle database backend, but PanDA  can also work with MySQL family of databases.
+
+\subsection{PanDA Pilot}
+ PanDA is a pilot based workload  management system.
+In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload processing on a worker  node) are submitted to sites.
+When these  pilot jobs start on a  worker node they contact a central  server  to retrieve  a real payload (i.e., an end-user job) and execute it.
+Using these pilot-based  workflows helps to improve job reliability, optimize resource utilization, allows for opportunistic  resources usage, and mitigates  many of the problems  associated with the inhomogeneities found on the Grid.
+The pilots themselves  do not contain all the functionality needed to request a payload  job from the Panda Server.
+First, the complete  set of pilot code is downloaded via HTTP from a  central Subversion repository.
+The repository works with Apache web server configured  with a  memory-based   web proxy (Squid).
+The purpose  of the cache  is to reduce  the
+
+request load on the back-end Subversion  server.
+That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server.
+The benefits of this system include a high-performance code download service combined with code updates still being immediately  available  as soon as they are committed to source code control.
+
+\subsection{PanDA Factory}
+As a pilot-based  system, PanDA  requires some way to get the initial PanDA pilot onto worker nodes at sites.
+This is done with the help of a component  called AutoPyFactory (APF).
+APF runs in a single daemonized process, launching  a separate thread for each internal  workflow.
+Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote.
+The behavior of these APF workflows is determined by the combination of a set  of plug-ins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.
+
+\subsection{PanDA Monitoring}
+The PanDA Monitor is a web-based dashboard-style  graphi- cal application.
+It runs on Apache and interacts with the back- end database for persistence.
+The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory.
+The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.
+
+\section{Evolution of PanDA WMS}
+While PanDa  currently uses  more than 100,000  cores at well over 100 Grid sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will  require more resources  than Grid computing can possibly provide.
+The Worldwide LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned analysis and data processing, but it will be insufficient  for Monte Carlo (MC) production and any extra activities.
+Additional computing and storage  resources are therefore required.
+To alleviate  these challenges, ATLAS is engaged  in an ambitious  program to expand  the current computing model to include  additional  resources such as the opportunistic   use of supercomputers   as well as commercial and academic clouds.
+In turn this activity drives evolution of the PanDA WMS.
+
+\subsection{Use of Supercomputers with PanDA}
+The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers.
+Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network.
+The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors.
+On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI).
+While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization.
+It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run.
+What matters is the total number of cycles that can be offloaded from the Grid.
+Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc.
+A reorganization  of the standard  workflow is therefore needed.
+The most suitable task for HPCs in HEP is event generation.
+Event generators are mostly computational,  stand-alone code, with few input requirements.
+There is no need  to stage-in much data.
+The event generation  in ATLAS corresponds  to
+10-15\% of all jobs on the Grid.
+A more difficult workload to adapt  for HPCs is detector  simulation.
+Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.
+However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
+\subsection{Interfacing PanDA with Titan}
+The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US.
+Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
+Integration with Titan is the current focus for PanDA develop- ers.
+The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface [6].
+This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world.
+The pilot can communicate with the PanDA server from the front-end machine.
+The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job.
+ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
+1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design.
+Time allocation on an HPC is competitive and large projects are often preferred.
+About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources.
+The worker nodes
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figure/PandaInterfaceWithHPC.png}
+\caption{Schematic view of the PanDA system\label{fig:architecture}}
+\end{center}
+\end{figure}
+
+-------------------------------------------------------------------------------------
+\begin{enumerate}
   \item PanDA WMS
-  \begin{itemize}
+  \begin{enumerate}
     \item State Diagrams
@@
     \item Software Module and Architecture diagrams.
-  \end{itemize}
+  \end{enumerate}
   \item PANDA Pilot
-\end{itemize}
+\end{enumerate}
 
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item Challenges
@@
   \item Take-aways, lessons learned
-\end{itemize}
+\end{enumerate}
 
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item Internal PanDA performance
   \item External PanDA performance (workload specific on underlying system/platform)
-\end{itemize}
+\end{enumerate}
 
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item Metadata performance issue, and how it’s resolved
@@
   \item ?
-\end{itemize}
+\end{enumerate}
 
@@
 
-% section panda_future_roadmap (end)
+ \section{Conclusion}\label{sec:conclusion}
+The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.  In the process,  the old batch job paradigm  of
+computing in HEP was  discarded  in favor  of a  far more flexible and scalable  model.
+The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments.
+PanDA  is the first exascale  workload manage- ment system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013.
+Next LHC run will pose massive computing  challenges.
+With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more.
+Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS.
+To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system.
+Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry.
+The technology base provided by the PanDA system will enhance the usage of a variety  of high- performance computing resources available to basic research.
 
-% \subsubsection{Subsubsection Heading Here}
-% Subsubsection text here.
+\input{ack}
 
-% \section{Type style and Fonts}
-% Wherever Times is specified, Times Roman or Times New Roman may be used. If
-% neither is available on your system, please use the font closest in
-% appearance to Times. Avoid using bit-mapped fonts if possible. True-Type 1
-% or Open Type fonts are preferred. Please embed symbol fonts, as well, for
-% math, etc.
+%\section*{Acknowledgment}
 
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex,
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation Results}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
-%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that IEEE does not put floats in the very first column - or typically
-% anywhere on the first page for that matter. Also, in-text middle ("here")
-% positioning is not used. Most IEEE journals/conferences use top floats
-% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
-% command of the stfloats package.
-
-% \section{Conclusion}
-% The conclusion goes here. this is more of the conclusion
-
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgement
-\section*{Acknowledgment}
-
-The authors would like to thank... More thanks here...
+%The authors would like to thank... More thanks here...
 

### 2016-09-19 paper.tex
Changed the path of the figures
First draft workflow/workload and related definitions. Ongoing.

--- 2016-09-06 paper.tex
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item Brief overview of distributed CI.
-  \item Workload vs Workflow
-  \item Distinguish between WFMS and WLMS!!!!
-  \begin{itemize}
-    \item Role of Workload MS: help “hide” and federate heterogeneous resources.
-  \end{itemize}
-\end{itemize}
+  \item Workload vs Workflow. (DONE)
+  \item Distinguish between WFMS and WLMS.
+  \begin{enumerate}
+    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
+  \end{enumerate}
+\end{enumerate}
+
+\paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
+disciplines with different meaning. In the field of scientific computing,
+``workflow'' assumes different meanings depending on the characteristics of
+the computation, of the software tools used to support this computation, and
+of the resources on which it is performed. Further, a workflow may indicate a
+whole application, a description of the computational process of that
+application or, more commonly, a series of tasks related by data
+dependences.
+
+The lack of a consistent and shared definition of ``workflow'' hinders the
+understanding of its properties and its relations with related concepts. For
+example, we need to clarify the difference between ``workflow'', ``workload'',
+``task'', or ``job'' but also between workflow template and instance, or
+data-flow and control-flow. This clarification is precondition to specify
+properly the design of software systems that support the execution of
+scientific applications.
+
+In this paper, a workflow is a set of tasks alongside a description of their
+properties and relations. Each task may have an arbitrary number of properties
+like number of cores, executables, or input/output files. Tasks may have
+precedence relations depending on their data dependences or any other type of
+dependency mandated by the application algorithms. Tasks with precedence
+relations have to be executed serially, otherwise tasks can be executed
+concurrently. A workload is defined as a set of tasks that can be executed
+concurrently. As such, a workflow can be composed by a set of workloads.
+
+The terms ``task'' and ``job'' are also used inconsistently across communities
+that perform scientific computing. In this paper, a task is a \ldots
+\mtnote{Are we going to use task and job as defined in Jedi and Panda Server?}
+
+Usually, workflows are represented as graphs where tasks are vertices and
+their relations edges~\cite{}. Often, graphs are supposed to be acyclic but
+graphs with cycles have been used to represent workflows of workflows. In this
+paper, a workflow template is a type of graph while a workflow instance is a
+type of graph with a specific set of vertices and edges. Further, a graph is
+``abstract'' when no resource properties are available for all vertices,
+``concrete'' otherwise~\cite{}.
+
+The distinction between an abstract and concrete graph highlights an important
+element of the execution process of a workflow. Workflows can be specified
+without indicating on what resources they should be executed. Accordingly, an
+abstract workflow can be interpreted as a formal description of the
+requirements for the workflow execution. Once these requirements are matched
+by a set of resource capabilities, an abstract workflow becomes a concrete
+workflow, i.e. a formal description of the workflow execution process and its
+execution environment.
 
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item Other workload management systems (WMS) (not exclusive)
-  \begin{itemize}
+  \begin{enumerate}
     \item glidein-WMS
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item PanDA WMS
-  \begin{itemize}
+  \begin{enumerate}
     \item State Diagrams

### 2016-09-19 WMSLiterature.tex

@@
 \subsection{HTCondor Glidein and GlideinWMS}
-The HTCondor Glidein system  as part of
-the HTCondor [153] software ecosystem. The HTCondor
-Glidein is a pilot-based system to aggregate DCRs with
-heterogeneous middleware into HTCondor resource pools.
-The logical components of HTCondor relevant to the
-Glidein system are: 
+The HTCondor Glidein system  as part of the HTCondor software ecosystem. The HTCondor Glidein is a pilot based system to aggregate DCRs with heterogeneous middleware into HTCondor resource pools.
+Condor is based on daemons collaborating by exchanging messages over the network. We can isolate four main logical components:
 \begin{itemize}
-\item \emph{Schedd}, a daemon that implement a queuing system that holds workload tasks;
-\item \emph{Startd}, a  daemon to control the DCR resources. 
+\item \emph{Schedd}, implements a queuing system that holds workload tasks;
+\item \emph{Startd}, controls the DCR resources. 
 \item \emph{Collector}, holds references to all the active
@@
 \end{itemize}
-In
-HTCondor Glidein has been complemented by Glidein-WMS, a Glidein-based workload management system thatautomates deployment and management of Glideins on multiple types of DCR middleware.
-GlideinWMS builds upon the
-HTCondor Glidein system by adding the following logical components: 
+Glidein-WMS has been developed to integrate HTCondor Glidein to  automate the deployment and management of Glideins on multiple types of DCR middleware. 
+The integration required three additional logical components: 
 \begin{itemize}
@@
 
-Both HTCondor Glidein and GlideWMS rely on one or more HTCondor Collectors to match task requirements and resource properties, represented as ClassAds. This
-matching can be evaluated right before the scheduling of the task. In this way, late binding is achieved but early binding remains unsupported.
+Both HTCondor Glidein and GlideWMS rely on one or more HTCondor Collectors to match task requirements and resource properties, represented as ClassAds.
+This matching can be evaluated right before the scheduling of the task.
+In this way, late binding is achieved but early binding remains unsupported.
+\begin{table*}
+\begin{center}
+\begin{tabular}{llllllll}
+  \hline
+Pilot System  &Logical Components& Execution Strategy & Binding  & Workload Definition  &  Broker  & \\
+\hline
+Alien & Central services, site services & pull, pilot via Co-pilot & Late & JDL & Condor (ClassAd attributes) &\\
+DIRAC & Services, Agents& pull, pilot-based & Late & JDL, WF (TMS) & Condor(ClassAd attributes) &\\
+glidein WMS& Daemons & pull, pilot-based  & Late & Pegaus, DAGMan & Condor(ClassAd attributes) &\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Comparison of the three Workload and Data Management Systems}\label{tab:Summary}
+\end{table*}

### 2016-09-19 paper.tex

--- 2016-09-13 paper.tex
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figure/DailyJobs.png}
+\includegraphics[width=\columnwidth]{figures/DailyJobs.png}
 \caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
@@
   \item Brief overview of distributed CI.
-  \item Workload vs Workflow
-  \item Distinguish between WFMS and WLMS!!!!
+  \item Workload vs Workflow. (DONE)
+  \item Distinguish between WFMS and WLMS.
   \begin{enumerate}
-    \item Role of Workload MS: help “hide” and federate heterogeneous resources.
+    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
   \end{enumerate}
 \end{enumerate}
+
+\paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
+disciplines with different meaning. In the field of scientific computing,
+``workflow'' assumes different meanings depending on the characteristics of
+the computation, of the software tools used to support this computation, and
+of the resources on which it is performed. Further, a workflow may indicate a
+whole application, a description of the computational process of that
+application or, more commonly, a series of tasks related by data
+dependences.
+
+The lack of a consistent and shared definition of ``workflow'' hinders the
+understanding of its properties and its relations with related concepts. For
+example, we need to clarify the difference between ``workflow'', ``workload'',
+``task'', or ``job'' but also between workflow template and instance, or
+data-flow and control-flow. This clarification is precondition to specify
+properly the design of software systems that support the execution of
+scientific applications.
+
+In this paper, a workflow is a set of tasks alongside a description of their
+properties and relations. Each task may have an arbitrary number of properties
+like number of cores, executables, or input/output files. Tasks may have
+precedence relations depending on their data dependences or any other type of
+dependency mandated by the application algorithms. Tasks with precedence
+relations have to be executed serially, otherwise tasks can be executed
+concurrently. A workload is defined as a set of tasks that can be executed
+concurrently. As such, a workflow can be composed by a set of workloads.
+
+The terms ``task'' and ``job'' are also used inconsistently across communities
+that perform scientific computing. In this paper, a task is a \ldots
+\mtnote{Are we going to use task and job as defined in Jedi and Panda Server?}
+
+Usually, workflows are represented as graphs where tasks are vertices and
+their relations edges~\cite{}. Often, graphs are supposed to be acyclic but
+graphs with cycles have been used to represent workflows of workflows. In this
+paper, a workflow template is a type of graph while a workflow instance is a
+type of graph with a specific set of vertices and edges. Further, a graph is
+``abstract'' when no resource properties are available for all vertices,
+``concrete'' otherwise~\cite{}.
+
+The distinction between an abstract and concrete graph highlights an important
+element of the execution process of a workflow. Workflows can be specified
+without indicating on what resources they should be executed. Accordingly, an
+abstract workflow can be interpreted as a formal description of the
+requirements for the workflow execution. Once these requirements are matched
+by a set of resource capabilities, an abstract workflow becomes a concrete
+workflow, i.e. a formal description of the workflow execution process and its
+execution environment.
 
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figure/PandaArch.jpg}
+\includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
 \caption{Schematic view of the PanDA system\label{fig:architecture}}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figure/PandaInterfaceWithHPC.png}
+\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
 \caption{Schematic view of the PanDA system\label{fig:architecture}}

### 2016-09-20 paper.tex
Added bibliography tex file, minor changes in Related works

@@
  Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.
-Heterogeneous  re- sources are distributed worldwide, thousands of physicists analyz- ing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year.
+Heterogeneous  resources are distributed worldwide, thousands of physicists analyzing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year.
 The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
@@
 Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers.
-Thousands of physicists analyze peta-scale  data volumes daily.
-One of  the LHC experiments  the ATLAS  [1]  utilizes the PanDA  workload management system [4] (WMS) for distributed data processing and analysis.
-The ATLAS Computing model [2] is based on a Grid paradigm [3], with multilevel, hierarchically distributed computing  and storage resources.
+Thousands of physicists analyze petascale  data volumes daily.
+One of  the LHC experiments  the ATLAS  \cite{Aad:2008}  utilizes the PanDA  workload management system \cite{Maeno2011} (WMS) for distributed data processing and analysis.
+The ATLAS Computing model \cite{Atlas2005} is based on a Grid paradigm \cite{Foster:1998}, with multilevel, hierarchically distributed computing  and storage resources.
 Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
@@
 Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade.
-PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middle- ware).
+PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middleware).
 This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
-PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user anal- ysis, while actively evolving to meet rapidly changing science needs.
+PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user analysis, while actively evolving to meet rapidly changing science needs.
 Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
@@
   \end{enumerate}
+{enumerate}
   \item Comparison and contrast
@@
 The repository works with Apache web server configured  with a  memory-based   web proxy (Squid).
-The purpose  of the cache  is to reduce  the
-
-request load on the back-end Subversion  server.
+The purpose  of the cache  is to reduce  the request load on the back-end Subversion  server.
 That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server.
@@
 Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote.
-The behavior of these APF workflows is determined by the combination of a set  of plug-ins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.
+The behavior of these APF workflows is determined by the combination of a set  of plugins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.
 
 \subsection{PanDA Monitoring}
-The PanDA Monitor is a web-based dashboard-style  graphi- cal application.
-It runs on Apache and interacts with the back- end database for persistence.
+The PanDA Monitor is a web-based dashboard-style  graphical application.
+It runs on Apache and interacts with the back-end database for persistence.
 The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory.
@@
 Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
-Integration with Titan is the current focus for PanDA develop- ers.
-The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface [6].
+Integration with Titan is the current focus for PanDA developers.
+The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}.
 This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world.
@@
  \section{Conclusion}\label{sec:conclusion}
-The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.  In the process,  the old batch job paradigm  of
-computing in HEP was  discarded  in favor  of a  far more flexible and scalable  model.
+The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
+In the process,  the old batch job paradigm  of computing in HEP was  discarded  in favor of a  far more flexible and scalable  model.
 The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA  is the first exascale  workload manage- ment system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013.
+PanDA  is the first exascale  workload management system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013.
 Next LHC run will pose massive computing  challenges.
@@
 Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry.
-The technology base provided by the PanDA system will enhance the usage of a variety  of high- performance computing resources available to basic research.
+The technology base provided by the PanDA system will enhance the usage of a variety  of high-performance computing resources available to basic research.
 
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
 
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
+\bibliographystyle{plain}
+\bibliography{bibliography}
 

### 2016-09-20 paper.tex
Iteration definitions section 2

--- 2016-09-19 paper.tex
@@
 
-In this paper, a workflow is a set of tasks alongside a description of their
-properties and relations. Each task may have an arbitrary number of properties
-like number of cores, executables, or input/output files. Tasks may have
-precedence relations depending on their data dependences or any other type of
-dependency mandated by the application algorithms. Tasks with precedence
-relations have to be executed serially, otherwise tasks can be executed
-concurrently. A workload is defined as a set of tasks that can be executed
-concurrently. As such, a workflow can be composed by a set of workloads.
+In this paper, we use the following definitions:
+
+\begin{description}
+  \item[Task]     A set of operations to be performed on a computing platform,
+                  alongside a description of the properties and dependences of
+                  those operations, and indications on how they should be
+                  executed and satisfied.
+  \item[Workload] A set of tasks that can be executed concurrently, possibly
+                  related by a set of relations. For example, tasks of a workload can share one or more input files or communicate during execution.
+  \item[Workflow] Set of workloads, related by a set of relations that define
+                  the order in which each workload can be executed. Data
+                  dependences are the most common relations among workloads,
+                  used to define the precedence among their executions. Note
+                  that, formally, a workload can have a single task.
+\end{description}
+
+Each task may have an arbitrary number of properties like number of cores,
+executables, or input/output files. Tasks may have precedence relations
+depending on their data dependences or any other type of dependency mandated
+by the application algorithms. Tasks with precedence relations have to be
+executed serially, otherwise tasks can be executed concurrently. A workload is
+defined as a set of tasks that can be executed concurrently. As such, a
+workflow can be composed by a set of workloads.
 
 The terms ``task'' and ``job'' are also used inconsistently across communities
-that perform scientific computing. In this paper, a task is a \ldots
-\mtnote{Are we going to use task and job as defined in Jedi and Panda Server?}
+that perform scientific computing. In this paper, ``job'' refers to the entity
+that is submitted to a local resource management system (LRMS), like a batch
+system or a scheduler. As such, a task can become a job when is scheduled on a
+resource that exposes a LRMS but can also become a virtual machine or a
+container when bootstrapped on an infrastructure supporting virtualization.
 
 Usually, workflows are represented as graphs where tasks are vertices and
-their relations edges~\cite{}. Often, graphs are supposed to be acyclic but
-graphs with cycles have been used to represent workflows of workflows. In this
-paper, a workflow template is a type of graph while a workflow instance is a
-type of graph with a specific set of vertices and edges. Further, a graph is
-``abstract'' when no resource properties are available for all vertices,
-``concrete'' otherwise~\cite{}.
+their relations are edges~\cite{}. Often, graphs are supposed to be acyclic
+but graphs with cycles have been used to represent workflows of
+workflows~\cite{}. In this paper, a workflow template is a type of graph while
+a workflow instance is a type of graph with a specific set of vertices and
+edges. Further, a graph is ``abstract'' when no resource properties are
+available for all vertices, ``concrete'' otherwise~\cite{}.
 
@@
 without indicating on what resources they should be executed. Accordingly, an
-abstract workflow can be interpreted as a formal description of the
-requirements for the workflow execution. Once these requirements are matched
-by a set of resource capabilities, an abstract workflow becomes a concrete
-workflow, i.e. a formal description of the workflow execution process and its
-execution environment.
+abstract graph can be interpreted as a formal description of the requirements
+for the workflow execution. Once these requirements are matched by a set of
+resource capabilities, the graph becomes concrete, i.e. a formal description
+of the workflow execution process and its execution environment.
 

### 2016-09-20 WMSLiterature.tex

@@
 Alien is a Workload and Data Management system composed of a 	set of middleware tools and services entirely based on web-services and standard protocols.
-Alien was originally developed for the ALICE experiment \cite{} but subsequently used by several virtual organizations \cite{}.
+The framework was originally developed for the ALICE experiment \cite{Alice1995} but subsequently used by several virtual organizations \cite{McClatechey2003,GPCALMA}.
 %The system has been deployed in 2001 for distributed producti	on of Monte Carlo data, detector simulation and reconstruction.
-Alien is composed of two type of services:
+
+Alien is composed of two type of services \cite{Bagnasco2010}:
 \begin{itemize}
@@
 
-The most important central services are:
+The most important Central services are:
 \begin{itemize}
-\item \emph{Job Manager}, a database that keeps track of all the jobs submitted to the system and their current execution status;
+\item \emph{Job Manager} (or \emph{Task Queue}), a database that keeps track of all the jobs submitted to the system and their current execution status;
 \item \emph{Brokers}, they are the core of task executions and data transfers; they receive tasks in form of JDL,  keep them ordered by priority and send them to the CE for execution;
@@
 
-The CE asks the Broker for jobs to execute by sending its JDL.
-Once received the JDL, the Broker will try to match it with the JDL of the jobs in queue.
-If a match exists then the Broker sends the jobs JDL to the CE.
+The CE asks the Broker for jobs to execute by sending its description in form of Job Description Language (JDL).
+Once received the JDL, the Broker will try to match it with the JDL of the jobs in queue. If a match exists then the Broker sends the jobs JDL to the CE.
 Immediately after receving a job JDL, the CE create a new service on the worker node called ProcessMonitor.
-This service allows the CE (and the rest of Alien services through the CE) to interact with the job while is running.
+This service allows the CE (and the rest of Alien services through the CE) to interact with the job while is running \cite{Saiz2003}.
 This execution strategy is called ``pull mode'' due to the fact that CE asks for jobs.
-It is worth to mention that more recent version of Alien implement exploit gLite and CE-CREAM. The latter allows the system to by-pass the broker during the job submission \cite{}.
+It is worth to mention that recent versions of Alien exploit gLite and CE-CREAM. The latter allows the system to by-pass the broker during the job submission \cite{Bagnasco2004}.
 
-Alien uses the Job Description Language (JDL) to allow users to describe their workload task by task;  i.e.,  users can specify features such as task priorities, the level of parallelism (one core, multi-core, MPI etc..) and  also the DCR that should be targeted for the execution.
+Since Alien exploits JDL, a  workload can be described task by task according to features such as: task priorities, the level of parallelism (one core, multi-core, MPI etc..) and  also the DCR that should be targeted for the execution.
 
@@
 \subsection{DIRAC} 
-DIRAC (Distributed Infrastructure with Remote Agent Control) Workload and Data Management System is a software product, developed within the CERN LHCb project, to manage the processing of detector data, Monte Carlo simulations, and end-user analyses.
-DIRAC  architecture relies on four entities:
+DIRAC (Distributed Infrastructure with Remote Agent Control) Workload and Data Management System has been developed within the CERN LHCb project to manage the processing of detector data, Monte Carlo simulations, and end-user analyses \cite{Tsaregorodtsev2004}.
+DIRAC  architecture relies on four entities \cite{Paterson2010}:
 \begin{itemize}
@@
 \item \emph{Agents}: perform repetitive tasks like querying file catalogs,  monitoring of jobs on resources.
-\item \emph{Resources}: they can be PC's, site cluters and Grids. Agents interact with them without distinction.
+\item \emph{Resources}: they can be PC's, site clusters and Grids. Agents interact with them without distinction.
 \end{itemize}
-In the same way of Alien, DIRAC implements a pull scheduling. Furthermore DIRAC was the first WMS to exploit the concept of Pilot Agent on the Grid. 
+In the same way of Alien, DIRAC implements a pull scheduling. Furthermore DIRAC was the first WMS to exploit the concept of Pilot Agent on the Grid \cite{Casajus2010}. 
 Pilots Agents are gLite jobs that are submitted to the grid when jobs arrive into the WMS. 
@@
 
+JDL and \emph{Transformation Management System} (TMS) are used for task description \cite{Tsaregorodtsev2006,Pacini2006}. The latter allows the user to describe complex data dependences. 
 \subsection{HTCondor Glidein and GlideinWMS}

### 2016-09-20 paper.tex
list of components to be discussed.
Fix latex

@@
 
--------------------------------------------------------------------------------------------
+--------------------------------------------------------------------
 \begin{enumerate}
@@
 
+--------------------------------------------------------------------
+
 \paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
@@
 \begin{description}
-  \item[Task]     A set of operations to be performed on a computing platform,
-                  alongside a description of the properties and dependences of
-                  those operations, and indications on how they should be
-                  executed and satisfied.
-  \item[Workload] A set of tasks that can be executed concurrently, possibly
-                  related by a set of relations. For example, tasks of a workload can share one or more input files or communicate during execution.
-  \item[Workflow] Set of workloads, related by a set of relations that define
-                  the order in which each workload can be executed. Data
-                  dependences are the most common relations among workloads,
-                  used to define the precedence among their executions. Note
-                  that, formally, a workload can have a single task.
+  \item[Task.] A set of operations to be performed on a computing platform,
+  alongside a description of the properties and dependences of those
+  operations, and indications on how they should be executed and satisfied.
+  \item[Workload.] A set of tasks that can be executed concurrently, possibly
+  related by a set of relations. For example, tasks of a workload can share
+  one or more input files or communicate during execution.
+  \item[Workflow.] Set of workloads, related by a set of relations that define
+  the order in which each workload can be executed. Data dependences are the
+  most common relations among workloads, used to define the precedence among
+  their executions. Note that, formally, a workload can have a single task.
 \end{description}
@@
 \input{WMSLiterature}
---------------------------------------------------------------------------------------------------
+--------------------------------------------------------------------
 \begin{enumerate}
@@
 \includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
-\caption{Schematic view of the PanDA system\label{fig:architecture}}
+\caption{Schematic view of the PanDA system\label{fig:interface}}
 \end{center}
@@
 
--------------------------------------------------------------------------------------
+--------------------------------------------------------------------
 \begin{enumerate}

### 2016-09-20 paper.tex
minor notes.

--- 2016-09-20 paper.tex
@@
 
+\begin{itemize}
+\item PanDA Server
+\item Pilot
+\item Factory
+\item PanDA monitoring
+\item JEDI ??
+\item What else?
+\end{itemize}
+
 \begin{figure}
@@
 \includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
-\caption{Schematic view of the PanDA system\label{fig:architecture}}
+\caption{Schematic view of the PanDA system\label{fig:architecture}. 
+  Originally PanDA was designed for grid infrastructure... In this paper
+  is focussed on the HPC use case; there are differences from  the grid use
+  case. This paper discusses how PanDA has been adapted to execute a workload on
+  HPC resources. The schematic overview is presented for workload X on Titan......}
 \end{center}

### 2016-09-22 overall.tex
basic restructuring

--- 2016-09-20 paper.tex
@@
-%% bare_conf.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
-% Add the compsocconf option for Computer Society conferences.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-\input{head}
-\input{include}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\newcommand{\sarp}[1]{{\color{green}\textit{sarp: {#1}}}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Everthing you wanted to know about PanDA but were truly truly afraid to ask}
-
-% author names and affiliations
-% use a multiple column layout for up to two different
-% affiliations
-
-\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-\and
-\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3},
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
- Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.
-Heterogeneous  resources are distributed worldwide, thousands of physicists analyzing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year.
-The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model.
-The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
-2013.
-We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
-\end{abstract}
-
-\begin{IEEEkeywords}
-TBD.
-\end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-% ----------------------------------------------------------------------------
-% I
-% ----------------------------------------------------------------------------
-\section{Introduction}\label{sec:intro}
-The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades.
-Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers.
-Thousands of physicists analyze petascale  data volumes daily.
-One of  the LHC experiments  the ATLAS  \cite{Aad:2008}  utilizes the PanDA  workload management system \cite{Maeno2011} (WMS) for distributed data processing and analysis.
-The ATLAS Computing model \cite{Atlas2005} is based on a Grid paradigm \cite{Foster:1998}, with multilevel, hierarchically distributed computing  and storage resources.
-Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
-PanDA has a highly scalable architecture.
-Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade.
-PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middleware).
-This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
-PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user analysis, while actively evolving to meet rapidly changing science needs.
-Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/DailyJobs.png}
-\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
-\end{center}
-\end{figure}
-
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
-  \item Why are WMS needed?
-  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
-  \item \ldots
-  \item Overview/Summary of paper.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% II
-% ----------------------------------------------------------------------------
-\section{Background}
-\label{sec:background}
-
-\begin{enumerate}
-  \item Brief overview of distributed CI.
-  \item Workload vs Workflow. (DONE)
-  \item Distinguish between WFMS and WLMS.
-  \begin{enumerate}
-    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
-  \end{enumerate}
-\end{enumerate}
-
---------------------------------------------------------------------
-
-\paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
-disciplines with different meaning. In the field of scientific computing,
-``workflow'' assumes different meanings depending on the characteristics of
-the computation, of the software tools used to support this computation, and
-of the resources on which it is performed. Further, a workflow may indicate a
-whole application, a description of the computational process of that
-application or, more commonly, a series of tasks related by data
-dependences.
-
-The lack of a consistent and shared definition of ``workflow'' hinders the
-understanding of its properties and its relations with related concepts. For
-example, we need to clarify the difference between ``workflow'', ``workload'',
-``task'', or ``job'' but also between workflow template and instance, or
-data-flow and control-flow. This clarification is precondition to specify
-properly the design of software systems that support the execution of
-scientific applications.
-
-In this paper, we use the following definitions:
-
-\begin{description}
-  \item[Task.] A set of operations to be performed on a computing platform,
-  alongside a description of the properties and dependences of those
-  operations, and indications on how they should be executed and satisfied.
-  \item[Workload.] A set of tasks that can be executed concurrently, possibly
-  related by a set of relations. For example, tasks of a workload can share
-  one or more input files or communicate during execution.
-  \item[Workflow.] Set of workloads, related by a set of relations that define
-  the order in which each workload can be executed. Data dependences are the
-  most common relations among workloads, used to define the precedence among
-  their executions. Note that, formally, a workload can have a single task.
-\end{description}
-
-Each task may have an arbitrary number of properties like number of cores,
-executables, or input/output files. Tasks may have precedence relations
-depending on their data dependences or any other type of dependency mandated
-by the application algorithms. Tasks with precedence relations have to be
-executed serially, otherwise tasks can be executed concurrently. A workload is
-defined as a set of tasks that can be executed concurrently. As such, a
-workflow can be composed by a set of workloads.
-
-The terms ``task'' and ``job'' are also used inconsistently across communities
-that perform scientific computing. In this paper, ``job'' refers to the entity
-that is submitted to a local resource management system (LRMS), like a batch
-system or a scheduler. As such, a task can become a job when is scheduled on a
-resource that exposes a LRMS but can also become a virtual machine or a
-container when bootstrapped on an infrastructure supporting virtualization.
-
-Usually, workflows are represented as graphs where tasks are vertices and
-their relations are edges~\cite{}. Often, graphs are supposed to be acyclic
-but graphs with cycles have been used to represent workflows of
-workflows~\cite{}. In this paper, a workflow template is a type of graph while
-a workflow instance is a type of graph with a specific set of vertices and
-edges. Further, a graph is ``abstract'' when no resource properties are
-available for all vertices, ``concrete'' otherwise~\cite{}.
-
-The distinction between an abstract and concrete graph highlights an important
-element of the execution process of a workflow. Workflows can be specified
-without indicating on what resources they should be executed. Accordingly, an
-abstract graph can be interpreted as a formal description of the requirements
-for the workflow execution. Once these requirements are matched by a set of
-resource capabilities, the graph becomes concrete, i.e. a formal description
-of the workflow execution process and its execution environment.
-
-% ----------------------------------------------------------------------------
-% III
-% ----------------------------------------------------------------------------
-\section{Related Work}
-\label{sec:related}
-\input{WMSLiterature}
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Other workload management systems (WMS) (not exclusive)
-  \begin{enumerate}
-    \item glidein-WMS
-    \item Dirac
-    \item ALICE-AlieN
-  \end{enumerate}
-{enumerate}
-  \item Comparison and contrast
-  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% IV
-% ----------------------------------------------------------------------------
-\section{PanDA Overview}\label{sec:panda_overview}
-PanDA consists of several interconnected subsystems, most of them are built from off-the-shelf and Open Source components. Figure \ref{fig:architecture} shows a schematic view of the PanDA system.
-In the following we will briefly describe system’s architecture and components.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
-\caption{Schematic view of the PanDA system\label{fig:architecture}}
-\end{center}
-\end{figure}
-\subsection{PanDA Server}
-The PanDA server is the main component of the system, It provides  a task queue managing all job information centrally.
-The PanDA  server receives jobs through the client interface into the task queue, upon which a brokerage module  operates to prioritize  and assign work on the basis of job type, priority, input data and its locality, and available CPU resources.
-The PanDA  server operates as a web service.
-It runs on Apache web server, interacting with the back-end database running  on separate servers.
-It uses the Apache worker  model, with many independent   processes  handling client requests  in  parallel.
-Since all  state is maintained  in  the central database,  the PanDA Server application instance itself is stateless.
-Currently production version of PanDA utilizes Oracle database backend, but PanDA  can also work with MySQL family of databases.
-
-\subsection{PanDA Pilot}
- PanDA is a pilot based workload  management system.
-In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload processing on a worker  node) are submitted to sites.
-When these  pilot jobs start on a  worker node they contact a central  server  to retrieve  a real payload (i.e., an end-user job) and execute it.
-Using these pilot-based  workflows helps to improve job reliability, optimize resource utilization, allows for opportunistic  resources usage, and mitigates  many of the problems  associated with the inhomogeneities found on the Grid.
-The pilots themselves  do not contain all the functionality needed to request a payload  job from the Panda Server.
-First, the complete  set of pilot code is downloaded via HTTP from a  central Subversion repository.
-The repository works with Apache web server configured  with a  memory-based   web proxy (Squid).
-The purpose  of the cache  is to reduce  the request load on the back-end Subversion  server.
-That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server.
-The benefits of this system include a high-performance code download service combined with code updates still being immediately  available  as soon as they are committed to source code control.
-
-\subsection{PanDA Factory}
-As a pilot-based  system, PanDA  requires some way to get the initial PanDA pilot onto worker nodes at sites.
-This is done with the help of a component  called AutoPyFactory (APF).
-APF runs in a single daemonized process, launching  a separate thread for each internal  workflow.
-Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote.
-The behavior of these APF workflows is determined by the combination of a set  of plugins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.
-
-\subsection{PanDA Monitoring}
-The PanDA Monitor is a web-based dashboard-style  graphical application.
-It runs on Apache and interacts with the back-end database for persistence.
-The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory.
-The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.
-
-\section{Evolution of PanDA WMS}
-While PanDa  currently uses  more than 100,000  cores at well over 100 Grid sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will  require more resources  than Grid computing can possibly provide.
-The Worldwide LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned analysis and data processing, but it will be insufficient  for Monte Carlo (MC) production and any extra activities.
-Additional computing and storage  resources are therefore required.
-To alleviate  these challenges, ATLAS is engaged  in an ambitious  program to expand  the current computing model to include  additional  resources such as the opportunistic   use of supercomputers   as well as commercial and academic clouds.
-In turn this activity drives evolution of the PanDA WMS.
-
-\subsection{Use of Supercomputers with PanDA}
-The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers.
-Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network.
-The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors.
-On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI).
-While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization.
-It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run.
-What matters is the total number of cycles that can be offloaded from the Grid.
-Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc.
-A reorganization  of the standard  workflow is therefore needed.
-The most suitable task for HPCs in HEP is event generation.
-Event generators are mostly computational,  stand-alone code, with few input requirements.
-There is no need  to stage-in much data.
-The event generation  in ATLAS corresponds  to
-10-15\% of all jobs on the Grid.
-A more difficult workload to adapt  for HPCs is detector  simulation.
-Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.
-However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
-\subsection{Interfacing PanDA with Titan}
-The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US.
-Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
-Integration with Titan is the current focus for PanDA developers.
-The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}.
-This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world.
-The pilot can communicate with the PanDA server from the front-end machine.
-The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job.
-ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
-1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design.
-Time allocation on an HPC is competitive and large projects are often preferred.
-About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources.
-The worker nodes
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
-\caption{Schematic view of the PanDA system\label{fig:interface}}
-\end{center}
-\end{figure}
-
---------------------------------------------------------------------
-\begin{enumerate}
-  \item PanDA WMS
-  \begin{enumerate}
-    \item State Diagrams
-    \item Control Flow Diagrams
-    \item Software Module and Architecture diagrams.
-  \end{enumerate}
-  \item PANDA Pilot
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% V
-% ----------------------------------------------------------------------------
-\section{Deploying PanDA on a Leadership-scale system}
-\label{sec:panda_deployment}
-
-\begin{enumerate}
-  \item Challenges
-  \item Opportunities
-  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
-  \item Take-aways, lessons learned
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VI
-% ----------------------------------------------------------------------------
-\section{Performance Characterization}
-\label{sec:panda_performance}
-
-\begin{enumerate}
-  \item Internal PanDA performance
-  \item External PanDA performance (workload specific on underlying system/platform)
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VII
-% ----------------------------------------------------------------------------
 \section{Putting it all together: Experience of PanDA on TITAN} % (fold)
 \label{sec:panda_titan}
+
+\subsection{PanDA Metadata Peformance Impact at OLCF}
 
@@
 \end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VIII
-% ----------------------------------------------------------------------------
-\section{PANDA: Future/RoadMap}
-\label{sec:panda_roadmap}
-
- \section{Conclusion}\label{sec:conclusion}
-The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process,  the old batch job paradigm  of computing in HEP was  discarded  in favor of a  far more flexible and scalable  model.
-The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA  is the first exascale  workload management system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013.
-Next LHC run will pose massive computing  challenges.
-With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more.
-Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS.
-To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system.
-Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry.
-The technology base provided by the PanDA system will enhance the usage of a variety  of high-performance computing resources available to basic research.
-
-\input{ack}
-
-%\section*{Acknowledgment}
-
-%The authors would like to thank... More thanks here...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-
-\bibliographystyle{plain}
-\bibliography{bibliography}
-
-% that's all folks
-\end{document}

### 2016-10-04 Background.tex
Splitting in input tex files, First Draft Planned Experiments

--- 2016-09-20 paper.tex
@@
-%% bare_conf.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
-% Add the compsocconf option for Computer Society conferences.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-\input{head}
-\input{include}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\newcommand{\sarp}[1]{{\color{green}\textit{sarp: {#1}}}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Everthing you wanted to know about PanDA but were truly truly afraid to ask}
-
-% author names and affiliations
-% use a multiple column layout for up to two different
-% affiliations
-
-\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-\and
-\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3},
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
- Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.
-Heterogeneous  resources are distributed worldwide, thousands of physicists analyzing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year.
-The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model.
-The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
-2013.
-We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
-\end{abstract}
-
-\begin{IEEEkeywords}
-TBD.
-\end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-% ----------------------------------------------------------------------------
-% I
-% ----------------------------------------------------------------------------
-\section{Introduction}\label{sec:intro}
-The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades.
-Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers.
-Thousands of physicists analyze petascale  data volumes daily.
-One of  the LHC experiments  the ATLAS  \cite{Aad:2008}  utilizes the PanDA  workload management system \cite{Maeno2011} (WMS) for distributed data processing and analysis.
-The ATLAS Computing model \cite{Atlas2005} is based on a Grid paradigm \cite{Foster:1998}, with multilevel, hierarchically distributed computing  and storage resources.
-Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
-PanDA has a highly scalable architecture.
-Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade.
-PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middleware).
-This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
-PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user analysis, while actively evolving to meet rapidly changing science needs.
-Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/DailyJobs.png}
-\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
-\end{center}
-\end{figure}
-
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
-  \item Why are WMS needed?
-  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
-  \item \ldots
-  \item Overview/Summary of paper.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% II
-% ----------------------------------------------------------------------------
 \section{Background}
 \label{sec:background}
-
-\begin{enumerate}
-  \item Brief overview of distributed CI.
-  \item Workload vs Workflow. (DONE)
-  \item Distinguish between WFMS and WLMS.
-  \begin{enumerate}
-    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
-  \end{enumerate}
-\end{enumerate}
-
---------------------------------------------------------------------
-
 \paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
@@
 of the workflow execution process and its execution environment.
-
-% ----------------------------------------------------------------------------
-% III
-% ----------------------------------------------------------------------------
-\section{Related Work}
-\label{sec:related}
-\input{WMSLiterature}
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Other workload management systems (WMS) (not exclusive)
-  \begin{enumerate}
-    \item glidein-WMS
-    \item Dirac
-    \item ALICE-AlieN
-  \end{enumerate}
-{enumerate}
-  \item Comparison and contrast
-  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% IV
-% ----------------------------------------------------------------------------
-\section{PanDA Overview}\label{sec:panda_overview}
-PanDA consists of several interconnected subsystems, most of them are built from off-the-shelf and Open Source components. Figure \ref{fig:architecture} shows a schematic view of the PanDA system.
-In the following we will briefly describe system’s architecture and components.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
-\caption{Schematic view of the PanDA system\label{fig:architecture}}
-\end{center}
-\end{figure}
-\subsection{PanDA Server}
-The PanDA server is the main component of the system, It provides  a task queue managing all job information centrally.
-The PanDA  server receives jobs through the client interface into the task queue, upon which a brokerage module  operates to prioritize  and assign work on the basis of job type, priority, input data and its locality, and available CPU resources.
-The PanDA  server operates as a web service.
-It runs on Apache web server, interacting with the back-end database running  on separate servers.
-It uses the Apache worker  model, with many independent   processes  handling client requests  in  parallel.
-Since all  state is maintained  in  the central database,  the PanDA Server application instance itself is stateless.
-Currently production version of PanDA utilizes Oracle database backend, but PanDA  can also work with MySQL family of databases.
-
-\subsection{PanDA Pilot}
- PanDA is a pilot based workload  management system.
-In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload processing on a worker  node) are submitted to sites.
-When these  pilot jobs start on a  worker node they contact a central  server  to retrieve  a real payload (i.e., an end-user job) and execute it.
-Using these pilot-based  workflows helps to improve job reliability, optimize resource utilization, allows for opportunistic  resources usage, and mitigates  many of the problems  associated with the inhomogeneities found on the Grid.
-The pilots themselves  do not contain all the functionality needed to request a payload  job from the Panda Server.
-First, the complete  set of pilot code is downloaded via HTTP from a  central Subversion repository.
-The repository works with Apache web server configured  with a  memory-based   web proxy (Squid).
-The purpose  of the cache  is to reduce  the request load on the back-end Subversion  server.
-That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server.
-The benefits of this system include a high-performance code download service combined with code updates still being immediately  available  as soon as they are committed to source code control.
-
-\subsection{PanDA Factory}
-As a pilot-based  system, PanDA  requires some way to get the initial PanDA pilot onto worker nodes at sites.
-This is done with the help of a component  called AutoPyFactory (APF).
-APF runs in a single daemonized process, launching  a separate thread for each internal  workflow.
-Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote.
-The behavior of these APF workflows is determined by the combination of a set  of plugins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.
-
-\subsection{PanDA Monitoring}
-The PanDA Monitor is a web-based dashboard-style  graphical application.
-It runs on Apache and interacts with the back-end database for persistence.
-The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory.
-The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.
-
-\section{Evolution of PanDA WMS}
-While PanDa  currently uses  more than 100,000  cores at well over 100 Grid sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will  require more resources  than Grid computing can possibly provide.
-The Worldwide LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned analysis and data processing, but it will be insufficient  for Monte Carlo (MC) production and any extra activities.
-Additional computing and storage  resources are therefore required.
-To alleviate  these challenges, ATLAS is engaged  in an ambitious  program to expand  the current computing model to include  additional  resources such as the opportunistic   use of supercomputers   as well as commercial and academic clouds.
-In turn this activity drives evolution of the PanDA WMS.
-
-\subsection{Use of Supercomputers with PanDA}
-The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers.
-Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network.
-The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors.
-On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI).
-While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization.
-It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run.
-What matters is the total number of cycles that can be offloaded from the Grid.
-Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc.
-A reorganization  of the standard  workflow is therefore needed.
-The most suitable task for HPCs in HEP is event generation.
-Event generators are mostly computational,  stand-alone code, with few input requirements.
-There is no need  to stage-in much data.
-The event generation  in ATLAS corresponds  to
-10-15\% of all jobs on the Grid.
-A more difficult workload to adapt  for HPCs is detector  simulation.
-Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.
-However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
-\subsection{Interfacing PanDA with Titan}
-The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US.
-Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
-Integration with Titan is the current focus for PanDA developers.
-The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}.
-This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world.
-The pilot can communicate with the PanDA server from the front-end machine.
-The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job.
-ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
-1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design.
-Time allocation on an HPC is competitive and large projects are often preferred.
-About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources.
-The worker nodes
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
-\caption{Schematic view of the PanDA system\label{fig:interface}}
-\end{center}
-\end{figure}
-
---------------------------------------------------------------------
-\begin{enumerate}
-  \item PanDA WMS
-  \begin{enumerate}
-    \item State Diagrams
-    \item Control Flow Diagrams
-    \item Software Module and Architecture diagrams.
-  \end{enumerate}
-  \item PANDA Pilot
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% V
-% ----------------------------------------------------------------------------
-\section{Deploying PanDA on a Leadership-scale system}
-\label{sec:panda_deployment}
-
-\begin{enumerate}
-  \item Challenges
-  \item Opportunities
-  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
-  \item Take-aways, lessons learned
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VI
-% ----------------------------------------------------------------------------
-\section{Performance Characterization}
-\label{sec:panda_performance}
-
-\begin{enumerate}
-  \item Internal PanDA performance
-  \item External PanDA performance (workload specific on underlying system/platform)
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VII
-% ----------------------------------------------------------------------------
-\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
-\label{sec:panda_titan}
-
-\begin{enumerate}
-  \item Metadata performance issue, and how it’s resolved
-  \item File I/O performance and impact
-  \item ?
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VIII
-% ----------------------------------------------------------------------------
-\section{PANDA: Future/RoadMap}
-\label{sec:panda_roadmap}
-
- \section{Conclusion}\label{sec:conclusion}
-The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process,  the old batch job paradigm  of computing in HEP was  discarded  in favor of a  far more flexible and scalable  model.
-The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA  is the first exascale  workload management system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013.
-Next LHC run will pose massive computing  challenges.
-With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more.
-Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS.
-To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system.
-Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry.
-The technology base provided by the PanDA system will enhance the usage of a variety  of high-performance computing resources available to basic research.
-
-\input{ack}
-
-%\section*{Acknowledgment}
-
-%The authors would like to thank... More thanks here...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-
-\bibliographystyle{plain}
-\bibliography{bibliography}
-
-% that's all folks
-\end{document}

### 2016-10-04 Experiments.tex
Splitting in input tex files, First Draft Planned Experiments

The execution of jobs on TITAN follows a three step scheme: i) as first, TITAN's Backfill queue communicates the presence of a new allocation i.e. a set of unused resources that can be used for a pre-determined time period\aanote{How many offers does PanDA receive per time unit? Average nodes per offer?}; ii) then, PanDA must accept or decline the allocation; iii) if the allocation is accepted, PanDA wraps a certain amount of events in a job and sends the job for execution.\aanote{I am supposing that the concept of event has been already introduced at this point}.
At least in principle, PanDA can reject an allocation because it does not fit the requirements in terms of number of nodes or wall-time unit; however, this situation is rare because events are single core Athena-MP threads that require on average twenty minutes for being executed. Thus, the decomposition in event is fine grained once compared to the average allocation \aanote{Which correspond to......}.

As a consequence, PanDA accepts the major part of the proposed allocations by selecting a number of events that maximize the utilization of the resources during the assigned time interval.

On one hand, this execution strategy leads to a drastic reduction of the queue time and, consequently, it guarantees high-performances. On the other hand, this strategy is tailored for the execution of ATLAS workload on TITAN by using the Backfill queue.

Questions such as \emph{``Is this the best strategy possible?''} or \emph{``Can PanDA achieve similar performances by using TITAN's batch queue?''}  remain unanswered.
For this reason, in the remaining of this section we propose experimental results to address them. The idea behind the experiments is to show the behaviour of ATLAS workload on TITAN's batch queue.  
As first, we investigate resource allocation as a function of time (wall-time) and space (number of nodes) in order to understand how TITAN's scheduler behaves under different allocation requests. This analysis includes two bulks of experiments: the first, generated ``in silico'', is based on simulations of the TITAN's scheduler by means of the Moab simulator, whereas the second, generated ``in vivo'',  considers real submissions of ATLAS workload on TITAN's batch queue.

By using the batch queue, we move in a situation where the queue time becomes dominant but, at the same time, we gain more freedom to decide the parameters of the allocation. For this reason, the second set of the experiments aims to find sub-optimal parameters for resource allocation.

Additionally, due to the dynamics of the Backfill queue, PanDA has a high probability to re-acquire an allocation immediately after it has released one \aanote{Are we able to quantify this ``immediately''?}.

### 2016-10-04 PandaEvolution.tex
Splitting in input tex files, First Draft Planned Experiments

--- 2016-09-20 paper.tex
@@
-%% bare_conf.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
-% Add the compsocconf option for Computer Society conferences.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-\input{head}
-\input{include}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\newcommand{\sarp}[1]{{\color{green}\textit{sarp: {#1}}}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Everthing you wanted to know about PanDA but were truly truly afraid to ask}
-
-% author names and affiliations
-% use a multiple column layout for up to two different
-% affiliations
-
-\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-\and
-\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3},
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
- Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.
-Heterogeneous  resources are distributed worldwide, thousands of physicists analyzing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year.
-The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model.
-The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
-2013.
-We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
-\end{abstract}
-
-\begin{IEEEkeywords}
-TBD.
-\end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-% ----------------------------------------------------------------------------
-% I
-% ----------------------------------------------------------------------------
-\section{Introduction}\label{sec:intro}
-The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades.
-Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers.
-Thousands of physicists analyze petascale  data volumes daily.
-One of  the LHC experiments  the ATLAS  \cite{Aad:2008}  utilizes the PanDA  workload management system \cite{Maeno2011} (WMS) for distributed data processing and analysis.
-The ATLAS Computing model \cite{Atlas2005} is based on a Grid paradigm \cite{Foster:1998}, with multilevel, hierarchically distributed computing  and storage resources.
-Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
-PanDA has a highly scalable architecture.
-Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade.
-PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middleware).
-This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
-PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user analysis, while actively evolving to meet rapidly changing science needs.
-Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/DailyJobs.png}
-\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
-\end{center}
-\end{figure}
-
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
-  \item Why are WMS needed?
-  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
-  \item \ldots
-  \item Overview/Summary of paper.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% II
-% ----------------------------------------------------------------------------
-\section{Background}
-\label{sec:background}
-
-\begin{enumerate}
-  \item Brief overview of distributed CI.
-  \item Workload vs Workflow. (DONE)
-  \item Distinguish between WFMS and WLMS.
-  \begin{enumerate}
-    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
-  \end{enumerate}
-\end{enumerate}
-
---------------------------------------------------------------------
-
-\paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
-disciplines with different meaning. In the field of scientific computing,
-``workflow'' assumes different meanings depending on the characteristics of
-the computation, of the software tools used to support this computation, and
-of the resources on which it is performed. Further, a workflow may indicate a
-whole application, a description of the computational process of that
-application or, more commonly, a series of tasks related by data
-dependences.
-
-The lack of a consistent and shared definition of ``workflow'' hinders the
-understanding of its properties and its relations with related concepts. For
-example, we need to clarify the difference between ``workflow'', ``workload'',
-``task'', or ``job'' but also between workflow template and instance, or
-data-flow and control-flow. This clarification is precondition to specify
-properly the design of software systems that support the execution of
-scientific applications.
-
-In this paper, we use the following definitions:
-
-\begin{description}
-  \item[Task.] A set of operations to be performed on a computing platform,
-  alongside a description of the properties and dependences of those
-  operations, and indications on how they should be executed and satisfied.
-  \item[Workload.] A set of tasks that can be executed concurrently, possibly
-  related by a set of relations. For example, tasks of a workload can share
-  one or more input files or communicate during execution.
-  \item[Workflow.] Set of workloads, related by a set of relations that define
-  the order in which each workload can be executed. Data dependences are the
-  most common relations among workloads, used to define the precedence among
-  their executions. Note that, formally, a workload can have a single task.
-\end{description}
-
-Each task may have an arbitrary number of properties like number of cores,
-executables, or input/output files. Tasks may have precedence relations
-depending on their data dependences or any other type of dependency mandated
-by the application algorithms. Tasks with precedence relations have to be
-executed serially, otherwise tasks can be executed concurrently. A workload is
-defined as a set of tasks that can be executed concurrently. As such, a
-workflow can be composed by a set of workloads.
-
-The terms ``task'' and ``job'' are also used inconsistently across communities
-that perform scientific computing. In this paper, ``job'' refers to the entity
-that is submitted to a local resource management system (LRMS), like a batch
-system or a scheduler. As such, a task can become a job when is scheduled on a
-resource that exposes a LRMS but can also become a virtual machine or a
-container when bootstrapped on an infrastructure supporting virtualization.
-
-Usually, workflows are represented as graphs where tasks are vertices and
-their relations are edges~\cite{}. Often, graphs are supposed to be acyclic
-but graphs with cycles have been used to represent workflows of
-workflows~\cite{}. In this paper, a workflow template is a type of graph while
-a workflow instance is a type of graph with a specific set of vertices and
-edges. Further, a graph is ``abstract'' when no resource properties are
-available for all vertices, ``concrete'' otherwise~\cite{}.
-
-The distinction between an abstract and concrete graph highlights an important
-element of the execution process of a workflow. Workflows can be specified
-without indicating on what resources they should be executed. Accordingly, an
-abstract graph can be interpreted as a formal description of the requirements
-for the workflow execution. Once these requirements are matched by a set of
-resource capabilities, the graph becomes concrete, i.e. a formal description
-of the workflow execution process and its execution environment.
-
-% ----------------------------------------------------------------------------
-% III
-% ----------------------------------------------------------------------------
-\section{Related Work}
-\label{sec:related}
-\input{WMSLiterature}
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Other workload management systems (WMS) (not exclusive)
-  \begin{enumerate}
-    \item glidein-WMS
-    \item Dirac
-    \item ALICE-AlieN
-  \end{enumerate}
-{enumerate}
-  \item Comparison and contrast
-  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% IV
-% ----------------------------------------------------------------------------
-\section{PanDA Overview}\label{sec:panda_overview}
-PanDA consists of several interconnected subsystems, most of them are built from off-the-shelf and Open Source components. Figure \ref{fig:architecture} shows a schematic view of the PanDA system.
-In the following we will briefly describe system’s architecture and components.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
-\caption{Schematic view of the PanDA system\label{fig:architecture}}
-\end{center}
-\end{figure}
-\subsection{PanDA Server}
-The PanDA server is the main component of the system, It provides  a task queue managing all job information centrally.
-The PanDA  server receives jobs through the client interface into the task queue, upon which a brokerage module  operates to prioritize  and assign work on the basis of job type, priority, input data and its locality, and available CPU resources.
-The PanDA  server operates as a web service.
-It runs on Apache web server, interacting with the back-end database running  on separate servers.
-It uses the Apache worker  model, with many independent   processes  handling client requests  in  parallel.
-Since all  state is maintained  in  the central database,  the PanDA Server application instance itself is stateless.
-Currently production version of PanDA utilizes Oracle database backend, but PanDA  can also work with MySQL family of databases.
-
-\subsection{PanDA Pilot}
- PanDA is a pilot based workload  management system.
-In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload processing on a worker  node) are submitted to sites.
-When these  pilot jobs start on a  worker node they contact a central  server  to retrieve  a real payload (i.e., an end-user job) and execute it.
-Using these pilot-based  workflows helps to improve job reliability, optimize resource utilization, allows for opportunistic  resources usage, and mitigates  many of the problems  associated with the inhomogeneities found on the Grid.
-The pilots themselves  do not contain all the functionality needed to request a payload  job from the Panda Server.
-First, the complete  set of pilot code is downloaded via HTTP from a  central Subversion repository.
-The repository works with Apache web server configured  with a  memory-based   web proxy (Squid).
-The purpose  of the cache  is to reduce  the request load on the back-end Subversion  server.
-That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server.
-The benefits of this system include a high-performance code download service combined with code updates still being immediately  available  as soon as they are committed to source code control.
-
-\subsection{PanDA Factory}
-As a pilot-based  system, PanDA  requires some way to get the initial PanDA pilot onto worker nodes at sites.
-This is done with the help of a component  called AutoPyFactory (APF).
-APF runs in a single daemonized process, launching  a separate thread for each internal  workflow.
-Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote.
-The behavior of these APF workflows is determined by the combination of a set  of plugins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.
-
-\subsection{PanDA Monitoring}
-The PanDA Monitor is a web-based dashboard-style  graphical application.
-It runs on Apache and interacts with the back-end database for persistence.
-The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory.
-The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.
-
 \section{Evolution of PanDA WMS}
+\subsection{Brief Introduction to Titan}
 While PanDa  currently uses  more than 100,000  cores at well over 100 Grid sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will  require more resources  than Grid computing can possibly provide.
@@
 Time allocation on an HPC is competitive and large projects are often preferred.
-About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources.
-The worker nodes
+About 10\% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources.
+The worker nodes sit idle because there are not enough of them or they do not have enough time to handle a large scale computing job.
+On a machine of the scale of Titan, these 10\% correspond to estimated 300M core hours per year.
+A system that can occupy those temporarily free nodes with smaller scale tasks would be very valuable.
+This offers a great possibility for PanDA to harvest the opportunistic resources on Titan and use a similar mechanism on other HPCs.
+Functionality has been added to the PanDA Pilot to collect information about available unused worker nodes on Titan in real time.
+This allows PanDA to define precisely size and duration of jobs submitted to Titan according to available free resources.
+First test of the system were quite successful and show great promise in increasing the resource utilization on Titan.
 \begin{figure}
@@
 \end{figure}
-
---------------------------------------------------------------------
-\begin{enumerate}
-  \item PanDA WMS
-  \begin{enumerate}
-    \item State Diagrams
-    \item Control Flow Diagrams
-    \item Software Module and Architecture diagrams.
-  \end{enumerate}
-  \item PANDA Pilot
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% V
-% ----------------------------------------------------------------------------
-\section{Deploying PanDA on a Leadership-scale system}
-\label{sec:panda_deployment}
-
-\begin{enumerate}
-  \item Challenges
-  \item Opportunities
-  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
-  \item Take-aways, lessons learned
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VI
-% ----------------------------------------------------------------------------
-\section{Performance Characterization}
-\label{sec:panda_performance}
-
-\begin{enumerate}
-  \item Internal PanDA performance
-  \item External PanDA performance (workload specific on underlying system/platform)
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VII
-% ----------------------------------------------------------------------------
-\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
-\label{sec:panda_titan}
-
-\begin{enumerate}
-  \item Metadata performance issue, and how it’s resolved
-  \item File I/O performance and impact
-  \item ?
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VIII
-% ----------------------------------------------------------------------------
-\section{PANDA: Future/RoadMap}
-\label{sec:panda_roadmap}
-
- \section{Conclusion}\label{sec:conclusion}
-The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process,  the old batch job paradigm  of computing in HEP was  discarded  in favor of a  far more flexible and scalable  model.
-The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA  is the first exascale  workload management system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013.
-Next LHC run will pose massive computing  challenges.
-With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more.
-Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS.
-To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system.
-Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry.
-The technology base provided by the PanDA system will enhance the usage of a variety  of high-performance computing resources available to basic research.
-
-\input{ack}
-
-%\section*{Acknowledgment}
-
-%The authors would like to thank... More thanks here...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-
-\bibliographystyle{plain}
-\bibliography{bibliography}
-
-% that's all folks
-\end{document}

### 2016-10-04 PandaOverview.tex
Splitting in input tex files, First Draft Planned Experiments

--- 2016-09-20 paper.tex
@@
-%% bare_conf.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
-% Add the compsocconf option for Computer Society conferences.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-\input{head}
-\input{include}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\newcommand{\sarp}[1]{{\color{green}\textit{sarp: {#1}}}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Everthing you wanted to know about PanDA but were truly truly afraid to ask}
-
-% author names and affiliations
-% use a multiple column layout for up to two different
-% affiliations
-
-\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-\and
-\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3},
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
- Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.
-Heterogeneous  resources are distributed worldwide, thousands of physicists analyzing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year.
-The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model.
-The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
-2013.
-We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
-\end{abstract}
-
-\begin{IEEEkeywords}
-TBD.
-\end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-% ----------------------------------------------------------------------------
-% I
-% ----------------------------------------------------------------------------
-\section{Introduction}\label{sec:intro}
-The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades.
-Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers.
-Thousands of physicists analyze petascale  data volumes daily.
-One of  the LHC experiments  the ATLAS  \cite{Aad:2008}  utilizes the PanDA  workload management system \cite{Maeno2011} (WMS) for distributed data processing and analysis.
-The ATLAS Computing model \cite{Atlas2005} is based on a Grid paradigm \cite{Foster:1998}, with multilevel, hierarchically distributed computing  and storage resources.
-Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
-PanDA has a highly scalable architecture.
-Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade.
-PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middleware).
-This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
-PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user analysis, while actively evolving to meet rapidly changing science needs.
-Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/DailyJobs.png}
-\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
-\end{center}
-\end{figure}
-
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
-  \item Why are WMS needed?
-  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
-  \item \ldots
-  \item Overview/Summary of paper.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% II
-% ----------------------------------------------------------------------------
-\section{Background}
-\label{sec:background}
-
-\begin{enumerate}
-  \item Brief overview of distributed CI.
-  \item Workload vs Workflow. (DONE)
-  \item Distinguish between WFMS and WLMS.
-  \begin{enumerate}
-    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
-  \end{enumerate}
-\end{enumerate}
-
---------------------------------------------------------------------
-
-\paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
-disciplines with different meaning. In the field of scientific computing,
-``workflow'' assumes different meanings depending on the characteristics of
-the computation, of the software tools used to support this computation, and
-of the resources on which it is performed. Further, a workflow may indicate a
-whole application, a description of the computational process of that
-application or, more commonly, a series of tasks related by data
-dependences.
-
-The lack of a consistent and shared definition of ``workflow'' hinders the
-understanding of its properties and its relations with related concepts. For
-example, we need to clarify the difference between ``workflow'', ``workload'',
-``task'', or ``job'' but also between workflow template and instance, or
-data-flow and control-flow. This clarification is precondition to specify
-properly the design of software systems that support the execution of
-scientific applications.
-
-In this paper, we use the following definitions:
-
-\begin{description}
-  \item[Task.] A set of operations to be performed on a computing platform,
-  alongside a description of the properties and dependences of those
-  operations, and indications on how they should be executed and satisfied.
-  \item[Workload.] A set of tasks that can be executed concurrently, possibly
-  related by a set of relations. For example, tasks of a workload can share
-  one or more input files or communicate during execution.
-  \item[Workflow.] Set of workloads, related by a set of relations that define
-  the order in which each workload can be executed. Data dependences are the
-  most common relations among workloads, used to define the precedence among
-  their executions. Note that, formally, a workload can have a single task.
-\end{description}
-
-Each task may have an arbitrary number of properties like number of cores,
-executables, or input/output files. Tasks may have precedence relations
-depending on their data dependences or any other type of dependency mandated
-by the application algorithms. Tasks with precedence relations have to be
-executed serially, otherwise tasks can be executed concurrently. A workload is
-defined as a set of tasks that can be executed concurrently. As such, a
-workflow can be composed by a set of workloads.
-
-The terms ``task'' and ``job'' are also used inconsistently across communities
-that perform scientific computing. In this paper, ``job'' refers to the entity
-that is submitted to a local resource management system (LRMS), like a batch
-system or a scheduler. As such, a task can become a job when is scheduled on a
-resource that exposes a LRMS but can also become a virtual machine or a
-container when bootstrapped on an infrastructure supporting virtualization.
-
-Usually, workflows are represented as graphs where tasks are vertices and
-their relations are edges~\cite{}. Often, graphs are supposed to be acyclic
-but graphs with cycles have been used to represent workflows of
-workflows~\cite{}. In this paper, a workflow template is a type of graph while
-a workflow instance is a type of graph with a specific set of vertices and
-edges. Further, a graph is ``abstract'' when no resource properties are
-available for all vertices, ``concrete'' otherwise~\cite{}.
-
-The distinction between an abstract and concrete graph highlights an important
-element of the execution process of a workflow. Workflows can be specified
-without indicating on what resources they should be executed. Accordingly, an
-abstract graph can be interpreted as a formal description of the requirements
-for the workflow execution. Once these requirements are matched by a set of
-resource capabilities, the graph becomes concrete, i.e. a formal description
-of the workflow execution process and its execution environment.
-
-% ----------------------------------------------------------------------------
-% III
-% ----------------------------------------------------------------------------
-\section{Related Work}
-\label{sec:related}
-\input{WMSLiterature}
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Other workload management systems (WMS) (not exclusive)
-  \begin{enumerate}
-    \item glidein-WMS
-    \item Dirac
-    \item ALICE-AlieN
-  \end{enumerate}
-{enumerate}
-  \item Comparison and contrast
-  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% IV
-% ----------------------------------------------------------------------------
-\section{PanDA Overview}\label{sec:panda_overview}
 PanDA consists of several interconnected subsystems, most of them are built from off-the-shelf and Open Source components. Figure \ref{fig:architecture} shows a schematic view of the PanDA system.
@@
 The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.
-
-\section{Evolution of PanDA WMS}
-While PanDa  currently uses  more than 100,000  cores at well over 100 Grid sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will  require more resources  than Grid computing can possibly provide.
-The Worldwide LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned analysis and data processing, but it will be insufficient  for Monte Carlo (MC) production and any extra activities.
-Additional computing and storage  resources are therefore required.
-To alleviate  these challenges, ATLAS is engaged  in an ambitious  program to expand  the current computing model to include  additional  resources such as the opportunistic   use of supercomputers   as well as commercial and academic clouds.
-In turn this activity drives evolution of the PanDA WMS.
-
-\subsection{Use of Supercomputers with PanDA}
-The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers.
-Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network.
-The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors.
-On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI).
-While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization.
-It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run.
-What matters is the total number of cycles that can be offloaded from the Grid.
-Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc.
-A reorganization  of the standard  workflow is therefore needed.
-The most suitable task for HPCs in HEP is event generation.
-Event generators are mostly computational,  stand-alone code, with few input requirements.
-There is no need  to stage-in much data.
-The event generation  in ATLAS corresponds  to
-10-15\% of all jobs on the Grid.
-A more difficult workload to adapt  for HPCs is detector  simulation.
-Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.
-However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
-\subsection{Interfacing PanDA with Titan}
-The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US.
-Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
-Integration with Titan is the current focus for PanDA developers.
-The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}.
-This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world.
-The pilot can communicate with the PanDA server from the front-end machine.
-The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job.
-ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
-1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design.
-Time allocation on an HPC is competitive and large projects are often preferred.
-About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources.
-The worker nodes
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
-\caption{Schematic view of the PanDA system\label{fig:interface}}
-\end{center}
-\end{figure}
-
---------------------------------------------------------------------
-\begin{enumerate}
-  \item PanDA WMS
-  \begin{enumerate}
-    \item State Diagrams
-    \item Control Flow Diagrams
-    \item Software Module and Architecture diagrams.
-  \end{enumerate}
-  \item PANDA Pilot
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% V
-% ----------------------------------------------------------------------------
-\section{Deploying PanDA on a Leadership-scale system}
-\label{sec:panda_deployment}
-
-\begin{enumerate}
-  \item Challenges
-  \item Opportunities
-  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
-  \item Take-aways, lessons learned
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VI
-% ----------------------------------------------------------------------------
-\section{Performance Characterization}
-\label{sec:panda_performance}
-
-\begin{enumerate}
-  \item Internal PanDA performance
-  \item External PanDA performance (workload specific on underlying system/platform)
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VII
-% ----------------------------------------------------------------------------
-\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
-\label{sec:panda_titan}
-
-\begin{enumerate}
-  \item Metadata performance issue, and how it’s resolved
-  \item File I/O performance and impact
-  \item ?
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VIII
-% ----------------------------------------------------------------------------
-\section{PANDA: Future/RoadMap}
-\label{sec:panda_roadmap}
-
- \section{Conclusion}\label{sec:conclusion}
-The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process,  the old batch job paradigm  of computing in HEP was  discarded  in favor of a  far more flexible and scalable  model.
-The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA  is the first exascale  workload management system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013.
-Next LHC run will pose massive computing  challenges.
-With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more.
-Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS.
-To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system.
-Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry.
-The technology base provided by the PanDA system will enhance the usage of a variety  of high-performance computing resources available to basic research.
-
-\input{ack}
-
-%\section*{Acknowledgment}
-
-%The authors would like to thank... More thanks here...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-
-\bibliographystyle{plain}
-\bibliography{bibliography}
-
-% that's all folks
-\end{document}

### 2016-10-04 introduction.tex
Splitting in input tex files, First Draft Planned Experiments

--- 2016-09-20 paper.tex
@@
-%% bare_conf.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
-% Add the compsocconf option for Computer Society conferences.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-\input{head}
-\input{include}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\newcommand{\sarp}[1]{{\color{green}\textit{sarp: {#1}}}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Everthing you wanted to know about PanDA but were truly truly afraid to ask}
-
-% author names and affiliations
-% use a multiple column layout for up to two different
-% affiliations
-
-\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-\and
-\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3},
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
- Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.
-Heterogeneous  resources are distributed worldwide, thousands of physicists analyzing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year.
-The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model.
-The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
-2013.
-We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
-\end{abstract}
-
-\begin{IEEEkeywords}
-TBD.
-\end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-% ----------------------------------------------------------------------------
-% I
-% ----------------------------------------------------------------------------
 \section{Introduction}\label{sec:intro}
@@
 \end{figure}
-
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
-  \item Why are WMS needed?
-  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
-  \item \ldots
-  \item Overview/Summary of paper.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% II
-% ----------------------------------------------------------------------------
-\section{Background}
-\label{sec:background}
-
-\begin{enumerate}
-  \item Brief overview of distributed CI.
-  \item Workload vs Workflow. (DONE)
-  \item Distinguish between WFMS and WLMS.
-  \begin{enumerate}
-    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
-  \end{enumerate}
-\end{enumerate}
-
---------------------------------------------------------------------
-
-\paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
-disciplines with different meaning. In the field of scientific computing,
-``workflow'' assumes different meanings depending on the characteristics of
-the computation, of the software tools used to support this computation, and
-of the resources on which it is performed. Further, a workflow may indicate a
-whole application, a description of the computational process of that
-application or, more commonly, a series of tasks related by data
-dependences.
-
-The lack of a consistent and shared definition of ``workflow'' hinders the
-understanding of its properties and its relations with related concepts. For
-example, we need to clarify the difference between ``workflow'', ``workload'',
-``task'', or ``job'' but also between workflow template and instance, or
-data-flow and control-flow. This clarification is precondition to specify
-properly the design of software systems that support the execution of
-scientific applications.
-
-In this paper, we use the following definitions:
-
-\begin{description}
-  \item[Task.] A set of operations to be performed on a computing platform,
-  alongside a description of the properties and dependences of those
-  operations, and indications on how they should be executed and satisfied.
-  \item[Workload.] A set of tasks that can be executed concurrently, possibly
-  related by a set of relations. For example, tasks of a workload can share
-  one or more input files or communicate during execution.
-  \item[Workflow.] Set of workloads, related by a set of relations that define
-  the order in which each workload can be executed. Data dependences are the
-  most common relations among workloads, used to define the precedence among
-  their executions. Note that, formally, a workload can have a single task.
-\end{description}
-
-Each task may have an arbitrary number of properties like number of cores,
-executables, or input/output files. Tasks may have precedence relations
-depending on their data dependences or any other type of dependency mandated
-by the application algorithms. Tasks with precedence relations have to be
-executed serially, otherwise tasks can be executed concurrently. A workload is
-defined as a set of tasks that can be executed concurrently. As such, a
-workflow can be composed by a set of workloads.
-
-The terms ``task'' and ``job'' are also used inconsistently across communities
-that perform scientific computing. In this paper, ``job'' refers to the entity
-that is submitted to a local resource management system (LRMS), like a batch
-system or a scheduler. As such, a task can become a job when is scheduled on a
-resource that exposes a LRMS but can also become a virtual machine or a
-container when bootstrapped on an infrastructure supporting virtualization.
-
-Usually, workflows are represented as graphs where tasks are vertices and
-their relations are edges~\cite{}. Often, graphs are supposed to be acyclic
-but graphs with cycles have been used to represent workflows of
-workflows~\cite{}. In this paper, a workflow template is a type of graph while
-a workflow instance is a type of graph with a specific set of vertices and
-edges. Further, a graph is ``abstract'' when no resource properties are
-available for all vertices, ``concrete'' otherwise~\cite{}.
-
-The distinction between an abstract and concrete graph highlights an important
-element of the execution process of a workflow. Workflows can be specified
-without indicating on what resources they should be executed. Accordingly, an
-abstract graph can be interpreted as a formal description of the requirements
-for the workflow execution. Once these requirements are matched by a set of
-resource capabilities, the graph becomes concrete, i.e. a formal description
-of the workflow execution process and its execution environment.
-
-% ----------------------------------------------------------------------------
-% III
-% ----------------------------------------------------------------------------
-\section{Related Work}
-\label{sec:related}
-\input{WMSLiterature}
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Other workload management systems (WMS) (not exclusive)
-  \begin{enumerate}
-    \item glidein-WMS
-    \item Dirac
-    \item ALICE-AlieN
-  \end{enumerate}
-{enumerate}
-  \item Comparison and contrast
-  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% IV
-% ----------------------------------------------------------------------------
-\section{PanDA Overview}\label{sec:panda_overview}
-PanDA consists of several interconnected subsystems, most of them are built from off-the-shelf and Open Source components. Figure \ref{fig:architecture} shows a schematic view of the PanDA system.
-In the following we will briefly describe system’s architecture and components.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
-\caption{Schematic view of the PanDA system\label{fig:architecture}}
-\end{center}
-\end{figure}
-\subsection{PanDA Server}
-The PanDA server is the main component of the system, It provides  a task queue managing all job information centrally.
-The PanDA  server receives jobs through the client interface into the task queue, upon which a brokerage module  operates to prioritize  and assign work on the basis of job type, priority, input data and its locality, and available CPU resources.
-The PanDA  server operates as a web service.
-It runs on Apache web server, interacting with the back-end database running  on separate servers.
-It uses the Apache worker  model, with many independent   processes  handling client requests  in  parallel.
-Since all  state is maintained  in  the central database,  the PanDA Server application instance itself is stateless.
-Currently production version of PanDA utilizes Oracle database backend, but PanDA  can also work with MySQL family of databases.
-
-\subsection{PanDA Pilot}
- PanDA is a pilot based workload  management system.
-In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload processing on a worker  node) are submitted to sites.
-When these  pilot jobs start on a  worker node they contact a central  server  to retrieve  a real payload (i.e., an end-user job) and execute it.
-Using these pilot-based  workflows helps to improve job reliability, optimize resource utilization, allows for opportunistic  resources usage, and mitigates  many of the problems  associated with the inhomogeneities found on the Grid.
-The pilots themselves  do not contain all the functionality needed to request a payload  job from the Panda Server.
-First, the complete  set of pilot code is downloaded via HTTP from a  central Subversion repository.
-The repository works with Apache web server configured  with a  memory-based   web proxy (Squid).
-The purpose  of the cache  is to reduce  the request load on the back-end Subversion  server.
-That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server.
-The benefits of this system include a high-performance code download service combined with code updates still being immediately  available  as soon as they are committed to source code control.
-
-\subsection{PanDA Factory}
-As a pilot-based  system, PanDA  requires some way to get the initial PanDA pilot onto worker nodes at sites.
-This is done with the help of a component  called AutoPyFactory (APF).
-APF runs in a single daemonized process, launching  a separate thread for each internal  workflow.
-Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote.
-The behavior of these APF workflows is determined by the combination of a set  of plugins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.
-
-\subsection{PanDA Monitoring}
-The PanDA Monitor is a web-based dashboard-style  graphical application.
-It runs on Apache and interacts with the back-end database for persistence.
-The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory.
-The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.
-
-\section{Evolution of PanDA WMS}
-While PanDa  currently uses  more than 100,000  cores at well over 100 Grid sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will  require more resources  than Grid computing can possibly provide.
-The Worldwide LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned analysis and data processing, but it will be insufficient  for Monte Carlo (MC) production and any extra activities.
-Additional computing and storage  resources are therefore required.
-To alleviate  these challenges, ATLAS is engaged  in an ambitious  program to expand  the current computing model to include  additional  resources such as the opportunistic   use of supercomputers   as well as commercial and academic clouds.
-In turn this activity drives evolution of the PanDA WMS.
-
-\subsection{Use of Supercomputers with PanDA}
-The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers.
-Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network.
-The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors.
-On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI).
-While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization.
-It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run.
-What matters is the total number of cycles that can be offloaded from the Grid.
-Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc.
-A reorganization  of the standard  workflow is therefore needed.
-The most suitable task for HPCs in HEP is event generation.
-Event generators are mostly computational,  stand-alone code, with few input requirements.
-There is no need  to stage-in much data.
-The event generation  in ATLAS corresponds  to
-10-15\% of all jobs on the Grid.
-A more difficult workload to adapt  for HPCs is detector  simulation.
-Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.
-However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
-\subsection{Interfacing PanDA with Titan}
-The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US.
-Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
-Integration with Titan is the current focus for PanDA developers.
-The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}.
-This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world.
-The pilot can communicate with the PanDA server from the front-end machine.
-The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job.
-ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
-1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design.
-Time allocation on an HPC is competitive and large projects are often preferred.
-About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources.
-The worker nodes
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
-\caption{Schematic view of the PanDA system\label{fig:interface}}
-\end{center}
-\end{figure}
-
---------------------------------------------------------------------
-\begin{enumerate}
-  \item PanDA WMS
-  \begin{enumerate}
-    \item State Diagrams
-    \item Control Flow Diagrams
-    \item Software Module and Architecture diagrams.
-  \end{enumerate}
-  \item PANDA Pilot
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% V
-% ----------------------------------------------------------------------------
-\section{Deploying PanDA on a Leadership-scale system}
-\label{sec:panda_deployment}
-
-\begin{enumerate}
-  \item Challenges
-  \item Opportunities
-  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
-  \item Take-aways, lessons learned
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VI
-% ----------------------------------------------------------------------------
-\section{Performance Characterization}
-\label{sec:panda_performance}
-
-\begin{enumerate}
-  \item Internal PanDA performance
-  \item External PanDA performance (workload specific on underlying system/platform)
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VII
-% ----------------------------------------------------------------------------
-\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
-\label{sec:panda_titan}
-
-\begin{enumerate}
-  \item Metadata performance issue, and how it’s resolved
-  \item File I/O performance and impact
-  \item ?
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VIII
-% ----------------------------------------------------------------------------
-\section{PANDA: Future/RoadMap}
-\label{sec:panda_roadmap}
-
- \section{Conclusion}\label{sec:conclusion}
-The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process,  the old batch job paradigm  of computing in HEP was  discarded  in favor of a  far more flexible and scalable  model.
-The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA  is the first exascale  workload management system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013.
-Next LHC run will pose massive computing  challenges.
-With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more.
-Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS.
-To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system.
-Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry.
-The technology base provided by the PanDA system will enhance the usage of a variety  of high-performance computing resources available to basic research.
-
-\input{ack}
-
-%\section*{Acknowledgment}
-
-%The authors would like to thank... More thanks here...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-
-\bibliographystyle{plain}
-\bibliography{bibliography}
-
-% that's all folks
-\end{document}

### 2016-10-04 paper.tex
basic restructuring
minor addition.
Splitting in input tex files, First Draft Planned Experiments

--- 2016-09-20 paper.tex
@@
 % ----------------------------------------------------------------------------
-\section{Introduction}\label{sec:intro}
-The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades.
-Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers.
-Thousands of physicists analyze petascale  data volumes daily.
-One of  the LHC experiments  the ATLAS  \cite{Aad:2008}  utilizes the PanDA  workload management system \cite{Maeno2011} (WMS) for distributed data processing and analysis.
-The ATLAS Computing model \cite{Atlas2005} is based on a Grid paradigm \cite{Foster:1998}, with multilevel, hierarchically distributed computing  and storage resources.
-Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
-PanDA has a highly scalable architecture.
-Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade.
-PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middleware).
-This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
-PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user analysis, while actively evolving to meet rapidly changing science needs.
-Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/DailyJobs.png}
-\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
-\end{center}
-\end{figure}
 
+\input{introduction}
 --------------------------------------------------------------------
@@
 % ----------------------------------------------------------------------------
-\section{Background}
-\label{sec:background}
 
+\input{Background}
 \begin{enumerate}
@@
 
-\paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
-disciplines with different meaning. In the field of scientific computing,
-``workflow'' assumes different meanings depending on the characteristics of
-the computation, of the software tools used to support this computation, and
-of the resources on which it is performed. Further, a workflow may indicate a
-whole application, a description of the computational process of that
-application or, more commonly, a series of tasks related by data
-dependences.
-
-The lack of a consistent and shared definition of ``workflow'' hinders the
-understanding of its properties and its relations with related concepts. For
-example, we need to clarify the difference between ``workflow'', ``workload'',
-``task'', or ``job'' but also between workflow template and instance, or
-data-flow and control-flow. This clarification is precondition to specify
-properly the design of software systems that support the execution of
-scientific applications.
-
-In this paper, we use the following definitions:
-
-\begin{description}
-  \item[Task.] A set of operations to be performed on a computing platform,
-  alongside a description of the properties and dependences of those
-  operations, and indications on how they should be executed and satisfied.
-  \item[Workload.] A set of tasks that can be executed concurrently, possibly
-  related by a set of relations. For example, tasks of a workload can share
-  one or more input files or communicate during execution.
-  \item[Workflow.] Set of workloads, related by a set of relations that define
-  the order in which each workload can be executed. Data dependences are the
-  most common relations among workloads, used to define the precedence among
-  their executions. Note that, formally, a workload can have a single task.
-\end{description}
-
-Each task may have an arbitrary number of properties like number of cores,
-executables, or input/output files. Tasks may have precedence relations
-depending on their data dependences or any other type of dependency mandated
-by the application algorithms. Tasks with precedence relations have to be
-executed serially, otherwise tasks can be executed concurrently. A workload is
-defined as a set of tasks that can be executed concurrently. As such, a
-workflow can be composed by a set of workloads.
-
-The terms ``task'' and ``job'' are also used inconsistently across communities
-that perform scientific computing. In this paper, ``job'' refers to the entity
-that is submitted to a local resource management system (LRMS), like a batch
-system or a scheduler. As such, a task can become a job when is scheduled on a
-resource that exposes a LRMS but can also become a virtual machine or a
-container when bootstrapped on an infrastructure supporting virtualization.
-
-Usually, workflows are represented as graphs where tasks are vertices and
-their relations are edges~\cite{}. Often, graphs are supposed to be acyclic
-but graphs with cycles have been used to represent workflows of
-workflows~\cite{}. In this paper, a workflow template is a type of graph while
-a workflow instance is a type of graph with a specific set of vertices and
-edges. Further, a graph is ``abstract'' when no resource properties are
-available for all vertices, ``concrete'' otherwise~\cite{}.
-
-The distinction between an abstract and concrete graph highlights an important
-element of the execution process of a workflow. Workflows can be specified
-without indicating on what resources they should be executed. Accordingly, an
-abstract graph can be interpreted as a formal description of the requirements
-for the workflow execution. Once these requirements are matched by a set of
-resource capabilities, the graph becomes concrete, i.e. a formal description
-of the workflow execution process and its execution environment.
-
 % ----------------------------------------------------------------------------
@@
 % ----------------------------------------------------------------------------
-\section{Related Work}
-\label{sec:related}
+
 \input{WMSLiterature}
@@
 \section{PanDA Overview}\label{sec:panda_overview}
-PanDA consists of several interconnected subsystems, most of them are built from off-the-shelf and Open Source components. Figure \ref{fig:architecture} shows a schematic view of the PanDA system.
-In the following we will briefly describe system’s architecture and components.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
-\caption{Schematic view of the PanDA system\label{fig:architecture}}
-\end{center}
-\end{figure}
-\subsection{PanDA Server}
-The PanDA server is the main component of the system, It provides  a task queue managing all job information centrally.
-The PanDA  server receives jobs through the client interface into the task queue, upon which a brokerage module  operates to prioritize  and assign work on the basis of job type, priority, input data and its locality, and available CPU resources.
-The PanDA  server operates as a web service.
-It runs on Apache web server, interacting with the back-end database running  on separate servers.
-It uses the Apache worker  model, with many independent   processes  handling client requests  in  parallel.
-Since all  state is maintained  in  the central database,  the PanDA Server application instance itself is stateless.
-Currently production version of PanDA utilizes Oracle database backend, but PanDA  can also work with MySQL family of databases.
-
-\subsection{PanDA Pilot}
- PanDA is a pilot based workload  management system.
-In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload processing on a worker  node) are submitted to sites.
-When these  pilot jobs start on a  worker node they contact a central  server  to retrieve  a real payload (i.e., an end-user job) and execute it.
-Using these pilot-based  workflows helps to improve job reliability, optimize resource utilization, allows for opportunistic  resources usage, and mitigates  many of the problems  associated with the inhomogeneities found on the Grid.
-The pilots themselves  do not contain all the functionality needed to request a payload  job from the Panda Server.
-First, the complete  set of pilot code is downloaded via HTTP from a  central Subversion repository.
-The repository works with Apache web server configured  with a  memory-based   web proxy (Squid).
-The purpose  of the cache  is to reduce  the request load on the back-end Subversion  server.
-That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server.
-The benefits of this system include a high-performance code download service combined with code updates still being immediately  available  as soon as they are committed to source code control.
-
-\subsection{PanDA Factory}
-As a pilot-based  system, PanDA  requires some way to get the initial PanDA pilot onto worker nodes at sites.
-This is done with the help of a component  called AutoPyFactory (APF).
-APF runs in a single daemonized process, launching  a separate thread for each internal  workflow.
-Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote.
-The behavior of these APF workflows is determined by the combination of a set  of plugins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.
-
-\subsection{PanDA Monitoring}
-The PanDA Monitor is a web-based dashboard-style  graphical application.
-It runs on Apache and interacts with the back-end database for persistence.
-The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory.
-The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.
-
-\section{Evolution of PanDA WMS}
-While PanDa  currently uses  more than 100,000  cores at well over 100 Grid sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will  require more resources  than Grid computing can possibly provide.
-The Worldwide LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned analysis and data processing, but it will be insufficient  for Monte Carlo (MC) production and any extra activities.
-Additional computing and storage  resources are therefore required.
-To alleviate  these challenges, ATLAS is engaged  in an ambitious  program to expand  the current computing model to include  additional  resources such as the opportunistic   use of supercomputers   as well as commercial and academic clouds.
-In turn this activity drives evolution of the PanDA WMS.
-
-\subsection{Use of Supercomputers with PanDA}
-The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers.
-Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network.
-The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors.
-On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI).
-While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization.
-It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run.
-What matters is the total number of cycles that can be offloaded from the Grid.
-Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc.
-A reorganization  of the standard  workflow is therefore needed.
-The most suitable task for HPCs in HEP is event generation.
-Event generators are mostly computational,  stand-alone code, with few input requirements.
-There is no need  to stage-in much data.
-The event generation  in ATLAS corresponds  to
-10-15\% of all jobs on the Grid.
-A more difficult workload to adapt  for HPCs is detector  simulation.
-Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.
-However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
-\subsection{Interfacing PanDA with Titan}
-The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US.
-Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
-Integration with Titan is the current focus for PanDA developers.
-The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}.
-This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world.
-The pilot can communicate with the PanDA server from the front-end machine.
-The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job.
-ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
-1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design.
-Time allocation on an HPC is competitive and large projects are often preferred.
-About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources.
-The worker nodes
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
-\caption{Schematic view of the PanDA system\label{fig:interface}}
-\end{center}
-\end{figure}
-
 --------------------------------------------------------------------
@@
 \end{enumerate}
+--------------------------------------------------------------------
+
+\input{titan}
+
+\input{PandaEvolution}
 
@@
 % ----------------------------------------------------------------------------
+
 \section{Deploying PanDA on a Leadership-scale system}
@@
 % ----------------------------------------------------------------------------
-\section{Performance Characterization}
-\label{sec:panda_performance}
+\input{perf}
 
-\begin{enumerate}
-  \item Internal PanDA performance
-  \item External PanDA performance (workload specific on underlying system/platform)
-\end{enumerate}
+%\section{Performance Characterization}
+%\label{sec:panda_performance}
+%
+%\begin{enumerate}
+%  \item Internal PanDA performance
+%  \item External PanDA performance (workload specific on underlying system/platform)
+%\end{enumerate}
 
@@
 % ----------------------------------------------------------------------------
-\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
-\label{sec:panda_titan}
+\input{overall}
 
-\begin{enumerate}
-  \item Metadata performance issue, and how it’s resolved
-  \item File I/O performance and impact
-  \item ?
-\end{enumerate}
+%\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
+%\label{sec:panda_titan}
+%
+%\begin{enumerate}
+%  \item Metadata performance issue, and how it’s resolved
+%  \item File I/O performance and impact
+%  \item ?
+%\end{enumerate}
 
@@
 % ----------------------------------------------------------------------------
+
 \section{PANDA: Future/RoadMap}
+\input{Experiments}
 \label{sec:panda_roadmap}

### 2016-10-04 Experiments.tex
Few notes on experiments

@@
-The execution of jobs on TITAN follows a three step scheme: i) as first, TITAN's Backfill queue communicates the presence of a new allocation i.e.
+The execution of jobs on TITAN follows a three step scheme: i) as first, TITAN's Backfill queue communicates the presence of a new allocation\mtnote{communicates the availability of idle resources?
+The term `allocation' may indicate the amount of resources available to the project.} i.e.
 a set of unused resources that can be used for a pre-determined time period\aanote{How many offers does PanDA receive per time unit?
 Average nodes per offer?}; ii) then, PanDA must accept or decline the allocation; iii) if the allocation is accepted, PanDA wraps a certain amount of events in a job and sends the job for execution.\aanote{I am supposing that the concept of event has been already introduced at this point}.
-At least in principle, PanDA can reject an allocation because it does not fit the requirements in terms of number of nodes or wall-time unit; however, this situation is rare because events are single core Athena-MP threads that require on average twenty minutes for being executed.
-Thus, the decomposition in event is fine grained once compared to the average allocation \aanote{Which correspond to......}.
+
+At least in principle, PanDA can reject an allocation because it does not fit the requirements in terms of number of nodes or wall-time unit; however, this situation is rare because events are single core Athena-MP threads\mtnote{Should we consider events inputs of Athena-MP jobs, with each event processed in a single thread by a---to be specified---module of Athena-MP?} that require on average twenty minutes for being executed.
+Thus, the decomposition in event is fine grained once compared to the average allocation \aanote{Which correspond to\ldots}.
 
@@
 On one hand, this execution strategy leads to a drastic reduction of the queue time and, consequently, it guarantees high-performances.
-On the other hand, this strategy is tailored for the execution of ATLAS workload on TITAN by using the Backfill queue.
+On the other hand, this strategy is tailored for the execution of ATLAS workload on TITAN by using the Backfill queue\mtnote{Do we have a queue dedicated to backfilling on Titan or should we speak of backfilling done on one or more queues?}.
 
 Questions such as \emph{``Is this the best strategy possible?''} or \emph{``Can PanDA achieve similar performances by using TITAN's batch queue?''}  remain unanswered.
+
 For this reason, in the remaining of this section we propose experimental results to address them. The idea behind the experiments is to show the behaviour of ATLAS workload on TITAN's batch queue.
-As first, we investigate resource allocation as a function of time (wall-time) and space (number of nodes) in order to understand how TITAN's scheduler behaves under different allocation requests.
+
+As first, we investigate resource allocation as a function of time (wall-time) and space (number of nodes\mtnote{cores?}) in order to understand how TITAN's scheduler and workload behave under different allocation requests.
 This analysis includes two bulks of experiments: the first, generated ``in silico'', is based on simulations of the TITAN's scheduler by means of the Moab simulator, whereas the second, generated ``in vivo'',  considers real submissions of ATLAS workload on TITAN's batch queue.

### 2016-10-04 Experiments.tex
minor comments and fixes.

@@
+\jhanote{Not sure this should be under ``PANDA:Future/ROAD MAP?''}
+
 The execution of jobs on TITAN follows a three step scheme: i) as first, TITAN's Backfill queue communicates the presence of a new allocation\mtnote{communicates the availability of idle resources?
@@
 
-On one hand, this execution strategy leads to a drastic reduction of the queue time and, consequently, it guarantees high-performances.
+On one hand, this \jhanote{I would use the word execution strategy carefully.
+Maybe a more general term such as execution model, a special type of which is execution strategy?} execution strategy leads to a drastic reduction of the queue time and, consequently, it guarantees high-performances.
 On the other hand, this strategy is tailored for the execution of ATLAS workload on TITAN by using the Backfill queue\mtnote{Do we have a queue dedicated to backfilling on Titan or should we speak of backfilling done on one or more queues?}.
@@
 
-As first, we investigate resource allocation as a function of time (wall-time) and space (number of nodes\mtnote{cores?}) in order to understand how TITAN's scheduler and workload behave under different allocation requests.
+First, we investigate resource allocation as a function of time (wall-time) and space (number of nodes\mtnote{cores?}) in order to understand how TITAN's scheduler and workload behave under different allocation requests.
 This analysis includes two bulks of experiments: the first, generated ``in silico'', is based on simulations of the TITAN's scheduler by means of the Moab simulator, whereas the second, generated ``in vivo'',  considers real submissions of ATLAS workload on TITAN's batch queue.

### 2016-10-04 Experiments.tex
Second Draft Experiments

--- 2016-10-04 Experiments.tex
@@
-The execution of jobs on TITAN follows a three step scheme: i) as first, TITAN's Backfill queue communicates the presence of a new allocation i.e.
-a set of unused resources that can be used for a pre-determined time period\aanote{How many offers does PanDA receive per time unit?
-Average nodes per offer?}; ii) then, PanDA must accept or decline the allocation; iii) if the allocation is accepted, PanDA wraps a certain amount of events in a job and sends the job for execution.\aanote{I am supposing that the concept of event has been already introduced at this point}.
-At least in principle, PanDA can reject an allocation because it does not fit the requirements in terms of number of nodes or wall-time unit; however, this situation is rare because events are single core Athena-MP threads that require on average twenty minutes for being executed.
-Thus, the decomposition in event is fine grained once compared to the average allocation \aanote{Which correspond to......}.
+%\jhanote{Not sure this should be under ``PANDA:Future/ROAD MAP?'' (cut and paste error it was not meant to be; FIXED)}
+\section{Experiments}
+The execution of jobs on TITAN follows a three step scheme: i) as first, TITAN's Backfill queue communicates  the availability of a slot i.e.
+a set of idle cores for a given time period\aanote{How many offers does PanDA receive per time unit?
+Average nodes per offer?}; ii) then, PanDA can decide to run on the slot or wait until a larger slot becomes available; iii) if the slot is taken, PanDA wraps a certain amount of events in a job and sends the job for execution.\aanote{I am supposing that the concept of event has been already introduced at this point}.
 
-As a consequence, PanDA accepts the major part of the proposed allocations by selecting a number of events that maximize the utilization of the resources during the assigned time interval.
+At least in principle, PanDA can reject the slot because it does not fit the requirements in terms of number of cores or wall-time unit; however, this situation is rare because events are single core Athena-MP threads\mtnote{Should we consider events inputs of Athena-MP jobs, with each event processed in a single thread by a---to be specified---module of Athena-MP?} that require on average twenty minutes for being executed.
+Thus, the decomposition in event is fine grained once compared to the average allocation \aanote{Which correspond to\ldots}.
 
-On one hand, this execution strategy leads to a drastic reduction of the queue time and, consequently, it guarantees high-performances.
-On the other hand, this strategy is tailored for the execution of ATLAS workload on TITAN by using the Backfill queue.
+As a consequence, PanDA accepts the major part of the proposed slots by selecting a number of events that maximize the utilization of the resources during the assigned time interval.
 
-Questions such as \emph{``Is this the best strategy possible?''} or \emph{``Can PanDA achieve similar performances by using TITAN's batch queue?''}  remain unanswered.
+On one hand, this %\jhanote{I would use the word execution strategy carefully. Maybe a more general term such as execution model, a special type of which is execution strategy?(I Agree, FIXED)} 
+execution model leads to a drastic reduction of the queue time and, consequently, it guarantees high-performances.
+On the other hand, this model is tailored for the execution of ATLAS workload on TITAN by using the Backfill queue\mtnote{Do we have a queue dedicated to backfilling on Titan or should we speak of backfilling done on one or more queues?}.
+
+Questions such as \emph{``Is this the best execution model?''} or \emph{``Can PanDA achieve similar performances by using TITAN's batch queue?''}  remain unanswered.
+
 For this reason, in the remaining of this section we propose experimental results to address them. The idea behind the experiments is to show the behaviour of ATLAS workload on TITAN's batch queue.
-As first, we investigate resource allocation as a function of time (wall-time) and space (number of nodes) in order to understand how TITAN's scheduler behaves under different allocation requests.
-This analysis includes two bulks of experiments: the first, generated ``in silico'', is based on simulations of the TITAN's scheduler by means of the Moab simulator, whereas the second, generated ``in vivo'',  considers real submissions of ATLAS workload on TITAN's batch queue.
 
-By using the batch queue, we move in a situation where the queue time becomes dominant but, at the same time, we gain more freedom to decide the parameters of the allocation.
-For this reason, the second set of the experiments aims to find sub-optimal parameters for resource allocation.
+First, we investigate the acquisition of slot as a function of time (wall-time) and space (number of cores) in order to understand how TITAN's scheduler and workload behave under different slot requests.
+This analysis includes two bulks of experiments: the first, generated ``in silico'', is based on simulations of the TITAN's scheduler by means of the Moab simulator \cite{}, whereas the second, generated ``in vivo'',  considers real submissions of ATLAS workload on TITAN's batch queue.
 
-Additionally, due to the dynamics of the Backfill queue, PanDA has a high probability to re-acquire an allocation immediately after it has released one \aanote{Are we able to quantify this ``immediately''?}.
+By using the batch queue, we move in a situation where the queue time becomes dominant but, at the same time, we have more freedom to decide the parameters of the slot.
+For this reason, the second set of the experiments aims to find sub-optimal parameters with which we can minimize the trade-off between the size of a slot and the time spent in queue waiting for that slot to become available.
+In other words, we aim to minimize the completion time by finding the best trade-off between execution time and queue time.
+
+This execution model introduces slot utilization as one of the key factors for high-performances.
+This happens because, in order to minimize the time spent in queue, we might asks for slots in advance and, then we could not br able to saturate them when they become available.
+Thus, this strategy requires a new functionality that allows the job to receive and execute new events while it is already running on the resources.
+In order to do that we perform the experiments by using a new generation executor that implements such functionality.
+
+As last observation, it is important to point out that the percentage of utilization of a slot is minor problem with the current implementation because, due to the dynamics of the Backfill queue, PanDA has a high probability to re-acquire a slot immediately after it has released one\aanote{Are we able to quantify this ``immediately''?}.

### 2016-10-04 background.tex
Splitting in input tex files, First Draft Planned Experiments

--- 2016-10-04 Background.tex
@@
-\section{Background}
-\label{sec:background}
 \paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
 disciplines with different meaning. In the field of scientific computing,
-``workflow'' assumes different meanings depending on the characteristics of
-the computation, of the software tools used to support this computation, and
-of the resources on which it is performed. Further, a workflow may indicate a
-whole application, a description of the computational process of that
-application or, more commonly, a series of tasks related by data
-dependences.
+``workflow'' assumes different meanings depending on the characteristics of the
+computation, of the software tools used to support this computation, and of the
+resources on which it is performed. Further, a workflow may indicate a whole
+application, a description of the computational process of that application or,
+more commonly, a series of tasks related by data dependences.
 
@@
 understanding of its properties and its relations with related concepts. For
-example, we need to clarify the difference between ``workflow'', ``workload'',
+example, we need to clarify the difference among ``workflow'', ``workload'',
 ``task'', or ``job'' but also between workflow template and instance, or
-data-flow and control-flow. This clarification is precondition to specify
-properly the design of software systems that support the execution of
-scientific applications.
+data-flow and control-flow. This is precondition to specify properly the design
+of software systems that support the execution of scientific applications.
 
@@
 \begin{description}
+
   \item[Task.] A set of operations to be performed on a computing platform,
-  alongside a description of the properties and dependences of those
-  operations, and indications on how they should be executed and satisfied.
+  alongside a description of the properties and dependences of those operations,
+  and indications on how the dependences should be satisfied and the operations
+  should be executed.
+
+  \item[Job.] A unit of execution that performs one or more unit of work. Jobs
+  relates to the resource on which they are executed. One or more tasks can be
+  the units of work executed by a job.
+
   \item[Workload.] A set of tasks that can be executed concurrently, possibly
-  related by a set of relations. For example, tasks of a workload can share
-  one or more input files or communicate during execution.
+  related by a set of relations. For example, tasks of a workload can share one
+  or more input files or communicate during execution.
+
   \item[Workflow.] Set of workloads, related by a set of relations that define
@@
   their executions. Note that, formally, a workload can have a single task.
+
 \end{description}
@@
 The terms ``task'' and ``job'' are also used inconsistently across communities
-that perform scientific computing. In this paper, ``job'' refers to the entity
-that is submitted to a local resource management system (LRMS), like a batch
-system or a scheduler. As such, a task can become a job when is scheduled on a
-resource that exposes a LRMS but can also become a virtual machine or a
+that perform scientific computing. In this paper, ``job'' refers to the unit of
+work that is submitted to a local resource management system (LRMS), like a
+batch system or a scheduler. As such, a task can become a job when is scheduled
+on a resource that exposes a LRMS but can also become a virtual machine or a
 container when bootstrapped on an infrastructure supporting virtualization.
+Tasks can be grouped into jobs, statically or dynamically, depending on the
+resource capabilities and the task requirements.
 
-Usually, workflows are represented as graphs where tasks are vertices and
-their relations are edges~\cite{}. Often, graphs are supposed to be acyclic
-but graphs with cycles have been used to represent workflows of
-workflows~\cite{}. In this paper, a workflow template is a type of graph while
-a workflow instance is a type of graph with a specific set of vertices and
-edges. Further, a graph is ``abstract'' when no resource properties are
-available for all vertices, ``concrete'' otherwise~\cite{}.
+Usually, workflows are represented as graphs in which tasks are vertices and
+relations are edges~\cite{}. Often, graphs are supposed to be acyclic but graphs
+with cycles have been used to represent workflows of workflows~\cite{}. In this
+paper, a workflow template is a type of graph while a workflow instance is a
+type of graph with a specific set of vertices and edges. Further, a graph is
+``abstract'' when no resource properties are available for all vertices,
+``concrete'' otherwise~\cite{}.
 
@@
 for the workflow execution. Once these requirements are matched by a set of
-resource capabilities, the graph becomes concrete, i.e. a formal description
+resource capabilities, the graph becomes concrete, i.e., a formal description
 of the workflow execution process and its execution environment.

### 2016-10-05 titan.tex
titan text

\subsection{OLCF and Titan}

In 2004, the Oak Ridge Leadership Computing Facility (OLCF) was established at
Oak Ridge National Laboratory (ORNL).The  OLCF's is mission to accelerate the
scientific discovery and engineering progress by providing outstanding
computing and data management resources to high-priority research and
development projects.

Titan, a leadership-class and the fastest open-science supercomputer in the
U.S., is located at OLCF. Titan enables scientists to evaluate and assess various
complex physical phenomena via large-scale computational simulations.

Titan is a Cray XK7 system with 18,688 compute nodes~\cite{top500}. Each
compute node is equipped with an AMD Opteron CPU and a Nvidia Tesla GPU. This
hybrid design provides improved energy efficiency, as well as an order of
magnitude in computational capacity over its predecessor. Titan has 710 TB of
total system memory. This leadership capability compute system is backed by a
center-wide parallel file system known as Spider II~\cite{spider2}. Spider II
is one of the world's fastest and largest POSIX complaint parallel file
systems. It is designed to serve write-heavy I/O workloads generated by Titan
compute clients and other OLCF resources.

The topology and architecture details of Spider II infrastructure are
illustrated in Figure~\ref{fig:xk7-compute-node} and
described as follows:

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/spider2arch.pdf}}\\
    \end{tabular}
    \caption{Infrastructure and I/O path between Titan (Cray XK7) and its backend storage}
    \label{fig:xk7-compute-node}
\end{figure}

Spider II is based on the Lustre technology~\cite{lustre}.  It is configured
and deployed as two independent, non-overlapping file systems, each with 144
Lustre Object Storage Servers (OSSs) and 1,008 Lustre Object Storage Targets
(OSTs). Spider II offers ~30 PB usable file system capacity to OLCF systems and
users and has over 1 TB/s aggregate write I/O performance.

### 2016-10-05 overall.tex
overall section
initial drop on atlas mds performance improvement

@@
 
-\subsection{PanDA Metadata Peformance Impact at OLCF}
+\subsection{PanDA Metadata Performance Impact at OLCF}
 
-\begin{enumerate}
-  \item Metadata performance issue, and how it’s resolved
-  \item File I/O performance and impact
-  \item ?
-\end{enumerate}
+The ATLAS framework has a configuration and initialization stage, which is
+called Athena. At this stage, the running job is assembled  on the fly from a
+large number of shared libraries.  Also, at this stage, every algorithm and
+service is being configured, at run time, by a corresponding set of Python
+scripts, which results in a large number of read I/O accesses to small python
+scripts, with many includes and imports Python calls.
+
+When ATLAS was originally deployed at the OLCF, all shared libraries were
+placed at the Spider 2 Lustre file system. However, it was quickly discovered
+that the I/O pattern at the Athena stage was detrimental to the overall Spider
+2 performance. Since Spider 2 is a center-wide file system and shared by all
+OLCF resources and users, this resulted in lower overall interactive and
+metadata performance at OLCF. As can be seen in Listing~\ref{mdstrace} the
+metadata I/O activity for ATLAS exhibits a spike corresponding to the beginning
+of the runs (i.e., the Athena stage) before tapering off.
+
+\begin{lstlisting}[language=bash,frame=single,basicstyle=\ttfamily\tiny,caption=ATLAS metadata trace,label=mdstrace] 
+XK7 Application 9205593
+      39012 RPCs from 300 of 300 nodes
+        ~69291.96 per sec
+          37851 LDLM_ENQUEUE RPCs    ~67229.83 per sec
+                pmin 13us pavg 42us pmax 4983us
+          611 LDLM_CANCEL RPCs    ~1085.24 per sec
+                pmin 10us pavg 16us pmax 32us
+          277 MDS_CLOSE RPCs    ~492.00 per sec
+                pmin 15us pavg 20us pmax 38us
+          154 MDS_READPAGE RPCs    ~273.53 per sec
+                pmin 170us pavg 292us pmax 671us
+          86 MDS_GETXATTR RPCs    ~152.75 per sec
+                pmin 15us pavg 21us pmax 65us
+          30 MDS_GETATTR RPCs    ~53.29 per sec
+                pmin 16us pavg 20us pmax 27us
+          3 MDS_REINT RPCs    ~5.33 per sec
+                pmin 103us pavg 136us pmax 196us
+          Overall times
+                pmin 10us pavg 42us pmax 4983us
+
+XK7 Application 9205355
+      8698 RPCs from 62 of 62 nodes
+        ~15449.13 per sec
+          8445 LDLM_ENQUEUE RPCs    ~14999.76 per sec
+                pmin 16us pavg 41us pmax 780us
+          92 MDS_CLOSE RPCs    ~163.41 per sec
+                pmin 15us pavg 23us pmax 52us
+          55 MDS_READPAGE RPCs    ~97.69 per sec
+                pmin 189us pavg 291us pmax 534us
+          52 LDLM_CANCEL RPCs    ~92.36 per sec
+                pmin 12us pavg 16us pmax 25us
+          41 MDS_GETXATTR RPCs    ~72.82 per sec
+                pmin 16us pavg 20us pmax 27us
+          13 MDS_GETATTR RPCs    ~23.09 per sec
+                pmin 16us pavg 21us pmax 38us
+          Overall times
+                pmin 12us pavg 42us pmax 780us
+
+\end{lstlisting}
+
+As this problem was identified, the OLCF staff has enabled read-only access to
+certain NFS-exported directories from Titan compute nodes. This in turn allowed
+the OLCF staff to install a software package from a Titan login node and have
+it available read-only on a Titan compute node.
+
+After moving the ATLAS release on Lustre to the NFS-exported read-only
+directories, the metadata contention problem has resolved and a two orders of
+magnitude metadata performance improvement (i.e., ~6,300 seconds on Lustre vs.
+~1,500 seconds on NFS) has been observed for ATLAS.  Also, for both cases the
+input file is placed on memory disk on a Titan compute nodes and for Lustre the
+validation step took on average 1,320 seconds, while it was 40 seconds on NFS,
+showing a three orders of magnitude performance improvement.
+
+Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
+improvement on Titan. The circled region illustrates the switch from Lustre to
+NFS-exported directory for hosting the ATLAS release.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{tabular}{cc}
+        {\includegraphics[width=0.48\textwidth]{figures/panda-completed-jobs-sw-move.pdf}}\\
+    \end{tabular}
+    \caption{ATLAS performance improvement on Titan. The circled region shows the switch from Lustre to NFS-exported directory for hosting the ATLAS release.}
+    \label{fig:atlas-perf-improvement}
+\end{figure}
+
+\subsection{PanDA File I/O Performance Impact at OLCF}
+
+File I/O performance and impact

### 2016-11-08 bckground.tex
Still issues with capitalizations
Iteration background, preliminary work on workflow

--- 2016-10-04 background.tex
@@
 example, we need to clarify the difference among ``workflow'', ``workload'',
-``task'', or ``job'' but also between workflow template and instance, or
-data-flow and control-flow. This is precondition to specify properly the design
-of software systems that support the execution of scientific applications.
+``task'', or ``job'' but also between workflow ``template'' and ``instance'', or
+``data-flow'' and ``control-flow''. This is precondition to specify properly the
+design of software systems that support the execution of scientific
+applications.
 
@@
 Each task may have an arbitrary number of properties like number of cores,
-executables, or input/output files. Tasks may have precedence relations
+executables, or input/output files. Tasks may have precedence interrelations,
 depending on their data dependences or any other type of dependency mandated
 by the application algorithms. Tasks with precedence relations have to be
-executed serially, otherwise tasks can be executed concurrently. A workload is
+executed serially; otherwise, tasks can be executed concurrently. A workload is
 defined as a set of tasks that can be executed concurrently. As such, a
-workflow can be composed by a set of workloads.
+workflow can be composed of a set of workloads.
 
@@
 that perform scientific computing. In this paper, ``job'' refers to the unit of
-work that is submitted to a local resource management system (LRMS), like a
-batch system or a scheduler. As such, a task can become a job when is scheduled
-on a resource that exposes a LRMS but can also become a virtual machine or a
-container when bootstrapped on an infrastructure supporting virtualization.
-Tasks can be grouped into jobs, statically or dynamically, depending on the
-resource capabilities and the task requirements.
+work that is submitted to a local resource management system (LRMS), like the
+Slurm or PBS batch system of a cluster. As such, a task can become a job when
+is scheduled on a resource that exposes a LRMS but can also become a virtual
+machine or a container when bootstrapped on an infrastructure supporting
+virtualization. Tasks can be statically or dynamically grouped into jobs,
+depending on the resource capabilities and the task requirements.
 
@@
 paper, a workflow template is a type of graph while a workflow instance is a
-type of graph with a specific set of vertices and edges. Further, a graph is
-``abstract'' when no resource properties are available for all vertices,
-``concrete'' otherwise~\cite{}.
+workflow template with specific vertices and edges. Further, a workflow instance
+is ``abstract'' when no resource properties are available for all vertices,
+``concrete'' otherwise~\cite{}. When not qualified, the term ``workflow''
+indicates a abstract workflow instance.
 
-The distinction between an abstract and concrete graph highlights an important
-element of the execution process of a workflow. Workflows can be specified
-without indicating on what resources they should be executed. Accordingly, an
-abstract graph can be interpreted as a formal description of the requirements
-for the workflow execution. Once these requirements are matched by a set of
-resource capabilities, the graph becomes concrete, i.e., a formal description
-of the workflow execution process and its execution environment.
+The distinction between an abstract and concrete workflow instances highlights
+an important element of the execution process of a workflow. Abstract workflows
+can be interpreted as a formal description of the requirements for the workflow
+execution, while concrete workflows can be interpreted as a formal description of the workflow's execution process.
+The process of resource selection matches the resource capabilities to the workflow requirements making possible the transition from execution requirements to execution process.
+
+Workload managers are those components of workflow systems or, more in general, of distributed applications that enable resource selection.
+
+\subsection{ATLAS Workflow}
+
+In this paper we focus on the ATLAS Monte Carlo workflow\mtnote{Are there other
+workflows for ATLAS?} and its evolving processes of resource selection.
+Traditionally, the ATLAS workflow has been based on single-core tasks, executed
+on Grid resources to process large amount (2TB) of input data divided into
+discrete I/O units called ``events''. This approach has many merits: (i)
+simplified process of resource selection under the assumption of homogeneous
+resources and execution environments; (ii) opportunistic distribution of compute
+tasks across resources based on contingent availability; (iii) relative
+robustness due to the ability of re-executing failed tasks; and (iv)
+multi-tiered distribution and storage of events' data.
+
+Despite all the advantages,
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figures/atlas_workflow.pdf}
+  \caption{ATLAS Workflow}
+\label{fig:atlas_workflow}
+\end{figure}

### 2016-11-08 bckground.tex

@@
-\paragraph*{Workload Vs Workflow} The term ``workflow'' is used in many
-disciplines with different meaning. In the field of scientific computing,
-``workflow'' assumes different meanings depending on the characteristics of the
-computation, of the software tools used to support this computation, and of the
-resources on which it is performed. Further, a workflow may indicate a whole
-application, a description of the computational process of that application or,
-more commonly, a series of tasks related by data dependences.
+\subsection{Workflows and Workloads}
+
+The term ``workflow'' is used in many disciplines with different meaning. In the
+field of scientific computing, ``workflow'' assumes different meanings depending
+on the characteristics of the computation, of the software tools used to support
+this computation, and of the resources on which it is performed. Further, a
+workflow may indicate a whole application, a description of the computational
+process of that application or, more commonly, a series of tasks related by data
+dependences.
 
@@
 
+\subsection{The ATLAS Monte Carlo Workflow}
+
 The distinction between an abstract and concrete workflow instances highlights
@@
 
-Workload managers are those components of workflow systems or, more in general, of distributed applications that enable resource selection.
+Workload managers are those components of workflow systems or, more in general,
+of distributed applications that enable resource selection, acquisition, and
+use. The set of capabilities required by a workload manager mostly depends on
+the heterogeneity and dynamicity of both the workflow and the resources.
+Heterogeneity is a measure of the diversity of the workflow requirements and
+resource capabilities that need to be matched to enable the execution of the
+workflow's tasks; dynamicity is a measure of how much these requirements and
+capabilities varies during execution.
 
-\subsection{ATLAS Workflow}
+The PanDA workload management system was initially designed to manage the execution of static, relatively homogeneous tasks on dynamic but homogeneous resources.
+Progressively, PanDA has evolved to enable the management of more heterogeneous tasks and resources.
+In this paper we describe how the design and architecture of PanDA has enabled executions of multi-threaded tasks on Titan, currently the largest high performance computing (HPC) resource available in the USA for scientific research.
 

### 2016-11-18 WMSLiterature.tex
Splitting in input tex files, First Draft Planned Experiments
minor re-structuring.

@@
-\subsection{Alien}
+\subsection{Related Work}
+\label{sec:related}
+\subsubsection{Alien}
 Alien is a Workload and Data Management system composed of a 	set of middleware tools and services entirely based on web-services and standard protocols.
@@
 
-\subsection{DIRAC} 
+\subsubsection{DIRAC} 
 DIRAC (Distributed Infrastructure with Remote Agent Control) Workload and Data Management System has been developed within the CERN LHCb project to manage the processing of detector data, Monte Carlo simulations, and end-user analyses \cite{Tsaregorodtsev2004}.
@@
 JDL and \emph{Transformation Management System} (TMS) are used for task description \cite{Tsaregorodtsev2006,Pacini2006}. The latter allows the user to describe complex data dependences. 
-\subsection{HTCondor Glidein and GlideinWMS}
+\subsubsection{HTCondor Glidein and GlideinWMS}
 The HTCondor Glidein system  as part of the HTCondor software ecosystem. The HTCondor Glidein is a pilot based system to aggregate DCRs with heterogeneous middleware into HTCondor resource pools.
-Condor is based on daemons collaborating by exchanging messages over the network. We can isolate four main logical components:
+Condor is based on daemons collaborating by exchanging messages over the network. We can isolate four main logical components \cite{Sfiligoi2008}:
 \begin{itemize}

### 2016-11-29 bckground.tex
Minor iteration

@@
 can be interpreted as a formal description of the requirements for the workflow
-execution, while concrete workflows can be interpreted as a formal description of the workflow's execution process.
-The process of resource selection matches the resource capabilities to the workflow requirements making possible the transition from execution requirements to execution process.
+execution, while concrete workflows can be interpreted as a formal description
+of the workflow's execution process. The process of resource selection matches
+the resource capabilities to the workflow requirements making possible the
+transition from execution requirements to execution process.
 
@@
 
-The PanDA workload management system was initially designed to manage the execution of static, relatively homogeneous tasks on dynamic but homogeneous resources.
-Progressively, PanDA has evolved to enable the management of more heterogeneous tasks and resources.
-In this paper we describe how the design and architecture of PanDA has enabled executions of multi-threaded tasks on Titan, currently the largest high performance computing (HPC) resource available in the USA for scientific research.
+The PanDA workload management system was initially designed to manage the
+execution of static, relatively homogeneous tasks on dynamic but homogeneous
+resources. Progressively, PanDA has evolved to enable the management of more
+heterogeneous tasks and resources. In this paper we describe how the design and
+architecture of PanDA has enabled executions of multi-threaded tasks on Titan,
+currently the largest high performance computing (HPC) resource available in the
+USA for scientific research.
 
-In this paper we focus on the ATLAS Monte Carlo workflow\mtnote{Are there other
-workflows for ATLAS?} and its evolving processes of resource selection.
 Traditionally, the ATLAS workflow has been based on single-core tasks, executed
@@
 
-Despite all the advantages,
+Despite all the advantages, ATLAS workflows have evolved to include
+compute-intense tasks that can benefit from multithreaded execution on HPC
+machines. The Monte Carlo workflow~\cite{} is one of the largest workflows
+executed by the ATLAS experiment and at least one of its stages can benefit from
+large amount of parallelism. The Geant4 toolkit is used to simulate the passage
+of particles through matter and\ldots
 
@@
   \includegraphics[width=\columnwidth]{figures/atlas_workflow.pdf}
-  \caption{ATLAS Workflow}
+  \caption{ATLAS Monte Carlo workflow.}
 \label{fig:atlas_workflow}

### 2016-11-29 overall.tex
data excel sheet
latest version

@@
 File I/O performance and impact
+
+\begin{figure}[!htb]
+    \centering
+    \begin{tabular}{cc}
+        {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_hist.pdf}}\\
+    \end{tabular}
+    \caption{ATLAS file read I/O histogram on Titan for week of 10/25/16.}
+    \label{fig:atlas-titan-io-read}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{tabular}{cc}
+        {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_hist.pdf}}\\
+    \end{tabular}
+    \caption{ATLAS file write I/O histogram on Titan for week of 10/25/16.}
+    \label{fig:atlas-titan-io-written}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{tabular}{cc}
+        {\includegraphics[width=0.48\textwidth]{figures/panda_file_open_hist.pdf}}\\
+    \end{tabular}
+    \caption{ATLAS file open() histogram on Titan for week of 10/25/16.}
+    \label{fig:atlas-titan-file-open}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{tabular}{cc}
+        {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
+    \end{tabular}
+    \caption{ATLAS file close() histogram on Titan for week of 10/25/16.}
+    \label{fig:atlas-titan_file_close}
+\end{figure}
+
+\begin{table*}[t]
+\centering
+\begin{tabular}{lllllllll}
+ & Num. Nodes & Duration (s) & Read (GB) & Written (GB) & GB Read/nodes & GB Written/nodes & open() & close() \\
+Min & 1 & 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\
+Max & 300 & 7,452 & 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\
+Average & 35.66 & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 & 34,155.74 \\
+Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376 & 231,346.55 & 53,799.08
+\end{tabular}
+\caption{The Statistical breakdown of the I/O impact of 1,175 PanDA jobs executed at OLCF for the week of 10/25/16}
+\label{panda-olcf-stats}
+\end{table*}

### 2016-12-05 overall.tex
finishing the write up for the I/O impact section

@@
 
-\subsection{PanDA Metadata Performance Impact at OLCF}
+\subsection{PanDA Shared Library I/O Performance Impact at OLCF}
 
@@
 that the I/O pattern at the Athena stage was detrimental to the overall Spider
-2 performance. Since Spider 2 is a center-wide file system and shared by all
-OLCF resources and users, this resulted in lower overall interactive and
-metadata performance at OLCF. As can be seen in Listing~\ref{mdstrace} the
-metadata I/O activity for ATLAS exhibits a spike corresponding to the beginning
-of the runs (i.e., the Athena stage) before tapering off.
+2 file system performance. Since Spider 2 is a center-wide file system and
+shared by all OLCF resources and users, this resulted in lower overall
+interactive and metadata performance at OLCF. As can be seen in
+Listing~\ref{mdstrace} the metadata I/O activity for ATLAS exhibits a spike
+corresponding to the beginning of the runs (i.e., the Athena stage) before
+tapering off.
 
+\begin{minipage}{\linewidth}
 \begin{lstlisting}[language=bash,frame=single,basicstyle=\ttfamily\tiny,caption=ATLAS metadata trace,label=mdstrace] 
@@
 \end{lstlisting}
+\end{minipage}
 
@@
 
-\subsection{PanDA File I/O Performance Impact at OLCF}
+\subsection{PanDA I/O Impact at OLCF}
 
-File I/O performance and impact
+To better understand the I/O impact of ATLAS PanDA project on Titan
+supercomputing environment we analyzed 1,175 jobs ran on the week of
+10/25/2016, for a total of 174 hours. Table~\ref{panda-olcf-stats} shows the
+overall statistical breakdown of the observed file I/O impact of ATLAS at OLCF.
+Figures \ref{fig:atlas-titan-io-read} and \ref{fig:atlas-titan-io-written} show
+the file read and write I/O histograms for these 1,175 jobs, respectively.
+Figures \ref{fig:atlas-titan-file-open} and \ref{fig:atlas-titan-file-close}
+show the file open() and close() metadata load histograms of the same 1,175
+ATLAS jobs, respectively.
+
+As can be seen from Table~\ref{panda-olcf-stats}, the number of nodes used by
+ATLAS jobs vary between 1 and 300, while the average is at 35. 75\% of the
+ATLAS jobs consume less than 25 and 92\% consume less than or equal to 100
+Titan compute nodes. During the 174 hours of data collection, we observed that
+6.75 ATLAS jobs were executed on average per hour on Titan and ran for
+1.74 hours on average.
+
+ATLAS jobs are file read I/O heavy, as can be seen in
+Table~\ref{panda-olcf-stats}. The maximum amount read by any ATLAS job in
+aggregate in this observed period was less than 250 GB and the maximum amount
+written in aggregate was less than 75 GB. The average amount read per job is 20
+GB and average amount written is 6 GB.
 
@@
 
+Per job read and write file I/O statistics show an interesting pattern.
+Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
+compute node on Titan is less than 400 MB on average, while the amount of data
+written per node is less than 170 MB on average. This correlates with our
+finding that ATLAS PanDA jobs are read heavy. However, as can be seen in
+Table~\ref{panda-olcf-stats} and figures {fig:atlas-titan-io-read} and
+{fig:atlas-titan-io-written}, the distribution between read and written amount
+of data per job is quite different from one another. The read I/O distribution
+per job shows a long tail, ranging from 25 GB to 250 GB, while the written
+amount of data has a very narrow distribution.
+
 \begin{figure}[!htb]
@@
     \caption{ATLAS file close() histogram on Titan for week of 10/25/16.}
-    \label{fig:atlas-titan_file_close}
+    \label{fig:atlas-titan-file-close}
 \end{figure}
+
+On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file open() operations
+and 5 file close() operations per second. As can be seen from figures
+\ref{fig:atlas-titan-file-open} and \ref{fig:atlas-titan-file-close}, they
+exhibit similar distributions. The maximum number of file open() operations are
+around 170 on average and the maximum number of file close() operations are 39
+on average per second per job. The total number of file open() operations is
+172,089,760 for the observed window of 174 hours, while the total number of
+file close() operations stand at 40,132,992 for the total 1,175 ATLAS PanDA
+jobs in the same observation window. The difference between these two values is
+puzzling and it is under investigation at the time of writing this paper.
+{\sarp{can someone take a look at this? I would expect at least similar values
+for number of opens and closes and there is vast difference. Can it be that
+ATLAS jobs do a lot of opens without following closes, or perhaps multiple
+opens for the same files over and over?}}
 
@@
 \end{table*}
+
+Overall, based on our experiments with real-world jobs, it can be safely
+concluded that the file and metadata I/O load of ATLAS PanDA project on the
+OLCF Titan supercomputing environment and the Spider 2 file system is not
+detrimental to the center operations and the overall impact is minimal, at the
+current scale of the project.

### 2016-12-30 bckground.tex
Preliminary description of Figure 2

@@
 \end{figure}
+
+\subsubsection{High-level Description}
+
+\paragraph{Figure~\ref{fig:atlas_workflow} (1)} Multiple end-users (i.e.,
+physicists) submit descriptions of ATLAS Monte Carlo Workflows to the JEDI
+component of ATLAS middleware. JEDI supervises the execution of these workflows
+through a set of eight predefined stages. Each stage requires a specific type of
+input and returns a specific type of output. In this paper, we focus on the
+third stage called ``Detector Simulator'': the only one, at the moment, executed
+on the HPC resources of Titan.
+
+\paragraph{Figure~\ref{fig:atlas_workflow} (2)} JEDI submits/hands over events
+to a PANDA Server. This server manages the execution of all the ATLAS workflows
+and their stages, including the ATLAS Monte Carlo Workflow. In
+Figure~\ref{fig:atlas_workflow} we depicted only how the events of the third
+stage are organized within PANDA Server. Tasks are container of jobs and each
+job contains itself a set of events. All the events of each job belongs to a
+specific submission of a workflow description and, therefore, to a specific user
+(i.e., virtual organization). Tasks should therefore be regarded as multitenant
+while jobs and their events as monotenant\mtnote{confirmed?}.
+
+\paragraph{Figure~\ref{fig:atlas_workflow} (3)} Sets of events belonging to
+multiple tasks and jobs are pushed/pulled? from the PANDA Server to/by one or
+more PANDA Pilot. Sixteen? PANDA Pilots are currently instantiated on a
+dedicated DSL node of Titan. These instances are responsible for the packaging
+of a set of events into a PBS job script. Packaging is performed on the base of
+the information gathered about the availability of backfill resources. Depending
+on the amount of nodes available, a specific number of events is packaged into
+the PBS script.\mtnote{can the events packaged into a PBS job belong to
+different submissions and therefore users?}
+
+\paragraph{Figure~\ref{fig:atlas_workflow} (4)} AthenaMP is used to perform the
+simulation of the events within the ATLAS detector, 1 AthenaMP thread for each
+core available on the Titan nodes. Currently, 16 threads are used for each node,
+each thread being able to simulate an average of 1 event every n??
+minutes?/second?. So far, no description has been given of the flow of data
+required to simulate each event.

### 2017-01-10 bckground.tex
Questions about PANDA and Monte Carlo workflow

@@
 required to simulate each event.
+
+\begin{itemize}
+    \item Does each job of a PANDA Server (and therefore all the tasks of that specific job) belong to the same `user'/VO?
+    \item Does a PANDA Pilot pull jobs/tasks from a PANDA Server or does a PANDA server push jobs/tasks to a PANDA Pilot?
+    \item Does a PANDA Pilot pull jobs or tasks from a PANDA Server? Are the jobs within a PANDA Pilot the same jobs (i.e., the same container of the same tasks) in the PANDA Server?
+    \item Do we have 16 PANDA Pilots on a Titan DSL (is this the right acronym?) node? How many DSL nodes?
+    \item Can the events packaged into a PBS job belong to different submissions, i.e., different jobs, i.e., different users?
+    \item Do we have the distribution of the time taken on titan to process 1 event of the Monte Carlo workflow?
+\end{itemize}

### 2017-01-31 PandaEvolution.tex
Reorganizing structure of the paper and starting to fix latex

@@
-\section{Evolution of PanDA WMS}
 \subsection{Brief Introduction to Titan}
-While PanDa  currently uses  more than 100,000  cores at well over 100 Grid sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will  require more resources  than Grid computing can possibly provide.
-The Worldwide LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned analysis and data processing, but it will be insufficient  for Monte Carlo (MC) production and any extra activities.
-Additional computing and storage  resources are therefore required.
-To alleviate  these challenges, ATLAS is engaged  in an ambitious  program to expand  the current computing model to include  additional  resources such as the opportunistic   use of supercomputers   as well as commercial and academic clouds.
-In turn this activity drives evolution of the PanDA WMS.
+
+While PanDa  currently uses  more than 100,000  cores at well over 100 Grid
+sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will
+require more resources  than Grid computing can possibly provide. The Worldwide
+LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned
+analysis and data processing, but it will be insufficient  for Monte Carlo (MC)
+production and any extra activities. Additional computing and storage  resources
+are therefore required.  To alleviate  these challenges, ATLAS is engaged  in an
+ambitious  program to expand  the current computing model to include  additional
+resources such as the opportunistic   use of supercomputers   as well as
+commercial and academic clouds. In turn this activity drives evolution of the
+PanDA WMS\@.
 
 \subsection{Use of Supercomputers with PanDA}
-The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers.
-Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network.
-The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors.
-On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI).
-While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization.
-It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run.
-What matters is the total number of cycles that can be offloaded from the Grid.
-Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc.
-A reorganization  of the standard  workflow is therefore needed.
-The most suitable task for HPCs in HEP is event generation.
-Event generators are mostly computational,  stand-alone code, with few input requirements.
-There is no need  to stage-in much data.
-The event generation  in ATLAS corresponds  to
-10-15\% of all jobs on the Grid.
-A more difficult workload to adapt  for HPCs is detector  simulation.
-Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.
-However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
+
+The idea is to extend the ATLAS Computing Model beyond the Grid into the domains
+of supercomputers. Modern High Performance Computing  (HPC) involves  a very
+large number of worker nodes  connected  through a  high-speed   network. The
+worker nodes can have  multicore CPUs that can be augmented with massively
+parallel Graphics Processing Units (GPUs) or other types of coprocessors. On an
+HPC machine, a  typical job is highly parallel and connected,  where each core
+calculates  a small part of the problem and coordinates the activity via message
+passing interface (MPI). While this is different from the HEP computing
+paradigm, where jobs are independent, it still shares common  features  such as
+the use of parallelization. It is not a requirement  that HPC machines are able
+to run any possible task, nor is it relevant how many kinds of job types that
+can be run. What matters is the total number of cycles that can be offloaded
+from the Grid. Standard  ATLAS  workflow  can not be easily ported to
+supercomputers   due  to   several   complications such  as specialized worker
+node  setups, no  outbound network connections,  limited memory per node, custom
+operating systems,  etc. A reorganization  of the standard  workflow is
+therefore needed.
+
+The most suitable task for HPCs in HEP is event generation. Event generators are
+mostly computational,  stand-alone code, with few input requirements.  There is
+no need  to stage-in much data. The event generation  in ATLAS corresponds  to
+10--15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs
+is detector  simulation. Simulation tasks tend to be tied to the framework of a
+particular simulation and often require  database access for geometry and
+relevant run conditions.  However,  there is a  lot to be gained  since
+simulation  tasks correspond to 2/3 of the Grid capacity.
+
 \subsection{Interfacing PanDA with Titan}
-The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US.
-Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
-Integration with Titan is the current focus for PanDA developers.
-The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}.
-This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world.
-The pilot can communicate with the PanDA server from the front-end machine.
-The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job.
-ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
-1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design.
-Time allocation on an HPC is competitive and large projects are often preferred.
-About 10\% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources.
-The worker nodes sit idle because there are not enough of them or they do not have enough time to handle a large scale computing job.
-On a machine of the scale of Titan, these 10\% correspond to estimated 300M core hours per year.
-A system that can occupy those temporarily free nodes with smaller scale tasks would be very valuable.
-This offers a great possibility for PanDA to harvest the opportunistic resources on Titan and use a similar mechanism on other HPCs.
-Functionality has been added to the PanDA Pilot to collect information about available unused worker nodes on Titan in real time.
-This allows PanDA to define precisely size and duration of jobs submitted to Titan according to available free resources.
-First test of the system were quite successful and show great promise in increasing the resource utilization on Titan.
+
+The Titan supercomputer, current number two (number one from November 2012 until
+June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing
+Facility (OLCF) in Oak Ridge National Laboratory, US\@. Titan, was the first
+large-scale system to use  a hybrid architecture that utilizes worker nodes with
+both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
+
+Integration with Titan is the current focus for PanDA developers. The project
+aims to integrate Titan with the PanDA system using an updated PanDA Pilot that
+runs on the front-end node and submits ATLAS payloads to the worker nodes using
+the local batch system (PBS) via the SAGA (Simple API for Grid Applications)
+interface~\cite{SAGA}. This solves several complications of  running on HPC
+worker nodes,  including the lack of connectivity to outside world. The pilot
+can communicate with the PanDA server from the front-end machine. The
+interactive front-end  machines and the worker nodes  use  a shared  file system
+which makes  it  easy  for the pilot to stage-in  any input files that are
+required by the payload and stage-out the produced output files after the
+payload has finished at the end of the job. ATLAS Tier 1 computing  center at
+Brookhaven National Lab is currently  used for data stage out from Titan but in
+principle  that can be any Grid site.
+
+\paragraph*{HPC Backfill} HPC facilities are geared towards  large scale jobs by
+design. Time allocation on an HPC is competitive and large projects are often
+preferred. About 10\% of capacity on a  typical HPC machine  is unused  due to
+mismatches between job sizes and available  resources. The worker nodes sit idle
+because there are not enough of them or they do not have enough time to handle a
+large scale computing job. On a machine of the scale of Titan, these 10\%
+correspond to estimated 300M core hours per year. A system that can occupy those
+temporarily free nodes with smaller scale tasks would be very valuable. This
+offers a great possibility for PanDA to harvest the opportunistic resources on
+Titan and use a similar mechanism on other HPCs. Functionality has been added to
+the PanDA Pilot to collect information about available unused worker nodes on
+Titan in real time. This allows PanDA to define precisely size and duration of
+jobs submitted to Titan according to available free resources. First test of the
+system were quite successful and show great promise in increasing the resource
+utilization on Titan.
+
 \begin{figure}

### 2017-01-31 bckground.tex
Reorganizing structure of the paper and starting to fix latex

--- 2016-11-29 bckground.tex
@@
-\subsection{Workflows and Workloads}
+\subsection{Terminology}
+
+\mtnote{To be moved to the end of the introduction.}
 
@@
 indicates a abstract workflow instance.
-
-\subsection{The ATLAS Monte Carlo Workflow}
-
-The distinction between an abstract and concrete workflow instances highlights
-an important element of the execution process of a workflow. Abstract workflows
-can be interpreted as a formal description of the requirements for the workflow
-execution, while concrete workflows can be interpreted as a formal description
-of the workflow's execution process. The process of resource selection matches
-the resource capabilities to the workflow requirements making possible the
-transition from execution requirements to execution process.
-
-Workload managers are those components of workflow systems or, more in general,
-of distributed applications that enable resource selection, acquisition, and
-use. The set of capabilities required by a workload manager mostly depends on
-the heterogeneity and dynamicity of both the workflow and the resources.
-Heterogeneity is a measure of the diversity of the workflow requirements and
-resource capabilities that need to be matched to enable the execution of the
-workflow's tasks; dynamicity is a measure of how much these requirements and
-capabilities varies during execution.
-
-The PanDA workload management system was initially designed to manage the
-execution of static, relatively homogeneous tasks on dynamic but homogeneous
-resources. Progressively, PanDA has evolved to enable the management of more
-heterogeneous tasks and resources. In this paper we describe how the design and
-architecture of PanDA has enabled executions of multi-threaded tasks on Titan,
-currently the largest high performance computing (HPC) resource available in the
-USA for scientific research.
-
-Traditionally, the ATLAS workflow has been based on single-core tasks, executed
-on Grid resources to process large amount (2TB) of input data divided into
-discrete I/O units called ``events''. This approach has many merits: (i)
-simplified process of resource selection under the assumption of homogeneous
-resources and execution environments; (ii) opportunistic distribution of compute
-tasks across resources based on contingent availability; (iii) relative
-robustness due to the ability of re-executing failed tasks; and (iv)
-multi-tiered distribution and storage of events' data.
-
-Despite all the advantages, ATLAS workflows have evolved to include
-compute-intense tasks that can benefit from multithreaded execution on HPC
-machines. The Monte Carlo workflow~\cite{} is one of the largest workflows
-executed by the ATLAS experiment and at least one of its stages can benefit from
-large amount of parallelism. The Geant4 toolkit is used to simulate the passage
-of particles through matter and\ldots
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figures/atlas_workflow.pdf}
-  \caption{ATLAS Monte Carlo workflow.}
-\label{fig:atlas_workflow}
-\end{figure}

### 2017-01-31 overall.tex
Reorganizing structure of the paper and starting to fix latex

@@
-\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
-\label{sec:panda_titan}
-
 \subsection{PanDA Shared Library I/O Performance Impact at OLCF}
@@
 shared by all OLCF resources and users, this resulted in lower overall
-interactive and metadata performance at OLCF. As can be seen in
+interactive and metadata performance at OLCF\@. As can be seen in
 Listing~\ref{mdstrace} the metadata I/O activity for ATLAS exhibits a spike
@@
 directories, the metadata contention problem has resolved and a two orders of
-magnitude metadata performance improvement (i.e., ~6,300 seconds on Lustre vs.
-~1,500 seconds on NFS) has been observed for ATLAS.  Also, for both cases the
+magnitude metadata performance improvement (i.e., \~6,300 seconds on Lustre vs.\
+\~1,500 seconds on NFS) has been observed for ATLAS\@.  Also, for both cases the
 input file is placed on memory disk on a Titan compute nodes and for Lustre the
@@
 To better understand the I/O impact of ATLAS PanDA project on Titan
-supercomputing environment we analyzed 1,175 jobs ran on the week of
-10/25/2016, for a total of 174 hours. Table~\ref{panda-olcf-stats} shows the
-overall statistical breakdown of the observed file I/O impact of ATLAS at OLCF.
-Figures \ref{fig:atlas-titan-io-read} and \ref{fig:atlas-titan-io-written} show
+supercomputing environment we analyzed 1,175 jobs ran on the week of 10/25/2016,
+for a total of 174 hours. Table~\ref{panda-olcf-stats} shows the overall
+statistical breakdown of the observed file I/O impact of ATLAS at OLCF\@.
+Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written} show
 the file read and write I/O histograms for these 1,175 jobs, respectively.
-Figures \ref{fig:atlas-titan-file-open} and \ref{fig:atlas-titan-file-close}
-show the file open() and close() metadata load histograms of the same 1,175
+Figures~\ref{fig:atlas-titan-file-open} and~\ref{fig:atlas-titan-file-close}
+show the file $open()$ and $close()$ metadata load histograms of the same 1,175
 ATLAS jobs, respectively.
@@
 aggregate in this observed period was less than 250 GB and the maximum amount
-written in aggregate was less than 75 GB. The average amount read per job is 20
-GB and average amount written is 6 GB.
+written in aggregate was less than 75 GB\@. The average amount read per job is
+20 GB and average amount written is 6 GB\@.
 
@@
     \end{tabular}
-    \caption{ATLAS file open() histogram on Titan for week of 10/25/16.}
+    \caption{ATLAS file $open()$ histogram on Titan for week of 10/25/16.}
 \label{fig:atlas-titan-file-open}
@@
     \end{tabular}
-    \caption{ATLAS file close() histogram on Titan for week of 10/25/16.}
+    \caption{ATLAS file $close()$ histogram on Titan for week of 10/25/16.}
 \label{fig:atlas-titan-file-close}
@@
 
-On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file open() operations
-and 5 file close() operations per second. As can be seen from figures
-\ref{fig:atlas-titan-file-open} and \ref{fig:atlas-titan-file-close}, they
-exhibit similar distributions. The maximum number of file open() operations are
-around 170 on average and the maximum number of file close() operations are 39
-on average per second per job. The total number of file open() operations is
-172,089,760 for the observed window of 174 hours, while the total number of
-file close() operations stand at 40,132,992 for the total 1,175 ATLAS PanDA
-jobs in the same observation window. The difference between these two values is
-puzzling and it is under investigation at the time of writing this paper.
-{\sarp{can someone take a look at this? I would expect at least similar values
-for number of opens and closes and there is vast difference. Can it be that
-ATLAS jobs do a lot of opens without following closes, or perhaps multiple
-opens for the same files over and over?}}
+On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file $open()$
+operations and 5 file $close()$ operations per second. As can be seen from
+figures~\ref{fig:atlas-titan-file-open} and~\ref{fig:atlas-titan-file-close},
+they exhibit similar distributions. The maximum number of file $open()$
+operations are around 170 on average and the maximum number of file $close()$
+operations are 39 on average per second per job. The total number of file
+$open()$ operations is 172,089,760 for the observed window of 174 hours, while
+the total number of file $close()$ operations stand at 40,132,992 for the total
+1,175 ATLAS PanDA jobs in the same observation window. The difference between
+these two values is puzzling and it is under investigation at the time of
+writing this paper. {\sarp{can someone take a look at this? I would expect at
+least similar values for number of opens and closes and there is vast
+difference. Can it be that ATLAS jobs do a lot of opens without following
+closes, or perhaps multiple opens for the same files over and over?}}
 
@@
 \begin{tabular}{lllllllll}
- & Num. Nodes & Duration (s) & Read (GB) & Written (GB) & GB Read/nodes & GB Written/nodes & open() & close() \\
+ & Num. Nodes & Duration (s) & Read (GB) & Written (GB) & GB Read/nodes & GB Written/nodes & $open()$ & $close()$ \\
 Min & 1 & 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\

### 2017-01-31 paper.tex
initial drop on atlas mds performance improvement
minor re-structuring.
finishing the write up for the I/O impact section
Preliminary description of Figure 2
Reorganizing structure of the paper and starting to fix latex

@@
-%% bare_conf.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
 \documentclass[10pt, conference, compsocconf]{IEEEtran}
-% Add the compsocconf option for Computer Society conferences.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
 
@@
 \input{include}
+
+\usepackage{listings}
+\usepackage{comment}
 
@@
 
-\newcommand{\sarp}[1]{{\color{green}\textit{sarp: {#1}}}}
+% -----------------------------------------------------------------------------
+% BEGIN DOCUMENT
+% -----------------------------------------------------------------------------
+\begin{document}
 
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Everthing you wanted to know about PanDA but were truly truly afraid to ask}
+\title{Everthing you wanted to know about PanDA but were truly truly afraid to
+ask}
 
@@
 % affiliations
-
 \author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
@@
 
+% ----------------------------------------------------------------------------
+% ABSTRACT
+% ----------------------------------------------------------------------------
 \begin{abstract}
- Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.
-Heterogeneous  resources are distributed worldwide, thousands of physicists analyzing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year.
-The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model.
-The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
-2013.
-We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
+Experiments at the Large Hadron Collider (LHC) face unprecedented computing
+challenges. Heterogeneous resources are distributed worldwide, thousands of
+physicists analyzing the data need remote access to hundreds of computing
+sites, the volume of processed data is beyond the exabyte scale, and data
+processing requires more than billions of hours of computing usage per year. The
+PanDA (Production and Distributed Analysis) system was developed to meet the
+scale and complexity of LHC distributed computing for the ATLAS experiment. In
+the process, the old batch job paradigm of computing in HEP was discarded in
+favor of a far more flexible and scalable model. The success of PanDA at the LHC
+is leading to widespread adoption and testing by other experiments. PanDA is
+the first exascale workload management system in HEP, already operating at
+a million computing jobs per day, and processing over an exabyte of data in
+2013. We will describe the design and implementation of PanDA, present data on
+the performance of PanDA at the LHC, and discuss plans for future evolution
+of the system to meet new challenges of scale, heterogeneity and increasing
+user base.
 \end{abstract}
@@
 % ----------------------------------------------------------------------------
-% I
+% I - INTRODUCTION
 % ----------------------------------------------------------------------------
+\input{introduction}
 
-\input{introduction}
---------------------------------------------------------------------
 \begin{enumerate}
-  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
+  \item Problem that has been solved: executing many applications on
+  heterogeneous distributed resources.
   \item Why are WMS needed?
-  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
+  \item Increasing importance of WMS w.r.t complex and scalable applications on
+  heterogeneous distributed resources.
   \item \ldots
@@
 % ----------------------------------------------------------------------------
-% II
+% II - BACKGROUND AND RELATED WORK
 % ----------------------------------------------------------------------------
+\section{Background and Related Work}
+\label{sec:bkgrd_literature}
 
-\input{Background}
+\begin{enumerate}
+  \item Other workload management systems (WMS) (not exclusive)
+  \begin{enumerate}
+    \item glidein-WMS
+    \item Dirac
+    \item ALICE-AlieN
+  \end{enumerate}
+  \begin{enumerate}
+    \item Comparison and contrast
+    \item Core features of a general WMS, i.e., minimal complete model of a WMS.
+  \end{enumerate}
+\end{enumerate}
+
+\input{WMSLiterature}
+
 \begin{enumerate}
@@
 
---------------------------------------------------------------------
+% Please do not correct bckground to background.
+\input{bckground}
 
 % ----------------------------------------------------------------------------
-% III
+% III - PanDA OVERVIEW
 % ----------------------------------------------------------------------------
+\section{PanDA Overview}
+\label{sec:panda_overview}
 
-\input{WMSLiterature}
---------------------------------------------------------------------
-\begin{enumerate}
-  \item Other workload management systems (WMS) (not exclusive)
-  \begin{enumerate}
-    \item glidein-WMS
-    \item Dirac
-    \item ALICE-AlieN
-  \end{enumerate}
-{enumerate}
-  \item Comparison and contrast
-  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% IV
-% ----------------------------------------------------------------------------
-\section{PanDA Overview}\label{sec:panda_overview}
---------------------------------------------------------------------
 \begin{enumerate}
@@
 \end{enumerate}
---------------------------------------------------------------------
-
-\input{titan}
-
-\input{PandaEvolution}
 
 % ----------------------------------------------------------------------------
-% V
+% IV - ATLAS MONTE CARLO WORKFLOW
 % ----------------------------------------------------------------------------
+\section{ATLAS Monte Carlo Workflow}
+\label{sec:mc_workflow}
 
+\input{workflow}
+
+% ----------------------------------------------------------------------------
+% V - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
+% ----------------------------------------------------------------------------
 \section{Deploying PanDA on a Leadership-scale system}
@@
 \begin{enumerate}
+  \item Why? To support a specific stage of the ATLAS Monte Carlo Workflow.
   \item Challenges
@@
 
-% ----------------------------------------------------------------------------
-% VI
-% ----------------------------------------------------------------------------
-\input{perf}
-
-%\section{Performance Characterization}
-%\label{sec:panda_performance}
-%
-%\begin{enumerate}
-%  \item Internal PanDA performance
-%  \item External PanDA performance (workload specific on underlying system/platform)
-%\end{enumerate}
+\input{titan}
 
 % ----------------------------------------------------------------------------
-% VII
+% VI - EVOLUTION OF PANDA WMS
 % ----------------------------------------------------------------------------
+% \section{Evolution of PanDA WMS}
+% \label{sec:panda_evolution}
+
+\input{PandaEvolution}
+
+% ----------------------------------------------------------------------------
+% VI - PERFORMANCE CHARACTERIZATION
+% ----------------------------------------------------------------------------
+% \section{Performance Characterization}
+% \label{sec:panda_performance}
+%
+% \input{perf}
+
+% ----------------------------------------------------------------------------
+% VI - PUTTING IT ALL TOGETHER: EXPERIENCE OF PANDA ON TITAN
+% ----------------------------------------------------------------------------
+\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
+\label{sec:panda_titan}
+
 \input{overall}
@@
 % ----------------------------------------------------------------------------
-% VIII
+% VII - THE NEXT GENERATION EXECUTOR
 % ----------------------------------------------------------------------------
-
-\section{PANDA: Future/RoadMap}
-\input{Experiments}
+% \section{PANDA: Future/RoadMap}
+\section{PANDA: The Next Generation Executor}
 \label{sec:panda_roadmap}
 
- \section{Conclusion}\label{sec:conclusion}
-The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.
-In the process,  the old batch job paradigm  of computing in HEP was  discarded  in favor of a  far more flexible and scalable  model.
-The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments.
-PanDA  is the first exascale  workload management system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013.
-Next LHC run will pose massive computing  challenges.
-With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more.
-Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS.
-To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system.
-Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry.
-The technology base provided by the PanDA system will enhance the usage of a variety  of high-performance computing resources available to basic research.
+\begin{enumerate}
+    \item Rationale and design (why)
+    \item Architecture (how)
+    \item Integration
+    \item Characterization (experiments)
+\end{enumerate}
 
+\input{Experiments}
+
+% ----------------------------------------------------------------------------
+% VIII - CONCLUSIONS
+% ----------------------------------------------------------------------------
+\section{Conclusion}
+\label{sec:conclusion}
+
+The PanDA  system was developed to meet the scale and complexity of LHC
+distributed computing for the ATLAS experiment.  In the process,  the old batch
+job paradigm  of computing in HEP was  discarded  in favor of a  far more
+flexible and scalable  model. The success  of PanDA  at the LHC is leading to
+widespread adoption and testing by other experiments. PanDA  is the first
+exascale  workload management system in HEP, already operating at a million
+computing jobs per day, and processing over an exabyte of data in 2013. Next LHC
+run will pose massive computing  challenges. With a  doubling of the beam
+energy  and luminosity as  well as an increased  need for  simulates  data, the
+data volume is expected to increase with a factor 5-6 or more. Storing and
+processing  this amount of data is a  challenge   that cannot be resolved with
+the currently existing  computing  resources in ATLAS. To resolve this
+challenge, ATLAS is turning to commercial  as well as academic Cloud services
+and HPCs via the PanDA system. Also the work underway is enabling the use of
+PanDA by new scientific collaborations and communities as a means  of leveraging
+extreme scale computing  resources with a low barrier of entry. The technology
+base provided by the PanDA system will enhance the usage of a variety  of
+high-performance computing resources available to basic research.
+
+% ----------------------------------------------------------------------------
+% ACKNOWLEDGEMENTS
+% ----------------------------------------------------------------------------
 \input{ack}
@@
 
+% ----------------------------------------------------------------------------
+% REFERENCES
+% ----------------------------------------------------------------------------
 \bibliographystyle{plain}

### 2017-01-31 workflow.tex
Reorganizing structure of the paper and starting to fix latex

The distinction between an abstract and concrete workflow instances highlights
an important element of the execution process of a workflow. Abstract workflows
can be interpreted as a formal description of the requirements for the workflow
execution, while concrete workflows can be interpreted as a formal description
of the workflow's execution process. The process of resource selection matches
the resource capabilities to the workflow requirements making possible the
transition from execution requirements to execution process.

Workload managers are those components of workflow systems or, more in general,
of distributed applications that enable resource selection, acquisition, and
use. The set of capabilities required by a workload manager mostly depends on
the heterogeneity and dynamicity of both the workflow and the resources.
Heterogeneity is a measure of the diversity of the workflow requirements and
resource capabilities that need to be matched to enable the execution of the
workflow's tasks; dynamicity is a measure of how much these requirements and
capabilities varies during execution.

The PanDA workload management system was initially designed to manage the
execution of static, relatively homogeneous tasks on dynamic but homogeneous
resources. Progressively, PanDA has evolved to enable the management of more
heterogeneous tasks and resources. In this paper we describe how the design and
architecture of PanDA has enabled executions of multi-threaded tasks on Titan,
currently the largest high performance computing (HPC) resource available in the
USA for scientific research.

Traditionally, the ATLAS workflow has been based on single-core tasks, executed
on Grid resources to process large amount (2TB) of input data divided into
discrete I/O units called ``events''. This approach has many merits: (i)
simplified process of resource selection under the assumption of homogeneous
resources and execution environments; (ii) opportunistic distribution of compute
tasks across resources based on contingent availability; (iii) relative
robustness due to the ability of re-executing failed tasks; and (iv)
multi-tiered distribution and storage of events' data.

Despite all the advantages, ATLAS workflows have evolved to include
compute-intense tasks that can benefit from multithreaded execution on HPC
machines. The Monte Carlo workflow~\cite{} is one of the largest workflows
executed by the ATLAS experiment and at least one of its stages can benefit from
large amount of parallelism. The Geant4 toolkit is used to simulate the passage
of particles through matter and\ldots

\begin{figure}
  \includegraphics[width=\columnwidth]{figures/atlas_workflow.pdf}
  \caption{ATLAS Monte Carlo workflow.}
\label{fig:atlas_workflow}
\end{figure}

\subsubsection{High-level Description}

\paragraph{Figure~\ref{fig:atlas_workflow} (1)} Multiple end-users (i.e.,
physicists) submit descriptions of ATLAS Monte Carlo Workflows to the JEDI
component of ATLAS middleware. JEDI supervises the execution of these workflows
through a set of eight predefined stages. Each stage requires a specific type of
input and returns a specific type of output. In this paper, we focus on the
third stage called ``Detector Simulator'': the only one, at the moment, executed
on the HPC resources of Titan.

\paragraph{Figure~\ref{fig:atlas_workflow} (2)} JEDI submits/hands over events
to a PANDA Server. This server manages the execution of all the ATLAS workflows
and their stages, including the ATLAS Monte Carlo Workflow. In
Figure~\ref{fig:atlas_workflow} we depicted only how the events of the third
stage are organized within PANDA Server. Tasks are container of jobs and each
job contains itself a set of events. All the events of each job belongs to a
specific submission of a workflow description and, therefore, to a specific user
(i.e., virtual organization). Tasks should therefore be regarded as multitenant
while jobs and their events as monotenant\mtnote{confirmed?}.

\paragraph{Figure~\ref{fig:atlas_workflow} (3)} Sets of events belonging to
multiple tasks and jobs are pushed/pulled? from the PANDA Server to/by one or
more PANDA Pilot. Sixteen? PANDA Pilots are currently instantiated on a
dedicated DSL node of Titan. These instances are responsible for the packaging
of a set of events into a PBS job script. Packaging is performed on the base of
the information gathered about the availability of backfill resources. Depending
on the amount of nodes available, a specific number of events is packaged into
the PBS script.\mtnote{can the events packaged into a PBS job belong to
different submissions and therefore users?}

\paragraph{Figure~\ref{fig:atlas_workflow} (4)} AthenaMP is used to perform the
simulation of the events within the ATLAS detector, 1 AthenaMP thread for each
core available on the Titan nodes. Currently, 16 threads are used for each node,
each thread being able to simulate an average of 1 event every n??
minutes?/second?. So far, no description has been given of the flow of data
required to simulate each event.

Questions for Danila/Sergey:
\begin{itemize}
    \item Does each job of a PANDA Server (and therefore all the tasks of that specific job) belong to the same `user'/VO\@?
    \item Does a PANDA Pilot pull jobs/tasks from a PANDA Server or does a PANDA server push jobs/tasks to a PANDA Pilot?
    \item Does a PANDA Pilot pull jobs or tasks from a PANDA Server? Are the jobs within a PANDA Pilot the same jobs (i.e., the same container of the same tasks) in the PANDA Server?
    \item Do we have 16 PANDA Pilots on a Titan DSL (is this the right acronym?) node? How many DSL nodes?
    \item Can the events packaged into a PBS job belong to different submissions, i.e., different jobs, i.e., different users?
    \item Do we have the distribution of the time taken on titan to process 1 event of the Monte Carlo workflow?
\end{itemize}

### 2017-01-31 Experiments.tex
Iteration organization

@@
 %\jhanote{Not sure this should be under ``PANDA:Future/ROAD MAP?'' (cut and paste error it was not meant to be; FIXED)}
-\section{Experiments}
-The execution of jobs on TITAN follows a three step scheme: i) as first, TITAN's Backfill queue communicates  the availability of a slot i.e.
-a set of idle cores for a given time period\aanote{How many offers does PanDA receive per time unit?
-Average nodes per offer?}; ii) then, PanDA can decide to run on the slot or wait until a larger slot becomes available; iii) if the slot is taken, PanDA wraps a certain amount of events in a job and sends the job for execution.\aanote{I am supposing that the concept of event has been already introduced at this point}.
 
-At least in principle, PanDA can reject the slot because it does not fit the requirements in terms of number of cores or wall-time unit; however, this situation is rare because events are single core Athena-MP threads\mtnote{Should we consider events inputs of Athena-MP jobs, with each event processed in a single thread by a---to be specified---module of Athena-MP?} that require on average twenty minutes for being executed.
-Thus, the decomposition in event is fine grained once compared to the average allocation \aanote{Which correspond to\ldots}.
+\subsection{Experiments}
 
-As a consequence, PanDA accepts the major part of the proposed slots by selecting a number of events that maximize the utilization of the resources during the assigned time interval.
+The execution of jobs on TITAN follows a three step scheme: (i) as first,
+TITAN's Backfill queue communicates  the availability of a slot i.e., a set of
+idle cores for a given time period\aanote{How many offers does PanDA receive per
+time unit? Average nodes per offer?}; (ii) then, PanDA can decide to run on the
+slot or wait until a larger slot becomes available; (iii) if the slot is taken,
+PanDA wraps a certain amount of events in a job and sends the job for
+execution.\aanote{I am supposing that the concept of event has been already
+introduced at this point}.
 
-On one hand, this %\jhanote{I would use the word execution strategy carefully. Maybe a more general term such as execution model, a special type of which is execution strategy?(I Agree, FIXED)} 
-execution model leads to a drastic reduction of the queue time and, consequently, it guarantees high-performances.
-On the other hand, this model is tailored for the execution of ATLAS workload on TITAN by using the Backfill queue\mtnote{Do we have a queue dedicated to backfilling on Titan or should we speak of backfilling done on one or more queues?}.
+At least in principle, PanDA can reject the slot because it does not fit the
+requirements in terms of number of cores or wall-time unit; however, this
+situation is rare because events are single core Athena-MP threads\mtnote{Should
+we consider events inputs of Athena-MP jobs, with each event processed in a
+single thread by a---to be specified---module of Athena-MP?} that require on
+average twenty minutes for being executed. Thus, the decomposition in event is
+fine grained once compared to the average allocation \aanote{Which correspond
+to\ldots}.
+
+As a consequence, PanDA accepts the major part of the proposed slots by
+selecting a number of events that maximize the utilization of the resources
+during the assigned time interval.
+
+On one hand, this
+% \jhanote{I would use the word execution strategy carefully. Maybe a more
+% general term such as execution model, a special type of which is execution
+% strategy?(I Agree, FIXED)}
+execution model leads to a drastic reduction of the queue time and,
+consequently, it guarantees high-performances. On the other hand, this model is
+tailored for the execution of ATLAS workload on TITAN by using the Backfill
+queue\mtnote{Do we have a queue dedicated to backfilling on Titan or should we
+speak of backfilling done on one or more queues?}.
 
@@
 
-For this reason, in the remaining of this section we propose experimental results to address them. The idea behind the experiments is to show the behaviour of ATLAS workload on TITAN's batch queue.
+For this reason, in the remaining of this section we propose experimental
+results to address them. The idea behind the experiments is to show the behavior
+of ATLAS workload on TITAN's batch queue.
 
-First, we investigate the acquisition of slot as a function of time (wall-time) and space (number of cores) in order to understand how TITAN's scheduler and workload behave under different slot requests.
-This analysis includes two bulks of experiments: the first, generated ``in silico'', is based on simulations of the TITAN's scheduler by means of the Moab simulator \cite{}, whereas the second, generated ``in vivo'',  considers real submissions of ATLAS workload on TITAN's batch queue.
+First, we investigate the acquisition of slot as a function of time (wall-time)
+and space (number of cores) in order to understand how TITAN's scheduler and
+workload behave under different slot requests. This analysis includes two bulks
+of experiments: the first, generated ``in silico'', is based on simulations of
+the TITAN's scheduler by means of the Moab simulator~\cite{}, whereas the
+second, generated ``in vivo'',  considers real submissions of ATLAS workload on
+TITAN's batch queue.
 
-By using the batch queue, we move in a situation where the queue time becomes dominant but, at the same time, we have more freedom to decide the parameters of the slot.
-For this reason, the second set of the experiments aims to find sub-optimal parameters with which we can minimize the trade-off between the size of a slot and the time spent in queue waiting for that slot to become available.
-In other words, we aim to minimize the completion time by finding the best trade-off between execution time and queue time.
+By using the batch queue, we move in a situation where the queue time becomes
+dominant but, at the same time, we have more freedom to decide the parameters of
+the slot. For this reason, the second set of the experiments aims to find
+sub-optimal parameters with which we can minimize the trade-off between the size
+of a slot and the time spent in queue waiting for that slot to become available.
+In other words, we aim to minimize the completion time by finding the best
+trade-off between execution time and queue time.
 
-This execution model introduces slot utilization as one of the key factors for high-performances.
-This happens because, in order to minimize the time spent in queue, we might asks for slots in advance and, then we could not br able to saturate them when they become available.
-Thus, this strategy requires a new functionality that allows the job to receive and execute new events while it is already running on the resources.
-In order to do that we perform the experiments by using a new generation executor that implements such functionality.
+This execution model introduces slot utilization as one of the key factors for
+high-performances. This happens because, in order to minimize the time spent in
+queue, we might asks for slots in advance and, then we could not be able to
+saturate them when they become available. Thus, this strategy requires a new
+functionality that allows the job to receive and execute new events while it is
+already running on the resources. In order to do that we perform the experiments
+by using a new generation executor that implements such functionality.
 
-As last observation, it is important to point out that the percentage of utilization of a slot is minor problem with the current implementation because, due to the dynamics of the Backfill queue, PanDA has a high probability to re-acquire a slot immediately after it has released one\aanote{Are we able to quantify this ``immediately''?}.
+As last observation, it is important to point out that the percentage of
+utilization of a slot is minor problem with the current implementation because,
+due to the dynamics of the Backfill queue, PanDA has a high probability to
+re-acquire a slot immediately after it has released one\aanote{Are we able to
+quantify this ``immediately''?}.

### 2017-02-02 paper.tex
Iteration organization
Include missing PandaOverview.tex

@@
 \author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
+\IEEEauthorblockA{line 1 (of Affiliation): dept.\ name of organization\\
 line 2: name of organization, acronyms acceptable\\
@@
 \IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
+\IEEEauthorblockA{line 1 (of Affiliation): dept.\ name of organization\\
 line 2: name of organization, acronyms acceptable\\
@@
 \begin{IEEEkeywords}
-TBD.
+TBD\@.
 \end{IEEEkeywords}
@@
     \item Comparison and contrast
-    \item Core features of a general WMS, i.e., minimal complete model of a WMS.
+    \item Core features of a general WMS, i.e., minimal complete model of a WMS\@.
   \end{enumerate}
@@
 \begin{enumerate}
-  \item Brief overview of distributed CI.
+  \item Brief overview of distributed CI\@.
   \item Workload vs Workflow. (DONE)
-  \item Distinguish between WFMS and WLMS.
+  \item Distinguish between WFMS and WLMS\@.
   \begin{enumerate}
-    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
+    \item Role of Workload MS\@: help ``hide'' and federate heterogeneous resources.
   \end{enumerate}
@@
 \end{enumerate}
+
+\input{PandaOverview}
 
@@
 % ----------------------------------------------------------------------------
-% VI - PUTTING IT ALL TOGETHER: EXPERIENCE OF PANDA ON TITAN
+% VI - PERFORMANCE CHARACTERIZATION
 % ----------------------------------------------------------------------------
-\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
+% \section{Putting it all together: Experience of PanDA on TITAN} % (fold)
+\section{Performance Characterization}
 \label{sec:panda_titan}
@@
 % \section{PANDA: Future/RoadMap}
-\section{PANDA: The Next Generation Executor}
+\section{PANDA\@: The Next Generation Executor}
 \label{sec:panda_roadmap}
@@
 \begin{enumerate}
+    \item Discussion of the limitations of the current state of the art
     \item Rationale and design (why)
@@
 energy  and luminosity as  well as an increased  need for  simulates  data, the
-data volume is expected to increase with a factor 5-6 or more. Storing and
+data volume is expected to increase with a factor 5--6 or more. Storing and
 processing  this amount of data is a  challenge   that cannot be resolved with
-the currently existing  computing  resources in ATLAS. To resolve this
+the currently existing  computing  resources in ATLAS\@. To resolve this
 challenge, ATLAS is turning to commercial  as well as academic Cloud services

### 2017-02-06 workflow.tex
Notes from meeting 01312017
Itartion workflow description

@@
 of the workflow's execution process. The process of resource selection matches
-the resource capabilities to the workflow requirements making possible the
-transition from execution requirements to execution process.
+the capabilities of resources to the requirements of a workflow, making possible
+the transition from execution requirements to execution process.
 
-Workload managers are those components of workflow systems or, more in general,
-of distributed applications that enable resource selection, acquisition, and
-use. The set of capabilities required by a workload manager mostly depends on
-the heterogeneity and dynamicity of both the workflow and the resources.
-Heterogeneity is a measure of the diversity of the workflow requirements and
-resource capabilities that need to be matched to enable the execution of the
-workflow's tasks; dynamicity is a measure of how much these requirements and
-capabilities varies during execution.
+As seen in \S\ref{sec:bkgrd}, Workload managers are those components of workflow
+systems or, more in general, of distributed applications that enable resource
+selection, acquisition, and use. The set of capabilities required by a workload
+manager mostly depends on the heterogeneity and dynamicity of both the workflow
+and the resources. Heterogeneity is a measure of the diversity of the workflow
+requirements and resource capabilities; dynamicity is a measure of how much
+these requirements and capabilities varies during execution.
 
 The PanDA workload management system was initially designed to manage the
-execution of static, relatively homogeneous tasks on dynamic but homogeneous
-resources. Progressively, PanDA has evolved to enable the management of more
-heterogeneous tasks and resources. In this paper we describe how the design and
-architecture of PanDA has enabled executions of multi-threaded tasks on Titan,
-currently the largest high performance computing (HPC) resource available in the
-USA for scientific research.
+execution of static, relatively homogeneous tasks on dynamic and heterogeneous
+resources~\cite{panda_grid}. Progressively, PanDA has evolved to enable the
+management of more heterogeneous tasks and resources. In this paper, we describe
+how the design and architecture of PanDA has enabled executions of
+multi-threaded tasks on Titan, currently the largest high performance computing
+(HPC) resource available in the USA to scientific research.
 
@@
 simplified process of resource selection under the assumption of homogeneous
-resources and execution environments; (ii) opportunistic distribution of compute
-tasks across resources based on contingent availability; (iii) relative
-robustness due to the ability of re-executing failed tasks; and (iv)
-multi-tiered distribution and storage of events' data.
+execution environment; (ii) opportunistic distribution of compute tasks across
+resources based on contingent availability; (iii) relative robustness due to the
+ability of re-executing failed tasks; and (iv) multi-tiered distribution and
+storage of events' data.
 
 Despite all the advantages, ATLAS workflows have evolved to include
-compute-intense tasks that can benefit from multithreaded execution on HPC
+compute-intense tasks that can benefit also from multithreaded execution on HPC
 machines. The Monte Carlo workflow~\cite{} is one of the largest workflows
-executed by the ATLAS experiment and at least one of its stages can benefit from
-large amount of parallelism. The Geant4 toolkit is used to simulate the passage
-of particles through matter and\ldots
+executed by the ATLAS experiment, and at least one of its stages can benefit
+from large amount of parallelism. The Geant4 toolkit is used to simulate the
+passage of particles through matter and\ldots
 
@@
 stage are organized within PANDA Server. Tasks are container of jobs and each
-job contains itself a set of events. All the events of each job belongs to a
-specific submission of a workflow description and, therefore, to a specific user
-(i.e., virtual organization). Tasks should therefore be regarded as multitenant
-while jobs and their events as monotenant\mtnote{confirmed?}.
+job contains itself a set of events. Each task, each job in that task and all
+the events of those jobs belongs to a specific user. This enables provenance
+across the PanDA stack and isolating the management of the data of each task by
+dedicated subsystems.
 
 \paragraph{Figure~\ref{fig:atlas_workflow} (3)} Sets of events belonging to
-multiple tasks and jobs are pushed/pulled? from the PANDA Server to/by one or
-more PANDA Pilot. Sixteen? PANDA Pilots are currently instantiated on a
-dedicated DSL node of Titan. These instances are responsible for the packaging
-of a set of events into a PBS job script. Packaging is performed on the base of
-the information gathered about the availability of backfill resources. Depending
-on the amount of nodes available, a specific number of events is packaged into
-the PBS script.\mtnote{can the events packaged into a PBS job belong to
-different submissions and therefore users?}
+multiple tasks and jobs are pulled from the PANDA Server by one or more PANDA
+Pilot. Currently, twenty PANDA Pilots are instantiated on four dedicated DSL
+nodes of Titan. These pilots are responsible for the packaging of a set of
+events into a PBS job script. Packaging is performed on the base of the
+information gathered about the availability of backfill resources on Titan.
+Depending on the amount of nodes available on Titan, a specific number of events
+is packaged into the PBS script. It should be noted that the events packaged
+into a PBS job can belong to different tasks and therefore users.
 
 \paragraph{Figure~\ref{fig:atlas_workflow} (4)} AthenaMP is used to perform the
-simulation of the events within the ATLAS detector, 1 AthenaMP thread for each
-core available on the Titan nodes. Currently, 16 threads are used for each node,
-each thread being able to simulate an average of 1 event every n??
-minutes?/second?. So far, no description has been given of the flow of data
-required to simulate each event.
+simulation of the events within the ATLAS detector, one AthenaMP thread for each
+core available on the Titan nodes. Currently, sixteen threads are used for each
+node, each thread being able to simulate an average of 1 event every n??
+minutes?/second? (see Figure~\ref{distribution_amp_titan}). So far, no
+description has been given of the flow of data required to simulate each event.
 
-Questions for Danila/Sergey:
-\begin{itemize}
-    \item Does each job of a PANDA Server (and therefore all the tasks of that specific job) belong to the same `user'/VO\@?
-    \item Does a PANDA Pilot pull jobs/tasks from a PANDA Server or does a PANDA server push jobs/tasks to a PANDA Pilot?
-    \item Does a PANDA Pilot pull jobs or tasks from a PANDA Server? Are the jobs within a PANDA Pilot the same jobs (i.e., the same container of the same tasks) in the PANDA Server?
-    \item Do we have 16 PANDA Pilots on a Titan DSL (is this the right acronym?) node? How many DSL nodes?
-    \item Can the events packaged into a PBS job belong to different submissions, i.e., different jobs, i.e., different users?
-    \item Do we have the distribution of the time taken on titan to process 1 event of the Monte Carlo workflow?
-\end{itemize}
+% Questions for Danila/Sergey:
+% \begin{itemize}
+%     \item Does each job of a PANDA Server (and therefore all the tasks of that specific job) belong to the same `user'/VO\@?
+%     \item Does a PANDA Pilot pull jobs/tasks from a PANDA Server or does a PANDA server push jobs/tasks to a PANDA Pilot?
+%     \item Does a PANDA Pilot pull jobs or tasks from a PANDA Server? Are the jobs within a PANDA Pilot the same jobs (i.e., the same container of the same tasks) in the PANDA Server?
+%     \item Do we have 16 PANDA Pilots on a Titan DSL (is this the right acronym?) node? How many DSL nodes?
+%     \item Can the events packaged into a PBS job belong to different submissions, i.e., different jobs, i.e., different users?
+%     \item Do we have the distribution of the time taken on titan to process 1 event of the Monte Carlo workflow?
+% \end{itemize}
+
+% Notes - answers
+%
+% - Does each job of a PANDA Server (and therefore all the tasks of that
+%   specific job) belong to the same ‘user’/VO? • Does a PANDA Pilot pull
+%   jobs/tasks from a PANDA Server or does a PANDA server push jobs/tasks to a
+%   PANDA Pilot?
+%
+% Same user task->job on the server.
+%
+% - Does a PANDA Pilot pull jobs or tasks from a PANDA Server? Are the jobs
+%   within a PANDA Pilot the same jobs (i.e., the same container of the same
+%   tasks) in the PANDA Server?
+%
+% Panda pilot is a edge service on titan, event-driven by backfill
+% availability. When available, it DOES THE PULL from the panda server.
+%
+% - Do we have 16 PANDA Pilots on a Titan DTN (is this the right acronym?)
+%   node? How many DTN nodes?
+%
+% Currently 20, thinking to go higher, on 4 DTN (Data Transfer Node), 5 pilots
+% for DTN, DTN that enable interactive data transfer. New DTN farm is being
+% rolled on. Our set up will change. More node, more pilots per node. This is
+% bound by DTN performance, overloading DTN (currently absence of data on why
+% these nodes are overladed. Network/memory should not be a problem, it is
+% probably purely CPU because of tools used to move files, e.g., GridFTP.)
+%
+% - Can the events packaged into a PBS job belong to different submissions,
+%   i.e., different jobs, i.e., different.
+%
+% They do belong to different tasks.
+%
+% - Do we have the distribution of the time taken on titan to process 1 event
+%   of the Monte Carlo workflow?
+%
+% To be added to Section VI about performance. See presentation:
+% https://drive.google.com/drive/u/0/folders/0BxWZQExiTsWNRTdMQXN4a1g3bHc

### 2017-02-06 WMSLiterature.tex
Related work reduced, fixed terminology errors in related works

@@
 \label{sec:related}
+
+In this section, we present other workload manager systems that have been developed to support LHC related but they have been used also to support other computing demanding projects.
+
+%\begin{itemize}
+%\item \emph{Clients}: consist in a set of APIs that allows users to submit job requests. Clients interact directly with DIRAC central services.
+%\item \emph{Resources}: they can be PC's, site clusters and Grids. Agents interact with them without distinction.
+%\end{itemize}
+
 \subsubsection{Alien}
@@
 \begin{itemize}
-\item \emph{Job Manager} (or \emph{Task Queue}), a database that keeps track of all the jobs submitted to the system and their current execution status;
+\item \emph{Task Queue}, a database that keeps track of all the tasks submitted to the system and their current execution status;
 \item \emph{Brokers}, they are the core of task executions and data transfers; they receive tasks in form of JDL,  keep them ordered by priority and send them to the CE for execution;
@@
 
-The job execution in Alien is usually distributed over several sites.
+The task execution in Alien is usually distributed over several sites.
 Each of these sites has at least one service called ClusterMonitor.
-On one side the Cluster Monitor is used to communicate with Central services (Job Manager and Broker), on the other side it can manage Computing Elements (CE) by starting and stopping them whenever it receives the signal.
-The CE is the resource in charge of the execution of the jobs. A CE usually is associated with a batch queue and can send the jobs to the worker nodes controlled by the queue.
+On one side the Cluster Monitor is used to communicate with Central services (Task queue and Broker), on the other side it can manage Computing Elements (CE) by starting and stopping them whenever it receives the signal.
+The CE is in charge of the execution of the tasks on the resources. A CE usually is associated with a batch queue and can send the tasks to the worker nodes controlled by the queue.
 
-The CE asks the Broker for jobs to execute by sending its description in form of Job Description Language (JDL).
-Once received the JDL, the Broker will try to match it with the JDL of the jobs in queue. If a match exists then the Broker sends the jobs JDL to the CE.
+The CE asks the Broker for tasks to execute by sending its description in form of Job Description Language (JDL).
+Once received the JDL, the Broker will try to match it with the JDL of thetasks in queue. If a match exists then the Broker sends the tasks JDL to the CE.
 Immediately after receving a job JDL, the CE create a new service on the worker node called ProcessMonitor.
 This service allows the CE (and the rest of Alien services through the CE) to interact with the job while is running \cite{Saiz2003}.
-This execution strategy is called ``pull mode'' due to the fact that CE asks for jobs.
-It is worth to mention that recent versions of Alien exploit gLite and CE-CREAM. The latter allows the system to by-pass the broker during the job submission \cite{Bagnasco2004}.
+This execution strategy is called ``pull mode'' due to the fact that CE asks for tasks.
 
@@
 DIRAC (Distributed Infrastructure with Remote Agent Control) Workload and Data Management System has been developed within the CERN LHCb project to manage the processing of detector data, Monte Carlo simulations, and end-user analyses \cite{Tsaregorodtsev2004}.
-DIRAC  architecture relies on four entities \cite{Paterson2010}:
+DIRAC  architecture relies on two entities \cite{Paterson2010}:
 \begin{itemize}
-\item \emph{Clients}: consist in a set of APIs that allows users to submit job requests. Clients interact directly with DIRAC central services.
-\item \emph{Services}: serve Clients and Agents by performing crucial operations such as Job Management, Configuration, Bookkeping and Accounting.
-\item \emph{Agents}: perform repetitive tasks like querying file catalogs,  monitoring of jobs on resources.
-\item \emph{Resources}: they can be PC's, site clusters and Grids. Agents interact with them without distinction.
+\item \emph{Services}: serve Clients and Agents by performing crucial operations such as Task Management, Configuration, Bookkeping and Accounting.
+\item \emph{Agents}: perform repetitive actions like querying file catalogs,  monitoring of jobs on resources.
 \end{itemize}
 In the same way of Alien, DIRAC implements a pull scheduling. Furthermore DIRAC was the first WMS to exploit the concept of Pilot Agent on the Grid \cite{Casajus2010}. 
-Pilots Agents are gLite jobs that are submitted to the grid when jobs arrive into the WMS. 
+Pilots Agents are gLite jobs that are submitted to the grid when tasks arrive into the WMS. 
 DIRAC pilot system has four main logical components:
@@
 \item a set of TaskQueueDirectors that submits JobWrappers to target DCRs;
-\item a MatchMaker that matches requests from JobWrappers to suitable tasks into TaskQeues.
+\item a MatchMaker that matches requests from JobWrappers to suitable tasks into TaskQueues.
 \end{itemize}
@@
 \item \emph{Startd}, controls the DCR resources. 
-\item \emph{Collector}, holds references to all the active
-Schedd/Startd daemons; 
+\item \emph{Collector}, holds references to all the activeSchedd/Startd daemons; 
 \item \emph{Negotiator} matches tasks queued in a Schedd to resources handled by a Startd.
@@
 From there on the execution model is the same as described for the HTCondor Glidein Service.
-
+\begin{comment}
 The resources managed by a single Glidein (i.e., pilot) are limited to compute resources.
@@
 \end{table*}
+\end{comment}

### 2017-02-06 WMSLiterature.tex
Inversion between WMSLiterature and Terminology

@@
 
-Since Alien exploits JDL, a  workload can be described task by task according to features such as: task priorities, the level of parallelism (one core, multi-core, MPI etc..) and  also the DCR that should be targeted for the execution.
+%Since Alien exploits JDL, a  workload can be described task by task according to features such as: task priorities, the level of parallelism (one core, multi-core, MPI etc..) and  also the DCR that should be targeted for the execution.
 
@@
 
-JDL and \emph{Transformation Management System} (TMS) are used for task description \cite{Tsaregorodtsev2006,Pacini2006}. The latter allows the user to describe complex data dependences. 
+%JDL and \emph{Transformation Management System} (TMS) are used for task description \cite{Tsaregorodtsev2006,Pacini2006}. The latter allows the user to describe complex data dependences. 
 \subsubsection{HTCondor Glidein and GlideinWMS}

### 2017-02-13 overall.tex
updates to the I/O section per Sergey's feedback and new graphs

@@
 
-The ATLAS framework has a configuration and initialization stage, which is
-called Athena. At this stage, the running job is assembled  on the fly from a
-large number of shared libraries.  Also, at this stage, every algorithm and
-service is being configured, at run time, by a corresponding set of Python
-scripts, which results in a large number of read I/O accesses to small python
-scripts, with many includes and imports Python calls.
+Athena, the ATLAS framework has a configuration and initialization stage. At
+this stage, the running job is assembled  on the fly from a large number of
+shared libraries.  Also, at this stage, every algorithm and service is being
+configured, at run time, by a corresponding set of Python scripts, which
+results in a large number of read operations accesses to small python scripts, with
+many includes and imports Python calls.
 
@@
 Listing~\ref{mdstrace} the metadata I/O activity for ATLAS exhibits a spike
-corresponding to the beginning of the runs (i.e., the Athena stage) before
-tapering off.
+corresponding to the beginning of the runs before tapering off.
 
@@
 
-Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
-improvement on Titan. The circled region illustrates the switch from Lustre to
-NFS-exported directory for hosting the ATLAS release.
+%Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
+%improvement on Titan. The circled region illustrates the switch from Lustre to
+%NFS-exported directory for hosting the ATLAS release.
 
-\begin{figure}[!htb]
-    \centering
-    \begin{tabular}{cc}
-        {\includegraphics[width=0.48\textwidth]{figures/panda-completed-jobs-sw-move.pdf}}\\
-    \end{tabular}
-    \caption{ATLAS performance improvement on Titan. The circled region shows the switch from Lustre to NFS-exported directory for hosting the ATLAS release.}
-\label{fig:atlas-perf-improvement}
-\end{figure}
+%\begin{figure}[!htb]
+%    \centering
+%    \begin{tabular}{cc}
+%        {\includegraphics[width=0.48\textwidth]{figures/panda-completed-jobs-sw-move.pdf}}\\
+%    \end{tabular}
+%    \caption{ATLAS performance improvement on Titan. The circled region shows the switch from Lustre to NFS-exported directory for hosting the ATLAS release.}
+%\label{fig:atlas-perf-improvement}
+%\end{figure}
 
@@
 
-ATLAS jobs are file read I/O heavy, as can be seen in
+ATLAS jobs issues a large number of file read operations, as can be seen in
 Table~\ref{panda-olcf-stats}. The maximum amount read by any ATLAS job in
@@
     \begin{tabular}{cc}
-        {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_hist.pdf}}\\
+        {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_finer_hist.pdf}}\\
     \end{tabular}
-    \caption{ATLAS file read I/O histogram on Titan for week of 10/25/16.}
+    \caption{ATLAS file read operation histogram on Titan for week of 10/25/16.}
 \label{fig:atlas-titan-io-read}
@@
     \begin{tabular}{cc}
-        {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_hist.pdf}}\\
+        {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_finer_hist.pdf}}\\
     \end{tabular}
@@
 {fig:atlas-titan-io-written}, the distribution between read and written amount
-of data per job is quite different from one another. The read I/O distribution
-per job shows a long tail, ranging from 25 GB to 250 GB, while the written
+of data per job is quite different from one another. The read operation distribution
+per job shows a long tail, ranging from 12.5 GB to 250 GB, while the written
 amount of data has a very narrow distribution.
@@
 On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file $open()$
-operations and 5 file $close()$ operations per second. As can be seen from
-figures~\ref{fig:atlas-titan-file-open} and~\ref{fig:atlas-titan-file-close},
-they exhibit similar distributions. The maximum number of file $open()$
-operations are around 170 on average and the maximum number of file $close()$
-operations are 39 on average per second per job. The total number of file
-$open()$ operations is 172,089,760 for the observed window of 174 hours, while
-the total number of file $close()$ operations stand at 40,132,992 for the total
-1,175 ATLAS PanDA jobs in the same observation window. The difference between
-these two values is puzzling and it is under investigation at the time of
-writing this paper. {\sarp{can someone take a look at this? I would expect at
-least similar values for number of opens and closes and there is vast
-difference. Can it be that ATLAS jobs do a lot of opens without following
-closes, or perhaps multiple opens for the same files over and over?}}
+operations and 5 file $close()$ operations per second. The file $open()$
+operations listed here don't include the file $stat()$ operations. As can be
+seen from figures~\ref{fig:atlas-titan-file-open}
+and~\ref{fig:atlas-titan-file-close}, they exhibit similar distributions. The
+maximum number of file $open()$ operations are around 170 on average and the
+maximum number of file $close()$ operations are 39 on average per second per
+job. The total number of file $open()$ operations is 172,089,760 for the
+observed window of 174 hours, while the total number of file $close()$
+operations stand at 40,132,992 for the total 1,175 ATLAS PanDA jobs in the same
+observation window. The difference between these two values is puzzling and it
+is under investigation at the time of writing this paper. One possible
+explanation is that ATLAS PanDA jobs perhaps don't call a file $close()$
+operation per every file $open()$ issued.
 

### 2017-02-14 Experiments.tex
Description experiments - Changed

--- 2016-10-04 Experiments.tex
@@
-\jhanote{Not sure this should be under ``PANDA:Future/ROAD MAP?''}
+In the batch queue,  queue time becomes dominant but, at the same time, we have more freedom to decide the parameters of the slot.
 
-The execution of jobs on TITAN follows a three step scheme: i) as first, TITAN's Backfill queue communicates the presence of a new allocation\mtnote{communicates the availability of idle resources?
-The term `allocation' may indicate the amount of resources available to the project.} i.e.
-a set of unused resources that can be used for a pre-determined time period\aanote{How many offers does PanDA receive per time unit?
-Average nodes per offer?}; ii) then, PanDA must accept or decline the allocation; iii) if the allocation is accepted, PanDA wraps a certain amount of events in a job and sends the job for execution.\aanote{I am supposing that the concept of event has been already introduced at this point}.
+\subsection{Experiments}
 
-At least in principle, PanDA can reject an allocation because it does not fit the requirements in terms of number of nodes or wall-time unit; however, this situation is rare because events are single core Athena-MP threads\mtnote{Should we consider events inputs of Athena-MP jobs, with each event processed in a single thread by a---to be specified---module of Athena-MP?} that require on average twenty minutes for being executed.
-Thus, the decomposition in event is fine grained once compared to the average allocation \aanote{Which correspond to\ldots}.
+In this section we describe experiments that have been performed by running the ATLAS workload on TITAN's batch queue by means of the next generation executor.   
+The importance of these experiments is twofold: on the one side, we illustrate the performances of the NGE; on the other side, they help to characterize TITAN's batch queue and we show that it can be used as alternative to backfilling.
+Since we aim to study the response time of batch queue as it is perceived by a common user,  special privileges have been asked to system administrator.
+%This is particularly important for the development of future execution strategies because queue time becomes dominant in
 
-As a consequence, PanDA accepts the major part of the proposed allocations by selecting a number of events that maximize the utilization of the resources during the assigned time interval.
+Experiments aim to show the number of completed events as function of the time. 
+Each experiment corresponds to a number of different and independent runs that is meaningful from a statistical point of view.
+Each run consists in the sustained execution of the ATLAS workload for a pre-determined period of time \aanote{A week?
+10 days?}.
+Since execution of events is carried on by means NGE jobs, each run is based on a script that maintains constant pressure on TITAN's batch queue.
+The main tasks of the script are to:  i) guarantee that there are constantly four NGE jobs eligible for running on TITAN's queue (note that four jobs is the maximum allowed in TITAN's batch queue \ref{}); ii)  bind Athena-MP jobs to the NGE jobs; iii) perfom garbage collection.
+No limit in the number of NGE jobs is imposed.
 
-On one hand, this \jhanote{I would use the word execution strategy carefully.
-Maybe a more general term such as execution model, a special type of which is execution strategy?} execution strategy leads to a drastic reduction of the queue time and, consequently, it guarantees high-performances.
-On the other hand, this strategy is tailored for the execution of ATLAS workload on TITAN by using the Backfill queue\mtnote{Do we have a queue dedicated to backfilling on Titan or should we speak of backfilling done on one or more queues?}.
+The simulation of events is due to the following execution model: the script binds a number of Athena-MP jobs that is equal to the number of nodes on which NGE job is running; then, NGE job spawns the Athena-MP jobs on the nodes; finally, the Athena-MP jobs will simulate a number of events that is proportional to the wall-time of the NGE job.
+As a consequence, the number of nodes determines the number of events that can be simulated concurrently whereas the wall-time determines the number of events that can be simulated one after the other.
+These two parameters together determine the size of the NGE job.
 
-Questions such as \emph{``Is this the best strategy possible?''} or \emph{``Can PanDA achieve similar performances by using TITAN's batch queue?''}  remain unanswered.
+The size of the NGE job is constant within an experiment. Thus, each experiment is fully characterized by the size of the NGE jobs that it send on the nodes.
 
-For this reason, in the remaining of this section we propose experimental results to address them. The idea behind the experiments is to show the behaviour of ATLAS workload on TITAN's batch queue.
+Each execution of Athena-MP is guaranteed to be successfull and, on average, comparable with those executed by PanDA in production.
 
-First, we investigate resource allocation as a function of time (wall-time) and space (number of nodes\mtnote{cores?}) in order to understand how TITAN's scheduler and workload behave under different allocation requests.
-This analysis includes two bulks of experiments: the first, generated ``in silico'', is based on simulations of the TITAN's scheduler by means of the Moab simulator, whereas the second, generated ``in vivo'',  considers real submissions of ATLAS workload on TITAN's batch queue.
+Table \ref{tab:experiments} provides the list of the experiments that have been performed.
 
-By using the batch queue, we move in a situation where the queue time becomes dominant but, at the same time, we gain more freedom to decide the parameters of the allocation.
-For this reason, the second set of the experiments aims to find sub-optimal parameters for resource allocation.
+\begin{table*}
+\begin{center}
+\begin{tabular}{|c|l|l|l|l|}
+  \hline
+Experiment &\# nodes &wall-time (hours) & run duration(days) & \# runs   \\
+\hline
+1 & &  & &   \\
+2 & &  & &  \\
+3 & &  & &  \\
+4 & &  & &  \\
+\hline
+\end{tabular}
+\end{center}
+\caption{List of experiments}\label{tab:experiments}
+\end{table*}
 
-Additionally, due to the dynamics of the Backfill queue, PanDA has a high probability to re-acquire an allocation immediately after it has released one \aanote{Are we able to quantify this ``immediately''?}.
+Each experiment has been designed by choosing the maximum number of nodes and the maximum wall-time of each TITAN bin \ref{} \aanote{Are we really doing this???}.
+The length of the experiment has been decided.....\aanote{It was 10 times the wall-time or something like that}.
+
+Figure \ref{fig:avtEvts} \aanote{(DRAFT)}depicts average number of completed events as function of the time.
+Figure \ref{fig:avtTimes} \aanote{(DRAFT)} depicts the average queue time and execution time per experiment.
+
+\begin{figure}[!htb]
+        \includegraphics[width=0.48\textwidth]{./figures/draft/mean.pdf}
+    \caption{Average number of completed as function of the time.}
+\label{fig:avtEvts}
+\end{figure}
+
+\begin{figure}[!htb]
+        \includegraphics[width=0.48\textwidth]{./figures/draft/AvgT.pdf}
+    \caption{Average queue time and execution time of the experiments.}
+\label{fig:avtTimes}
+\end{figure}
+
+Because of the definition of the experiment the results shown in Figure \ref{} are alike (but not entirely comparable to) those provided by in Figure \ref{} where the performances of backfilling are shown.
+
+% For this reason, the second set of the experiments aims to find
+%sub-optimal parameters with which we can minimize the trade-off between the size
+%of a slot and the time spent in queue waiting for that slot to become available.
+%In other words, we aim to minimize the completion time by finding the best
+%trade-off between execution time and queue time.
+%
+%This execution model introduces slot utilization as one of the key factors for
+%high-performances. This happens because, in order to minimize the time spent in
+%queue, we might asks for slots in advance and, then we could not be able to
+%saturate them when they become available. Thus, this strategy requires a new
+%functionality that allows the job to receive and execute new events while it is
+%already running on the resources. In order to do that we perform the experiments
+%by using a new generation executor that implements such functionality.
+%
+%As last observation, it is important to point out that the percentage of
+%utilization of a slot is minor problem with the current implementation because,
+%due to the dynamics of the Backfill queue, PanDA has a high probability to
+%re-acquire a slot immediately after it has released one\aanote{Are we able to
+%quantify this ``immediately''?}.

### 2017-02-14 workflow.tex
Added Backfill description -- First attempt plus questions

@@
 % https://drive.google.com/drive/u/0/folders/0BxWZQExiTsWNRTdMQXN4a1g3bHc
+
+\subsubsection{Backfill - High-level Description}
+PanDA pilots  interrogate TITAN's queue continuously in order to find unused resources.
+This is done by using \emph{showbf} which provides the current number of unused resources and the duration of their idle time.
+
+If PanDA server queue is not empty, each PanDA pilot tries to have five PanDA jobs constantly running on the resources.
+Therefore, if unused resources are available, a PanDA pilot sends a new PanDA job on TITAN's queue every time that one of its PanDA jobs terminates its execution.
+
+PanDA pilot implements a greedy approach i.e.
+it tries to obtain all the available resources with a single PAnDA job\aanote{Verify this}.
+The size of a PanDA job cannot exceed 300 nodes \aanote{why???}.
+
+After the submission of a PanDA job,  the PanDA pilot waits up to five minutes.
+If the PanDA job does not start in such time interval then the PanDA pilot deletes the job from TITAN's queue and try to resubmit the job.
+
+The resubmission might change the size of the PAnDA job is the number of available nodes is changed in the meantime.
+
+No job is submitted to TITAN's queue If less than \emph{XXX} nodes are available or if they are not  available for at least 2 hours.
+
+Once the PAnDA job is running on the nodes, it spawns an Athena-MP job per node.
+Each Athena-MP job runs 16 concurrent processes that are in charge for the simulation of 100 events.
+No Athena-MP job is submitted after another.
+Therefore, PanDA job terminates immediately after all the Athena-MP jobs completed their simulation.
+
+Questions:
+\begin{itemize}
+\item Which is the component that performs showbf? The Pilot?
+\item Do pilots compete for the same available nodes?
+\item There is any restriction in terms of minimum wall-time or minimum number of nodes?
+\item If yes, Are there restrictrions imposed by ORNL or by PanDA server?
+\item Is PanDA job walltime fixed or flexible? Two hours?
+\item Does the pilot initiate input staging as a consequence of available nodes? if yes, what happens if input data is not there? 
+\end{itemize}

### 2017-02-14 Experiments.tex
some comments in Experiment Design.

@@
 
-In this section we describe experiments that have been performed by running the ATLAS workload on TITAN's batch queue by means of the next generation executor.   
-The importance of these experiments is twofold: on the one side, we illustrate the performances of the NGE; on the other side, they help to characterize TITAN's batch queue and we show that it can be used as alternative to backfilling.
+In this section we describe experiments that have been performed by running the ATLAS workload on TITAN's batch queue by means of the next generation executor (NGE) \jhanote{We'll need to reference the section/sub-section where this is discussed}.
+
+The importance \jhanote{aim?} of these experiments is twofold: on the one side, we illustrate the performances of the NGE; on the other side, they help to characterize TITAN's batch queue and we show that it can be used as alternative to backfilling.
 Since we aim to study the response time of batch queue as it is perceived by a common user,  special privileges have been asked to system administrator.
+\jhanote{I don't think the second point captures accurately why we are interested in the NGE.
+Please refine}
+
 %This is particularly important for the development of future execution strategies because queue time becomes dominant in
 
-Experiments aim to show the number of completed events as function of the time. 
+Experiments aim to show \jhanote{calculate? determine?} the number of completed events as function of the time. 
 Each experiment corresponds to a number of different and independent runs that is meaningful from a statistical point of view.
-Each run consists in the sustained execution of the ATLAS workload for a pre-determined period of time \aanote{A week?
+\jhanote{previous sentence needs fixing} Each run consists of a sustained execution of the ATLAS workload for a pre-determined period of time \aanote{A week?
 10 days?}.
-Since execution of events is carried on by means NGE jobs, each run is based on a script that maintains constant pressure on TITAN's batch queue.
-The main tasks of the script are to:  i) guarantee that there are constantly four NGE jobs eligible for running on TITAN's queue (note that four jobs is the maximum allowed in TITAN's batch queue \ref{}); ii)  bind Athena-MP jobs to the NGE jobs; iii) perfom garbage collection.
+\jhanote{How do we determine this time window?} Since execution of events is carried on by means \jhanote{what is an NGE job?} NGE jobs, each run is based on a script that maintains constant pressure on TITAN's batch queue.
+\jhanote{please separate what we are trying to achieve from how..}
+
+The main tasks of the script are to:  i) guarantee that there are constantly four NGE jobs eligible for running on TITAN's queue (note that four jobs is the maximum allowed in TITAN's batch queue \ref{}); ii)  bind Athena-MP jobs to the NGE jobs; iii) perform garbage collection.
 No limit in the number of NGE jobs is imposed.
 
-The simulation of events is due to the following execution model: the script binds a number of Athena-MP jobs that is equal to the number of nodes on which NGE job is running; then, NGE job spawns the Athena-MP jobs on the nodes; finally, the Athena-MP jobs will simulate a number of events that is proportional to the wall-time of the NGE job.
+The simulation of events is due to the following execution model: the script\jhanote{the script is too casual/imprecise} binds a number of Athena-MP jobs that is equal to the number of nodes on which NGE job is running; then, NGE job spawns the Athena-MP jobs on the nodes; finally, the Athena-MP jobs will simulate a number of events that is proportional to the wall-time of the NGE job.
 As a consequence, the number of nodes determines the number of events that can be simulated concurrently whereas the wall-time determines the number of events that can be simulated one after the other.
 These two parameters together determine the size of the NGE job.
+\jhanote{please refine.
+difficult to understand}
 

### 2017-02-14 introduction.tex
Preparatory work to move definitions

@@
 \section{Introduction}\label{sec:intro}
-The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades.
-Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers.
-Thousands of physicists analyze petascale  data volumes daily.
-One of  the LHC experiments  the ATLAS  \cite{Aad:2008}  utilizes the PanDA  workload management system \cite{Maeno2011} (WMS) for distributed data processing and analysis.
-The ATLAS Computing model \cite{Atlas2005} is based on a Grid paradigm \cite{Foster:1998}, with multilevel, hierarchically distributed computing  and storage resources.
-Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
-PanDA has a highly scalable architecture.
-Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade.
-PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middleware).
-This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
-PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user analysis, while actively evolving to meet rapidly changing science needs.
-Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
+
+The Large Hadron Collider (LHC) is created to explore the fundamental properties
+of matter for the next decades. Since LHC start-up in 2009, the experiments
+collected and distributed hundreds of petabytes of data worldwide to hundreds of
+computer centers. Thousands of physicists analyze petascale data volumes daily.
+One of the LHC experiments, the ATLAS~\cite{Aad:2008}, utilizes the Production
+and Distributed Analysis (PanDA) workload management system~\cite{Maeno2011}
+(WMS) for distributed data processing and analysis. The ATLAS Computing
+model~\cite{Atlas2005} is based on a Grid paradigm~\cite{Foster:1998}, with
+multilevel, hierarchically distributed computing and storage resources. PanDA
+has been developed to meet growing ATLAS production and analysis requirements
+for a data-driven workload management system capable of operating at LHC data
+processing scale.
+
+PanDA has a highly scalable architecture. Scalability has been demonstrated in
+ATLAS through the rapid increase in usage over the past several years of
+operations, and PanDA is expected to meet the continuously growing computing
+requirements of ATLAS over the next decade. PanDA was designed to have the
+flexibility to adapt to emerging computing technologies in processing, storage,
+networking as well as the underlying software stack. This flexibility has also
+been successfully demonstrated through the past six years: computing centers in
+ATLAS, spanning many continents, were seamlessly integrated into PanDA\@.
+
+PanDA manages a wide spectrum of workloads, ranging from raw data processing to
+Monte Carlo simulation and user analysis, while constantly evolving to meet
+rapidly changing science needs. Today, PanDA serves several thousand users,
+managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU
+cores which process more than a million jobs per day (Figure~\ref{fig:daily}).
+
 \begin{figure}
@@
         \includegraphics[width=\columnwidth]{figures/DailyJobs.png}
-\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
+        \caption{Daily completed jobs on ATLAS Grid for the past 12 month}
     \end{center}
+\label{fig:daily}
 \end{figure}
+
+In this paper, we describe how PanDA has been engineered to execute a specific
+stage of the ATLAS Monte Carlo workflow on Titan, the larger high-performance
+computing HPC system currently available in the USA\@.\mtnote{Explain the
+benefits offered by Titan in terms of multithreading per node and possibly large
+amount of concurrent nodes. Introduce also the notion of backfill.} This extends
+the scope of PanDA's compute model, integrating both high-throughput and
+high-performance computing resources and enabling the concurrent execution of
+both  single and multi-core jobs. The integration of PanDA and Titan went
+through three main engineering phases: (i) feasibility study and rapid
+prototyping of an initial solution; (ii) progressing scaling of the  prototype
+to saturate the available resources; (iii) study of a product-grade architecture
+for generic HPC resources. Both phase i and ii have been completed enabling the
+execution of up to eight million jobs a week on Titan. A prototype has been
+engineered to support phase iii and experimental data are being collected.
+
+In the next section we introduce \ldots.

### 2017-03-01 PandaOverview.tex
Section III

@@
-PanDA consists of several interconnected subsystems, most of them are built from off-the-shelf and Open Source components. Figure \ref{fig:architecture} shows a schematic view of the PanDA system.
-In the following we will briefly describe system’s architecture and components.
+\mtnote{TODO: replace `resource' with `site'; `http' to `HTTP'?}
+
+The design of the Production and Distributed Analysis (PanDA) system started in
+2005 to develop a WMS for the ATLAS experiment. PanDA went into production for
+the LHC run 1 between 2009 and 2013 and was then partially redesigned to be
+deployed on run 2, between 2015 and 2018. New design features are constantly
+prototyped and tested on small scale but, in this section, we summarize the
+design, architecture, and execution process of the current state of the
+practice.
+
+\subsection{Design}
+
+% \begin{itemize}
+%   \item Serve both individual and production workflows
+%   \item hide WLCG resources behind a single interface
+%   \item queue-based interface to the application layer
+%   \item pilot-based
+%   \item job-based
+%   \item global view of the properties and states of all the jobs
+%   \item global view of the available resources as a single worldwide
+%      processing resource.
+%   \item centralize data management
+%   \item X509 AAA
+%   \item Multitenancy
+% \end{itemize}
 
 \begin{itemize}
-\item PanDA Server
-\item Pilot
-\item Factory
-\item PanDA monitoring
-\item JEDI ??
-\item What else?
+  \item No MPI tasks?
+  \item Pilot multitenancy?
+  \item PanDA Server Brokerage component and Event service interaction?
+  \item PanDA Server Data Service component and Event Service interaction?
 \end{itemize}
+
+PanDA's application model primarily assumes tasks, as single entities or
+organized into workloads or workflows. Since 2005, a certain amount of
+parallelism has been progressively introduced assuming both single core and
+multi-threaded tasks~\cite{multithreaded_tasks} but, so far, no MPI tasks have
+been considered for production. Tasks are assumed to have at least two types of
+data dependences: one or more input file and one or more data sources. Task are
+also supposed to be relatively self-contained, capable of setting up their own
+execution environment or having a minimal set of common dependences. Different
+types of task executable are assumed, both in terms of user-defined batch
+scripts or application frameworks for group of users.
+
+PanDA's user model is based on multitenancy and the support of at least two
+types of HEP users: individual researchers and groups executing production-grade
+workflows. Users are considered free to submit tasks and workflows at any point
+in time, directly or via dedicated application frameworks. Consistently, PanDA's
+security model is based on separation between authentication, authorization and
+accounting for both single users and group of users. Both authentication and
+authorization are based on digital certificates and the X.509 standard.
+
+Currently, PanDA's execution model is based on five main abstractions: task,
+job, queue, pilot, and event. Tasks are assumed to be partitioned into one or
+more job, possibly of variable sizes. Jobs have both attributes and states and
+are queued into a global queue for execution. Prioritization and binding of jobs
+are assumed to depend on the attributes of each job. Pilot abstracts resource
+capabilities and jobs are bound to pilots and executed by consuming their
+resources. Consistently, pilots are considered to be multitenant as jobs
+belonging to multiple workflows and/or users could be bound and executed on
+them.
+
+PanDA' data model uses events as the unit of data. Each event represents a
+collision captured in the detector and one or more events can be packaged into
+files or data source with other abstractions. As with jobs, data have both
+attributes and states and some of the attributes are linked to those of jobs so
+to create an association between the two otherwise separated entities. Data are
+assumed to be stored centrally, replicated when necessary and transferred over
+the network. This assumptions applies to both input and output data.
+
+PanDA design supports provenance and traceability for both jobs and data.
+Attributes enable provenance by linking jobs and data items, providing
+information like ownership or project affiliation. States enable traceability by
+providing information about the stage of the execution in which each job or data
+item is or have been. Some attributes are assumed to be immutable across
+execution and jobs and data items are assumed to be always in one and only one
+state.
+
+\subsection{Architecture}
+
+The implementation of PanDA consists of several interconnected subsystems, most
+of them built from off-the-shelf and Open Source components. At a high level,
+PanDA's architecture has six subsystems: PanDA Server, AutoPyFactory, PanDA
+Pilot, PanDA Monitoring, JEDI, and Event Service. Subsystems communicate via
+dedicated interfaces, implemented as API or http messaging. Databases are used
+to store eventful entities like tasks, jobs and events, and to store information
+about resources, logs, and accounting. Each subsystems is implemented as one or
+more modules as shown in Figure~\ref{fig:architecture}. In the following we
+briefly describe the overall architecture of PanDA and its components. The
+description provides all the details required to understand how HPC resources
+have been integrated within PanDA. For a full description of PanDA architecture
+see~\cite{wiki, papers}.
+
+Capabilities:
+\begin{itemize}
+  \item interoperable across job submission systems
+  \item asynchronous distribution of input/out data organized as file collections associated to (set of) jobs.
+  \item policy-aware scheduling
+  \item Supporting multiple types of executable code.
+\end{itemize}
+
+% Components:
+% \begin{itemize}
+%   \item monitoring system
+%   \item information system (static/dynamic information)
+%   \item data management system
+%   \item
+% \end{itemize}
+
+% \begin{itemize}
+% \item PanDA Server
+% \item Pilot
+% \item Factory
+% \item PanDA monitoring
+% \item JEDI ??
+% \item What else?
+% \end{itemize}
 
@@
     \includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
-\caption{Schematic view of the PanDA system\label{fig:architecture}. 
-  Originally PanDA was designed for grid infrastructure... In this paper
-  is focussed on the HPC use case; there are differences from  the grid use
-  case. This paper discusses how PanDA has been adapted to execute a workload on
-  HPC resources. The schematic overview is presented for workload X on Titan......}
   \end{center}
+\caption{\mtnote{This diagram is a placeholder. I am working on a substitute
+that will be referenced when describing PanDA's subsystems.}Schematic view of
+the PanDA system\label{fig:architecture}, as deployed for production in 2017.
+Black boxes indicate PanDA's subsystems; Orange boxes components of a subsystem;
+purple boxes databases; green arrows http messaging interfaces; red arrows API;
+purple arrows database communications.}
+% Originally PanDA was designed for grid infrastructure... In this
+% paper is focussed on the HPC use case; there are differences from
+% the grid use case. This paper discusses how PanDA has been adapted
+% to execute a workload % on HPC resources. The schematic overview is
+% presented for workload X on Titan......
 \end{figure}
-\subsection{PanDA Server}
-The PanDA server is the main component of the system, It provides  a task queue managing all job information centrally.
-The PanDA  server receives jobs through the client interface into the task queue, upon which a brokerage module  operates to prioritize  and assign work on the basis of job type, priority, input data and its locality, and available CPU resources.
-The PanDA  server operates as a web service.
-It runs on Apache web server, interacting with the back-end database running  on separate servers.
-It uses the Apache worker  model, with many independent   processes  handling client requests  in  parallel.
-Since all  state is maintained  in  the central database,  the PanDA Server application instance itself is stateless.
-Currently production version of PanDA utilizes Oracle database backend, but PanDA  can also work with MySQL family of databases.
 
-\subsection{PanDA Pilot}
- PanDA is a pilot based workload  management system.
-In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload processing on a worker  node) are submitted to sites.
-When these  pilot jobs start on a  worker node they contact a central  server  to retrieve  a real payload (i.e., an end-user job) and execute it.
-Using these pilot-based  workflows helps to improve job reliability, optimize resource utilization, allows for opportunistic  resources usage, and mitigates  many of the problems  associated with the inhomogeneities found on the Grid.
-The pilots themselves  do not contain all the functionality needed to request a payload  job from the Panda Server.
-First, the complete  set of pilot code is downloaded via HTTP from a  central Subversion repository.
-The repository works with Apache web server configured  with a  memory-based   web proxy (Squid).
-The purpose  of the cache  is to reduce  the request load on the back-end Subversion  server.
-That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server.
-The benefits of this system include a high-performance code download service combined with code updates still being immediately  available  as soon as they are committed to source code control.
+\paragraph{\textbf{PanDA Server}} this subsystem is the central hub service of
+PanDA and it operates as a web service. It runs on Apache web server,
+interacting with the back-end database running on separate servers. It uses the
+Apache worker model, with many independent  processes handling client requests
+in parallel. Since all state is maintained in the central database, the PanDA
+Server application instance itself is stateless. Currently production version of
+PanDA utilizes Oracle database backend, but PanDA can also work with MySQL
+family of databases.
 
-\subsection{PanDA Factory}
-As a pilot-based  system, PanDA  requires some way to get the initial PanDA pilot onto worker nodes at sites.
-This is done with the help of a component  called AutoPyFactory (APF).
-APF runs in a single daemonized process, launching  a separate thread for each internal  workflow.
-Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote.
-The behavior of these APF workflows is determined by the combination of a set  of plugins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.
+% . It provides a task queue
+% managing all job information centrally. The PanDA server receives jobs through
+% the client interface into the task queue, upon which a brokerage module
+% operates to prioritize and assign work on the basis of job type, priority,
+% input data and its locality, and available CPU resources. The PanDA server
 
-\subsection{PanDA Monitoring}
-The PanDA Monitor is a web-based dashboard-style  graphical application.
-It runs on Apache and interacts with the back-end database for persistence.
-The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory.
-The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.
+PanDA Server implementation consists of four components, all implemented in
+Python and communicating via Python APIs: Task Buffer, Brokerage, Job
+Dispatcher, and Data Service. Task Buffer is the management component of PanDA's
+global queue and offers functionalities to: (i) store jobs into the global
+queue; (ii) provide information about the state of each stored job; and (iii)
+update the state of each job upon communication with the other three components
+of PanDA Server. The Brokerage component enables (i) matching job attributes
+with site and pilot attributes; (ii) managing the dispatch of input data to
+processing sites; and (iii) deriving PanDA's data pre-placement requirement.
+% The job/pilot matching algorithm assigns a weight to each candidate pilot and
+% after applying a set of policies, the best ten candidates are taken.
+The Job Dispatcher implements functionalities to: (i) receive requests from
+active pilots for new jobs to process; (ii) retrieving from Task Buffer the jobs
+with highest priority that can be executed on the resource on which the
+requesting pilot has been instantiated; (iii) creating a wrapper for the
+selected job, tailored to the resource of the requesting pilot; and (iv)
+monitoring the state of the job execution on the pilot, updating the Task Buffer
+component. Finally, the Data Service component offers functionalities to: (i)
+retrieve information from Task Buffer and Brokerage about the data requirements
+of each job and the resource to which each job has been assigned; (ii) enable
+interaction with the ATLAS Distributed Data Management (DDM)~\cite{atlas_ddm} to
+make input files available to the jobs and manage their output data.
+
+\paragraph{\textbf{PanDA Pilot}} this subsystem is execution environment for
+PanDA's jobs. Implemented in Python, a PanDA Pilot is launched by a wrapper
+submitted to a target resource. Wrappers are tailored to the specific interfaces
+and middleware exposed by each site so that PanDA Pilot remains
+site-independent. PanDA Pilot implements four main functionalities: (i)
+preparing the execution environment for PanDA's jobs, including recovering and
+cleaning up previously failed jobs on that site and collecting information about
+the compute, data, and network capabilities of the site; (ii) requesting a job
+for execution to the Job Dispatcher component of PanDA Server; (iii) initiating
+and monitoring the execution of pulled jobs, including setting up the job,
+transferring its input when required, executing the payload, transferring
+output, and checking log files and local space; and (iv) clean up when the
+payload has finished to execute.
+
+% PanDA is a pilot based workload management system.
+% In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload
+% processing on a worker node) are submitted to sites. When these pilot jobs
+% start on a worker node they contact a central server to retrieve a real
+% payload (i.e., an end-user job) and execute it.
+
+PanDA Pilot implements the pilot paradigm~\cite{pilot_review} and, as such, it
+decouples resource acquisition from job execution via multi-stage scheduling.
+This has several benefits, including increasing the number of completed jobs per
+unit of resource by lowering the time taken between finishing the execution of a
+job an starting a the execution of a new job; enabling both concurrent and
+sequential job execution depending on the number of cores available to the pilot
+and the time for which the remain available; and exposing a unified and
+consistent interface for the scheduling of jobs, independent from the type of
+resource on which the pilot has been instantiated and the middleware installed
+on the site. The specifics implementation of PanDA Pilot enables also better
+reliability by real-time monitoring of the job execution process, and a certain
+degree of fault-tolerance by recovering information and, possibly, output files
+from jobs that failed, even before instantiating the PanDA Pilot.
+
+% Using these pilot-based workflows helps to improve job reliability, optimize
+% resource utilization, allows for opportunistic resources usage, and mitigates
+% many of the problems associated with the inhomogeneities found on the Grid.
+
+It should be noted that upon bootstrapping, a PanDA Pilot download major parts
+of its runtime code a central Subversion repository via HTTP. The repository
+works with an Apache web server, configured with a memory-based web proxy
+(Squid). The purpose of the cache is to reduce the request load on the
+back-end Subversion server. That improves performance \mtnote{Do we have
+data/numbers?} since only occasional queries trigger a full lookup on the
+back-end Subversion system, and most external queries are pulled from memory
+on the front-end web server. The benefits of this system include a
+high-performance code download service combined with code updates still being
+immediately available as soon as they are committed to source code control.
+
+\paragraph{\textbf{AutoPyFactory}} this subsystem implements a factory to submit
+pilots locally and to both grid and cloud sites. Implemented to be executed as a
+single demonized process, AutoPyFactory enable the concurrent spawning of
+multiple job submission processes called APFQueue. Each APFQueue can communicate
+with PanDA Server and the Brokerage module to collect information about the
+state of the jobs it manages and, on the base of this information, submit new
+pilots to one site per APFQueue. AutoPyFactory implements functionalities to
+evaluate the number of pilots already instantiated on each site, and to apply a
+consistent and adequate pressure on their batch queues. Submission is performed
+via the Condor-G batch system~\cite{}, enabling submission to multiple type of
+sites. AutoPyFactory exposes a dedicated monitoring service, feeding information
+to the PanDA Monitoring subsystem. A proxy manager is also run, enabling
+certificate-based authentication and authorization.
+
+% As a pilot-based system, PanDA requires some way to get the initial PanDA
+% pilot onto worker nodes at sites. This is done with the help of a component
+% called AutoPyFactory (APF). APF runs in a single daemonized process,
+% launching a separate thread for each internal workflow. Each one of these
+% internal workflows typically serves a single job queue as defined in WMS, and
+% delivers pilots to a single batch queue, either local or remote. The behavior
+% of these APF workflows is determined by the combination of a set of plugins,
+% invoked in a fixed order, in a loop, each one in charge of the performance of
+% a well defined action.
+
+\paragraph{\textbf{PanDA Monitoring}} this subsystem is implemented as a
+web-based, dashboard-style graphical application. It runs on Apache and
+interacts with the back-end database of PanDA Server to enable persistence.
+PanDA Monitoring enables users and site administrators to gather information
+about the status of current and past jobs, data movement, and pilot factory.
+PanDA Monitoring allows users access to all job log files on job by job basis,
+greatly simplifying code debugging and failure analysis in a distributed
+computing environment.
+
+\paragraph{JEDI} \mtnote{Should we include it?}
+
+\paragraph{Event Service} \mtnote{Should we include it?}
+
+\subsection{Execution Process}
+
+\mtnote{Probably move this to the workflow}
+\mtnote{Think whether a comparison with other WMS based on the review we did
+makes sense in this context.}
+
+\mtnote{Summary: JEDI retrieves task descriptions and source files, converting
+them into job descriptions and events sets. Jobs are submitted to PanDA Server
+while events to the Event Service. The PanDA Server schedules jobs to pilots
+that have been submitted to resources via the PanDA Factory. Once executed, jobs
+pulls events from the Event Service as input and push output back to JEDI.}

### 2017-03-01 introduction.tex
Notes on introduction
Motivation for HPC and Simulations

@@
 In the next section we introduce \ldots.
+
+Why Titan?
+\begin{itemize}
+    \item A lot of slow (relative the grid) and homogeneous cores together
+    \item Grid is saturated
+    \item Enable HPC as a calss of computing resource
+    \item Enable future DOE experimental and observational capabilities on HPC
+    \item Why is so important: more data, run 2 and run 3.
+    \item growth of grid is flat (economic model) and saturated. We need more CPU because we have more data.
+    \item ATLAS spend most of the time on simulations this is why we want to offload simulations - that happen happen to be performed via AthenaMP.
+\end{itemize}

### 2017-03-04 PandaEvolution.tex
JEDI and double execution process

@@
 \subsection{Use of Supercomputers with PanDA}
+
+\begin{itemize}
+    \item Preserved logic by adapting PanDA Pilot to work as a (user land) persistent service instead of being submitted to the worknode.
+This was necessary because workernodes would not be able to communicate with PanDA Server.
+    \item SAGA is used to submit jobs to the local batch system. Submitting jobs replaced spawning processes in the PanDA Pilot.
+    \item Sergey provided architecture and reference paper.
+\end{itemize}
 

### 2017-03-04 PandaOverview.tex
JEDI and double execution process

@@
 2005 to develop a WMS for the ATLAS experiment. PanDA went into production for
-the LHC run 1 between 2009 and 2013 and was then partially redesigned to be
-deployed on run 2, between 2015 and 2018. New design features are constantly
-prototyped and tested on small scale but, in this section, we summarize the
+the LHC Run 1 between 2009 and 2013 and was then partially redesigned to be
+deployed on Run 2, between 2015 and 2018. In this section, we summarize the
 design, architecture, and execution process of the current state of the
@@
 
+% -----------------------------------------------------------------------------
 \subsection{Design}
-
-% \begin{itemize}
-%   \item Serve both individual and production workflows
-%   \item hide WLCG resources behind a single interface
-%   \item queue-based interface to the application layer
-%   \item pilot-based
-%   \item job-based
-%   \item global view of the properties and states of all the jobs
-%   \item global view of the available resources as a single worldwide
-%      processing resource.
-%   \item centralize data management
-%   \item X509 AAA
-%   \item Multitenancy
-% \end{itemize}
+\label{ssec:panda_design}
 
@@
 
-PanDA's application model primarily assumes tasks, as single entities or
-organized into workloads or workflows. Since 2005, a certain amount of
-parallelism has been progressively introduced assuming both single core and
-multi-threaded tasks~\cite{multithreaded_tasks} but, so far, no MPI tasks have
-been considered for production. Tasks are assumed to have at least two types of
-data dependences: one or more input file and one or more data sources. Task are
-also supposed to be relatively self-contained, capable of setting up their own
-execution environment or having a minimal set of common dependences. Different
-types of task executable are assumed, both in terms of user-defined batch
-scripts or application frameworks for group of users.
+PanDA's application model primarily assumes single tasks, workloads and
+workflows. Tasks represent a set of operations performed on a set of events,
+stored across input files. Tasks can be decomposed into jobs, where each job
+represents the task's set of operations and a partition of the task's events.
+Since 2005, a certain amount of parallelism has been progressively introduced
+for jobs~\cite{multithreaded_jobs} but, so far, no MPI jobs have been considered
+for production. Jobs are supposed to be relatively self-contained, capable of
+setting up their own execution environment or having a minimal set of common
+dependences. Different types of job payload (i.e., executable) are assumed, both
+in terms of user-defined batch scripts or application frameworks for group of
+users.
 
@@
 types of HEP users: individual researchers and groups executing production-grade
-workflows. Users are considered free to submit tasks and workflows at any point
-in time, directly or via dedicated application frameworks. Consistently, PanDA's
-security model is based on separation between authentication, authorization and
-accounting for both single users and group of users. Both authentication and
-authorization are based on digital certificates and the X.509 standard.
+workflows. Users are considered free to submit tasks and workflows to the PanDA
+WMS at any point in time, directly or via dedicated application frameworks.
+Consistently, PanDA's security model is based on separation between
+authentication, authorization and accounting for both single users and group of
+users. Both authentication and authorization are based on digital certificates
+and the X.509 standard and on the virtual organization abstraction.
 
 Currently, PanDA's execution model is based on five main abstractions: task,
-job, queue, pilot, and event. Tasks are assumed to be partitioned into one or
-more job, possibly of variable sizes. Jobs have both attributes and states and
-are queued into a global queue for execution. Prioritization and binding of jobs
-are assumed to depend on the attributes of each job. Pilot abstracts resource
-capabilities and jobs are bound to pilots and executed by consuming their
-resources. Consistently, pilots are considered to be multitenant as jobs
-belonging to multiple workflows and/or users could be bound and executed on
-them.
+job, queue, pilot, and event. Both tasks and jobs are assumed to have attributes
+and states and to be queued into a global queue for execution. Prioritization
+and binding of jobs are assumed to depend on the attributes of each job. Pilot
+is used to indicate the abstraction of resource capabilities. Jobs are thought
+to be bound to pilots and the execution of their payload is performed on the
+site where the pilots has been instantiated. Consistently, pilots are considered
+to be multitenant as jobs belonging to multiple workflows and/or users can be
+bound and executed on them.
 
@@
 collision captured in the detector and one or more events can be packaged into
-files or data source with other abstractions. As with jobs, data have both
-attributes and states and some of the attributes are linked to those of jobs so
-to create an association between the two otherwise separated entities. Data are
-assumed to be stored centrally, replicated when necessary and transferred over
-the network. This assumptions applies to both input and output data.
+files or data source with other container abstractions. As with jobs, data have
+both attributes and states, and some of the attributes are shared between events
+and jobs so to create associations. Data are assumed to be stored centrally,
+replicated when necessary and transferred over the network. This assumptions
+applies to both input and output data.
 
@@
 
+% -----------------------------------------------------------------------------
 \subsection{Architecture}
+\label{ssec:panda_arch}
 
@@
 of them built from off-the-shelf and Open Source components. At a high level,
-PanDA's architecture has six subsystems: PanDA Server, AutoPyFactory, PanDA
+PanDA's architecture has six main subsystems: PanDA Server, AutoPyFactory, PanDA
 Pilot, PanDA Monitoring, JEDI, and Event Service. Subsystems communicate via
-dedicated interfaces, implemented as API or http messaging. Databases are used
-to store eventful entities like tasks, jobs and events, and to store information
-about resources, logs, and accounting. Each subsystems is implemented as one or
-more modules as shown in Figure~\ref{fig:architecture}. In the following we
-briefly describe the overall architecture of PanDA and its components. The
-description provides all the details required to understand how HPC resources
-have been integrated within PanDA. For a full description of PanDA architecture
-see~\cite{wiki, papers}.
-
-Capabilities:
-\begin{itemize}
-  \item interoperable across job submission systems
-  \item asynchronous distribution of input/out data organized as file collections associated to (set of) jobs.
-  \item policy-aware scheduling
-  \item Supporting multiple types of executable code.
-\end{itemize}
-
-% Components:
-% \begin{itemize}
-%   \item monitoring system
-%   \item information system (static/dynamic information)
-%   \item data management system
-%   \item
-% \end{itemize}
-
-% \begin{itemize}
-% \item PanDA Server
-% \item Pilot
-% \item Factory
-% \item PanDA monitoring
-% \item JEDI ??
-% \item What else?
-% \end{itemize}
+dedicated API or http messaging. Databases are used to store eventful entities
+like tasks, jobs and events, and to store information about resources, logs, and
+accounting. Each subsystems is implemented by one or more modules as shown in
+Figure~\ref{fig:architecture}. In the following, we briefly describe the overall
+architecture of PanDA and its components. The description provides all the
+details required to understand how HPC resources have been integrated within
+PanDA (\ref{sec:integration}). For a more detailed description of PanDA
+architecture see~\cite{wiki, papers}.
 
@@
 that will be referenced when describing PanDA's subsystems.}Schematic view of
-the PanDA system\label{fig:architecture}, as deployed for production in 2017.
-Black boxes indicate PanDA's subsystems; Orange boxes components of a subsystem;
-purple boxes databases; green arrows http messaging interfaces; red arrows API;
-purple arrows database communications.}
+the PanDA system, as deployed for production in 2017. Black boxes indicate
+PanDA's subsystems; Orange boxes components of a subsystem; purple boxes
+databases; green arrows http messaging interfaces; red arrows API; purple arrows
+database communications.}
 % Originally PanDA was designed for grid infrastructure... In this
@@
 % presented for workload X on Titan......
+\label{fig:architecture}
 \end{figure}
@@
 PanDA and it operates as a web service. It runs on Apache web server,
-interacting with the back-end database running on separate servers. It uses the
-Apache worker model, with many independent  processes handling client requests
-in parallel. Since all state is maintained in the central database, the PanDA
-Server application instance itself is stateless. Currently production version of
-PanDA utilizes Oracle database backend, but PanDA can also work with MySQL
-family of databases.
+interacting with the back-end database running on separate servers. It
+implements the Apache worker model, with many independent processes handling
+client requests in parallel. Since all state is maintained in the central
+database, the PanDA Server application instance itself is stateless. Currently
+production version of PanDA utilizes Oracle database backend, but PanDA can also
+work with MySQL family of databases.
 
@@
 global queue and offers functionalities to: (i) store jobs into the global
-queue; (ii) provide information about the state of each stored job; and (iii)
-update the state of each job upon communication with the other three components
-of PanDA Server. The Brokerage component enables (i) matching job attributes
-with site and pilot attributes; (ii) managing the dispatch of input data to
-processing sites; and (iii) deriving PanDA's data pre-placement requirement.
+queue; (ii) provide information about and update the state of each stored job.
+The Brokerage component enables (i) matching job attributes with site and pilot
+attributes; (ii) managing the dispatch of input data to processing sites; and
+(iii) deriving PanDA's data pre-placement requirement.
 % The job/pilot matching algorithm assigns a weight to each candidate pilot and
@@
 with highest priority that can be executed on the resource on which the
-requesting pilot has been instantiated; (iii) creating a wrapper for the
+requesting pilots have been instantiated; (iii) creating a wrapper for the
 selected job, tailored to the resource of the requesting pilot; and (iv)
@@
 
-\paragraph{\textbf{PanDA Pilot}} this subsystem is execution environment for
-PanDA's jobs. Implemented in Python, a PanDA Pilot is launched by a wrapper
-submitted to a target resource. Wrappers are tailored to the specific interfaces
-and middleware exposed by each site so that PanDA Pilot remains
-site-independent. PanDA Pilot implements four main functionalities: (i)
-preparing the execution environment for PanDA's jobs, including recovering and
-cleaning up previously failed jobs on that site and collecting information about
-the compute, data, and network capabilities of the site; (ii) requesting a job
-for execution to the Job Dispatcher component of PanDA Server; (iii) initiating
-and monitoring the execution of pulled jobs, including setting up the job,
-transferring its input when required, executing the payload, transferring
-output, and checking log files and local space; and (iv) clean up when the
-payload has finished to execute.
+\paragraph{\textbf{PanDA Pilot}} this subsystem implements the execution
+environment for PanDA's jobs. Implemented in Python, a PanDA Pilot is launched
+by a wrapper submitted to a site. Wrappers are tailored to the
+specific interfaces and middleware exposed by each site so that PanDA Pilot
+remains site-independent. PanDA Pilot implements four main functionalities: (i)
+preparing the execution environment for PanDA's jobs,
+% including  recovering and cleaning up previously failed jobs on that site and
+collecting information about the compute, data, and network capabilities of the
+site; (ii) requesting a job for execution to the Job Dispatcher component of
+PanDA Server; (iii) initiating and monitoring the execution of pulled jobs,
+including setting up the job, transferring its input when required, executing
+the payload, transferring output, and checking log files and local space; and
+(iv) clean up when the payload has finished to execute.
 
@@
 
-PanDA Pilot implements the pilot paradigm~\cite{pilot_review} and, as such, it
-decouples resource acquisition from job execution via multi-stage scheduling.
-This has several benefits, including increasing the number of completed jobs per
-unit of resource by lowering the time taken between finishing the execution of a
-job an starting a the execution of a new job; enabling both concurrent and
-sequential job execution depending on the number of cores available to the pilot
-and the time for which the remain available; and exposing a unified and
-consistent interface for the scheduling of jobs, independent from the type of
-resource on which the pilot has been instantiated and the middleware installed
-on the site. The specifics implementation of PanDA Pilot enables also better
-reliability by real-time monitoring of the job execution process, and a certain
-degree of fault-tolerance by recovering information and, possibly, output files
-from jobs that failed, even before instantiating the PanDA Pilot.
+PanDA Pilot implements the pilot paradigm~\cite{pilot_review}, decoupling
+resource acquisition from job execution via multi-stage scheduling. This has
+several benefits, including increasing throughput by lowering the time taken to
+queue and schedule jobs; enabling both concurrent and sequential job execution
+depending on the number of cores available to the pilot and the time for which
+it remains available; and exposing a unified and consistent interface for the
+scheduling of jobs, independent from the middleware of the site on which the
+pilot has been instantiated.
+
+% The specific implementation of PanDA Pilot promotes reliability by
+% monitoring the execution of the job's payload, and a certain degree of
+% fault-tolerance by recovering information and, possibly, output files from
+% jobs that failed, even before instantiating the PanDA Pilot.
 
@@
 It should be noted that upon bootstrapping, a PanDA Pilot download major parts
-of its runtime code a central Subversion repository via HTTP. The repository
-works with an Apache web server, configured with a memory-based web proxy
-(Squid). The purpose of the cache is to reduce the request load on the
+of its runtime code via HTTP from a central Subversion repository. The
+repository works with an Apache web server, configured with a memory-based web
+proxy (Squid). The purpose of the cache is to reduce the request load on the
 back-end Subversion server. That improves performance \mtnote{Do we have
 data/numbers?} since only occasional queries trigger a full lookup on the
-back-end Subversion system, and most external queries are pulled from memory
-on the front-end web server. The benefits of this system include a
-high-performance code download service combined with code updates still being
-immediately available as soon as they are committed to source code control.
+back-end Subversion system, and most external queries are pulled from memory on
+the front-end web server. The benefits of this system include a high-performance
+code download service, combined with code updates still being immediately
+available as soon as they are committed to source code control.
 
@@
 pilots locally and to both grid and cloud sites. Implemented to be executed as a
-single demonized process, AutoPyFactory enable the concurrent spawning of
+single demonized process, AutoPyFactory enables the concurrent spawning of
 multiple job submission processes called APFQueue. Each APFQueue can communicate
@@
 consistent and adequate pressure on their batch queues. Submission is performed
-via the Condor-G batch system~\cite{}, enabling submission to multiple type of
-sites. AutoPyFactory exposes a dedicated monitoring service, feeding information
-to the PanDA Monitoring subsystem. A proxy manager is also run, enabling
-certificate-based authentication and authorization.
+via the Condor-G batch system~\cite{condorg}, enabling submission to multiple
+type of sites. AutoPyFactory exposes a dedicated monitoring service, feeding
+information to the PanDA Monitoring subsystem. A proxy manager is also run,
+enabling certificate-based authentication and authorization.
 
@@
 about the status of current and past jobs, data movement, and pilot factory.
-PanDA Monitoring allows users access to all job log files on job by job basis,
-greatly simplifying code debugging and failure analysis in a distributed
-computing environment.
+PanDA Monitoring allows users access to all log files of each job, greatly
+simplifying code debugging and failure analysis in a distributed computing
+environment.
 
-\paragraph{JEDI} \mtnote{Should we include it?}
+\paragraph{\textbf{JEDI}} this subsystem implements a workload and workflow
+management for tasks instead of jobs, as defined by HEP users. Implemented as a
+standalone service that interacts with Panda Server via its REST interface, JEDI
+offers two main functionalities: partitioning tasks to jobs, and binding jobs to
+resources via dedicated queues. JEDI partition tasks to jobs by matching tasks
+requirements to resource capabilities. This matching is based on tasks
+attributes, and static and dynamic information about the resources of each
+available site. Dynamic information is acquired via a small number of `scout'
+jobs that collect information about current resources and their performance.
+Functionally, JEDI defines a set of queues associated to groups of resources
+with specific capabilities. Tasks are assigned to groups of these queues, and
+the jobs in which each task is partitioned are assigned to each queue, depending
+on their requirements. JEDI replaces the job brokerage functionalities of the
+Brokerage module of the PanDA Server but this module still handles communication
+between the Task Buffer and the Data Service modules.
+%, using it to enact the execution of the jobs defined by JEDI
+% on the resources and sites decided by JEDI.
 
-\paragraph{Event Service} \mtnote{Should we include it?}
+% -----------------------------------------------------------------------------
+\subsection{Management and Execution}
+\label{ssec:panda_exec}
 
-\subsection{Execution Process}
+% The tasks of workloads and workflows submitted to PanDA are partitioned
+% to one or more jobs. Each job is then executed on one of the available
+% PanDA Pilots, chosen among those that the PanDA Server manages across multiple
+% sites. Tasks processes a certain amount of events that are stored into files,
+% and when a task is partitioned to multiple jobs, subsets of events need to be
+% assigned to each job and made available to the pilot on which that job will be
+% executed. Finally, the output of each job need to be retrieved and aggregated
+% into the output of the original task (Fig.~\ref{fig:architecture}).
 
-\mtnote{Probably move this to the workflow}
+The relation between tasks and jobs can be one-to-one or one-to-many, and the
+conversion between the two can by static or dynamic. During the LHC Run 1, PanDA
+required users or the application layer to perform a static conversion between
+tasks and jobs: tasks were described as a set of jobs with a fixed number of
+events and then submitted to the PanDA Server (Fig.~\ref{fig:architecture}:A1).
+
+This approach introduced inefficiency both with usability and resource
+utilization. Ideally, users should not have to reason in terms of jobs: Users
+conceive analyses in terms of one or more, possibly related tasks; the `job'
+abstraction is required exclusively by the execution middleware, i.e. PanDA.
+Further, a static partitioning of tasks into jobs does not take into account the
+heterogeneity and dynamicity of the resources of the pilots on which each job
+will be executed.
+
+Another problem of static job sizing is that PanDA instantiates pilots on sites
+with different type of resources and different models of availability of those
+resources. An optimal sizing of each job should take into account these
+properties. For example, sites may offer cores with different speed, networking
+with different amount of bandwidth, and resources could be guaranteed to be
+available for a defined amount of time or could disappear at any point in time
+as with opportunistic models of resource provision.
+
+JEDI was deployed for the LHC Run 2 to address these inefficiencies with a
+machine-driven conversion of tasks into jobs. Users or the application layer (i)
+submits tasks descriptions to JEDI; (ii) JEDI partitions tasks to jobs of
+different size, depending on both static and dynamic information about available
+resources; (iii) jobs are bound to sites with resources that best match jobs'
+requirements; and (iv) jobs are submitted to the PanDA Server for execution
+(Fig.~\ref{fig:architecture}:A2-B).
+
+Once submitted to the PanDA Server, jobs are stored by the Task Buffer component
+into a global queue implemented as a relational database
+(Fig.~\ref{fig:architecture}:C). When jobs are submitted directly to the PanDA
+Server, the Brokerage component is used to bind jobs to available sites,
+depending on static information about the resources available for each site.
+Jobs submitted by JEDI are instead already bound to resources so no further
+brokerage is needed. Once jobs are bound to sites, the Brokerage module
+communicates to the Data Service module what data sets need to be made available
+on what site. The Data Service communicates these requirements to the ATLAS DDM
+that, when needed, takes care of aggregating files into datasets and containers,
+replicating them on the required sites.
+
+The AutoPyFactory subsystem communicates with the Task Buffer component of the
+PanDA Server, acquiring information about jobs that have been bound and are
+ready for execution. On the base of this information, AutoPyFactory defines a
+set of suitable PanDA Pilots, queues them into APFQueues, and concurrently
+submits them to multiple sites.
+
+% AutoPyFactory submits pilots only when similar pilots are not yet available
+% and maintains a predefined amount of pressure on the queue of each site. In
+% this way, resource availability and utilizations is optimized.
+
+When PanDA Pilots become available on sites, they
+% bootstrap, download their code, collect information about the execution
+% environment and
+call the Job Dispatcher module of the PanDA Server, requesting for a job to
+execute \mtnote{Can a pilot request multiple jobs to execute concurrently?}. The
+Job Dispatcher communicates with the Task Buffer, requesting for a job that is
+bound to the site of that pilot and ready to be executed. Task Buffer checks the
+global queue (i.e., the PanDA database) and if such a job is available, it
+passes the job handler to the Job Dispatcher. The Job Dispatcher dispatches the
+job to the PanDA Pilot.
+
+Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
+subprocess for the execution of the job's payload. The job is setup, input data
+are transferred from the designated staging in location, and the job's payload
+is executed. Once completed, output is transferred to the staging out location.
+% During the job's payload execution, the monitoring process of PanDA Pilot
+% checks the status of the job execution and its environment.
+
+The Data Service module of the PanDA Server tracks and collects the output
+generated by each job, updating jobs' attributes via the Task Buffer module.
+When the output of all the jobs of a task are retrieved, they are aggregated and
+made available to the user either via the PanDA Server of the JEDI service,
+depending on how tasks or jobs were initially submitted.
+
 \mtnote{Think whether a comparison with other WMS based on the review we did
 makes sense in this context.}
-
-\mtnote{Summary: JEDI retrieves task descriptions and source files, converting
-them into job descriptions and events sets. Jobs are submitted to PanDA Server
-while events to the Event Service. The PanDA Server schedules jobs to pilots
-that have been submitted to resources via the PanDA Factory. Once executed, jobs
-pulls events from the Event Service as input and push output back to JEDI.}

### 2017-03-04 workflow.tex
JEDI and double execution process

@@
-The distinction between an abstract and concrete workflow instances highlights
-an important element of the execution process of a workflow. Abstract workflows
-can be interpreted as a formal description of the requirements for the workflow
-execution, while concrete workflows can be interpreted as a formal description
-of the workflow's execution process. The process of resource selection matches
-the capabilities of resources to the requirements of a workflow, making possible
-the transition from execution requirements to execution process.
+% The distinction between an abstract and concrete workflow instances highlights
+% an important element of the execution process of a workflow. Abstract
+% workflows can be interpreted as a formal description of the requirements for
+% the workflow execution, while concrete workflows can be interpreted as a
+% formal description of the workflow's execution process. The process of
+% resource selection matches the capabilities of resources to the requirements
+% of a workflow, making possible the transition from execution requirements to
+% execution process.
+
+\begin{itemize}
+    \item Is the Montecarlo Workflow a production workflow?
+\end{itemize}
+
+The design and implementation of PanDA enables the execution of diverse HEP workflows.
+Here we describe the execution process of the Monte Carlo workflow~\cite{montecarlo_workflow}, one of the largest workflows
+executed by the ATLAS experiment
+
+\subsection{NOTES - Please ignore}
 
@@
 selection, acquisition, and use. The set of capabilities required by a workload
-manager mostly depends on the heterogeneity and dynamicity of both the workflow
+manager mostly depends on the heterogeneity and dynamicity of both the workload
 and the resources. Heterogeneity is a measure of the diversity of the workflow
@@
 resources~\cite{panda_grid}. Progressively, PanDA has evolved to enable the
-management of more heterogeneous tasks and resources. In this paper, we describe
-how the design and architecture of PanDA has enabled executions of
-multi-threaded tasks on Titan, currently the largest high performance computing
-(HPC) resource available in the USA to scientific research.
+management of more heterogeneous tasks and resources.
+
+% In this paper, we describe how the design and architecture of PanDA has
+% enabled executions of multi-threaded tasks on Titan, currently the largest
+% high performance computing (HPC) resource available in the USA to scientific
+% research.
 

### 2017-03-05 PandaEvolution.tex
First pass IV (former IV and V) + restructuring

--- 2017-01-31 PandaEvolution.tex
@@
-\subsection{Brief Introduction to Titan}
+The next LHC data taking run (Run 3) will require more resources than the
+Worldwide LHC Computing Grid (WLCG) can possibly provide. Currently, PanDA WMS
+uses more than 100,000 cores at over 100 Grid sites, with a peak performance of
+0.3 petaFLOPS. This capacity will be sufficient for the planned analysis and
+data processing, but it will be insufficient for Monte Carlo (MC) production and
+any extra activities. To alleviate these challenges, ATLAS is engaged in a
+program to expand the current computing model to include additional resources
+such as the opportunistic use of supercomputers as well as commercial and
+academic clouds.
 
-While PanDa  currently uses  more than 100,000  cores at well over 100 Grid
-sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will
-require more resources  than Grid computing can possibly provide. The Worldwide
-LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned
-analysis and data processing, but it will be insufficient  for Monte Carlo (MC)
-production and any extra activities. Additional computing and storage  resources
-are therefore required.  To alleviate  these challenges, ATLAS is engaged  in an
-ambitious  program to expand  the current computing model to include  additional
-resources such as the opportunistic   use of supercomputers   as well as
-commercial and academic clouds. In turn this activity drives evolution of the
-PanDA WMS\@.
+% \subsection{Use of Supercomputers with PanDA}
 
-\subsection{Use of Supercomputers with PanDA}
+Modern supercomputers have been designed to mainly support parallel computation
+requiring runtime communication. Job execution is parallelized across multiple
+cores, each core calculating a small part of the problem and communicating with
+other cores via message passing interface (MPI). Accordingly, supercomputers
+have large number of worker nodes, connected through a high-speed, low-latency
+network. Each worker node has multicore CPUs, usually augmented with massively
+parallel Graphics Processing Units (GPUs) or other types of coprocessors.
 
-The idea is to extend the ATLAS Computing Model beyond the Grid into the domains
-of supercomputers. Modern High Performance Computing  (HPC) involves  a very
-large number of worker nodes  connected  through a  high-speed   network. The
-worker nodes can have  multicore CPUs that can be augmented with massively
-parallel Graphics Processing Units (GPUs) or other types of coprocessors. On an
-HPC machine, a  typical job is highly parallel and connected,  where each core
-calculates  a small part of the problem and coordinates the activity via message
-passing interface (MPI). While this is different from the HEP computing
-paradigm, where jobs are independent, it still shares common  features  such as
-the use of parallelization. It is not a requirement  that HPC machines are able
-to run any possible task, nor is it relevant how many kinds of job types that
-can be run. What matters is the total number of cycles that can be offloaded
-from the Grid. Standard  ATLAS  workflow  can not be easily ported to
-supercomputers   due  to   several   complications such  as specialized worker
-node  setups, no  outbound network connections,  limited memory per node, custom
-operating systems,  etc. A reorganization  of the standard  workflow is
-therefore needed.
+PanDA WMS has been designed to support distributed computing. Workload or
+workflow executions involves concurrent and/or sequential runs of possibly large
+amount of jobs, requiring no or minimal parallelization and no communication at
+runtime. Accordingly, computing infrastructure like WLCG are designed to
+aggregate large amount of computing resources across sites, possibly dislocated
+across multiple geographical allocation. While each site may deploy runtime
+message-passing capabilities, usually these are not used to perform distributed
+computations.
 
-The most suitable task for HPCs in HEP is event generation. Event generators are
-mostly computational,  stand-alone code, with few input requirements.  There is
-no need  to stage-in much data. The event generation  in ATLAS corresponds  to
-10--15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs
-is detector  simulation. Simulation tasks tend to be tied to the framework of a
-particular simulation and often require  database access for geometry and
-relevant run conditions.  However,  there is a  lot to be gained  since
-simulation  tasks correspond to 2/3 of the Grid capacity.
+There are at least two approaches to enable PanDA WMS to execute ATLAS jobs on
+supercomputers: using the subset of resources and capabilities shared by both
+supercomputers and grid; reconciling the two paradigms of computation by means
+prototype a working solution that enables computation and better understanding
+of the problem space. The latter approach is principled and better suited for a
+production-grade solution developed on the base of the experience accumulated
+with the prototype and capable of supporting general-purpose workloads and
+workflows on both supercomputers and grid infrastructures.
 
+This section illustrates the design, architecture and execution process of the
+prototype developed by the PanDA team to support execution of some of the ATLAS
+workload on the Titan supercomputer. After a critical analysis of the results
+obtained and the lessons learned, the following section introduces the design
+and first experimental characterization of a next generation executor. This
+components is designed to abstract resources and capabilities enabling the
+concurrent execution of both parallel and distributed computing.
+
+% -----------------------------------------------------------------------------
 \subsection{Interfacing PanDA with Titan}
+\label{ssec:panda_titan}
 
 The Titan supercomputer, current number two (number one from November 2012 until
-June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing
-Facility (OLCF) in Oak Ridge National Laboratory, US\@. Titan, was the first
-large-scale system to use  a hybrid architecture that utilizes worker nodes with
-both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
+June 2013) on the Top 500 list~\cite{top500} is located at the Oak Ridge
+Leadership Computing Facility (OLCF) in Oak Ridge National Laboratory, US\@.
+Titan is a Cray XK7 system with 18,688 compute nodes and a total of 299,008 CPU
+cores. Each compute node has an AMD Opteron CPU, a Nvidia Tesla GPU, 32 GB of
+RAM and no local storage, though a 16 GB RAM disk can be set up. Compute nodes
+use Cray’s Gemini interconnect for inter-node MPI messaging but have no network
+connection to the outside world. Titan is served by the Spider
+II~\cite{spider2}, a Lustre filesystem with 32 PB of disk storage, and by a 29
+PB HPSS tape storage system. Titan’s worker nodes run Compute Node Linux, a run
+time environment based on SUSE Linux Enterprise Server.
 
-Integration with Titan is the current focus for PanDA developers. The project
-aims to integrate Titan with the PanDA system using an updated PanDA Pilot that
-runs on the front-end node and submits ATLAS payloads to the worker nodes using
-the local batch system (PBS) via the SAGA (Simple API for Grid Applications)
-interface~\cite{SAGA}. This solves several complications of  running on HPC
-worker nodes,  including the lack of connectivity to outside world. The pilot
-can communicate with the PanDA server from the front-end machine. The
-interactive front-end  machines and the worker nodes  use  a shared  file system
-which makes  it  easy  for the pilot to stage-in  any input files that are
-required by the payload and stage-out the produced output files after the
-payload has finished at the end of the job. ATLAS Tier 1 computing  center at
-Brookhaven National Lab is currently  used for data stage out from Titan but in
-principle  that can be any Grid site.
+% Titan, was the first large-scale system to use a hybrid architecture that
+% utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla
+% K20 GPU accelerators.
+% This hybrid design provides improved energy efficiency, as well as an order of
+% magnitude in computational capacity over its predecessor.
 
-\paragraph*{HPC Backfill} HPC facilities are geared towards  large scale jobs by
-design. Time allocation on an HPC is competitive and large projects are often
-preferred. About 10\% of capacity on a  typical HPC machine  is unused  due to
-mismatches between job sizes and available  resources. The worker nodes sit idle
-because there are not enough of them or they do not have enough time to handle a
-large scale computing job. On a machine of the scale of Titan, these 10\%
-correspond to estimated 300M core hours per year. A system that can occupy those
-temporarily free nodes with smaller scale tasks would be very valuable. This
-offers a great possibility for PanDA to harvest the opportunistic resources on
-Titan and use a similar mechanism on other HPCs. Functionality has been added to
-the PanDA Pilot to collect information about available unused worker nodes on
-Titan in real time. This allows PanDA to define precisely size and duration of
-jobs submitted to Titan according to available free resources. First test of the
-system were quite successful and show great promise in increasing the resource
-utilization on Titan.
+Titan offers to its users login nodes and data transfer nodes (DTN). Users have
+to log into one of these nodes via ssh to submit jobs to Titan's PBS scheduler.
+Titan's authentication and authorization model is based on two-factor
+authentication with a RSA SecurID key, generated every 30 seconds. Login nodes
+and DTN are the only nodes with out/inbound wide area network connectivity. Work
+nodes have only local network access. Data staging is supported via the DTNs,
+from the wide area network and the OLCF high performance storage system (HPSS).
+DTNs are shared across all Titan's users that can schedule jobs to a dedicates
+PBS batch system to automate data staging. Fair-share policies are in place both
+for the batch system that for the use of DTN resources.
+
+Titan's architecture, configuration and policies poses several challenges to the
+integration with PanDA. PanDA Pilot requires to contact the Job Dispatcher
+module of the PanDA Server to pull jobs to execute. Titan's work nodes do not
+offer outbound network connectivity making the current deployment model of PanDA
+pilots unfeasible. Titan does not support PanDA's security model based on
+certificates and virtual organizations, making the PanDA's approach to identity
+management also unfeasible. While DTNs offer wide area network data transfer, an
+integration with ATLAS DDM is beyond the functional and administrative scope of
+the current prototyping phase. Finally, the specific characteristics of the
+execution environment, especially the dependences on a Lustre filesystem and on
+modules tailored to Compute Node Linux, require reengineering of ATLAS
+application frameworks.
+
+Currently, no HEP application can benefit from Titan's GPUs but some
+computationally-intensive, non memory-intensive tasks can be offloaded from the
+grid to Titan's large amount of cores. Further, when HEP tasks can be
+partitioned to independent jobs, Titan work nodes can be used to execute up to
+16 concurrent jobs' payload. Given these constraints and challenges, the type of
+task most suitable for execution on Titan is HEP event generation. Event
+generators are mostly computational and stand-alone code, requiring less than
+1GB of RAM at runtime and with small input data requirements. The event
+generation in ATLAS accounts for 10--15\% of all jobs on the Grid, corresponding
+to 2/3 of the current grid capacity. A more difficult workload to adapt for
+Titan would be that for detector simulation. Simulation tasks tend to be tied to
+the framework of a particular simulation, often requiring database access for
+geometry and relevant run conditions.
+
+\mtnote{Describe Monte Carlo workflow and Geant4 here.}
+
+% While this is different from the HEP computing paradigm, where jobs are
+% independent, it still shares common features such as the use of
+% parallelization. It is not a requirement that HPC machines are able to run
+% any possible task, nor is it relevant how many kinds of job types that can be
+% run. What matters is the total number of cycles that can be offloaded from
+% the Grid.
+
+% Standard ATLAS workflow can not be easily ported to supercomputers due to
+% several complications such as specialized worker node setups, no outbound
+% network connections, limited memory per node, custom operating systems, etc. A
+% reorganization of the standard workflow is therefore needed.
+
+% -----------------------------------------------------------------------------
+\subsection{PanDA Broker on Titan}
+\label{ssec:panda_titan}
+
+The most relevant design constraint for integrating PanDA WMS and Titan is the
+lack of connectivity from the work nodes. This makes impossible to port PanDA
+Pilot while maintaining the defining feature of the pilot abstraction:
+multi-stage scheduling to decouple resource acquisition from workload scheduling
+and execution. As direct communication between PanDA Pilot and PanDA Server is
+required, PanDA Pilots cannot be scheduled on Titan and jobs cannot be pulled by
+a pilot from the work nodes.
+
+Usually, this design limitation would make the integration of PanDA and Titan
+unfeasible. Pilot-based executions are necessary to avoid waiting in the
+supercomputer's queue to execute every single job of a distributed workload.
+This is particularly true for Titan where, consistently with the intended use
+case of the machine, queuing policies privilege parallel jobs on the base of the
+quantities of work nodes requested. It would still be possible to execute HEP
+event generation tasks but the time spent waiting in the queue would be too
+large compared to the execution time.
+
+About 10\% of Titan's capacity is unused due to mismatches between job sizes and
+available resources~\cite{titan_utilization}. Worker nodes are unused because
+there are not enough nodes to execute large jobs or the required amount of nodes
+would not be available for a long enough time. This makes available around
+30,000 cores for a total of approximately 270M core hours per
+year~\cite{titan_free}, roughly 30\% of the overall capacity of WLCG. A system
+that can occupy those temporarily free nodes with smaller scale tasks would be
+very valuable. For this reason, Titan offers backfill functionalities: users can
+interrogate Titan's Moab scheduler about how many free nodes are currently
+available and for how long. If a user submit a request for that number of nodes,
+queue time will be minimal.
+
+\mtnote{Please feel free to add details about backfill as needed.}
+
+Backfill functionality offers the opportunity not to use the pilot abstraction
+while still avoiding too much queue time overhead. On Titan, PanDA Pilot has
+been redesigned to serve as a job broker. Architected as a stand-alone service
+to be executed on a DTN, this PanDA Broker implements functionalities to: (i)
+interrogate Titan about backfill availability; (ii) pull Geant4 jobs and events;
+(iii) wrap Geant4 payloads into PBS MPI job; (iv) submitting PBS jobs to Titan's
+batch system; and (v) staging input/output files. The new functionalities of
+PanDA Broker involve backfill querying, payload wrapping, and job submission.
+Functionalities for Geant4 job and events pulling, and file staging were mostly
+inherited from PanDA Pilot.
 
@@
   \begin{center}
-\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
-\caption{Schematic view of the PanDA system\label{fig:interface}}
+    \includegraphics[width=\columnwidth]{figures/PanDA_setup_at_OLCF.png}
+    \caption{Schematic view of the PanDA Broker.}
   \end{center}
+\label{fig:panda_broker}
 \end{figure}
+
+Backfill querying is performed via the dedicated Moab command while payload
+wrapping is done via a tailored Python MPI script. This MPI wrapper serves as a
+``container'' for non-MPI jobs, enabling execution of unmodified Grid-centric
+jobs on Titan. Typically, a MPI wrapper is workload-specific as it sets up the
+execution environment for a specific payload. This involves organization of
+worker directories, data management, optional input parameters modification, and
+cleanup on exit. Upon submission, a copy of the MPI wrapper runs on every
+available work node. Each script knows its MPI rank (an index that runs from
+zero to a maximum number of nodes or script copies) as well as the total number
+of ranks in the current submission. When activated on worker nodes, each copy of
+the wrapper script starts the execution of the payload in a subprocess and waits
+until its completion.
+
+MPI wrappers are submitted to Titan's PBS batch system via
+RADICAL-SAGA~\cite{radical-saga}. SAGA (Simple API for Grid Applications)
+provides a unified interface to diverse job schedulers and file transferring
+services. SAGA provides an interoperability layer that lowers the complexity and
+improves the simplicity of using distributed infrastructure. RADICAL-SAGA is a
+Python module, compliant with the OGF GFD.90 SAGA
+specification~\cite{saga-spec}. Behind the API façade, RADICAL-SAGA implements a
+adaptor architecture: each adaptor interface the SAGA API with different
+middleware systems and services, including the batch scheduler of Titan.
+
+PanDA Broker stages capabilities are implemented via a file system shared among
+DTNs and work nods. Input files that are required by the payload are downloaded
+via the ATLAS DDM service and then stored in a directory on the shared
+filesystem. Part of the MPI wrapper setup process is to make the location of
+these files known to the payload. Payloads' output files are located on the
+shared filesystem by the broker and transferred from Titan to the ATLAS Tier 1
+computing center at Brookhaven National Lab (BNL). In principle any Grid site
+could be used.
+
+The execution process of Geant4 on Titan by PanDA has X steps: (i) PanDA Broker
+queries Titan's Moab scheduler about currently unused nodes and checks whether
+the time and size of the available backfill resources are suitable to execute
+jobs of Geant4 tasks; (ii) The PanDA Broker transmits this information to the
+Job Dispatcher module of the PanDA server, and in response gets a list of Geant4
+jobs bound to Titan; (iii) based on the jobs attributes, the broker requests to
+the Data Service module of the PanDA Server to transfer the necessary input data
+via the ATLAS DDM service; (iv) once all the data is transferred, the PanDA
+Broker calculates the number of cores and duration of jobs that can be submitted
+to Titan on the base of the information collected about backfill resource
+availability; (v) the broker creates one or more PBS MPI scripts to wrap enough
+Geant4 payload to fit the calculated number and duration of cores; (vi) the
+broker submits jobs to Titan's PBS batch system via RADICAL-SAGA; (vii) the PBS
+jobs waits a minimal amount of time into the queue until is scheduled for
+execution; (viii) upon execution on the worker nodes, the MPI script sets up the
+execution environment required by the wrapped Geant4 payload, including
+assembling the AthenaMP framework in ramdisk; (ix) the payload is then executed
+into dedicated processes, 16 for each work node, 1 for each core; (x) the PanDA
+Broker locates on the shared filesystem the output files produced by the payload
+and transfer them to BNL ATLAS Tier 1 computing centre.
+
+% from the Job Dispatcher module of PanDA Server (as done by a PanDA Pilot);
+% (iii) downloading events via PanDA Server's Data Service from ATLAS DDM to
+% Titan's DTN; (iv) wrap the Geant4 payload into a PBS job; (v) submit jobs to
+% Titan's PBS batch scheduler;
+
+% Integration with Titan is the current focus for PanDA developers.
+
+% The project aims to integrate Titan with the PanDA system using an updated
+% PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the
+% worker nodes using the local batch system (PBS) via the SAGA (Simple API for
+% Grid Applications) interface~\cite{SAGA}. This solves several complications
+% of running on HPC worker nodes, including the lack of connectivity to outside
+% world. The pilot can communicate with the PanDA server from the front-end
+% machine.
+
+% \paragraph*{HPC Backfill} HPC facilities are geared towards large scale jobs
+% by design. Time allocation on an HPC is competitive and large projects are
+% often preferred. About 10\% of capacity on a typical HPC machine is unused
+% due to mismatches between job sizes and available resources. The worker nodes
+% sit idle because there are not enough of them or they do not have enough time
+% to handle a large scale computing job. On a machine of the scale of Titan,
+% these 10\% correspond to estimated 300M core hours per year. A system that
+% can occupy those temporarily free nodes with smaller scale tasks would be
+% very valuable.
+
+% This offers a great possibility for PanDA to harvest the opportunistic
+% resources on Titan and use a similar mechanism on other HPCs. Functionality
+% has been added to the PanDA Pilot to collect information about available
+% unused worker nodes on Titan in real time.
+
+% First test of the system were quite successful and show great promise in
+% increasing the resource utilization on Titan.

### 2017-03-05 overall.tex
First pass IV (former IV and V) + restructuring

@@
+% -----------------------------------------------------------------------------
+\subsection{Geant4 Tasks Performance Analysis on Titan}
+\label{ssec:panda_titan}
+
+Currently, 20 instances of PanDA Broker are concurrently deployed on a total of
+4 DTNs, 5 instances for each DTN. Each broker controls from 15 to 300 ATLAS
+simulation jobs per submission, for a maximum total of 96,000 cores. Since
+November 2015, the brokers have operated only in backfill mode, without a
+defined time allocation, and running at the lowest priority on Titan.
+Figure~\ref{core-hours-utilization} shows Titan core hours used by ATLAS from
+September 2015 to March 2017. During that period, ATLAS consumed about XX
+million Titan core hours. About XX million detector simulation jobs were
+completed, and more than 132 million events processed. In February 2017, we
+exceeded X million core hours per month and by March 2017 we ran at more than
+X million core hours per month.
+
+Figure~\ref{backfill-utilization} shows backfill utilization efficiency, defined
+as a fraction of Titan’s total free resources that were utilized by ATLAS, for
+the period from September 2015 to March 2017. Due to continuos improvements in
+the system, the efficiency grew from less than 5\% to more than XX\%, reaching
+XX\% in February 2017. At the peak, this corresponds to more than X\%
+improvement in Titan’s overall utilization.
+
+\mtnote{Add new diagram as discussed in F2F meeting at Rutgers.}
+
+% -----------------------------------------------------------------------------
 \subsection{PanDA Shared Library I/O Performance Impact at OLCF}

### 2017-03-05 titan.tex
First pass IV (former IV and V) + restructuring

@@
 In 2004, the Oak Ridge Leadership Computing Facility (OLCF) was established at
-Oak Ridge National Laboratory (ORNL).The  OLCF's is mission to accelerate the
-scientific discovery and engineering progress by providing outstanding
-computing and data management resources to high-priority research and
-development projects.
+Oak Ridge National Laboratory (ORNL). The OLCF's mission is to accelerate
+scientific discovery and engineering progress by providing outstanding computing
+and data management resources to high-priority research and development
+projects. Among other resources, OLCF manages Titan, a leadership-class and the
+fastest open-science supercomputer in the US. Titan enables scientists to
+evaluate and assess various complex physical phenomena via large-scale
+computational simulations.
 
-Titan, a leadership-class and the fastest open-science supercomputer in the
-U.S., is located at OLCF. Titan enables scientists to evaluate and assess various
-complex physical phenomena via large-scale computational simulations.
-
-Titan is a Cray XK7 system with 18,688 compute nodes~\cite{top500}. Each
-compute node is equipped with an AMD Opteron CPU and a Nvidia Tesla GPU. This
-hybrid design provides improved energy efficiency, as well as an order of
-magnitude in computational capacity over its predecessor. Titan has 710 TB of
-total system memory. This leadership capability compute system is backed by a
-center-wide parallel file system known as Spider II~\cite{spider2}. Spider II
-is one of the world's fastest and largest POSIX complaint parallel file
-systems. It is designed to serve write-heavy I/O workloads generated by Titan
+Titan is a Cray XK7 system with 18,688 compute nodes~\cite{top500}. Each compute
+node is equipped with an AMD Opteron CPU and a Nvidia Tesla GPU. This hybrid
+design provides improved energy efficiency, as well as an order of magnitude in
+computational capacity over its predecessor. Titan has 710 TB of total system
+memory and a center-wide parallel file system known as Spider II~\cite{spider2}.
+Spider II is one of the world's fastest and largest POSIX complaint parallel
+file systems, designed to serve write-heavy I/O workloads generated by Titan
 compute clients and other OLCF resources.

### 2017-03-06 PandaEvolution.tex
Iteration IV

@@
-The next LHC data taking run (Run 3) will require more resources than the
+The next LHC run for taking data (Run 3) will require more resources than the
 Worldwide LHC Computing Grid (WLCG) can possibly provide. Currently, PanDA WMS
@@
 0.3 petaFLOPS. This capacity will be sufficient for the planned analysis and
-data processing, but it will be insufficient for Monte Carlo (MC) production and
-any extra activities. To alleviate these challenges, ATLAS is engaged in a
-program to expand the current computing model to include additional resources
-such as the opportunistic use of supercomputers as well as commercial and
-academic clouds.
+data processing, but it will be insufficient for the Monte Carlo (MC) production
+workflow and any extra activity. To alleviate these challenges, ATLAS is engaged
+in a program to expand the current computing model to include additional
+resources such as the opportunistic use of supercomputers as well as commercial
+and academic clouds.
 
-% \subsection{Use of Supercomputers with PanDA}
+\subsection{Use of Supercomputers with PanDA}
 
 Modern supercomputers have been designed to mainly support parallel computation
-requiring runtime communication. Job execution is parallelized across multiple
-cores, each core calculating a small part of the problem and communicating with
-other cores via message passing interface (MPI). Accordingly, supercomputers
-have large number of worker nodes, connected through a high-speed, low-latency
-network. Each worker node has multicore CPUs, usually augmented with massively
-parallel Graphics Processing Units (GPUs) or other types of coprocessors.
+that requires runtime communication. Job execution is parallelized across
+multiple cores, each core calculating a small part of the problem and
+communicating with other cores via a message passing interface (MPI).
+Accordingly, supercomputers have large number of worker nodes, connected through
+a high-speed, low-latency dedicated network. Each worker node has multicore
+CPUs, usually augmented with massively parallel Graphics Processing Units (GPUs)
+or other types of specialized coprocessors.
 
-PanDA WMS has been designed to support distributed computing. Workload or
-workflow executions involves concurrent and/or sequential runs of possibly large
-amount of jobs, requiring no or minimal parallelization and no communication at
-runtime. Accordingly, computing infrastructure like WLCG are designed to
-aggregate large amount of computing resources across sites, possibly dislocated
-across multiple geographical allocation. While each site may deploy runtime
-message-passing capabilities, usually these are not used to perform distributed
-computations.
+PanDA WMS has been designed to support distributed computing. Executing
+Workloads or workflows involves concurrent and/or sequential runs of possibly
+large amount of jobs, each requiring no or minimal parallelization and no
+communication at runtime. Accordingly, computing infrastructure like WLCG have
+been designed to aggregate large amount of computing resources across sites,
+often dislocated across geographical locations. While each site may deploy
+runtime message-passing capabilities, usually these are not used to perform
+distributed computations.
 
 There are at least two approaches to enable PanDA WMS to execute ATLAS jobs on
-supercomputers: using the subset of resources and capabilities shared by both
-supercomputers and grid; reconciling the two paradigms of computation by means
-prototype a working solution that enables computation and better understanding
-of the problem space. The latter approach is principled and better suited for a
-production-grade solution developed on the base of the experience accumulated
-with the prototype and capable of supporting general-purpose workloads and
-workflows on both supercomputers and grid infrastructures.
+supercomputers: (i) using the subset of resources and capabilities shared by
+both supercomputers and Grid; (ii) reconciling the parallel and distributed
+computing paradigms by means of dedicated abstractions. The former is a
+pragmatic approach that enables the execution of specific workloads by
+prototyping single-point solutions. The latter approach is principled and better
+suited for a production-grade solution, capable of supporting general-purpose
+workloads and workflows on both supercomputers and Grid infrastructures. These
+two approaches are not mutually exclusive: developing a single-point solution
+gives the opportunity to better understand the problem space, supporting the
+creation of abstractions for the integration of computing paradigms.
 
-This section illustrates the design, architecture and execution process of the
-prototype developed by the PanDA team to support execution of some of the ATLAS
-workload on the Titan supercomputer. After a critical analysis of the results
-obtained and the lessons learned, the following section introduces the design
-and first experimental characterization of a next generation executor. This
-components is designed to abstract resources and capabilities enabling the
-concurrent execution of both parallel and distributed computing.
+This section illustrates the design, architecture and execution process of a
+prototype developed by the PanDA team to support execution of the Geant4 stage
+of the ATLAS production Monte Carlo workflow on the Titan supercomputer, managed
+by the Oak Ridge Leadership Computing Facility (OLCF) at the Oak Ridge National
+Laboratory (ORNL). After an analysis of the results obtained and the lessons
+learned, the following section introduces the design and first experimental
+characterization of a next generation executor (NGE). NGE is designed to
+abstract resources and capabilities, enabling the concurrent execution of both
+parallel and distributed computing.
 
@@
 
-The Titan supercomputer, current number two (number one from November 2012 until
-June 2013) on the Top 500 list~\cite{top500} is located at the Oak Ridge
-Leadership Computing Facility (OLCF) in Oak Ridge National Laboratory, US\@.
-Titan is a Cray XK7 system with 18,688 compute nodes and a total of 299,008 CPU
-cores. Each compute node has an AMD Opteron CPU, a Nvidia Tesla GPU, 32 GB of
-RAM and no local storage, though a 16 GB RAM disk can be set up. Compute nodes
-use Cray’s Gemini interconnect for inter-node MPI messaging but have no network
-connection to the outside world. Titan is served by the Spider
-II~\cite{spider2}, a Lustre filesystem with 32 PB of disk storage, and by a 29
-PB HPSS tape storage system. Titan’s worker nodes run Compute Node Linux, a run
-time environment based on SUSE Linux Enterprise Server.
+The Titan supercomputer, current number two
+% (number one from November 2012 until June 2013)
+on the Top 500 list~\cite{top500}, is a Cray XK7 system with 18,688 compute
+nodes and a total of 299,008 CPU cores. Each compute node has an AMD Opteron
+CPU, a Nvidia Tesla GPU, 32 GB of RAM and no local storage, though a 16 GB RAM
+disk can be set up. Compute nodes use Cray’s Gemini interconnect for inter-node
+MPI messaging but have no network connection to the outside world. Titan is
+served by the Spider II~\cite{spider2}, a Lustre filesystem with 32 PB of disk
+storage, and by a 29 PB HPSS tape storage system. Titan’s worker nodes run
+Compute Node Linux, a run time environment based on SUSE Linux Enterprise
+Server.
 
@@
 
-Titan offers to its users login nodes and data transfer nodes (DTN). Users have
+Titan offers to its users login nodes and data transfer nodes (DTNs). Users have
 to log into one of these nodes via ssh to submit jobs to Titan's PBS scheduler.
@@
 authentication with a RSA SecurID key, generated every 30 seconds. Login nodes
-and DTN are the only nodes with out/inbound wide area network connectivity. Work
-nodes have only local network access. Data staging is supported via the DTNs,
-from the wide area network and the OLCF high performance storage system (HPSS).
-DTNs are shared across all Titan's users that can schedule jobs to a dedicates
-PBS batch system to automate data staging. Fair-share policies are in place both
-for the batch system that for the use of DTN resources.
+and DTNs are the only nodes with out/inbound wide area network connectivity.
+Work nodes have only local network access. Data staging is supported via the
+DTNs, from the wide area network and the OLCF high performance storage system
+(HPSS). DTNs are shared across all Titan's users that can schedule jobs to a
+dedicates PBS batch system to automate data staging. Fair-share policies are in
+place both for the batch system and for the use of DTN resources.
 
@@
 module of the PanDA Server to pull jobs to execute. Titan's work nodes do not
-offer outbound network connectivity making the current deployment model of PanDA
-pilots unfeasible. Titan does not support PanDA's security model based on
+offer outbound network connectivity, making the current deployment model of
+PanDA pilots unfeasible. Titan does not support PanDA's security model based on
 certificates and virtual organizations, making the PanDA's approach to identity
@@
 the current prototyping phase. Finally, the specific characteristics of the
-execution environment, especially the dependences on a Lustre filesystem and on
-modules tailored to Compute Node Linux, require reengineering of ATLAS
+execution environment, especially the absence of local storage on the work nodes
+and modules tailored to Compute Node Linux, require reengineering of ATLAS
 application frameworks.
@@
 Currently, no HEP application can benefit from Titan's GPUs but some
-computationally-intensive, non memory-intensive tasks can be offloaded from the
-grid to Titan's large amount of cores. Further, when HEP tasks can be
+computationally-intensive and non memory-intensive tasks can be offloaded from
+the Grid to Titan's large amount of cores. Further, when HEP tasks can be
 partitioned to independent jobs, Titan work nodes can be used to execute up to
-16 concurrent jobs' payload. Given these constraints and challenges, the type of
-task most suitable for execution on Titan is HEP event generation. Event
-generators are mostly computational and stand-alone code, requiring less than
-1GB of RAM at runtime and with small input data requirements. The event
-generation in ATLAS accounts for 10--15\% of all jobs on the Grid, corresponding
-to 2/3 of the current grid capacity. A more difficult workload to adapt for
-Titan would be that for detector simulation. Simulation tasks tend to be tied to
-the framework of a particular simulation, often requiring database access for
-geometry and relevant run conditions.
+16 concurrent jobs' payload, one for each core. Given these constraints and
+challenges, the type of task most suitable for execution on Titan is HEP event
+generation. Event generations are mostly computational and stand-alone code,
+requiring less than 1GB of RAM at runtime and with small input data
+requirements. The event generation in ATLAS accounts for 10--15\% of all jobs on
+the Grid, corresponding to 2/3 of the current Grid capacity. A more difficult
+workload to adapt for Titan would be that for detector simulation. Simulation
+tasks tend to be tied to the framework of a particular simulation, often
+requiring database access for geometry and relevant run conditions.
 
-\mtnote{Describe Monte Carlo workflow and Geant4 here.}
+\mtnote{Currently, is PanDA running event generation, simulation or both on Titain?}
+\mtnote{Please provide a diagram illustrating event generation and simulation
+stats compared to the whole ATLAS computation activity.}
+
+Event generation and detector simulation are part of the ATLAS production Monte
+Carlo (MC) workflow (also known as MC production chain)~\cite{mcworkflow}. The
+MC workflow consists of four main stages: event generation, detector simulation,
+digitization, and reconstruction. Even generation creates sets of particle
+four-momenta via different generators, e.g., PYTHIA~\cite{herwig} and
+HERWIG~\cite{herwig}. The detector simulator is called Geant4~\cite{geant4} and
+simulates the interaction of these particles with the sensitive material of the
+ATLAS detector. Each interaction creates a so-called hit and all hits are
+collected and passed to the digitalization stage. There, hits are further
+process to mimic the readout of the detector, including subtle events specific
+to each detector's sub-component. Finally, the reconstruction stage operates
+local pattern recognition, creating high-level objects like particles and jets.
 
@@
 
-The most relevant design constraint for integrating PanDA WMS and Titan is the
-lack of connectivity from the work nodes. This makes impossible to port PanDA
-Pilot while maintaining the defining feature of the pilot abstraction:
-multi-stage scheduling to decouple resource acquisition from workload scheduling
-and execution. As direct communication between PanDA Pilot and PanDA Server is
-required, PanDA Pilots cannot be scheduled on Titan and jobs cannot be pulled by
-a pilot from the work nodes.
+The lack of wide area network connectivity on Titan's work nodes is the most
+relevant challenge for integrating PanDA WMS and Titan. Without connectivity,
+Panda Pilots cannot be scheduled on work nodes because they would not be able to
+communicate with PanDA Server and therefore pull and execute jobs. This makes
+impossible to port PanDA Pilot to Titan while maintaining the defining feature
+of the pilot abstraction: decoupling resource acquisition from workload
+execution via multi-stage scheduling.
 
-Usually, this design limitation would make the integration of PanDA and Titan
-unfeasible. Pilot-based executions are necessary to avoid waiting in the
-supercomputer's queue to execute every single job of a distributed workload.
-This is particularly true for Titan where, consistently with the intended use
-case of the machine, queuing policies privilege parallel jobs on the base of the
-quantities of work nodes requested. It would still be possible to execute HEP
-event generation tasks but the time spent waiting in the queue would be too
-large compared to the execution time.
+The unavailability of pilots is a potential drawback when executing distributed
+workloads like MC event generation and simulation. Pilots are used to increase
+the throughput of distributed workloads: while pilots have to wait in the
+supercomputer's queue, once scheduled, they can pull and execute jobs avoiding
+the supercomputer's queue. This is particularly relevant for machines like Titan
+where queue policies privilege parallel jobs on the base of the number of work
+nodes they request: the higher the number of nodes, the shorter the amount of
+queue time (modulo fair share policies).
 
-About 10\% of Titan's capacity is unused due to mismatches between job sizes and
-available resources~\cite{titan_utilization}. Worker nodes are unused because
-there are not enough nodes to execute large jobs or the required amount of nodes
-would not be available for a long enough time. This makes available around
-30,000 cores for a total of approximately 270M core hours per
-year~\cite{titan_free}, roughly 30\% of the overall capacity of WLCG. A system
-that can occupy those temporarily free nodes with smaller scale tasks would be
-very valuable. For this reason, Titan offers backfill functionalities: users can
-interrogate Titan's Moab scheduler about how many free nodes are currently
-available and for how long. If a user submit a request for that number of nodes,
-queue time will be minimal.
+% MC event generation and simulation tasks can be executed without pilots but
+% the time spent waiting in the queue would be too large compared to the
+% execution time of each task.
+
+Titan's backfill functionality offers the opportunity not to use the pilot
+abstraction while still avoiding queue time overheads. At any point in time,
+Titan's Moab scheduler shows how many work nodes cannot be used by the jobs
+already queued: All these jobs are either too large or their walltime is too
+long. By submitting a job requesting the number of free work nodes and walltime,
+queue time can be avoided.
+
+Compared to pilots, backfill has the disadvantage of limiting the amount of work
+nodes that can be requested. Pilots are normal jobs: they can request as many
+work nodes for as much time as a queue can offer. On the contrary, jobs sized on
+the base of the backfill availability depend on the amount of work nodes that
+cannot be given to any other job. Usually, this amount of work nodes is a
+fraction of the total capacity of the queue but the size of Titan mitigates this
+limitation. Every year, about 10\% of Titan's capacity remains
+unused~\cite{titan_utilization}, for a total of 30,000 cores that correspond to
+270M core hours per year, roughly 30\% of the overall capacity of WLCG.
+
+% A system that can use those temporarily free work nodes with smaller
+% scale tasks would be very valuable. For this reason, Titan offers backfill
+% functionalities: users can interrogate Titan's Moab scheduler about how many
+% free nodes are currently available and for how long. If a user submit a
+% requestvfor that number of nodes and walltime, queue time will be minimal.
 
@@
 
-Backfill functionality offers the opportunity not to use the pilot abstraction
-while still avoiding too much queue time overhead. On Titan, PanDA Pilot has
-been redesigned to serve as a job broker. Architected as a stand-alone service
-to be executed on a DTN, this PanDA Broker implements functionalities to: (i)
-interrogate Titan about backfill availability; (ii) pull Geant4 jobs and events;
-(iii) wrap Geant4 payloads into PBS MPI job; (iv) submitting PBS jobs to Titan's
-batch system; and (v) staging input/output files. The new functionalities of
-PanDA Broker involve backfill querying, payload wrapping, and job submission.
-Functionalities for Geant4 job and events pulling, and file staging were mostly
-inherited from PanDA Pilot.
+Accordingly, PanDA Pilot has been redesigned to serve as a job broker On Titan.
+Architected as a stand-alone service to be executed on a DTN, this PanDA Broker
+implements functionalities to: (i) interrogate Titan about backfill
+availability; (ii) pull MC jobs and events; (iii) wrap these jobs' payload into
+PBS MPI jobs; (iv) submitting PBS jobs to Titan's batch system; and (v) staging
+input/output files. Backfill querying, payload wrapping, and job submission
+required a new implementation while job and events pulling, and file staging
+were mostly inherited from PanDA Pilot.
 
@@
 
-Backfill querying is performed via the dedicated Moab command while payload
-wrapping is done via a tailored Python MPI script. This MPI wrapper serves as a
+Backfill querying is performed via a dedicated Moab command while a tailored
+Python MPI script is used to wrap job's payload. This MPI wrapper serves as a
 ``container'' for non-MPI jobs, enabling execution of unmodified Grid-centric
@@
 MPI wrappers are submitted to Titan's PBS batch system via
-RADICAL-SAGA~\cite{radical-saga}. SAGA (Simple API for Grid Applications)
-provides a unified interface to diverse job schedulers and file transferring
-services. SAGA provides an interoperability layer that lowers the complexity and
-improves the simplicity of using distributed infrastructure. RADICAL-SAGA is a
-Python module, compliant with the OGF GFD.90 SAGA
-specification~\cite{saga-spec}. Behind the API façade, RADICAL-SAGA implements a
-adaptor architecture: each adaptor interface the SAGA API with different
-middleware systems and services, including the batch scheduler of Titan.
+RADICAL-SAGA~\cite{radical-saga}. The Simple API for Grid Applications (SAGA)
+offers a unified interface to diverse job schedulers and file transferring
+services. In this way, SAGA provides an interoperability layer that lowers the
+complexity of using distributed infrastructure. RADICAL-SAGA is a Python module,
+compliant with the OGF GFD.90 SAGA specification~\cite{saga-spec}. Behind the
+API façade, RADICAL-SAGA implements a adaptor architecture: each adaptor
+interface the SAGA API with different middleware systems and services, including
+the batch scheduler of Titan.
 
-PanDA Broker stages capabilities are implemented via a file system shared among
-DTNs and work nods. Input files that are required by the payload are downloaded
-via the ATLAS DDM service and then stored in a directory on the shared
-filesystem. Part of the MPI wrapper setup process is to make the location of
-these files known to the payload. Payloads' output files are located on the
-shared filesystem by the broker and transferred from Titan to the ATLAS Tier 1
-computing center at Brookhaven National Lab (BNL). In principle any Grid site
-could be used.
+The data staging capabilities of the PanDA Broker are implemented via a file
+system that is shared among DTNs and work nodes. The input files required by the
+payload are downloaded via the ATLAS DDM service and then stored in a directory
+on the shared filesystem. The MPI wrapper setup process includes making the
+location of these files known to the payload. The PanDA Broker can locate the
+Payloads' output files on the shared filesystem and transfer them from Titan to
+the ATLAS Tier 1 computing center at Brookhaven National Lab (BNL) or any other
+Grid site.
 
-The execution process of Geant4 on Titan by PanDA has X steps: (i) PanDA Broker
-queries Titan's Moab scheduler about currently unused nodes and checks whether
-the time and size of the available backfill resources are suitable to execute
-jobs of Geant4 tasks; (ii) The PanDA Broker transmits this information to the
-Job Dispatcher module of the PanDA server, and in response gets a list of Geant4
-jobs bound to Titan; (iii) based on the jobs attributes, the broker requests to
-the Data Service module of the PanDA Server to transfer the necessary input data
-via the ATLAS DDM service; (iv) once all the data is transferred, the PanDA
-Broker calculates the number of cores and duration of jobs that can be submitted
-to Titan on the base of the information collected about backfill resource
-availability; (v) the broker creates one or more PBS MPI scripts to wrap enough
-Geant4 payload to fit the calculated number and duration of cores; (vi) the
-broker submits jobs to Titan's PBS batch system via RADICAL-SAGA; (vii) the PBS
-jobs waits a minimal amount of time into the queue until is scheduled for
-execution; (viii) upon execution on the worker nodes, the MPI script sets up the
-execution environment required by the wrapped Geant4 payload, including
-assembling the AthenaMP framework in ramdisk; (ix) the payload is then executed
-into dedicated processes, 16 for each work node, 1 for each core; (x) the PanDA
-Broker locates on the shared filesystem the output files produced by the payload
-and transfer them to BNL ATLAS Tier 1 computing centre.
+Once deployed, every PanDA Broker supports the execution of MC event generation
+and detector simulation on Titan in 10 steps: (i) PanDA Broker queries Titan's
+Moab scheduler about currently unused worker nodes;
+% and checks whether the time and size of the available backfill resources are
+% suitable to execute jobs;
+(ii) The PanDA Broker transmits this information to the Job Dispatcher module of
+the PanDA server and, in response, it gets a list of jobs that have been bound
+to Titan by JEDI; (iii) based on the jobs attributes, the PanDA Broker requests
+to the Data Service module of the PanDA Server to transfer the necessary input
+data to the DTN via the ATLAS DDM service; (iv) once all the data are
+transferred, the PanDA Broker calculates the number of cores and duration of
+jobs that can be submitted to Titan on the base of the information collected
+about backfill worker nodes availability; (v) the PanDA Broker creates one or
+more PBS MPI scripts to wrap enough jobs' payload to fit the work nodes
+available via backfill; (vi) the PanDA Broker submits these MPI jobs to the
+Titan's PBS batch system via RADICAL-SAGA; (vii) the MPI jobs waits a minimal
+amount of time into the queue until are scheduled for execution; (viii) upon
+execution on the worker nodes, each MPI script sets up the execution environment
+required by the wrapped payload, assembling a specific AthenaMP configuration in
+ramdisk; (ix) the payload is executed into dedicated processes, 16 for each work
+node, 1 for each core; (x) the PanDA Broker locates the output files produced by
+the payload on the shared filesystem and transfer them to BNL ATLAS Tier 1
+computing centre.
 

### 2017-03-06 paper.tex
Itartion workflow description
Inversion between WMSLiterature and Terminology
Section III
First pass IV (former IV and V) + restructuring
Iteration IV

@@
 
-\title{Everthing you wanted to know about PanDA but were truly truly afraid to
+\title{Everything you wanted to know about PanDA but were truly truly afraid to
 ask}
@@
 \section{Background and Related Work}
-\label{sec:bkgrd_literature}
+\label{sec:bkgrd}
+
+\begin{enumerate}
+  \item Brief overview of distributed CI\@.
+  \item Workload vs Workflow. (DONE)
+  \item Distinguish between WFMS and WLMS\@.
+  \begin{enumerate}
+    \item Role of Workload MS\@: help ``hide'' and federate heterogeneous resources.
+  \end{enumerate}
+\end{enumerate}
+
+% Please do not correct bckground to background.
+\input{bckground}
 
@@
 
-\begin{enumerate}
-  \item Brief overview of distributed CI\@.
-  \item Workload vs Workflow. (DONE)
-  \item Distinguish between WFMS and WLMS\@.
-  \begin{enumerate}
-    \item Role of Workload MS\@: help ``hide'' and federate heterogeneous resources.
-  \end{enumerate}
-\end{enumerate}
-
-% Please do not correct bckground to background.
-\input{bckground}
-
 % ----------------------------------------------------------------------------
@@
 
-\begin{enumerate}
-  \item PanDA WMS
-  \begin{enumerate}
-    \item State Diagrams
-    \item Control Flow Diagrams
-    \item Software Module and Architecture diagrams.
-  \end{enumerate}
-  \item PANDA Pilot
-\end{enumerate}
-
 \input{PandaOverview}
@@
 % ----------------------------------------------------------------------------
-% IV - ATLAS MONTE CARLO WORKFLOW
-% ----------------------------------------------------------------------------
-\section{ATLAS Monte Carlo Workflow}
-\label{sec:mc_workflow}
-
-\input{workflow}
-
-% ----------------------------------------------------------------------------
-% V - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
+% IV - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
 % ----------------------------------------------------------------------------
@@
 
-\begin{enumerate}
-  \item Why? To support a specific stage of the ATLAS Monte Carlo Workflow.
-  \item Challenges
-  \item Opportunities
-  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
-  \item Take-aways, lessons learned
-\end{enumerate}
-
-\input{titan}
-
-% ----------------------------------------------------------------------------
-% VI - EVOLUTION OF PANDA WMS
-% ----------------------------------------------------------------------------
-% \section{Evolution of PanDA WMS}
-% \label{sec:panda_evolution}
+% \begin{enumerate}
+%   \item Why? To support a specific stage of the ATLAS Monte Carlo Workflow.
+%   \item Challenges
+%   \item Opportunities
+%   \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
+%   \item Take-aways, lessons learned
+% \end{enumerate}
 
@@
 % ----------------------------------------------------------------------------
-% \section{Performance Characterization}
-% \label{sec:panda_performance}
-%
-% \input{perf}
-
-% ----------------------------------------------------------------------------
-% VI - PERFORMANCE CHARACTERIZATION
-% ----------------------------------------------------------------------------
-% \section{Putting it all together: Experience of PanDA on TITAN} % (fold)
 \section{Performance Characterization}

### 2017-03-06 titan.tex
Iteration IV

@@
 
-The topology and architecture details of Spider II infrastructure are
-illustrated in Figure~\ref{fig:xk7-compute-node} and
-described as follows:
-
-\begin{figure}[!htb]
-    \centering
-    \begin{tabular}{cc}
-        {\includegraphics[width=0.48\textwidth]{figures/spider2arch.pdf}}\\
-    \end{tabular}
-    \caption{Infrastructure and I/O path between Titan (Cray XK7) and its backend storage}
-    \label{fig:xk7-compute-node}
-\end{figure}
-
-Spider II is based on the Lustre technology~\cite{lustre}.  It is configured
-and deployed as two independent, non-overlapping file systems, each with 144
-Lustre Object Storage Servers (OSSs) and 1,008 Lustre Object Storage Targets
-(OSTs). Spider II offers ~30 PB usable file system capacity to OLCF systems and
-users and has over 1 TB/s aggregate write I/O performance.
+% The topology and architecture details of Spider II infrastructure are
+% illustrated in Figure~\ref{fig:xk7-compute-node} and
+% described as follows:
+%
+%
+% \begin{figure}[!htb]
+%     \centering
+%     \begin{tabular}{cc}
+%         {\includegraphics[width=0.48\textwidth]{figures/spider2arch.pdf}}\\
+%     \end{tabular}
+%     \caption{Infrastructure and I/O path between Titan (Cray XK7) and its backend storage}
+%     \label{fig:xk7-compute-node}
+% \end{figure}
+%
+%
+%
+% Spider II is based on the Lustre technology~\cite{lustre}.  It is configured
+% and deployed as two independent, non-overlapping file systems, each with 144
+% Lustre Object Storage Servers (OSSs) and 1,008 Lustre Object Storage Targets
+% (OSTs). Spider II offers ~30 PB usable file system capacity to OLCF systems and
+% users and has over 1 TB/s aggregate write I/O performance.

### 2017-03-06 PandaEvolution.tex
Second iteration IV

@@
 0.3 petaFLOPS. This capacity will be sufficient for the planned analysis and
-data processing, but it will be insufficient for the Monte Carlo (MC) production
+data processing, but it will be insufficient for the Monte Carlo production
 workflow and any extra activity. To alleviate these challenges, ATLAS is engaged
@@
 
-Modern supercomputers have been designed to mainly support parallel computation
+Modern supercomputers have been designed mainly to support parallel computation
 that requires runtime communication. Job execution is parallelized across
@@
 
-PanDA WMS has been designed to support distributed computing. Executing
-Workloads or workflows involves concurrent and/or sequential runs of possibly
+PanDA WMS has been designed to support distributed computing. Executing ATLAS
+workloads or workflows involves concurrent and/or sequential runs of possibly
 large amount of jobs, each requiring no or minimal parallelization and no
-communication at runtime. Accordingly, computing infrastructure like WLCG have
-been designed to aggregate large amount of computing resources across sites,
-often dislocated across geographical locations. While each site may deploy
-runtime message-passing capabilities, usually these are not used to perform
-distributed computations.
+runtime communication. Thus, computing infrastructure like WLCG have been
+designed to aggregate large amount of computing resources across sites, often
+geographically dislocated. While each site may deploy runtime message-passing
+capabilities, usually these are not used to perform distributed computations.
 
-There are at least two approaches to enable PanDA WMS to execute ATLAS jobs on
-supercomputers: (i) using the subset of resources and capabilities shared by
-both supercomputers and Grid; (ii) reconciling the parallel and distributed
-computing paradigms by means of dedicated abstractions. The former is a
-pragmatic approach that enables the execution of specific workloads by
-prototyping single-point solutions. The latter approach is principled and better
-suited for a production-grade solution, capable of supporting general-purpose
-workloads and workflows on both supercomputers and Grid infrastructures. These
-two approaches are not mutually exclusive: developing a single-point solution
-gives the opportunity to better understand the problem space, supporting the
-creation of abstractions for the integration of computing paradigms.
+There are at least two approaches to enable PanDA WMS to execute ATLAS workloads
+or workflows on supercomputers: (i) using the subset of resources and
+capabilities shared by both supercomputers and WLCG; (ii) reconciling the
+parallel and distributed computing paradigms by means of dedicated abstractions.
+The former is a pragmatic approach that enables the execution of specific
+workloads by prototyping single-point solutions. The latter is a principled
+approach, better suited for a production-grade solution, capable of supporting
+general-purpose workloads and workflows on both supercomputers and Grid
+infrastructures. These two approaches are not mutually exclusive: developing a
+single-point solution gives the opportunity to better understand the problem
+space, supporting the creation of abstractions for the integration of computing
+paradigms.
 
-This section illustrates the design, architecture and execution process of a
-prototype developed by the PanDA team to support execution of the Geant4 stage
-of the ATLAS production Monte Carlo workflow on the Titan supercomputer, managed
-by the Oak Ridge Leadership Computing Facility (OLCF) at the Oak Ridge National
-Laboratory (ORNL). After an analysis of the results obtained and the lessons
-learned, the following section introduces the design and first experimental
-characterization of a next generation executor (NGE). NGE is designed to
-abstract resources and capabilities, enabling the concurrent execution of both
-parallel and distributed computing.
+This section illustrates the design and architecture of a job broker prototyped
+by the PanDA team. This broker supports execution of part of the ATLAS
+production Monte Carlo workflow on Titan, a leadership-class supercomputer
+managed by the Oak Ridge Leadership Computing Facility (OLCF) at the Oak Ridge
+National Laboratory (ORNL). After an analysis of the results obtained and the
+lessons learned, the following section introduces the design and first
+experimental characterization of a next generation executor (NGE). NGE is
+designed to abstract resources and capabilities, enabling the concurrent
+execution of both parallel and distributed computing on generic high performance
+computing (HPC) machines.
 
@@
 
-The Titan supercomputer, current number two
+The Titan supercomputer, current number three
 % (number one from November 2012 until June 2013)
-on the Top 500 list~\cite{top500}, is a Cray XK7 system with 18,688 compute
-nodes and a total of 299,008 CPU cores. Each compute node has an AMD Opteron
-CPU, a Nvidia Tesla GPU, 32 GB of RAM and no local storage, though a 16 GB RAM
-disk can be set up. Compute nodes use Cray’s Gemini interconnect for inter-node
-MPI messaging but have no network connection to the outside world. Titan is
-served by the Spider II~\cite{spider2}, a Lustre filesystem with 32 PB of disk
-storage, and by a 29 PB HPSS tape storage system. Titan’s worker nodes run
-Compute Node Linux, a run time environment based on SUSE Linux Enterprise
-Server.
+on the Top 500 list~\cite{top500}, is a Cray XK7 system with 18,688 work nodes
+and a total of 299,008 CPU cores. Each work node has an AMD Opteron CPU, a
+Nvidia Tesla GPU, 32 GB of RAM and no local storage, though a 16 GB RAM disk can
+be set up. Work nodes use Cray’s Gemini interconnect for inter-node MPI
+messaging. Titan is served by the Spider II~\cite{spider2}, a Lustre filesystem
+with 32 PB of disk storage, and by a 29 PB HPSS tape storage system. Titan’s
+work nodes run Compute Node Linux, a run time environment based on SUSE Linux
+Enterprise Server.
 
@@
 (HPSS). DTNs are shared across all Titan's users that can schedule jobs to a
-dedicates PBS batch system to automate data staging. Fair-share policies are in
+dedicates batch system to automate data staging. Fair-share policies are in
 place both for the batch system and for the use of DTN resources.
@@
 partitioned to independent jobs, Titan work nodes can be used to execute up to
-16 concurrent jobs' payload, one for each core. Given these constraints and
-challenges, the type of task most suitable for execution on Titan is HEP event
-generation. Event generations are mostly computational and stand-alone code,
-requiring less than 1GB of RAM at runtime and with small input data
-requirements. The event generation in ATLAS accounts for 10--15\% of all jobs on
-the Grid, corresponding to 2/3 of the current Grid capacity. A more difficult
-workload to adapt for Titan would be that for detector simulation. Simulation
-tasks tend to be tied to the framework of a particular simulation, often
-requiring database access for geometry and relevant run conditions.
+16 concurrent jobs' payload, one for each available core. Given these
+constraints and challenges, the types of task most suitable for execution on
+Titan is Monte Carlo event generation and detector simulation. Both types of
+task are mostly computational and stand-alone code, requiring less than 1GB of
+RAM at runtime and with small input data requirements. The event generation in
+ATLAS accounts for 10--15\% of all jobs on WLCG, corresponding to 2/3 of the
+current Grid capacity. Detector simulation accounts for \ldots .
 
-\mtnote{Currently, is PanDA running event generation, simulation or both on Titain?}
-\mtnote{Please provide a diagram illustrating event generation and simulation
-stats compared to the whole ATLAS computation activity.}
+\mtnote{Currently, is PanDA running event generation, detector simulation or
+both on Titain?}
+\mtnote{Please provide a diagram illustrating event generation and detector
+simulation stats compared to the whole ATLAS computation activity.}
 
@@
 ATLAS detector. Each interaction creates a so-called hit and all hits are
-collected and passed to the digitalization stage. There, hits are further
-process to mimic the readout of the detector, including subtle events specific
-to each detector's sub-component. Finally, the reconstruction stage operates
-local pattern recognition, creating high-level objects like particles and jets.
+collected and passed on for digitalization where hits are further process to
+mimic the readout of the detector. Finally, reconstruction operates local
+pattern recognition, creating high-level objects like particles and jets.
 
@@
 The unavailability of pilots is a potential drawback when executing distributed
-workloads like MC event generation and simulation. Pilots are used to increase
-the throughput of distributed workloads: while pilots have to wait in the
-supercomputer's queue, once scheduled, they can pull and execute jobs avoiding
-the supercomputer's queue. This is particularly relevant for machines like Titan
-where queue policies privilege parallel jobs on the base of the number of work
-nodes they request: the higher the number of nodes, the shorter the amount of
-queue time (modulo fair share policies).
+workloads like MC event generation and detector simulation. Pilots are used to
+increase the throughput of distributed workloads: while pilots have to wait in
+the supercomputer's queue, once scheduled, they can pull and execute jobs
+independently from the system's queue. This is particularly relevant for
+machines like Titan where queue policies privilege parallel jobs on the base of
+the number of work nodes they request: the higher the number of nodes, the
+shorter the amount of queue time (modulo fair share policies).
 
@@
 Titan's backfill functionality offers the opportunity not to use the pilot
-abstraction while still avoiding queue time overheads. At any point in time,
-Titan's Moab scheduler shows how many work nodes cannot be used by the jobs
-already queued: All these jobs are either too large or their walltime is too
-long. By submitting a job requesting the number of free work nodes and walltime,
-queue time can be avoided.
+abstraction while still avoiding queue time overheads. Backfill availability is
+the number of work nodes that cannot be used by any of the jobs already queued
+on Titan: All queued jobs are either too large or their walltime is too long. At
+any point in time, Titan's Moab scheduler can be queried for backfill
+availability and a job submitted shortly after placing this query and requesting
+for no more than the backfill availability, is usually scheduled immediately
+spending almost no time in the queue.
 
@@
 the base of the backfill availability depend on the amount of work nodes that
-cannot be given to any other job. Usually, this amount of work nodes is a
-fraction of the total capacity of the queue but the size of Titan mitigates this
+cannot be given to any other job. Usually, backfill availability is a fraction
+of the total capacity of the queue but the size of Titan mitigates this
 limitation. Every year, about 10\% of Titan's capacity remains
-unused~\cite{titan_utilization}, for a total of 30,000 cores that correspond to
-270M core hours per year, roughly 30\% of the overall capacity of WLCG.
+unused~\cite{titan_utilization}, for a total of 30,000 cores, corresponding to
+around 270M core hours per year, roughly 30\% of the overall capacity of WLCG.
 
@@
 
-Accordingly, PanDA Pilot has been redesigned to serve as a job broker On Titan.
-Architected as a stand-alone service to be executed on a DTN, this PanDA Broker
-implements functionalities to: (i) interrogate Titan about backfill
-availability; (ii) pull MC jobs and events; (iii) wrap these jobs' payload into
-PBS MPI jobs; (iv) submitting PBS jobs to Titan's batch system; and (v) staging
-input/output files. Backfill querying, payload wrapping, and job submission
-required a new implementation while job and events pulling, and file staging
-were mostly inherited from PanDA Pilot.
+Given the communication requirements of PanDA Pilots and the unused capacity of
+Titan, PanDA pilot was redesigned to serve as a job broker on Titan. Architected
+as a stand-alone service to be executed on a DTN, this prototype called `PanDA
+Broker' implements functionalities to: (i) interrogate Titan about backfill
+availability; (ii) pull MC jobs and events; (iii) wrap jobs' payload into PBS
+MPI jobs; (iv) submitting MPI PBS jobs to Titan's PBS batch system; and (v)
+staging input/output files. Backfill querying, payload wrapping, and job
+submission required a new implementation while job and events pulling, and file
+staging were mostly inherited from PanDA Pilot.
 
@@
 MPI wrappers are submitted to Titan's PBS batch system via
-RADICAL-SAGA~\cite{radical-saga}. The Simple API for Grid Applications (SAGA)
+RADICAL-SAGA~\cite{radical-saga}, a Python module, compliant with the OGF GFD.90
+SAGA specification~\cite{saga-spec}. The Simple API for Grid Applications (SAGA)
 offers a unified interface to diverse job schedulers and file transferring
 services. In this way, SAGA provides an interoperability layer that lowers the
-complexity of using distributed infrastructure. RADICAL-SAGA is a Python module,
-compliant with the OGF GFD.90 SAGA specification~\cite{saga-spec}. Behind the
-API façade, RADICAL-SAGA implements a adaptor architecture: each adaptor
-interface the SAGA API with different middleware systems and services, including
-the batch scheduler of Titan.
+complexity of using distributed infrastructures. Behind the API façade,
+RADICAL-SAGA implements a adaptor architecture: each adaptor interface the SAGA
+API with different middleware systems and services, including the batch
+scheduler of Titan.
 
@@
 on the shared filesystem. The MPI wrapper setup process includes making the
-location of these files known to the payload. The PanDA Broker can locate the
-Payloads' output files on the shared filesystem and transfer them from Titan to
-the ATLAS Tier 1 computing center at Brookhaven National Lab (BNL) or any other
-Grid site.
+location of these files available to the payload. The PanDA Broker can locate
+the Payloads' output files on the shared filesystem and transfer them from Titan
+to the ATLAS Tier 1 computing center at Brookhaven National Lab (BNL) or any
+other Grid site.
 
 Once deployed, every PanDA Broker supports the execution of MC event generation
-and detector simulation on Titan in 10 steps: (i) PanDA Broker queries Titan's
+and detector simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's
 Moab scheduler about currently unused worker nodes;
@@
 % suitable to execute jobs;
-(ii) The PanDA Broker transmits this information to the Job Dispatcher module of
-the PanDA server and, in response, it gets a list of jobs that have been bound
-to Titan by JEDI; (iii) based on the jobs attributes, the PanDA Broker requests
-to the Data Service module of the PanDA Server to transfer the necessary input
-data to the DTN via the ATLAS DDM service; (iv) once all the data are
-transferred, the PanDA Broker calculates the number of cores and duration of
-jobs that can be submitted to Titan on the base of the information collected
-about backfill worker nodes availability; (v) the PanDA Broker creates one or
-more PBS MPI scripts to wrap enough jobs' payload to fit the work nodes
-available via backfill; (vi) the PanDA Broker submits these MPI jobs to the
-Titan's PBS batch system via RADICAL-SAGA; (vii) the MPI jobs waits a minimal
-amount of time into the queue until are scheduled for execution; (viii) upon
-execution on the worker nodes, each MPI script sets up the execution environment
-required by the wrapped payload, assembling a specific AthenaMP configuration in
-ramdisk; (ix) the payload is executed into dedicated processes, 16 for each work
-node, 1 for each core; (x) the PanDA Broker locates the output files produced by
-the payload on the shared filesystem and transfer them to BNL ATLAS Tier 1
-computing centre.
+(ii) the PanDA Broker queries the Job Dispatcher module of the PanDA server for
+jobs that have been bound to Titan by JEDI; (iii)  the PanDA Broker requests to
+the Data Service module of the PanDA Server what input data these jobs require
+and transfer them to the DTN via the ATLAS DDM service; (iv) the PanDA Broker
+creates one or more PBS MPI script, wrapping enough jobs' payload to fit
+backfill availability; (v) the PanDA Broker submits these MPI jobs to the
+Titan's PBS batch system via RADICAL-SAGA; (vi) the MPI jobs waits a minimal
+amount of time into the queue until they are scheduled for execution; (vii) upon
+execution on the worker nodes, each MPI script creates a ramdisk and sets up the
+specific AthenaMP configuration required by the wrapped payload; (viii) the
+payload is executed into dedicated processes, 16 for each work node, 1 for each
+available core; (ix) upon completion of an MPI job, the PanDA Broker locates the
+produced output files on the shared filesystem and transfer them to BNL ATLAS
+Tier 1 computing centre.
 

### 2017-03-06 PandaEvolution.tex
Small editing
Eliminate event generation

@@
 16 concurrent jobs' payload, one for each available core. Given these
-constraints and challenges, the types of task most suitable for execution on
-Titan is Monte Carlo event generation and detector simulation. Both types of
-task are mostly computational and stand-alone code, requiring less than 1GB of
-RAM at runtime and with small input data requirements. The event generation in
-ATLAS accounts for 10--15\% of all jobs on WLCG, corresponding to 2/3 of the
-current Grid capacity. Detector simulation accounts for \ldots .
+constraints and challenges, the type of task most suitable for execution at the moment on Titan is Monte Carlo
+% event generation and
+detector simulation.
+% Both types of task are
+This type of task is mostly computational-intensive, requiring less than 1GB of
+RAM at runtime and with small input data requirements. Detector simulation tasks
+in ATLAS account for XX--XX\% of all jobs on WLCG, corresponding to X/X of the
+current Grid capacity.\mtnote{Please provide numbers to replace `XX' and,
+possibly, a diagram illustrating detector simulation stats compared to the whole
+ATLAS computation activity.}
 
-\mtnote{Currently, is PanDA running event generation, detector simulation or
-both on Titain?}
-\mtnote{Please provide a diagram illustrating event generation and detector
-simulation stats compared to the whole ATLAS computation activity.}
+% The event generation in ATLAS accounts for 10--15\% of all jobs on WLCG,
+% corresponding to 2/3 of the current Grid capacity.
 
-Event generation and detector simulation are part of the ATLAS production Monte
-Carlo (MC) workflow (also known as MC production chain)~\cite{mcworkflow}. The
-MC workflow consists of four main stages: event generation, detector simulation,
-digitization, and reconstruction. Even generation creates sets of particle
-four-momenta via different generators, e.g., PYTHIA~\cite{herwig} and
-HERWIG~\cite{herwig}. The detector simulator is called Geant4~\cite{geant4} and
-simulates the interaction of these particles with the sensitive material of the
-ATLAS detector. Each interaction creates a so-called hit and all hits are
-collected and passed on for digitalization where hits are further process to
-mimic the readout of the detector. Finally, reconstruction operates local
-pattern recognition, creating high-level objects like particles and jets.
+% Event generation and
+Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
+(also known as MC production chain)~\cite{mcworkflow}. The MC workflow consists
+of four main stages: event generation, detector simulation, digitization, and
+reconstruction. Even generation creates sets of particle four-momenta via
+different generators, e.g., PYTHIA~\cite{herwig} and HERWIG~\cite{herwig}. The
+detector simulator is called Geant4~\cite{geant4} and simulates the interaction
+of these particles with the sensitive material of the ATLAS detector. Each
+interaction creates a so-called hit and all hits are collected and passed on for
+digitalization where hits are further process to mimic the readout of the
+detector. Finally, reconstruction operates local pattern recognition, creating
+high-level objects like particles and jets.
 
@@
 The unavailability of pilots is a potential drawback when executing distributed
-workloads like MC event generation and detector simulation. Pilots are used to
-increase the throughput of distributed workloads: while pilots have to wait in
-the supercomputer's queue, once scheduled, they can pull and execute jobs
-independently from the system's queue. This is particularly relevant for
-machines like Titan where queue policies privilege parallel jobs on the base of
-the number of work nodes they request: the higher the number of nodes, the
-shorter the amount of queue time (modulo fair share policies).
+workloads like MC
+% event generation and
+detector simulation. Pilots are used to increase the throughput of distributed
+workloads: while pilots have to wait in the supercomputer's queue, once
+scheduled, they can pull and execute jobs independently from the system's queue.
+Jobs can be concurrently on every core available to the pilot and multiple
+generations of concurrent executions can be performed until the pilot's walltime
+is exhausted. This is particularly relevant for machines like Titan where queue
+policies privilege parallel jobs on the base of the number of work nodes they
+request: the higher the number of nodes, the shorter the amount of queue time
+(modulo fair share policies).
 
@@
 \mtnote{Please feel free to add details about backfill as needed.}
+\mtnote{Please note: I know the following departs for the current name given to
+the PanDA subsystem installed on Titan. I think there is a good reason to
+call it a broker instead of a pilot, and I think I explained it in the previous
+paragraphs. Please  take this just as a suggestion, something I would like to
+discuss in our meeting.}
 
@@
 
-Once deployed, every PanDA Broker supports the execution of MC event generation
-and detector simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's
+Once deployed, every PanDA Broker supports the execution of MC
+% event generation and
+detector simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's
 Moab scheduler about currently unused worker nodes;

### 2017-03-07 PandaEvolution.tex
Copy and address Sergey's comments + bibliography

@@
 
-PanDA WMS has been designed to support distributed computing. Executing ATLAS
-workloads or workflows involves concurrent and/or sequential runs of possibly
-large amount of jobs, each requiring no or minimal parallelization and no
-runtime communication. Thus, computing infrastructure like WLCG have been
-designed to aggregate large amount of computing resources across sites, often
-geographically dislocated. While each site may deploy runtime message-passing
-capabilities, usually these are not used to perform distributed computations.
+PanDA WMS has been designed to support distributed Grid
+computing\sergeynote{specifically "distributed Grid computing"}\mtnote{done.}.
+Executing ATLAS workloads or workflows involves concurrent and/or sequential
+runs of possibly large amount of jobs, each requiring no or minimal
+parallelization and no runtime communication. Thus, computing infrastructure
+like WLCG have been designed to aggregate large amount of computing resources
+across multiple sites.
+%, often geographically dislocated.
+\sergeynote{geographically distributed?}\mtnote{Better?} While each site may
+deploy runtime message-passing capabilities, usually these are not used to
+perform distributed computations.
 
@@
 % Both types of task are
-This type of task is mostly computational-intensive, requiring less than 1GB of
+This type of task is mostly computational-intensive, requiring less than 2GB of
 RAM at runtime and with small input data requirements. Detector simulation tasks
@@
 Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
-(also known as MC production chain)~\cite{mcworkflow}. The MC workflow consists
-of four main stages: event generation, detector simulation, digitization, and
-reconstruction. Even generation creates sets of particle four-momenta via
-different generators, e.g., PYTHIA~\cite{herwig} and HERWIG~\cite{herwig}. The
-detector simulator is called Geant4~\cite{geant4} and simulates the interaction
-of these particles with the sensitive material of the ATLAS detector. Each
-interaction creates a so-called hit and all hits are collected and passed on for
+(also known as MC production
+chain)~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
+consists of four main stages: event generation, detector simulation,
+digitization, and reconstruction. Even generation creates sets of particle
+four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia}
+and HERWIG~\cite{corcella2001herwig}. The detector simulator is called
+Geant4~\cite{agostinelli2003geant4} and simulates the interaction of these
+particles with the sensitive material of the ATLAS detector. Each interaction
+creates a so-called hit and all hits are collected and passed on for
 digitalization where hits are further process to mimic the readout of the
@@
 work nodes for as much time as a queue can offer. On the contrary, jobs sized on
-the base of the backfill availability depend on the amount of work nodes that
-cannot be given to any other job. Usually, backfill availability is a fraction
-of the total capacity of the queue but the size of Titan mitigates this
-limitation. Every year, about 10\% of Titan's capacity remains
+the base of the backfill availability depend on the
+% amount of work nodes that cannot be given to any other job.
+any other job in the Titan's queue\sergeynote{any other job currently in the
+Titan's queue?}\mtnote{Better?}. Usually, backfill availability is a fraction of
+the total capacity of the queue but the size of Titan mitigates this limitation.
+Every year, about 10\% of Titan's capacity remains
 unused~\cite{titan_utilization}, for a total of 30,000 cores, corresponding to
@@
 \mtnote{Please note: I know the following departs for the current name given to
-the PanDA subsystem installed on Titan. I think there is a good reason to
-call it a broker instead of a pilot, and I think I explained it in the previous
+the PanDA subsystem installed on Titan. I think there is a good reason to call
+it a broker instead of a pilot, and I think I explained it in the previous
 paragraphs. Please  take this just as a suggestion, something I would like to
-discuss in our meeting.}
+discuss in our meeting.} \sergeynote{It's not a pilot in conventional sense. I
+call it an agent. It just happened that we were able to use full PanDA pilot's
+code base to serve our purposes on Titan. that's why, by inertia, we still call
+it a pilot.}\mtnote{Thank you. We would have a problem calling it `agent' as we
+use that term to name the pilot of NGE. Would PanDA Broker work?}
 
@@
 availability; (ii) pull MC jobs and events; (iii) wrap jobs' payload into PBS
-MPI jobs; (iv) submitting MPI PBS jobs to Titan's PBS batch system; and (v)
-staging input/output files. Backfill querying, payload wrapping, and job
-submission required a new implementation while job and events pulling, and file
-staging were mostly inherited from PanDA Pilot.
+MPI jobs; (iv) submitting MPI PBS jobs to Titan's PBS batch system and monitor
+their execution;\sergeynote{Monitor job state and state transitions using
+SAGA}\mtnote{Better?} and (v) staging input/output files. Backfill querying,
+payload wrapping, and job submission required a new implementation while job and
+events pulling, and file staging were mostly inherited from PanDA Pilot.
 

### 2017-03-07 PandaOverview.tex
First pass IV (former IV and V) + restructuring
Copy and address Sergey's comments + bibliography

@@
 2005 to develop a WMS for the ATLAS experiment. PanDA went into production for
-the LHC Run 1 between 2009 and 2013 and was then partially redesigned to be
-deployed on Run 2, between 2015 and 2018. In this section, we summarize the
-design, architecture, and execution process of the current state of the
-practice.
+the LHC Run 1 on 2009 and was then partially redesigned to be deployed on Run 2,
+on 2015. In this section, we summarize the design, architecture, and execution
+process of the current state of the practice.
 
@@
 both attributes and states, and some of the attributes are shared between events
-and jobs so to create associations. Data are assumed to be stored centrally,
-replicated when necessary and transferred over the network. This assumptions
-applies to both input and output data.
+and jobs so to create associations. Raw, reconstruction, and simulation data are
+assumed to be distributed across multiple and tiered sites and managed by the
+ATLAS Distributed Data Management (DDM)~\cite{garonne2012atlas}. Datasets
+required by each job are assumed to be replicated when necessary over the
+network. This applies to both input and output data. \sergeynote{Raw data from
+the ATLAS detector is captured at CERN and goes through reconstruction step to
+produce data in secondary formats, smaller in volume and with more and more
+physics content instead of raw detector hits. Consequently a second, custodial,
+copy of raw data is being distributed in predetermined volumes to National T1
+centers around the world (like BNL  T1 center). This insures low raw data
+redundancy and safety. Multiple copies of chunks of secondary data sets (derived
+from real raw data) are also distributed to ~100 centers around the word for
+ease of access and improved availability for analysis. Then raw detector data is
+stored on Tapes at CERN, just  in case. Most users, who do physics analysis,
+rarely work with raw data, only a small group of experts needs to work with raw
+data for specialized tasks like detector calibrations and such. There may be a
+need to redo a reconstruction step, when errors in previous reconstruction
+campaign are uncovered or improved detector understanding will lead for a better
+quality physics. Then data stored on tape is recovered and re-processed again.
+Simulation data is produced in many places anyway and is not centralized to
+begin with. Here's the paper that describes current DDM state in ATLAS, data
+replication strategies and such.
+http://iopscience.iop.org/article/10.1088/1742-6596/396/3/032045/pdf DDM is a
+complex and evolving topic in ATLAS.}\mtnote{Thank you. As we are describing the
+design of PanDA I thought we should not enter into a too detailed description of
+the DDM design. I eliminated the mistake I had made about `centralized data' but
+left DDM just as a reference. Any better?}
 
@@
 of each job and the resource to which each job has been assigned; (ii) enable
-interaction with the ATLAS Distributed Data Management (DDM)~\cite{atlas_ddm} to
-make input files available to the jobs and manage their output data.
+interaction with the ATLAS DDM to make input files available to the jobs and
+manage their output data.
 
@@
 \paragraph{\textbf{AutoPyFactory}} this subsystem implements a factory to submit
-pilots locally and to both grid and cloud sites. Implemented to be executed as a
-single demonized process, AutoPyFactory enables the concurrent spawning of
-multiple job submission processes called APFQueue. Each APFQueue can communicate
-with PanDA Server and the Brokerage module to collect information about the
-state of the jobs it manages and, on the base of this information, submit new
-pilots to one site per APFQueue. AutoPyFactory implements functionalities to
-evaluate the number of pilots already instantiated on each site, and to apply a
-consistent and adequate pressure on their batch queues. Submission is performed
-via the Condor-G batch system~\cite{condorg}, enabling submission to multiple
+pilots locally and to both grid and cloud
+sites~\cite{caballero2012autopyfactory}. Implemented to be executed as a single
+demonized process, AutoPyFactory enables the concurrent spawning of multiple job
+submission processes called APFQueue. Each APFQueue can communicate with PanDA
+Server and the Brokerage module to collect information about the state of the
+jobs it manages and, on the base of this information, submit new pilots to one
+site per APFQueue. AutoPyFactory implements functionalities to evaluate the
+number of pilots already instantiated on each site, and to apply a consistent
+and adequate pressure on their batch queues. Submission is performed via the
+Condor-G batch system~\cite{frey2002condor}, enabling submission to multiple
 type of sites. AutoPyFactory exposes a dedicated monitoring service, feeding
@@
 \paragraph{\textbf{PanDA Monitoring}} this subsystem is implemented as a
-web-based, dashboard-style graphical application. It runs on Apache and
-interacts with the back-end database of PanDA Server to enable persistence.
-PanDA Monitoring enables users and site administrators to gather information
-about the status of current and past jobs, data movement, and pilot factory.
-PanDA Monitoring allows users access to all log files of each job, greatly
-simplifying code debugging and failure analysis in a distributed computing
-environment.
+web-based, dashboard-style graphical application~\cite{klimentov2011atlas}. It
+runs on Apache and interacts with the back-end database of PanDA Server to
+enable persistence. PanDA Monitoring enables users and site administrators to
+gather information about the status of current and past jobs, data movement, and
+pilot factory. PanDA Monitoring allows users access to all log files of each
+job, greatly simplifying code debugging and failure analysis in a distributed
+computing environment.
 
 \paragraph{\textbf{JEDI}} this subsystem implements a workload and workflow
-management for tasks instead of jobs, as defined by HEP users. Implemented as a
-standalone service that interacts with Panda Server via its REST interface, JEDI
-offers two main functionalities: partitioning tasks to jobs, and binding jobs to
-resources via dedicated queues. JEDI partition tasks to jobs by matching tasks
-requirements to resource capabilities. This matching is based on tasks
-attributes, and static and dynamic information about the resources of each
-available site. Dynamic information is acquired via a small number of `scout'
-jobs that collect information about current resources and their performance.
-Functionally, JEDI defines a set of queues associated to groups of resources
-with specific capabilities. Tasks are assigned to groups of these queues, and
-the jobs in which each task is partitioned are assigned to each queue, depending
-on their requirements. JEDI replaces the job brokerage functionalities of the
-Brokerage module of the PanDA Server but this module still handles communication
-between the Task Buffer and the Data Service modules.
+management for tasks instead of jobs, as defined by HEP
+users~\cite{borodin2015scaling}. Implemented as a standalone service that
+interacts with Panda Server via its REST interface, JEDI offers two main
+functionalities: partitioning tasks to jobs, and binding jobs to resources via
+dedicated queues. JEDI partition tasks to jobs by matching tasks requirements to
+resource capabilities. This matching is based on tasks attributes, and static
+and dynamic information about the resources of each available site. Dynamic
+information is acquired via a small number of `scout' jobs that collect
+information about current resources and their performance. Functionally, JEDI
+defines a set of queues associated to groups of resources with specific
+capabilities. Tasks are assigned to groups of these queues, and the jobs in
+which each task is partitioned are assigned to each queue, depending on their
+requirements. JEDI replaces the job brokerage functionalities of the Brokerage
+module of the PanDA Server but this module still handles communication between
+the Task Buffer and the Data Service modules.
 %, using it to enact the execution of the jobs defined by JEDI
@@
 % -----------------------------------------------------------------------------
-\subsection{Management and Execution}
+\subsection{Execution}
 \label{ssec:panda_exec}
@@
 This approach introduced inefficiency both with usability and resource
-utilization. Ideally, users should not have to reason in terms of jobs: Users
-conceive analyses in terms of one or more, possibly related tasks; the `job'
-abstraction is required exclusively by the execution middleware, i.e. PanDA.
-Further, a static partitioning of tasks into jobs does not take into account the
-heterogeneity and dynamicity of the resources of the pilots on which each job
-will be executed.
+utilization~\cite{borodin2015unified}. Ideally, users should not have to reason
+in terms of jobs: Users conceive analyses in terms of one or more, possibly
+related tasks; the `job' abstraction is required exclusively by the execution
+middleware, i.e. PanDA. Further, a static partitioning of tasks into jobs does
+not take into account the heterogeneity and dynamicity of the resources of the
+pilots on which each job will be executed.
 
@@
 call the Job Dispatcher module of the PanDA Server, requesting for a job to
-execute \mtnote{Can a pilot request multiple jobs to execute concurrently?}. The
-Job Dispatcher communicates with the Task Buffer, requesting for a job that is
-bound to the site of that pilot and ready to be executed. Task Buffer checks the
-global queue (i.e., the PanDA database) and if such a job is available, it
-passes the job handler to the Job Dispatcher. The Job Dispatcher dispatches the
-job to the PanDA Pilot.
+execute \mtnote{Can a pilot request multiple jobs to execute
+concurrently?}\sergeynote{In principle it should be able to do that, but that's not
+how it is operated. One pilot - one job is a cleaner model in an environment
+where resources are "owned" by the organization that runs pilots. Simplifies
+many things. At least as far as I know.}. The Job Dispatcher communicates with
+the Task Buffer, requesting for a job that is bound to the site of that pilot
+and ready to be executed. Task Buffer checks the global queue (i.e., the PanDA
+database) and if such a job is available, it passes the job handler to the Job
+Dispatcher. The Job Dispatcher dispatches the job to the PanDA Pilot.
 
@@
 \mtnote{Think whether a comparison with other WMS based on the review we did
-makes sense in this context.}
+makes sense in this context.}\sergeynote{You provided a nice description of the
+PanDA system here.}

### 2017-03-08 overall.tex
Ongoing section V

--- 2017-02-13 overall.tex
@@
-\subsection{PanDA Shared Library I/O Performance Impact at OLCF}
+% -----------------------------------------------------------------------------
+% \subsection{Geant4 Tasks Performance Analysis on Titan}
+% \label{ssec:panda_titan}
 
-Athena, the ATLAS framework has a configuration and initialization stage. At
-this stage, the running job is assembled  on the fly from a large number of
-shared libraries.  Also, at this stage, every algorithm and service is being
-configured, at run time, by a corresponding set of Python scripts, which
-results in a large number of read operations accesses to small python scripts, with
-many includes and imports Python calls.
+Currently, 20 instances of PanDA Broker are deployed on a total of 4 DTNs, 5
+instances for each DTN. Each broker controls from 15 to 300 detector simulation
+jobs per submission, for a theoretical maximum concurrent use of 96,000
+cores. Since November 2015, the PanDA Brokers have operated only in backfill
+mode, without a defined time allocation, and running at the lowest priority on
+Titan.
 
-When ATLAS was originally deployed at the OLCF, all shared libraries were
-placed at the Spider 2 Lustre file system. However, it was quickly discovered
-that the I/O pattern at the Athena stage was detrimental to the overall Spider
-2 file system performance. Since Spider 2 is a center-wide file system and
-shared by all OLCF resources and users, this resulted in lower overall
-interactive and metadata performance at OLCF\@. As can be seen in
-Listing~\ref{mdstrace} the metadata I/O activity for ATLAS exhibits a spike
-corresponding to the beginning of the runs before tapering off.
+Figure~\ref{fig:core-hours-utilization} shows Titan core hours used by ATLAS
+from January 2016 to February 2017. During that period, ATLAS consumed a total
+of 73.8 million Titan core hours, for an average of 7M hours a month, with a
+minimum of 3.3M hours in April 2016 and a maximum 14.8M hours in February 2017.
+On February, PanDA Brokers used almost twice as much backfill availability than
+in any other month. This is likely due to the hardware upgrades made to the
+DTNs. The absence of continuous monitoring of the DTNs does not allow for the
+specific identification of the bottleneck but spot measurements of the DTNs load
+indicate that a faster CPU and better networking were likely responsible for the
+improved performance. No relevant code update was made between January and
+February 2017 and PanDA Brokers logs indicated that they were able to respond
+more promptly to backfill availability.
 
-\begin{minipage}{\linewidth}
-\begin{lstlisting}[language=bash,frame=single,basicstyle=\ttfamily\tiny,caption=ATLAS metadata trace,label=mdstrace]
-XK7 Application 9205593
-      39012 RPCs from 300 of 300 nodes
-        ~69291.96 per sec
-          37851 LDLM_ENQUEUE RPCs    ~67229.83 per sec
-                pmin 13us pavg 42us pmax 4983us
-          611 LDLM_CANCEL RPCs    ~1085.24 per sec
-                pmin 10us pavg 16us pmax 32us
-          277 MDS_CLOSE RPCs    ~492.00 per sec
-                pmin 15us pavg 20us pmax 38us
-          154 MDS_READPAGE RPCs    ~273.53 per sec
-                pmin 170us pavg 292us pmax 671us
-          86 MDS_GETXATTR RPCs    ~152.75 per sec
-                pmin 15us pavg 21us pmax 65us
-          30 MDS_GETATTR RPCs    ~53.29 per sec
-                pmin 16us pavg 20us pmax 27us
-          3 MDS_REINT RPCs    ~5.33 per sec
-                pmin 103us pavg 136us pmax 196us
-          Overall times
-                pmin 10us pavg 42us pmax 4983us
+\begin{figure}[htp]
+    \subfloat[]{
+        \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
+        \label{fig:core-hours-utilization}
+    }
 
-XK7 Application 9205355
-      8698 RPCs from 62 of 62 nodes
-        ~15449.13 per sec
-          8445 LDLM_ENQUEUE RPCs    ~14999.76 per sec
-                pmin 16us pavg 41us pmax 780us
-          92 MDS_CLOSE RPCs    ~163.41 per sec
-                pmin 15us pavg 23us pmax 52us
-          55 MDS_READPAGE RPCs    ~97.69 per sec
-                pmin 189us pavg 291us pmax 534us
-          52 LDLM_CANCEL RPCs    ~92.36 per sec
-                pmin 12us pavg 16us pmax 25us
-          41 MDS_GETXATTR RPCs    ~72.82 per sec
-                pmin 16us pavg 20us pmax 27us
-          13 MDS_GETATTR RPCs    ~23.09 per sec
-                pmin 16us pavg 21us pmax 38us
-          Overall times
-                pmin 12us pavg 42us pmax 780us
+    \subfloat[]{
+        \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
+        \label{fig:backfill-utilization}
+    }
+\caption{(a).(b).}
+\end{figure}
 
-\end{lstlisting}
-\end{minipage}
+Investigations of the CPU load of the upgraded DTNs shows an average
+XX\mtnote{ask Danila for data} load. This indicates that further hardware
+upgrades would not be likely to improve significantly the performance of the
+PanDA Brokers. Nonetheless, the current load suggests that the number of brokers
+per DTN could be increased. This would enable the submission of a larger number
+of concurrent jobs to the Titan's PBS queue, allowing for PanDA to consume a
+higher percentage of the overall backfill availability.
 
-As this problem was identified, the OLCF staff has enabled read-only access to
-certain NFS-exported directories from Titan compute nodes. This in turn allowed
-the OLCF staff to install a software package from a Titan login node and have
-it available read-only on a Titan compute node.
+Figure~\ref{fig:backfill-utilization} shows backfill utilization efficiency,
+defined as the fraction of Titan’s total cores available via backfill utilized
+by ATLAS, from January 2016 to February 2017. ATLAS reached 18\% average
+efficiency with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\%
+efficiency on February 2017. The number of total backfill cores available in
+April 2016 was 38.1M and in February 2017 33.1M. This shows that the improvement
+in the efficiency was determine by the different in total backfill availability.
 
-After moving the ATLAS release on Lustre to the NFS-exported read-only
-directories, the metadata contention problem has resolved and a two orders of
-magnitude metadata performance improvement (i.e., \~6,300 seconds on Lustre vs.\
-\~1,500 seconds on NFS) has been observed for ATLAS\@.  Also, for both cases the
-input file is placed on memory disk on a Titan compute nodes and for Lustre the
-validation step took on average 1,320 seconds, while it was 40 seconds on NFS,
-showing a three orders of magnitude performance improvement.
+Figure~\ref{fig:hpc-workload-utilization} shows that between January 2016 and
+February 2017, about XX million detector simulation jobs were completed on
+Titan, for a total of XXXM events processed. This is equivalent to XX\% of the
+total amount of detector simulations performed by ATLAS in the same period of
+time, and XX\% of the total number of events processed. Comparatively, Titan
+contributed between XX\% and XX\% of the total WLCG availability. When
+accounting for the amount of unused backfill availability and the rate of
+improvement of PanDA efficiency, these figures confirms the fundamental role
+that supercomputers' resources can play for the LHC Run 3.
+
+\begin{figure}[htp]
+    \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
+\caption{\mtnote{Placeholder for the diagram/data I asked to Danila and
+Sergey}}
+\label{fig:hpc-workload-utilization}
+\end{figure}
+
+The scalability of PanDA Broker depends on the DTNs' resources. Each PanDA
+Broker can execute 1 job on up to 300 work nodes for a total of 4,800 cores.
+Scaling above this threshold requires instantiating multiple PanDA Brokers on
+the same DTN and, once the DTN resources are saturated, on multiple DTNs. In
+principle, this is not a problem on Titan as DTNs are supposed to be available
+and can be periodically upgraded. Assuming the efficiency reached on February as
+reference, using all the available backfill resources on Titan will require to
+instantiate around 60 PanDA Brokers. It is not clear whether this is within the
+scope of the current DTN capabilities, especially when considering the increased
+staging in and out of files from the DTNs and ATLAS Grid sites.
+
+The current design and architecture of the PanDA Broker proven to be very
+robust. The failure rate of jobs produced by issues with PanDA Broker and PanDA
+in general is XX\%. This confirms the benefits given by reusing most of the code
+base of the PanDA Pilot and adopting RADICAL-SAGA for communicating with Titan's
+PBS batch system and monitoring jobs execution. During XX months of executing
+production jobs, most of the failures were due to: \ldots, \ldots, and \ldots.
+The usual corrective measure were put in place and the progressive decline of
+the failure rate confirms that PanDA Broker reached production-grade stability
+on Titan.\mtnote{I requested figures about failure types and rate to Sergey and
+Danila.}
+
+\begin{itemize}
+    \item Comparison hardware performance Titan/Grid
+    \item problem with lustre addressed with ramdisk
+    \item Summary overall figure of performance analogous to those produced with NGE. Diagram as discussed in F2F meeting at Rutgers.
+    \item Competition for cache in AMD processors.
+\end{itemize}
+
+Currently, two main parameters measure the performance of the detector
+simulation jobs submitted to Titan: (i) the time taken to setup
+AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
+GEANT4 simulation toolkit; and (ii) the distribution of the time taken by the
+Geant4 toolkit to simulate a certain number of events.
+
+AthenaMP has an initialization and configuration stage. At initialization time,
+AthenaMP is  assembled from a large number of shared libraries. Once
+bootstrapped, every algorithm and service is configured by a set of Python
+scripts. Both these operations result in a large number of read operations,
+including those required to  access of small python scripts.
+
+% AthenaMP is a multipurpose framework that needs to be configured depending on
+% the type of payalod that needs to be executed. For Geant4, this configuration
+% process links up to 200 libraries, all requiring filesystem read and write
+% operations.
+
+% The execution of Geant4 is mostly compute-intensive requiring to write around
+% XXKB per 100 events on disk and no more than 2GB of memory.
+
+% -----------------------------------------------------------------------------
+% \subsection{PanDA Shared Library I/O Performance Impact at OLCF}
+
+% Athena, the ATLAS framework has a configuration and initialization stage. At
+% this stage, the running job is assembled on the fly from a large number of
+% shared libraries. Also, at this stage, every algorithm and service is being
+% configured, at run time, by a corresponding set of Python scripts, which
+% results in a large number of read operations accesses to small python
+% scripts, with many includes and imports Python calls.
+
+Originally, all shared libraries of AthenaMP and the python scripts for job
+configuration were stored on the Spider 2 Lustre file system. However, the I/O
+patterns of the initialization and configuration stages degraded the performance
+of the filesystem. Since Spider 2 is a center-wide file system, shared by all
+OLCF resources and users, this resulted in lower overall interactive and
+metadata performance at OLCF\@. \mtnote{I am afraid the details about the trace
+are too specific given the space constraints of a SC submission. Please feel
+free to uncomment it if you disagree.}
+% As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
+% exhibits a spike corresponding to the beginning of the runs before tapering
+% off.
+
+% \begin{minipage}{\linewidth}
+% \begin{lstlisting}[language=bash,frame=single,basicstyle=\ttfamily\tiny,caption=ATLAS metadata trace,label=mdstrace]
+% XK7 Application 9205593
+%       39012 RPCs from 300 of 300 nodes
+%         ~69291.96 per sec
+%           37851 LDLM_ENQUEUE RPCs    ~67229.83 per sec
+%                 pmin 13us pavg 42us pmax 4983us
+%           611 LDLM_CANCEL RPCs    ~1085.24 per sec
+%                 pmin 10us pavg 16us pmax 32us
+%           277 MDS_CLOSE RPCs    ~492.00 per sec
+%                 pmin 15us pavg 20us pmax 38us
+%           154 MDS_READPAGE RPCs    ~273.53 per sec
+%                 pmin 170us pavg 292us pmax 671us
+%           86 MDS_GETXATTR RPCs    ~152.75 per sec
+%                 pmin 15us pavg 21us pmax 65us
+%           30 MDS_GETATTR RPCs    ~53.29 per sec
+%                 pmin 16us pavg 20us pmax 27us
+%           3 MDS_REINT RPCs    ~5.33 per sec
+%                 pmin 103us pavg 136us pmax 196us
+%           Overall times
+%                 pmin 10us pavg 42us pmax 4983us
+%
+% XK7 Application 9205355
+%       8698 RPCs from 62 of 62 nodes
+%         ~15449.13 per sec
+%           8445 LDLM_ENQUEUE RPCs    ~14999.76 per sec
+%                 pmin 16us pavg 41us pmax 780us
+%           92 MDS_CLOSE RPCs    ~163.41 per sec
+%                 pmin 15us pavg 23us pmax 52us
+%           55 MDS_READPAGE RPCs    ~97.69 per sec
+%                 pmin 189us pavg 291us pmax 534us
+%           52 LDLM_CANCEL RPCs    ~92.36 per sec
+%                 pmin 12us pavg 16us pmax 25us
+%           41 MDS_GETXATTR RPCs    ~72.82 per sec
+%                 pmin 16us pavg 20us pmax 27us
+%           13 MDS_GETATTR RPCs    ~23.09 per sec
+%                 pmin 16us pavg 21us pmax 38us
+%           Overall times
+%                 pmin 12us pavg 42us pmax 780us
+%
+% \end{lstlisting}
+% \end{minipage}
+
+% As this problem was identified, the OLCF staff has enabled read-only access to
+% certain NFS-exported directories from Titan compute nodes. This in turn
+% allowed the OLCF staff to install a software package from a Titan login node
+% and have it available read-only on a Titan compute node.
+
+This issue was addressed by moving the AthenaMP distribution to a read-only NFS
+directory made accessible from the Titan work nodes. This eliminated the
+metadata contention problem improving metadata read performance of two orders of
+magnitude: from \~6,300 seconds on Lustre to \~1,500 seconds on NFS. Further, at
+configuration time, the input files of each job were stored to a ramdisk on the
+work nodes. This offered three orders of magnitude performance improvement in
+the file validation step: from 1,320 seconds on Lustre to 40 seconds on NFS.
 
@@
 
+% -----------------------------------------------------------------------------
 \subsection{PanDA I/O Impact at OLCF}
@@
 As can be seen from Table~\ref{panda-olcf-stats}, the number of nodes used by
-ATLAS jobs vary between 1 and 300, while the average is at 35. 75\% of the
-ATLAS jobs consume less than 25 and 92\% consume less than or equal to 100
-Titan compute nodes. During the 174 hours of data collection, we observed that
-6.75 ATLAS jobs were executed on average per hour on Titan and ran for
-1.74 hours on average.
+ATLAS jobs vary between 1 and 300, while the average is at 35. 75\% of the ATLAS
+jobs consume less than 25 and 92\% consume less than or equal to 100 Titan
+compute nodes. During the 174 hours of data collection, we observed that 6.75
+ATLAS jobs were executed on average per hour on Titan and ran for 1.74 hours on
+average.
 
@@
 {fig:atlas-titan-io-written}, the distribution between read and written amount
-of data per job is quite different from one another. The read operation distribution
-per job shows a long tail, ranging from 12.5 GB to 250 GB, while the written
-amount of data has a very narrow distribution.
+of data per job is quite different from one another. The read operation
+distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
+the written amount of data has a very narrow distribution.
 
@@
 Overall, based on our experiments with real-world jobs, it can be safely
-concluded that the file and metadata I/O load of ATLAS PanDA project on the
-OLCF Titan supercomputing environment and the Spider 2 file system is not
-detrimental to the center operations and the overall impact is minimal, at the
-current scale of the project.
+concluded that the file and metadata I/O load of ATLAS PanDA project on the OLCF
+Titan supercomputing environment and the Spider 2 file system is not detrimental
+to the center operations and the overall impact is minimal, at the current scale
+of the project.

### 2017-03-09 overall.tex
Iteration section V

@@
-% -----------------------------------------------------------------------------
-% \subsection{Geant4 Tasks Performance Analysis on Titan}
-% \label{ssec:panda_titan}
-
 Currently, 20 instances of PanDA Broker are deployed on a total of 4 DTNs, 5
 instances for each DTN. Each broker controls from 15 to 300 detector simulation
-jobs per submission, for a theoretical maximum concurrent use of 96,000
-cores. Since November 2015, the PanDA Brokers have operated only in backfill
-mode, without a defined time allocation, and running at the lowest priority on
-Titan.
+jobs per submission, one simulation job for each Titan's work node and a
+theoretical maximum concurrent use of 96,000 cores. Since November 2015, the
+PanDA Brokers have operated only in backfill mode, without a defined time
+allocation, and running at the lowest priority on Titan.
 
-Figure~\ref{fig:core-hours-utilization} shows Titan core hours used by ATLAS
-from January 2016 to February 2017. During that period, ATLAS consumed a total
-of 73.8 million Titan core hours, for an average of 7M hours a month, with a
+We evaluated the efficiency, scalability and reliability of the PanDA Brokers
+deployed on Titan along two dimensions: (i) amount of backfill availability
+utilized by the PanDA Broker; and (ii) amount of jobs walltime spent performing
+detector simulations. We based these evaluations on measuring the number of
+cores utilized by ATLAS on Titan, the overall backfill availability on Titan,
+the number of detector simulations performed on Titan and on all the resources
+available to ATLAS, and the failure rate of these simulations. All the
+measurements were performed between January 2016 and February 2017.
+
+% -----------------------------------------------------------------------------
+\subsection{Efficiency and Scalability of PanDA Broker on Titan}
+\label{ssec:panda_titan}
+
+Figure~\ref{fig:core-hours-utilization} shows the number of core hours used by
+ATLAS on Titan, from January 2016 to February 2017. During that period, ATLAS
+consumed a total of 73.8M core hours, for an average of 7M hours a month, with a
 minimum of 3.3M hours in April 2016 and a maximum 14.8M hours in February 2017.
-On February, PanDA Brokers used almost twice as much backfill availability than
-in any other month. This is likely due to the hardware upgrades made to the
-DTNs. The absence of continuous monitoring of the DTNs does not allow for the
-specific identification of the bottleneck but spot measurements of the DTNs load
-indicate that a faster CPU and better networking were likely responsible for the
-improved performance. No relevant code update was made between January and
-February 2017 and PanDA Brokers logs indicated that they were able to respond
-more promptly to backfill availability.
+
+% \begin{figure}[htp]
+%     \subfloat[]{
+%         \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
+%         \label{fig:core-hours-utilization}
+%     }
+%
+%     \subfloat[]{
+%         \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
+%         \label{fig:backfill-utilization}
+%     }
+% \caption{(a).(b).}
+% \end{figure}
 
 \begin{figure}[htp]
-    \subfloat[]{
 \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
+\caption{Caption.}
 \label{fig:core-hours-utilization}
-    }
-
-    \subfloat[]{
-        \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
-        \label{fig:backfill-utilization}
-    }
-\caption{(a).(b).}
 \end{figure}
 
+On February 2017, PanDA Brokers used almost twice as much backfill availability
+than in any other month. No relevant code update was made during that period and
+logs indicated that the brokers were able to respond more promptly to backfill
+availability. This is likely due to hardware upgrades on the DTNs. The absence
+of continuous monitoring of the DTNs does not allow to quantify bottlenecks but
+spot measurements of the DTNs load indicate that a faster CPU and better
+networking were likely responsible for the improved performance.
+
 Investigations of the CPU load of the upgraded DTNs shows an average
-XX\mtnote{ask Danila for data} load. This indicates that further hardware
-upgrades would not be likely to improve significantly the performance of the
-PanDA Brokers. Nonetheless, the current load suggests that the number of brokers
-per DTN could be increased. This would enable the submission of a larger number
-of concurrent jobs to the Titan's PBS queue, allowing for PanDA to consume a
-higher percentage of the overall backfill availability.
+XX\mtnote{asked Danila for data} load. As such, further hardware upgrades seem
+unlikely to improve significantly the performance of PanDA Brokers. Nonetheless,
+the current load suggests that the number of brokers per DTN could be increased.
+This would enable the submission of a larger number of concurrent jobs to
+Titan's PBS queue, allowing for PanDA to consume a higher percentage of
+backfill availability.
 
-Figure~\ref{fig:backfill-utilization} shows backfill utilization efficiency,
-defined as the fraction of Titan’s total cores available via backfill utilized
-by ATLAS, from January 2016 to February 2017. ATLAS reached 18\% average
-efficiency with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\%
-efficiency on February 2017. The number of total backfill cores available in
-April 2016 was 38.1M and in February 2017 33.1M. This shows that the improvement
-in the efficiency was determine by the different in total backfill availability.
+Figure~\ref{fig:backfill-utilization} shows backfill utilization efficiency from
+January 2016 to February 2017. Efficiency is defined as the fraction of Titan’s
+cores available via backfill that was utilized by the PanDA Brokers. The brokers
+reached 18\% average efficiency with a minimum 8.9\% efficiency on April 2016
+and a maximum 33.5\% efficiency on February 2017. The number of total backfill
+cores available was 38.1M in April 2016, and 33.1M in February 2017. This shows
+that the difference in total backfill availability was not responsible for the measured difference in efficiency.
+
+\begin{figure}[htp]
+\includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
+\caption{Caption.}
+\label{fig:backfill-utilization}
+\end{figure}
 
 Figure~\ref{fig:hpc-workload-utilization} shows that between January 2016 and
-February 2017, about XX million detector simulation jobs were completed on
-Titan, for a total of XXXM events processed. This is equivalent to XX\% of the
-total amount of detector simulations performed by ATLAS in the same period of
-time, and XX\% of the total number of events processed. Comparatively, Titan
-contributed between XX\% and XX\% of the total WLCG availability. When
-accounting for the amount of unused backfill availability and the rate of
-improvement of PanDA efficiency, these figures confirms the fundamental role
-that supercomputers' resources can play for the LHC Run 3.
+February 2017, about 2.25M million detector simulation jobs were completed on
+Titan, for a total of 225M events processed. This is equivalent to 9\% of all
+the 250M detector simulations performed by ATLAS in the same period of time, and
+3.5\% of the 6.6B events processed by those jobs. Comparatively, Titan
+contributed 3.9\% of the total of around 200K cores available to ATLAS on Grid,
+Cloud, and HPC infrastructures together. When accounting for the amount of
+unused backfill availability and the rate of improvement of PanDA efficiency on
+Titan, these figures confirms the relevance of supercomputers' resource
+contribution to the LHC Run 3.
 
@@
 \caption{\mtnote{Placeholder for the diagram/data I asked to Danila and
-Sergey}}
+Sergey. Consider a table instead of a figure.}}
 \label{fig:hpc-workload-utilization}
@@
 
-The scalability of PanDA Broker depends on the DTNs' resources. Each PanDA
-Broker can execute 1 job on up to 300 work nodes for a total of 4,800 cores.
-Scaling above this threshold requires instantiating multiple PanDA Brokers on
-the same DTN and, once the DTN resources are saturated, on multiple DTNs. In
-principle, this is not a problem on Titan as DTNs are supposed to be available
-and can be periodically upgraded. Assuming the efficiency reached on February as
-reference, using all the available backfill resources on Titan will require to
-instantiate around 60 PanDA Brokers. It is not clear whether this is within the
-scope of the current DTN capabilities, especially when considering the increased
-staging in and out of files from the DTNs and ATLAS Grid sites.
+Detector simulations executed on Titan, all process 100 events. This number has
+been derived empirically by observing the average walltime of backfill
+resources: too many events would \ldots\mtnote{I need to speak further to Sergey
+to clarify this.}. Usually, detector simulations performed on the Grid process a
+higher number of events than those performed on Titan. This explains the
+difference between the percentage of detector simulations performed on Titan
+(9\%) and of events computed on Titan for those simulations (3.9\%). PanDA Event
+Service~\cite{panda-event-service} would address this limitation: each PanDA
+Broker would be able to pull a dynamic number of events for each simulation,
+depending on the available walltime. This functionality is currently being
+implemented.
 
-The current design and architecture of the PanDA Broker proven to be very
-robust. The failure rate of jobs produced by issues with PanDA Broker and PanDA
-in general is XX\%. This confirms the benefits given by reusing most of the code
-base of the PanDA Pilot and adopting RADICAL-SAGA for communicating with Titan's
-PBS batch system and monitoring jobs execution. During XX months of executing
-production jobs, most of the failures were due to: \ldots, \ldots, and \ldots.
-The usual corrective measure were put in place and the progressive decline of
-the failure rate confirms that PanDA Broker reached production-grade stability
-on Titan.\mtnote{I requested figures about failure types and rate to Sergey and
-Danila.}
+Given the current design and implementation, the utilization of a larger portion
+of Titan's backfill availability depends on the the number of concurrent PanDA
+Brokers instantiated on the DTNs.
+% The number of concurrent detector simulation jobs that PanDA Broker can manage
+% depends on the DTNs' resources. Each PanDA Broker can execute 1 MPI PBS job on
+% up to 300 work nodes for a total of 4,800 cores. Scaling above this threshold
+% requires instantiating multiple PanDA Brokers on the same DTN and, once the
+% DTN resources are saturated, on multiple DTNs.
+In principle, this is not a problem on Titan as multiple DTNs are supposed to be
+available and periodically upgraded. Assuming the efficiency reached on February
+as reference, using all the available backfill resources on Titan would require
+around 60 concurrent PanDA Brokers. Further testing and better monitoring
+capabilities are required to establish whether the current DTN capabilities can
+support this load, especially when considering the increased volume of files
+staged in and out the DTNs.
+
+% -----------------------------------------------------------------------------
+\subsection{Efficiency and Scalability of Detector Simulation on Titan}
+\label{ssec:panda_titan}
 
@@
 
+% ----------------------------------------------------------------------------
+\subsection{Reliability of PanDA Broker and Detector Simulations on Titan}
+\label{ssec:reliability}
+
+The current design and architecture of the PanDA Broker is proving to be as
+reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
+the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
+while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
+Brokers were responsible for around the 19\% of the failures, compared to the
+29\% of failures produced by the JobDispatcher module of the PanDA Server, and
+the 13\% failures produced by Geant4. The current failure rate confirms the
+benefits given by reusing most of the code base of the PanDA Pilot for
+implementing the PanDA Broker. It also shows that adopting third-party libraries
+like RADICAL-SAGA did not have a measurable adverse effect on reliability.
+
+Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
+recorded via the PanDA Brokers between January 2016 and February 2017. The
+initial preponderance of unknown failures was progressively addressed by a more
+accurate failure mode developed for the PanDA Broker and Titan. It also shows
+the progressive improvement of the broker's reliability: In Februrary 2017, most
+of the failures were produced by the initialization and configuration phases of
+AthenaMP or by Geant4 failed executions.\mtnote{If we want to keep this diagram,
+we will need to plot it from the CSV to make it readable.}
+
+\begin{figure}[htp]
+    \includegraphics[clip,width=\columnwidth]{figures/failures_titan.png}
+\caption{Caption.}
+\label{fig:failures-titan}
+\end{figure}
+
 % -----------------------------------------------------------------------------

### 2017-03-10 overall.tex
Iteration V

@@
 \begin{itemize}
+    \item problem with lustre addressed with ramdisk
+    \item Competition for cache in AMD processors.
+    \item Summary overall figure of performance analogous to those produced with NGE. Diagram as discussed in F2F meeting at Rutgers.
     \item Comparison hardware performance Titan/Grid
-    \item problem with lustre addressed with ramdisk
-    \item Summary overall figure of performance analogous to those produced with NGE. Diagram as discussed in F2F meeting at Rutgers.
-    \item Competition for cache in AMD processors.
 \end{itemize}
 
-Currently, two main parameters measure the performance of the detector
-simulation jobs submitted to Titan: (i) the time taken to setup
+We use two main parameters to measure the performance of the detector simulation
+jobs submitted to Titan: (i) the time taken to setup
 AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
@@
 AthenaMP has an initialization and configuration stage. At initialization time,
-AthenaMP is  assembled from a large number of shared libraries. Once
-bootstrapped, every algorithm and service is configured by a set of Python
-scripts. Both these operations result in a large number of read operations,
-including those required to  access of small python scripts.
+AthenaMP is assembled from a large number of shared libraries. Once initialized,
+every algorithm and service of AthenaMP is configured by a set of Python
+scripts. Both these operations result in a large number of read operations on
+the filesystem, including those required to access of small python scripts.
 
@@
 
-Originally, all shared libraries of AthenaMP and the python scripts for job
-configuration were stored on the Spider 2 Lustre file system. However, the I/O
-patterns of the initialization and configuration stages degraded the performance
-of the filesystem. Since Spider 2 is a center-wide file system, shared by all
-OLCF resources and users, this resulted in lower overall interactive and
-metadata performance at OLCF\@. \mtnote{I am afraid the details about the trace
-are too specific given the space constraints of a SC submission. Please feel
-free to uncomment it if you disagree.}
+Initially, all shared libraries of AthenaMP and the configuration python scripts
+were stored on the Spider 2 Lustre file system. However, the I/O patterns of the
+initialization and configuration stages degraded the performance of the
+filesystem (Figure~\ref{}). Since Spider 2 is a center-wide file system, this
+resulted in lower overall interactive and metadata performance for all OLCF
+resources and users. \mtnote{I am afraid the details about the trace are too
+specific given the space constraints of a SC submission. Please feel free to
+uncomment it if you disagree.}
 % As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
@@
 This issue was addressed by moving the AthenaMP distribution to a read-only NFS
-directory made accessible from the Titan work nodes. This eliminated the
-metadata contention problem improving metadata read performance of two orders of
-magnitude: from \~6,300 seconds on Lustre to \~1,500 seconds on NFS. Further, at
-configuration time, the input files of each job were stored to a ramdisk on the
-work nodes. This offered three orders of magnitude performance improvement in
-the file validation step: from 1,320 seconds on Lustre to 40 seconds on NFS.
+directory made accessible from the Titan work nodes. This eliminated the problem
+of metadata contention, improving metadata read performance from \~6,300 seconds
+on Lustre to \~1,500 seconds on NFS. Further, at configuration time, the input
+files of each detector simulation job were stored in a ramdisk on the work
+nodes. This offered a marked performance improvement: from 1,320 seconds on
+Lustre to 40 seconds on NFS.
 
@@
 %\end{figure}
+
+At runtime, detector simulation jobs showed performance degradation when all the
+16 cores of Titan's work nodes were used. As shows in Figure~\ref{}, the mean
+runtime for 8 concurrent jobs computing 100 events is XXXs, with each job
+executed in 8 distinct processes, one per core. The mean runtime for 16
+concurrent jobs computing 100 events is instead XXXs. This can be explained by
+looking at the architecture of the cache in the AMD Opteron 6274 CPUs. This
+processor shares the level 1 instruction cache between couples of cores,
+impacting the performance of \ldots computation.
+
+The compute performance of Titan for detector simulation jobs is, on average,
+lower that that of LCWG. Figure~\ref{} shows the difference in execution time of
+the same type of simulation between Titan and \ldots. This is not an issue with
+the design of PanDA Brokers and, more in general, with the migration of part of
+the ATLAS workload to supercomputers: Titan has been active since 2012 and it is
+to be expected that more modern CPUs offer better performance. A pilot project
+is already in place to test PanDA Broker on Summit, the next generation
+supercomputer that will become active in 2018.
 

### 2017-03-11 overall.tex
Section V ready for comments

@@
-Currently, 20 instances of PanDA Broker are deployed on a total of 4 DTNs, 5
-instances for each DTN. Each broker controls from 15 to 300 detector simulation
-jobs per submission, one simulation job for each Titan's work node and a
-theoretical maximum concurrent use of 96,000 cores. Since November 2015, the
-PanDA Brokers have operated only in backfill mode, without a defined time
-allocation, and running at the lowest priority on Titan.
+Currently, ATLAS deploys 20 instances of PanDA Broker on 4 Titan's DTNs, 5
+instances per DTN. Each broker submits and manages the execution of 15 to 300
+detector simulation jobs, one job for each work node, and a theoretical maximum
+concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have operated
+only in backfill mode, without a defined time allocation, and running at the
+lowest priority on Titan.
 
 We evaluated the efficiency, scalability and reliability of the PanDA Brokers
-deployed on Titan along two dimensions: (i) amount of backfill availability
-utilized by the PanDA Broker; and (ii) amount of jobs walltime spent performing
-detector simulations. We based these evaluations on measuring the number of
-cores utilized by ATLAS on Titan, the overall backfill availability on Titan,
-the number of detector simulations performed on Titan and on all the resources
-available to ATLAS, and the failure rate of these simulations. All the
-measurements were performed between January 2016 and February 2017.
+along two dimensions: (i) the amount of backfill availability utilized by the
+PanDA Brokers; and (ii) the amount of jobs walltime spent performing detector
+simulations. We based these evaluations on measuring the number of cores
+utilized by ATLAS on Titan, the overall backfill availability, the number of
+detector simulations performed on all the resources available to ATLAS and on
+Titan, and the failure rate of these simulations. All the measurements were
+performed between January 2016 and February 2017.
 
@@
 availability. This is likely due to hardware upgrades on the DTNs. The absence
-of continuous monitoring of the DTNs does not allow to quantify bottlenecks but
+of continuous monitoring of the nodes does not allow to quantify bottlenecks but
 spot measurements of the DTNs load indicate that a faster CPU and better
@@
 
-Investigations of the CPU load of the upgraded DTNs shows an average
-XX\mtnote{asked Danila for data} load. As such, further hardware upgrades seem
-unlikely to improve significantly the performance of PanDA Brokers. Nonetheless,
-the current load suggests that the number of brokers per DTN could be increased.
-This would enable the submission of a larger number of concurrent jobs to
-Titan's PBS queue, allowing for PanDA to consume a higher percentage of
-backfill availability.
+Investigations showed an average CPU load of 3.6\% on the upgraded DTNs. As
+such, further hardware upgrades seem unlikely to improve significantly the
+performance of PanDA Brokers. Nonetheless, the current load suggests that the
+number of brokers per DTN could be increased. This would enable the submission
+of a larger number of concurrent jobs to Titan's PBS queue, allowing for PanDA
+to consume a higher percentage of backfill availability.
 
@@
 cores available was 38.1M in April 2016, and 33.1M in February 2017. This shows
-that the difference in total backfill availability was not responsible for the measured difference in efficiency.
+that the measured difference in efficiency did not depend on a comparable
+difference in total backfill availability.
 
@@
 contributed 3.9\% of the total of around 200K cores available to ATLAS on Grid,
-Cloud, and HPC infrastructures together. When accounting for the amount of
-unused backfill availability and the rate of improvement of PanDA efficiency on
-Titan, these figures confirms the relevance of supercomputers' resource
-contribution to the LHC Run 3.
+Cloud, and HPC infrastructures together. These figures confirms the relevance of
+supercomputers' resource contribution to the LHC Run 3, especially when
+accounting for the amount of unused backfill availability and the rate of
+improvement of PanDA efficiency.
 
@@
 
-Detector simulations executed on Titan, all process 100 events. This number has
-been derived empirically by observing the average walltime of backfill
-resources: too many events would \ldots\mtnote{I need to speak further to Sergey
-to clarify this.}. Usually, detector simulations performed on the Grid process a
-higher number of events than those performed on Titan. This explains the
-difference between the percentage of detector simulations performed on Titan
-(9\%) and of events computed on Titan for those simulations (3.9\%). PanDA Event
-Service~\cite{panda-event-service} would address this limitation: each PanDA
-Broker would be able to pull a dynamic number of events for each simulation,
-depending on the available walltime. This functionality is currently being
-implemented.
+All the detector simulations executed on Titan process 100 events. This number
+has been chosen by observing the average walltime of backfill resources: too
+many events would \ldots\mtnote{I need to speak further to Sergey to clarify
+this.}.
+
+Usually, detector simulations performed on the Grid process a higher number of
+events than those on Titan. This explains the difference between the percentage
+of detector simulations (9\%) and of events computed by those simulations
+(3.9\%) on Titan. PanDA Event Service~\cite{panda-event-service} will contribute
+to address this limitation: each PanDA Broker will be able to pull a variable
+number of events for each simulation, depending on the available walltime. This
+functionality is currently being implemented.
 
@@
 AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
-GEANT4 simulation toolkit; and (ii) the distribution of the time taken by the
-Geant4 toolkit to simulate a certain number of events.
+GEANT4 simulation toolkit~\cite{agostinelli2003geant4}; and (ii) the
+distribution of the time taken by the Geant4 toolkit to simulate a certain
+number of events.
 
 AthenaMP has an initialization and configuration stage. At initialization time,
-AthenaMP is assembled from a large number of shared libraries. Once initialized,
-every algorithm and service of AthenaMP is configured by a set of Python
-scripts. Both these operations result in a large number of read operations on
-the filesystem, including those required to access of small python scripts.
+AthenaMP is assembled from a large number of shared libraries, depending on the
+type of payload that will have to be computed. Once initialized, every algorithm
+and service of AthenaMP is configured by a set of Python scripts. Both these
+operations result in a large number of read operations on the filesystem,
+including those required to access of small python scripts.
 
@@
 
-Initially, all shared libraries of AthenaMP and the configuration python scripts
-were stored on the Spider 2 Lustre file system. However, the I/O patterns of the
-initialization and configuration stages degraded the performance of the
-filesystem (Figure~\ref{}). Since Spider 2 is a center-wide file system, this
-resulted in lower overall interactive and metadata performance for all OLCF
+Initially, all the shared libraries of AthenaMP and the python scripts for the
+configuration stage were stored on the Spider 2 Lustre file system. However, the
+I/O patterns of the initialization and configuration stages degraded the
+performance of the filesystem (Figure~\ref{}\mtnote{We may want to aggregate
+some of the diagrams about lustre's experiments here}). Since Spider 2 is a
+center-wide file system, this resulted in  performance degradation for all OLCF
 resources and users. \mtnote{I am afraid the details about the trace are too
@@
 
-This issue was addressed by moving the AthenaMP distribution to a read-only NFS
-directory made accessible from the Titan work nodes. This eliminated the problem
-of metadata contention, improving metadata read performance from \~6,300 seconds
-on Lustre to \~1,500 seconds on NFS. Further, at configuration time, the input
-files of each detector simulation job were stored in a ramdisk on the work
-nodes. This offered a marked performance improvement: from 1,320 seconds on
+After careful characterization of the impact of ATLAS jobs on Lustre, the
+performance degradation was addressed by moving the AthenaMP distribution to a
+read-only NFS directory, shared among DTNs and work nodes. This eliminated the
+problem of metadata contention, improving metadata read performance from \~6,300
+seconds on Lustre to \~1,500 seconds on NFS. Further, at configuration time, the
+input files of each detector simulation job were stored to a ramdisk on the work
+nodes. This offered a marked improvement of reading time: from 1,320 seconds on
 Lustre to 40 seconds on NFS.
@@
 At runtime, detector simulation jobs showed performance degradation when all the
-16 cores of Titan's work nodes were used. As shows in Figure~\ref{}, the mean
-runtime for 8 concurrent jobs computing 100 events is XXXs, with each job
-executed in 8 distinct processes, one per core. The mean runtime for 16
-concurrent jobs computing 100 events is instead XXXs. This can be explained by
-looking at the architecture of the cache in the AMD Opteron 6274 CPUs. This
-processor shares the level 1 instruction cache between couples of cores,
-impacting the performance of \ldots computation.
+16 cores of Titan's work nodes were used. As shows in
+Figure~\ref{}\mtnote{Sergery: do you have a diagram showing the difference in
+performance between using 16 and 8 cores for a detector simulation job with 100
+events?}, the mean runtime for 8 concurrent jobs computing 100 events is XXXs,
+with each job executed in 8 distinct processes, one per core. The mean runtime
+for 16 concurrent jobs computing 100 events is instead XXXs. This can be
+explained by looking at the cache architecture of the AMD Opteron 6274 CPU. In
+this processor, every two cores share the level 1 instruction cache, impacting
+the performance of \ldots\mtnote{Do we kow what instruction set is affected?}
+computation.
 
 The compute performance of Titan for detector simulation jobs is, on average,
-lower that that of LCWG. Figure~\ref{} shows the difference in execution time of
-the same type of simulation between Titan and \ldots. This is not an issue with
+lower that that of LCWG. Figure~\ref{}\mtnote{Sergey: I remember you showed us
+some plots comparing Geant4 simulation's performance on Titan and some of the
+Grid sites. Could we use one or two of the here?} shows the difference in
+execution time of the same type of simulation between Titan and
+\ldots\mtnote{Add the name of the grid sites here}. This is not an issue with
 the design of PanDA Brokers and, more in general, with the migration of part of
-the ATLAS workload to supercomputers: Titan has been active since 2012 and it is
-to be expected that more modern CPUs offer better performance. A pilot project
-is already in place to test PanDA Broker on Summit, the next generation
-supercomputer that will become active in 2018.
+the ATLAS workload to supercomputers: Titan has been active since 2012 and more
+modern CPUs are expected to offer better performance. A pilot project is already
+in place to test PanDA Broker on Summit, the next generation supercomputer that
+will become active in 2018 at OLCF.
 
@@
 29\% of failures produced by the JobDispatcher module of the PanDA Server, and
-the 13\% failures produced by Geant4. The current failure rate confirms the
-benefits given by reusing most of the code base of the PanDA Pilot for
-implementing the PanDA Broker. It also shows that adopting third-party libraries
-like RADICAL-SAGA did not have a measurable adverse effect on reliability.
+the 13\% failures produced by the Geant4 toolkit. The current failure rate of
+the PanDA Brokers confirms the benefits of reusing most of the code base of the
+PanDA Pilot for implementing the PanDA Broker. It also shows that adopting
+third-party libraries like RADICAL-SAGA did not have a measurable adverse effect
+on reliability.
 
 Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
-recorded via the PanDA Brokers between January 2016 and February 2017. The
+recorded by the PanDA Brokers between January 2016 and February 2017. The
 initial preponderance of unknown failures was progressively addressed by a more
-accurate failure mode developed for the PanDA Broker and Titan. It also shows
-the progressive improvement of the broker's reliability: In Februrary 2017, most
-of the failures were produced by the initialization and configuration phases of
-AthenaMP or by Geant4 failed executions.\mtnote{If we want to keep this diagram,
-we will need to plot it from the CSV to make it readable.}
+accurate failure model, specifically tailored to the PanDA Broker and Titan. It
+also shows the progressive improvement of the broker's reliability: In Februrary
+2017, most of the failures were produced by the initialization and configuration
+phases of AthenaMP or by failed executions of the payload.\mtnote{If we want to
+keep this diagram, we will need to plot it from the CSV to make it readable.}
 

### 2017-03-11 overall.tex
Notes on backfill dynamism

@@
 staged in and out the DTNs.
+
+\mtnote{Extend/change previous paragraph explaining also the problem of the
+`spike'/`wave' of backfill availability. The problem without enough brokers to
+cope with variable amounts of backfill availability is that we may exhaust all
+20 pilots to use a portion of backfill smaller than the one that becomes
+available when no more brokers are available.}
 

### 2017-03-11 overall.tex
typo

--- 2017-03-11 overall.tex
@@
 staged in and out the DTNs.
+
+\mtnote{Extend/change previous paragraph explaining also the problem of the
+`spike'/`wave' of backfill availability. The problem without enough brokers to
+cope with variable amounts of backfill availability is that we may exhaust all
+20 brokers to use a portion of backfill smaller than the one that becomes
+available when no more brokers are available.}
 

### 2017-03-11 nge.tex
Stud: connection V to VI

As seen in \S\ref{sec:panda_deployment}, PanDA Broker has been designed to
submit PBS jobs to Titan's batch system instead of pilots due to the absence of
wide area network connectivity on Titan's work nodes. As discussed in
\S\ref{sec:panda_titan}, the unavailability of pilots imposes three main
limitations: (i) static coupling between PBS jobs and detector simulations and
(ii) support for a single type of payload.

The static coupling between PBS jobs and detector simulations makes impossible
to schedule multiple generations of workload on the same PBS job. Specifically,
once a number of detector simulations are packaged into a PBS job and this job
is queued on Titan, no further simulations can be added to that job. New
simulations have to be packaged into a new PBS job that need to be submitted to
Titan on the base of the backfill availability of that moment.

The support of workload generations would enable a more efficient use of the
backfill availability walltime. Currently, when a set of simulations ends also
the PBS ends, independently on whether more walltime would still be available.
With a pilot, more simulations could be executed so to utilize all the walltime
while avoiding further job packaging and submission overheads.\mtnote{more?}

Pilots offer a general scheduling interface while hiding the mechanics of
coordinating resources across multiple work nodes. This eliminates the need of
packaging payload into PBS jobs within the broker, greatly simplifying the
submission process. This simplification would make it possible to submit
different types of payload without having to develop a specific PBS script for
every payload.

### 2017-03-11 nge.tex
Transition V to VI

@@
-As seen in \S\ref{sec:panda_deployment}, PanDA Broker has been designed to
-submit PBS jobs to Titan's batch system instead of pilots due to the absence of
-wide area network connectivity on Titan's work nodes. As discussed in
-\S\ref{sec:panda_titan}, the unavailability of pilots imposes three main
-limitations: (i) static coupling between PBS jobs and detector simulations and
-(ii) support for a single type of payload.
+As seen in \S\ref{sec:panda_deployment}, the absence of wide area network
+connectivity on Titan's work nodes was one of the main reasons for PanDA Broker
+not to implement pilots. As discussed in \S\ref{sec:panda_titan}, the absence of
+pilots imposes the static coupling between PBS jobs and detector simulations,
+and development tailored to each type of payload that needs to be executed.
 
@@
 backfill availability walltime. Currently, when a set of simulations ends also
-the PBS ends, independently on whether more walltime would still be available.
-With a pilot, more simulations could be executed so to utilize all the walltime
-while avoiding further job packaging and submission overheads.\mtnote{more?}
+the PBS job ends, independently on whether more walltime would still be
+available. With a pilot, more simulations could be executed so to utilize all
+the available walltime, while avoiding further job packaging and submission
+overheads.\mtnote{Are there more reasons for wanting multiple generations?}
 
-Pilots offer a general scheduling interface while hiding the mechanics of
-coordinating resources across multiple work nodes. This eliminates the need of
-packaging payload into PBS jobs within the broker, greatly simplifying the
-submission process. This simplification would make it possible to submit
-different types of payload without having to develop a specific PBS script for
-every payload.
+Pilots can offer a payload-independent scheduling interface while hiding the
+mechanics of coordination and communication among multiple work nodes. This
+could eliminate the need for packaging payload into PBS jobs within the broker,
+greatly simplifying the submission process. This simplification would also
+enable the submission of different types of payload, without having to develop a
+specific PBS script for every payload. Finally, the submission process would be
+MPI-independent, as MPI is used for coordination among multiple work nodes, not
+by the payload.
+
+\mtnote{NOTE: I do not mention backfill availability Vs normal availability (i.e., submitting to the queue a `normal' job) because, in theory, the design of PanDA Broker enable this capability.
+I would speak about about backfill/not backfill in the discussion of NGE, when speaking about its generality.
+Part of this generality, is being able to submit to whatever batch system is supported by SAGA, including Titan's PBS and all its queues.}
+
+\mtnote{NOTE: NGE is not able to use backfill on Titan as we do not have specific functionalities to interrogate the Moab scheduler about backfill availability.
+This is why NGE should be presented as the pilot for PanDA Broker and not as an alternative to PanDA Broker.
+NGE alone would not be able to speak directly to PanDA Server.}
+
+% -----------------------------------------------------------------------------
+\subsection{Next Generation Executer}
+\label{ssec:nge}
+
+\begin{enumerate}
+    \item Rationale and design (why)
+    \item Architecture (how)
+    \item Integration
+    \item Characterization (experiments)
+\end{enumerate}
+
+We are developing a prototype of Next Generation Executer (NGE) to add pilot
+capabilities to the PanDA Broker. NGE is \ldots

### 2017-03-11 abstract.tex
Clean up

--- 2017-03-06 paper.tex
@@
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
-
-\input{head}
-\input{include}
-
-\usepackage{listings}
-\usepackage{comment}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-% -----------------------------------------------------------------------------
-% BEGIN DOCUMENT
-% -----------------------------------------------------------------------------
-\begin{document}
-
-\title{Everything you wanted to know about PanDA but were truly truly afraid to
-ask}
-
-% author names and affiliations
-% use a multiple column layout for up to two different
-% affiliations
-\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept.\ name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-\and
-\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept.\ name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3},
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
-\maketitle
-
-% ----------------------------------------------------------------------------
-% ABSTRACT
-% ----------------------------------------------------------------------------
-\begin{abstract}
 Experiments at the Large Hadron Collider (LHC) face unprecedented computing
@@
 user base.
-\end{abstract}
-
-\begin{IEEEkeywords}
-TBD\@.
-\end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-% ----------------------------------------------------------------------------
-% I - INTRODUCTION
-% ----------------------------------------------------------------------------
-\input{introduction}
-
-\begin{enumerate}
-  \item Problem that has been solved: executing many applications on
-  heterogeneous distributed resources.
-  \item Why are WMS needed?
-  \item Increasing importance of WMS w.r.t complex and scalable applications on
-  heterogeneous distributed resources.
-  \item \ldots
-  \item Overview/Summary of paper.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% II - BACKGROUND AND RELATED WORK
-% ----------------------------------------------------------------------------
-\section{Background and Related Work}
-\label{sec:bkgrd}
-
-\begin{enumerate}
-  \item Brief overview of distributed CI\@.
-  \item Workload vs Workflow. (DONE)
-  \item Distinguish between WFMS and WLMS\@.
-  \begin{enumerate}
-    \item Role of Workload MS\@: help ``hide'' and federate heterogeneous resources.
-  \end{enumerate}
-\end{enumerate}
-
-% Please do not correct bckground to background.
-\input{bckground}
-
-\begin{enumerate}
-  \item Other workload management systems (WMS) (not exclusive)
-  \begin{enumerate}
-    \item glidein-WMS
-    \item Dirac
-    \item ALICE-AlieN
-  \end{enumerate}
-  \begin{enumerate}
-    \item Comparison and contrast
-    \item Core features of a general WMS, i.e., minimal complete model of a WMS\@.
-  \end{enumerate}
-\end{enumerate}
-
-\input{WMSLiterature}
-
-% ----------------------------------------------------------------------------
-% III - PanDA OVERVIEW
-% ----------------------------------------------------------------------------
-\section{PanDA Overview}
-\label{sec:panda_overview}
-
-\input{PandaOverview}
-
-% ----------------------------------------------------------------------------
-% IV - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
-% ----------------------------------------------------------------------------
-\section{Deploying PanDA on a Leadership-scale system}
-\label{sec:panda_deployment}
-
-% \begin{enumerate}
-%   \item Why? To support a specific stage of the ATLAS Monte Carlo Workflow.
-%   \item Challenges
-%   \item Opportunities
-%   \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
-%   \item Take-aways, lessons learned
-% \end{enumerate}
-
-\input{PandaEvolution}
-
-% ----------------------------------------------------------------------------
-% VI - PERFORMANCE CHARACTERIZATION
-% ----------------------------------------------------------------------------
-\section{Performance Characterization}
-\label{sec:panda_titan}
-
-\input{overall}
-
-%\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
-%\label{sec:panda_titan}
-%
-%\begin{enumerate}
-%  \item Metadata performance issue, and how it’s resolved
-%  \item File I/O performance and impact
-%  \item ?
-%\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VII - THE NEXT GENERATION EXECUTOR
-% ----------------------------------------------------------------------------
-% \section{PANDA: Future/RoadMap}
-\section{PANDA\@: The Next Generation Executor}
-\label{sec:panda_roadmap}
-
-\begin{enumerate}
-    \item Discussion of the limitations of the current state of the art
-    \item Rationale and design (why)
-    \item Architecture (how)
-    \item Integration
-    \item Characterization (experiments)
-\end{enumerate}
-
-\input{Experiments}
-
-% ----------------------------------------------------------------------------
-% VIII - CONCLUSIONS
-% ----------------------------------------------------------------------------
-\section{Conclusion}
-\label{sec:conclusion}
-
-The PanDA  system was developed to meet the scale and complexity of LHC
-distributed computing for the ATLAS experiment.  In the process,  the old batch
-job paradigm  of computing in HEP was  discarded  in favor of a  far more
-flexible and scalable  model. The success  of PanDA  at the LHC is leading to
-widespread adoption and testing by other experiments. PanDA  is the first
-exascale  workload management system in HEP, already operating at a million
-computing jobs per day, and processing over an exabyte of data in 2013. Next LHC
-run will pose massive computing  challenges. With a  doubling of the beam
-energy  and luminosity as  well as an increased  need for  simulates  data, the
-data volume is expected to increase with a factor 5--6 or more. Storing and
-processing  this amount of data is a  challenge   that cannot be resolved with
-the currently existing  computing  resources in ATLAS\@. To resolve this
-challenge, ATLAS is turning to commercial  as well as academic Cloud services
-and HPCs via the PanDA system. Also the work underway is enabling the use of
-PanDA by new scientific collaborations and communities as a means  of leveraging
-extreme scale computing  resources with a low barrier of entry. The technology
-base provided by the PanDA system will enhance the usage of a variety  of
-high-performance computing resources available to basic research.
-
-% ----------------------------------------------------------------------------
-% ACKNOWLEDGEMENTS
-% ----------------------------------------------------------------------------
-\input{ack}
-
-%\section*{Acknowledgment}
-
-%The authors would like to thank... More thanks here...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-
-% ----------------------------------------------------------------------------
-% REFERENCES
-% ----------------------------------------------------------------------------
-\bibliographystyle{plain}
-\bibliography{bibliography}
-
-% that's all folks
-\end{document}

### 2017-03-11 conclusion.tex
Clean up

--- 2017-03-06 paper.tex
@@
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
-
-\input{head}
-\input{include}
-
-\usepackage{listings}
-\usepackage{comment}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-% -----------------------------------------------------------------------------
-% BEGIN DOCUMENT
-% -----------------------------------------------------------------------------
-\begin{document}
-
-\title{Everything you wanted to know about PanDA but were truly truly afraid to
-ask}
-
-% author names and affiliations
-% use a multiple column layout for up to two different
-% affiliations
-\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept.\ name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-\and
-\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept.\ name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3},
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
-\maketitle
-
-% ----------------------------------------------------------------------------
-% ABSTRACT
-% ----------------------------------------------------------------------------
-\begin{abstract}
-Experiments at the Large Hadron Collider (LHC) face unprecedented computing
-challenges. Heterogeneous resources are distributed worldwide, thousands of
-physicists analyzing the data need remote access to hundreds of computing
-sites, the volume of processed data is beyond the exabyte scale, and data
-processing requires more than billions of hours of computing usage per year. The
-PanDA (Production and Distributed Analysis) system was developed to meet the
-scale and complexity of LHC distributed computing for the ATLAS experiment. In
-the process, the old batch job paradigm of computing in HEP was discarded in
-favor of a far more flexible and scalable model. The success of PanDA at the LHC
-is leading to widespread adoption and testing by other experiments. PanDA is
-the first exascale workload management system in HEP, already operating at
-a million computing jobs per day, and processing over an exabyte of data in
-2013. We will describe the design and implementation of PanDA, present data on
-the performance of PanDA at the LHC, and discuss plans for future evolution
-of the system to meet new challenges of scale, heterogeneity and increasing
-user base.
-\end{abstract}
-
-\begin{IEEEkeywords}
-TBD\@.
-\end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-% ----------------------------------------------------------------------------
-% I - INTRODUCTION
-% ----------------------------------------------------------------------------
-\input{introduction}
-
-\begin{enumerate}
-  \item Problem that has been solved: executing many applications on
-  heterogeneous distributed resources.
-  \item Why are WMS needed?
-  \item Increasing importance of WMS w.r.t complex and scalable applications on
-  heterogeneous distributed resources.
-  \item \ldots
-  \item Overview/Summary of paper.
-\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% II - BACKGROUND AND RELATED WORK
-% ----------------------------------------------------------------------------
-\section{Background and Related Work}
-\label{sec:bkgrd}
-
-\begin{enumerate}
-  \item Brief overview of distributed CI\@.
-  \item Workload vs Workflow. (DONE)
-  \item Distinguish between WFMS and WLMS\@.
-  \begin{enumerate}
-    \item Role of Workload MS\@: help ``hide'' and federate heterogeneous resources.
-  \end{enumerate}
-\end{enumerate}
-
-% Please do not correct bckground to background.
-\input{bckground}
-
-\begin{enumerate}
-  \item Other workload management systems (WMS) (not exclusive)
-  \begin{enumerate}
-    \item glidein-WMS
-    \item Dirac
-    \item ALICE-AlieN
-  \end{enumerate}
-  \begin{enumerate}
-    \item Comparison and contrast
-    \item Core features of a general WMS, i.e., minimal complete model of a WMS\@.
-  \end{enumerate}
-\end{enumerate}
-
-\input{WMSLiterature}
-
-% ----------------------------------------------------------------------------
-% III - PanDA OVERVIEW
-% ----------------------------------------------------------------------------
-\section{PanDA Overview}
-\label{sec:panda_overview}
-
-\input{PandaOverview}
-
-% ----------------------------------------------------------------------------
-% IV - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
-% ----------------------------------------------------------------------------
-\section{Deploying PanDA on a Leadership-scale system}
-\label{sec:panda_deployment}
-
-% \begin{enumerate}
-%   \item Why? To support a specific stage of the ATLAS Monte Carlo Workflow.
-%   \item Challenges
-%   \item Opportunities
-%   \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
-%   \item Take-aways, lessons learned
-% \end{enumerate}
-
-\input{PandaEvolution}
-
-% ----------------------------------------------------------------------------
-% VI - PERFORMANCE CHARACTERIZATION
-% ----------------------------------------------------------------------------
-\section{Performance Characterization}
-\label{sec:panda_titan}
-
-\input{overall}
-
-%\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
-%\label{sec:panda_titan}
-%
-%\begin{enumerate}
-%  \item Metadata performance issue, and how it’s resolved
-%  \item File I/O performance and impact
-%  \item ?
-%\end{enumerate}
-
-% ----------------------------------------------------------------------------
-% VII - THE NEXT GENERATION EXECUTOR
-% ----------------------------------------------------------------------------
-% \section{PANDA: Future/RoadMap}
-\section{PANDA\@: The Next Generation Executor}
-\label{sec:panda_roadmap}
-
-\begin{enumerate}
-    \item Discussion of the limitations of the current state of the art
-    \item Rationale and design (why)
-    \item Architecture (how)
-    \item Integration
-    \item Characterization (experiments)
-\end{enumerate}
-
-\input{Experiments}
-
-% ----------------------------------------------------------------------------
-% VIII - CONCLUSIONS
-% ----------------------------------------------------------------------------
-\section{Conclusion}
-\label{sec:conclusion}
-
 The PanDA  system was developed to meet the scale and complexity of LHC
@@
 high-performance computing resources available to basic research.
-
-% ----------------------------------------------------------------------------
-% ACKNOWLEDGEMENTS
-% ----------------------------------------------------------------------------
-\input{ack}
-
-%\section*{Acknowledgment}
-
-%The authors would like to thank... More thanks here...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-
-% ----------------------------------------------------------------------------
-% REFERENCES
-% ----------------------------------------------------------------------------
-\bibliographystyle{plain}
-\bibliography{bibliography}
-
-% that's all folks
-\end{document}

### 2017-03-11 paper.tex
Ongoing section V
Stud: connection V to VI
Clean up

@@
 \input{include}
-
-\usepackage{listings}
-\usepackage{comment}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
 
@@
 \begin{abstract}
-Experiments at the Large Hadron Collider (LHC) face unprecedented computing
-challenges. Heterogeneous resources are distributed worldwide, thousands of
-physicists analyzing the data need remote access to hundreds of computing
-sites, the volume of processed data is beyond the exabyte scale, and data
-processing requires more than billions of hours of computing usage per year. The
-PanDA (Production and Distributed Analysis) system was developed to meet the
-scale and complexity of LHC distributed computing for the ATLAS experiment. In
-the process, the old batch job paradigm of computing in HEP was discarded in
-favor of a far more flexible and scalable model. The success of PanDA at the LHC
-is leading to widespread adoption and testing by other experiments. PanDA is
-the first exascale workload management system in HEP, already operating at
-a million computing jobs per day, and processing over an exabyte of data in
-2013. We will describe the design and implementation of PanDA, present data on
-the performance of PanDA at the LHC, and discuss plans for future evolution
-of the system to meet new challenges of scale, heterogeneity and increasing
-user base.
+\input{abstract}
 \end{abstract}
@@
 % ----------------------------------------------------------------------------
+\section{Introduction}
+\label{sec:intro}
+
 \input{introduction}
-
-\begin{enumerate}
-  \item Problem that has been solved: executing many applications on
-  heterogeneous distributed resources.
-  \item Why are WMS needed?
-  \item Increasing importance of WMS w.r.t complex and scalable applications on
-  heterogeneous distributed resources.
-  \item \ldots
-  \item Overview/Summary of paper.
-\end{enumerate}
 
@@
 
-\begin{enumerate}
-  \item Brief overview of distributed CI\@.
-  \item Workload vs Workflow. (DONE)
-  \item Distinguish between WFMS and WLMS\@.
-  \begin{enumerate}
-    \item Role of Workload MS\@: help ``hide'' and federate heterogeneous resources.
-  \end{enumerate}
-\end{enumerate}
-
+\input{WMSLiterature}
 % Please do not correct bckground to background.
 \input{bckground}
-
-\begin{enumerate}
-  \item Other workload management systems (WMS) (not exclusive)
-  \begin{enumerate}
-    \item glidein-WMS
-    \item Dirac
-    \item ALICE-AlieN
-  \end{enumerate}
-  \begin{enumerate}
-    \item Comparison and contrast
-    \item Core features of a general WMS, i.e., minimal complete model of a WMS\@.
-  \end{enumerate}
-\end{enumerate}
-
-\input{WMSLiterature}
 
@@
 
-% \begin{enumerate}
-%   \item Why? To support a specific stage of the ATLAS Monte Carlo Workflow.
-%   \item Challenges
-%   \item Opportunities
-%   \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
-%   \item Take-aways, lessons learned
-% \end{enumerate}
-
 \input{PandaEvolution}
@@
 % ----------------------------------------------------------------------------
-% VI - PERFORMANCE CHARACTERIZATION
+% V - ANALYSIS AND DISCUSSION
 % ----------------------------------------------------------------------------
-\section{Performance Characterization}
+\section{Analysis and Discussion}
 \label{sec:panda_titan}
@@
 \input{overall}
-
-%\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
-%\label{sec:panda_titan}
-%
-%\begin{enumerate}
-%  \item Metadata performance issue, and how it’s resolved
-%  \item File I/O performance and impact
-%  \item ?
-%\end{enumerate}
 
@@
 
-\begin{enumerate}
-    \item Discussion of the limitations of the current state of the art
-    \item Rationale and design (why)
-    \item Architecture (how)
-    \item Integration
-    \item Characterization (experiments)
-\end{enumerate}
-
+\input{nge}
 \input{Experiments}
@@
 % ----------------------------------------------------------------------------
-% VIII - CONCLUSIONS
+% VIII - CONCLUSION
 % ----------------------------------------------------------------------------
@@
 
-The PanDA  system was developed to meet the scale and complexity of LHC
-distributed computing for the ATLAS experiment.  In the process,  the old batch
-job paradigm  of computing in HEP was  discarded  in favor of a  far more
-flexible and scalable  model. The success  of PanDA  at the LHC is leading to
-widespread adoption and testing by other experiments. PanDA  is the first
-exascale  workload management system in HEP, already operating at a million
-computing jobs per day, and processing over an exabyte of data in 2013. Next LHC
-run will pose massive computing  challenges. With a  doubling of the beam
-energy  and luminosity as  well as an increased  need for  simulates  data, the
-data volume is expected to increase with a factor 5--6 or more. Storing and
-processing  this amount of data is a  challenge   that cannot be resolved with
-the currently existing  computing  resources in ATLAS\@. To resolve this
-challenge, ATLAS is turning to commercial  as well as academic Cloud services
-and HPCs via the PanDA system. Also the work underway is enabling the use of
-PanDA by new scientific collaborations and communities as a means  of leveraging
-extreme scale computing  resources with a low barrier of entry. The technology
-base provided by the PanDA system will enhance the usage of a variety  of
-high-performance computing resources available to basic research.
+\input{conclusion}
 
@@
 % ----------------------------------------------------------------------------
+\section*{Acknowledgements}
+\label{sec:ack}
+
 \input{ack}
-
-%\section*{Acknowledgment}
-
-%The authors would like to thank... More thanks here...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
 

### 2017-03-12 WMSLiterature.tex
Ongoing section II

--- 2017-02-06 WMSLiterature.tex
@@
-\subsection{Related Work}
-\label{sec:related}
+The Production and Distributed Analysis (PanDA) is a Workload Management System
+(WMS) developed to support the operations of the LHC ATLAS experiment. PanDA is
+specifically design to manage the execution of distributed workloads and
+workflows via pilots.
 
-In this section, we present other workload manager systems that have been developed to support LHC related but they have been used also to support other computing demanding projects.
+WMS for distributed executions are a type of middlware, designed to coordinate
+specific activities about resources and workloads: discovering and selecting
+resources, submitting the tasks of a workload, and monitoring the  execution of
+those tasks~\cite{marco2009glite}. Pilot is an abstraction that enables
+multi-level scheduling by decoupling resource acquisition from workload
+scheduling~\cite{turilli2015comprehensive}. Pilots are implemented by resource
+placeholders: a request for a certain amount of resources is scheduled on a site
+(e.g., by submitting a job or requesting a virtual machine) and once the
+resources are acquired, tasks are scheduled directly to the pilot, not to site's
+scheduler.
+
+WMS can implement the pilot abstraction: WMS control the acquisition of
+resources by means of pilots and then manage the execution of workloads or
+workflows' tasks on those pilots. In tis way, pilots expose a uniform scheduling
+interface and execution environment, isolating the differences among the
+heterogeneous resources of multiple sites; WMS centralize the management of the
+execution process, offering a unified interface to the application layer.
+Pilot-enabled WMS are particularly relevant for LHC experiments, where large
+amount of heterogeneous resources, distributed across multiple sites have to be
+coordinated to execute tens of million of tasks per months while producing and
+storing petabytes of data.
+
+Several pilot-enabled WMS were developed for the LHC experiments:
+AliEn~\cite{Bagnasco2010} for ALICE; DIRAC~\cite{Paterson2010} for LHCb;
+GlideinWMS~\cite{sfiligoi2008glideinwms} for CMS; and
+PanDA~\cite{maeno2014evolution} for ATLAS. These systems implement similar
+design and architectural principles: separation of concern between management
+and execution, data and compute; centralization of the management capabilities;
+distributed execution of the execution capabilities across multiple sites;
+unification of the application interface; hiding of resource heterogeneity;
+centralized monitoring and accounting of both resources and execution; and
+collection of static and sometimes dynamic information about resources.
+
+For example, \mtnote{Summarize design and architecture descriptions of the four
+WMS stressing some of their main similarities. Use a subset of the `principles'
+listed above.}
+
+The implementations of pilot-enabled WMS mainly differ in how specialized they
+are for a type of resource middleware and workflow. All the WMS developed for
+LHC assume to use sites exposing different flavors of Grid middleware. As such,
+all the WMS implementations support Grid-like authentication and authorization
+and a computational model based on distributing a large amount of single-core
+tasks across hundreds of sites\mtnote{Is this true?}.
+
+% Differences among workflow supported.
+
+AliEn, DIRAC, and PanDA are implemented to support specific experimental
+workflows, while GlideinWMS was developed as component of the Condor software
+framework. ALICE workload is \ldots. LHCb workload is \ldots. CMS workload is
+\ldots.
+
+The ATLAS project requires data, compute, and memory-intensive workflows, both
+for single users and so called `production' research groups. Generally, PanDA
+was designed to support mostly single-core tasks, requiring a variable \ldots.
+
+% In this section, we present other workload manager systems that have been
+% developed to support LHC related but they have been used also to support other
+% computing demanding projects.
 
 %\begin{itemize}
-%\item \emph{Clients}: consist in a set of APIs that allows users to submit job requests. Clients interact directly with DIRAC central services.
-%\item \emph{Resources}: they can be PC's, site clusters and Grids. Agents interact with them without distinction.
+%   \item \emph{Clients}: consist in a set of APIs that allows users to submit
+% job requests. Clients interact directly with DIRAC central services.
+%   \item \emph{Resources}: they can be PC's, site clusters and Grids. Agents
+% interact with them without distinction.
 %\end{itemize}
 
+% -----------------------------------------------------------------------------
 \subsubsection{Alien}
-Alien is a Workload and Data Management system composed of a 	set of middleware tools and services entirely based on web-services and standard protocols.
-The framework was originally developed for the ALICE experiment \cite{Alice1995} but subsequently used by several virtual organizations \cite{McClatechey2003,GPCALMA}.
-%The system has been deployed in 2001 for distributed producti	on of Monte Carlo data, detector simulation and reconstruction.
 
-Alien is composed of two type of services \cite{Bagnasco2010}:
+Alien is a Workload and Data Management system composed of a set of middleware
+tools and services entirely based on web-services and standard protocols. The
+framework was originally developed for the ALICE experiment~\cite{Alice1995} but
+subsequently used by several virtual organizations
+\cite{McClatechey2003,GPCALMA}.
+
+% The system has been deployed in 2001 for distributed production of Monte Carlo
+% data, detector simulation and reconstruction.
+
+Alien is composed of two type of services~\cite{Bagnasco2010}:
+
 \begin{itemize}
-\item
-\emph{Central services}, these services are unique for each virtual organisation, therefore there is only one  configuration point for the management;
+    \item \emph{Central services}, these services are unique for each virtual organization, therefore there is only one configuration point for the management;
     \item \emph{Site services}, they provide the interfacing to local resources and Grid services running on a VO-box;
@@
 
-%The Computing Agents are instead site services that monitor the local Computing Element, advertise site's capabilities and are responsible for submitting the JobAgents.
+%The Computing Agents are instead site services that monitor the local
+% Computing Element, advertise site's capabilities and are responsible for
+% submitting the JobAgents.
 
-%Information about the status of the sites and central services, full job statistics and monitoring information are kept in a MonALISA repository.
+% Information about the status of the sites and central services, full job
+% statistics and monitoring information are kept in a MonALISA repository.
 %% CLUSTER MONITOR SHOULD BE EQUAL TO COMPUTING AGENT
@@
 On one side the Cluster Monitor is used to communicate with Central services (Task queue and Broker), on the other side it can manage Computing Elements (CE) by starting and stopping them whenever it receives the signal.
-The CE is in charge of the execution of the tasks on the resources. A CE usually is associated with a batch queue and can send the tasks to the worker nodes controlled by the queue.
 
-The CE asks the Broker for tasks to execute by sending its description in form of Job Description Language (JDL).
-Once received the JDL, the Broker will try to match it with the JDL of thetasks in queue. If a match exists then the Broker sends the tasks JDL to the CE.
-Immediately after receving a job JDL, the CE create a new service on the worker node called ProcessMonitor.
-This service allows the CE (and the rest of Alien services through the CE) to interact with the job while is running \cite{Saiz2003}.
-This execution strategy is called ``pull mode'' due to the fact that CE asks for tasks.
+The CE is in charge of the execution of the tasks on the resources. A CE usually
+is associated with a batch queue and can send the tasks to the worker nodes
+controlled by the queue. The CE asks the Broker for tasks to execute by sending
+its description in form of Job Description Language (JDL). Once received the
+JDL, the Broker will try to match it with the JDL of the tasks in queue. If a
+match exists then the Broker sends the tasks JDL to the CE. Immediately after
+receiving a job JDL, the CE creates a new service on the worker node called
+ProcessMonitor. This service allows the CE (and the rest of Alien services
+through the CE) to interact with the job while is running \cite{Saiz2003}. This
+execution strategy is called ``pull mode'' due to the fact that CE asks for
+tasks.
 
-Since Alien exploits JDL, a  workload can be described task by task according to features such as: task priorities, the level of parallelism (one core, multi-core, MPI etc..) and  also the DCR that should be targeted for the execution.
+% Since Alien exploits JDL, a workload can be described task by task according
+% to features such as: task priorities, the level of parallelism (one core,
+% multi-core, MPI etc..) and also the DCR that should be targeted for the
+% execution.
 
-%Job submission is implemented by following the so-called ``pull mode'' which is composed of the following steps: 
-%\begin{enumerate}
-% \item the VO-Box monitors the status of the site queues through polls to the resource running on the CE; 
-%\item  the Job Broker receives a report everytime slots become available; 
-%\item  if the Task Queue is not empty, the Job Broker asks the VO-box to submit a number of Agents;
-%\item finally, the JobAgents are submitted  to the site Computing Element either by way of that sends them back to the site Computing Element or, wherever available, directly through the CREAM interface on the CE itself.
-%\end{enumerate}
+% Job submission is implemented by following the so-called ``pull mode'' which
+% is composed of the following steps:
 
+% \begin{enumerate}
+%   \item the VO-Box monitors the status of the site queues through polls to
+% the resource running on the CE;
+%   \item the Job Broker receives a report everytime slots become available;
+%   \item if the Task Queue is not empty, the Job Broker asks the VO-box to
+% submit a number of Agents;
+%   \item finally, the JobAgents are submitted  to the site Computing Element
+% either by way of that sends them back to the site Computing Element or,
+% wherever available, directly through the CREAM interface on the CE itself.
+% \end{enumerate}
+
+% -----------------------------------------------------------------------------
 \subsubsection{DIRAC}
-DIRAC (Distributed Infrastructure with Remote Agent Control) Workload and Data Management System has been developed within the CERN LHCb project to manage the processing of detector data, Monte Carlo simulations, and end-user analyses \cite{Tsaregorodtsev2004}.
-DIRAC  architecture relies on two entities \cite{Paterson2010}:
+
+DIRAC (Distributed Infrastructure with Remote Agent Control) Workload and Data
+Management System has been developed within the CERN LHCb project to manage the
+processing of detector data, Monte Carlo simulations, and end-user analyses
+\cite{Tsaregorodtsev2004}. DIRAC's architecture relies on two entities
+\cite{Paterson2010}:
+
 \begin{itemize}
-\item \emph{Services}: serve Clients and Agents by performing crucial operations such as Task Management, Configuration, Bookkeping and Accounting.
-\item \emph{Agents}: perform repetitive actions like querying file catalogs,  monitoring of jobs on resources.
+    \item \emph{Services}: serve Clients and Agents by performing crucial
+    operations such as Task Management, Configuration, Bookkeeping and
+    Accounting.
+    \item \emph{Agents}: perform repetitive actions like querying file
+    catalogs, monitoring of jobs on resources.
 \end{itemize}
-In the same way of Alien, DIRAC implements a pull scheduling. Furthermore DIRAC was the first WMS to exploit the concept of Pilot Agent on the Grid \cite{Casajus2010}. 
-Pilots Agents are gLite jobs that are submitted to the grid when tasks arrive into the WMS. 
-DIRAC pilot system has four main logical components:
+
+In the same way of Alien, DIRAC implements a pull scheduling. Furthermore DIRAC
+was the first WMS to exploit the concept of Pilot Agent on the Grid
+\cite{Casajus2010}, i.e., gLite jobs that are submitted to the grid when tasks
+arrive into the WMS. DIRAC pilot system has four main logical components:
+
 \begin{itemize}
-\item a set of TaskQueues that collect tasks submitted by users, multiple TaskQueue being created depending on the requirements and ownership of the tasks;
-\item a set of JobWrappers that are executed on the DCR to bind compute resources and execute tasks submitted by the users;
+    \item a set of TaskQueues that collect tasks submitted by users, multiple
+    TaskQueue being created depending on the requirements and ownership of the
+    tasks;
+    \item a set of JobWrappers that are executed on the DCR to bind compute
+    resources and execute tasks submitted by the users;
     \item a set of TaskQueueDirectors that submits JobWrappers to target DCRs;
-\item a MatchMaker that matches requests from JobWrappers to suitable tasks into TaskQueues.
+    \item a MatchMaker that matches requests from JobWrappers to suitable tasks
+    into TaskQueues.
 \end{itemize}
-The DIRAC execution model can be summarized in five
-steps: 1.
-a user submits its workload in form of tasks to the WMS Job Manager; 2.
-submitted tasks are validated and added to a new or an existing TaskQueue, depending on the task properties; 3.
-TaskQueueDirector evaluates TaskQueues and a suitable number of JobWrappers are submitted to available
-DCRs; 4. JobWrappers get instantiated on the DCRs and, then,  ask for tasks to the MatchMaker; 5. JobWrappers execute tasks while JobWrapper’s Watchdog monitor them.
-TaskQueueDirectors deploy Pilots by getting a list of TaskQueues and calculating the number of pilot to submit  according to user priorities.
-Once deployed on the compute resource, Pilots, a.k.a.
-JobWrappers, hold the resource in the form of single or multiple cores, spanning portions, whole, or multiple compute nodes.
-Pilots do not expose data capabilities although the system allows the user to perform both data staging and data replication.
-TaskQueues, TaskQueueDirectors, and the MatchMaker are implemented as services whereas the JobWrapper is implemented within the Agents together with the WatchDog.
 
-JDL and \emph{Transformation Management System} (TMS) are used for task description \cite{Tsaregorodtsev2006,Pacini2006}. The latter allows the user to describe complex data dependences. 
+The DIRAC execution model can be summarized in five steps: 1. a user submits its
+workload in form of tasks to the WMS Job Manager; 2. submitted tasks are
+validated and added to a new or an existing TaskQueue, depending on the task
+properties; 3. TaskQueueDirector evaluates TaskQueues and a suitable number of
+JobWrappers are submitted to available DCRs; 4. JobWrappers get instantiated on
+the DCRs and, then, ask for tasks to the MatchMaker; 5. JobWrappers execute
+tasks while JobWrapper’s Watchdog monitor them.
+
+TaskQueueDirectors deploy Pilots by getting a list of TaskQueues and calculating
+the number of pilot to submit according to user priorities. Once deployed on the
+compute resource, Pilots, a.k.a. JobWrappers, hold the resource in the form of
+single or multiple cores, spanning portions, whole, or multiple compute nodes.
+Pilots do not expose data capabilities although the system allows the user to
+perform both data staging and data replication. TaskQueues, TaskQueueDirectors,
+and the MatchMaker are implemented as services whereas the JobWrapper is
+implemented within the Agents together with the WatchDog.
+
+% JDL and \emph{Transformation Management System} (TMS) are used for task
+% description \cite{Tsaregorodtsev2006,Pacini2006}. The latter allows the user
+% to describe complex data dependences.
+
+% -----------------------------------------------------------------------------
 \subsubsection{HTCondor Glidein and GlideinWMS}
-The HTCondor Glidein system  as part of the HTCondor software ecosystem. The HTCondor Glidein is a pilot based system to aggregate DCRs with heterogeneous middleware into HTCondor resource pools.
-Condor is based on daemons collaborating by exchanging messages over the network. We can isolate four main logical components \cite{Sfiligoi2008}:
+
+The HTCondor Glidein system as part of the HTCondor software ecosystem. The
+HTCondor Glidein is a pilot based system to aggregate DCRs with heterogeneous
+middleware into HTCondor resource pools. Condor is based on daemons
+collaborating by exchanging messages over the network. We can isolate four main
+logical components~\cite{Sfiligoi2008}:
+
 \begin{itemize}
@@
 \end{itemize}
-Glidein-WMS has been developed to integrate HTCondor Glidein to  automate the deployment and management of Glideins on multiple types of DCR middleware. 
-The integration required three additional logical components: 
+
+Glidein-WMS has been developed to integrate HTCondor Glidein to  automate the
+deployment and management of Glideins on multiple types of DCR middleware. The
+integration required three additional logical components:
+
 \begin{itemize}
     \item \emph{Glidein Factories} that submit tasks to the DCRs middleware;
-\item a set of \emph{Virtual Organizations (VO) Frontend} daemons that match the tasks on one or more Schedd to the resource attributes;
-\item a \emph{Collector} that holds references to all the active Glidein Factories and VO Frontend daemons. 
+    \item a set of \emph{Virtual Organizations (VO) Frontend} daemons that
+    match the tasks on one or more Schedd to the resource attributes;
+    \item a \emph{Collector} that holds references to all the active Glidein
+    Factories and VO Frontend daemons.
 \end{itemize}
 
- The execution model of the HTCondor Glidein system can be summarized in nine steps: 1.
-the user submits a Glidein (i.e., a job) to a DCR batch scheduler; 2.
-once executed, this Glidein bootstraps a Startd daemon; 3.
-the Startd daemon advertises itself with the Collector; 4.
-the user submits the tasks of the workload to the Schedd daemon; 5.
-the Schedd advertises these tasks to the Collector; 6.
-the Negotiator matches the requirements of the tasks to the properties of one of the available Startd daemon (i.e., a Glidein); 7.
-the Negotiator communicates the match to the Schedd; 8.
-the Schedd submits the tasks to the Startd daemon indicated by
-the Negotiator; 9. the task is executed.
+The execution model of the HTCondor Glidein system can be summarized in nine
+steps: 1. the user submits a Glidein (i.e., a job) to a DCR batch scheduler; 2.
+once executed, this Glidein bootstraps a Startd daemon; 3. the Startd daemon
+advertises itself with the Collector; 4. the user submits the tasks of the
+workload to the Schedd daemon; 5. the Schedd advertises these tasks to the
+Collector; 6. the Negotiator matches the requirements of the tasks to the
+properties of one of the available Startd daemon (i.e., a Glidein); 7. the
+Negotiator communicates the match to the Schedd; 8. the Schedd submits the tasks
+to the Startd daemon indicated by the Negotiator; 9. the task is executed.
 
-By using GlideinWMS, the user does not have to submit Glidein directly but only tasks to Schedd.
-From there: 1.
-every Schedd advertises its tasks with the VO Frontend; 2.
-the VO Frontend matches the tasks’ requirements to the resource properties advertised by the WMS Connector; 3.
-the VO Frontend places requests for Glideins instantiation to the WMS Collector; 4.
-the WMS Collector contacts the appropriate Glidein Factory to execute the requested Glideins; 5.
-the requested Glideins become active on the DCRs; and 6.
-the Glideins advertise their availability to the (HTCondor) Collector.
-From there on the execution model is the same as described for the HTCondor Glidein Service.
-\begin{comment}
-The resources managed by a single Glidein (i.e., pilot) are limited to compute resources.
-Glideins may bind one or more cores, depending on the target DCRs.
-For example, heterogeneous HTCondor pools with resources for desktops, workstations, small campus clusters, and some larger clusters will run mostly single core Glideins.
-More specialized pools that hold, for example, only DCRs with HTC, Grid, or Cloud middleware may instantiate Glideins with a larger number of cores.
-Both HTCondor Glidein and GlideinWMS provide abstractions for file staging but pilots are not used to hold data or network resources.
-The process of pilot deployment is the main difference between HTCondor Glidein and GlideinWMS. While the
-HTCondor Glidein system requires users to submit the pilots to the DCRs, GlideinWMS automates and optimizes pilot provisioning.
-GlideinWMS attempts to maximize the throughput of task execution by continuously instantiating Glideins until the queues of the available Schedd are emptied.
-Once all the tasks have been executed, the remaining Glideins are terminated.
-HTCondor Glidein and GlideWMS expose the interfaces of HTCondor to the application layer and no theoretical
-limitations are posed on the type and complexity of the workloads that can be executed. For example, DAGMan
-(Directed Acyclic Graph Manager) has been designed to execute workflows by submitting tasks to Schedd, and a tool is available to design applications based on the master-worker coordination pattern.
+By using GlideinWMS, the user does not have to submit Glidein directly but only
+tasks to Schedd. From there: 1. every Schedd advertises its tasks with the VO
+Frontend; 2. the VO Frontend matches the tasks’ requirements to the resource
+properties advertised by the WMS Connector; 3. the VO Frontend places requests
+for Glideins instantiation to the WMS Collector; 4. the WMS Collector contacts
+the appropriate Glidein Factory to execute the requested Glideins; 5. the
+requested Glideins become active on the DCRs; and 6. the Glideins advertise
+their availability to the (HTCondor) Collector. From there on the execution
+model is the same as described for the HTCondor Glidein Service.
 
-Both HTCondor Glidein and GlideWMS rely on one or more HTCondor Collectors to match task requirements and resource properties, represented as ClassAds.
-This matching can be evaluated right before the scheduling of the task.
-In this way, late binding is achieved but early binding remains unsupported.
-\begin{table*}
-\begin{center}
-\begin{tabular}{llllllll}
-  \hline
-Pilot System  &Logical Components& Execution Strategy & Binding  & Workload Definition  &  Broker  & \\
-\hline
-Alien & Central services, site services & pull, pilot via Co-pilot & Late & JDL & Condor (ClassAd attributes) &\\
-DIRAC & Services, Agents& pull, pilot-based & Late & JDL, WF (TMS) & Condor(ClassAd attributes) &\\
-glidein WMS& Daemons & pull, pilot-based  & Late & Pegaus, DAGMan & Condor(ClassAd attributes) &\\
-\hline
-\end{tabular}
-\end{center}
-\caption{Comparison of the three Workload and Data Management Systems}\label{tab:Summary}
-\end{table*}
-\end{comment}
+% The resources managed by a single Glidein (i.e., pilot) are limited to compute
+% resources. Glideins may bind one or more cores, depending on the target DCRs.
+% For example, heterogeneous HTCondor pools with resources for desktops,
+% workstations, small campus clusters, and some larger clusters will run mostly
+% single core Glideins. More specialized pools that hold, for example, only DCRs
+% with HTC, Grid, or Cloud middleware may instantiate Glideins with a larger
+% number of cores. Both HTCondor Glidein and GlideinWMS provide abstractions for
+% file staging but pilots are not used to hold data or network resources. The
+% process of pilot deployment is the main difference between HTCondor Glidein
+% and GlideinWMS. While the
+%
+% HTCondor Glidein system requires users to submit the pilots to the DCRs,
+% GlideinWMS automates and optimizes pilot provisioning. GlideinWMS attempts to
+% maximize the throughput of task execution by continuously instantiating
+% Glideins until the queues of the available Schedd are emptied. Once all the
+% tasks have been executed, the remaining Glideins are terminated. HTCondor
+% Glidein and GlideWMS expose the interfaces of HTCondor to the application
+% layer and no theoretical limitations are posed on the type and complexity of
+% the workloads that can be executed. For example, DAGMan (Directed Acyclic
+% Graph Manager) has been designed to execute workflows by submitting tasks to
+% Schedd, and a tool is available to design applications based on the
+% master-worker coordination pattern.
+%
+% Both HTCondor Glidein and GlideWMS rely on one or more HTCondor Collectors to
+% match task requirements and resource properties, represented as ClassAds. This
+% matching can be evaluated right before the scheduling of the task. In this
+% way, late binding is achieved but early binding remains unsupported.
+%
+% \begin{table*}
+% \begin{center}
+% \begin{tabular}{llllllll}
+%   \hline
+%     Pilot System  &Logical Components& Execution Strategy & Binding  & Workload Definition  &  Broker  & \\
+%   \hline
+%     Alien & Central services, site services & pull, pilot via Co-pilot & Late & JDL & Condor (ClassAd attributes) &\\
+%     DIRAC & Services, Agents& pull, pilot-based & Late & JDL, WF (TMS) & Condor(ClassAd attributes) &\\
+%     glidein WMS& Daemons & pull, pilot-based  & Late & Pegaus, DAGMan & Condor(ClassAd attributes) &\\
+%   \hline
+% \end{tabular}
+% \end{center}
+% \caption{Comparison of the three Workload and Data Management Systems}\label{tab:Summary}
+% \end{table*}

### 2017-03-12 overall.tex
Address #3

@@
 Figure~\ref{fig:hpc-workload-utilization} shows that between January 2016 and
-February 2017, about 2.25M million detector simulation jobs were completed on
-Titan, for a total of 225M events processed. This is equivalent to 9\% of all
-the 250M detector simulations performed by ATLAS in the same period of time, and
-3.5\% of the 6.6B events processed by those jobs. Comparatively, Titan
-contributed 3.9\% of the total of around 200K cores available to ATLAS on Grid,
-Cloud, and HPC infrastructures together. These figures confirms the relevance of
+February 2017, about 2.25M detector simulation jobs were completed on Titan, for
+a total of 225M events processed. This is equivalent to 0.9\% of all the 250M
+detector simulations performed by ATLAS in the same period of time, and 3.5\% of
+the 6.6B events processed by those jobs. Comparatively, Titan contributed 3.9\%
+of the total of around 200K cores available to ATLAS on Grid, Cloud, and HPC
+infrastructures together. These figures confirms the relevance of
 supercomputers' resource contribution to the LHC Run 3, especially when

### 2017-03-12 PandaOverview.tex
Clean up
scattered changes.

@@
 
-\begin{itemize}
-  \item No MPI tasks?
-  \item Pilot multitenancy?
-  \item PanDA Server Brokerage component and Event service interaction?
-  \item PanDA Server Data Service component and Event Service interaction?
-\end{itemize}
+% \begin{itemize}
+%   \item No MPI tasks?
+%   \item Pilot multitenancy?
+%   \item PanDA Server Brokerage component and Event service interaction?
+%   \item PanDA Server Data Service component and Event Service interaction?
+% \end{itemize}
 
@@
 Server and the Brokerage module to collect information about the state of the
-jobs it manages and, on the base of this information, submit new pilots to one
+jobs it manages and, on the basis of this information, submit new pilots to one
 site per APFQueue. AutoPyFactory implements functionalities to evaluate the

### 2017-03-12 overall.tex
scattered changes.

--- 2017-03-11 overall.tex
@@
 ATLAS on Titan, from January 2016 to February 2017. During that period, ATLAS
-consumed a total of 73.8M core hours, for an average of 7M hours a month, with a
-minimum of 3.3M hours in April 2016 and a maximum 14.8M hours in February 2017.
+consumed a total of 73.8M core-hours, for an average of 7M core-hours a month, with a
+minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours in February 2017.
 

### 2017-03-12 overall.tex
Address #4

--- 2017-03-12 overall.tex
@@
 
-Figure~\ref{fig:core-hours-utilization} shows the number of core hours used by
+Figure~\ref{fig:core-hours-utilization} shows the number of core-hours used by
 ATLAS on Titan, from January 2016 to February 2017. During that period, ATLAS
-consumed a total of 73.8M core hours, for an average of 7M hours a month, with a
+consumed a total of 73.8M core-hours, for an average of 7M hours a month, with a
 minimum of 3.3M hours in April 2016 and a maximum 14.8M hours in February 2017.

### 2017-03-12 PandaEvolution.tex

@@
 
-Compared to pilots, backfill has the disadvantage of limiting the amount of work
-nodes that can be requested. Pilots are normal jobs: they can request as many
-work nodes for as much time as a queue can offer. On the contrary, jobs sized on
-the base of the backfill availability depend on the
-% amount of work nodes that cannot be given to any other job.
-any other job in the Titan's queue\sergeynote{any other job currently in the
-Titan's queue?}\mtnote{Better?}. Usually, backfill availability is a fraction of
-the total capacity of the queue but the size of Titan mitigates this limitation.
+Compared to pilots, backfill has the disadvantage of limiting the amount of
+work nodes that can be requested. Pilots are normal jobs: they can request as
+many work nodes for as much time as a queue can offer. On the contrary, jobs
+sized on the basis of backfill information availability \jhanote{should be
+resource availability or backfill queue information?} depend on the % amount
+of work nodes that cannot be given to any other job. any other job in the
+Titan's queue\sergeynote{any other job currently in the Titan's
+queue?}\mtnote{Better?}. Usually, backfill availability is a fraction of the
+total capacity of the queue but the size of Titan mitigates this limitation.
 Every year, about 10\% of Titan's capacity remains
-unused~\cite{titan_utilization}, for a total of 30,000 cores, corresponding to
-around 270M core hours per year, roughly 30\% of the overall capacity of WLCG.
+unused~\cite{titan_utilization}, corresponding to an average of 30,000 unused
+cores. This equals approximately 270M core hours per year, roughly 30\% of the
+overall capacity of WLCG.
 

### 2017-03-12 PandaEvolution.tex
Address #5

@@
 
-Titan's backfill functionality offers the opportunity not to use the pilot
-abstraction while still avoiding queue time overheads. Backfill availability is
-the number of work nodes that cannot be used by any of the jobs already queued
-on Titan: All queued jobs are either too large or their walltime is too long. At
-any point in time, Titan's Moab scheduler can be queried for backfill
-availability and a job submitted shortly after placing this query and requesting
-for no more than the backfill availability, is usually scheduled immediately
-spending almost no time in the queue.
+Titan’s backfill functionality offers the opportunity to avoid the overhead of
+queue wait times without using pilot abstraction. Backfill availability is the
+number of work nodes that cannot be used by any of the jobs already queued on
+Titan: All queued jobs are either too large or their walltime is too long. At
+any point in time, Titan’s Moab scheduler can be queried for backfill
+availability. Based upon the result of this query, a job can be shaped to
+request no more than the backfill availability, and when submitted as such is
+usually scheduled immediately spending almost no time in the queue.
 

### 2017-03-12 overall.tex
comments and changes in V-A.

@@
 
-We evaluated the efficiency, scalability and reliability of the PanDA Brokers
+We evaluate the efficiency, scalability and reliability of the PanDA Brokers
 along two dimensions: (i) the amount of backfill availability utilized by the
-PanDA Brokers; and (ii) the amount of jobs walltime spent performing detector
-simulations. We based these evaluations on measuring the number of cores
-utilized by ATLAS on Titan, the overall backfill availability, the number of
-detector simulations performed on all the resources available to ATLAS and on
-Titan, and the failure rate of these simulations. All the measurements were
-performed between January 2016 and February 2017.
+PanDA Brokers; and (ii) the amount of job walltime spent performing detector
+simulations. \jhanote{it is difficult to parse what (ii)?}. We based these
+evaluations on measuring the number of cores utilized by ATLAS on Titan, the
+overall backfill availability, the number of detector simulations performed on
+all the resources available to ATLAS and on Titan, and the failure rate of
+these simulations. All the measurements were performed between January 2016
+and February 2017.
 
@@
 ATLAS on Titan, from January 2016 to February 2017. During that period, ATLAS
-consumed a total of 73.8M core-hours, for an average of 7M hours a month, with a
-minimum of 3.3M hours in April 2016 and a maximum 14.8M hours in February 2017.
+consumed a total of 73.8M core-hours, for an average of $\approx$ 7M core-hours a month,
+with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours
+in February 2017.
 
@@
 
-On February 2017, PanDA Brokers used almost twice as much backfill availability
-than in any other month. No relevant code update was made during that period and
-logs indicated that the brokers were able to respond more promptly to backfill
-availability. This is likely due to hardware upgrades on the DTNs. The absence
-of continuous monitoring of the nodes does not allow to quantify bottlenecks but
-spot measurements of the DTNs load indicate that a faster CPU and better
-networking were likely responsible for the improved performance.
-
-Investigations showed an average CPU load of 3.6\% on the upgraded DTNs. As
-such, further hardware upgrades seem unlikely to improve significantly the
-performance of PanDA Brokers. Nonetheless, the current load suggests that the
-number of brokers per DTN could be increased. This would enable the submission
-of a larger number of concurrent jobs to Titan's PBS queue, allowing for PanDA
-to consume a higher percentage of backfill availability.
-
-Figure~\ref{fig:backfill-utilization} shows backfill utilization efficiency from
-January 2016 to February 2017. Efficiency is defined as the fraction of Titan’s
-cores available via backfill that was utilized by the PanDA Brokers. The brokers
-reached 18\% average efficiency with a minimum 8.9\% efficiency on April 2016
+Figure~\ref{fig:backfill-utilization} shows backfill utilization from
+January 2016 to February 2017. \jhanote{Neither figure show efficiency. Both show
+utilization.} Efficiency is defined as the fraction of Titan’s
+cores available via backfill that was utilized by the PanDA Brokers. \jhanote{I would plot the efficiency on the Y2 axis for both.} The brokers
+reached 18\% average efficiency \jhanote{this sentence is a bit difficult to parse. Is the 18\% overall efficiency?} with a minimum 8.9\% efficiency on April 2016
 and a maximum 33.5\% efficiency on February 2017. The number of total backfill
@@
 \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
-\caption{Caption.}
+\caption{Caption.
+\jhanote{I would suggest similar style of X-axis as previous figure.
+Dificult to parse which is April if we are going to call out a specific month.
+Also consistenty in style is generally good to keep cognitive burden low.}}
 \label{fig:backfill-utilization}
@@
 \end{figure}
+
+On February 2017, PanDA Brokers used almost twice as much backfill availability
+than in any other month. No relevant code update was made during that period and
+logs indicated that the brokers were able to respond more promptly to backfill
+availability. This is likely due to hardware upgrades on the DTNs. The absence
+of continuous monitoring of the nodes does not allow to quantify bottlenecks but
+spot measurements of the DTNs load indicate that a faster CPU and better
+networking were likely responsible for the improved performance.
+
+Investigations showed an average CPU load of 3.6\% on the upgraded DTNs. As
+such, further hardware upgrades seem unlikely to improve significantly the
+performance of PanDA Brokers. Nonetheless, the current load suggests that the
+number of brokers per DTN could be increased. This would enable the submission
+of a larger number of concurrent jobs to Titan's PBS queue, allowing for PanDA
+to consume a higher percentage of backfill availability.
 

### 2017-03-12 PandaEvolution.tex

@@
 sized on the basis of backfill information availability \jhanote{should be
-resource availability or backfill queue information?} depend on the % amount
+resource availability or backfill queue information?}\mtnote{AFAIH, there is no dedicated queue to backfill} depend on the number
+% amount
 of work nodes that cannot be given to any other job. any other job in the

### 2017-03-12 overall.tex
Addressing/answering comments V-A

@@
 along two dimensions: (i) the amount of backfill availability utilized by the
-PanDA Brokers; and (ii) the amount of job walltime spent performing detector
-simulations. \jhanote{it is difficult to parse what (ii)?}. We based these
-evaluations on measuring the number of cores utilized by ATLAS on Titan, the
-overall backfill availability, the number of detector simulations performed on
-all the resources available to ATLAS and on Titan, and the failure rate of
-these simulations. All the measurements were performed between January 2016
-and February 2017.
+PanDA Brokers; and (ii) the amount of % job walltime
+runtime spent performing detector simulations. \jhanote{it is difficult to parse
+what (ii)?}\mtnote{Better?} We based these evaluations on measuring the number
+of cores utilized by ATLAS on Titan, the overall backfill availability, the
+number of detector simulations performed on all the resources available to ATLAS
+and on Titan, and the failure rate of these simulations. All the measurements
+were performed between January 2016 and February 2017.
 
@@
 
-Figure~\ref{fig:backfill-utilization} shows backfill utilization from
-January 2016 to February 2017. \jhanote{Neither figure show efficiency. Both show
-utilization.} Efficiency is defined as the fraction of Titan’s
-cores available via backfill that was utilized by the PanDA Brokers. \jhanote{I would plot the efficiency on the Y2 axis for both.} The brokers
-reached 18\% average efficiency \jhanote{this sentence is a bit difficult to parse. Is the 18\% overall efficiency?} with a minimum 8.9\% efficiency on April 2016
-and a maximum 33.5\% efficiency on February 2017. The number of total backfill
-cores available was 38.1M in April 2016, and 33.1M in February 2017. This shows
-that the measured difference in efficiency did not depend on a comparable
-difference in total backfill availability.
+Figure~\ref{fig:backfill-utilization} shows backfill utilization from January
+2016 to February 2017. \jhanote{Neither figure show efficiency. Both show
+utilization.} Efficiency is defined as the fraction of Titan’s cores available
+via backfill that was utilized by the PanDA Brokers. \jhanote{I would plot the
+efficiency on the Y2 axis for both.} The brokers reached 18\% average efficiency
+\jhanote{this sentence is a bit difficult to parse. Is the 18\% overall
+efficiency?}\mtnote{average across all months considered. Should I call it
+``overall''?} with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\%
+efficiency on February 2017. The number of total backfill cores available was
+38.1M in April 2016, and 33.1M in February 2017. This shows that the measured
+difference in efficiency did not depend on a comparable difference in total
+backfill availability.
 
@@
 \jhanote{I would suggest similar style of X-axis as previous figure.
-Dificult to parse which is April if we are going to call out a specific month.
-Also consistenty in style is generally good to keep cognitive burden low.}}
+Difficult to parse which is April if we are going to call out a specific month.
+Also consistency in style is generally good to keep cognitive burden low.}}
 \label{fig:backfill-utilization}

### 2017-03-12 overall.tex
Address #6
fix, removed comments.

--- 2017-03-12 overall.tex
@@
 along two dimensions: (i) the amount of backfill availability utilized by the
-PanDA Brokers; and (ii) the amount of job walltime spent performing detector
-simulations. \jhanote{it is difficult to parse what (ii)?}. We based these
-evaluations on measuring the number of cores utilized by ATLAS on Titan, the
-overall backfill availability, the number of detector simulations performed on
-all the resources available to ATLAS and on Titan, and the failure rate of
-these simulations. All the measurements were performed between January 2016
-and February 2017.
+PanDA Brokers; and (ii) the amount of 
+runtime spent performing detector simulations. We based these evaluations on measuring the number
+of cores utilized by ATLAS on Titan, the overall backfill availability, the
+number of detector simulations performed on all the resources available to ATLAS
+and on Titan, and the failure rate of these simulations. All the measurements
+were performed between January 2016 and February 2017.
 
@@
 
-Figure~\ref{fig:backfill-utilization} shows backfill utilization from
-January 2016 to February 2017. \jhanote{Neither figure show efficiency. Both show
-utilization.} Efficiency is defined as the fraction of Titan’s
-cores available via backfill that was utilized by the PanDA Brokers. \jhanote{I would plot the efficiency on the Y2 axis for both.} The brokers
-reached 18\% average efficiency \jhanote{this sentence is a bit difficult to parse. Is the 18\% overall efficiency?} with a minimum 8.9\% efficiency on April 2016
-and a maximum 33.5\% efficiency on February 2017. The number of total backfill
-cores available was 38.1M in April 2016, and 33.1M in February 2017. This shows
-that the measured difference in efficiency did not depend on a comparable
-difference in total backfill availability.
+Figure~\ref{fig:backfill-utilization} shows backfill utilization from January
+2016 to February 2017 (hereafter called experiment time window). \jhanote{Neither figure show efficiency. Both show
+utilization.} Efficiency is defined as the fraction of Titan’s cores available
+via backfill that was utilized by the PanDA Brokers. \jhanote{I would plot the
+efficiency on the Y2 axis for both.} The brokers reached 18\% average efficiency
+when measured across the experiment time window,
+with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\%
+efficiency on February 2017. The number of total backfill cores available was
+38.1M in April 2016, and 33.1M in February 2017. This shows that the measured
+difference in efficiency did not depend on a comparable difference in total
+backfill availability.
 

### 2017-03-12 nge.tex
Typos

@@
 
-\mtnote{NOTE: I do not mention backfill availability Vs normal availability (i.e., submitting to the queue a `normal' job) because, in theory, the design of PanDA Broker enable this capability.
-I would speak about about backfill/not backfill in the discussion of NGE, when speaking about its generality.
-Part of this generality, is being able to submit to whatever batch system is supported by SAGA, including Titan's PBS and all its queues.}
+\mtnote{NOTE: I do not mention backfill availability Vs normal availability
+(i.e., submitting to the queue a `normal' job) because, in theory, the design of
+PanDA Broker enables this capability. I would speak about about backfill/not
+backfill in the discussion of NGE, when speaking about its generality. Part of
+this generality is being able to submit to whatever batch system is supported by
+SAGA, including Titan's PBS and all its queues.}
 

### 2017-03-13 PandaEvolution.tex
Question IV; transition V-VI

--- 2017-03-12 PandaEvolution.tex
@@
 
-Compared to pilots, backfill has the disadvantage of limiting the amount of
-work nodes that can be requested. Pilots are normal jobs: they can request as
-many work nodes for as much time as a queue can offer. On the contrary, jobs
-sized on the basis of backfill information availability \jhanote{should be
-resource availability or backfill queue information?} depend on the % amount
+Compared to pilots, backfill has the disadvantage of limiting the amount of work
+nodes that can be requested. Pilots are normal jobs: they can request as many
+work nodes for as much time as a queue can offer. On the contrary, jobs sized on
+the basis of backfill information availability \jhanote{should be resource
+availability or backfill queue information?}\mtnote{AFAIH, there is no dedicated
+queue to backfill} depend on the number % amount
 of work nodes that cannot be given to any other job. any other job in the
@@
 it a pilot.}\mtnote{Thank you. We would have a problem calling it `agent' as we
-use that term to name the pilot of NGE. Would PanDA Broker work?}
+use that term to name the pilot of NGE. Would PanDA Broker work?}\mtnote{Sergey
+agreed that we should not use `pilot'. He is quite convinced by `broker' but we
+are both open to better suggestions.}
 
@@
 
-Once deployed, every PanDA Broker supports the execution of MC
-% event generation and
-detector simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's
-Moab scheduler about currently unused worker nodes;
-% and checks whether the time and size of the available backfill resources are
-% suitable to execute jobs;
-(ii) the PanDA Broker queries the Job Dispatcher module of the PanDA server for
-jobs that have been bound to Titan by JEDI; (iii)  the PanDA Broker requests to
-the Data Service module of the PanDA Server what input data these jobs require
-and transfer them to the DTN via the ATLAS DDM service; (iv) the PanDA Broker
-creates one or more PBS MPI script, wrapping enough jobs' payload to fit
-backfill availability; (v) the PanDA Broker submits these MPI jobs to the
-Titan's PBS batch system via RADICAL-SAGA; (vi) the MPI jobs waits a minimal
-amount of time into the queue until they are scheduled for execution; (vii) upon
-execution on the worker nodes, each MPI script creates a ramdisk and sets up the
-specific AthenaMP configuration required by the wrapped payload; (viii) the
-payload is executed into dedicated processes, 16 for each work node, 1 for each
-available core; (ix) upon completion of an MPI job, the PanDA Broker locates the
-produced output files on the shared filesystem and transfer them to BNL ATLAS
-Tier 1 computing centre.
+\mtnote{Does one PanDA Broker support the submission and monitoring of one job
+at a time? Or does it support the concurrent submission and monitoring of
+multiple jobs (assuming enough backfill availability)?}
+
+Once deployed, every PanDA Broker supports the execution of MC detector
+simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's Moab scheduler
+about currently unused worker nodes; (ii) the PanDA Broker queries the Job
+Dispatcher module of the PanDA server for jobs that have been bound to Titan by
+JEDI; (iii)  the PanDA Broker requests to the Data Service module of the PanDA
+Server what input data these jobs require and transfer them to the DTN via the
+ATLAS DDM service; (iv) the PanDA Broker creates one or more PBS MPI script,
+wrapping enough jobs' payload to fit backfill availability; (v) the PanDA Broker
+submits these MPI jobs to the Titan's PBS batch system via RADICAL-SAGA; (vi)
+the MPI jobs waits a minimal amount of time into the queue until they are
+scheduled for execution; (vii) upon execution on the worker nodes, each MPI
+script creates a ramdisk and sets up the specific AthenaMP configuration
+required by the wrapped payload; (viii) the payload is executed into dedicated
+processes, 16 for each work node, 1 for each available core; (ix) upon
+completion of an MPI job, the PanDA Broker locates the produced output files on
+the shared filesystem and transfer them to BNL ATLAS Tier 1 computing centre.
 

### 2017-03-13 nge.tex
Question IV; transition V-VI

@@
 
+Multiple generations would also help with the transition from an execution model
+based on a fixed amount of event per simulation (100 at the moment) to one in
+which the number of events is decided on the base of the available walltime.
+Pilots would enable to package simulations with a homogeneous number of events,
+especially when core and walltime availabilities would diverge as when a small
+amount of cores is available for a long amount of time or \textit{vice versa}.
+In this situations, a PBS job would would require all simulation to have either
+a large or a small number of events; a pilot would instead support multiple
+generations of simulations with a different number of events and therefore
+duration.
+
 Pilots can offer a payload-independent scheduling interface while hiding the
@@
 enable the submission of different types of payload, without having to develop a
-specific PBS script for every payload. Finally, the submission process would be
+specific PBS script for each payload. The submission process would also be
 MPI-independent, as MPI is used for coordination among multiple work nodes, not
@@
 
-\mtnote{NOTE: I do not mention backfill availability Vs normal availability
-(i.e., submitting to the queue a `normal' job) because, in theory, the design of
-PanDA Broker enables this capability. I would speak about about backfill/not
-backfill in the discussion of NGE, when speaking about its generality. Part of
-this generality is being able to submit to whatever batch system is supported by
-SAGA, including Titan's PBS and all its queues.}
+\mtnote{I tried to argue that pilots facilitates the submission to a normal
+queue but I was not able to find a convincing argumentation. When we consider
+only Titan, a PBS job is the only way to submit a job to its queue(s). With NGE,
+we translate a pilot description to a PBS job and then submit it to a queue via
+a PBS command. PanDA Brokers do something very similar: they package detector
+simulations into a PBS job and they submit it via a PBS command. The fact that
+PanDA Brokers define the number of cores and walltime of a PBS job on the base
+of backfill information, does not make the packaging or the submission processes
+any different: PanDA Brokers are already submitting to the nomal queue. If I got
+it right, PanDA Brokers could already submit a job with, say, 12000 cores and 2
+hours walltime. They don't do it because they have no allocation and they would
+wait a long time in the queue (maybe, let's see what our experiments say about
+that). I do not see how NGE could make all this any different.}
 
-\mtnote{NOTE: NGE is not able to use backfill on Titan as we do not have specific functionalities to interrogate the Moab scheduler about backfill availability.
-This is why NGE should be presented as the pilot for PanDA Broker and not as an alternative to PanDA Broker.
-NGE alone would not be able to speak directly to PanDA Server.}
+\mtnote{NOTE: I would speak about about backfill/not backfill in the discussion
+of NGE, when speaking about its generality towards resources, i.e., unified
+submission and scheduling process across different resources and multiple
+resources. Part of this generality is being able to submit to whatever batch
+system is supported by SAGA (including Titan's PBS and all its queues) and, in
+case, to multiple resources at the same time.}
+
+\mtnote{NOTE: NGE is not able to use backfill on Titan as we do not have
+specific functionalities to interrogate the Moab scheduler about backfill
+availability. This is why NGE should be presented as the pilot for PanDA Broker
+and not as an alternative to PanDA Broker. Further, NGE alone would not be able
+to speak directly to PanDA Server.}
 

### 2017-03-16 PandaEvolution.tex
Iteration 4 and 5 adding Sergey's data

@@
 
+% -----------------------------------------------------------------------------
 \subsection{Use of Supercomputers with PanDA}
+\label{ssec:panda-supercomputers}
 
@@
 \subsection{Interfacing PanDA with Titan}
-\label{ssec:panda_titan}
+\label{ssec:panda-titan}
 
@@
 
-\mtnote{Does one PanDA Broker support the submission and monitoring of one job
-at a time? Or does it support the concurrent submission and monitoring of
-multiple jobs (assuming enough backfill availability)?}
-
 Once deployed, every PanDA Broker supports the execution of MC detector
-simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's Moab scheduler
-about currently unused worker nodes; (ii) the PanDA Broker queries the Job
+simulation on Titan in 8 steps: (i) PanDA Broker queries Titan's Moab scheduler
+about current backfill availability; (ii) the PanDA Broker queries the Job
 Dispatcher module of the PanDA server for jobs that have been bound to Titan by
 JEDI; (iii)  the PanDA Broker requests to the Data Service module of the PanDA
-Server what input data these jobs require and transfer them to the DTN via the
-ATLAS DDM service; (iv) the PanDA Broker creates one or more PBS MPI script,
-wrapping enough jobs' payload to fit backfill availability; (v) the PanDA Broker
-submits these MPI jobs to the Titan's PBS batch system via RADICAL-SAGA; (vi)
-the MPI jobs waits a minimal amount of time into the queue until they are
-scheduled for execution; (vii) upon execution on the worker nodes, each MPI
-script creates a ramdisk and sets up the specific AthenaMP configuration
-required by the wrapped payload; (viii) the payload is executed into dedicated
-processes, 16 for each work node, 1 for each available core; (ix) upon
-completion of an MPI job, the PanDA Broker locates the produced output files on
-the shared filesystem and transfer them to BNL ATLAS Tier 1 computing centre.
+Server what input data these jobs require and transfer the files to the DTN via
+the ATLAS DDM service; (iv) the PanDA Broker creates a PBS MPI script, wrapping
+enough jobs' payload to fit backfill availability; (v) the PanDA Broker submits
+the PBS MPI script to the Titan's PBS batch system via RADICAL-SAGA; (vi) upon
+execution on the worker node(s), the MPI script creates a ramdisk and configure
+AthenaMP to execute the wrapped payload; (vii) the payload is executed into
+dedicated processes, 16 for each work node, 1 for each available core; (viii)
+upon completion of the MPI script, the PanDA Broker locates the output on the
+shared NFS filesystem and transfer it to BNL ATLAS Tier 1 computing centre.
 

### 2017-03-16 overall.tex
Iteration 4 and 5 adding Sergey's data

@@
 along two dimensions: (i) the amount of backfill availability utilized by the
-PanDA Brokers; and (ii) the amount of 
-runtime spent performing detector simulations. We based these evaluations on measuring the number
-of cores utilized by ATLAS on Titan, the overall backfill availability, the
-number of detector simulations performed on all the resources available to ATLAS
-and on Titan, and the failure rate of these simulations. All the measurements
-were performed between January 2016 and February 2017.
+PanDA Brokers; and (ii) the amount of runtime spent performing detector
+simulations. We based these evaluations on measuring the number of cores
+utilized by ATLAS on Titan, the overall backfill availability, the number of
+detector simulations performed on all the resources available to ATLAS and on
+Titan, and the failure rate of these simulations. All the measurements were
+performed between January 2016 and February 2017.
 
 % -----------------------------------------------------------------------------
-\subsection{Efficiency and Scalability of PanDA Broker on Titan}
+\subsection{Utilization and Efficiency of PanDA Broker on Titan}
 \label{ssec:panda_titan}
@@
 ATLAS on Titan, from January 2016 to February 2017. During that period, ATLAS
-consumed a total of 73.8M core-hours, for an average of $\approx$ 7M core-hours a month,
-with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours
-in February 2017.
+consumed a total of 73.8M core-hours, for an average of $\approx$ 7M core-hours
+a month, with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M
+core-hours in February 2017.
 
@@
 Figure~\ref{fig:backfill-utilization} shows backfill utilization from January
-2016 to February 2017 (hereafter called experiment time window). \jhanote{Neither figure show efficiency. Both show
-utilization.} Efficiency is defined as the fraction of Titan’s cores available
-via backfill that was utilized by the PanDA Brokers. \jhanote{I would plot the
-efficiency on the Y2 axis for both.} The brokers reached 18\% average efficiency
-when measured across the experiment time window,
-with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\%
-efficiency on February 2017. The number of total backfill cores available was
-38.1M in April 2016, and 33.1M in February 2017. This shows that the measured
-difference in efficiency did not depend on a comparable difference in total
-backfill availability.
+2016 to February 2017 (hereafter called experiment time window).
+\jhanote{Neither figure show efficiency. Both show utilization.} Efficiency is
+defined as the fraction of Titan’s cores available via backfill that was
+utilized by the PanDA Brokers. \jhanote{I would plot the efficiency on the Y2
+axis for both.} The brokers reached 18\% average efficiency when measured across
+the experiment time window, with a minimum 8.9\% efficiency on April 2016 and a
+maximum 33.5\% efficiency on February 2017. The number of total backfill cores
+available was 38.1M in April 2016, and 33.1M in February 2017. This shows that
+the measured difference in efficiency did not depend on a comparable difference
+in total backfill availability.
 
@@
 All the detector simulations executed on Titan process 100 events. This number
-has been chosen by observing the average walltime of backfill resources: too
-many events would \ldots\mtnote{I need to speak further to Sergey to clarify
-this.}.
+of events is consistent with the physics of the use case and with the average
+duration of backfill availability. The duration of a detector simulation is a
+function of the number of events: the more events, the longer the simulation.
+Not all events take the same time to be simulated:
+Figure~\ref{fig:100event-distrib} shows the distribution of that duration for
+XXXXX events simulated on Titan. When compared to the distribution of backfill
+availability on Titan in Figure~\ref{fig:backfill-distrib}, 100 events appear to
+be a good compromise.
+
+\begin{figure}[htp]
+    \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
+\caption{Distribution of the time taken by a Geant4 detector simulation to simulate one event.
+3000 Events; 30 Titan work nodes; 16 simulations per node; 100 events per node.
+Average time per event \~14 min.
+Broad distribution from \~1 to \~40 minutes.}
+\label{fig:100event-distrib}
+\end{figure}
+
+\begin{figure}[htp]
+    \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
+\caption{\mtnote{Placeholder for the distribution of backfill walltime availability on Titan.}}
+\label{fig:backfill-distrib}
+\end{figure}
 
@@
 of detector simulations (9\%) and of events computed by those simulations
-(3.9\%) on Titan. PanDA Event Service~\cite{panda-event-service} will contribute
-to address this limitation: each PanDA Broker will be able to pull a variable
-number of events for each simulation, depending on the available walltime. This
-functionality is currently being implemented.
+(3.9\%) on Titan. PanDA Broker could fit the number of events to the walltime of
+the backfill availability on the base of the distribution of
+Figure~\ref{fig:backfill-distrib}. That specific number of event could then be
+pulled from the PanDA Event service~\cite{panda-event-service} and given as
+input to one or more simulations. Once packaged into a MPI PBS job, these
+simulations would better fit backfill availability, increasing the efficiency of
+PanDA Brokers.
 
-Given the current design and implementation, the utilization of a larger portion
-of Titan's backfill availability depends on the the number of concurrent PanDA
-Brokers instantiated on the DTNs.
+The transition from a homogeneous to a heterogeneous number of events per
+detector simulation has implications for the application layer. An even number
+of events across simulations makes it easier to partition, track and package
+events across simulations, especially when they are performed on both the Grid
+and Titan. A homogeneous number of events also helps to keep the size and
+duration of other stages of the ATLAS MC workflow (\S\ref{ssec:panda-titan})
+more uniform. Further analysis is needed to evaluate the trade offs between
+increased efficiency of resource utilization and the complexity that would be
+introduced at the application layer.
+
+Currently, each PanDA Broker creates, submits, and monitor a single MPI PBS
+script at a time. This design is inherited from PanDA Pilot where a single
+process is spawn at a time to execute the payload. As a consequence, the
+utilization of a larger portion of Titan's backfill availability depends on the
+the number of concurrent PanDA Brokers instantiated on the DTNs: When all the 20
+PanDA Brokers have submitted a MPI PBS Job, further backfill availability cannot
+be used.
+
+Based on the current data, increasing PanDA Brokers efficiency on Titan would
+require around 60 concurrent brokers. Further testing and better monitoring
+capabilities are required to establish whether the DTN capabilities can support
+this load, especially when considering the volume of files staged in and out the
+DTNs. Alternatively, a design of a PanDA Broker capable of managing multiple MPI
+scripts at a time is being evaluated.
+
 % The number of concurrent detector simulation jobs that PanDA Broker can manage
@@
 % DTN resources are saturated, on multiple DTNs.
-In principle, this is not a problem on Titan as multiple DTNs are supposed to be
-available and periodically upgraded. Assuming the efficiency reached on February
-as reference, using all the available backfill resources on Titan would require
-around 60 concurrent PanDA Brokers. Further testing and better monitoring
-capabilities are required to establish whether the current DTN capabilities can
-support this load, especially when considering the increased volume of files
-staged in and out the DTNs.
-
-\mtnote{Extend/change previous paragraph explaining also the problem of the
-`spike'/`wave' of backfill availability. The problem without enough brokers to
-cope with variable amounts of backfill availability is that we may exhaust all
-20 brokers to use a portion of backfill smaller than the one that becomes
-available when no more brokers are available.}
 
@@
 
-At runtime, detector simulation jobs showed performance degradation when all the
-16 cores of Titan's work nodes were used. As shows in
-Figure~\ref{}\mtnote{Sergery: do you have a diagram showing the difference in
-performance between using 16 and 8 cores for a detector simulation job with 100
-events?}, the mean runtime for 8 concurrent jobs computing 100 events is XXXs,
-with each job executed in 8 distinct processes, one per core. The mean runtime
-for 16 concurrent jobs computing 100 events is instead XXXs. This can be
-explained by looking at the cache architecture of the AMD Opteron 6274 CPU. In
-this processor, every two cores share the level 1 instruction cache, impacting
-the performance of \ldots\mtnote{Do we kow what instruction set is affected?}
-computation.
+The AMD Opteron 6274 CPU used on Titan has 16 cores, divided into 8 compute
+units. Each compute units has 1 floating point (FP) scheduler shared between 2
+cores. When using 16 cores for FP-intensive calculations, each pair of cores
+competes for a single FP scheduler. This creates the overhead shown in
+Figure~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
+concurrent simulations computing 50 events is 10.8 minutes, while for 16
+simulations is 14.25 minutes (consistent with
+Figure~\ref{fig:100event-distrib}). The allocation policy based on work nodes
+and not cores used, does not justify the use of 1/2 of the cores available.
+
+\begin{figure}[htp]
+    \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison.pdf}
+\caption{Comparison between distributions of the time taken by a Geant4 detector simulation to simulate one event when placing 2 simulations (h1) or 1 simulation (h2) per CPU.
+2 simulation use 16 cores per node, 1 simulation 8, 1 per compute unit.
+50 Events; 1 Titan work nodes; 16 work threads per node; 100 events per node.
+Average time per event \~14 min.
+Broad distribution from \~1 to \~40 minutes.}
+\label{fig:comparison-8-16cores}
+\end{figure}
 

### 2017-03-17 PandaEvolution.tex
Iteration and shortening IV and V; add new data from Sergey

@@
 The next LHC run for taking data (Run 3) will require more resources than the
-Worldwide LHC Computing Grid (WLCG) can possibly provide. Currently, PanDA WMS
-uses more than 100,000 cores at over 100 Grid sites, with a peak performance of
-0.3 petaFLOPS. This capacity will be sufficient for the planned analysis and
-data processing, but it will be insufficient for the Monte Carlo production
-workflow and any extra activity. To alleviate these challenges, ATLAS is engaged
-in a program to expand the current computing model to include additional
-resources such as the opportunistic use of supercomputers as well as commercial
-and academic clouds.
+Worldwide LHC Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more
+than 100,000 cores at over 100 Grid sites, with a peak performance of 0.3
+petaFLOPS. This capacity will be sufficient for the planned analysis and data
+processing, but it will be insufficient for the Monte Carlo production workflow
+and any extra activity. To alleviate these challenges, ATLAS is engaged in a
+program to expand the current computing model to include additional resources
+such as the opportunistic use of supercomputers as well as commercial and
+academic clouds.
 
@@
 a high-speed, low-latency dedicated network. Each worker node has multicore
-CPUs, usually augmented with massively parallel Graphics Processing Units (GPUs)
-or other types of specialized coprocessors.
+CPUs, usually augmented with parallel Graphics Processing Units (GPUs) or other
+types of specialized coprocessors.
 
-PanDA WMS has been designed to support distributed Grid
-computing\sergeynote{specifically "distributed Grid computing"}\mtnote{done.}.
-Executing ATLAS workloads or workflows involves concurrent and/or sequential
-runs of possibly large amount of jobs, each requiring no or minimal
-parallelization and no runtime communication. Thus, computing infrastructure
-like WLCG have been designed to aggregate large amount of computing resources
-across multiple sites.
-%, often geographically dislocated.
-\sergeynote{geographically distributed?}\mtnote{Better?} While each site may
-deploy runtime message-passing capabilities, usually these are not used to
-perform distributed computations.
+% \sergeynote{specifically "distributed Grid computing"}\mtnote{done.}.
+% \sergeynote{geographically distributed?}\mtnote{Better?}
+
+PanDA WMS has been designed to support distributed Grid computing. Executing
+ATLAS workloads or workflows involves concurrent and/or sequential runs of
+possibly large amount of jobs, each requiring no or minimal parallelization and
+no runtime communication. Thus, computing infrastructure like WLCG have been
+designed to aggregate large amount of computing resources across multiple sites.
+While each site may deploy runtime message-passing capabilities, usually these
+are not used to perform distributed computations.
 
@@
 
-The Titan supercomputer, current number three
-% (number one from November 2012 until June 2013)
-on the Top 500 list~\cite{top500}, is a Cray XK7 system with 18,688 work nodes
-and a total of 299,008 CPU cores. Each work node has an AMD Opteron CPU, a
-Nvidia Tesla GPU, 32 GB of RAM and no local storage, though a 16 GB RAM disk can
-be set up. Work nodes use Cray’s Gemini interconnect for inter-node MPI
-messaging. Titan is served by the Spider II~\cite{spider2}, a Lustre filesystem
-with 32 PB of disk storage, and by a 29 PB HPSS tape storage system. Titan’s
-work nodes run Compute Node Linux, a run time environment based on SUSE Linux
-Enterprise Server.
+The Titan supercomputer, current number three on the Top 500 list~\cite{top500},
+is a Cray XK7 system with 18,688 worker nodes and a total of 299,008 CPU cores.
+Each worker node has an AMD Opteron  6274 16-core CPU, a Nvidia Tesla K20X GPU,
+32 GB of RAM and no local storage, though a 16 GB RAM disk can be set up. Work
+nodes use Cray’s Gemini interconnect for inter-node MPI messaging. Titan is
+served by the Spider II~\cite{spider2}, a Lustre filesystem with 32 PB of disk
+storage, and by a 29 PB HPSS tape storage system. Titan’s worker nodes run
+Compute Node Linux, a run time environment based on SUSE Linux Enterprise
+Server.
 
@@
 % utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla
-% K20 GPU accelerators.
-% This hybrid design provides improved energy efficiency, as well as an order of
-% magnitude in computational capacity over its predecessor.
+% K20 GPU accelerators. This hybrid design provides improved energy efficiency,
+% as well as an order of magnitude in computational capacity over its
+% predecessor.
 
-Titan offers to its users login nodes and data transfer nodes (DTNs). Users have
-to log into one of these nodes via ssh to submit jobs to Titan's PBS scheduler.
-Titan's authentication and authorization model is based on two-factor
-authentication with a RSA SecurID key, generated every 30 seconds. Login nodes
-and DTNs are the only nodes with out/inbound wide area network connectivity.
-Work nodes have only local network access. Data staging is supported via the
-DTNs, from the wide area network and the OLCF high performance storage system
-(HPSS). DTNs are shared across all Titan's users that can schedule jobs to a
-dedicates batch system to automate data staging. Fair-share policies are in
-place both for the batch system and for the use of DTN resources.
+Titan's users submit jobs to Titan's PBS scheduler by logging into login or data
+transfer nodes (DTNs). Titan's authentication and authorization model is based
+on two-factor authentication with a RSA SecurID key, generated every 30 seconds.
+Login nodes and DTNs have out/inbound wide area network connectivity while
+worker nodes have only local network access. Fair-share and allocation policies
+are in place both for the PBS batch system and shared file systems.
+
+% Data staging is supported via the DTNs, from the wide area network and from
+% the OLCF high performance storage system (HPSS). DTNs are shared across all
+% Titan's users that can schedule jobs to a dedicates batch system to automate
+% data staging. Fair-share policies are in place both for the batch system and
+ %for the use of DTN resources.
 
 Titan's architecture, configuration and policies poses several challenges to the
-integration with PanDA. PanDA Pilot requires to contact the Job Dispatcher
-module of the PanDA Server to pull jobs to execute. Titan's work nodes do not
-offer outbound network connectivity, making the current deployment model of
-PanDA pilots unfeasible. Titan does not support PanDA's security model based on
-certificates and virtual organizations, making the PanDA's approach to identity
-management also unfeasible. While DTNs offer wide area network data transfer, an
-integration with ATLAS DDM is beyond the functional and administrative scope of
-the current prototyping phase. Finally, the specific characteristics of the
-execution environment, especially the absence of local storage on the work nodes
-and modules tailored to Compute Node Linux, require reengineering of ATLAS
-application frameworks.
+integration with PanDA. The deployment model of PanDA Pilot is unfeasible on
+Titan: PanDA Pilot requires to contact the Job Dispatcher of the PanDA Server to
+pull jobs to execute but this is not possible on Titan because worker nodes do
+not offer outbound network connectivity. Further, Titan does not support PanDA's
+security model based on certificates and virtual organizations, making the
+PanDA's approach to identity management also unfeasible. While DTNs offer wide
+area network data transfer, an integration with ATLAS DDM is beyond the
+functional and administrative scope of the current prototyping phase. Finally,
+the specific characteristics of the execution environment, especially the
+absence of local storage on the worker nodes and modules tailored to Compute
+Node Linux, require reengineering of ATLAS application frameworks.
 
@@
 the Grid to Titan's large amount of cores. Further, when HEP tasks can be
-partitioned to independent jobs, Titan work nodes can be used to execute up to
+partitioned to independent jobs, Titan worker nodes can be used to execute up to
 16 concurrent jobs' payload, one for each available core. Given these
-constraints and challenges, the type of task most suitable for execution at the moment on Titan is Monte Carlo
+constraints and challenges, the type of task most suitable for execution at the
+moment on Titan is Monte Carlo
 % event generation and
@@
 
-% The event generation in ATLAS accounts for 10--15\% of all jobs on WLCG,
-% corresponding to 2/3 of the current Grid capacity.
-
-% Event generation and
 Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
@@
 consists of four main stages: event generation, detector simulation,
-digitization, and reconstruction. Even generation creates sets of particle
+digitization, and reconstruction. Event generation creates sets of particle
 four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia}
@@
 
-The lack of wide area network connectivity on Titan's work nodes is the most
+The lack of wide area network connectivity on Titan's worker nodes is the most
 relevant challenge for integrating PanDA WMS and Titan. Without connectivity,
-Panda Pilots cannot be scheduled on work nodes because they would not be able to
-communicate with PanDA Server and therefore pull and execute jobs. This makes
+Panda Pilots cannot be scheduled on worker nodes because they would not be able
+to communicate with PanDA Server and therefore pull and execute jobs. This makes
 impossible to port PanDA Pilot to Titan while maintaining the defining feature
@@
 The unavailability of pilots is a potential drawback when executing distributed
-workloads like MC
-% event generation and
-detector simulation. Pilots are used to increase the throughput of distributed
-workloads: while pilots have to wait in the supercomputer's queue, once
-scheduled, they can pull and execute jobs independently from the system's queue.
-Jobs can be concurrently on every core available to the pilot and multiple
-generations of concurrent executions can be performed until the pilot's walltime
-is exhausted. This is particularly relevant for machines like Titan where queue
-policies privilege parallel jobs on the base of the number of work nodes they
-request: the higher the number of nodes, the shorter the amount of queue time
-(modulo fair share policies).
+workloads like MC detector simulation. Pilots are used to increase the
+throughput of distributed workloads: while pilots have to wait in the
+supercomputer's queue, once scheduled, they can pull and execute jobs
+independently from the system's queue. Jobs can be concurrently executed on
+every core available to the pilot, and multiple generations of concurrent
+executions can be performed until the pilot's walltime is exhausted. This is
+particularly relevant for machines like Titan where queue policies privilege
+parallel jobs on the base of the number of worker nodes they request: the higher
+the number of nodes, the shorter the amount of queue time (modulo fair-share and
+allocation policies).
 
@@
 queue wait times without using pilot abstraction. Backfill availability is the
-number of work nodes that cannot be used by any of the jobs already queued on
-Titan: All queued jobs are either too large or their walltime is too long. At
-any point in time, Titan’s Moab scheduler can be queried for backfill
-availability. Based upon the result of this query, a job can be shaped to
-request no more than the backfill availability, and when submitted as such is
-usually scheduled immediately spending almost no time in the queue.
+number of worker nodes that cannot be used for a certain amount of time by any
+of the jobs already queued on Titan: All queued jobs are either too large or
+their walltime is too long. At any point in time, Titan’s Moab scheduler can be
+queried for backfill availability. Based upon the result of this query, a job
+can be shaped to request no more than the backfill availability. As such, when
+submitted this job is usually scheduled immediately, spending almost no time in
+the queue.
 
@@
 nodes that can be requested. Pilots are normal jobs: they can request as many
-work nodes for as much time as a queue can offer. On the contrary, jobs sized on
-the basis of backfill information availability \jhanote{should be resource
+worker nodes for as much time as a queue can offer. On the contrary, jobs sized
+on the basis of backfill information availability \jhanote{should be resource
 availability or backfill queue information?}\mtnote{AFAIH, there is no dedicated
-queue to backfill} depend on the number % amount
-of work nodes that cannot be given to any other job. any other job in the
-Titan's queue\sergeynote{any other job currently in the Titan's
-queue?}\mtnote{Better?}. Usually, backfill availability is a fraction of the
-total capacity of the queue but the size of Titan mitigates this limitation.
-Every year, about 10\% of Titan's capacity remains
-unused~\cite{titan_utilization}, corresponding to an average of 30,000 unused
-cores. This equals approximately 270M core hours per year, roughly 30\% of the
-overall capacity of WLCG.
+queue to backfill} depend on the number of worker nodes that cannot be given to
+any other job in the Titan's queue.
+% \sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.
 
-% A system that can use those temporarily free work nodes with smaller
+Usually, backfill availability is a fraction of the total capacity of the queue
+but the size of Titan mitigates this limitation. Every year, about 10\% of
+Titan's capacity remains unused~\cite{titan_utilization}, corresponding to an
+average of 30,000 unused cores. This equals approximately 270M core hours per
+year, roughly 30\% of the overall capacity of WLCG.
+
+% A system that can use those temporarily free worker nodes with smaller
 % scale tasks would be very valuable. For this reason, Titan offers backfill
@@
 Broker' implements functionalities to: (i) interrogate Titan about backfill
-availability; (ii) pull MC jobs and events; (iii) wrap jobs' payload into PBS
-MPI jobs; (iv) submitting MPI PBS jobs to Titan's PBS batch system and monitor
-their execution;\sergeynote{Monitor job state and state transitions using
-SAGA}\mtnote{Better?} and (v) staging input/output files. Backfill querying,
-payload wrapping, and job submission required a new implementation while job and
-events pulling, and file staging were mostly inherited from PanDA Pilot.
+availability; (ii) pull MC jobs and events; (iii) wrap the payload of ATLAS jobs
+into MPI scripts; (iv) submitting MPI scripts to Titan's PBS batch system and
+monitor their execution; and (v) staging input/output files. Backfill querying,
+payload wrapping, and scripts submission required a new implementation while
+pulling ATLAS job and events, and file staging were mostly inherited from PanDA
+Pilot.
+
+% \sergeynote{Monitor job state and state transitions using
+% SAGA}\mtnote{Better?}
 
@@
 
-Backfill querying is performed via a dedicated Moab command while a tailored
-Python MPI script is used to wrap job's payload. This MPI wrapper serves as a
-``container'' for non-MPI jobs, enabling execution of unmodified Grid-centric
-jobs on Titan. Typically, a MPI wrapper is workload-specific as it sets up the
-execution environment for a specific payload. This involves organization of
-worker directories, data management, optional input parameters modification, and
-cleanup on exit. Upon submission, a copy of the MPI wrapper runs on every
-available work node. Each script knows its MPI rank (an index that runs from
-zero to a maximum number of nodes or script copies) as well as the total number
-of ranks in the current submission. When activated on worker nodes, each copy of
-the wrapper script starts the execution of the payload in a subprocess and waits
+Backfill querying is performed via a dedicated Moab scheduler command while a
+tailored Python MPI script is used to execute the payload of ATLAS jobs. This
+MPI script enables the execution of unmodified Grid-centric, ATLAS jobs on
+Titan. Typically, a MPI script is workload-specific as it sets up the execution
+environment for a specific payload. This involves organization of worker
+directories, data management, optional input parameters modification, and
+cleanup on exit. Upon submission, a copy of the MPI script runs on every
+available worker node,
+% Each script knows its MPI rank (an index that runs from zero to a maximum
+% number of nodes or script copies) as well as the total number of ranks in the
+% current submission. When activated on worker nodes, each copy of the MPI
+% script
+starting the execution of the ATLAS job's payload in a subprocess and waits
 until its completion.
 
-MPI wrappers are submitted to Titan's PBS batch system via
+MPI scripts are submitted to Titan's PBS batch system via
 RADICAL-SAGA~\cite{radical-saga}, a Python module, compliant with the OGF GFD.90
@@
 RADICAL-SAGA implements a adaptor architecture: each adaptor interface the SAGA
-API with different middleware systems and services, including the batch
+API with different middleware systems and services, including the PBS batch
 scheduler of Titan.
@@
 The data staging capabilities of the PanDA Broker are implemented via a file
-system that is shared among DTNs and work nodes. The input files required by the
-payload are downloaded via the ATLAS DDM service and then stored in a directory
-on the shared filesystem. The MPI wrapper setup process includes making the
-location of these files available to the payload. The PanDA Broker can locate
-the Payloads' output files on the shared filesystem and transfer them from Titan
-to the ATLAS Tier 1 computing center at Brookhaven National Lab (BNL) or any
-other Grid site.
+system that is shared among DTNs and worker nodes. The input files with the
+events of the ATLAS jobs are downloaded on the shared filesystem via the ATLAS
+DDM service. The MPI script setup process includes making the location of these
+files available to the payload of the ATLAS's jobs. The PanDA Broker can locate
+the payload's output files on the shared filesystem and transfer them from Titan
+to any computing center used by ATLAS.
+% the ATLAS Tier 1 computing center at Brookhaven National Lab (BNL) or any
+% other Grid site.
 
-Once deployed, every PanDA Broker supports the execution of MC detector
-simulation on Titan in 8 steps: (i) PanDA Broker queries Titan's Moab scheduler
-about current backfill availability; (ii) the PanDA Broker queries the Job
-Dispatcher module of the PanDA server for jobs that have been bound to Titan by
-JEDI; (iii)  the PanDA Broker requests to the Data Service module of the PanDA
-Server what input data these jobs require and transfer the files to the DTN via
-the ATLAS DDM service; (iv) the PanDA Broker creates a PBS MPI script, wrapping
-enough jobs' payload to fit backfill availability; (v) the PanDA Broker submits
-the PBS MPI script to the Titan's PBS batch system via RADICAL-SAGA; (vi) upon
-execution on the worker node(s), the MPI script creates a ramdisk and configure
-AthenaMP to execute the wrapped payload; (vii) the payload is executed into
-dedicated processes, 16 for each work node, 1 for each available core; (viii)
-upon completion of the MPI script, the PanDA Broker locates the output on the
-shared NFS filesystem and transfer it to BNL ATLAS Tier 1 computing centre.
+Once deployed on Titan, every PanDA Broker supports the execution of MC detector
+simulation in 8 steps: (i) PanDA Broker queries Titan's Moab scheduler about
+current backfill availability; (ii) the PanDA Broker queries the Job Dispatcher
+module of the PanDA server for ATLAS jobs that have been bound to Titan by JEDI;
+(iii) the PanDA Broker requests to the Data Service module of the PanDA Server
+what input data these ATLAS jobs require and transfer the files to the DTN via
+the ATLAS DDM service; (iv) the PanDA Broker creates an MPI script, wrapping
+enough ATLAS jobs' payload to fit backfill availability; (v) the PanDA Broker
+submits the MPI script to the Titan's PBS batch system via RADICAL-SAGA; (vi)
+upon execution on the worker node(s), the MPI script creates a ramdisk and
+configure AthenaMP to execute the ATLAS jobs' payload; (vii) the payload is
+executed into dedicated processes, 16 for each worker node, 1 for each available
+core; (viii) upon completion of the payload execution, the PanDA Broker locates
+the output on the shared filesystem and transfer it to designated computing
+centre.
 

### 2017-03-17 overall.tex
Iteration and shortening IV and V; add new data from Sergey

@@
 instances per DTN. Each broker submits and manages the execution of 15 to 300
-detector simulation jobs, one job for each work node, and a theoretical maximum
-concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have operated
-only in backfill mode, without a defined time allocation, and running at the
-lowest priority on Titan.
+ATLAS jobs, one job for each worker node, and a theoretical maximum concurrent
+use of 96,000 cores. Since November 2015, PanDA Brokers have operated only in
+backfill mode, without a defined time allocation, and running at the lowest
+priority on Titan.
 
@@
 Titan, and the failure rate of these simulations. All the measurements were
-performed between January 2016 and February 2017.
+performed between January 2016 and February 2017, hereafter called `experiment
+time window'.
 
@@
 Figure~\ref{fig:core-hours-utilization} shows the number of core-hours used by
-ATLAS on Titan, from January 2016 to February 2017. During that period, ATLAS
-consumed a total of 73.8M core-hours, for an average of $\approx$ 7M core-hours
-a month, with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M
-core-hours in February 2017.
+ATLAS on Titan during the experiment time window. ATLAS consumed a total of
+73.8M core-hours, for an average of $\approx$ 7M core-hours a month, with a
+minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours in
+February 2017.
 
@@
 
-Figure~\ref{fig:backfill-utilization} shows backfill utilization from January
-2016 to February 2017 (hereafter called experiment time window).
-\jhanote{Neither figure show efficiency. Both show utilization.} Efficiency is
-defined as the fraction of Titan’s cores available via backfill that was
-utilized by the PanDA Brokers. \jhanote{I would plot the efficiency on the Y2
-axis for both.} The brokers reached 18\% average efficiency when measured across
-the experiment time window, with a minimum 8.9\% efficiency on April 2016 and a
-maximum 33.5\% efficiency on February 2017. The number of total backfill cores
-available was 38.1M in April 2016, and 33.1M in February 2017. This shows that
-the measured difference in efficiency did not depend on a comparable difference
-in total backfill availability.
+Figure~\ref{fig:backfill-utilization} shows backfill utilization during the
+experiment time window.\jhanote{Neither figure show efficiency. Both show
+utilization.} Efficiency is defined as the fraction of Titan’s cores available
+via backfill that was utilized by the PanDA Brokers. \jhanote{I would plot the
+efficiency on the Y2 axis for both.} The brokers reached 18\% average
+efficiency, with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\%
+efficiency on February 2017. The number of total backfill cores available was
+38.1M in April 2016, and 33.1M in February 2017. This shows that the measured
+difference in efficiency did not depend on a comparable difference in total
+backfill availability.
 
@@
 
-Figure~\ref{fig:hpc-workload-utilization} shows that between January 2016 and
-February 2017, about 2.25M detector simulation jobs were completed on Titan, for
-a total of 225M events processed. This is equivalent to 0.9\% of all the 250M
+Figure~\ref{fig:hpc-workload-utilization} shows that during the experiment time
+window, about 2.25M detector simulation jobs were completed on Titan, for a
+total of 225M events processed. This is equivalent to 0.9\% of all the 250M
 detector simulations performed by ATLAS in the same period of time, and 3.5\% of
@@
 availability. This is likely due to hardware upgrades on the DTNs. The absence
-of continuous monitoring of the nodes does not allow to quantify bottlenecks but
-spot measurements of the DTNs load indicate that a faster CPU and better
+of continuous monitoring of those nodes does not allow to quantify bottlenecks
+but spot measurements of their load indicate that a faster CPU and better
 networking were likely responsible for the improved performance.
@@
 
-All the detector simulations executed on Titan process 100 events. This number
-of events is consistent with the physics of the use case and with the average
+Every detector simulation executed on Titan process 100 events. This number of
+events is consistent with the physics of the use case and with the average
 duration of backfill availability. The duration of a detector simulation is a
@@
 Not all events take the same time to be simulated:
-Figure~\ref{fig:100event-distrib} shows the distribution of that duration for
-XXXXX events simulated on Titan. When compared to the distribution of backfill
-availability on Titan in Figure~\ref{fig:backfill-distrib}, 100 events appear to
-be a good compromise.
+Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time of
+1 event from $\approx$ 2 to $\approx$ 40 minutes, with a mean of $\approx$ 14
+minutes. Considering that each worker node process up to 16 events concurrently,
+100 events takes around 2 hours to process.
 
@@
 \caption{Distribution of the time taken by a Geant4 detector simulation to simulate one event.
-3000 Events; 30 Titan work nodes; 16 simulations per node; 100 events per node.
-Average time per event \~14 min.
-Broad distribution from \~1 to \~40 minutes.}
+3000 Events; 30 Titan worker nodes; 16 simulations per node; 100 events per node.
+Average time per event $\approx$ 14 min.
+Broad distribution from $\approx$ 2 to $\approx$ 40 minutes.}
 \label{fig:100event-distrib}
@@
 
+Figure~\ref{fig:backfill-distrib} shows the distribution of backfill
+availability on Titan as a function of number of nodes and the time of their
+availability (i.e., walltime). The mean number of nodes over the experiment
+window time was $\approx$ 267, and their mean walltime was 1445 minutes.
+Detector simulations of 100 events, enable to use down to 1 node for 1/4 of the
+mean walltime of backfill availability. As such, it offers a good compromise for
+the PanDA Broker efficiency.
+
 \begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
-\caption{\mtnote{Placeholder for the distribution of backfill walltime availability on Titan.}}
+    \includegraphics[clip,width=\columnwidth]{figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.png}
+\caption{\mtnote{Waiting to know whether we have data for the experiment time window.}}
 \label{fig:backfill-distrib}
@@
 
-Usually, detector simulations performed on the Grid process a higher number of
-events than those on Titan. This explains the difference between the percentage
-of detector simulations (9\%) and of events computed by those simulations
-(3.9\%) on Titan. PanDA Broker could fit the number of events to the walltime of
-the backfill availability on the base of the distribution of
+Usually, detector simulations performed on the Grid process 10 times the number
+of events than those on Titan. This explains the difference between the
+percentage of detector simulations (9\%) and of events computed by those
+simulations (3.9\%) on Titan. PanDA Broker could fit the number of events to the
+walltime of the backfill availability on the base of the distribution of
 Figure~\ref{fig:backfill-distrib}. That specific number of event could then be
 pulled from the PanDA Event service~\cite{panda-event-service} and given as
-input to one or more simulations. Once packaged into a MPI PBS job, these
-simulations would better fit backfill availability, increasing the efficiency of
-PanDA Brokers.
+input to one or more simulations. Once packaged into the MPI script submitted to
+titan's PBS batch system, these simulations would better fit backfill
+availability, increasing the efficiency of PanDA Brokers.
 
@@
 and Titan. A homogeneous number of events also helps to keep the size and
-duration of other stages of the ATLAS MC workflow (\S\ref{ssec:panda-titan})
-more uniform. Further analysis is needed to evaluate the trade offs between
-increased efficiency of resource utilization and the complexity that would be
-introduced at the application layer.
+duration of other stages of the MC workflow (\S\ref{ssec:panda-titan}) more
+uniform. Further analysis is needed to evaluate the trade offs between increased
+efficiency of resource utilization and the complexity that would be introduced
+at the application layer.
 
@@
 % depends on the DTNs' resources. Each PanDA Broker can execute 1 MPI PBS job on
-% up to 300 work nodes for a total of 4,800 cores. Scaling above this threshold
-% requires instantiating multiple PanDA Brokers on the same DTN and, once the
-% DTN resources are saturated, on multiple DTNs.
+% up to 300 worker nodes for a total of 4,800 cores. Scaling above this
+% threshold requires instantiating multiple PanDA Brokers on the same DTN and,
+% once the DTN resources are saturated, on multiple DTNs.
 
@@
 \begin{itemize}
-    \item problem with lustre addressed with ramdisk
-    \item Competition for cache in AMD processors.
-    \item Summary overall figure of performance analogous to those produced with NGE. Diagram as discussed in F2F meeting at Rutgers.
-    \item Comparison hardware performance Titan/Grid
+    \item problem with lustre addressed with ramdisk (done)
+    \item Competition for cache in AMD processors. (done)
+    \item Comparison hardware performance Titan/Grid (data acquired)
+    \item Summary overall figure of performance analogous to those produced with NGE. Diagram as discussed in F2F meeting at Rutgers. (waiting)
 \end{itemize}
@@
 uncomment it if you disagree.}
+
 % As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
@@
 performance degradation was addressed by moving the AthenaMP distribution to a
-read-only NFS directory, shared among DTNs and work nodes. This eliminated the
-problem of metadata contention, improving metadata read performance from \~6,300
-seconds on Lustre to \~1,500 seconds on NFS. Further, at configuration time, the
-input files of each detector simulation job were stored to a ramdisk on the work
-nodes. This offered a marked improvement of reading time: from 1,320 seconds on
-Lustre to 40 seconds on NFS.
+read-only NFS directory, shared among DTNs and worker nodes. This eliminated the
+problem of metadata contention, improving metadata read performance from
+$\approx$ 6,300 seconds on Lustre to $\approx$ 1,500 seconds on NFS. Further, at
+configuration time, the input files of each detector simulation job were stored
+to a ramdisk on the work nodes. This offered a marked improvement of reading
+time: from 1,320 seconds on Lustre to 40 seconds on NFS.
 
@@
 simulations is 14.25 minutes (consistent with
-Figure~\ref{fig:100event-distrib}). The allocation policy based on work nodes
-and not cores used, does not justify the use of 1/2 of the cores available.
+Figure~\ref{fig:100event-distrib}). Despite an inefficiency of almost 30\%,
+Titan's allocation policy based on number of worker nodes used instead of number
+of cores does not justify the use of 1/2 of the cores available.
 
@@
     \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison.pdf}
-\caption{Comparison between distributions of the time taken by a Geant4 detector simulation to simulate one event when placing 2 simulations (h1) or 1 simulation (h2) per CPU.
-2 simulation use 16 cores per node, 1 simulation 8, 1 per compute unit.
-50 Events; 1 Titan work nodes; 16 work threads per node; 100 events per node.
-Average time per event \~14 min.
-Broad distribution from \~1 to \~40 minutes.}
+    \caption{Comparison between distributions of the time taken by a Geant4
+    detector simulation to simulate one event when placing 2 simulations (h1) or
+    1 simulation (h2) per CPU. 2 simulation use 16 cores per node, 1 simulation
+    8, 1 per compute unit. 50 Events; 1 Titan worker nodes; 16 work threads per
+    node; 100 events per node.}
 \label{fig:comparison-8-16cores}
@@
 The compute performance of Titan for detector simulation jobs is, on average,
-lower that that of LCWG. Figure~\ref{}\mtnote{Sergey: I remember you showed us
-some plots comparing Geant4 simulation's performance on Titan and some of the
-Grid sites. Could we use one or two of the here?} shows the difference in
-execution time of the same type of simulation between Titan and
+lower that that of LCWG. Figure~\ref{fig:comparison-titan-grid} shows the
+difference in execution time of the same type of simulation between Titan and
 \ldots\mtnote{Add the name of the grid sites here}. This is not an issue with
@@
 will become active in 2018 at OLCF.
+
+\begin{figure}[htp]
+    \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison.pdf}
+    \caption{}
+\label{fig:comparison-titan-grid}
+\end{figure}
 

### 2017-03-17 Experiments.tex
Transition V to VI

--- 2017-02-14 Experiments.tex
@@
-In the batch queue,  queue time becomes dominant but, at the same time, we have more freedom to decide the parameters of the slot.
+% In the batch queue,  queue time becomes dominant but, at the same time, we
+% have more freedom to decide the parameters of the slot.
 
@@
 
-In this section we describe experiments that have been performed by running the ATLAS workload on TITAN's batch queue by means of the next generation executor.   
-The importance of these experiments is twofold: on the one side, we illustrate the performances of the NGE; on the other side, they help to characterize TITAN's batch queue and we show that it can be used as alternative to backfilling.
-Since we aim to study the response time of batch queue as it is perceived by a common user,  special privileges have been asked to system administrator.
-%This is particularly important for the development of future execution strategies because queue time becomes dominant in
+In this section we illustrate the performances of the NGE by means of experiments.
+Experiments aim to determine the performances of the NGE and to show how it provides a negligible overhead while introducing new functionalities.
 
-Experiments aim to show the number of completed events as function of the time. 
-Each experiment corresponds to a number of different and independent runs that is meaningful from a statistical point of view.
-Each run consists in the sustained execution of the ATLAS workload for a pre-determined period of time \aanote{A week?
-10 days?}.
-Since execution of events is carried on by means NGE jobs, each run is based on a script that maintains constant pressure on TITAN's batch queue.
-The main tasks of the script are to:  i) guarantee that there are constantly four NGE jobs eligible for running on TITAN's queue (note that four jobs is the maximum allowed in TITAN's batch queue \ref{}); ii)  bind Athena-MP jobs to the NGE jobs; iii) perfom garbage collection.
-No limit in the number of NGE jobs is imposed.
+All the experiments have been performed on TITAN's batch queue.
+This allowed us to decide the pilot size in terms of number of nodes and wall-time.
+For the same reason, the completion time of each run of our experiments is dominated by the queue time.
+However, since we are interested only in the performances of the NGE, we present only data related to the execution time of the pilot and the tasks that are run within it.
 
-The simulation of events is due to the following execution model: the script binds a number of Athena-MP jobs that is equal to the number of nodes on which NGE job is running; then, NGE job spawns the Athena-MP jobs on the nodes; finally, the Athena-MP jobs will simulate a number of events that is proportional to the wall-time of the NGE job.
-As a consequence, the number of nodes determines the number of events that can be simulated concurrently whereas the wall-time determines the number of events that can be simulated one after the other.
-These two parameters together determine the size of the NGE job.
+We test the NGE pilot for weak scalability, weak scalability with resubmission, strong scalability and we provide a proof of concept about how the NGE is able to run heterogeous workloads.
 
-The size of the NGE job is constant within an experiment. Thus, each experiment is fully characterized by the size of the NGE jobs that it send on the nodes.
+\subsubsection{Weak scalability}
+\aanote{I am using the term simulation....I will change it once we decide how to call the CU}
+The experiment consists in running as many tasks as the number of nodes controlled by the pilot.
+Tasks consists in Athena-MP job that simulate 100 events taken from ATLAS workload.
+Each Athena-MP job uses 16 cores which coincide with the number of cores available on TITAN's nodes.
+In order to minimize measurement noises, we used the very same simulation for all the tasks. Therefore, ideally all the tasks should have the same execution time which corresponds to ~$80$ minutes.
 
-Each execution of Athena-MP is guaranteed to be successfull and, on average, comparable with those executed by PanDA in production.
+Tasks do not experience queue within the pilot since there is one node for each ATLAS simulation.
+Therefore, delays are of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all of simulations on the nodes.
 
-Table \ref{tab:experiments} provides the list of the experiments that have been performed.
+We tested different pilot sizes, i.e. : 250, 500, 1000 and 2000. For all of them the walltime was 2 hours.
 
-\begin{table*}
-\begin{center}
-\begin{tabular}{|c|l|l|l|l|}
-  \hline
-Experiment &\# nodes &wall-time (hours) & run duration(days) & \# runs   \\
-\hline
-1 & &  & &   \\
-2 & &  & &  \\
-3 & &  & &  \\
-4 & &  & &  \\
-\hline
-\end{tabular}
-\end{center}
-\caption{List of experiments}\label{tab:experiments}
-\end{table*}
-
-Each experiment has been designed by choosing the maximum number of nodes and the maximum wall-time of each TITAN bin \ref{} \aanote{Are we really doing this???}.
-The length of the experiment has been decided.....\aanote{It was 10 times the wall-time or something like that}.
-
-Figure \ref{fig:avtEvts} \aanote{(DRAFT)}depicts average number of completed events as function of the time.
-Figure \ref{fig:avtTimes} \aanote{(DRAFT)} depicts the average queue time and execution time per experiment.
+Figure \ref{fig:weakScal1} depicts the average pilot duration, the average execution time of each tasks and the average time required to send a tasks for execution.
 
 \begin{figure}[!htb]
-        \includegraphics[width=0.48\textwidth]{./figures/draft/mean.pdf}
-    \caption{Average number of completed as function of the time.}
-\label{fig:avtEvts}
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/Weak.pdf}
+    \caption{Average pilot execution time, average tasks execution time and average launching time for pilot sizes: 250, 500, 1000 and 2000.}
+\label{fig:weakScal1}
 \end{figure}
+\aanote{TO be commented as soon as the figure becomes definitive. Note that It would be better if we show the ratio between pilot duration and execution time}
+\subsubsection{Weak scalability with resubmission }
+This experiment is similar to the one presented above but in this case we want to test also the impact of submitting new simulations in place of those that end.
+Thus, we send to the pilot a number of tasks that is five times larger than the number of nodes.
+The reason behind this experiment is to stress the manager and the agent by pushing new tasks while others are ending their execution in such a way to increase the concurrency of these two components.
+
+In order to perform the experiments in a reasonable amount of time, we reduced  the number of events that are simulated by Athena-MP to sixteen.
+This allows us to complete a task in ~25 minutes only.
+Note that the number of events has been chosen in such a way that all the cores of a node run one event.
+
+We tested different pilot sizes, i.e. : 250, 500, 1000 and 2000. For all of them the walltime was 3 hours.
+
+Figure \ref{fig:weakScal2} \aanote{(DRAFT)}  depicts the average pilot duration, the average execution time of each tasks and the average time required to send a tasks for execution.
 
@@
         \includegraphics[width=0.48\textwidth]{./figures/draft/AvgT.pdf}
-    \caption{Average queue time and execution time of the experiments.}
-\label{fig:avtTimes}
+    \caption{Average pilot execution time, average tasks execution time and average launching time for pilot sizes: 250, 500, 1000 and 2000.}
+\label{fig:weakScal2}
 \end{figure}
 
-Because of the definition of the experiment the results shown in Figure \ref{} are alike (but not entirely comparable to) those provided by in Figure \ref{} where the performances of backfilling are shown.
+\subsubsection{Strong scability}
+
+\subsubsection{Heterogeneous execution}
 

### 2017-03-17 introduction.tex
Clean up

@@
-\section{Introduction}\label{sec:intro}
-
-The Large Hadron Collider (LHC) is created to explore the fundamental properties
-of matter for the next decades. Since LHC start-up in 2009, the experiments
-collected and distributed hundreds of petabytes of data worldwide to hundreds of
-computer centers. Thousands of physicists analyze petascale data volumes daily.
+The Large Hadron Collider (LHC) was created to explore the fundamental
+properties of matter for the next decades.  Since LHC start-up in 2009,
+multiple experiments  at LHC have collected and distributed hundreds of
+petabytes of data worldwide to hundreds of computer centers. 
+Thousands of physicists analyze petascale data volumes daily.
 One of the LHC experiments, the ATLAS~\cite{Aad:2008}, utilizes the Production
@@
 
-PanDA has a highly scalable architecture. Scalability has been demonstrated in
-ATLAS through the rapid increase in usage over the past several years of
-operations, and PanDA is expected to meet the continuously growing computing
-requirements of ATLAS over the next decade. PanDA was designed to have the
-flexibility to adapt to emerging computing technologies in processing, storage,
-networking as well as the underlying software stack. This flexibility has also
-been successfully demonstrated through the past six years: computing centers in
-ATLAS, spanning many continents, were seamlessly integrated into PanDA\@.
+% PanDA has a highly scalable architecture. Scalability has been demonstrated in
+% ATLAS through the rapid increase in usage over the past several years of
+% operations, and PanDA is expected to meet the continuously growing computing
+% requirements of ATLAS over the next decade. PanDA was designed to have the
+% flexibility to adapt to emerging computing technologies in processing, storage,
+% networking as well as the underlying software stack. This flexibility has also
+% been successfully demonstrated through the past six years: computing centers in
+% ATLAS, spanning many continents, were seamlessly integrated into PanDA\@.
 
-PanDA manages a wide spectrum of workloads, ranging from raw data processing to
-Monte Carlo simulation and user analysis, while constantly evolving to meet
-rapidly changing science needs. Today, PanDA serves several thousand users,
-managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU
-cores which process more than a million jobs per day (Figure~\ref{fig:daily}).
+% PanDA manages a wide spectrum of workloads, ranging from raw data processing to
+% Monte Carlo simulation and user analysis, while constantly evolving to meet
+% rapidly changing science needs.
+
+Today, PanDA serves several thousand users, managing job distribution to
+hundreds of ATLAS sites with more than 100,000 CPU cores which process more
+than a million jobs per day (Figure~\ref{fig:daily}).
 

### 2017-03-18 overall.tex
Section V ready for review. Iteration, figures, captions and addressed comments

@@
 
-Figure~\ref{fig:core-hours-utilization} shows the number of core-hours used by
-ATLAS on Titan during the experiment time window. ATLAS consumed a total of
-73.8M core-hours, for an average of $\approx$ 7M core-hours a month, with a
+Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of core-hours
+used by ATLAS on Titan during the experiment time window. ATLAS consumed a total
+of 73.8M core-hours, for an average of $\approx$7M core-hours a month, with a
 minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours in
@@
 
-\begin{figure}[htp]
-\includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
-\caption{Caption.}
-\label{fig:core-hours-utilization}
-\end{figure}
-
-Figure~\ref{fig:backfill-utilization} shows backfill utilization during the
-experiment time window.\jhanote{Neither figure show efficiency. Both show
-utilization.} Efficiency is defined as the fraction of Titan’s cores available
-via backfill that was utilized by the PanDA Brokers. \jhanote{I would plot the
-efficiency on the Y2 axis for both.} The brokers reached 18\% average
-efficiency, with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\%
-efficiency on February 2017. The number of total backfill cores available was
-38.1M in April 2016, and 33.1M in February 2017. This shows that the measured
-difference in efficiency did not depend on a comparable difference in total
-backfill availability.
+% \begin{figure}[htp]
+% \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
+% \caption{Caption.}
+% \label{fig:core-hours-utilization}
+% \end{figure}
 
 \begin{figure}[htp]
-\includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
-\caption{Caption.
-\jhanote{I would suggest similar style of X-axis as previous figure.
-Difficult to parse which is April if we are going to call out a specific month.
-Also consistency in style is generally good to keep cognitive burden low.}}
+    \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.pdf}
+    \caption{Gray bars: Number of unused cores of Titan's backfill availability;
+    green bars: Number of cores of backfill availability used by ATLAS via PanDA
+    Brokers on Titan; red labels: Efficiency of PanDA Brokers on Titan defined
+    as percentage of backfill availability used every month of the experiment
+    window.}
 \label{fig:backfill-utilization}
@@
 
-Figure~\ref{fig:hpc-workload-utilization} shows that during the experiment time
-window, about 2.25M detector simulation jobs were completed on Titan, for a
-total of 225M events processed. This is equivalent to 0.9\% of all the 250M
-detector simulations performed by ATLAS in the same period of time, and 3.5\% of
-the 6.6B events processed by those jobs. Comparatively, Titan contributed 3.9\%
-of the total of around 200K cores available to ATLAS on Grid, Cloud, and HPC
-infrastructures together. These figures confirms the relevance of
-supercomputers' resource contribution to the LHC Run 3, especially when
-accounting for the amount of unused backfill availability and the rate of
+Figure~\ref{fig:backfill-utilization} (green bars) shows backfill utilization
+during the experiment time window. Efficiency
+(Figure~\ref{fig:backfill-utilization}, red labels) is defined as the fraction
+of Titan’s cores available via backfill that was utilized by the PanDA Brokers.
+The brokers reached 18\% average efficiency, with a minimum 8.9\% efficiency on
+April 2016 and a maximum 33.5\% efficiency on February 2017. The number of total
+backfill cores available was 38.1M in April 2016, and 33.1M in February 2017.
+This shows that the measured difference in efficiency did not depend on a
+comparable difference in total backfill availability.
+
+\jhanote{I would plot the efficiency on the Y2 axis for both
+[diagrams].}\mtnote{Done but then the first diagram ended up plotting a subset
+of the data of the second diagram. I used only the second diagram.} \jhanote{I
+would suggest similar style of X-axis as previous figure. Difficult to parse
+which is April if we are going to call out a specific month. Also consistency in
+style is generally good to keep cognitive burden low.}\mtnote{Done.}
+
+% Figure~\ref{fig:hpc-workload-utilization} shows that d
+
+During the experiment time window, about 2.25M detector simulation jobs were
+completed on Titan, for a total of 225M events processed. This is equivalent to
+0.9\% of all the 250M detector simulations performed by ATLAS in the same period
+of time, and 3.5\% of the 6.6B events processed by those jobs. Comparatively,
+Titan contributed 3.9\% of the total of around 200K cores available to ATLAS on
+Grid, Cloud, and HPC infrastructures together. These figures confirms the
+relevance of supercomputers' resource contribution to the LHC Run 3, especially
+when accounting for the amount of unused backfill availability and the rate of
 improvement of PanDA efficiency.
 
-\begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
-\caption{\mtnote{Placeholder for the diagram/data I asked to Danila and
-Sergey. Consider a table instead of a figure.}}
-\label{fig:hpc-workload-utilization}
-\end{figure}
+% \begin{figure}[htp]
+%     \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
+% \caption{\mtnote{Placeholder for the diagram/data I asked to Danila and
+% Sergey. Consider a table instead of a figure.}}
+% \label{fig:hpc-workload-utilization}
+% \end{figure}
 
@@
 Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time of
-1 event from $\approx$ 2 to $\approx$ 40 minutes, with a mean of $\approx$ 14
+1 event from $\approx$2 to $\approx$40 minutes, with a mean of $\approx$14
 minutes. Considering that each worker node process up to 16 events concurrently,
@@
     \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
-\caption{Distribution of the time taken by a Geant4 detector simulation to simulate one event.
-3000 Events; 30 Titan worker nodes; 16 simulations per node; 100 events per node.
-Average time per event $\approx$ 14 min.
-Broad distribution from $\approx$ 2 to $\approx$ 40 minutes.}
+    \caption{Distribution of the time taken by a Geant4 detector simulation to
+    simulate one event. 3000 Events; 30 Titan worker nodes; 16 simulations per
+    node; 100 events per node. Average time per event $\approx$14 min. Broad
+    distribution from $\approx$2 to $\approx$40 minutes.}
 \label{fig:100event-distrib}
@@
 availability on Titan as a function of number of nodes and the time of their
-availability (i.e., walltime). The mean number of nodes over the experiment
-window time was $\approx$ 267, and their mean walltime was 1445 minutes.
-Detector simulations of 100 events, enable to use down to 1 node for 1/4 of the
-mean walltime of backfill availability. As such, it offers a good compromise for
-the PanDA Broker efficiency.
+availability (i.e., walltime). We recorded these data by polling Titan's Moab
+scheduler at regular intervals between Oct 28 and Nov 3 2015, while developing
+PanDA Broker. The mean number of nodes was 267.1, and their mean walltime was
+1445 minutes. Detector simulations of 100 events, enable to use down to 1 node
+for 1/4 of the mean walltime of backfill availability. As such, it offers a good
+compromise for PanDA Broker efficiency.
 
 \begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.png}
-\caption{\mtnote{Waiting to know whether we have data for the experiment time window.}}
+    \includegraphics[clip,width=\columnwidth]{figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
+    \caption{Backfill availability measures as walltime vs number of cores
+    between Oct 28 and Nov 2015. Availability was measured 5312 times: mean
+    number of work nodes available 267.1; mean walltime available 1445
+    minutes.\mtnote{Waiting to know whether we have data for the experiment
+    time window.}}
 \label{fig:backfill-distrib}
@@
 
-\begin{itemize}
-    \item problem with lustre addressed with ramdisk (done)
-    \item Competition for cache in AMD processors. (done)
-    \item Comparison hardware performance Titan/Grid (data acquired)
-    \item Summary overall figure of performance analogous to those produced with NGE. Diagram as discussed in F2F meeting at Rutgers. (waiting)
-\end{itemize}
+% \begin{itemize}
+%     \item problem with lustre addressed with ramdisk (done)
+%     \item Competition for cache in AMD processors. (done)
+%     \item Comparison hardware performance Titan/Grid (data acquired)
+%     \item Summary overall figure of performance analogous to those produced
+%     with NGE. Diagram as discussed in F2F meeting at Rutgers. (waiting)
+% \end{itemize}
 
@@
 problem of metadata contention, improving metadata read performance from
-$\approx$ 6,300 seconds on Lustre to $\approx$ 1,500 seconds on NFS. Further, at
+$\approx$6,300 seconds on Lustre to $\approx$1,500 seconds on NFS. Further, at
 configuration time, the input files of each detector simulation job were stored
@@
 
-The compute performance of Titan for detector simulation jobs is, on average,
-lower that that of LCWG. Figure~\ref{fig:comparison-titan-grid} shows the
-difference in execution time of the same type of simulation between Titan and
-\ldots\mtnote{Add the name of the grid sites here}. This is not an issue with
-the design of PanDA Brokers and, more in general, with the migration of part of
-the ATLAS workload to supercomputers: Titan has been active since 2012 and more
-modern CPUs are expected to offer better performance. A pilot project is already
-in place to test PanDA Broker on Summit, the next generation supercomputer that
-will become active in 2018 at OLCF.
+The performance analysis of Titan's AMD CPUs for detector simulations helps also
+to compare Titan and Grid site performances. Usually, Grid sites exposes
+resources with heterogeneous CPU architectures and a maximum of 8 (virtual)
+cores per worker node, while Titan's offer an homogeneous 16 cores architecture.
+We used the rate of events processes per minute as a measure of the efficiency
+of executing the same detector simulation on both Titan or Grid sites.
+Figure~\ref{fig:comparison-titan-grid} compares the efficiencies of Titan to the
+BNL and SIGNET Grid sites, normalized for 8 cores. Effective performance
+per-core at Titan is $\approx$0.57 event per minute, roughly 1/2 of BNL and 1/3
+of SIGNET performances.
 
 \begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison.pdf}
-    \caption{}
+    \includegraphics[clip,width=\columnwidth]{figures/titan_grid_event_rate.pdf}
+    \caption{Comparison of the event rate for finished jobs at Titan (h1), BNL
+    (h2) and GRIDNET (h3). Titan histogram shows 1/2 of the measured values to
+    account for the core difference between Titan (16) and the Grid sites (8).
+    Performance per core: BNL 1.126; SIGNET 0.885; Titan 0.577.}
 \label{fig:comparison-titan-grid}
 \end{figure}
+
+The differences in performance between Titan and the BNL and SIGNET Grid sites
+are due to the FP scheduler competition and the availability of newer
+processors. The CPUs at the Grid sites have one FP scheduler per core and are on
+average newer than the CPU of Titan. The heterogeneity of the Grid sites' CPUs
+explain the higher performance variance compared to the performance consistency
+measured on Titan.
+
+\mtnote{Figure summarizing job performance on Titan analogous to those produced
+with NGE, as discussed in F2F meeting at Rutgers?}
 
@@
 initial preponderance of unknown failures was progressively addressed by a more
-accurate failure model, specifically tailored to the PanDA Broker and Titan. It
-also shows the progressive improvement of the broker's reliability: In Februrary
-2017, most of the failures were produced by the initialization and configuration
-phases of AthenaMP or by failed executions of the payload.\mtnote{If we want to
-keep this diagram, we will need to plot it from the CSV to make it readable.}
+accurate failure model, specifically tailored to the PanDA Broker and Titan.
+Figure~\ref{fig:failures-titan} also shows the progressive improvement of the
+broker's reliability: In Februrary 2017, errors related to input files and
+initialization were almost eliminated and execution errors were comparable to
+those recorded in January 2017 even if almost twice as many simulations were
+performed (Figure~\ref{fig:backfill-utilization}). The increased amount of
+unknown failure in February indicates that furter work is needed to improve the
+PanDA Broker failure model.
 
 \begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/failures_titan.png}
-\caption{Caption.}
+    \includegraphics[clip,width=\columnwidth]{figures/failures_titan.pdf}
+    \caption{PanDA Broker failures during the experiment time window. Failures
+    were recorded with 52 exit codes, here aggregated into 6 categories. The
+    number of failures should be compared to the number of simulations executed
+    in each month as plotted in Figure~\ref{fig:backfill-utilization}.
+    Improvements in the failure model of PanDA Broker shows a decrease in
+    unknown failure and a the progressive addressing of other categories of
+    failure.}
 \label{fig:failures-titan}

### 2017-03-19 WMSLiterature.tex
Iteration 4 and 5 adding Sergey's data
First pass section II; moved definitions to end of introduction

@@
 
-WMS can implement the pilot abstraction: WMS control the acquisition of
+WMS like PanDA implement the pilot abstraction: WMS control the acquisition of
 resources by means of pilots and then manage the execution of workloads or
 workflows' tasks on those pilots. In tis way, pilots expose a uniform scheduling
-interface and execution environment, isolating the differences among the
+interface and execution environment, abstracting the differences among the
 heterogeneous resources of multiple sites; WMS centralize the management of the
@@
 design and architectural principles: separation of concern between management
-and execution, data and compute; centralization of the management capabilities;
-distributed execution of the execution capabilities across multiple sites;
-unification of the application interface; hiding of resource heterogeneity;
-centralized monitoring and accounting of both resources and execution; and
-collection of static and sometimes dynamic information about resources.
+and execution, and data and compute; centralization of the management
+capabilities; distributed execution of the execution capabilities across
+multiple sites; unification of the application interface; hiding of resource
+heterogeneity; centralized monitoring and accounting of both resources and
+execution; and collection of static and sometimes dynamic information about
+resources.
 
-For example, \mtnote{Summarize design and architecture descriptions of the four
-WMS stressing some of their main similarities. Use a subset of the `principles'
-listed above.}
+AliEn, DIRAC, GlideinWMS and PanDA all share a similar design with two types of
+components: the management ones facing the application layer and centralizing
+the capabilities required to acquire tasks' descriptions and matching them to
+resource capabilities; and resource components required to acquire information
+about resource capabilities and the resources themselves. Architecturally, the
+management components include one or more queue and a scheduler capable of
+coordinating via push/pull protocols with the resource modules. All resource
+components include middleware-specific APIs for submitting resource requests
+and, in the case of the pilot-enabled WMS of the LHC projects, a pilot capable
+of pulling tasks from the management modules and scheduling them on their
+resources (i.e., multi-level scheduling).
 
-The implementations of pilot-enabled WMS mainly differ in how specialized they
-are for a type of resource middleware and workflow. All the WMS developed for
-LHC assume to use sites exposing different flavors of Grid middleware. As such,
-all the WMS implementations support Grid-like authentication and authorization
-and a computational model based on distributing a large amount of single-core
-tasks across hundreds of sites\mtnote{Is this true?}.
+AliEn, DIRAC, GlideinWMS and PanDA present several similarities also in their
+implementations. All were initially implemented to use Grid resources, using one
+or more components to the Condor software ecosystem~\cite{thain2005distributed}
+and, as with GlideinWMS, contributing to its development. Accordingly, all LHC
+WMS implemented Grid-like authentication and authorization systems and adopted a
+computational model based on distributing a large amount of single/few-cores
+tasks across hundreds of sites\mtnote{Is this true?}. Tasks are assumed to fail
+or being held and all WMS implements subsystems dedicated to manage these
+failures or simply resubmitting failed tasks.
+
+All the experiments at LHC produces and process large amount of data both from
+actual collisions in the accelerator and from their simulations. Dedicated,
+multi-tiered data systems have been built to store, replicate, and distributed
+these data. All LHC WMS interface with these systems to move data to the sites
+where related compute tasks are executed or to schedule compute tasks where
+(large amount of) data are already stored.
+
+Currently, the implementation of PanDA WMS has at least two distinguishing
+features: the scale at which it manages computation, and network-aware
+scheduling of tasks to sites. PanDA concurrently supports distributed
+computations on up to 240,000 cores, submitting more than 55 million jobs in
+2016 across more than 100 sites, and processing up to XXXXPB of data
+distributed worldwide across XX Tier 1-3 facilities. As for 2017, PanDA serves
+XXXX users, including XX production groups. The scale of this operation is
+unprecedented, making PanDA the WMS managing the largest computational campaign
+in the world.
+
+PanDA was initially implemented to support computations across an unreliable and
+relatively expensive networking infrastructure. Accordingly, PanDA assumed a
+rigid model of data placement based on replication and long-term storage of data
+on the Grid sites. This model proven too wasteful, requiring storage space for
+unused and obsolete data. PanDA implementation was evolved to benefit from the
+development of the internet and of the networking infrastructure for scientific
+research. Today, PanDA constantly monitors network throughput and latency  among
+its management components, Grid sites and data facilities. PanDA uses this
+information to schedule tasks on sites, depending on the estimated time required
+to download and/or replicate input data and to stage out output data.
+
+The dynamic management of data replication and download is an example of a
+generalized paradigm shift that distinguishes the implementation of PanDA from
+other WMS. New components are being implemented to support the dynamic sizing of
+input dataset based on resource capabilities, the dynamic sizing of the
+resources held by pilots both in terms of number of cores and duration of the
+pilot, and further abstraction of different types of resources to reconcile
+High-Throughput and High-Performance Computing paradigms.
 
@@
 
-AliEn, DIRAC, and PanDA are implemented to support specific experimental
-workflows, while GlideinWMS was developed as component of the Condor software
-framework. ALICE workload is \ldots. LHCb workload is \ldots. CMS workload is
-\ldots.
-
-The ATLAS project requires data, compute, and memory-intensive workflows, both
-for single users and so called `production' research groups. Generally, PanDA
-was designed to support mostly single-core tasks, requiring a variable \ldots.
-
+% AliEn, DIRAC, and PanDA are implemented to support specific experimental
+% workflows, while GlideinWMS was developed as component of the Condor software
+% framework. ALICE workload is \ldots. LHCb workload is \ldots. CMS workload is
+% \ldots.
+%
+% The ATLAS project requires data, compute, and memory-intensive workflows, both
+% for single users and so called `production' research groups. Generally, PanDA
+% was designed to support mostly single-core tasks, requiring a variable \ldots.
+%
 % In this section, we present other workload manager systems that have been
@@
 
-The task execution in Alien is usually distributed over several sites.
-Each of these sites has at least one service called ClusterMonitor.
-On one side the Cluster Monitor is used to communicate with Central services (Task queue and Broker), on the other side it can manage Computing Elements (CE) by starting and stopping them whenever it receives the signal.
+The task execution in Alien is usually distributed over several sites. Each of
+these sites has at least one service called ClusterMonitor. On one side the
+Cluster Monitor is used to communicate with Central services (Task queue and
+Broker), on the other side it can manage Computing Elements (CE) by starting and
+stopping them whenever it receives the signal.
 

### 2017-03-19 bckground.tex
Notes
First pass section II; moved definitions to end of introduction

@@
-\subsection{Terminology}
+The Production and Distributed Analysis (PanDA) is a Workload Management System
+(WMS) developed to support the operations of the LHC ATLAS experiment. PanDA is
+specifically design to manage the execution of distributed workloads and
+workflows via pilots.
 
-\mtnote{To be moved to the end of the introduction.}
+WMS for distributed executions are a type of middlware, designed to coordinate
+specific activities about resources and workloads: discovering and selecting
+resources, submitting the tasks of a workload, and monitoring the execution of
+those tasks~\cite{marco2009glite}. Pilot is an abstraction that enables
+multi-level scheduling by decoupling resource acquisition from workload
+scheduling~\cite{turilli2015comprehensive}. Pilots are implemented by resource
+placeholders: a request for a certain amount of resources is scheduled on a site
+(e.g., by submitting a job or requesting a virtual machine) and once the
+resources are acquired, tasks are scheduled directly to the pilot, not to site's
+scheduler.
 
-The term ``workflow'' is used in many disciplines with different meaning. In the
-field of scientific computing, ``workflow'' assumes different meanings depending
-on the characteristics of the computation, of the software tools used to support
-this computation, and of the resources on which it is performed. Further, a
-workflow may indicate a whole application, a description of the computational
-process of that application or, more commonly, a series of tasks related by data
-dependences.
+WMS like PanDA implement the pilot abstraction: WMS control the acquisition of
+resources by means of pilots and then manage the execution of workloads or
+workflows' tasks on those pilots. In tis way, pilots expose a uniform scheduling
+interface and execution environment, abstracting the differences among the
+heterogeneous resources of multiple sites; WMS centralize the management of the
+execution process, offering a unified interface to the application layer.
+Pilot-enabled WMS are particularly relevant for LHC experiments, where large
+amount of heterogeneous resources, distributed across multiple sites have to be
+coordinated to execute tens of million of tasks per months while producing and
+storing petabytes of data.
 
-The lack of a consistent and shared definition of ``workflow'' hinders the
-understanding of its properties and its relations with related concepts. For
-example, we need to clarify the difference among ``workflow'', ``workload'',
-``task'', or ``job'' but also between workflow ``template'' and ``instance'', or
-``data-flow'' and ``control-flow''. This is precondition to specify properly the
-design of software systems that support the execution of scientific
-applications.
+Several pilot-enabled WMS were developed for the LHC experiments:
+AliEn~\cite{Bagnasco2010} for ALICE; DIRAC~\cite{Paterson2010} for LHCb;
+GlideinWMS~\cite{sfiligoi2008glideinwms} for CMS; and
+PanDA~\cite{maeno2014evolution} for ATLAS. These systems implement similar
+design and architectural principles: separation of concern between management
+and execution, and data and compute; centralization of the management
+capabilities; distributed execution of the execution capabilities across
+multiple sites; unification of the application interface; hiding of resource
+heterogeneity; centralized monitoring and accounting of both resources and
+execution; and collection of static and sometimes dynamic information about
+resources.
 
-In this paper, we use the following definitions:
+AliEn, DIRAC, GlideinWMS and PanDA all share a similar design with two types of
+components: the management ones facing the application layer and centralizing
+the capabilities required to acquire tasks' descriptions and matching them to
+resource capabilities; and resource components required to acquire information
+about resource capabilities and the resources themselves. Architecturally, the
+management components include one or more queue and a scheduler capable of
+coordinating via push/pull protocols with the resource modules. All resource
+components include middleware-specific APIs for submitting resource requests
+and, in the case of the pilot-enabled WMS of the LHC projects, a pilot capable
+of pulling tasks from the management modules and scheduling them on their
+resources (i.e., multi-level scheduling).
 
-\begin{description}
+AliEn, DIRAC, GlideinWMS and PanDA present several similarities also in their
+implementations. All were initially implemented to use Grid resources, using one
+or more components to the Condor software ecosystem~\cite{thain2005distributed}
+and, as with GlideinWMS, contributing to its development. Accordingly, all LHC
+WMS implemented Grid-like authentication and authorization systems and adopted a
+computational model based on distributing a large amount of single/few-cores
+tasks across hundreds of sites\mtnote{Is this true?}. Tasks are assumed to fail
+or being held and all WMS implements subsystems dedicated to manage these
+failures or simply resubmitting failed tasks.
 
-  \item[Task.] A set of operations to be performed on a computing platform,
-  alongside a description of the properties and dependences of those operations,
-  and indications on how the dependences should be satisfied and the operations
-  should be executed.
+All the experiments at LHC produces and process large amount of data both from
+actual collisions in the accelerator and from their simulations. Dedicated,
+multi-tiered data systems have been built to store, replicate, and distributed
+these data. All LHC WMS interface with these systems to move data to the sites
+where related compute tasks are executed or to schedule compute tasks where
+(large amount of) data are already stored.
 
-  \item[Job.] A unit of execution that performs one or more unit of work. Jobs
-  relates to the resource on which they are executed. One or more tasks can be
-  the units of work executed by a job.
+Currently, the implementation of PanDA WMS has at least two distinguishing
+features: the scale at which it manages computation, and network-aware
+scheduling of tasks to sites. PanDA concurrently supports distributed
+computations on up to 240,000 cores, submitting more than 365 million jobs in
+2016 across more than 100 sites, and processing up to XXXXPB of data distributed
+worldwide across XX Tier 1-3 facilities. As for 2017, PanDA serves several
+thousand users, including XX production groups. The scale of this operation is
+unprecedented, making PanDA the WMS managing the largest computational campaign
+in the world.
 
-  \item[Workload.] A set of tasks that can be executed concurrently, possibly
-  related by a set of relations. For example, tasks of a workload can share one
-  or more input files or communicate during execution.
+PanDA was initially implemented to support computations across an unreliable and
+relatively expensive networking infrastructure. Accordingly, PanDA assumed a
+rigid model of data placement based on replication and long-term storage of data
+on the Grid sites. This model proven too wasteful, requiring storage space for
+unused and obsolete data. PanDA implementation was evolved to benefit from the
+development of the internet and of the networking infrastructure for scientific
+research. Today, PanDA constantly monitors network throughput and latency  among
+its management components, Grid sites and data facilities. PanDA uses this
+information to schedule tasks on sites, depending on the estimated time required
+to download and/or replicate input data and to stage out output data.
 
-  \item[Workflow.] Set of workloads, related by a set of relations that define
-  the order in which each workload can be executed. Data dependences are the
-  most common relations among workloads, used to define the precedence among
-  their executions. Note that, formally, a workload can have a single task.
+The dynamic management of data replication and download is an example of a
+generalized paradigm shift that distinguishes the implementation of PanDA from
+other WMS. New components are being implemented to support the dynamic sizing of
+input dataset based on resource capabilities, the dynamic sizing of the
+resources held by pilots both in terms of number of cores and duration of the
+pilot, and further abstraction of different types of resources to reconcile
+High-Throughput and High-Performance Computing paradigms.
 
-\end{description}
+% \subsection{Terminology}
+%
+% \mtnote{To be moved to the end of the introduction.}
 
-Each task may have an arbitrary number of properties like number of cores,
-executables, or input/output files. Tasks may have precedence interrelations,
-depending on their data dependences or any other type of dependency mandated
-by the application algorithms. Tasks with precedence relations have to be
-executed serially; otherwise, tasks can be executed concurrently. A workload is
-defined as a set of tasks that can be executed concurrently. As such, a
-workflow can be composed of a set of workloads.
+% Each task may have an arbitrary number of properties like number of cores,
+% executables, or input/output files. Tasks may have precedence interrelations,
+% depending on their data dependences or any other type of dependency mandated
+% by the application algorithms. Tasks with precedence relations have to be
+% executed serially; otherwise, tasks can be executed concurrently. A workload
+% is defined as a set of tasks that can be executed concurrently. As such, a
+% workflow can be composed of a set of workloads.
 
-The terms ``task'' and ``job'' are also used inconsistently across communities
-that perform scientific computing. In this paper, ``job'' refers to the unit of
-work that is submitted to a local resource management system (LRMS), like the
-Slurm or PBS batch system of a cluster. As such, a task can become a job when
-is scheduled on a resource that exposes a LRMS but can also become a virtual
-machine or a container when bootstrapped on an infrastructure supporting
-virtualization. Tasks can be statically or dynamically grouped into jobs,
-depending on the resource capabilities and the task requirements.
-
-Usually, workflows are represented as graphs in which tasks are vertices and
-relations are edges~\cite{}. Often, graphs are supposed to be acyclic but graphs
-with cycles have been used to represent workflows of workflows~\cite{}. In this
-paper, a workflow template is a type of graph while a workflow instance is a
-workflow template with specific vertices and edges. Further, a workflow instance
-is ``abstract'' when no resource properties are available for all vertices,
-``concrete'' otherwise~\cite{}. When not qualified, the term ``workflow''
-indicates a abstract workflow instance.
+% The terms ``task'' and ``job'' are also used inconsistently across communities
+% that perform scientific computing. In this paper, ``job'' refers to the unit of
+% work that is submitted to a local resource management system (LRMS), like the
+% Slurm or PBS batch system of a cluster. As such, a task can become a job when
+% is scheduled on a resource that exposes a LRMS but can also become a virtual
+% machine or a container when bootstrapped on an infrastructure supporting
+% virtualization. Tasks can be statically or dynamically grouped into jobs,
+% depending on the resource capabilities and the task requirements.
+%
+% Usually, workflows are represented as graphs in which tasks are vertices and
+% relations are edges~\cite{}. Often, graphs are supposed to be acyclic but graphs
+% with cycles have been used to represent workflows of workflows~\cite{}. In this
+% paper, a workflow template is a type of graph while a workflow instance is a
+% workflow template with specific vertices and edges. Further, a workflow instance
+% is ``abstract'' when no resource properties are available for all vertices,
+% ``concrete'' otherwise~\cite{}. When not qualified, the term ``workflow''
+% indicates a abstract workflow instance.
+%
+% Notes for refactoring:
+% \begin{itemize}
+%     \item Naming and nomenclature is a contribution
+%     \item 1, 1.5 pages.
+% \end{itemize}

### 2017-03-19 introduction.tex
First pass section II; moved definitions to end of introduction

@@
 
-Today, PanDA serves several thousand users, managing job distribution to
-hundreds of ATLAS sites with more than 100,000 CPU cores which process more
-than a million jobs per day (Figure~\ref{fig:daily}).
+\mtnote{Used the following in related work. We may want to reduce/remove it from
+the introduction: ``Today, PanDA serves several thousand users, managing job
+distribution to hundreds of ATLAS sites with more than 100,000 CPU cores which
+process more than a million jobs per day.''} % (Figure~\ref{fig:daily}).
 
-\begin{figure}
-    \begin{center}
-        \includegraphics[width=\columnwidth]{figures/DailyJobs.png}
-        \caption{Daily completed jobs on ATLAS Grid for the past 12 month}
-    \end{center}
-\label{fig:daily}
-\end{figure}
+% \begin{figure}
+%     \begin{center}
+%         \includegraphics[width=\columnwidth]{figures/DailyJobs.png}
+%         \caption{Daily completed jobs on ATLAS Grid for the past 12 month}
+%     \end{center}
+% \label{fig:daily}
+% \end{figure}
 
@@
 \end{itemize}
+
+\mtnote{Moved from related work were we use already this terminology. To be
+iterated/adjusted for consistency once the first draft will be ready.}
+
+The term ``workflow'' is used in many disciplines with different meaning. In the
+field of scientific computing, ``workflow'' assumes different meanings depending
+on the characteristics of the computation, of the software tools used to support
+this computation, and of the resources on which it is performed. Further, a
+workflow may indicate a whole application, a description of the computational
+process of that application or, more commonly, a series of tasks related by data
+dependences.
+
+The lack of a consistent and shared definition of ``workflow'' hinders the
+understanding of its properties and its relations with related concepts. For
+example, we need to clarify the difference among ``workflow'', ``workload'',
+``task'', or ``job'' but also between workflow ``template'' and ``instance'', or
+``data-flow'' and ``control-flow''. This is precondition to specify properly the
+design of software systems that support the execution of scientific
+applications.
+
+In this paper, we use the following definitions:
+
+\begin{description}
+
+  \item[Task.] A set of operations to be performed on a computing platform,
+  alongside a description of the properties and dependences of those operations,
+  and indications on how the dependences should be satisfied and the operations
+  should be executed.
+
+  \item[Job.] A unit of  work performed by submitting a script to a resource
+  management system (LRMS), like  Slurm or PBS, or by requesting a virtual
+  machine or a container to a site supporting virtualization. One or more jobs
+  can perform the operations described with a task.
+
+  \item[Workload.] A set of jobs that can be executed concurrently, possibly
+  related by a set of relations. For example, jobs of a workload can share one
+  or more input files or communicate during execution.
+
+  \item[Workflow.] Set of jobs, related by a set of relations that define the
+  order in which each task can be executed. Data dependences are the most common
+  relations among workloads, used to define the precedence among their
+  executions.
+
+\end{description}

### 2017-03-19 PandaEvolution.tex
Ongoing references

@@
 but the size of Titan mitigates this limitation. Every year, about 10\% of
-Titan's capacity remains unused~\cite{titan_utilization}, corresponding to an
-average of 30,000 unused cores. This equals approximately 270M core hours per
-year, roughly 30\% of the overall capacity of WLCG.
+Titan's capacity remains unused, corresponding to an average of 30,000 unused
+cores. This equals approximately 270M core hours per year, roughly 30\% of the
+overall capacity of WLCG.
 
@@
 MPI scripts are submitted to Titan's PBS batch system via
-RADICAL-SAGA~\cite{radical-saga}, a Python module, compliant with the OGF GFD.90
-SAGA specification~\cite{saga-spec}. The Simple API for Grid Applications (SAGA)
-offers a unified interface to diverse job schedulers and file transferring
-services. In this way, SAGA provides an interoperability layer that lowers the
-complexity of using distributed infrastructures. Behind the API façade,
-RADICAL-SAGA implements a adaptor architecture: each adaptor interface the SAGA
-API with different middleware systems and services, including the PBS batch
-scheduler of Titan.
+RADICAL-SAGA~\cite{radical-saga_url}, a Python module, compliant with the OGF
+GFD.90 SAGA specification~\cite{goodale2008simple}. The Simple API for Grid
+Applications (SAGA) offers a unified interface to diverse job schedulers and
+file transferring services. In this way, SAGA provides an interoperability layer
+that lowers the complexity of using distributed infrastructures. Behind the API
+façade, RADICAL-SAGA implements a adaptor architecture: each adaptor interface
+the SAGA API with different middleware systems and services, including the PBS
+batch scheduler of Titan.
 

### 2017-03-19 PandaOverview.tex
Iteration 4 and 5 adding Sergey's data
Ongoing references

@@
 Since 2005, a certain amount of parallelism has been progressively introduced
-for jobs~\cite{multithreaded_jobs} but, so far, no MPI jobs have been considered
+for jobs~\cite{crooks2012multi} but, so far, no MPI jobs have been considered
 for production. Jobs are supposed to be relatively self-contained, capable of
@@
 details required to understand how HPC resources have been integrated within
-PanDA (\ref{sec:integration}). For a more detailed description of PanDA
-architecture see~\cite{wiki, papers}.
+PanDA (\ref{sec:integration}).
+
+% For a more detailed description of PanDA architecture see~\cite{wiki, papers}.
 
@@
 
-PanDA Pilot implements the pilot paradigm~\cite{pilot_review}, decoupling
-resource acquisition from job execution via multi-stage scheduling. This has
-several benefits, including increasing throughput by lowering the time taken to
-queue and schedule jobs; enabling both concurrent and sequential job execution
-depending on the number of cores available to the pilot and the time for which
-it remains available; and exposing a unified and consistent interface for the
-scheduling of jobs, independent from the middleware of the site on which the
-pilot has been instantiated.
+PanDA Pilot implements the pilot paradigm~\cite{turilli2015comprehensive},
+decoupling resource acquisition from job execution via multi-stage scheduling.
+This has several benefits, including increasing throughput by lowering the time
+taken to queue and schedule jobs; enabling both concurrent and sequential job
+execution depending on the number of cores available to the pilot and the time
+for which it remains available; and exposing a unified and consistent interface
+for the scheduling of jobs, independent from the middleware of the site on which
+the pilot has been instantiated.
 
@@
 execute \mtnote{Can a pilot request multiple jobs to execute
-concurrently?}\sergeynote{In principle it should be able to do that, but that's not
-how it is operated. One pilot - one job is a cleaner model in an environment
+concurrently?}\sergeynote{In principle it should be able to do that, but that's
+not how it is operated. One pilot - one job is a cleaner model in an environment
 where resources are "owned" by the organization that runs pilots. Simplifies

### 2017-03-19 Experiments.tex
A little bit less rough version

@@
 
-In this section we illustrate the performances of the NGE by means of experiments.
-Experiments aim to determine the performances of the NGE and to show how it provides a negligible overhead while introducing new functionalities.
+In this section we present experiments that aim to determine the performances of the NGE and to show how it provides a negligible overhead while introducing new functionalities.
 
-All the experiments have been performed on TITAN's batch queue.
-This allowed us to decide the pilot size in terms of number of nodes and wall-time.
-For the same reason, the completion time of each run of our experiments is dominated by the queue time.
-However, since we are interested only in the performances of the NGE, we present only data related to the execution time of the pilot and the tasks that are run within it.
+Experiments consist in running AthenaMP instances by using NGE pilots where each AthenaMP simulates a pre-determined number of events taken from ATLAS workload.	
+We present four groups of experiments in which we test the NGE for weak scalability, weak scalability with resubmission, strong scalability and heterogeous workloads.
+For each group we vary the size of the pilots and the number of AthenaMP that are sent for execution. 
+We collect data about the execution time of the pilots and the AthenaMP that are executed within them by considering all stages that these entities pass through during the execution.
+Experiments have been performed out of production by using NGE's pilots on TITAN's batch queue. 
+Because of TITAN's batch queue, the turnaround time of each run of our experiments is dominated by the time spent on TITAN's queue.
+Since we are interested only in the performances of the NGE, we do not provide data about queue time.
 
-We test the NGE pilot for weak scalability, weak scalability with resubmission, strong scalability and we provide a proof of concept about how the NGE is able to run heterogeous workloads.
+For the experiment, AthenaMP has been set to use all the 16 cores present on TITAN's nodes.
 
 \subsubsection{Weak scalability}
-\aanote{I am using the term simulation....I will change it once we decide how to call the CU}
-The experiment consists in running as many tasks as the number of nodes controlled by the pilot.
-Tasks consists in Athena-MP job that simulate 100 events taken from ATLAS workload.
-Each Athena-MP job uses 16 cores which coincide with the number of cores available on TITAN's nodes.
-In order to minimize measurement noises, we used the very same simulation for all the tasks. Therefore, ideally all the tasks should have the same execution time which corresponds to ~$80$ minutes.
+
+The experiment consists in running as many AthenaMP instances (also referred as tasks from now-on) as the number of nodes controlled by the pilot.
+Each AthenaMP have been assigned with 100 events whose simulation requires  ~$80$ minutes.
 
 Tasks do not experience queue within the pilot since there is one node for each ATLAS simulation.
-Therefore, delays are of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all of simulations on the nodes.
+Therefore, delays are consequence of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all of tasks on the nodes.
 
@@
 
-\subsubsection{Strong scability}
+\subsubsection{Strong scalability}
 

### 2017-03-19 bckground.tex

@@
-The Production and Distributed Analysis (PanDA) is a Workload Management System
-(WMS) developed to support the operations of the LHC ATLAS experiment. PanDA is
-specifically design to manage the execution of distributed workloads and
-workflows via pilots.
+The Production and Distributed Analysis (PanDA) software is a Workload
+Management System (WMS) developed to support the operations of the LHC ATLAS
+experiment. PanDA is specifically design to manage the execution of distributed
+workloads and workflows via pilots.
 
-WMS for distributed executions are a type of middlware, designed to coordinate
+WMS for distributed executions are a type of middleware, designed to coordinate
 specific activities about resources and workloads: discovering and selecting
-resources, submitting the tasks of a workload, and monitoring the execution of
-those tasks~\cite{marco2009glite}. Pilot is an abstraction that enables
-multi-level scheduling by decoupling resource acquisition from workload
+resources, submitting tasks of workloads and workflows, and monitoring the
+execution of those tasks~\cite{marco2009glite}. Pilot is an abstraction that
+enables multi-level scheduling by decoupling resource acquisition from workload
 scheduling~\cite{turilli2015comprehensive}. Pilots are implemented by resource
@@
 resources by means of pilots and then manage the execution of workloads or
-workflows' tasks on those pilots. In tis way, pilots expose a uniform scheduling
-interface and execution environment, abstracting the differences among the
-heterogeneous resources of multiple sites; WMS centralize the management of the
-execution process, offering a unified interface to the application layer.
-Pilot-enabled WMS are particularly relevant for LHC experiments, where large
-amount of heterogeneous resources, distributed across multiple sites have to be
-coordinated to execute tens of million of tasks per months while producing and
-storing petabytes of data.
+workflows' tasks on those pilots. WMS centralize the management of the execution
+process, offering a unified interface to the application layer. Pilots expose a
+uniform scheduling interface and execution environment, abstracting the
+differences among the heterogeneous resources of multiple sites. Pilot-enabled
+WMS are particularly relevant for LHC experiments, where millions of tasks are
+executed across multiple sites every month, analyzing and producing petabytes of
+data.
 
@@
 PanDA~\cite{maeno2014evolution} for ATLAS. These systems implement similar
-design and architectural principles: separation of concern between management
-and execution, and data and compute; centralization of the management
-capabilities; distributed execution of the execution capabilities across
-multiple sites; unification of the application interface; hiding of resource
-heterogeneity; centralized monitoring and accounting of both resources and
-execution; and collection of static and sometimes dynamic information about
-resources.
+design and architectural principles: centralization of task and resource
+management, and of monitoring and accounting; distribution of task execution
+across multiple sites; unification of the application interface; hiding of
+resource heterogeneity; and collection of static and sometimes dynamic
+information about resources.
 
@@
 the capabilities required to acquire tasks' descriptions and matching them to
-resource capabilities; and resource components required to acquire information
-about resource capabilities and the resources themselves. Architecturally, the
-management components include one or more queue and a scheduler capable of
-coordinating via push/pull protocols with the resource modules. All resource
-components include middleware-specific APIs for submitting resource requests
-and, in the case of the pilot-enabled WMS of the LHC projects, a pilot capable
-of pulling tasks from the management modules and scheduling them on their
-resources (i.e., multi-level scheduling).
+resource capabilities; and resource components used to acquire compute and data
+resources and information about their capabilities. Architecturally, the
+management components include one or more queue and a scheduler that coordinates
+with the resource modules via push/pull protocols. All resource components
+include middleware-specific APIs to request for resources, and a pilot capable
+of pulling tasks from the management modules and executing them on its
+resources.
 
-AliEn, DIRAC, GlideinWMS and PanDA present several similarities also in their
-implementations. All were initially implemented to use Grid resources, using one
-or more components to the Condor software ecosystem~\cite{thain2005distributed}
-and, as with GlideinWMS, contributing to its development. Accordingly, all LHC
-WMS implemented Grid-like authentication and authorization systems and adopted a
+AliEn, DIRAC, GlideinWMS and PanDA also have similar implementations. These WMS
+were initially implemented to use Grid resources, using one or more components
+to the Condor software ecosystem~\cite{thain2005distributed} and, as with
+GlideinWMS, contributing to its development. Accordingly, all LHC WMS
+implemented Grid-like authentication and authorization systems and adopted a
 computational model based on distributing a large amount of single/few-cores
-tasks across hundreds of sites\mtnote{Is this true?}. Tasks are assumed to fail
-or being held and all WMS implements subsystems dedicated to manage these
-failures or simply resubmitting failed tasks.
+tasks across hundreds of sites\mtnote{Is this true?}.
 
@@
 Currently, the implementation of PanDA WMS has at least two distinguishing
-features: the scale at which it manages computation, and network-aware
-scheduling of tasks to sites. PanDA concurrently supports distributed
-computations on up to 240,000 cores, submitting more than 365 million jobs in
-2016 across more than 100 sites, and processing up to XXXXPB of data distributed
-worldwide across XX Tier 1-3 facilities. As for 2017, PanDA serves several
-thousand users, including XX production groups. The scale of this operation is
-unprecedented, making PanDA the WMS managing the largest computational campaign
-in the world.
+features: the scale at which it operates distributed computing, and
+network-aware scheduling of tasks to sites\mtnote{PanDA team: please let me know
+whether more unique (i.e., not present in other LHC WMS) features of PanDA
+implementation should be indicated}. PanDA concurrently supports distributed
+computations on up to 240,000 cores, submitting more than 365 million jobs a
+year across 100 sites, and processing up to XXXXPB of data distributed worldwide
+on XX Tier 1-3 facilities. As for 2017, PanDA serves several thousand users,
+including XX production groups. The scale of this operation is unprecedented,
+making PanDA the WMS that manages the largest computational campaign in the
+world.
 
@@
 generalized paradigm shift that distinguishes the implementation of PanDA from
-other WMS. New components are being implemented to support the dynamic sizing of
-input dataset based on resource capabilities, the dynamic sizing of the
+other WMS. New components are being implemented to support: (i) dynamic sizing
+of input dataset based on resource capabilities; (ii) dynamic sizing of the
 resources held by pilots both in terms of number of cores and duration of the
-pilot, and further abstraction of different types of resources to reconcile
-High-Throughput and High-Performance Computing paradigms.
+pilot; and (iii) further abstraction of different types of resources to
+reconcile high-throughput and high-performance computing paradigms.
 
@@
 % The terms ``task'' and ``job'' are also used inconsistently across communities
-% that perform scientific computing. In this paper, ``job'' refers to the unit of
-% work that is submitted to a local resource management system (LRMS), like the
-% Slurm or PBS batch system of a cluster. As such, a task can become a job when
-% is scheduled on a resource that exposes a LRMS but can also become a virtual
-% machine or a container when bootstrapped on an infrastructure supporting
-% virtualization. Tasks can be statically or dynamically grouped into jobs,
-% depending on the resource capabilities and the task requirements.
-%
+% that perform scientific computing. In this paper, ``job'' refers to the unit
+% of work that is submitted to a local resource management system (LRMS), like
+% the Slurm or PBS batch system of a cluster. As such, a task can become a job
+% when is scheduled on a resource that exposes a LRMS but can also become a
+% virtual machine or a container when bootstrapped on an infrastructure
+% supporting virtualization. Tasks can be statically or dynamically grouped
+% into jobs, depending on the resource capabilities and the task requirements.
+
 % Usually, workflows are represented as graphs in which tasks are vertices and
-% relations are edges~\cite{}. Often, graphs are supposed to be acyclic but graphs
-% with cycles have been used to represent workflows of workflows~\cite{}. In this
-% paper, a workflow template is a type of graph while a workflow instance is a
-% workflow template with specific vertices and edges. Further, a workflow instance
-% is ``abstract'' when no resource properties are available for all vertices,
-% ``concrete'' otherwise~\cite{}. When not qualified, the term ``workflow''
-% indicates a abstract workflow instance.
-%
+% relations are edges~\cite{}. Often, graphs are supposed to be acyclic but
+% graphs with cycles have been used to represent workflows of
+% workflows~\cite{}. In this paper, a workflow template is a type of graph
+% while a workflow instance is a workflow template with specific vertices and
+% edges. Further, a workflow instance is ``abstract'' when no resource
+% properties are available for all vertices, ``concrete'' otherwise~\cite{}.
+% When not qualified, the term ``workflow'' indicates a abstract workflow
+% instance.
+
 % Notes for refactoring:

### 2017-03-19 nge.tex

@@
 As seen in \S\ref{sec:panda_deployment}, the absence of wide area network
-connectivity on Titan's work nodes was one of the main reasons for PanDA Broker
-not to implement pilots. As discussed in \S\ref{sec:panda_titan}, the absence of
-pilots imposes the static coupling between PBS jobs and detector simulations,
-and development tailored to each type of payload that needs to be executed.
+connectivity on Titan's worker nodes was one of the main reasons for PanDA
+Broker not to implement pilots. As discussed in \S\ref{sec:panda_titan}, the
+absence of pilots imposes the static coupling between PBS jobs and detector
+simulations, and development tailored to each type of payload that needs to be
+executed.
 
@@
 Pilots can offer a payload-independent scheduling interface while hiding the
-mechanics of coordination and communication among multiple work nodes. This
+mechanics of coordination and communication among multiple worker nodes. This
 could eliminate the need for packaging payload into PBS jobs within the broker,
@@
 specific PBS script for each payload. The submission process would also be
-MPI-independent, as MPI is used for coordination among multiple work nodes, not
-by the payload.
+MPI-independent, as MPI is used for coordination among multiple worker nodes,
+not by the payload.
 

### 2017-03-19 overall.tex

@@
 Figure~\ref{fig:backfill-distrib}. That specific number of event could then be
-pulled from the PanDA Event service~\cite{panda-event-service} and given as
-input to one or more simulations. Once packaged into the MPI script submitted to
+pulled from the PanDA Event service~\cite{calafiura2015atlas} and given as input
+to one or more simulations. Once packaged into the MPI script submitted to
 titan's PBS batch system, these simulations would better fit backfill

### 2017-03-19 paper.tex
Ongoing section II
proposed new title.
First pass section II; moved definitions to end of introduction

@@
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
+\documentclass[sigconf, screen]{acmart}
 
@@
 \input{include}
+
+% DOI
+% \acmDOI{10.475/123_4}
+
+% ISBN
+% \acmISBN{123-4567-24-567/08/06}
+
+%Conference
+\acmConference[SC2017]{The International Conference for High Performance Computing, Networking, Storage and Analysis}{November 2017}{Denver, Colorado USA}
+\acmYear{2017}
+\copyrightyear{2017}
+
+% \acmPrice{15.00}
 
@@
 
-\title{Everything you wanted to know about PanDA but were truly truly afraid to
-ask}
+\title{A Case Study on Abstractions to Converge High-Throughput and High-Performance Computing}
 
-% author names and affiliations
-% use a multiple column layout for up to two different
-% affiliations
-\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept.\ name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-\and
-\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
-\IEEEauthorblockA{line 1 (of Affiliation): dept.\ name of organization\\
-line 2: name of organization, acronyms acceptable\\
-line 3: City, Country\\
-line 4: Email: name@xyz.com}
-}
+\author{Author 1}
+\affiliation{\institution{Institution 1}}
+\email{email1}
 
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
+\author{Author 2}
+\affiliation{\institution{Institution 2}}
+\email{email2}
 
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3},
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
+\author{Author 3}
+\affiliation{\institution{Institution 3}}
+\email{email3}
 
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
+\author{Author 4}
+\affiliation{\institution{Institution 4}}
+\email{email4}
 
-% make the title area
-\maketitle
+\author{Author 5}
+\affiliation{\institution{Institution 5}}
+\email{email5}
+
+\author{Author 6}
+\affiliation{\institution{Institution 6}}
+\email{email6}
+
+% The default list of authors is too long for headers}
+\renewcommand{\shortauthors}{Author 1 et al.}
 
@@
 
-\begin{IEEEkeywords}
-TBD\@.
-\end{IEEEkeywords}
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below.
+%
+\begin{CCSXML}
+\end{CCSXML}
 
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
+% \ccsdesc[500]{Computer systems organization~Embedded systems}
+% \ccsdesc[300]{Computer systems organization~Redundancy}
+% \ccsdesc{Computer systems organization~Robotics}
+% \ccsdesc[100]{Networks~Network reliability}
+
+\keywords{ACM proceedings, \LaTeX, text tagging}
+
+\maketitle
 
@@
 % ----------------------------------------------------------------------------
-% II - BACKGROUND AND RELATED WORK
+% II - RELATED WORK
 % ----------------------------------------------------------------------------
-\section{Background and Related Work}
-\label{sec:bkgrd}
+\section{Related Work}
+\label{sec:related}
 
-\input{WMSLiterature}
+% \input{WMSLiterature}
 % Please do not correct bckground to background.
@@
 % ----------------------------------------------------------------------------
-\bibliographystyle{plain}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{bibliography}

### 2017-03-19 PandaOverview.tex
Ongoing iteration III, preparing for new diagram

@@
 
-The design of the Production and Distributed Analysis (PanDA) system started in
-2005 to develop a WMS for the ATLAS experiment. PanDA went into production for
-the LHC Run 1 on 2009 and was then partially redesigned to be deployed on Run 2,
-on 2015. In this section, we summarize the design, architecture, and execution
-process of the current state of the practice.
+The design of the PanDA WMS started in 2005 and its implementation went into
+production for the LHC Run 1 on 2009. PanDA was then partially redesigned to be
+deployed on Run 2, on 2015. In this section, we summarize the design,
+architecture, and execution process of the current state of the practice.
 
@@
 
-PanDA's application model primarily assumes single tasks, workloads and
-workflows. Tasks represent a set of operations performed on a set of events,
-stored across input files. Tasks can be decomposed into jobs, where each job
-represents the task's set of operations and a partition of the task's events.
-Since 2005, a certain amount of parallelism has been progressively introduced
-for jobs~\cite{crooks2012multi} but, so far, no MPI jobs have been considered
+PanDA's application model assumes single tasks, workloads and workflows. Tasks
+represent a set of operations performed on a set of events, stored in one or
+more input files. Tasks are decomposed into jobs, where each job represents the
+task's set of operations and a partition of the task's events. Since 2005, a
+certain amount of parallelism has been progressively introduced for job
+execution~\cite{crooks2012multi} but, so far, no MPI jobs have been considered
 for production. Jobs are supposed to be relatively self-contained, capable of
 setting up their own execution environment or having a minimal set of common
-dependences. Different types of job payload (i.e., executable) are assumed, both
-in terms of user-defined batch scripts or application frameworks for group of
-users.
+dependences.
+
+% Different types of job payload (i.e., executable) are assumed, both in terms
+% of user-defined batch scripts or application frameworks for group of users.
 
 PanDA's user model is based on multitenancy and the support of at least two
-types of HEP users: individual researchers and groups executing production-grade
-workflows. Users are considered free to submit tasks and workflows to the PanDA
-WMS at any point in time, directly or via dedicated application frameworks.
-Consistently, PanDA's security model is based on separation between
+types of HEP users: individual researchers and groups executing so called
+`production' workflows. Users are free to submit tasks and workflows to the
+PanDA WMS at any point in time, directly or via dedicated application
+frameworks. Consistently, PanDA's security model is based on separation between
 authentication, authorization and accounting for both single users and group of
 users. Both authentication and authorization are based on digital certificates
-and the X.509 standard and on the virtual organization abstraction.
+and the X.509 standard and on the virtual organization (VO) abstraction.
 
@@
 and binding of jobs are assumed to depend on the attributes of each job. Pilot
-is used to indicate the abstraction of resource capabilities. Jobs are thought
-to be bound to pilots and the execution of their payload is performed on the
-site where the pilots has been instantiated. Consistently, pilots are considered
-to be multitenant as jobs belonging to multiple workflows and/or users can be
-bound and executed on them.
+is used to indicate the abstraction of resource capabilities. Each job is
+thought to be bound to one pilot and executed on the site where the pilot has
+been instantiated.
 
-PanDA' data model uses events as the unit of data. Each event represents a
-collision captured in the detector and one or more events can be packaged into
-files or data source with other container abstractions. As with jobs, data have
-both attributes and states, and some of the attributes are shared between events
-and jobs so to create associations. Raw, reconstruction, and simulation data are
-assumed to be distributed across multiple and tiered sites and managed by the
-ATLAS Distributed Data Management (DDM)~\cite{garonne2012atlas}. Datasets
-required by each job are assumed to be replicated when necessary over the
-network. This applies to both input and output data. \sergeynote{Raw data from
-the ATLAS detector is captured at CERN and goes through reconstruction step to
-produce data in secondary formats, smaller in volume and with more and more
-physics content instead of raw detector hits. Consequently a second, custodial,
-copy of raw data is being distributed in predetermined volumes to National T1
-centers around the world (like BNL  T1 center). This insures low raw data
-redundancy and safety. Multiple copies of chunks of secondary data sets (derived
-from real raw data) are also distributed to ~100 centers around the word for
-ease of access and improved availability for analysis. Then raw detector data is
-stored on Tapes at CERN, just  in case. Most users, who do physics analysis,
-rarely work with raw data, only a small group of experts needs to work with raw
-data for specialized tasks like detector calibrations and such. There may be a
-need to redo a reconstruction step, when errors in previous reconstruction
-campaign are uncovered or improved detector understanding will lead for a better
-quality physics. Then data stored on tape is recovered and re-processed again.
-Simulation data is produced in many places anyway and is not centralized to
-begin with. Here's the paper that describes current DDM state in ATLAS, data
-replication strategies and such.
-http://iopscience.iop.org/article/10.1088/1742-6596/396/3/032045/pdf DDM is a
-complex and evolving topic in ATLAS.}\mtnote{Thank you. As we are describing the
-design of PanDA I thought we should not enter into a too detailed description of
-the DDM design. I eliminated the mistake I had made about `centralized data' but
-left DDM just as a reference. Any better?}
+% Consistently, pilots are considered to be multitenant as jobs belonging to
+% multiple workflows and/or users can be bound and executed on them.
+
+PanDA' data model uses events as the unit of data\mtnote{Is event an abstraction
+specific to the MD workflow?}. Each event represents a collision captured in the
+detector and one or more events can be packaged into files or data source with
+other container abstractions. As with jobs, data have both attributes and
+states, and some of the attributes are shared between events and jobs so to
+create associations. Raw, reconstruction, and simulation data are assumed to be
+distributed across multiple storage facilities and managed by the ATLAS
+Distributed Data Management (DDM)~\cite{garonne2012atlas}. Datasets required by
+each job are assumed to be replicated when necessary over the network, both for
+input and output data.
+
+% \sergeynote{Raw data from the ATLAS detector is captured at CERN and goes
+% through reconstruction step to produce data in secondary formats, smaller in
+% volume and with more and more physics content instead of raw detector hits.
+% Consequently a second, custodial, copy of raw data is being distributed in
+% predetermined volumes to National T1 centers around the world (like BNL  T1
+% center). This insures low raw data redundancy and safety. Multiple copies of
+% chunks of secondary data sets (derived from real raw data) are also
+% distributed to ~100 centers around the word for ease of access and improved
+% availability for analysis. Then raw detector data is stored on Tapes at CERN,
+% just  in case. Most users, who do physics analysis, rarely work with raw
+% data, only a small group of experts needs to work with raw data for
+% specialized tasks like detector calibrations and such. There may be a need to
+% redo a reconstruction step, when errors in previous reconstruction campaign
+% are uncovered or improved detector understanding will lead for a better
+% quality physics. Then data stored on tape is recovered and re-processed
+% again. Simulation data is produced in many places anyway and is not
+% centralized to begin with. Here's the paper that describes current DDM state
+% in ATLAS, data replication strategies and such.
+% http://iopscience.iop.org/article/10.1088/1742-6596/396/3/032045/pdf DDM is a
+% complex and evolving topic in ATLAS.}\mtnote{Thank you. As we are describing
+% the design of PanDA I thought we should not enter into a too detailed
+% description of the DDM design. I eliminated the mistake I had made about
+% `centralized data' but left DDM just as a reference. Any better?}
 
@@
 providing information about the stage of the execution in which each job or data
-item is or have been. Some attributes are assumed to be immutable across
+item is or has been. Some attributes are assumed to be immutable across
 execution and jobs and data items are assumed to be always in one and only one
@@
 of them built from off-the-shelf and Open Source components. At a high level,
-PanDA's architecture has six main subsystems: PanDA Server, AutoPyFactory, PanDA
-Pilot, PanDA Monitoring, JEDI, and Event Service. Subsystems communicate via
-dedicated API or http messaging. Databases are used to store eventful entities
-like tasks, jobs and events, and to store information about resources, logs, and
-accounting. Each subsystems is implemented by one or more modules as shown in
-Figure~\ref{fig:architecture}. In the following, we briefly describe the overall
-architecture of PanDA and its components. The description provides all the
-details required to understand how HPC resources have been integrated within
-PanDA (\ref{sec:integration}).
+PanDA's architecture has five main subsystems: PanDA Server, AutoPyFactory,
+PanDA Pilot, PanDA Monitoring, and JEDI. Subsystems communicate via dedicated
+API or http messaging. Databases are used to store eventful entities like tasks,
+jobs and events, and to store information about resources, logs, and accounting.
+Each subsystems is implemented by one or more modules as shown in
+Figure~\ref{fig:architecture}.
+
+% In the following, we briefly describe the overall architecture of PanDA and
+% its components. The description provides all the details required to
+% understand how HPC resources have been integrated within PanDA
+% (\ref{sec:integration}).
 
@@
 \paragraph{\textbf{PanDA Server}} this subsystem is the central hub service of
-PanDA and it operates as a web service. It runs on Apache web server,
-interacting with the back-end database running on separate servers. It
-implements the Apache worker model, with many independent processes handling
-client requests in parallel. Since all state is maintained in the central
-database, the PanDA Server application instance itself is stateless. Currently
-production version of PanDA utilizes Oracle database backend, but PanDA can also
-work with MySQL family of databases.
+PanDA and it operates as a web service~\cite{maeno2011overview}. It runs on
+Apache web server, interacting with the back-end database running on separate
+servers. It implements the Apache worker model, with many independent processes
+handling client requests in parallel. Since all state is maintained in the
+central database, the PanDA Server application instance itself is stateless.
+Currently, the production version of PanDA utilizes an Oracle database as
+backend, but PanDA can also work with MySQL family of databases.
 
@@
 
-PanDA Server implementation consists of four components, all implemented in
-Python and communicating via Python APIs: Task Buffer, Brokerage, Job
-Dispatcher, and Data Service. Task Buffer is the management component of PanDA's
-global queue and offers functionalities to: (i) store jobs into the global
-queue; (ii) provide information about and update the state of each stored job.
-The Brokerage component enables (i) matching job attributes with site and pilot
-attributes; (ii) managing the dispatch of input data to processing sites; and
-(iii) deriving PanDA's data pre-placement requirement.
+PanDA Server consists of four components implemented in Python:
+% and communicating via Python APIs
+Task Buffer, Brokerage, Job Dispatcher, and Data Service. Task Buffer manages
+PanDA's global queue by: (i) storing jobs into the global queue; (ii) providing
+information about and updating the state of each stored job. The Brokerage
+component: (i) matches job attributes with site and pilot attributes; (ii)
+manages the dispatch of input data to processing sites; and (iii) derives
+PanDA's data pre-placement requirement. The Job Dispatcher: (i) receives
+requests from active pilots for new jobs to process; (ii) retrieves from Task
+Buffer the jobs with highest priority and that can be executed on the site of
+the requesting pilot; (iii) creates a wrapper script for the selected job,
+tailored to the resource of the requesting pilot; and (iv) monitors the state of
+the job execution on the pilot, updating the Task Buffer component. Finally, the
+Data Service component: (i) retrieves information from Task Buffer and Brokerage
+about the data requirements of each job and the site to which each job has been
+assigned; (ii) interacts with the ATLAS DDM to make input files available to the
+jobs and manage their output data.
+
 % The job/pilot matching algorithm assigns a weight to each candidate pilot and
 % after applying a set of policies, the best ten candidates are taken.
-The Job Dispatcher implements functionalities to: (i) receive requests from
-active pilots for new jobs to process; (ii) retrieving from Task Buffer the jobs
-with highest priority that can be executed on the resource on which the
-requesting pilots have been instantiated; (iii) creating a wrapper for the
-selected job, tailored to the resource of the requesting pilot; and (iv)
-monitoring the state of the job execution on the pilot, updating the Task Buffer
-component. Finally, the Data Service component offers functionalities to: (i)
-retrieve information from Task Buffer and Brokerage about the data requirements
-of each job and the resource to which each job has been assigned; (ii) enable
-interaction with the ATLAS DDM to make input files available to the jobs and
-manage their output data.
 
 \paragraph{\textbf{PanDA Pilot}} this subsystem implements the execution
-environment for PanDA's jobs. Implemented in Python, a PanDA Pilot is launched
-by a wrapper submitted to a site. Wrappers are tailored to the
-specific interfaces and middleware exposed by each site so that PanDA Pilot
-remains site-independent. PanDA Pilot implements four main functionalities: (i)
-preparing the execution environment for PanDA's jobs,
+environment for PanDA's jobs\cite{nilsson2011atlas}. Implemented in Python, a
+PanDA Pilot is launched by a wrapper script submitted to a site. Wrapper scripts
+are tailored to the middleware exposed by each site so to maintain PanDA Pilot
+site-independent. PanDA Pilot: (i) prepares the execution environment for
+PanDA's jobs,
 % including  recovering and cleaning up previously failed jobs on that site and
 collecting information about the compute, data, and network capabilities of the
-site; (ii) requesting a job for execution to the Job Dispatcher component of
-PanDA Server; (iii) initiating and monitoring the execution of pulled jobs,
-including setting up the job, transferring its input when required, executing
-the payload, transferring output, and checking log files and local space; and
-(iv) clean up when the payload has finished to execute.
+site; (ii) requests a job for execution to the Job Dispatcher component of PanDA
+Server; (iii) initiates and monitors the execution of pulled jobs, including
+setting up the job, transferring its input when required, executing the payload,
+transferring output, and checking log files and local space; and (iv) cleans up
+when the payload has finished to execute.
 
@@
 
-PanDA Pilot implements the pilot paradigm~\cite{turilli2015comprehensive},
-decoupling resource acquisition from job execution via multi-stage scheduling.
-This has several benefits, including increasing throughput by lowering the time
-taken to queue and schedule jobs; enabling both concurrent and sequential job
-execution depending on the number of cores available to the pilot and the time
-for which it remains available; and exposing a unified and consistent interface
-for the scheduling of jobs, independent from the middleware of the site on which
-the pilot has been instantiated.
+% -----------------------------------------------------------------------------
+% PILOT DESCRIPTION: COMMENTED OUT BECAUSE WE DESCRIBE IT LATER ON.
+% PanDA Pilot implements the pilot paradigm~\cite{turilli2015comprehensive},
+% decoupling resource acquisition from job execution via multi-stage scheduling.
+% This has several benefits, including increasing throughput by lowering the
+% time taken to queue and schedule jobs; enabling both concurrent and
+% sequential job execution depending on the number of cores available to the
+% pilot and the time for which it remains available; and exposing a unified and
+% consistent interface for the scheduling of jobs, independent from the
+% middleware of the site on which the pilot has been instantiated.
+% -----------------------------------------------------------------------------
 
@@
 
-It should be noted that upon bootstrapping, a PanDA Pilot download major parts
-of its runtime code via HTTP from a central Subversion repository. The
-repository works with an Apache web server, configured with a memory-based web
-proxy (Squid). The purpose of the cache is to reduce the request load on the
-back-end Subversion server. That improves performance \mtnote{Do we have
-data/numbers?} since only occasional queries trigger a full lookup on the
-back-end Subversion system, and most external queries are pulled from memory on
-the front-end web server. The benefits of this system include a high-performance
-code download service, combined with code updates still being immediately
-available as soon as they are committed to source code control.
+Upon bootstrapping, PanDA Pilot downloads major parts of its runtime code via
+HTTP from a central Subversion repository. The repository works with an Apache
+web server, configured with a memory-based web proxy (Squid). The purpose of the
+cache is to reduce the request load on the back-end Subversion server. That
+improves performance \mtnote{Do we have data/numbers?} since only occasional
+queries trigger a full lookup on the back-end Subversion system, and most
+external queries are pulled from memory on the front-end web server. The
+benefits of this system include a high-performance code download service,
+combined with code updates still being immediately available as soon as they are
+committed to source code control.
 
@@
 sites~\cite{caballero2012autopyfactory}. Implemented to be executed as a single
-demonized process, AutoPyFactory enables the concurrent spawning of multiple job
-submission processes called APFQueue. Each APFQueue can communicate with PanDA
-Server and the Brokerage module to collect information about the state of the
-jobs it manages and, on the basis of this information, submit new pilots to one
-site per APFQueue. AutoPyFactory implements functionalities to evaluate the
-number of pilots already instantiated on each site, and to apply a consistent
-and adequate pressure on their batch queues. Submission is performed via the
-Condor-G batch system~\cite{frey2002condor}, enabling submission to multiple
-type of sites. AutoPyFactory exposes a dedicated monitoring service, feeding
-information to the PanDA Monitoring subsystem. A proxy manager is also run,
-enabling certificate-based authentication and authorization.
+demonized process, AutoPyFactory concurrently spawns multiple job submission
+processes called APFQueue. Each APFQueue communicates with PanDA Server and the
+Brokerage module to collect information about the state of the jobs it manages.
+On the basis of this information, an APFQueue submits new pilots to one site.
+AutoPyFactory implements functionalities to evaluate the number of pilots
+already instantiated on each site, and to apply a consistent and adequate
+pressure on their batch queues. Submission is performed via the Condor-G batch
+system~\cite{frey2002condor}, enabling submission to multiple type of sites.
+AutoPyFactory exposes a dedicated monitoring service, feeding information to the
+PanDA Monitoring subsystem. A proxy manager enables certificate-based
+authentication and authorization for each site.
 
@@
 
-\paragraph{\textbf{JEDI}} this subsystem implements a workload and workflow
-management for tasks instead of jobs, as defined by HEP
-users~\cite{borodin2015scaling}. Implemented as a standalone service that
-interacts with Panda Server via its REST interface, JEDI offers two main
-functionalities: partitioning tasks to jobs, and binding jobs to resources via
-dedicated queues. JEDI partition tasks to jobs by matching tasks requirements to
-resource capabilities. This matching is based on tasks attributes, and static
-and dynamic information about the resources of each available site. Dynamic
-information is acquired via a small number of `scout' jobs that collect
-information about current resources and their performance. Functionally, JEDI
-defines a set of queues associated to groups of resources with specific
-capabilities. Tasks are assigned to groups of these queues, and the jobs in
-which each task is partitioned are assigned to each queue, depending on their
-requirements. JEDI replaces the job brokerage functionalities of the Brokerage
-module of the PanDA Server but this module still handles communication between
-the Task Buffer and the Data Service modules.
-%, using it to enact the execution of the jobs defined by JEDI
-% on the resources and sites decided by JEDI.
+\paragraph{\textbf{JEDI}} this subsystem exposes a task abstraction for workload
+and workflows specification and interfaces with PanDA Server to manage their
+execution on remote sites~\cite{borodin2015scaling}. Implemented as a standalone
+service, JEDI partitions tasks to jobs, and binds these jobs to sites via
+dedicated queues. JEDI collects information about sites' resource capability,
+availability, and performance. Based on this information, JEDI associates sites
+with similar capabilities to a queue. Each job is assigned to a queue by
+matching the job's requirements to the capability, availability and performance
+of the sites of that queue. In this way, JEDI replaces the Brokerage module of
+the PanDA Server.
+% This module still handles communication between the Task
+% Buffer and the Data Service modules.
 

### 2017-03-20 abstract.tex
Reduce stub abstract to 150 words required by SC

@@
+% Experiments at the Large Hadron Collider (LHC) face unprecedented computing
+% challenges. Heterogeneous resources are distributed worldwide, thousands of
+% physicists analyzing the data need remote access to hundreds of computing
+% sites, the volume of processed data is beyond the exabyte scale, and data
+% processing requires more than billions of hours of computing usage per year. The
+% PanDA (Production and Distributed Analysis) system was developed to meet the
+% scale and complexity of LHC distributed computing for the ATLAS experiment. In
+% the process, the old batch job paradigm of computing in HEP was discarded in
+% favor of a far more flexible and scalable model. The success of PanDA at the LHC
+% is leading to widespread adoption and testing by other experiments. PanDA is
+% the first exascale workload management system in HEP, already operating at
+% a million computing jobs per day, and processing over an exabyte of data in
+% 2013. We will describe the design and implementation of PanDA, present data on
+% the performance of PanDA at the LHC, and discuss plans for future evolution
+% of the system to meet new challenges of scale, heterogeneity and increasing
+% user base.
+
 Experiments at the Large Hadron Collider (LHC) face unprecedented computing
-challenges. Heterogeneous resources are distributed worldwide, thousands of
-physicists analyzing the data need remote access to hundreds of computing
-sites, the volume of processed data is beyond the exabyte scale, and data
-processing requires more than billions of hours of computing usage per year. The
-PanDA (Production and Distributed Analysis) system was developed to meet the
-scale and complexity of LHC distributed computing for the ATLAS experiment. In
-the process, the old batch job paradigm of computing in HEP was discarded in
-favor of a far more flexible and scalable model. The success of PanDA at the LHC
-is leading to widespread adoption and testing by other experiments. PanDA is
-the first exascale workload management system in HEP, already operating at
-a million computing jobs per day, and processing over an exabyte of data in
-2013. We will describe the design and implementation of PanDA, present data on
-the performance of PanDA at the LHC, and discuss plans for future evolution
-of the system to meet new challenges of scale, heterogeneity and increasing
-user base.
+challenges. Thousands of physicists analyze exabytes of data every year, using
+billions of computing hours on hundreds of computing sites worldwide. PanDA
+(Production and Distributed Analysis) is a workload management system (WMS)
+developed to meet the scale and complexity of LHC distributed computing for the
+ATLAS experiment. PanDA is the first exascale workload management system in HEP,
+executing millions of computing jobs per day, and processing over an exabyte of
+data in 2016. In this paper, we introduce the design and implementation of
+PanDA, describing its deployment on Titan, the third biggest supercomputer in
+the world. We analyze scalability, reliability and performance of PanDA on
+Titan, highlighting the challenges addressed by its architecture and
+implementation. We present preliminary results of experiments performed with the
+Next Generation Executer, a prototype we developed to meet new challenges of
+scale and resource heterogeneity.

### 2017-03-21 overall.tex
Backfill availability 2016 plot and iteration in V

@@
 availability (i.e., walltime). We recorded these data by polling Titan's Moab
-scheduler at regular intervals between Oct 28 and Nov 3 2015, while developing
-PanDA Broker. The mean number of nodes was 267.1, and their mean walltime was
-1445 minutes. Detector simulations of 100 events, enable to use down to 1 node
-for 1/4 of the mean walltime of backfill availability. As such, it offers a good
-compromise for PanDA Broker efficiency.
+scheduler at regular intervals during the experiment window time, while
+developing and deploying PanDA Brokers. The mean number of nodes was 680, and
+their mean walltime was 680 minutes. Detector simulations of 100 events, enable
+to use down to 1 node for 1/2 of the mean walltime of backfill availability. As
+such, it offers a good compromise for PanDA Broker efficiency.
 
 \begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
-    \caption{Backfill availability measures as walltime vs number of cores
-    between Oct 28 and Nov 2015. Availability was measured 5312 times: mean
-    number of work nodes available 267.1; mean walltime available 1445
-    minutes.\mtnote{Waiting to know whether we have data for the experiment
-    time window.}}
+    \includegraphics[clip,width=\columnwidth]{figures/titan_backfill_avail.pdf}
+    %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
+    \caption{Backfill availability measures as walltime in minutes vs number of
+    work nodes during the experiment window time. Availability was measured
+    62175 times: mean number of work nodes available 680; mean walltime
+    available 680 minutes.}
 \label{fig:backfill-distrib}

### 2017-03-21 introduction.tex
Fix bibliography, the joy of latex

@@
 The Large Hadron Collider (LHC) was created to explore the fundamental
-properties of matter for the next decades.  Since LHC start-up in 2009,
-multiple experiments  at LHC have collected and distributed hundreds of
-petabytes of data worldwide to hundreds of computer centers.
-Thousands of physicists analyze petascale data volumes daily.
-One of the LHC experiments, the ATLAS~\cite{Aad:2008}, utilizes the Production
-and Distributed Analysis (PanDA) workload management system~\cite{Maeno2011}
-(WMS) for distributed data processing and analysis. The ATLAS Computing
-model~\cite{Atlas2005} is based on a Grid paradigm~\cite{Foster:1998}, with
-multilevel, hierarchically distributed computing and storage resources. PanDA
-has been developed to meet growing ATLAS production and analysis requirements
-for a data-driven workload management system capable of operating at LHC data
-processing scale.
+properties of matter for the next decades.  Since LHC start-up in 2009, multiple
+experiments  at LHC have collected and distributed hundreds of petabytes of data
+worldwide to hundreds of computer centers. Thousands of physicists analyze
+petascale data volumes daily. One of the LHC experiments, the
+ATLAS~\cite{Aad:2008}, utilizes the Production and Distributed Analysis (PanDA)
+workload management system~\cite{Maeno2011} (WMS) for distributed data
+processing and analysis. The ATLAS Computing model~\cite{jones2008atlas} is
+based on a Grid paradigm~\cite{foster2003grid}, with multilevel, hierarchically
+distributed computing and storage resources. PanDA has been developed to meet
+growing ATLAS production and analysis requirements for a data-driven workload
+management system capable of operating at LHC data processing scale.
 

### 2017-03-21 Experiments.tex
New iteration experiments

@@
 
-In this section we present experiments that aim to determine the performances of the NGE and to show how it provides a negligible overhead while introducing new functionalities.
+In this section we present experiments that aim to determine the performances of the NGE and to show how it provides a minimal overhead while introducing new functionalities.
 
@@
 We present four groups of experiments in which we test the NGE for weak scalability, weak scalability with resubmission, strong scalability and heterogeous workloads.
-For each group we vary the size of the pilots and the number of AthenaMP that are sent for execution. 
+
 We collect data about the execution time of the pilots and the AthenaMP that are executed within them by considering all stages that these entities pass through during the execution.
@@
 Because of TITAN's batch queue, the turnaround time of each run of our experiments is dominated by the time spent on TITAN's queue.
-Since we are interested only in the performances of the NGE, we do not provide data about queue time.
+Since we are interested only in the performances of the NGE, we removed queue time from our statistics.
 
-For the experiment, AthenaMP has been set to use all the 16 cores present on TITAN's nodes.
+All the experiments have been performed by setting AthenaMP in such a way that it uses all the 16 cores present on TITAN's nodes.
 
@@
 
-Tasks do not experience queue within the pilot since there is one node for each ATLAS simulation.
-Therefore, delays are consequence of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all of tasks on the nodes.
+Tasks do not experience queue within the pilot since there is one node for each AthenaMP instance.
+Therefore, delays are consequence of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager, as defined in Section \ref{}, that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all the tasks on the nodes.
 
@@
 
-Figure \ref{fig:weakScal1} depicts the average pilot duration, the average execution time of each tasks and the average time required to send a tasks for execution.
+Figure \ref{fig:weakScal1a} depicts the average pilot duration and the average execution time of AthenaMP as function of the pilot size.
 
 \begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/Weak.pdf}
-    \caption{Average pilot execution time, average tasks execution time and average launching time for pilot sizes: 250, 500, 1000 and 2000.}
-\label{fig:weakScal1}
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakET1.pdf}
+    \caption{Average pilot execution time against average AthenaMP execution times for pilot sizes: 250, 500, 1000 and 2000.}
+\label{fig:weakScal1a}
 \end{figure}
-\aanote{TO be commented as soon as the figure becomes definitive. Note that It would be better if we show the ratio between pilot duration and execution time}
-\subsubsection{Weak scalability with resubmission }
-This experiment is similar to the one presented above but in this case we want to test also the impact of submitting new simulations in place of those that end.
-Thus, we send to the pilot a number of tasks that is five times larger than the number of nodes.
-The reason behind this experiment is to stress the manager and the agent by pushing new tasks while others are ending their execution in such a way to increase the concurrency of these two components.
 
-In order to perform the experiments in a reasonable amount of time, we reduced  the number of events that are simulated by Athena-MP to sixteen.
-This allows us to complete a task in ~25 minutes only.
-Note that the number of events has been chosen in such a way that all the cores of a node run one event.
+On the one side, we can observe that the average pilot duration is insensitive to the pilot size.
+This indicates that, on average, the overhead introduced by the pilot does not grow significatively with the size of the pilot.
+On the other side, we can notice that the average execution time of AthenaMP is smaller than the average pilot duration.
+This indicates that the overhead of the pilot is not null and can be quantified in ~5 minutes.
 
-We tested different pilot sizes, i.e. : 250, 500, 1000 and 2000. For all of them the walltime was 3 hours.
+\aanote{I think we should say something about the execution time of AthenaMP that is smaller than the others.}
 
-Figure \ref{fig:weakScal2} \aanote{(DRAFT)}  depicts the average pilot duration, the average execution time of each tasks and the average time required to send a tasks for execution.
+In Figure \ref{fig:weakScal1b}, we provide the time required by the pilot to start running AthenaMP instances.  
+We can notice that the larger is the number of units and the larger is the time required by the pilot to start running them.
+However, we can observe that the time does not grow linearly with the pilot size.
 
 \begin{figure}[!htb]
-        \includegraphics[width=0.48\textwidth]{./figures/draft/AvgT.pdf}
-    \caption{Average pilot execution time, average tasks execution time and average launching time for pilot sizes: 250, 500, 1000 and 2000.}
-\label{fig:weakScal2}
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakStart1.pdf}
+    \caption{Average time to start running AthenaMP for pilot sizes: 250, 500, 1000 and 2000.}
+\label{fig:weakScal1b}
 \end{figure}
 
+\subsubsection{Weak scalability with resubmission }
+This experiment is similar to the one presented above but in this case we want to test also the impact of submitting new AthenaMP instances in place of those that end.
+The reason behind this experiment is to stress pilot's components by pushing new tasks while others are ending their execution.
+
+In order to perform the experiments in less than two hours per run, we decided to send for execution a number of AthenaMP instances equal to five times the number of nodes and we reduced  the number of events that are simulated by each Athena-MP to sixteen\footnote{Note that the number of events has been chosen in such a way that all the cores of a node run one event.}.
+This allows us to complete a single AthenaMP in ~20 minutes only.
+
+We tested different pilot sizes, i.e. : 256, 512, 1024 and 2048. For all of them the walltime was 3 hours.
+
+Figure \ref{fig:weakScal2a} depicts the average pilot duration and  the average execution time of AthenaMP.
+
+\begin{figure}[!htb]
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakET2.pdf}
+    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
+\label{fig:weakScal2a}
+\end{figure}
+\begin{figure}[!htb]
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakStart2.pdf}
+    \caption{Average time to start running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
+\label{fig:weakScal2b}
+\end{figure}
 \subsubsection{Strong scalability}
+This experiments aim to test strong scalability. Therefore, we run the same amount of tasks for different pilot sizes. 
+We used a number of AthenaMP instances equal to 2048 and tested pilots with 256, 512, 1024 and 2048 nodes. 
+Each AthenaMP instance simulates sixteen events as the previous experiment.
+
+Figure \ref{fig:strongScala} depicts the average pilot duration and  the average execution time of AthenaMP.
+
+%\begin{figure}[!htb]
+%        \includegraphics[width=0.5\textwidth]{./figures/NGE/strongET.pdf}
+%    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
+%\label{fig:strongScala}
+%\end{figure}
+%\begin{figure}[!htb]
+%        \includegraphics[width=0.5\textwidth]{./figures/NGE/strongStart.pdf}
+%    \caption{Average time to start running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
+%\label{fig:strongScalb}
+%\end{figure}
 
 \subsubsection{Heterogeneous execution}
+
+The last experiment provides a proof of concept about the ability of the NGE to execute heterogeneous workload.
+In particular, we coupled the execution of AthenaMP with the execution of Gromacs to simulate molecular dynamics. 
+We performed the experiment by executing at first one AthenaMP per node and then, submitting a Gromacs simulation per core. Each Gromacs simulation requires ~20 minutes.
+We tested the following pilot sizes : 8, 32, 64.
 

### 2017-03-21 data/NGE/Analyzer.py

@@ -1,0 +1,141 @@
+import numpy
+import dataAnalysis as da
+#hashMap={'exp1':1000, 'exp2':2000,'exp3':250,'exp4':500}
+hashMap={'exp1':256, 'exp2':512,'exp3':1024,'exp4':2048}
+sessionFile ="sessions.csv"
+pilotFile= "pilots.csv"
+CUFile="units.csv"
+path ="../weak2/"
+
+QTFile = "PilotTimes2.csv"
+UnitFile ="UnitFiles2.csv"
+OverheadFile = "Overhead2.csv"
+
+PSIZE_PF_I=6
+PRUN_PF_I=2
+QT_PF_I=1
+
+EFFTIME_CU_I=1
+LASTTIME_I = 6
+
+SES_ST =8
+SES_SP =13
+SES_NCU =19
+
+PIL_ST =3
+PIL_SP =6
+PIL_NCU =9
+
+def ComputeOverhead2(inputPath,outputPath):
+
+	infile = open(inputPath)
+	infile.readline() # Consunme first line
+	outfile= open("./"+outputPath,"w+")
+	Overhead = {}
+	for key in hashMap.keys():
+		Overhead[key] = []
+	for line in infile:
+		tokens = line.split(',')
+		sum = 0
+		for i in range(PIL_ST,PIL_SP):
+			sum+=float(tokens[i])
+		print(str(sum) +  " "+ tokens[PIL_SP])		
+		Overhead[tokens[PIL_SP]] +=[sum]
+	for k in Overhead.keys():
+		(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(Overhead[k],1.0,perc=0.05)
+		outfile.write(str(hashMap[k])  + " " + str(avg) + " " + str(sd) + " " + str(hmin)  + " " + str(hmax) + " "+ k +" ")
+		(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(Overhead[k],float(tokens[PIL_NCU]),perc=0.05)
+		outfile.write(str(hashMap[k])  + " " + str(avg) + " " + str(sd) + " " + str(hmin)  + " " + str(hmax) + " "+ k +"\n")	
+	outfile.close()
+	infile.close()
+
+
+
+
+
+
+def ComputeOverhead(inputPath,outputPath):
+
+	infile = open(inputPath)
+	infile.readline() # Consunme first line
+	outfile= open("./"+outputPath,"w+")
+	Overhead = {}
+	for key in hashMap.keys():
[…]

### 2017-03-21 data/NGE/dataAnalysis.py

@@ -1,0 +1,207 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import math
+import scipy.stats as st
+M1 = 'm1'
+M2 = 'm2'
+COUNT = 'samplePoints'
+MIN = 'min'
+MAX = 'max' 
+
+def splitInIntervals(nodeStart,nodeEnd,nodeBinSize,walltimeStart,wallTimeEnd,walltimeBinSize,infilePath,outfilePath):
+	'''
+	Read a file and group data in bins according node size and walltime. <data>Start indicates the starting point, <data>End the last point, <data>BinSize the width of the bin. infilePath indicates the path of the input file. outfilePath indicates where the output will be written
+	Empty bin are not written in the file
+	The method returns a list of lists of dictionary
+'''
+	inputfile = open(infilePath,'r')
+	#Initialized data collection variable
+	data = list()
+	for i in range(nodeStart,nodeEnd+1,nodeBinSize):
+		item = []	
+		for j in range(walltimeStart,wallTimeEnd+1,walltimeBinSize):
+			item2 = []
+			item.append(item2)
+		data.append(item)
+	
+	for line in inputfile:
+		tokens = line.split()
+		nnodes = int(float(tokens[1]))
+		walltime = int(tokens[2])
+		row = (nnodes-nodeStart)/nodeBinSize
+		if row >= len(data):
+			row -=1
+	
+		column = (walltime-walltimeStart)/walltimeBinSize
+		if column >= len(data[row]):
+			column-=1
+		#print(str(row) +  " "+ str(column) + " "+ str(nnodes) +  " "+ str(walltime))
+		data[row][column].append(tokens)
+		
+	nnodeBin = nodeStart
+	walltimeBin = walltimeStart
+	for row in data:
+		walltimeBin = walltimeStart
+		for col in row:
+			# Write on the file only if the bin is not empty
+			
+			if len(col) !=0:
+				outfile = open(outfilePath+"-"+str(nnodeBin)+"-"+str(walltimeBin),'w+')
+				for record in col:
+					for item in record:
+						outfile.write(item + " ")
+					outfile.write("\n")	
+				outfile.close()
+			walltimeBin+=walltimeBinSize
+		nnodeBin+=nodeBinSize
+			
+
+	return data
+
+
+
+
[…]

### 2017-03-23 nge.tex

@@
+\subsection{Rationale}
+\aanote{I think that the names of the states should be removed. They are meaningless without the figures. I left them just in case we decide to reintroduce the figures.}
 As seen in \S\ref{sec:panda_deployment}, the absence of wide area network
@@
 
+\subsection{Architecture}
+\label{sec:arch}
+We developed and we are currently integrating a prototype of Next Generation Executor (NGE) to add pilot
+capabilities to the PanDA Broker. NGE is a runtime system designed to execute heteroge\-neous and dynamic
+workloads on diverse resources (Fig.~\ref{fig:arch-overview}). Workloads and
+pilots are described via the Pilot API and passed to the NGE runtime system. NGE
+launches the pilots and executes the tasks of the workload on them.
+
+\begin{figure}
+  \centering
+   \includegraphics[width=\columnwidth]{figures/architecture_compact_rp_paper.pdf}
+  % \caption{RP overview. An application uses the Pilot API to describe pilots
+  %   (green squares) and units (red circles). The PilotManager (purple box)
+  %   instantiates (dash arrow) pilots, the UnitManager (purple box)
+  %   instantiates (solid arrow) units. Both managers are executed on the user
+  %   workstation. Pilots are launched (dashed arrows) on resources A and B
+  %   via SAGA API, an Agent (purple box) is bootstrapped for each pilot
+  %   (green box), units (solid red circles) are scheduled to the Agents via
+  %   MongoDB and executed by the Agent's Executer (yellow box).
+  \caption{NGE overview. An application uses the Pilot API to describe pilots
+    (green squares) and units (red circles). The PilotManager instantiates
+    (dash arrow) pilots, the UnitManager instantiates (solid arrow) units.
+    Both managers are executed on the user workstation. Pilots are launched
+    (dashed arrows) on resources A and B via SAGA API, an Agent is
+    bootstrapped for each pilot, units are scheduled (solid arrow) to the
+    Agents via MongoDB and executed by the Agent's Executer. Boxes color
+    coding: gray for entities external to NGE, white for APIs, purple for NGE's
+    modules, green for pilots, yellow for module's components.
+\label{fig:arch-overview}
+  }
+\end{figure}
+
+Internally, NGE represents pilots as aggregates of resources independent from
+the architecture and topology of the target machines, and workloads as a set
+of units to be executed on the resources of the pilot. Both pilots and units
+are stateful entities, each with a well-defined state model and life cycle.
+Their states and state transitions are managed via the three modules of the NGE
+architecture: PilotManager, UnitManager, and Agent
+(Fig.~\ref{fig:arch-overview}, purple boxes).
+
+The PilotManager launches pilots on resources via the SAGA API\@\cite{radical-saga_url}.
+The UnitManager schedules units to pilots for execution. A MongoDB database is used to communicate the workload between UnitManager and Agents.  For this reason, the database instance needs to be
+accessible both from the user workstation and the target resources. The Agent
+bootstraps on a remote resource, pulls units from the MongoDB instance, and
+manages their execution on the cores held by the pilot.
+
+The lifespan of pilots has four states distributed among the PilotManager,
+resource, and pilot instance.% (Fig.~\ref{fig:pilot-state-model}). 
+Pilots are instantiated in the state \T{NEW} by the PilotManager, wait in a queue to be
+launched, and transition to \T{PM\_LAUNCH} when submitted to a Resource Manager
+(RM) via the SAGA API\@. Pilots wait in the queue of the RM and, once scheduled,
+become \T{P\_ACTIVE}. They remain in this state until the end of their
+lifetime, when they transition to \T{DONE}.
+
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=\columnwidth]{figures/pilot_state_model_compact_rp_paper.pdf}
+%  \caption{Pilot State Model. Instantiated in state \T{NEW}, each pilot is
+%    launched (\T{PM\_LAUNCH}) via SAGA API on a resource manager. After
+%    becoming \T{P\_ACTIVE}, it exhausts its duration ending in state
+%    \T{DONE}. The transition between states is sequential and each transition
+%    can be canceled or fail ending in the states \T{CANCELED} or \T{FAILED} (not
+%    depicted to improve the diagram clarity). Boxes color coding as per
+%    Fig.~\ref{fig:arch-overview}, with: orange for queues and blue for states.
+%\label{fig:pilot-state-model}}
+%\end{figure}
+%
+The unit state model has nine states distributed across the UnitManager,
+MongoDB instance, and Agent.% (Fig.~\ref{fig:unit-state-model}). 
+Instantiated in the state \T{NEW} by the UnitManager, every unit is scheduled on an Agent
+(\T{UM\_SCHEDULING}) via a queue on a MongoDB instance. The unit is then
+scheduled on the required number of cores held by the Agent's pilot
+(\T{A\_SCHEDULING}), and finally executed (\T{A\_EXECUTING}).
+
+The unit state model pertains also to the input and output data of the units.
+When required, the input data of a unit are either pushed to the Agent
+(\T{U\_STAGING\_IN}) or pulled from the Agent (\T{A\_STAGING\_IN}), depending
+on data locality and sharing requirements. Similarly, the output data of the
+unit are staged out by the Agent and UnitManager (\T{A\_STAGING\_OUT},
+\T{U\_STAGING\_OUT}) to a specified destination, e.g., the user workstation.
+Both input and output staging are optional, depending on the requirements of
+the units. The actual file transfers are enacted via SAGA, and support
+(gsi)-scp, (gsi)-sftp, and Globus Online.
+
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=\columnwidth]{figures/unit_state_model_compact_rp_paper.pdf}
+%  \caption{Unit State Model. Instantiated in state \T{NEW}, each unit is
+%    scheduled to an Agent (\T{UM\_SCHEDULING}) via MongoDB, and then scheduled
+%    to an Agent's Executer (\T{A\_SCHEDULING}). When required, unit input data
+%    are staged to the Agent's Pilot (\T{UM\_STAGING\_IN} or
+%    \T{A\_STAGING\_IN}). Each unit is executed (\T{A\_EXECUTING}), unit output
+%    data are staged out (\T{A\_STAGING\_OUT} and \T{UM\_STAGING\_OUT}), and
+%    each unit ends in state (\T{DONE}). Color coding and state omissions
+%    as per Fig.~\ref{fig:pilot-state-model}.
+%\label{fig:unit-state-model}}
+%\end{figure}
+
+%The state transitions represented in Figures~\ref{fig:pilot-state-model}
+%and~\ref{fig:unit-state-model} are sequential and every transition can fail or
+%be canceled by the PilotManager or UnitManager. 
+All state transitions are managed by the PilotManager, UnitManager, and Agent components. The only
+special case is the transition of the pilots to the state \T{P\_ACTIVE} which
+is dictated by the resource's RM, but managed by the PilotManager.
+
+Above all NGE's components, the Agent is the one that is worth to mention the most.
+It runs on MOM nodes and it is based on \emph{Open Run-Time Environment (ORTE)}.
+ORTE is a spin-off from the Open-MPI project and is a critical component of the OpenMPI implementation.
+It was developed to support distributed high-performance computing applications operating in a heterogeneous environment.
+The system transparently provides support for interprocess communication, resource discovery and allocation, and process launch across a variety of platforms.
+ORTE provides a mechanism similar to the Pilot concept - it allows the user to create a \emph{“dynamic virtual machine”} (DVM) that spans multiple nodes.
+ORTE provides libraries to allow the user to submit, monitor and manage tasks by avoiding filesystem bottlenecks and race conditions with network sockets.
+As a consequence, it is able to minimize the system overhead while submitting tasks.
+
+%Each component of RP has a independent semantic scope. This enables modularity
+%isolating implementation complexity and supporting diverse use cases and
+%environments.  For example, unit scheduling can be implemented by exchangeable
+%Scheduler components, suitable for applications of diverse scales, with
+%different coordination patterns, and executed on Beowulf clusters or Cray
+%machines.
+%
+%Components are also designed to be stateless and instantiated concurrently. In
+%this way, RP can manage multiple pilots and units at the same time, resulting
+%in scalable throughput and tolerance to failing components. Concurrent
+%components are coordinated via a dedicated communication mesh which incurs
+%infrastructure and runtime overhead, offset by the lower component complexity
+%and improved overall scalability of the system.
+
 % -----------------------------------------------------------------------------
-\subsection{Next Generation Executer}
-\label{ssec:nge}
 
@@
     \item Architecture (how)
-    \item Integration
+    \item Integration (Not present)
     \item Characterization (experiments)
@@
 
-We are developing a prototype of Next Generation Executer (NGE) to add pilot
-capabilities to the PanDA Broker. NGE is \ldots
+ \ldots

### 2017-03-24 PandaEvolution.tex
Fix bibliography, the joy of latex
Address Sergey's comments and diagram PanDA architecture

@@
 nodes use Cray’s Gemini interconnect for inter-node MPI messaging. Titan is
-served by the Spider II~\cite{spider2}, a Lustre filesystem with 32 PB of disk
-storage, and by a 29 PB HPSS tape storage system. Titan’s worker nodes run
+served by the Spider II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of
+disk storage, and by a 29 PB HPSS tape storage system. Titan’s worker nodes run
 Compute Node Linux, a run time environment based on SUSE Linux Enterprise
@@
 
-Currently, no HEP application can benefit from Titan's GPUs but some
-computationally-intensive and non memory-intensive tasks can be offloaded from
-the Grid to Titan's large amount of cores. Further, when HEP tasks can be
-partitioned to independent jobs, Titan worker nodes can be used to execute up to
-16 concurrent jobs' payload, one for each available core. Given these
+Currently, very few HEP applications could benefit from Titan's GPUs but some
+computationally-intensive and non memory-intensive tasks of ATLAS workflows can
+be offloaded from the Grid to Titan's large amount of cores. Further, when HEP
+tasks can be partitioned to independent jobs, Titan worker nodes can be used to
+execute up to 16 concurrent payloads, one for each available core. Given these
 constraints and challenges, the type of task most suitable for execution at the
-moment on Titan is Monte Carlo
-% event generation and
-detector simulation.
-% Both types of task are
-This type of task is mostly computational-intensive, requiring less than 2GB of
-RAM at runtime and with small input data requirements. Detector simulation tasks
-in ATLAS account for XX--XX\% of all jobs on WLCG, corresponding to X/X of the
-current Grid capacity.\mtnote{Please provide numbers to replace `XX' and,
-possibly, a diagram illustrating detector simulation stats compared to the whole
-ATLAS computation activity.}
+moment on Titan is Monte Carlo detector simulation. This type of task is mostly
+computational-intensive, requiring less than 2GB of RAM at runtime and with
+small input data requirements. Detector simulation tasks in ATLAS account for
+$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
+offloading.
 
@@
 digitization, and reconstruction. Event generation creates sets of particle
-four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia}
-and HERWIG~\cite{corcella2001herwig}. The detector simulator is called
-Geant4~\cite{agostinelli2003geant4} and simulates the interaction of these
-particles with the sensitive material of the ATLAS detector. Each interaction
-creates a so-called hit and all hits are collected and passed on for
+four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
+HERWIG~\cite{corcella2001herwig} and many others. The detector simulator is
+called Geant4~\cite{agostinelli2003geant4} and simulates the interaction of
+these particles with the sensitive material of the ATLAS detector. Each
+interaction creates a so-called hit and all hits are collected and passed on for
 digitalization where hits are further process to mimic the readout of the
@@
 it a pilot.}\mtnote{Thank you. We would have a problem calling it `agent' as we
-use that term to name the pilot of NGE. Would PanDA Broker work?}\mtnote{Sergey
-agreed that we should not use `pilot'. He is quite convinced by `broker' but we
-are both open to better suggestions.}
+use that term to name the pilot of NGE. Would PanDA Broker work?}\sergeynote{OK
+broker it is for this paper.}
 
 Given the communication requirements of PanDA Pilots and the unused capacity of
-Titan, PanDA pilot was redesigned to serve as a job broker on Titan. Architected
-as a stand-alone service to be executed on a DTN, this prototype called `PanDA
-Broker' implements functionalities to: (i) interrogate Titan about backfill
-availability; (ii) pull MC jobs and events; (iii) wrap the payload of ATLAS jobs
-into MPI scripts; (iv) submitting MPI scripts to Titan's PBS batch system and
-monitor their execution; and (v) staging input/output files. Backfill querying,
-payload wrapping, and scripts submission required a new implementation while
-pulling ATLAS job and events, and file staging were mostly inherited from PanDA
-Pilot.
+Titan, PanDA pilot was repurposed to serve as a job broker on the DTN nodes of
+Titan. Maintaining the core modules of PanDA Pilot and its stand-alone
+architecture, this prototype called `PanDA Broker' implements functionalities
+to: (i) interrogate Titan about backfill availability; (ii) pull MC jobs and
+events; (iii) wrap the payload of ATLAS jobs into MPI scripts; (iv) submitting
+MPI scripts to Titan's PBS batch system and monitor their execution; and (v)
+staging input/output files. Backfill querying, payload wrapping, and scripts
+submission required a new implementation while pulling ATLAS job and events, and
+file staging were inherited from PanDA Pilot.
 
@@
 module of the PanDA server for ATLAS jobs that have been bound to Titan by JEDI;
-(iii) the PanDA Broker requests to the Data Service module of the PanDA Server
-what input data these ATLAS jobs require and transfer the files to the DTN via
-the ATLAS DDM service; (iv) the PanDA Broker creates an MPI script, wrapping
-enough ATLAS jobs' payload to fit backfill availability; (v) the PanDA Broker
-submits the MPI script to the Titan's PBS batch system via RADICAL-SAGA; (vi)
-upon execution on the worker node(s), the MPI script creates a ramdisk and
-configure AthenaMP to execute the ATLAS jobs' payload; (vii) the payload is
-executed into dedicated processes, 16 for each worker node, 1 for each available
-core; (viii) upon completion of the payload execution, the PanDA Broker locates
-the output on the shared filesystem and transfer it to designated computing
-centre.
+(iii) Upon receiving the descriptions of those jobs, PanDA Broker pulls their
+input files from the ATLAS DDM service to the DTN; (iv) the PanDA Broker creates
+an MPI script, wrapping enough ATLAS jobs' payload to fit backfill availability;
+(v) the PanDA Broker submits the MPI script to the Titan's PBS batch system via
+RADICAL-SAGA; (vi) upon execution on the worker node(s), the MPI script creates
+configures and executes one AthenaMP for each work node available; (vi) AthenaMP
+spawns 1 event simulation process on each available core (16); (vii) during
+execution, PanDA Broker monitors execution progress and sends PanDA Server
+``heart beats'' for each job.; and (viii) upon completion of each simulation,
+the PanDA Broker locates the output on the shared filesystem, transfer it to
+designated computing centre, and performs cleanup.
 

### 2017-03-24 PandaOverview.tex
Address Sergey's comments and diagram PanDA architecture

@@
   \begin{center}
-    \includegraphics[width=\columnwidth]{figures/PandaArch.jpg}
+    \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
   \end{center}
-\caption{\mtnote{This diagram is a placeholder. I am working on a substitute
-that will be referenced when describing PanDA's subsystems.}Schematic view of
-the PanDA system, as deployed for production in 2017. Black boxes indicate
-PanDA's subsystems; Orange boxes components of a subsystem; purple boxes
-databases; green arrows http messaging interfaces; red arrows API; purple arrows
-database communications.}
+\caption{PanDA architecture representing the subsystems and components relevant
+to the integration of PanDA with Titan. Numbers indicates the execution process
+based on JEDI, from submission of tasks (1) to retrieval of their output (20).
+The monitoring subsystem, the architectural details of PanDA Pilot and the
+communication among subsystem's components are abstracted to improve clarity.
+% \mtnote{This diagram is a placeholder. I am working on a substitute
+% that will be referenced when describing PanDA's subsystems.}Schematic view of
+% the PanDA system, as deployed for production in 2017. Black boxes indicate
+% PanDA's subsystems; Orange boxes components of a subsystem; purple boxes
+% databases; green arrows http messaging interfaces; red arrows API; purple
+% arrows database communications.
+}
 % Originally PanDA was designed for grid infrastructure... In this
@@
 cache is to reduce the request load on the back-end Subversion server. That
-improves performance \mtnote{Do we have data/numbers?} since only occasional
-queries trigger a full lookup on the back-end Subversion system, and most
-external queries are pulled from memory on the front-end web server. The
-benefits of this system include a high-performance code download service,
-combined with code updates still being immediately available as soon as they are
-committed to source code control.
+improves performance since only occasional queries trigger a full lookup on the
+back-end Subversion system, and most external queries are pulled from memory on
+the front-end web server. The benefits of this system include a high-performance
+code download service, combined with code updates still being immediately
+available as soon as they are committed to source code control.
 

### 2017-03-25 PandaOverview.tex
Iteration III

@@
-\mtnote{TODO: replace `resource' with `site'; `http' to `HTTP'?}
+The design of PanDA WMS started in 2005 and its implementation went into
+production for the LHC Run 1 on 2009. PanDA was then extended with new
+subsystems to be deployed on Run 2, on 2015. In this section, we summarize the
+design, architecture, and execution process of the current state of the
+practice. We focus on those subsystems and components that are most relevant to
+the the integration of PanDA with supercomputers, the main topic of this paper.
 
-The design of the PanDA WMS started in 2005 and its implementation went into
-production for the LHC Run 1 on 2009. PanDA was then partially redesigned to be
-deployed on Run 2, on 2015. In this section, we summarize the design,
-architecture, and execution process of the current state of the practice.
+% \mtnote{TODO: replace `resource' with `site'; `http' to `HTTP'?}
 
@@
 % multiple workflows and/or users can be bound and executed on them.
+% PanDA' data model mostly uses events as the unit of data
+% \mtnote{Is event an abstraction specific to the MD workflow?}.
 
-PanDA' data model uses events as the unit of data\mtnote{Is event an abstraction
-specific to the MD workflow?}. Each event represents a collision captured in the
+In PanDA's data model, each event represents a collision captured in the
 detector and one or more events can be packaged into files or data source with
 other container abstractions. As with jobs, data have both attributes and
-states, and some of the attributes are shared between events and jobs so to
-create associations. Raw, reconstruction, and simulation data are assumed to be
-distributed across multiple storage facilities and managed by the ATLAS
-Distributed Data Management (DDM)~\cite{garonne2012atlas}. Datasets required by
-each job are assumed to be replicated when necessary over the network, both for
-input and output data.
+states, and some of the attributes are shared between events and jobs.
+% so to create associations.
+Raw, reconstruction, and simulation data are assumed to be distributed across
+multiple storage facilities and managed by the ATLAS Distributed Data Management
+(DDM)~\cite{garonne2012atlas}. Datasets required by each job are assumed to be
+replicated when necessary over the network, both for input and output data.
 
@@
 The implementation of PanDA consists of several interconnected subsystems, most
-of them built from off-the-shelf and Open Source components. At a high level,
-PanDA's architecture has five main subsystems: PanDA Server, AutoPyFactory,
-PanDA Pilot, PanDA Monitoring, and JEDI. Subsystems communicate via dedicated
-API or http messaging. Databases are used to store eventful entities like tasks,
-jobs and events, and to store information about resources, logs, and accounting.
-Each subsystems is implemented by one or more modules as shown in
-Figure~\ref{fig:architecture}.
+of them built from off-the-shelf and Open Source components. Subsystems
+communicate via dedicated API or HTTP messaging and each subsystems is
+implemented by one or more modules. Databases are used to store eventful
+entities like tasks, jobs and events, and to store information about sites,
+resources, logs, and accounting.
+
+The execution of specific tasks of ATLAS workflows on supercomputers involves
+four main subsystems of the PanDA architecture: PanDA Server, AutoPyFactory,
+PanDA Pilot, and JEDI. Figure~\ref{fig:architecture} shows a diagrammatic
+representation of these subsystems' architecture. For a full list of subsystems
+see Ref.~\cite{panda-wiki_url}.
 
@@
   \end{center}
-\caption{PanDA architecture representing the subsystems and components relevant
-to the integration of PanDA with Titan. Numbers indicates the execution process
-based on JEDI, from submission of tasks (1) to retrieval of their output (20).
-The monitoring subsystem, the architectural details of PanDA Pilot and the
-communication among subsystem's components are abstracted to improve clarity.
-% \mtnote{This diagram is a placeholder. I am working on a substitute
-% that will be referenced when describing PanDA's subsystems.}Schematic view of
-% the PanDA system, as deployed for production in 2017. Black boxes indicate
-% PanDA's subsystems; Orange boxes components of a subsystem; purple boxes
-% databases; green arrows http messaging interfaces; red arrows API; purple
-% arrows database communications.
-}
+    \caption{PanDA architecture with the subsystems and components
+    relevant to the integration of PanDA with supercomputers. Numbers indicates
+    the execution process based on JEDI: From the submission of tasks (1) to the
+    retrieval of their output (20). The monitoring subsystem, the architectural
+    details of PanDA Pilot and the communication among subsystem's components
+    are abstracted to improve clarity.}
+\label{fig:architecture}
+\end{figure}
+
 % Originally PanDA was designed for grid infrastructure... In this
@@
 % presented for workload X on Titan......
-\label{fig:architecture}
-\end{figure}
 
@@
 Apache web server, interacting with the back-end database running on separate
-servers. It implements the Apache worker model, with many independent processes
-handling client requests in parallel. Since all state is maintained in the
-central database, the PanDA Server application instance itself is stateless.
-Currently, the production version of PanDA utilizes an Oracle database as
-backend, but PanDA can also work with MySQL family of databases.
+machines. PanDA Server implements the Apache worker model, with many independent
+processes handling client requests in parallel. Since all state is maintained in
+the central database, the PanDA Server instance itself is stateless. Currently,
+the production version of PanDA utilizes an Oracle database as backend, but
+PanDA Server can also work with the MySQL family of databases.
 
@@
 
-PanDA Server consists of four components implemented in Python:
 % and communicating via Python APIs
-Task Buffer, Brokerage, Job Dispatcher, and Data Service. Task Buffer manages
-PanDA's global queue by: (i) storing jobs into the global queue; (ii) providing
+
+PanDA Server consists of four main components implemented in Python: Task
+Buffer, Brokerage, Job Dispatcher, and Data Service. Task Buffer manages PanDA's
+global queue by: (i) storing jobs into the global queue; (ii) providing
 information about and updating the state of each stored job. The Brokerage
@@
 the requesting pilot; (iii) creates a wrapper script for the selected job,
-tailored to the resource of the requesting pilot; and (iv) monitors the state of
+tailored to the site of the requesting pilot; and (iv) monitors the state of
 the job execution on the pilot, updating the Task Buffer component. Finally, the
@@
 
+% including  recovering and cleaning up previously failed jobs on that site and
+
 \paragraph{\textbf{PanDA Pilot}} this subsystem implements the execution
@@
 site-independent. PanDA Pilot: (i) prepares the execution environment for
-PanDA's jobs,
-% including  recovering and cleaning up previously failed jobs on that site and
-collecting information about the compute, data, and network capabilities of the
-site; (ii) requests a job for execution to the Job Dispatcher component of PanDA
-Server; (iii) initiates and monitors the execution of pulled jobs, including
-setting up the job, transferring its input when required, executing the payload,
-transferring output, and checking log files and local space; and (iv) cleans up
-when the payload has finished to execute.
+PanDA's jobs, collecting information about the compute, data, and network
+capabilities of the site; (ii) requests a job for execution to the Job
+Dispatcher component of PanDA Server; (iii) initiates and monitors the execution
+of pulled jobs, including setting up the job, transferring its input when
+required, executing the payload, transferring output, and checking log files and
+local space; and (iv) cleans up when the payload has finished to execute.
 
@@
 % PILOT DESCRIPTION: COMMENTED OUT BECAUSE WE DESCRIBE IT LATER ON.
+% -----------------------------------------------------------------------------
 % PanDA Pilot implements the pilot paradigm~\cite{turilli2015comprehensive},
@@
 
-\paragraph{\textbf{PanDA Monitoring}} this subsystem is implemented as a
-web-based, dashboard-style graphical application~\cite{klimentov2011atlas}. It
-runs on Apache and interacts with the back-end database of PanDA Server to
-enable persistence. PanDA Monitoring enables users and site administrators to
-gather information about the status of current and past jobs, data movement, and
-pilot factory. PanDA Monitoring allows users access to all log files of each
-job, greatly simplifying code debugging and failure analysis in a distributed
-computing environment.
+% -----------------------------------------------------------------------------
+% PANDA MONITORING - SPACE SAVING
+% -----------------------------------------------------------------------------
+% \paragraph{\textbf{PanDA Monitoring}} this subsystem is implemented as a
+% web-based, dashboard-style graphical application~\cite{klimentov2011atlas}. It
+% runs on Apache and interacts with the back-end database of PanDA Server to
+% enable persistence. PanDA Monitoring enables users and site administrators to
+% gather information about the status of current and past jobs, data movement,
+% and pilot factory. PanDA Monitoring allows users access to all log files of
+% each job, greatly simplifying code debugging and failure analysis in a
+% distributed computing environment.
+% -----------------------------------------------------------------------------
 
@@
 conversion between the two can by static or dynamic. During the LHC Run 1, PanDA
-required users or the application layer to perform a static conversion between
-tasks and jobs: tasks were described as a set of jobs with a fixed number of
-events and then submitted to the PanDA Server (Fig.~\ref{fig:architecture}:A1).
+required users or applications to perform a static conversion between tasks and
+jobs: tasks were described as a set of jobs with a fixed number of events and
+then submitted to the PanDA Server (Fig.~\ref{fig:architecture}:A1).
 
@@
 in terms of jobs: Users conceive analyses in terms of one or more, possibly
-related tasks; the `job' abstraction is required exclusively by the execution
-middleware, i.e. PanDA. Further, a static partitioning of tasks into jobs does
-not take into account the heterogeneity and dynamicity of the resources of the
-pilots on which each job will be executed.
+related tasks; the `job' abstraction is required by the execution middleware,
+i.e. PanDA. Further, a static partitioning of tasks into jobs does not take into
+account the heterogeneity and dynamicity of the resources of the pilots on which
+each job will be executed.
 
@@
 
-JEDI was deployed for the LHC Run 2 to address these inefficiencies with a
-machine-driven conversion of tasks into jobs. Users or the application layer (i)
-submits tasks descriptions to JEDI; (ii) JEDI partitions tasks to jobs of
+JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users or
+the application layer submits tasks descriptions to JEDI
+(Fig.~\ref{fig:architecture}:1) that stores them into a queue implemented by a
+database (Fig.~\ref{fig:architecture}:2). Tasks are partitioned into jobs of
 different size, depending on both static and dynamic information about available
-resources; (iii) jobs are bound to sites with resources that best match jobs'
-requirements; and (iv) jobs are submitted to the PanDA Server for execution
-(Fig.~\ref{fig:architecture}:A2-B).
+resources (Fig.~\ref{fig:architecture}:3). Jobs are bound to sites with
+resources that best match jobs' requirements, and submitted to the PanDA Server
+for execution (Fig.~\ref{fig:architecture}:4).
 
 Once submitted to the PanDA Server, jobs are stored by the Task Buffer component
-into a global queue implemented as a relational database
-(Fig.~\ref{fig:architecture}:C). When jobs are submitted directly to the PanDA
-Server, the Brokerage component is used to bind jobs to available sites,
-depending on static information about the resources available for each site.
-Jobs submitted by JEDI are instead already bound to resources so no further
-brokerage is needed. Once jobs are bound to sites, the Brokerage module
-communicates to the Data Service module what data sets need to be made available
-on what site. The Data Service communicates these requirements to the ATLAS DDM
+into a global queue implemented as a  database (Fig.~\ref{fig:architecture}:5).
+When jobs are submitted directly to the PanDA Server, the Brokerage component is
+used to bind jobs to available sites, depending on static information about the
+resources available for each site. Jobs submitted by JEDI are already bound to
+sites so no further brokerage is needed. Once jobs are bound to sites, the
+Brokerage module communicates to the Data Service module what data sets need to
+be made available on what site (Fig.~\ref{fig:architecture}:6). The Data Service
+communicates these requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7)
 that, when needed, takes care of aggregating files into datasets and containers,
-replicating them on the required sites.
+replicating them on the required sites (Fig.~\ref{fig:architecture}:8).
 
 The AutoPyFactory subsystem communicates with the Task Buffer component of the
-PanDA Server, acquiring information about jobs that have been bound and are
-ready for execution. On the base of this information, AutoPyFactory defines a
-set of suitable PanDA Pilots, queues them into APFQueues, and concurrently
-submits them to multiple sites.
+PanDA Server, acquiring information about jobs that are ready for execution on
+specific (type of) sites (Fig.~\ref{fig:architecture}:9). On the base of this
+information, AutoPyFactory defines a set of suitable PanDA Pilots and
+concurrently submits them a Condor-G broker (Fig.~\ref{fig:architecture}:10).
+This broker submits these pilots wrapped as jobs or VMs to the required sites
+(Fig.~\ref{fig:architecture}:11).
 
@@
 
-When PanDA Pilots become available on sites, they
+When a PanDA Pilot becomes available on a site, it
 % bootstrap, download their code, collect information about the execution
 % environment and
-call the Job Dispatcher module of the PanDA Server, requesting for a job to
-execute \mtnote{Can a pilot request multiple jobs to execute
-concurrently?}\sergeynote{In principle it should be able to do that, but that's
-not how it is operated. One pilot - one job is a cleaner model in an environment
-where resources are "owned" by the organization that runs pilots. Simplifies
-many things. At least as far as I know.}. The Job Dispatcher communicates with
+calls the Job Dispatcher module of the PanDA Server, requesting for a job to
+execute (Fig.~\ref{fig:architecture}:12). The Job Dispatcher communicates with
 the Task Buffer, requesting for a job that is bound to the site of that pilot
@@
 database) and if such a job is available, it passes the job handler to the Job
-Dispatcher. The Job Dispatcher dispatches the job to the PanDA Pilot.
+Dispatcher. The Job Dispatcher dispatches the job to the PanDA Pilot
+(Fig.~\ref{fig:architecture}:13).
+
+% \mtnote{Can a pilot request multiple jobs to execute
+% concurrently?}\sergeynote{In principle it should be able to do that, but
+% that's not how it is operated. One pilot - one job is a cleaner model in an
+% environment where resources are ``owned'' by the organization that runs
+% pilots. Simplifies many things. At least as far as I know.}
 
@@
 subprocess for the execution of the job's payload. The job is setup, input data
-are transferred from the designated staging in location, and the job's payload
-is executed. Once completed, output is transferred to the staging out location.
+are transferred from the designated staging in location
+(Fig.~\ref{fig:architecture}:14), and the job's payload is executed
+(Fig.~\ref{fig:architecture}:15). Once completed, output is transferred to the
+staging out location (Fig.~\ref{fig:architecture}:16).
+
 % During the job's payload execution, the monitoring process of PanDA Pilot
@@
 The Data Service module of the PanDA Server tracks and collects the output
-generated by each job, updating jobs' attributes via the Task Buffer module.
-When the output of all the jobs of a task are retrieved, they are aggregated and
-made available to the user either via the PanDA Server of the JEDI service,
-depending on how tasks or jobs were initially submitted.
+generated by each job (Fig.~\ref{fig:architecture}:17), updating jobs'
+attributes via the Task Buffer module (Fig.~\ref{fig:architecture}:18). When the
+output of all the jobs of a task are retrieved, it is made available to the user
+via PanDA Server. When a task is submitted to JEDI, task is marked as done
+(Fig.~\ref{fig:architecture}:19) and the result of its execution is made
+available to the user by JEDI (Fig.~\ref{fig:architecture}:20).
 
-\mtnote{Think whether a comparison with other WMS based on the review we did
-makes sense in this context.}\sergeynote{You provided a nice description of the
-PanDA system here.}
+% \mtnote{Think whether a comparison with other WMS based on the review we did
+% makes sense in this context.}\sergeynote{You provided a nice description of
+% the PanDA system here.}

### 2017-03-25 paper.tex
limits: 10 pages without references
modified title, as the "on Abstractions".. was incorrect.   open to further refinement.
Fix header

@@
 \documentclass[sigconf, screen]{acmart}
+
+% -----------------------------------------------------------------------------
+% PAGE LIMIT
+% Papers must be submitted in PDF format (readable by Adobe Acrobat Reader 5.0
+% and higher) and formatted for 8.5" x 11" (U.S. Letter). Submissions are
+% limited to 10 pages in the ACM format (see
+% http://www.acm.org/sigs/publications/proceedings-templates), using the
+% sample-sigconf template. The 10-page limit includes figures, tables, and
+% appendices, but does not include references, for which there is no page limit.
+% -----------------------------------------------------------------------------
 
@@
 %Conference
-\acmConference[SC2017]{The International Conference for High Performance Computing, Networking, Storage and Analysis}{November 2017}{Denver, Colorado USA}
+\acmConference[SC2017]{The International Conference for High Performance Computing, Networking, Storage and Analysis}{Nov 2017}{Denver, Colorado}
 \acmYear{2017}
@@
 
-\title{A Case Study on Abstractions to Converge High-Throughput and High-Performance Computing}
+\title{Converging High-Throughput and High-Performance Computing: A Case Study}
 

### 2017-03-26 introduction.tex
ongoing work in the introduction.

--- 2017-03-19 introduction.tex
@@
 multiple experiments  at LHC have collected and distributed hundreds of
-petabytes of data worldwide to hundreds of computer centers.
-Thousands of physicists analyze petascale data volumes daily.
-One of the LHC experiments, the ATLAS~\cite{Aad:2008}, utilizes the Production
-and Distributed Analysis (PanDA) workload management system~\cite{Maeno2011}
-(WMS) for distributed data processing and analysis. The ATLAS Computing
-model~\cite{Atlas2005} is based on a Grid paradigm~\cite{Foster:1998}, with
-multilevel, hierarchically distributed computing and storage resources. PanDA
-has been developed to meet growing ATLAS production and analysis requirements
-for a data-driven workload management system capable of operating at LHC data
-processing scale.
+petabytes of data worldwide to hundreds of computer centers. Thousands of
+physicists analyze petascale data volumes daily. The detection of the Higgs
+Boson speaks to the success of the detector and experiment design, as well as
+the sophistication of computing systems devised to analyze the data.
+
+Historically the computing systems consisted of the federation of a hundreds
+to thousands of distributed resources \textemdash{} ranging in scale from
+small to mid-size resource~\cite{foster2003grid}. Although the workloads to be
+executed are independent, the management of the distribution of extremely-
+large workloads across many heterogeneous resources to ensure the effective
+utilization of resources and efficient execution of workloads presents non-
+trivial challenges.
+
+Many software solutions have been developed in response to these challenges.
+One of the LHC experiments, the CMS experiment devised a solution based around
+the HTCondor~\cite{XX} software ecosystem. The ATLAS~\cite{Aad:2008}
+experiment, utilizes the Production and Distributed Analysis (PanDA) workload
+management system~\cite{Maeno2011} (WMS) for distributed data processing and
+analysis. The CMS and ATLAS experiments represent arguably the largest
+production grade distributed computing solutions and have symbolized the
+paradigm of {\it high-throughput computing} viz., the effective execution of
+many independent workloads.
+
+% Historically, the ATLAS Computing model~\cite{jones2008atlas} is based on a
+% Grid paradigm~\cite{foster2003grid}, with multilevel, hierarchically
+% distributed computing and storage resources.
+
+% PanDA has been developed to meet
+% growing ATLAS production and analysis requirements for a data-driven workload
+% management system capable of operating at LHC data processing scale.
 
@@
 
+\subsubsection*{The following is a candidate for major compression into a single paragraph}
+
 \mtnote{Moved from related work were we use already this terminology. To be

### 2017-03-26 nge.tex

--- 2017-03-19 nge.tex
@@
-As seen in \S\ref{sec:panda_deployment}, the absence of wide area network
-connectivity on Titan's worker nodes was one of the main reasons for PanDA
-Broker not to implement pilots. As discussed in \S\ref{sec:panda_titan}, the
-absence of pilots imposes the static coupling between PBS jobs and detector
-simulations, and development tailored to each type of payload that needs to be
-executed.
+\subsection{Rationale} \aanote{I think that the names of the states should be
+removed. They are meaningless without the figures. I left them just in case we
+decide to reintroduce the figures.}\mtnote{Done.} As seen in
+\S\ref{sec:panda_deployment}, the absence of wide area network connectivity on
+Titan's worker nodes was one of the main reasons for PanDA Broker not to
+implement pilots. As discussed in \S\ref{sec:panda_titan}, the absence of pilots
+imposes the static coupling between MPI scripts submitted to the PBS batch
+system and detector simulations, and development tailored to each type of
+payload that needs to be executed.
 
-The static coupling between PBS jobs and detector simulations makes impossible
-to schedule multiple generations of workload on the same PBS job. Specifically,
-once a number of detector simulations are packaged into a PBS job and this job
-is queued on Titan, no further simulations can be added to that job. New
-simulations have to be packaged into a new PBS job that need to be submitted to
-Titan on the base of the backfill availability of that moment.
+This static coupling makes impossible to schedule multiple generations of
+workload on the same PBS job. Specifically, once a number of detector
+simulations are packaged into a PBS job and this job is queued on Titan, no
+further simulations can be added to that job. New simulations have to be
+packaged into a new PBS job that needs to be submitted to Titan on the base of
+the backfill availability of that moment.
 
 The support of workload generations would enable a more efficient use of the
-backfill availability walltime. Currently, when a set of simulations ends also
-the PBS job ends, independently on whether more walltime would still be
+backfill availability of walltime. Currently, when a set of simulations ends
+also the PBS job ends, independently on whether more walltime would still be
 available. With a pilot, more simulations could be executed so to utilize all
 the available walltime, while avoiding further job packaging and submission
-overheads.\mtnote{Are there more reasons for wanting multiple generations?}
+overheads.
 
-Multiple generations would also help with the transition from an execution model
-based on a fixed amount of event per simulation (100 at the moment) to one in
-which the number of events is decided on the base of the available walltime.
-Pilots would enable to package simulations with a homogeneous number of events,
-especially when core and walltime availabilities would diverge as when a small
-amount of cores is available for a long amount of time or \textit{vice versa}.
-In this situations, a PBS job would would require all simulation to have either
-a large or a small number of events; a pilot would instead support multiple
-generations of simulations with a different number of events and therefore
-duration.
+% \mtnote{Are there more reasons for wanting multiple generations?}
+
+Multiple generations would also help with relaxing two assumptions of the
+current execution model: knowing the number of simulations before submitting the
+MPI script, having a fixed amount of events per simulation (100 at the moment).
+Pilots would enable the scheduling of simulations independently from whether
+they were available at the moment of submitting the pilot. Further, simulations
+with a varying number of events could be scheduled on a pilot, depending on the
+amount of remaining walltime and the distribution of execution time per event,
+as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event-distrib}. This
+capabilities would be particularly useful to increase the PanDA Broker
+efficiency for availabilities with a wide delta between the number of cores and
+walltime.
+
+% to one in which the number of events is decided on the base of the available
+% walltime. Pilots would enable to package simulations with a heterogeneous
+% number of events, especially when a small amount of cores is available for a
+% long period of time and \textit{vice versa}. In these situations, a PBS job
+% would require all simulations to have a small number of events that would not
+% use all the available walltime.
+
+% a pilot would instead support multiple generations of simulations with a
+% different number of events and therefore duration.
 
@@
 mechanics of coordination and communication among multiple worker nodes. This
-could eliminate the need for packaging payload into PBS jobs within the broker,
-greatly simplifying the submission process. This simplification would also
-enable the submission of different types of payload, without having to develop a
-specific PBS script for each payload. The submission process would also be
-MPI-independent, as MPI is used for coordination among multiple worker nodes,
-not by the payload.
+could eliminate the need for packaging payload into MPI scripts within the
+broker, greatly simplifying the submission process. This simplification would
+also enable the submission of different types of payload, without having to
+develop a specific PBS script for each payload. The submission process would
+also be MPI-independent, as MPI is used for coordination among multiple worker
+nodes, not by the payload.
 
-\mtnote{I tried to argue that pilots facilitates the submission to a normal
-queue but I was not able to find a convincing argumentation. When we consider
-only Titan, a PBS job is the only way to submit a job to its queue(s). With NGE,
-we translate a pilot description to a PBS job and then submit it to a queue via
-a PBS command. PanDA Brokers do something very similar: they package detector
-simulations into a PBS job and they submit it via a PBS command. The fact that
-PanDA Brokers define the number of cores and walltime of a PBS job on the base
-of backfill information, does not make the packaging or the submission processes
-any different: PanDA Brokers are already submitting to the nomal queue. If I got
-it right, PanDA Brokers could already submit a job with, say, 12000 cores and 2
-hours walltime. They don't do it because they have no allocation and they would
-wait a long time in the queue (maybe, let's see what our experiments say about
-that). I do not see how NGE could make all this any different.}
+% \mtnote{I tried to argue that pilots facilitates the submission to a normal
+% queue but I was not able to find a convincing argumentation. When we consider
+% only Titan, a PBS job is the only way to submit a job to its queue(s). With %
+% NGE, we translate a pilot description to a PBS job and then submit it to a
+% queue via a PBS command. PanDA Brokers do something very similar: they
+% package detector simulations into a PBS job and they submit it via a PBS
+% command. The fact that PanDA Brokers define the number of cores and walltime
+% of a PBS job on the base of backfill information, does not make the packaging
+% or the submission processes any different: PanDA Brokers are already
+% submitting to the nomal queue. If I got it right, PanDA Brokers could already
+% submit a job with, say, 12000 cores and 2 hours walltime. They don't do it
+% because they have no allocation and they would wait a long time in the queue
+% (maybe, let's see what our experiments say about that). I do not see how NGE
+% could make all this any different.}
 
@@
 % -----------------------------------------------------------------------------
-\subsection{Next Generation Executer}
-\label{ssec:nge}
+\subsection{Architecture}
+\label{sec:arch}
 
-\begin{enumerate}
-    \item Rationale and design (why)
-    \item Architecture (how)
-    \item Integration
-    \item Characterization (experiments)
-\end{enumerate}
+The implementation of pilot capabilities within the current PanDA Broker require
+quantification of the effective benefits that it could yield and, on the base of
+this analysis, a dedicated engineering effort. We developed a prototype of a
+pilot system capable of executing on Titan to study experimentally the
+quantitative and qualitative benefits that it could bring to PanDA. We called
+this prototype Next Generation Executor (NGE).
 
-We are developing a prototype of Next Generation Executer (NGE) to add pilot
-capabilities to the PanDA Broker. NGE is \ldots
+% We developed and we are currently integrating a prototype of Next Generation
+% Executor (NGE) to add pilot capabilities to the PanDA Broker.
+
+NGE is a runtime system designed to execute heterogeneous and dynamic workloads
+on diverse resources. Fig.~\ref{fig:arch-overview} illustrates its current
+architecture as deployed on Titan: the two management modules represent a
+simplified version of the PanDA Broker while the agent module is the pilot
+submitted to Titan and executed on its worker nodes. The communication between
+PanDA Broker and PanDA Server is abstracted away as not immediately useful to
+evaluate the performance and capabilities of a pilot on Titan.
+
+% Thus, workloads and pilots are described via the Pilot API and passed to the
+% NGE runtime system. NGE launches the pilots and executes the tasks of the
+% workload on them.
+
+\begin{figure}
+  \centering
+   \includegraphics[width=\columnwidth]{figures/architecture_compact_rp_paper.pdf}
+  % \caption{RP overview. An application uses the Pilot API to describe pilots
+  %   (green squares) and units (red circles). The PilotManager (purple box)
+  %   instantiates (dash arrow) pilots, the UnitManager (purple box)
+  %   instantiates (solid arrow) units. Both managers are executed on the user
+  %   workstation. Pilots are launched (dashed arrows) on resources A and B
+  %   via SAGA API, an Agent (purple box) is bootstrapped for each pilot
+  %   (green box), units (solid red circles) are scheduled to the Agents via
+  %   MongoDB and executed by the Agent's Executer (yellow box).
+  \caption{NGE Architecture as deployed on Titan. The PilotManager and the
+  UnitManager reside on one of Titan's DTN while the Agent is executed on
+  Titan's worker nodes. Boxes color coding: gray for entities external to NGE,
+  white for APIs, purple for NGE's modules, green for pilots, yellow for
+  module's components.\mtnote{Change the picture eliminating Resource B.}}
+\label{fig:arch-overview}
+\end{figure}
+
+% Internally, NGE represents pilots as aggregates of resources independent from
+% the architecture and topology of the target machines, and workloads as a set
+% of units to be executed on the resources of the pilot. Both pilots and units
+% are stateful entities, each with a well-defined state model and life cycle.
+% Their states and state transitions are managed via the three modules of the
+% NGE architecture: PilotManager, UnitManager, and Agent
+% (Fig.~\ref{fig:arch-overview}, purple boxes).
+
+NGE exposes an API for the application layer to describe workloads
+(Fig.~\ref{fig:arch-overview}, green squares) and pilots
+(Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
+and a UnitManager. The PilotManager submits (Fig.~\ref{fig:arch-overview}, dash
+arrow) pilots to Titan's PBS batch system via SAGA API, as done by PanDA Broker
+to submit MPI scripts. Once scheduled, the pilot's Agent is bootstrapped on
+Titan's workers nodes, and  the UnitManager schedules
+(Fig.~\ref{fig:arch-overview}, solid arrow) units to the Agent's Executer for
+execution. The UnitManager and the Agent communicate via a database that is
+instantiated on one of Titan's DTN so to be reachable by both modules. A similar
+approach would be used to enable PanDA Broker to schedule pilots instead of MPI
+scripts.
+
+% The lifespan of pilots has four states distributed among the PilotManager,
+% resource, and pilot instance.% (Fig.~\ref{fig:pilot-state-model}).
+% Pilots are instantiated in the state \T{NEW} by the PilotManager, wait in a
+% queue to be launched, and transition to \T{PM\_LAUNCH} when submitted to a
+% Resource Manager (RM) via the SAGA API\@. Pilots wait in the queue of the RM
+% and, once scheduled, become \T{P\_ACTIVE}. They remain in this state until the
+% end of their lifetime, when they transition to \T{DONE}.
+
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=\columnwidth]{figures/pilot_state_model_compact_rp_paper.pdf}
+%  \caption{Pilot State Model. Instantiated in state \T{NEW}, each pilot is
+%    launched (\T{PM\_LAUNCH}) via SAGA API on a resource manager. After
+%    becoming \T{P\_ACTIVE}, it exhausts its duration ending in state
+%    \T{DONE}. The transition between states is sequential and each transition
+%    can be canceled or fail ending in the states \T{CANCELED} or \T{FAILED} (not
+%    depicted to improve the diagram clarity). Boxes color coding as per
+%    Fig.~\ref{fig:arch-overview}, with: orange for queues and blue for states.
+%\label{fig:pilot-state-model}}
+%\end{figure}
+%
+
+% The unit state model has nine states distributed across the UnitManager,
+% MongoDB instance, and Agent.% (Fig.~\ref{fig:unit-state-model}).
+% Instantiated in the state \T{NEW} by the UnitManager, every unit is scheduled
+% on an Agent (\T{UM\_SCHEDULING}) via a queue on a MongoDB instance. The unit
+% is then scheduled on the required number of cores held by the Agent's pilot
+% (\T{A\_SCHEDULING}), and finally executed (\T{A\_EXECUTING}).
+
+% The unit state model pertains also to the input and output data of the units.
+% When required, the input data of a unit are either pushed to the Agent
+% (\T{U\_STAGING\_IN}) or pulled from the Agent (\T{A\_STAGING\_IN}), depending
+% on data locality and sharing requirements. Similarly, the output data of the
+% unit are staged out by the Agent and UnitManager (\T{A\_STAGING\_OUT},
+% \T{U\_STAGING\_OUT}) to a specified destination, e.g., the user workstation.
+% Both input and output staging are optional, depending on the requirements of
+% the units. The actual file transfers are enacted via SAGA, and support
+% (gsi)-scp, (gsi)-sftp, and Globus Online.
+
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=\columnwidth]{figures/unit_state_model_compact_rp_paper.pdf}
+%  \caption{Unit State Model. Instantiated in state \T{NEW}, each unit is
+%    scheduled to an Agent (\T{UM\_SCHEDULING}) via MongoDB, and then scheduled
+%    to an Agent's Executer (\T{A\_SCHEDULING}). When required, unit input data
+%    are staged to the Agent's Pilot (\T{UM\_STAGING\_IN} or
+%    \T{A\_STAGING\_IN}). Each unit is executed (\T{A\_EXECUTING}), unit output
+%    data are staged out (\T{A\_STAGING\_OUT} and \T{UM\_STAGING\_OUT}), and
+%    each unit ends in state (\T{DONE}). Color coding and state omissions
+%    as per Fig.~\ref{fig:pilot-state-model}.
+%\label{fig:unit-state-model}}
+%\end{figure}
+
+%The state transitions represented in Figures~\ref{fig:pilot-state-model}
+%and~\ref{fig:unit-state-model} are sequential and every transition can fail or
+%be canceled by the PilotManager or UnitManager.
+
+% All state transitions are managed by the PilotManager, UnitManager, and Agent
+% components. The only special case is the transition of the pilots to the state
+% \T{P\_ACTIVE} which is dictated by the resource's RM, but managed by the
+% PilotManager.
+
+% Above all NGE's components, the Agent is the one that is worth to mention the
+% most.
+
+Agents of NGE run on MOM nodes or worker nodes, using the \emph{Open Run-Time
+Environment (ORTE)}. ORTE is a spin-off from the Open-MPI project and is a
+critical component of the OpenMPI implementation. It was developed to support
+distributed high-performance computing applications operating in a heterogeneous
+environment. The system transparently provides support for interprocess
+communication, resource discovery and allocation, and process launching across a
+variety of platforms. ORTE provides a mechanism similar to the Pilot concept -
+it allows the user to create a \emph{``dynamic virtual machine''} (DVM) that
+spans multiple nodes. ORTE provides libraries to enable the submission,
+monitoring and managing of tasks, avoiding filesystem bottlenecks and race
+conditions with network sockets. As a consequence, it is able to minimize the
+system overhead while submitting tasks.
+
+%Each component of RP has a independent semantic scope. This enables modularity
+%isolating implementation complexity and supporting diverse use cases and
+%environments.  For example, unit scheduling can be implemented by exchangeable
+%Scheduler components, suitable for applications of diverse scales, with
+%different coordination patterns, and executed on Beowulf clusters or Cray
+%machines.
+%
+%Components are also designed to be stateless and instantiated concurrently. In
+%this way, RP can manage multiple pilots and units at the same time, resulting
+%in scalable throughput and tolerance to failing components. Concurrent
+%components are coordinated via a dedicated communication mesh which incurs
+%infrastructure and runtime overhead, offset by the lower component complexity
+%and improved overall scalability of the system.
+
+% \begin{enumerate}
+%     \item Rationale and design (why)
+%     \item Architecture (how)
+%     \item Integration (Not present)
+%     \item Characterization (experiments)
+% \end{enumerate}
+
+ % \ldots

### 2017-03-26 introduction.tex
introduction. first quick pass.

@@
 many independent workloads.
+
+As the LHC prepares for the high-luminosity era (Run 3 in $~\approx$ 2022), it
+is anticipated that the data volumes that will need analyzing will increase by
+10X compared to the current phase (Run 2).   The data will be larger in volume
+but will also require heterogeneous computational processes. In spite of the
+impressive scale of the ATLAS distributed computing system, demand for
+computing systems will significantly outstrip supply (availability).
+
+There are multiple levels at which this problem needs to be addressed
+urgently, e.g., the utilization of emerging parallel architectures (e.g.,
+platforms), algorithmic and advances in analytical methods (e.g., use of
+Machine Learning) and the ability to exploit different platforms (e.g., clouds
+and supercomputers).
+
+This paper is a case study of how the ATLAS experiment has "broken free" of
+the traditional computational approach of high-throughput computing on
+distributed resources to embrace new platforms, in particular high-performance
+computers. Specifically, we discuss the experience of integrating the PanDA
+workload management system with Titan ~\textemdash{} a DOE leadership
+computing facility and scaled to analyze up to \jhanote{XX\%} of the ATLAS
+experiment workload. We present the design of PanDA and show how localized
+substitution to use well established pilot-concepts allow enhanced support for
+heterogeneous workloads (such as molecular dynamics) and advanced execution
+modes.
+
+This state-of-practice paper provides multiple contributions.  (i) Documents
+the many design and operational considerations that have been  taken to
+support the sustained, scalable and production usage of Titan for historically
+high-throughput workloads, (ii) Extensions to PanDA to support non-traditional
+heterogeneous workloads and execution modes,  and (iii) As the community looks
+forward to designing the next generation of online analytical platforms
+~\cite{foap} the lessons learnt from our project provide some guidance for how
+current and future experimental and observational systems can be integrated
+with supercomputers in production.
 

### 2017-03-26 introduction.tex
introduction - some pruning.

@@
 
-% Historically, the ATLAS Computing model~\cite{jones2008atlas} is based on a
-% Grid paradigm~\cite{foster2003grid}, with multilevel, hierarchically
-% distributed computing and storage resources.
+% \subsubsection*{The following is a candidate for major compression into a single paragraph}
 
-% PanDA has been developed to meet
-% growing ATLAS production and analysis requirements for a data-driven workload
-% management system capable of operating at LHC data processing scale.
+% \mtnote{Used the following in related work. We may want to reduce/remove it from
+% the introduction: ``Today, PanDA serves several thousand users, managing job
+% distribution to hundreds of ATLAS sites with more than 100,000 CPU cores which
+% process more than a million jobs per day.''} % (Figure~\ref{fig:daily}).
 
-% PanDA has a highly scalable architecture. Scalability has been demonstrated in
-% ATLAS through the rapid increase in usage over the past several years of
-% operations, and PanDA is expected to meet the continuously growing computing
-% requirements of ATLAS over the next decade. PanDA was designed to have the
-% flexibility to adapt to emerging computing technologies in processing, storage,
-% networking as well as the underlying software stack. This flexibility has also
-% been successfully demonstrated through the past six years: computing centers in
-% ATLAS, spanning many continents, were seamlessly integrated into PanDA\@.
+% In this paper, we describe how PanDA has been engineered to execute a specific
+% stage of the ATLAS Monte Carlo workflow on Titan, the larger high-performance
+% computing HPC system currently available in the USA\@.\mtnote{Explain the
+% benefits offered by Titan in terms of multithreading per node and possibly large
+% amount of concurrent nodes. Introduce also the notion of backfill.} This extends
+% the scope of PanDA's compute model, integrating both high-throughput and
+% high-performance computing resources and enabling the concurrent execution of
+% both  single and multi-core jobs. The integration of PanDA and Titan went
+% through three main engineering phases: (i) feasibility study and rapid
+% prototyping of an initial solution; (ii) progressing scaling of the  prototype
+% to saturate the available resources; (iii) study of a product-grade architecture
+% for generic HPC resources. Both phase i and ii have been completed enabling the
+% execution of up to eight million jobs a week on Titan. A prototype has been
+% engineered to support phase iii and experimental data are being collected.
 
-% PanDA manages a wide spectrum of workloads, ranging from raw data processing to
-% Monte Carlo simulation and user analysis, while constantly evolving to meet
-% rapidly changing science needs.
+% In the next section we introduce \ldots.
 
-\mtnote{Used the following in related work. We may want to reduce/remove it from
-the introduction: ``Today, PanDA serves several thousand users, managing job
-distribution to hundreds of ATLAS sites with more than 100,000 CPU cores which
-process more than a million jobs per day.''} % (Figure~\ref{fig:daily}).
+% Why Titan?
+% \begin{itemize}
+%     \item A lot of slow (relative the grid) and homogeneous cores together
+%     \item Grid is saturated
+%     \item Enable HPC as a calss of computing resource
+%     \item Enable future DOE experimental and observational capabilities on HPC
+%     \item Why is so important: more data, run 2 and run 3.
+%     \item growth of grid is flat (economic model) and saturated. We need more CPU because we have more data.
+%     \item ATLAS spend most of the time on simulations this is why we want to offload simulations - that happen happen to be performed via AthenaMP.
+% \end{itemize}
 
-% \begin{figure}
-%     \begin{center}
-%         \includegraphics[width=\columnwidth]{figures/DailyJobs.png}
-%         \caption{Daily completed jobs on ATLAS Grid for the past 12 month}
-%     \end{center}
-% \label{fig:daily}
-% \end{figure}
+% \mtnote{Moved from related work were we use already this terminology. To be
+% iterated/adjusted for consistency once the first draft will be ready.}
 
-In this paper, we describe how PanDA has been engineered to execute a specific
-stage of the ATLAS Monte Carlo workflow on Titan, the larger high-performance
-computing HPC system currently available in the USA\@.\mtnote{Explain the
-benefits offered by Titan in terms of multithreading per node and possibly large
-amount of concurrent nodes. Introduce also the notion of backfill.} This extends
-the scope of PanDA's compute model, integrating both high-throughput and
-high-performance computing resources and enabling the concurrent execution of
-both  single and multi-core jobs. The integration of PanDA and Titan went
-through three main engineering phases: (i) feasibility study and rapid
-prototyping of an initial solution; (ii) progressing scaling of the  prototype
-to saturate the available resources; (iii) study of a product-grade architecture
-for generic HPC resources. Both phase i and ii have been completed enabling the
-execution of up to eight million jobs a week on Titan. A prototype has been
-engineered to support phase iii and experimental data are being collected.
+% The term ``workflow'' is used in many disciplines with different meaning. In the
+% field of scientific computing, ``workflow'' assumes different meanings depending
+% on the characteristics of the computation, of the software tools used to support
+% this computation, and of the resources on which it is performed. Further, a
+% workflow may indicate a whole application, a description of the computational
+% process of that application or, more commonly, a series of tasks related by data
+% dependences.
 
-In the next section we introduce \ldots.
+% The lack of a consistent and shared definition of ``workflow'' hinders the
+% understanding of its properties and its relations with related concepts. For
+% example, we need to clarify the difference among ``workflow'', ``workload'',
+% ``task'', or ``job'' but also between workflow ``template'' and ``instance'', or
+% ``data-flow'' and ``control-flow''. This is precondition to specify properly the
+% design of software systems that support the execution of scientific
+% applications.
 
-Why Titan?
-\begin{itemize}
-    \item A lot of slow (relative the grid) and homogeneous cores together
-    \item Grid is saturated
-    \item Enable HPC as a calss of computing resource
-    \item Enable future DOE experimental and observational capabilities on HPC
-    \item Why is so important: more data, run 2 and run 3.
-    \item growth of grid is flat (economic model) and saturated. We need more CPU because we have more data.
-    \item ATLAS spend most of the time on simulations this is why we want to offload simulations - that happen happen to be performed via AthenaMP.
-\end{itemize}
+% In this paper, we use the following definitions:
 
-\subsubsection*{The following is a candidate for major compression into a single paragraph}
+% \begin{description}
 
-\mtnote{Moved from related work were we use already this terminology. To be
-iterated/adjusted for consistency once the first draft will be ready.}
+%   \item[Task.] A set of operations to be performed on a computing platform,
+%   alongside a description of the properties and dependences of those operations,
+%   and indications on how the dependences should be satisfied and the operations
+%   should be executed.
 
-The term ``workflow'' is used in many disciplines with different meaning. In the
-field of scientific computing, ``workflow'' assumes different meanings depending
-on the characteristics of the computation, of the software tools used to support
-this computation, and of the resources on which it is performed. Further, a
-workflow may indicate a whole application, a description of the computational
-process of that application or, more commonly, a series of tasks related by data
-dependences.
+%   \item[Job.] A unit of  work performed by submitting a script to a resource
+%   management system (LRMS), like  Slurm or PBS, or by requesting a virtual
+%   machine or a container to a site supporting virtualization. One or more jobs
+%   can perform the operations described with a task.
 
-The lack of a consistent and shared definition of ``workflow'' hinders the
-understanding of its properties and its relations with related concepts. For
-example, we need to clarify the difference among ``workflow'', ``workload'',
-``task'', or ``job'' but also between workflow ``template'' and ``instance'', or
-``data-flow'' and ``control-flow''. This is precondition to specify properly the
-design of software systems that support the execution of scientific
-applications.
+%   \item[Workload.] A set of jobs that can be executed concurrently, possibly
+%   related by a set of relations. For example, jobs of a workload can share one
+%   or more input files or communicate during execution.
 
-In this paper, we use the following definitions:
+%   \item[Workflow.] Set of jobs, related by a set of relations that define the
+%   order in which each task can be executed. Data dependences are the most common
+%   relations among workloads, used to define the precedence among their
+%   executions.
 
-\begin{description}
-
-  \item[Task.] A set of operations to be performed on a computing platform,
-  alongside a description of the properties and dependences of those operations,
-  and indications on how the dependences should be satisfied and the operations
-  should be executed.
-
-  \item[Job.] A unit of  work performed by submitting a script to a resource
-  management system (LRMS), like  Slurm or PBS, or by requesting a virtual
-  machine or a container to a site supporting virtualization. One or more jobs
-  can perform the operations described with a task.
-
-  \item[Workload.] A set of jobs that can be executed concurrently, possibly
-  related by a set of relations. For example, jobs of a workload can share one
-  or more input files or communicate during execution.
-
-  \item[Workflow.] Set of jobs, related by a set of relations that define the
-  order in which each task can be executed. Data dependences are the most common
-  relations among workloads, used to define the precedence among their
-  executions.
-
-\end{description}
+% \end{description}

### 2017-03-26 appendix.tex
First iteration appendix template

Reproducibility initiative appendices: Artifact Description (AD) and Computational Results Analysis (CRA). Description of the organization of the appendix.

% -----------------------------------------------------------------------------
\subsection{Section 5.1 - Figure 3}
\label{apndx:fig3}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 5.1 - Figure 4}
\label{apndx:fig4}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 5.1 - Figure 5}
\label{apndx:fig5}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 5.2 - Figure 6}
\label{apndx:fig6}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 5.2 - Figure 7}
\label{apndx:fig7}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 5.3 - Figure 8}
\label{apndx:fig8}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 14}
\label{apndx:fig14}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 15}
\label{apndx:fig15}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 17}
\label{apndx:fig16}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 17}
\label{apndx:fig17}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

### 2017-03-26 data/ATLAS/plot_run_time_for_different_cpus_BNL_PROD_MCORE.py
ATLAS raw data

@@ -1,0 +1,526 @@
+#!/bin/env python
+#from ROOT import *
+from  ROOT import TFile
+from  ROOT import TH1F
+from  ROOT import TCanvas
+from  ROOT import TText
+from  ROOT import TGraph
+from  ROOT import TMultiGraph
+from  ROOT import TLegend
+from  ROOT import gStyle
+from  ROOT import gPad
+
+from array import array
+
+from datetime import datetime
+
+
+#
+# option to save a canvas
+#
+save_option = False
+#save_option = True
+
+import json
+from pprint import pprint
+
+input_file = 'task_5387358__data.json'
+input_file = "task_5890530_data.json"
+input_file = "task_7353647_data_cleaned.json"
+
+input_file = "task_7244729_data_SIGNET.json"
+input_file = "task_7244951_data_BNL_PROD_MCORE.json"
+
+
+# Output file name
+out_name = input_file.split(".")[0] + ".root"
+
+# Task ID
+
+task_id = input_file.split(".")[0].split("_")[1]
+print task_id
+
+print out_name
+
+
+
+with open(input_file) as data_file:    
+    data = json.load(data_file)
+
+#pprint(data)
+
+n_tests = len(data)
+print n_tests
+
+#exit()
+#---------------------------------------------------------------
+system ='titan'
+
+task   = task_id
+
+n_tests = 180
+t_limit = 1800.
+
+hist_name_1 = 'h1'
+hist_title_1 = 'Stage-In time for finished jobs'
+
+hist_name_1a = 'h1a'
+hist_title_1a = 'Stage-In time for finished jobs. T_stage_in > 0 s'
+
+hist_name_2 = 'h2'
+hist_title_2 = 'Run time for finished jobs'
+
+hist_name_3 = 'h3'
+hist_title_3 = 'Stage-Out time for finished jobs'
+
+hist_name_4 = 'h4'
+hist_title_4 = 'Stage-Out time for finished jobs. T_stage_out > 600 s'
+
+
+h_1=TH1F(hist_name_1, hist_title_1, n_tests, 0., t_limit)
[…]

### 2017-03-26 data/ATLAS/plot_run_time_for_different_cpus_SIGNET.py

@@ -1,0 +1,365 @@
+#!/bin/env python
+#from ROOT import *
+from  ROOT import TFile
+from  ROOT import TH1F
+from  ROOT import TCanvas
+from  ROOT import TText
+from  ROOT import TGraph
+from  ROOT import TMultiGraph
+from  ROOT import TLegend
+from  ROOT import gStyle
+from  ROOT import gPad
+
+from array import array
+
+from datetime import datetime
+
+
+#
+# option to save a canvas
+#
+save_option = False
+#save_option = True
+
+import json
+from pprint import pprint
+
+input_file = 'task_5387358__data.json'
+input_file = "task_5890530_data.json"
+input_file = "task_7353647_data_cleaned.json"
+
+input_file = "task_7244729_data_SIGNET.json"
+#input_file = "task_7244951_data_BNL_PROD_MCORE.json"
+
+
+# Output file name
+out_name = input_file.split(".")[0] + ".root"
+
+# Task ID
+
+task_id = input_file.split(".")[0].split("_")[1]
+print task_id
+
+print out_name
+
+
+
+with open(input_file) as data_file:    
+    data = json.load(data_file)
+
+#pprint(data)
+
+n_tests = len(data)
+print n_tests
+
+#exit()
+#---------------------------------------------------------------
+system ='titan'
+
+task   = task_id
+
+n_tests = 180
+t_limit = 1800.
+
+hist_name_1 = 'h1'
+hist_title_1 = 'Stage-In time for finished jobs'
+
+hist_name_1a = 'h1a'
+hist_title_1a = 'Stage-In time for finished jobs. T_stage_in > 0 s'
+
+hist_name_2 = 'h2'
+hist_title_2 = 'Run time for finished jobs'
+
+hist_name_3 = 'h3'
+hist_title_3 = 'Stage-Out time for finished jobs'
+
+hist_name_4 = 'h4'
+hist_title_4 = 'Stage-Out time for finished jobs. T_stage_out > 600 s'
+
+
+h_1=TH1F(hist_name_1, hist_title_1, n_tests, 0., t_limit)
[…]

### 2017-03-27 appendix.tex
Appendix Figure 8 and partial Figure 5

@@
 
+Replicability:
+
+\begin{enumerate}
+    \item
+\end{enumerate}
+
 % -----------------------------------------------------------------------------
@@
 
-Artifacts description.
+Correlation between backfill availability of cores and walltime on Titan during the experiment time window.
 
 \begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
+    \item Raw data: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.txt}, as provided by the PanDA Brokers logs on Titan.
+    \item Wrangling: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.csv}
+    \item Analysis and Plotting: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}
 \end{itemize}
+
+Replicability:
+
+\begin{enumerate}
+    \item
+\end{enumerate}
 
@@
 
-Artifacts description.
+Panda failures on Titan by exit code during the experiment window.
 
 \begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
+    \item Raw data: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.csv}
+    \item Wrangling: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Pivot Data \& Wrangled Data tabs)
+    \item Analysis: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Aggregated Data tab)
+    \item Plotting: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Plots tab)
 \end{itemize}
+
+Replicability:
+
+\begin{enumerate}
+    \item Download raw data download in csv format from the ATLAS jobs dashboard.
+Download link: \url{http://dashb-atlas-job.cern.ch/dashboard/request.py/terminatedjobsstatuscsv?sites=All%20T3210&sitesCat=All%20Countries&resourcetype=All&pandares=ORNL_Titan_MCORE&activities=simul&sitesSort=7&sitesCatSort=0&start=2016-01-01&end=2017-02-28&timeRange=daily&sortBy=16&granularity=Monthly&generic=0&series=30&type=abcb}
+    \item Load the csv file into Excel (or any other software/language).
+    \item Create a pivot table with failure types as columns and months as rows.
+    \item Aggregate data of error codes of the same type by merge columns and adding the values of their raws according to this table:
+    \item Plot the resulting table as a stacked area diagram.
+\end{enumerate}
 
@@
 \end{itemize}
+
+Replicability:
+
+\begin{enumerate}
+    \item
+\end{enumerate}
 
@@
 \end{itemize}
+
+Replicability:
+
+\begin{enumerate}
+    \item
+\end{enumerate}

### 2017-03-27 appendix.tex
Appendix Figure 5 and updated notebook

@@
 \begin{enumerate}
-    \item
+    \item Get a log file including regular polling of backfill availability from one or more PanDA Brokers on titan.
+    \item Wrangle the ascii file with the find/replace regular expressions listed at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}, cell \#1.
+    \item Load the Jupyter workbook from a python virtual environment with the module Jupyter installed and all the modules listed at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}, cell \#2.
+    \item Execute each cell of the Jupyter notebook to replicate Figure 5.
 \end{enumerate}

### 2017-03-27 PandaOverview.tex
minor scattered changes in Section 3.   unlock.

@@
 
-PanDA's user model is based on multitenancy and the support of at least two
-types of HEP users: individual researchers and groups executing so called
-`production' workflows. Users are free to submit tasks and workflows to the
-PanDA WMS at any point in time, directly or via dedicated application
-frameworks. Consistently, PanDA's security model is based on separation between
-authentication, authorization and accounting for both single users and group of
-users. Both authentication and authorization are based on digital certificates
-and the X.509 standard and on the virtual organization (VO) abstraction.
+PanDA's user \jhanote{usage} model is based on multitenancy \jhanote{of what?}
+and the support of at least two types of HEP users: individual researchers and
+groups executing so called `production' workflows. Users are free to submit
+tasks and workflows to the PanDA WMS at any point in time, directly or via
+dedicated application frameworks. Consistently, PanDA's security model is
+based on separation between authentication, authorization and accounting for
+both single users and group of users. Both authentication and authorization
+are based on digital certificates and the X.509 standard and on the virtual
+organization (VO) abstraction.
 
@@
 
-PanDA design supports provenance and traceability for both jobs and data.
+PanDA's design supports provenance and traceability for both jobs and data.
 Attributes enable provenance by linking jobs and data items, providing
@@
 The implementation of PanDA consists of several interconnected subsystems, most
-of them built from off-the-shelf and Open Source components. Subsystems
-communicate via dedicated API or HTTP messaging and each subsystems is
+of them built from off-the-shelf and Open Source components. Sbub-systems
+communicate via dedicated API or HTTP messaging and each sub-system is
 implemented by one or more modules. Databases are used to store eventful
@@
 PanDA Pilot, and JEDI. Figure~\ref{fig:architecture} shows a diagrammatic
-representation of these subsystems' architecture. For a full list of subsystems
+representation of these sub-systems' architecture. For a full list of subsystems
 see Ref.~\cite{panda-wiki_url}.
@@
 PanDA Server, acquiring information about jobs that are ready for execution on
-specific (type of) sites (Fig.~\ref{fig:architecture}:9). On the base of this
+specific (type of) sites (Fig.~\ref{fig:architecture}:9). On the basis of this
 information, AutoPyFactory defines a set of suitable PanDA Pilots and
-concurrently submits them a Condor-G broker (Fig.~\ref{fig:architecture}:10).
+concurrently submits them to a Condor-G broker (Fig.~\ref{fig:architecture}:10).
 This broker submits these pilots wrapped as jobs or VMs to the required sites

### 2017-03-27 bckground.tex
comments on related work.

@@
+\jhanote{In this section we should talk less about PanDA and more about other
+WMS -- highlighting the fact that (i) there isn't a comprehensive systems
+(design, architecture and execution properties) discussion of other WMS so
+difficult to compare, (ii) real comparison should be to the Gliden-in WMS
+ecosystem and use this as opportunity to narrate the reasons for the
+divergence. Also might this section be moved towards the rear, as this is 
+a practice paper and not a traditional research paper?}
+
 The Production and Distributed Analysis (PanDA) software is a Workload

### 2017-03-27 nge.tex
quick read, minor tweaks and a comment in 6.1 and 6.2
  did not touch 6.3

@@
 decide to reintroduce the figures.}\mtnote{Done.} As seen in
-\S\ref{sec:panda_deployment}, the absence of wide area network connectivity on
-Titan's worker nodes was one of the main reasons for PanDA Broker not to
-implement pilots. As discussed in \S\ref{sec:panda_titan}, the absence of pilots
-imposes the static coupling between MPI scripts submitted to the PBS batch
-system and detector simulations, and development tailored to each type of
-payload that needs to be executed.
+\S\ref{sec:panda_deployment}, the absence of wide area network
+\jhanote{admittedly i haven't read earlier section in question  carefully, but
+i'm a bit confused about the use of "wide area". wide area is typically used
+for cross data center} connectivity on Titan's worker nodes was one of the
+main reasons for PanDA Broker not to implement pilots. As discussed in
+\S\ref{sec:panda_titan}, the absence of pilots imposes the static coupling
+between MPI scripts submitted to the PBS batch system and detector
+simulations, and development tailored to each type of payload that needs to be
+executed.
 
@@
 
-The support of workload generations would enable a more efficient use of the
-backfill availability of walltime. Currently, when a set of simulations ends
-also the PBS job ends, independently on whether more walltime would still be
-available. With a pilot, more simulations could be executed so to utilize all
-the available walltime, while avoiding further job packaging and submission
-overheads.
+The support of  multiple workload generations would enable a more efficient
+use of the backfill availability of walltime. Currently, when a set of
+simulations ends, the PBS job also ends, independent of whether more wall-time
+would still be available. With a pilot, more simulations could be executed so
+to utilize all the available wall-time, while avoiding further job packaging
+and submission overheads.
 

### 2017-03-27 bckground.tex
Answer Shantenu's comments in II and VI

--- 2017-03-19 bckground.tex
@@
+\jhanote{In this section we should talk less about PanDA and more about other
+WMS -- highlighting the fact that (i) there isn't a comprehensive systems
+(design, architecture and execution properties) discussion of other WMS so
+difficult to compare}\mtnote{What dimension(s) should we use to argue that WMS are difficult to compare?}
+\jhanote{(ii) real comparison should be to the Gliden-in WMS ecosystem and use
+this as opportunity to narrate the reasons for the divergence.}\mtnote{From our
+meetings with Sergey: Before Run 1, ATLAS management become uncomfortable with
+the pace at which other WMS were being developed, loosing trust in their ability
+to deliver production-grade capabilities. ATLAS USA decided to go solo and
+developed their own solution. PanDA proven to be able to deliver the
+usual pilot-based WMS capabilities at the right scale. For that, it was chosen
+over other solutions.}
+\jhanote{Also might this section be moved towards the rear, as this is  a
+practice paper and not a traditional research paper?}\mtnote{Fine with me.}
+
 The Production and Distributed Analysis (PanDA) software is a Workload

### 2017-03-27 nge.tex
Answer Shantenu's comments in II and VI

@@
 \jhanote{admittedly i haven't read earlier section in question  carefully, but
-i'm a bit confused about the use of "wide area". wide area is typically used
-for cross data center} connectivity on Titan's worker nodes was one of the
-main reasons for PanDA Broker not to implement pilots. As discussed in
-\S\ref{sec:panda_titan}, the absence of pilots imposes the static coupling
-between MPI scripts submitted to the PBS batch system and detector
-simulations, and development tailored to each type of payload that needs to be
-executed.
+i'm a bit confused about the use of "wide area". wide area is typically used for
+cross data center}\mtnote{Every PanDA Pilot has to communicate with PanDA
+Server. PanDA Server runs on dedicated resources, outside OLCF. Titan's worker
+nodes cannot communicate with PanDA Server because their connectivity is limited
+to Titan's network. Do you have a suggestion for an alternative to ``wide
+area''?} connectivity on Titan's worker nodes was one of the main reasons for
+PanDA Broker not to implement pilots. As discussed in \S\ref{sec:panda_titan},
+the absence of pilots imposes the static coupling between MPI scripts submitted
+to the PBS batch system and detector simulations, and development tailored to
+each type of payload that needs to be executed.
 

### 2017-03-27 nge.tex
Iterate 6.1

@@
 removed. They are meaningless without the figures. I left them just in case we
-decide to reintroduce the figures.}\mtnote{Done.} As seen in
-\S\ref{sec:panda_deployment}, the absence of wide area network
+decide to reintroduce the figures.}\mtnote{Done.}
+
 \jhanote{admittedly i haven't read earlier section in question  carefully, but
@@
 to Titan's network. Do you have a suggestion for an alternative to ``wide
-area''?} connectivity on Titan's worker nodes was one of the main reasons for
-PanDA Broker not to implement pilots. As discussed in \S\ref{sec:panda_titan},
-the absence of pilots imposes the static coupling between MPI scripts submitted
-to the PBS batch system and detector simulations, and development tailored to
-each type of payload that needs to be executed.
+area''?}
 
-This static coupling makes impossible to schedule multiple generations of
-workload on the same PBS job. Specifically, once a number of detector
-simulations are packaged into a PBS job and this job is queued on Titan, no
-further simulations can be added to that job. New simulations have to be
-packaged into a new PBS job that needs to be submitted to Titan on the base of
-the backfill availability of that moment.
+As seen in \S\ref{sec:panda_deployment}, PanDA Broker was deployed on DTNs
+because of the absence of wide area network connectivity on Titan's worker
+nodes. The lack of pilot capabilities impacts both the efficiency and the
+flexibility of PanDA's execution process. Pilots could help to improve
+efficiency by increasing throughput and by enabling better use of backfill
+utilization. Further, pilots makes easier to support heterogeneous workloads.
+
+As discussed in \S\ref{sec:panda_titan}, the absence of pilots imposes the
+static coupling between MPI scripts submitted to the PBS batch system and
+detector simulations. This makes impossible to schedule multiple generations of
+workload on the same PBS job: Once a number of detector simulations are packaged
+into a PBS job and this job is queued on Titan, no further simulations can be
+added to that job. New simulations have to be packaged into a new PBS job that
+needs to be submitted to Titan on the base of the backfill availability of that
+moment.
 
@@
 current execution model: knowing the number of simulations before submitting the
-MPI script, having a fixed amount of events per simulation (100 at the moment).
-Pilots would enable the scheduling of simulations independently from whether
-they were available at the moment of submitting the pilot. Further, simulations
-with a varying number of events could be scheduled on a pilot, depending on the
-amount of remaining walltime and the distribution of execution time per event,
-as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event-distrib}. This
-capabilities would be particularly useful to increase the PanDA Broker
-efficiency for availabilities with a wide delta between the number of cores and
-walltime.
+MPI script, and having a fixed amount of events per simulation (100 at the
+moment). Pilots would enable the scheduling of simulations independently from
+whether they were available at the moment of submitting the pilot. Further,
+simulations with a varying number of events could be scheduled on a pilot,
+depending on the amount of remaining walltime and the distribution of execution
+time per event, as shown in \S\ref{ssec:panda_titan},
+Figure~\ref{fig:100event-distrib}. This capabilities would be particularly
+useful to increase the PanDA Broker efficiency for availabilities with a wide
+delta between the number of cores and walltime.
 

### 2017-03-27 Experiments.tex
New figures and new data -- experiments

@@
 Experiments consist in running AthenaMP instances by using NGE pilots where each AthenaMP simulates a pre-determined number of events taken from ATLAS workload.	
-We present four groups of experiments in which we test the NGE for weak scalability, weak scalability with resubmission, strong scalability and heterogeous workloads.
+We present four groups of experiments in which we test the NGE for weak scalability, weak scalability with multiple generation, strong scalability and heterogeous workloads.
 
@@
 \begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakStart1.pdf}
-    \caption{Average time to start running AthenaMP for pilot sizes: 250, 500, 1000 and 2000.}
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakOver1.pdf}
+    \caption{Average overhead running AthenaMP for pilot sizes: 250, 500, 1000 and 2000.}
 \label{fig:weakScal1b}
@@
 
-\subsubsection{Weak scalability with resubmission }
+\subsubsection{Weak scalability with multiple generation }
 This experiment is similar to the one presented above but in this case we want to test also the impact of submitting new AthenaMP instances in place of those that end.
@@
 \begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakStart2.pdf}
-    \caption{Average time to start running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakOver2.pdf}
+    \caption{Average overhead running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
 \label{fig:weakScal2b}
@@
 
-%\begin{figure}[!htb]
-%        \includegraphics[width=0.5\textwidth]{./figures/NGE/strongET.pdf}
-%    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
-%\label{fig:strongScala}
-%\end{figure}
-%\begin{figure}[!htb]
-%        \includegraphics[width=0.5\textwidth]{./figures/NGE/strongStart.pdf}
-%    \caption{Average time to start running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
-%\label{fig:strongScalb}
-%\end{figure}
+\begin{figure}[!htb]
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/strongET.pdf}
+    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
+\label{fig:strongScala}
+\end{figure}
+\begin{figure}[!htb]
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/strongOver.pdf}
+    \caption{Average overhead running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
+\label{fig:strongScalb}
+\end{figure}
 
@@
 We performed the experiment by executing at first one AthenaMP per node and then, submitting a Gromacs simulation per core. Each Gromacs simulation requires ~20 minutes.
-We tested the following pilot sizes : 8, 32, 64.
+We tested the following pilot sizes : 8, 16, 32, 64.
+
+\begin{figure}[!htb]
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/MDET.pdf}
+    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
+\label{fig:strongScala}
+\end{figure}
+\begin{figure}[!htb]
+        \includegraphics[width=0.5\textwidth]{./figures/NGE/MDOver.pdf}
+    \caption{Average overhead running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
+\label{fig:MDScal}
+\end{figure}
 

### 2017-03-28 Experiments.tex
Refinement experiments

@@
 Experiments consist in running AthenaMP instances by using NGE pilots where each AthenaMP simulates a pre-determined number of events taken from ATLAS workload.	
-We present four groups of experiments in which we test the NGE for weak scalability, weak scalability with multiple generation, strong scalability and heterogeous workloads.
+We present three groups of experiments in which we test the NGE for weak scalability, weak scalability with multiple generation and strong scalability.
 
-We collect data about the execution time of the pilots and the AthenaMP that are executed within them by considering all stages that these entities pass through during the execution.
+We collected data about the execution time of the pilots and the AthenaMP that have been executed within them by considering all stages of the execution.
 Experiments have been performed out of production by using NGE's pilots on TITAN's batch queue. 
@@
 The experiment consists in running as many AthenaMP instances (also referred as tasks from now-on) as the number of nodes controlled by the pilot.
-Each AthenaMP have been assigned with 100 events whose simulation requires  ~$80$ minutes.
+Each AthenaMP have been assigned with 100 events whose simulation requires  ~$70$ minutes.
 
 Tasks do not experience queue within the pilot since there is one node for each AthenaMP instance.
-Therefore, delays are consequence of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager, as defined in Section \ref{}, that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all the tasks on the nodes.
+Therefore, delays are consequence of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager, as defined in Section \ref{sec:arch}, that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all the tasks on the nodes.
 
@@
 \begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakET1.pdf}
-    \caption{Average pilot execution time against average AthenaMP execution times for pilot sizes: 250, 500, 1000 and 2000.}
+        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakET1.pdf}
+    \caption{Weak scalability: average pilot duration and average duration of a single AthenaMP execution by running on 200, 500, 1000 and 2000 nodes.}
 \label{fig:weakScal1a}
@@
 
-On the one side, we can observe that the average pilot duration is insensitive to the pilot size.
-This indicates that, on average, the overhead introduced by the pilot does not grow significatively with the size of the pilot.
-On the other side, we can notice that the average execution time of AthenaMP is smaller than the average pilot duration.
-This indicates that the overhead of the pilot is not null and can be quantified in ~5 minutes.
-
-\aanote{I think we should say something about the execution time of AthenaMP that is smaller than the others.}
-
-In Figure \ref{fig:weakScal1b}, we provide the time required by the pilot to start running AthenaMP instances.  
-We can notice that the larger is the number of units and the larger is the time required by the pilot to start running them.
-However, we can observe that the time does not grow linearly with the pilot size.
+Firstly, we can observe that, despite some fluctuations due to external factors\footnote{such as the Titan's shared filesystem and the shared database used by the NGE}, the average execution time of AthenaMP oscillates between 4200 and 4800 seconds.
+Secondly, we can observe that in all the cases the gap between AthenaMP execution time and the pilot duration is minimal although it slightly increases with the pilot size.
+In order to allow a better comparison, NGE's overhead is shown in Figure \ref{fig:weakScal1b} where we can notice that the overhead does not grow linearly with the number of units.
 
 \begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakOver1.pdf}
-    \caption{Average overhead running AthenaMP for pilot sizes: 250, 500, 1000 and 2000.}
+        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakOver1.pdf}
+    \caption{Weak scalability: average overhead by running AthenaMP on 200, 500, 1000 and 2000 nodes.}
 \label{fig:weakScal1b}
@@
 
-Figure \ref{fig:weakScal2a} depicts the average pilot duration and  the average execution time of AthenaMP.
+Figure \ref{fig:weakScal2a} depicts the average pilot duration and  the average execution time of five sequential instances of AthenaMP.  
+We can observe that the difference between the two durations is more marked than in the previous experiments.
+Despite this, by observing Figure \ref{fig:weakScal2b}, we can notice that the growth of the overhead is still less than linear.
 
 \begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakET2.pdf}
-    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
+        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakET2.pdf}
+    \caption{Weak scalability with multiple generations: average pilot duration and average duration of five sequential AthenaMP executions by running on 256, 512, 1024 and 2048 nodes.}
 \label{fig:weakScal2a}
@@
 \begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakOver2.pdf}
-    \caption{Average overhead running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
+        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakOver2.pdf}
+    \caption{Weak scalability with multiple generations: average overhead by running AthenaMP on 256, 512, 1024 and 2048 nodes.}
 \label{fig:weakScal2b}
@@
 \subsubsection{Strong scalability}
-This experiments aim to test strong scalability. Therefore, we run the same amount of tasks for different pilot sizes. 
+The last experiments deals with strong scalability. Therefore, we run the same amount of tasks for different pilot sizes. 
 We used a number of AthenaMP instances equal to 2048 and tested pilots with 256, 512, 1024 and 2048 nodes.
+Thus, the number of sequential AthenaMP instances is equal to eight for the smallest pilot and is null for the largest.
 Each AthenaMP instance simulates sixteen events as the previous experiment.
 
-Figure \ref{fig:strongScala} depicts the average pilot duration and  the average execution time of AthenaMP.
+Figure \ref{fig:strongScala} depicts the average pilot duration and  the average execution time of possibly sequential AthenaMP instances.
+We can notice that the difference between the pilot duration and the AthenaMP execution times is almost constant for all the pilot sizes although the overall duration of the pilot decreases linearly with the pilot size.
+This is confirmed by Figure \ref{fig:strongScalb} which depicts the overhead of the pilot.
 
 \begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/strongET.pdf}
-    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
+        \includegraphics[width=0.4\textwidth]{./figures/NGE/strongET.pdf}
+    \caption{Strong scalability: average pilot duration and average duration sequential AthenaMP executions by running on 256, 512, 1024 and 2048 nodes.}
 \label{fig:strongScala}
 \end{figure}
+
 \begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/strongOver.pdf}
-    \caption{Average overhead running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
+        \includegraphics[width=0.4\textwidth]{./figures/NGE/strongOver.pdf}
+    \caption{Strong scalability: average overhead running AthenaMP on 256, 512, 1024 and 2048 nodes.}
 \label{fig:strongScalb}
@@
 
-\subsubsection{Heterogeneous execution}
-
-The last experiment provides a proof of concept about the ability of the NGE to execute heterogeneous workload.
-In particular, we coupled the execution of AthenaMP with the execution of Gromacs to simulate molecular dynamics. 
-We performed the experiment by executing at first one AthenaMP per node and then, submitting a Gromacs simulation per core. Each Gromacs simulation requires ~20 minutes.
-We tested the following pilot sizes : 8, 16, 32, 64.
-
-\begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/MDET.pdf}
-    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
-\label{fig:strongScala}
-\end{figure}
-\begin{figure}[!htb]
-        \includegraphics[width=0.5\textwidth]{./figures/NGE/MDOver.pdf}
-    \caption{Average overhead running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
-\label{fig:MDScal}
-\end{figure}
+%\subsubsection{Heterogeneous execution}
+%
+%The last experiment provides a proof of concept about the ability of the NGE to execute heterogeneous workload.
+%In particular, we coupled the execution of AthenaMP with the execution of Gromacs to simulate molecular dynamics. 
+%We performed the experiment by executing at first one AthenaMP per node and then, submitting a Gromacs simulation per core. Each Gromacs simulation requires ~20 minutes.
+%We tested the following pilot sizes : 8, 16, 32, 64. 
+%
+%\begin{figure}[!htb]
+%        \includegraphics[width=0.5\textwidth]{./figures/NGE/MDET.pdf}
+%    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
+%\label{fig:strongScala}
+%\end{figure}
+%\begin{figure}[!htb]
+%        \includegraphics[width=0.5\textwidth]{./figures/NGE/MDOver.pdf}
+%    \caption{Average overhead running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
+%\label{fig:MDScal}
+%\end{figure}
 

### 2017-03-28 data/NGE/script2.py

@@ -1,0 +1,55 @@
+import dataAnalysis as da
+
+path = "strong"
+states = [da.REC,da.SCHED,da.QUEUE,da.EXEC]
+sampling = 0.1
+output = path+".csv"
+outfile = open(output,"w+")
+#256
+folder = "exp1"
+print("Processing "+folder)
+times = da.extractTimes(path+"/"+folder,states,sampling)
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[5],1.0,perc=0.05)
+outfile.write("1 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[4],1.0,perc=0.05)
+outfile.write("1 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[2],1.0,perc=0.05)
+outfile.write("1 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[3],1.0,perc=0.05)
+outfile.write("1 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +"\n")
+#512
+folder = "exp2"
+print("Processing "+folder)
+times = da.extractTimes(path+"/"+folder,states,sampling)
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[5],1.0,perc=0.05)
+outfile.write("2 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[4],1.0,perc=0.05)
+outfile.write("2 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[2],1.0,perc=0.05)
+outfile.write("2 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[3],1.0,perc=0.05)
[…]

### 2017-03-28 nge.tex
Update rp on titan architecture

@@
   \centering
-   \includegraphics[width=\columnwidth]{figures/architecture_compact_rp_paper.pdf}
+   \includegraphics[width=\columnwidth]{figures/rp_architecture_compact_atlaswms_paper.pdf}
   % \caption{RP overview. An application uses the Pilot API to describe pilots
@@
   white for APIs, purple for NGE's modules, green for pilots, yellow for
-  module's components.\mtnote{Change the picture eliminating Resource B.}}
+  module's components.}
 \label{fig:arch-overview}
@@
 (Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
-and a UnitManager. The PilotManager submits (Fig.~\ref{fig:arch-overview}, dash
-arrow) pilots to Titan's PBS batch system via SAGA API, as done by PanDA Broker
+and a UnitManager. The PilotManager submits pilots to Titan's PBS batch system
+via SAGA API (Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker
 to submit MPI scripts. Once scheduled, the pilot's Agent is bootstrapped on
-Titan's workers nodes, and  the UnitManager schedules
-(Fig.~\ref{fig:arch-overview}, solid arrow) units to the Agent's Executer for
-execution. The UnitManager and the Agent communicate via a database that is
-instantiated on one of Titan's DTN so to be reachable by both modules. A similar
-approach would be used to enable PanDA Broker to schedule pilots instead of MPI
-scripts.
+Titan's MOM node and worker nodes, and  the UnitManager schedules units to the
+Agent's Scheduler (Fig.~\ref{fig:arch-overview}, solid arrow). The Agent's
+Scheduler schedules the units on one or more Agent's Executor for execution. The
+Agent's executors can manage one or more worker nodes, depending on performance
+evaluations.
+
+The UnitManager and the Agent communicate via a database that is instantiated on
+one of Titan's DTN so to be reachable by both modules. A similar approach would
+be used to enable PanDA Broker to schedule pilots instead of MPI
+scripts.\mtnote{Rewrite this after checking it with Andre.}
 
@@
 
-Agents of NGE run on MOM nodes or worker nodes, using the \emph{Open Run-Time
-Environment (ORTE)}. ORTE is a spin-off from the Open-MPI project and is a
-critical component of the OpenMPI implementation. It was developed to support
-distributed high-performance computing applications operating in a heterogeneous
-environment. The system transparently provides support for interprocess
-communication, resource discovery and allocation, and process launching across a
-variety of platforms. ORTE provides a mechanism similar to the Pilot concept -
-it allows the user to create a \emph{``dynamic virtual machine''} (DVM) that
-spans multiple nodes. ORTE provides libraries to enable the submission,
-monitoring and managing of tasks, avoiding filesystem bottlenecks and race
-conditions with network sockets. As a consequence, it is able to minimize the
-system overhead while submitting tasks.
+The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for communication
+and coordination of units execution. ORTE is a spin-off from the Open-MPI
+project and is a critical component of the OpenMPI implementation. It was
+developed to support distributed high-performance computing applications
+operating in a heterogeneous environment. The system transparently provides
+support for interprocess communication, resource discovery and allocation, and
+process launching across a variety of platforms. ORTE provides a mechanism
+similar to the Pilot concept - it allows the user to create a \emph{``dynamic
+virtual machine''} (DVM) that spans multiple nodes. ORTE provides libraries to
+enable the submission, monitoring and managing of tasks, avoiding filesystem
+bottlenecks and race conditions with network sockets. As a consequence, it is
+able to minimize the system overhead while submitting tasks.
 

### 2017-03-28 appendix.tex
Update appendix template with the one provided by SC

@@
+% -----------------------------------------------------------------------------
+\subsection{Abstract}
+
+\input{abstract}
+
+% -----------------------------------------------------------------------------
+\subsection{Description}
+
+% -----------------------------------------------------------------------------
+\subsubsection{Check-list (artifact meta information)}
+
+{\small
+\begin{itemize}
+%  \item {\bf Algorithm: }
+  \item {\bf Program: ATLAS Monte Carlo Workflow, Geant4, AthenaMP, GROMACS}
+%  \item {\bf Compilation: }
+%  \item {\bf Transformations: }
+%  \item {\bf Binary: }
+  \item {\bf Data set: Available at \\ \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/tree/master/data}}
+  \item {\bf Run-time environment: PanDA Workload Management System, RADICAL-Pilot Pilot System (NGE)}
+  \item {\bf Hardware: OLCF Titan Cray XK7}
+%  \item {\bf Run-time state: }
+%  \item {\bf Execution: }
+  \item {\bf Output: Available at \\ \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/tree/master/data}}
+  \item {\bf Experiment workflow: Raw data acquisition, data wrangling and filtering, plotting, analysis}
+%  \item {\bf Experiment customization: }
+  \item {\bf Publicly available?: Yes}
+\end{itemize}
+}
+
+% -----------------------------------------------------------------------------
+\subsubsection{How software can be obtained (if available)}
+
+\begin{itemize}
+    \item PanDA Workload Management System: \\ \url{https://github.com/PanDAWMS}.
+    \item RADICAL-Pilot Pilot System (NGE): \\ \url{https://github.com/radical-cybertools}.
+\end{itemize}
+
+% -----------------------------------------------------------------------------
+\subsubsection{Hardware dependencies}
+
+Access and allocation on OLCF Titan Cray XK7, workstation with at least 8GB of RAM.
+
+% -----------------------------------------------------------------------------
+\subsubsection{Software dependencies}
+
+Python, jupyter, pandas, matplotlib, gnuplot, excel.
+
+% -----------------------------------------------------------------------------
+\subsubsection{Datasets}
+
+\begin{description}[align=left]
+    \item [Section 5.1 -- Figure 3] \url{}
+    \item [Section 5.1 -- Figure 4] \url{}
+    \item [Section 5.1 -- Figure 5] \url{}
+    \item [Section 5.1 -- Figure 6] \url{}
+    \item [Section 5.2 -- Figure 7] \url{}
+    \item [Section 5.2 -- Figure 8] \url{}
+    \item [Section 5.3 -- Figure 9] \url{}
+    \item [Section 6.3 -- Figure 14] \url{}
+    \item [Section 6.3 -- Figure 15] \url{}
+    \item [Section 6.3 -- Figure 16] \url{}
+    \item [Section 6.3 -- Figure 17] \url{}
+    \item [Section 6.3 -- Figure 18] \url{}
+    \item [Section 6.3 -- Figure 19] \url{}
+    \item [Section 6.3 -- Figure 20] \url{}
+    \item [Section 6.3 -- Figure 21] \url{}
+\end{description}
+
+% -----------------------------------------------------------------------------
+\subsection{Installation}
+
+\begin{description}[align=left]
+    \item [Section 5.1 -- Figure 3] \url{}
+    \item [Section 5.1 -- Figure 4] \url{}
+    \item [Section 5.1 -- Figure 5] \url{}
+    \item [Section 5.1 -- Figure 6] \url{}
+    \item [Section 5.2 -- Figure 7] \url{}
+    \item [Section 5.2 -- Figure 8] \url{}
+    \item [Section 5.3 -- Figure 9] \url{}
+    \item [Section 6.3 -- Figure 14] \url{}
+    \item [Section 6.3 -- Figure 15] \url{}
+    \item [Section 6.3 -- Figure 16] \url{}
+    \item [Section 6.3 -- Figure 17] \url{}
+    \item [Section 6.3 -- Figure 18] \url{}
+    \item [Section 6.3 -- Figure 19] \url{}
+    \item [Section 6.3 -- Figure 20] \url{}
+    \item [Section 6.3 -- Figure 21] \url{}
+\end{description}
+
+% -----------------------------------------------------------------------------
+\subsection{Experiment workflow}
+
+\begin{description}[align=left]
+    \item [Section 5.1 -- Figure 3] \url{}
+    \item [Section 5.1 -- Figure 4] \url{}
+    \item [Section 5.1 -- Figure 5] \url{}
+    \item [Section 5.1 -- Figure 6] \url{}
+    \item [Section 5.2 -- Figure 7] \url{}
+    \item [Section 5.2 -- Figure 8] \url{}
+    \item [Section 5.3 -- Figure 9] \url{}
+    \item [Section 6.3 -- Figure 14] \url{}
+    \item [Section 6.3 -- Figure 15] \url{}
+    \item [Section 6.3 -- Figure 16] \url{}
+    \item [Section 6.3 -- Figure 17] \url{}
+    \item [Section 6.3 -- Figure 18] \url{}
+    \item [Section 6.3 -- Figure 19] \url{}
+    \item [Section 6.3 -- Figure 20] \url{}
+    \item [Section 6.3 -- Figure 21] \url{}
+\end{description}
+
+% -----------------------------------------------------------------------------
+\subsection{Evaluation and expected result}
+
+\begin{description}[align=left]
+    \item [Section 5.1 -- Figure 3] \url{}
+    \item [Section 5.1 -- Figure 4] \url{}
+    \item [Section 5.1 -- Figure 5] \url{}
+    \item [Section 5.1 -- Figure 6] \url{}
+    \item [Section 5.2 -- Figure 7] \url{}
+    \item [Section 5.2 -- Figure 8] \url{}
+    \item [Section 5.3 -- Figure 9] \url{}
+    \item [Section 6.3 -- Figure 14] \url{}
+    \item [Section 6.3 -- Figure 15] \url{}
+    \item [Section 6.3 -- Figure 16] \url{}
+    \item [Section 6.3 -- Figure 17] \url{}
+    \item [Section 6.3 -- Figure 18] \url{}
+    \item [Section 6.3 -- Figure 19] \url{}
+    \item [Section 6.3 -- Figure 20] \url{}
+    \item [Section 6.3 -- Figure 21] \url{}
+\end{description}
+
+% ----------------------------------------------------------------------------- % \subsection{Experiment customization}
+
+% -----------------------------------------------------------------------------
+% \subsection{Notes}
+
+% =============================================================================
+\section{OLD -- REMOVE AFTER COMPLETING ABOVE}
 Reproducibility initiative appendices: Artifact Description (AD) and Computational Results Analysis (CRA). Description of the organization of the appendix.
@@
 % -----------------------------------------------------------------------------
-\subsection{Section 6.3 - Figure 17}
+\subsection{Section 6.3 - Figure 16}
 \label{apndx:fig16}

### 2017-03-28 paper.tex
First iteration appendix template
Fix Appendix numbering
Update appendix template with the one provided by SC

@@
 %Conference
-\acmConference[SC2017]{The International Conference for High Performance Computing, Networking, Storage and Analysis}{Nov 2017}{Denver, Colorado}
+\acmConference[SC2017]{The International Conference for High Performance Computing, Networking, Storage and Analysis}{Nov 2017}{Denver, Colorado, USA}
 \acmYear{2017}
@@
 
+% ----------------------------------------------------------------------------
+% APPENDIX
+% ----------------------------------------------------------------------------
+\appendix
+\section{Artifact Description --- Converging High-Throughput and High-Performance Computing: A Case Study}
+\label{apnx:ac}
+
+\input{appendix}
+
 % that's all folks

### 2017-03-28 PandaOverview.tex
Iterate and move related work

@@
+The Production and Distributed Analysis (PanDA) software is a Workload
+Management System (WMS) developed to support the operations of the LHC ATLAS
+experiment. PanDA is specifically design to manage the execution of distributed
+workloads and workflows via pilots.
+
+WMS for distributed executions are a type of middleware, designed to coordinate
+specific activities about resources and workloads: discovering and selecting
+resources, submitting tasks of workloads and workflows, and monitoring the
+execution of those tasks~\cite{marco2009glite}. Pilot is an abstraction that
+enables multi-level scheduling by decoupling resource acquisition from workload
+scheduling~\cite{turilli2015comprehensive}. Pilots are implemented by resource
+placeholders: a request for a certain amount of resources is scheduled on a site
+(e.g., by submitting a job or requesting a virtual machine) and once the
+resources are acquired, tasks are scheduled directly to the pilot, not to site's
+scheduler.
+
+WMS like PanDA implement the pilot abstraction: WMS control the acquisition of
+resources by means of pilots and then manage the execution of workloads or
+workflows' tasks on those pilots. WMS centralize the management of the execution
+process, offering a unified interface to the application layer. Pilots expose a
+uniform scheduling interface and execution environment, abstracting the
+differences among the heterogeneous resources of multiple sites. Pilot-enabled
+WMS are particularly relevant for LHC experiments, where millions of tasks are
+executed across multiple sites every month, analyzing and producing petabytes of
+data.
+
 The design of PanDA WMS started in 2005 and its implementation went into

### 2017-03-28 bckground.tex
Iterate and move related work

@@
 (design, architecture and execution properties) discussion of other WMS so
-difficult to compare}\mtnote{What dimension(s) should we use to argue that WMS are difficult to compare?}
+difficult to compare}\mtnote{Better?}
 \jhanote{(ii) real comparison should be to the Gliden-in WMS ecosystem and use
@@
 \jhanote{Also might this section be moved towards the rear, as this is  a
-practice paper and not a traditional research paper?}\mtnote{Fine with me.}
-
-The Production and Distributed Analysis (PanDA) software is a Workload
-Management System (WMS) developed to support the operations of the LHC ATLAS
-experiment. PanDA is specifically design to manage the execution of distributed
-workloads and workflows via pilots.
-
-WMS for distributed executions are a type of middleware, designed to coordinate
-specific activities about resources and workloads: discovering and selecting
-resources, submitting tasks of workloads and workflows, and monitoring the
-execution of those tasks~\cite{marco2009glite}. Pilot is an abstraction that
-enables multi-level scheduling by decoupling resource acquisition from workload
-scheduling~\cite{turilli2015comprehensive}. Pilots are implemented by resource
-placeholders: a request for a certain amount of resources is scheduled on a site
-(e.g., by submitting a job or requesting a virtual machine) and once the
-resources are acquired, tasks are scheduled directly to the pilot, not to site's
-scheduler.
-
-WMS like PanDA implement the pilot abstraction: WMS control the acquisition of
-resources by means of pilots and then manage the execution of workloads or
-workflows' tasks on those pilots. WMS centralize the management of the execution
-process, offering a unified interface to the application layer. Pilots expose a
-uniform scheduling interface and execution environment, abstracting the
-differences among the heterogeneous resources of multiple sites. Pilot-enabled
-WMS are particularly relevant for LHC experiments, where millions of tasks are
-executed across multiple sites every month, analyzing and producing petabytes of
-data.
+practice paper and not a traditional research paper?}\mtnote{Done.}
 
@@
 
-Currently, the implementation of PanDA WMS has at least two distinguishing
-features: the scale at which it operates distributed computing, and
-network-aware scheduling of tasks to sites\mtnote{PanDA team: please let me know
-whether more unique (i.e., not present in other LHC WMS) features of PanDA
-implementation should be indicated}. PanDA concurrently supports distributed
-computations on up to 240,000 cores, submitting more than 365 million jobs a
-year across 100 sites, and processing up to XXXXPB of data distributed worldwide
-on XX Tier 1-3 facilities. As for 2017, PanDA serves several thousand users,
-including XX production groups. The scale of this operation is unprecedented,
-making PanDA the WMS that manages the largest computational campaign in the
-world.
+% -----------------------------------------------------------------------------
+% NOTES ON TASK, JOBS, AND WORKFLOW TERMINOLOGY
+% -----------------------------------------------------------------------------
+% Currently, the implementation of PanDA WMS has at least two distinguishing
+% features: the scale at which it operates distributed computing, and
+% network-aware scheduling of tasks to sites\mtnote{PanDA team: please let me
+% know whether more unique (i.e., not present in other LHC WMS) features of
+% PanDA implementation should be indicated}. PanDA concurrently supports
+% distributed computations on up to 240,000 cores, submitting more than 365
+% million jobs a year across 100 sites, and processing up to XXXXPB of data
+% distributed worldwide on XX Tier 1-3 facilities. As for 2017, PanDA serves
+% several thousand users, including XX production groups. The scale of this
+% operation is unprecedented, making PanDA the WMS that manages the largest
+% computational campaign in the world.
 
-PanDA was initially implemented to support computations across an unreliable and
-relatively expensive networking infrastructure. Accordingly, PanDA assumed a
-rigid model of data placement based on replication and long-term storage of data
-on the Grid sites. This model proven too wasteful, requiring storage space for
-unused and obsolete data. PanDA implementation was evolved to benefit from the
-development of the internet and of the networking infrastructure for scientific
-research. Today, PanDA constantly monitors network throughput and latency among
-its management components, Grid sites and data facilities. PanDA uses this
-information to schedule tasks on sites, depending on the estimated time required
-to download and/or replicate input data and to stage out output data.
+% PanDA was initially implemented to support computations across an unreliable
+% and relatively expensive networking infrastructure. Accordingly, PanDA
+% assumed a rigid model of data placement based on replication and long-term
+% storage of data on the Grid sites. This model proven too wasteful, requiring
+% storage space for unused and obsolete data. PanDA implementation was evolved
+% to benefit from the development of the internet and of the networking
+% infrastructure for scientific research. Today, PanDA constantly monitors
+% network throughput and latency among its management components, Grid sites and
+% data facilities. PanDA uses this information to schedule tasks on sites,
+% depending on the estimated time required to download and/or replicate input
+% data and to stage out output data.
 
-The dynamic management of data replication and download is an example of a
-generalized paradigm shift that distinguishes the implementation of PanDA from
-other WMS. New components are being implemented to support: (i) dynamic sizing
-of input dataset based on resource capabilities; (ii) dynamic sizing of the
-resources held by pilots both in terms of number of cores and duration of the
-pilot; and (iii) further abstraction of different types of resources to
-reconcile high-throughput and high-performance computing paradigms.
+% The dynamic management of data replication and download is an example of a
+% generalized paradigm shift that distinguishes the implementation of PanDA from
+% other WMS. New components are being implemented to support: (i) dynamic sizing
+% of input dataset based on resource capabilities; (ii) dynamic sizing of the
+% resources held by pilots both in terms of number of cores and duration of the
+% pilot; and (iii) further abstraction of different types of resources to
+% reconcile high-throughput and high-performance computing paradigms.
+% -----------------------------------------------------------------------------
 
-% \subsection{Terminology}
-%
-% \mtnote{To be moved to the end of the introduction.}
-
+% -----------------------------------------------------------------------------
+% NOTES ON TASK, JOBS, AND WORKFLOW TERMINOLOGY
+% -----------------------------------------------------------------------------
 % Each task may have an arbitrary number of properties like number of cores,
@@
 % \end{itemize}
+% -----------------------------------------------------------------------------

### 2017-03-28 overall.tex
Iteration PanDA I/O Impact at OLCF

@@
 statistical breakdown of the observed file I/O impact of ATLAS at OLCF\@.
-Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written} show
-the file read and write I/O histograms for these 1,175 jobs, respectively.
-Figures~\ref{fig:atlas-titan-file-open} and~\ref{fig:atlas-titan-file-close}
-show the file $open()$ and $close()$ metadata load histograms of the same 1,175
-ATLAS jobs, respectively.
+% Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written}
+% show the file read and write I/O histograms for these 1,175 jobs,
+% respectively. Figures~\ref{fig:atlas-titan-file-open}
+% and~\ref{fig:atlas-titan-file-close} show the file $open()$ and $close()$
+% metadata load histograms of the same  1,175 ATLAS jobs, respectively.
+%
+% As can be seen from Table~\ref{panda-olcf-stats},
 
-As can be seen from Table~\ref{panda-olcf-stats}, the number of nodes used by
-ATLAS jobs vary between 1 and 300, while the average is at 35. 75\% of the ATLAS
-jobs consume less than 25 and 92\% consume less than or equal to 100 Titan
-compute nodes. During the 174 hours of data collection, we observed that 6.75
-ATLAS jobs were executed on average per hour on Titan and ran for 1.74 hours on
-average.
-
-ATLAS jobs issues a large number of file read operations, as can be seen in
-Table~\ref{panda-olcf-stats}. The maximum amount read by any ATLAS job in
-aggregate in this observed period was less than 250 GB and the maximum amount
-written in aggregate was less than 75 GB\@. The average amount read per job is
-20 GB and average amount written is 6 GB\@.
-
-\begin{figure}[!htb]
-    \centering
-    \begin{tabular}{cc}
-        {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_finer_hist.pdf}}\\
-    \end{tabular}
-    \caption{ATLAS file read operation histogram on Titan for week of 10/25/16.}
-\label{fig:atlas-titan-io-read}
-\end{figure}
-
-\begin{figure}[!htb]
-    \centering
-    \begin{tabular}{cc}
-        {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_finer_hist.pdf}}\\
-    \end{tabular}
-    \caption{ATLAS file write I/O histogram on Titan for week of 10/25/16.}
-\label{fig:atlas-titan-io-written}
-\end{figure}
-
-Per job read and write file I/O statistics show an interesting pattern.
-Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
-compute node on Titan is less than 400 MB on average, while the amount of data
-written per node is less than 170 MB on average. This correlates with our
-finding that ATLAS PanDA jobs are read heavy. However, as can be seen in
-Table~\ref{panda-olcf-stats} and figures {fig:atlas-titan-io-read} and
-{fig:atlas-titan-io-written}, the distribution between read and written amount
-of data per job is quite different from one another. The read operation
-distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
-the written amount of data has a very narrow distribution.
-
-\begin{figure}[!htb]
-    \centering
-    \begin{tabular}{cc}
-        {\includegraphics[width=0.48\textwidth]{figures/panda_file_open_hist.pdf}}\\
-    \end{tabular}
-    \caption{ATLAS file $open()$ histogram on Titan for week of 10/25/16.}
-\label{fig:atlas-titan-file-open}
-\end{figure}
-
-\begin{figure}[!htb]
-    \centering
-    \begin{tabular}{cc}
-        {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
-    \end{tabular}
-    \caption{ATLAS file $close()$ histogram on Titan for week of 10/25/16.}
-\label{fig:atlas-titan-file-close}
-\end{figure}
-
-On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file $open()$
-operations and 5 file $close()$ operations per second. The file $open()$
-operations listed here don't include the file $stat()$ operations. As can be
-seen from figures~\ref{fig:atlas-titan-file-open}
-and~\ref{fig:atlas-titan-file-close}, they exhibit similar distributions. The
-maximum number of file $open()$ operations are around 170 on average and the
-maximum number of file $close()$ operations are 39 on average per second per
-job. The total number of file $open()$ operations is 172,089,760 for the
-observed window of 174 hours, while the total number of file $close()$
-operations stand at 40,132,992 for the total 1,175 ATLAS PanDA jobs in the same
-observation window. The difference between these two values is puzzling and it
-is under investigation at the time of writing this paper. One possible
-explanation is that ATLAS PanDA jobs perhaps don't call a file $close()$
-operation per every file $open()$ issued.
+ATLAS jobs use between 1 and 300 worker nodes, and 35 on average. 75\% of the
+ATLAS jobs consume less than 25 nodes and 92\% less than 100. During the 174
+hours of data collection, 6.75 ATLAS jobs were executed on average per hour,
+each job running for an average of 1.74 hours.
+% ATLAS jobs issues a large number of file read operations, as can be seen in
+% Table~\ref{panda-olcf-stats}.
+Every job read less than 250 GB and wrote less than 75 GB of data and, on
+average, each job read 20 GB and wrote 6 GB of data.
 
@@
 
-Overall, based on our experiments with real-world jobs, it can be safely
-concluded that the file and metadata I/O load of ATLAS PanDA project on the OLCF
-Titan supercomputing environment and the Spider 2 file system is not detrimental
-to the center operations and the overall impact is minimal, at the current scale
-of the project.
+% \begin{figure}[!htb]
+%     \centering
+%     \begin{tabular}{cc}
+%         {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_finer_hist.pdf}}\\
+%     \end{tabular}
+%     \caption{ATLAS file read operation histogram on Titan for week of 10/25/16.}
+% \label{fig:atlas-titan-io-read}
+% \end{figure}
+%
+% \begin{figure}[!htb]
+%     \centering
+%     \begin{tabular}{cc}
+%         {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_finer_hist.pdf}}\\
+%     \end{tabular}
+%     \caption{ATLAS file write I/O histogram on Titan for week of 10/25/16.}
+% \label{fig:atlas-titan-io-written}
+% \end{figure}
+
+The I/O of the ATLAS job executed on Titan show an interesting pattern.
+Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
+worker node is less than 400 MB on average, while the amount of data written per
+node is less than 170 MB on average. This correlates with our finding that ATLAS
+PanDA jobs are read heavy. However,
+
+ATLAS PanDA jobs are read heavy: On average, the amount of data read per worker
+node is less than 400 MB, while the amount of data written is less than 170 MB.
+This correlates with our finding that . However,
+% as can be seen in Table~\ref{panda-olcf-stats} and figures
+% {fig:atlas-titan-io-read} and {fig:atlas-titan-io-written},
+the distributions of read and written data are quite different: The read
+operation distribution per job shows a long tail, ranging from 12.5 GB to 250
+GB, while the written amount of data has a very narrow distribution.\mtnote{Should we say why?}
+
+% \begin{figure}[!htb]
+%     \centering
+%     \begin{tabular}{cc}
+%         {\includegraphics[width=0.48\textwidth]{figures/panda_file_open_hist.pdf}}\\
+%     \end{tabular}
+%     \caption{ATLAS file $open()$ histogram on Titan for week of 10/25/16.}
+% \label{fig:atlas-titan-file-open}
+% \end{figure}
+%
+% \begin{figure}[!htb]
+%     \centering
+%     \begin{tabular}{cc}
+%         {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
+%     \end{tabular}
+%     \caption{ATLAS file $close()$ histogram on Titan for week of 10/25/16.}
+% \label{fig:atlas-titan-file-close}
+% \end{figure}
+
+The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
+operations per second (not including file $stat()$ operations) and 5 file $close()$ operations per second, with similar distributions.
+% As can be seen from figures~\ref{fig:atlas-titan-file-open}
+% and~\ref{fig:atlas-titan-file-close}, they .
+On average, the maximum number of file $open()$ operations per job is
+$\approx$170/s and the maximum number of file $close()$ operations is
+$\approx$39/s.
+% on average per second per job.
+For the 1,175 ATLAS PanDA jobs observed, the total number of file $open()$
+operations is 172,089,760 and the total number of file $close()$ operations is
+40,132,992. The difference between these two values is still under
+investigation: One possible explanation is that ATLAS PanDA jobs don't call a
+file $close()$ operation per every file $open()$ issued.
+
+Overall, based on our experiments with ATLAS jobs, it can be safely concluded
+that the file and metadata I/O load of ATLAS PanDA project on the OLCF Titan
+supercomputing environment and the Spider 2 file system is not detrimental to
+the center operations. At the current scale of the project, the overall impact
+of ATLAS operations on Titan is minimal.

### 2017-03-28 overall.tex
Eliminate subsection on reliability, second paragraph and figure

@@
 
-\mtnote{Figure summarizing job performance on Titan analogous to those produced
-with NGE, as discussed in F2F meeting at Rutgers?}
+% \mtnote{Figure summarizing job performance on Titan analogous to those
+% produced with NGE, as discussed in F2F meeting at Rutgers?}
 
 % ----------------------------------------------------------------------------
-\subsection{Reliability of PanDA Broker and Detector Simulations on Titan}
-\label{ssec:reliability}
+% \subsection{Reliability of PanDA Broker and Detector Simulations on Titan}
+% \label{ssec:reliability}
 
@@
 
-Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
-recorded by the PanDA Brokers between January 2016 and February 2017. The
-initial preponderance of unknown failures was progressively addressed by a more
-accurate failure model, specifically tailored to the PanDA Broker and Titan.
-Figure~\ref{fig:failures-titan} also shows the progressive improvement of the
-broker's reliability: In Februrary 2017, errors related to input files and
-initialization were almost eliminated and execution errors were comparable to
-those recorded in January 2017 even if almost twice as many simulations were
-performed (Figure~\ref{fig:backfill-utilization}). The increased amount of
-unknown failure in February indicates that furter work is needed to improve the
-PanDA Broker failure model.
+% Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
+% recorded by the PanDA Brokers between January 2016 and February 2017. The
+% initial preponderance of unknown failures was progressively addressed by a
+% more accurate failure model, specifically tailored to the PanDA Broker and
+% Titan. Figure~\ref{fig:failures-titan} also shows the progressive improvement
+% of the broker's reliability: In Februrary 2017, errors related to input files
+% and initialization were almost eliminated and execution errors were comparable
+% to those recorded in January 2017 even if almost twice as many simulations
+% were performed (Figure~\ref{fig:backfill-utilization}). The increased amount
+% of unknown failure in February indicates that furter work is needed to improve
+% the PanDA Broker failure model.
 
-\begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/failures_titan.pdf}
-    \caption{PanDA Broker failures during the experiment time window. Failures
-    were recorded with 52 exit codes, here aggregated into 6 categories. The
-    number of failures should be compared to the number of simulations executed
-    in each month as plotted in Figure~\ref{fig:backfill-utilization}.
-    Improvements in the failure model of PanDA Broker shows a decrease in
-    unknown failure and a the progressive addressing of other categories of
-    failure.}
-\label{fig:failures-titan}
-\end{figure}
+% \begin{figure}[htp]
+%     \includegraphics[clip,width=\columnwidth]{figures/failures_titan.pdf}
+    % \caption{PanDA Broker failures during the experiment time window. Failures
+    % were recorded with 52 exit codes, here aggregated into 6 categories. The
+    % number of failures should be compared to the number of simulations
+    % executed in each month as plotted in
+    % Figure~\ref{fig:backfill-utilization}. Improvements in the failure model
+    % of PanDA Broker shows a decrease in unknown failure and a the progressive
+    % addressing of other categories of failure.}
+% \label{fig:failures-titan}
+% \end{figure}
 

### 2017-03-29 Experiments.tex
Merging figures of the experiments

@@
 
-Figure \ref{fig:weakScal1a} depicts the average pilot duration and the average execution time of AthenaMP as function of the pilot size.
+Figure \ref{fig:weakScal1a} depicts the average pilot duration, the average execution time of AthenaMP and pilot's overhead as function of the pilot size.
 
 \begin{figure}[!htb]
-        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakET1.pdf}
-    \caption{Weak scalability: average pilot duration and average duration of a single AthenaMP execution by running on 200, 500, 1000 and 2000 nodes.}
+        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
+    \caption{Weak scalability: average pilot duration, average duration of a single AthenaMP execution and pilot's overhead for different pilot sizes (250, 500, 1000 and 2000 nodes).
+Note that the overhead refers to the axis on the right and it is expressed in tenths of a second.}
 \label{fig:weakScal1a}
@@
 Firstly, we can observe that, despite some fluctuations due to external factors\footnote{such as the Titan's shared filesystem and the shared database used by the NGE}, the average execution time of AthenaMP oscillates between 4200 and 4800 seconds.
-Secondly, we can observe that in all the cases the gap between AthenaMP execution time and the pilot duration is minimal although it slightly increases with the pilot size.
-In order to allow a better comparison, NGE's overhead is shown in Figure \ref{fig:weakScal1b} where we can notice that the overhead does not grow linearly with the number of units.
+Secondly, we can observe that in all the cases the gap between AthenaMP execution times and the pilot durations is minimal although it slightly increases with the pilot size.
+By observing NGE's overhead, we can notice the overhead does not grow linearly with the number of units.
 
-\begin{figure}[!htb]
-        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakOver1.pdf}
-    \caption{Weak scalability: average overhead by running AthenaMP on 200, 500, 1000 and 2000 nodes.}
-\label{fig:weakScal1b}
-\end{figure}
+%\begin{figure}[!htb]
+%        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakOver1.pdf}
+%    \caption{Weak scalability: average overhead by running AthenaMP on 200, 500, 1000 and 2000 nodes.}
+%\label{fig:weakScal1b}
+%\end{figure}
 
@@
 
-Figure \ref{fig:weakScal2a} depicts the average pilot duration and  the average execution time of five sequential instances of AthenaMP.  
-We can observe that the difference between the two durations is more marked than in the previous experiments.
-Despite this, by observing Figure \ref{fig:weakScal2b}, we can notice that the growth of the overhead is still less than linear.
+Figure \ref{fig:weakScal2a} depicts the average pilot duration, the average execution time of five sequential instances of AthenaMP and the corresponding overhead.  
+We can observe that the difference between the two durations is more marked than in the previous experiments. Despite this we can notice that the growth of the overhead is still less than linear.
 
 \begin{figure}[!htb]
-        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakET2.pdf}
-    \caption{Weak scalability with multiple generations: average pilot duration and average duration of five sequential AthenaMP executions by running on 256, 512, 1024 and 2048 nodes.}
+        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak2.pdf}
+    \caption{Weak scalability with multiple generations: average pilot duration, average duration of five sequential AthenaMP executions and pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes).
+Note that the overhead refers to the axis on the right and it is expressed in tenths of a second.}
 \label{fig:weakScal2a}
 \end{figure}
-\begin{figure}[!htb]
-        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakOver2.pdf}
-    \caption{Weak scalability with multiple generations: average overhead by running AthenaMP on 256, 512, 1024 and 2048 nodes.}
-\label{fig:weakScal2b}
-\end{figure}
+%\begin{figure}[!htb]
+%        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakOver2.pdf}
+%    \caption{Weak scalability with multiple generations: average overhead by running AthenaMP on 256, 512, 1024 and 2048 nodes.}
+%\label{fig:weakScal2b}
+%\end{figure}
 \subsubsection{Strong scalability}
@@
 We can notice that the difference between the pilot duration and the AthenaMP execution times is almost constant for all the pilot sizes although the overall duration of the pilot decreases linearly with the pilot size.
-This is confirmed by Figure \ref{fig:strongScalb} which depicts the overhead of the pilot.
 
 \begin{figure}[!htb]
-        \includegraphics[width=0.4\textwidth]{./figures/NGE/strongET.pdf}
-    \caption{Strong scalability: average pilot duration and average duration sequential AthenaMP executions by running on 256, 512, 1024 and 2048 nodes.}
+        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/strong.pdf}
+    \caption{Strong scalability:  average pilot duration, average duration of sequential AthenaMP executions and pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes).
+Note that the overhead refers to the axis on the right and it is expressed in tenths of a second.}
 \label{fig:strongScala}
@@
 
-\begin{figure}[!htb]
-        \includegraphics[width=0.4\textwidth]{./figures/NGE/strongOver.pdf}
-    \caption{Strong scalability: average overhead running AthenaMP on 256, 512, 1024 and 2048 nodes.}
-\label{fig:strongScalb}
-\end{figure}
+%
+%\begin{figure}[!htb]
+%        \includegraphics[width=0.4\textwidth]{./figures/NGE/strongOver.pdf}
+%    \caption{Strong scalability: average overhead running AthenaMP on 256, 512, 1024 and 2048 nodes.}
+%\label{fig:strongScalb}
+%\end{figure}
 

### 2017-03-29 PandaOverview.tex
Reduction III

--- 2017-03-27 PandaOverview.tex
@@
-The design of PanDA WMS started in 2005 and its implementation went into
-production for the LHC Run 1 on 2009. PanDA was then extended with new
-subsystems to be deployed on Run 2, on 2015. In this section, we summarize the
-design, architecture, and execution process of the current state of the
-practice. We focus on those subsystems and components that are most relevant to
-the the integration of PanDA with supercomputers, the main topic of this paper.
+PanDA is a Workload Management System (WMS) designed to support the execution of
+distributed workloads and workflows via pilots. WMS is middleware for
+discovering and selecting resources, submitting tasks of workloads and
+workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
+abstraction that enables multi-level scheduling by decoupling resource
+acquisition from tasks scheduling~\cite{turilli2015comprehensive}. Once
+implemented, a pilot is scheduled on a site (e.g., by submitting a job or
+requesting a virtual machine) and, once active, tasks are scheduled to the
+pilot, not to the site's scheduler.
+
+Pilot-enabled WMS are particularly relevant for LHC experiments, where millions
+of tasks are executed across multiple sites every month, analyzing and producing
+petabytes of data. The design of PanDA WMS started in 2005 and its
+implementation went into production for the LHC Run 1 on 2009. PanDA was then
+extended with new subsystems to be deployed on Run 2, on 2015. We summarize
+PanDA's design, architecture, and execution process, focussing on those
+subsystems and components that are most relevant to the the integration of PanDA
+with supercomputers.
 
@@
 
-% \begin{itemize}
-%   \item No MPI tasks?
-%   \item Pilot multitenancy?
-%   \item PanDA Server Brokerage component and Event service interaction?
-%   \item PanDA Server Data Service component and Event Service interaction?
-% \end{itemize}
+PanDA's application model assumes tasks, workloads and workflows. Tasks
+represent a set of operations performed on a set of events that are stored in
+one or more input files. Tasks are decomposed into jobs, where each job
+represents the task's set of operations and a partition of the task's events.
+Since 2005, a certain amount of parallelism has been progressively introduced
+for job execution~\cite{crooks2012multi} but, so far, no MPI jobs have been
+considered for production. Jobs are supposed to be relatively self-contained,
+capable of setting up their own execution environment or having a minimal set of
+common dependences.
 
-PanDA's application model assumes single tasks, workloads and workflows. Tasks
-represent a set of operations performed on a set of events, stored in one or
-more input files. Tasks are decomposed into jobs, where each job represents the
-task's set of operations and a partition of the task's events. Since 2005, a
-certain amount of parallelism has been progressively introduced for job
-execution~\cite{crooks2012multi} but, so far, no MPI jobs have been considered
-for production. Jobs are supposed to be relatively self-contained, capable of
-setting up their own execution environment or having a minimal set of common
-dependences.
-
-% Different types of job payload (i.e., executable) are assumed, both in terms
-% of user-defined batch scripts or application frameworks for group of users.
-
-PanDA's user \jhanote{usage} model is based on multitenancy \jhanote{of what?}
-and the support of at least two types of HEP users: individual researchers and
-groups executing so called `production' workflows. Users are free to submit
-tasks and workflows to the PanDA WMS at any point in time, directly or via
-dedicated application frameworks. Consistently, PanDA's security model is
-based on separation between authentication, authorization and accounting for
-both single users and group of users. Both authentication and authorization
-are based on digital certificates and the X.509 standard and on the virtual
-organization (VO) abstraction.
+PanDA's usage model is based on multitenancy of resources and the support of at
+least two types of HEP users: individual researchers and groups executing so
+called `production' workflows. Users are free to submit tasks and workflows to
+the PanDA WMS at any point in time, directly or via dedicated application
+frameworks. Consistently, PanDA's security model is based on separation between
+authentication, authorization and accounting for both single users and group of
+users. Both authentication and authorization are based on digital certificates
+and the X.509 standard and on the virtual organization (VO) abstraction.
 
@@
 
+In PanDA's data model, each event refers to the recorded or simulated
+measurement of a physical event. One or more events can be packaged into files
+or other data containers. As with jobs, data have both attributes and states,
+and some of the attributes are shared between events and jobs. Raw,
+reconstruction, and simulation data are assumed to be distributed across
+multiple storage facilities and managed by the ATLAS Distributed Data Management
+(DDM)~\cite{garonne2012atlas}. When necessary, datasets required by each job are
+assumed to be replicated over the network, both for input and output data.
+
+PanDA's design supports provenance and traceability for both jobs and data.
+Attributes enable provenance by linking jobs and data items, providing
+information like ownership or project affiliation. States enable traceability by
+providing information about the stage of the execution in which each job or data
+item is or has been. Some attributes are assumed to be immutable across
+execution and jobs and data items are assumed to be always in one and only one
+state.
+
+% \begin{itemize}
+%   \item No MPI tasks?
+%   \item Pilot multitenancy?
+%   \item PanDA Server Brokerage component and Event service interaction?
+%   \item PanDA Server Data Service component and Event Service interaction?
+% \end{itemize}
+
+% Different types of job payload (i.e., executable) are assumed, both in terms
+% of user-defined batch scripts or application frameworks for group of users.
+
 % Consistently, pilots are considered to be multitenant as jobs belonging to
@@
 % \mtnote{Is event an abstraction specific to the MD workflow?}.
-
-In PanDA's data model, each event represents a collision captured in the
-detector and one or more events can be packaged into files or data source with
-other container abstractions. As with jobs, data have both attributes and
-states, and some of the attributes are shared between events and jobs.
-% so to create associations.
-Raw, reconstruction, and simulation data are assumed to be distributed across
-multiple storage facilities and managed by the ATLAS Distributed Data Management
-(DDM)~\cite{garonne2012atlas}. Datasets required by each job are assumed to be
-replicated when necessary over the network, both for input and output data.
 
@@
 
-PanDA's design supports provenance and traceability for both jobs and data.
-Attributes enable provenance by linking jobs and data items, providing
-information like ownership or project affiliation. States enable traceability by
-providing information about the stage of the execution in which each job or data
-item is or has been. Some attributes are assumed to be immutable across
-execution and jobs and data items are assumed to be always in one and only one
-state.
-
 % -----------------------------------------------------------------------------
-\subsection{Architecture}
+\subsection{Implementation and Execution}
 \label{ssec:panda_arch}
@@
 The implementation of PanDA consists of several interconnected subsystems, most
-of them built from off-the-shelf and Open Source components. Sbub-systems
-communicate via dedicated API or HTTP messaging and each sub-system is
+of them built from off-the-shelf and Open Source components. Subsystems
+communicate via dedicated API or HTTP messaging, and each subsystem is
 implemented by one or more modules. Databases are used to store eventful
@@
 
-The execution of specific tasks of ATLAS workflows on supercomputers involves
-four main subsystems of the PanDA architecture: PanDA Server, AutoPyFactory,
-PanDA Pilot, and JEDI. Figure~\ref{fig:architecture} shows a diagrammatic
-representation of these sub-systems' architecture. For a full list of subsystems
-see Ref.~\cite{panda-wiki_url}.
-
-% In the following, we briefly describe the overall architecture of PanDA and
-% its components. The description provides all the details required to
-% understand how HPC resources have been integrated within PanDA
-% (\ref{sec:integration}).
-
-% For a more detailed description of PanDA architecture see~\cite{wiki, papers}.
+Currently, PanDA's architecture has five main subsystems: PanDA
+Server~\cite{maeno2011overview},
+AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
+Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
+Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of ATLAS
+workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}) but, at the
+moment, they are not relevant to understand how PanDA has been ported to
+supercomputers. For a full list of subsystems see Ref.~\cite{panda-wiki_url}.
+Figure~\ref{fig:architecture} shows a diagrammatic representation of PanDA main
+subsystems, highlighting the execution process of tasks while omitting
+monitoring details to improve readability.
 
@@
 
-% Originally PanDA was designed for grid infrastructure... In this
-% paper is focussed on the HPC use case; there are differences from
-% the grid use case. This paper discusses how PanDA has been adapted
-% to execute a workload % on HPC resources. The schematic overview is
-% presented for workload X on Titan......
+% -----------------------------------------------------------------------------
+% PANDA SERVER
+% -----------------------------------------------------------------------------
+% \textbf{PanDA Server} is the central hub service of PanDA WMS and it operates
+% as a web service~\cite{maeno2011overview}. It consists of four main components
+% implemented in Python: Task Buffer, Brokerage, Job Dispatcher, and Data
+% Service. Together, these components implement a global queue where to store
+% jobs, share information about jobs and their dataset properties with the other
+% subsystems, and update their states at execution time.
+% -----------------------------------------------------------------------------
 
-\paragraph{\textbf{PanDA Server}} this subsystem is the central hub service of
-PanDA and it operates as a web service~\cite{maeno2011overview}. It runs on
-Apache web server, interacting with the back-end database running on separate
-machines. PanDA Server implements the Apache worker model, with many independent
-processes handling client requests in parallel. Since all state is maintained in
-the central database, the PanDA Server instance itself is stateless. Currently,
-the production version of PanDA utilizes an Oracle database as backend, but
-PanDA Server can also work with the MySQL family of databases.
+% It runs on Apache web server, interacting with the back-end database running
+% on separate machines. PanDA Server implements the Apache worker model, with
+% many independent processes handling client requests in parallel. Since all
+% state is maintained in the central database, the PanDA Server instance itself
+% is stateless. Currently, the production version of PanDA utilizes an Oracle
+% database as backend, but PanDA Server can also work with the MySQL family of
+% databases.
 
@@
 % input data and its locality, and available CPU resources. The PanDA server
-
+%
 % and communicating via Python APIs
 
-PanDA Server consists of four main components implemented in Python: Task
-Buffer, Brokerage, Job Dispatcher, and Data Service. Task Buffer manages PanDA's
-global queue by: (i) storing jobs into the global queue; (ii) providing
-information about and updating the state of each stored job. The Brokerage
-component: (i) matches job attributes with site and pilot attributes; (ii)
-manages the dispatch of input data to processing sites; and (iii) derives
-PanDA's data pre-placement requirement. The Job Dispatcher: (i) receives
-requests from active pilots for new jobs to process; (ii) retrieves from Task
-Buffer the jobs with highest priority and that can be executed on the site of
-the requesting pilot; (iii) creates a wrapper script for the selected job,
-tailored to the site of the requesting pilot; and (iv) monitors the state of
-the job execution on the pilot, updating the Task Buffer component. Finally, the
-Data Service component: (i) retrieves information from Task Buffer and Brokerage
-about the data requirements of each job and the site to which each job has been
-assigned; (ii) interacts with the ATLAS DDM to make input files available to the
-jobs and manage their output data.
+% Task Buffer manages PanDA's global queue by: (i) storing jobs into the global
+% queue; (ii) providing information about and updating the state of each stored
+% job. The Brokerage component: (i) matches job attributes with site and pilot
+% attributes; (ii) manages the dispatch of input data to processing sites; and
+% (iii) derives PanDA's data pre-placement requirement. The Job Dispatcher: (i)
+% receives requests from active pilots for new jobs to process; (ii) retrieves
+% from Task Buffer the jobs with highest priority and that can be executed on
+% the site of the requesting pilot; (iii) creates a wrapper script for the
+% selected job, tailored to the site of the requesting pilot; and (iv) monitors
+% the state of the job execution on the pilot, updating the Task Buffer
+% component. Finally, the Data Service component: (i) retrieves information from
+% Task Buffer and Brokerage about the data requirements of each job and the site
+% to which each job has been assigned; (ii) interacts with the ATLAS DDM to make
+% input files available to the jobs and manage their output data.
 
-% The job/pilot matching algorithm assigns a weight to each candidate pilot and
-% after applying a set of policies, the best ten candidates are taken.
+% -----------------------------------------------------------------------------
+% PANDA PILOT
+% -----------------------------------------------------------------------------
+% \textbf{PanDA Pilot} implements the execution environment for PanDA's
+% jobs\cite{nilsson2011atlas}. Implemented in Python, a PanDA Pilot is launched
+% by a wrapper script submitted to a site. Wrapper scripts are tailored to the
+% middleware exposed by each site so to maintain PanDA Pilot site-independent.
+% PanDA Pilot pulls a job for execution, sets up the execution environment
+% including the required input files, spawns the jobs's executable, monitors its
+% execution, and supports the output's staging.
+% -----------------------------------------------------------------------------
 
-% including  recovering and cleaning up previously failed jobs on that site and
-
-\paragraph{\textbf{PanDA Pilot}} this subsystem implements the execution
-environment for PanDA's jobs\cite{nilsson2011atlas}. Implemented in Python, a
-PanDA Pilot is launched by a wrapper script submitted to a site. Wrapper scripts
-are tailored to the middleware exposed by each site so to maintain PanDA Pilot
-site-independent. PanDA Pilot: (i) prepares the execution environment for
-PanDA's jobs, collecting information about the compute, data, and network
-capabilities of the site; (ii) requests a job for execution to the Job
-Dispatcher component of PanDA Server; (iii) initiates and monitors the execution
-of pulled jobs, including setting up the job, transferring its input when
-required, executing the payload, transferring output, and checking log files and
-local space; and (iv) cleans up when the payload has finished to execute.
+% t: (i) prepares the execution environment for PanDA's jobs, collecting
+% information about the compute, data, and network capabilities of the site;
+% (ii) requests a job for execution to the Job Dispatcher component of PanDA
+% Server; (iii) initiates and monitors the execution of pulled jobs, including
+% setting up the job, transferring its input when required, executing the
+% payload, transferring output, and checking log files and local space; and (iv)
+% cleans up when the payload has finished to execute.
 
@@
 
-Upon bootstrapping, PanDA Pilot downloads major parts of its runtime code via
-HTTP from a central Subversion repository. The repository works with an Apache
-web server, configured with a memory-based web proxy (Squid). The purpose of the
-cache is to reduce the request load on the back-end Subversion server. That
-improves performance since only occasional queries trigger a full lookup on the
-back-end Subversion system, and most external queries are pulled from memory on
-the front-end web server. The benefits of this system include a high-performance
-code download service, combined with code updates still being immediately
-available as soon as they are committed to source code control.
+% Upon bootstrapping, PanDA Pilot downloads major parts of its runtime code via
+% HTTP from a central Subversion repository. The repository works with an Apache
+% web server, configured with a memory-based web proxy (Squid). The purpose of
+% the cache is to reduce the request load on the back-end Subversion server.
+% That improves performance since only occasional queries trigger a full lookup
+% on the back-end Subversion system, and most external queries are pulled from
+% memory on the front-end web server. The benefits of this system include a
+% high-performance code download service, combined with code updates still being
+% immediately available as soon as they are committed to source code control.
 
-\paragraph{\textbf{AutoPyFactory}} this subsystem implements a factory to submit
-pilots locally and to both grid and cloud
-sites~\cite{caballero2012autopyfactory}. Implemented to be executed as a single
-demonized process, AutoPyFactory concurrently spawns multiple job submission
-processes called APFQueue. Each APFQueue communicates with PanDA Server and the
-Brokerage module to collect information about the state of the jobs it manages.
-On the basis of this information, an APFQueue submits new pilots to one site.
-AutoPyFactory implements functionalities to evaluate the number of pilots
-already instantiated on each site, and to apply a consistent and adequate
-pressure on their batch queues. Submission is performed via the Condor-G batch
-system~\cite{frey2002condor}, enabling submission to multiple type of sites.
-AutoPyFactory exposes a dedicated monitoring service, feeding information to the
-PanDA Monitoring subsystem. A proxy manager enables certificate-based
-authentication and authorization for each site.
+% -----------------------------------------------------------------------------
+% AUTOPYFACTORY
+% -----------------------------------------------------------------------------
+% \textbf{AutoPyFactory} implements a factory to submit pilots locally and to
+% both grid and cloud sites~\cite{caballero2012autopyfactory}. Implemented to be
+% executed as a single demonized process, AutoPyFactory concurrently spawns
+% multiple processes called APFQueue that, on the base of information about jobs
+% queued on the PanDA server and on the state and capabilities of resources,
+% submit pilots to multiple sites with Condor-G~\cite{frey2002condor}.
+% -----------------------------------------------------------------------------
+
+% Each APFQueue communicates with PanDA Server and the Brokerage module to
+% collect information about the state of the jobs it manages. On the basis of
+% this information, an APFQueue submits new pilots to one site. AutoPyFactory
+% implements functionalities to evaluate the number of pilots already
+% instantiated on each site, and to apply a consistent and adequate pressure on
+% their batch queues. Submission is performed via the Condor-G batch
+% system~\cite{frey2002condor}, enabling submission to multiple type of sites.
+% AutoPyFactory exposes a dedicated monitoring service, feeding information to
+% the PanDA Monitoring subsystem. A proxy manager enables certificate-based
+% authentication and authorization for each site.
 
@@
 
-\paragraph{\textbf{JEDI}} this subsystem exposes a task abstraction for workload
-and workflows specification and interfaces with PanDA Server to manage their
-execution on remote sites~\cite{borodin2015scaling}. Implemented as a standalone
-service, JEDI partitions tasks to jobs, and binds these jobs to sites via
-dedicated queues. JEDI collects information about sites' resource capability,
-availability, and performance. Based on this information, JEDI associates sites
-with similar capabilities to a queue. Each job is assigned to a queue by
-matching the job's requirements to the capability, availability and performance
-of the sites of that queue. In this way, JEDI replaces the Brokerage module of
-the PanDA Server.
+% -----------------------------------------------------------------------------
+% JEDI
+% -----------------------------------------------------------------------------
+% \textbf{JEDI} exposes a task abstraction for workload and workflows
+% specification and interfaces with PanDA Server to manage their execution on
+% remote sites~\cite{borodin2015scaling}. Implemented as a standalone service,
+% JEDI partitions tasks to jobs, and binds these jobs to sites via dedicated
+% queues. JEDI collects information about sites' resource capability,
+% availability, and performance. Based on this information, JEDI associates
+% sites with similar capabilities to a queue. JEDI replaces and extends the
+% Brokerage module of the PanDA Server.
+% -----------------------------------------------------------------------------
+
+% Each job is assigned to a queue by matching the job's requirements to the
+% capability, availability and performance of the sites of that queue. In this
+% way,
+
 % This module still handles communication between the Task
@@
 % -----------------------------------------------------------------------------
-\subsection{Execution}
-\label{ssec:panda_exec}
+% \subsection{Execution}
+% \label{ssec:panda_exec}
 
@@
 jobs: tasks were described as a set of jobs with a fixed number of events and
-then submitted to the PanDA Server (Fig.~\ref{fig:architecture}:A1).
+then submitted to the PanDA Server.
 

### 2017-03-29 Experiments.tex
Pass on NGE experiments and scattered space savings

@@
 
-In this section we present experiments that aim to determine the performances of the NGE and to show how it provides a minimal overhead while introducing new functionalities.
+In this section we present experiments % that aim
+to determine the performances of the NGE and to show how it provides a minimal
+overhead while introducing new functionalities.
 
-Experiments consist in running AthenaMP instances by using NGE pilots where each AthenaMP simulates a pre-determined number of events taken from ATLAS workload.	
-We present three groups of experiments in which we test the NGE for weak scalability, weak scalability with multiple generation and strong scalability.
+Experiments % consist in running
+run instances of AthenaMP by using NGE pilots, % where each AthenaMP simulates
+simulating a pre-determined number of events in the ATLAS detector.\mtnote{are
+they 100 events?}
+% taken from ATLAS workload.
+We present three groups of experiments in which we test the NGE for weak
+scalability, weak scalability with multiple generation, and strong scalability.
 
-We collected data about the execution time of the pilots and the AthenaMP that have been executed within them by considering all stages of the execution.
-Experiments have been performed out of production by using NGE's pilots on TITAN's batch queue. 
-Because of TITAN's batch queue, the turnaround time of each run of our experiments is dominated by the time spent on TITAN's queue.
-Since we are interested only in the performances of the NGE, we removed queue time from our statistics.
+We % collected data about
+measured the execution time of the pilots and of the AthenaMP % that have been
+executed within them, collecting timestamps at % by considering
+all stages of the execution. Experiments have been performed % out of production
+by % using
+submitting NGE's pilots % on
+to TITAN's batch queue. Because of TITAN's % batch queue
+policies, the turnaround time of each run of our experiments is dominated by
+% the time spent on TITAN's queue
+queue time. Since we are interested only in the performances of the NGE, we
+removed queue time from our statistics.
 
-All the experiments have been performed by setting AthenaMP in such a way that it uses all the 16 cores present on TITAN's nodes.
+All the experiments have been performed by % setting
+configuring AthenaMP
+% in such a way that it uses
+to use all the 16 cores % present on
+of TITAN's worker nodes.
 
@@
 
-The experiment consists in running as many AthenaMP instances (also referred as tasks from now-on) as the number of nodes controlled by the pilot.
-Each AthenaMP have been assigned with 100 events whose simulation requires  ~$70$ minutes.
+% The
+In this experiment % consists in running
+we run as many AthenaMP instances (also referred as tasks from now-on) as the number of nodes controlled by the pilot. Each AthenaMP
+% have been assigned with 100 events whose simulation requires
+simulates 100 events, requiring ~$70$ minutes.
 
-Tasks do not experience queue within the pilot since there is one node for each AthenaMP instance.
-Therefore, delays are consequence of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager, as defined in Section \ref{sec:arch}, that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all the tasks on the nodes.
+Tasks do not % experience
+wait within the NGE Agent's queue % within the pilot
+since % there is
+one node %for
+is available to each AthenaMP instance. % Therefore, d
+Delays in tasks execution are consequence of % only
+three other factors: (i) the bootstrapping of the pilot on the nodes; (ii) the
+UnitManager, as defined in Section~\ref{sec:arch}, that has to dispatch tasks to
+the agent; and (iii) the time that the agent requires to spawn all the tasks on
+the nodes.
 
-We tested different pilot sizes, i.e. : 250, 500, 1000 and 2000. For all of them the walltime was 2 hours.
-
-Figure \ref{fig:weakScal1a} depicts the average pilot duration, the average execution time of AthenaMP and pilot's overhead as function of the pilot size.
+We tested % different
+pilots % sizes, i.e. :
+with 250, 500, 1000 and 2000 worker nodes and
+%. For all of them, the walltime was
+2 hours walltime. Figure~\ref{fig:weakScal1a} depicts the average pilot
+duration, the average execution time of AthenaMP, and the pilot's overhead as
+function of the pilot size.
 
@@
         \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
-    \caption{Weak scalability: average pilot duration, average duration of a single AthenaMP execution and pilot's overhead for different pilot sizes (250, 500, 1000 and 2000 nodes).
-Note that the overhead refers to the axis on the right and it is expressed in tenths of a second.}
+    \caption{Weak scalability: average pilot duration, average duration of a
+    single AthenaMP execution, and pilot's overhead for different pilot sizes
+    (250, 500, 1000 and 2000 worker nodes). Note: the overhead refers to
+    the axis on the right and it is expressed in tenths of a second.}
 \label{fig:weakScal1a}
@@
 
-Firstly, we can observe that, despite some fluctuations due to external factors\footnote{such as the Titan's shared filesystem and the shared database used by the NGE}, the average execution time of AthenaMP oscillates between 4200 and 4800 seconds.
-Secondly, we can observe that in all the cases the gap between AthenaMP execution times and the pilot durations is minimal although it slightly increases with the pilot size.
-By observing NGE's overhead, we can notice the overhead does not grow linearly with the number of units.
+% Firstly, w
+We % can
+observe that, despite some fluctuations due to external factors
+% \footnote{such as the
+(e.g., Titan's shared filesystem and the shared database used by the NGE), the
+average execution time of AthenaMP % oscillates
+ranges between 4200 and 4800 seconds. % Secondly, w
+We % can
+also observe that in all the cases the gap between AthenaMP execution times and
+the pilot durations is minimal, although it slightly increases with the pilot
+size. % By observing NGE's overhead, w
+We % can
+notice that NGE's overhead does not grow linearly with the number of units.
 
@@
 \subsubsection{Weak scalability with multiple generation }
-This experiment is similar to the one presented above but in this case we want to test also the impact of submitting new AthenaMP instances in place of those that end.
-The reason behind this experiment is to stress pilot's components by pushing new tasks while others are ending their execution.
 
-In order to perform the experiments in less than two hours per run, we decided to send for execution a number of AthenaMP instances equal to five times the number of nodes and we reduced  the number of events that are simulated by each Athena-MP to sixteen\footnote{Note that the number of events has been chosen in such a way that all the cores of a node run one event.}.
-This allows us to complete a single AthenaMP in ~20 minutes only.
+This experiment is similar to the one presented above but, in this case, we want
+to test also the impact of submitting multiple generations of % new
+AthenaMP to the same pilot.\mtnote{we used `generations' at the beginning of 6,
+I am assuming the reader understands the term by now. Please change if you
+disagree.}
+% instances in place of those that end.
+% The reason behind this experiment is to
+In this way, we stress the pilot's components, % by pushing new tasks while others are ending their execution
+as new tasks are scheduled for execution on the Agent while other tasks are still running.
 
-We tested different pilot sizes, i.e. : 256, 512, 1024 and 2048. For all of them the walltime was 3 hours.
+In order to perform the experiments in less than 2 hours per run\mtnote{we do
+we need to keep it below 2 hours?}, we
+% decided to send for execution a number of AthenaMP instances equal to
+used five times AthenaMP instances than the number of worker nodes. % and w
+We also reduced  the number of events
+% that are simulated
+simulated by each AthenaMP to sixteen.
+% \footnote{Note that the number of events has been chosen in such a way
+% that all the cores of a node run one event.}
+In this way, every core of each worker node was used to simulate one event,
+%. This allows us to complete
+reducing the running time of each AthenaMP to ~$20$ minutes. % only.
 
-Figure \ref{fig:weakScal2a} depicts the average pilot duration, the average execution time of five sequential instances of AthenaMP and the corresponding overhead.  
-We can observe that the difference between the two durations is more marked than in the previous experiments. Despite this we can notice that the growth of the overhead is still less than linear.
+We tested
+% different
+pilots
+% sizes, i.e. :
+with 256, 512, 1024 and 2048 worker nodes and
+%. For all of them, the walltime was
+3\mtnote{2?} hours walltime. Figure~\ref{fig:weakScal2a} depicts the average
+pilot duration, the average execution time of five sequential
+instances\mtnote{generations?} of AthenaMP, and the corresponding overhead. We
+% can
+observe that the difference between the two durations is more marked than in the
+previous experiments \mtnote{mostly due to the increased overhead?}. Despite
+this, we
+% can
+notice that the growth of the overhead is still less than linear.
 
@@
         \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak2.pdf}
-    \caption{Weak scalability with multiple generations: average pilot duration, average duration of five sequential AthenaMP executions and pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes).
-Note that the overhead refers to the axis on the right and it is expressed in tenths of a second.}
+    \caption{Weak scalability with multiple generations: average pilot
+    duration, average duration of five sequential AthenaMP executions, and
+    pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes).
+    Note that the overhead refers to the axis on the right and it is expressed
+    in tenths of a second.}
 \label{fig:weakScal2a}
 \end{figure}
+
 %\begin{figure}[!htb]
@@
 %\end{figure}
+
 \subsubsection{Strong scalability}
-The last experiments deals with strong scalability. Therefore, we run the same amount of tasks for different pilot sizes. 
-We used a number of AthenaMP instances equal to 2048 and tested pilots with 256, 512, 1024 and 2048 nodes.
-Thus, the number of sequential AthenaMP instances is equal to eight for the smallest pilot and is null for the largest.
-Each AthenaMP instance simulates sixteen events as the previous experiment.
 
-Figure \ref{fig:strongScala} depicts the average pilot duration and  the average execution time of possibly sequential AthenaMP instances.
-We can notice that the difference between the pilot duration and the AthenaMP execution times is almost constant for all the pilot sizes although the overall duration of the pilot decreases linearly with the pilot size.
+The last experiments % deals with
+study strong scalability. % Therefore, we
+by running the same amount of tasks for different pilot sizes. We used
+% a number of AthenaMP instances equal to
+2048 AthenaMP instances and % tested
+pilots with 256, 512, 1024 and 2048 nodes. Thus, the number of sequential
+AthenaMP instances is equal to 8 for the smallest pilot and is null for the
+largest\mtnote{Not sure I understand this sentence}. Each AthenaMP instance
+simulates sixteen events as in the previous experiment.
+
+Figure \ref{fig:strongScala} % depicts
+shows the average pilot duration and the average execution time of possibly
+sequential AthenaMP instances.  We % can
+notice that the difference between the pilot duration and the AthenaMP execution
+times is almost constant for all the pilot sizes, although the overall duration
+of the pilot decreases linearly with the pilot size.
 
@@
         \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/strong.pdf}
-    \caption{Strong scalability:  average pilot duration, average duration of sequential AthenaMP executions and pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes).
-Note that the overhead refers to the axis on the right and it is expressed in tenths of a second.}
+    \caption{Strong scalability:  average pilot duration, average duration of
+    sequential AthenaMP executions, and pilot's overhead for different pilot
+    sizes (256, 512, 1024 and 2048 nodes). Note that the overhead refers to the
+    axis on the right and it is expressed in tenths of a second.}
 \label{fig:strongScala}

### 2017-03-29 PandaOverview.tex
Pass on NGE experiments and scattered space savings

@@
 abstraction that enables multi-level scheduling by decoupling resource
-acquisition from tasks scheduling~\cite{turilli2015comprehensive}. Once
-implemented, a pilot is scheduled on a site (e.g., by submitting a job or
-requesting a virtual machine) and, once active, tasks are scheduled to the
-pilot, not to the site's scheduler.
+acquisition from tasks scheduling~\cite{turilli2015comprehensive}. When
+implemented, a pilot is scheduled on a site and, once active, tasks are
+scheduled to the pilot, not to the site's scheduler.
 
-Pilot-enabled WMS are particularly relevant for LHC experiments, where millions
-of tasks are executed across multiple sites every month, analyzing and producing
-petabytes of data. The design of PanDA WMS started in 2005 and its
-implementation went into production for the LHC Run 1 on 2009. PanDA was then
-extended with new subsystems to be deployed on Run 2, on 2015. We summarize
-PanDA's design, architecture, and execution process, focussing on those
-subsystems and components that are most relevant to the the integration of PanDA
-with supercomputers.
+Pilot-enabled WMS enable high throughput of tasks execution while supporting
+interoperability across multiple sites. This is particularly relevant for LHC
+experiments, where millions of tasks are executed across multiple sites every
+month, analyzing and producing petabytes of data. The design of PanDA WMS
+started in 2005 and its implementation went into production for the LHC Run 1 on
+2009. PanDA was then extended with new subsystems to be deployed on Run 2, on
+2015.
+
+% We summarize PanDA's design, architecture, and execution process, focussing on
+% those subsystems and components that are most relevant to the the integration
+% of PanDA with supercomputers.
 

### 2017-03-29 Experiments.tex
Modfied figures

@@
     single AthenaMP execution, and pilot's overhead for different pilot sizes
-    (250, 500, 1000 and 2000 worker nodes). Note: the overhead refers to
-    the axis on the right and it is expressed in tenths of a second.}
+    (250, 500, 1000 and 2000 worker nodes). }
 \label{fig:weakScal1a}
@@
     duration, average duration of five sequential AthenaMP executions, and
-    pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes).
-    Note that the overhead refers to the axis on the right and it is expressed
-    in tenths of a second.}
+    pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes).}
 \label{fig:weakScal2a}
@@
     sequential AthenaMP executions, and pilot's overhead for different pilot
-    sizes (256, 512, 1024 and 2048 nodes). Note that the overhead refers to the
-    axis on the right and it is expressed in tenths of a second.}
+    sizes (256, 512, 1024 and 2048 nodes).}
 \label{fig:strongScala}

### 2017-03-30 Experiments.tex
Added Answers to Matteo's questions and  refinement of Experiments

@@
 simulating a pre-determined number of events in the ATLAS detector.\mtnote{are
-they 100 events?}
+they 100 events?}\aanote{`Not always.'}
 % taken from ATLAS workload.
@@
 % have been assigned with 100 events whose simulation requires
-simulates 100 events, requiring ~$70$ minutes.
+simulates 100 events, requiring $\sim 70$ minutes on average.
 
@@
 I am assuming the reader understands the term by now. Please change if you
-disagree.}
+disagree.}\aanote{I agree}
 % instances in place of those that end.
@@
 
-In order to perform the experiments in less than 2 hours per run\mtnote{we do
-we need to keep it below 2 hours?}, we
+\mtnote{we do
+we need to keep it below 2 hours?}\aanote{You are right. It was misleadind. I Changed. Check if you like it.}
 % decided to send for execution a number of AthenaMP instances equal to
-used five times AthenaMP instances than the number of worker nodes. % and w
-We also reduced  the number of events
-% that are simulated
-simulated by each AthenaMP to sixteen.
+We performed the experiment by running five AthenaMP instances per node. % and w
+Since we are focusing on the overhead generated by the scheduling and bootstrap of AthenaMP instances, we also reduced  the number of events simulated by each AthenaMP to sixteen in such a way that the running time of each AthenaMP is, on average, $\sim 20$ minutes.
+This choice has been made to avoid the wasting of Titan's core hours but it does not affect the aim of the experiment.  
 % \footnote{Note that the number of events has been chosen in such a way
 % that all the cores of a node run one event.}
-In this way, every core of each worker node was used to simulate one event,
+%In this way, every core of each worker node was used to simulate one event,
 %. This allows us to complete
-reducing the running time of each AthenaMP to ~$20$ minutes. % only.
+%reducing the running time of each AthenaMP to ~$20$ minutes. % only.
 
@@
 pilot duration, the average execution time of five sequential
-instances\mtnote{generations?} of AthenaMP, and the corresponding overhead. We
+generations\mtnote{generations?}\aanote{Changed} of AthenaMP, and the corresponding overhead. We
 % can
 observe that the difference between the two durations is more marked than in the
-previous experiments \mtnote{mostly due to the increased overhead?}. Despite
-this, we
-% can
-notice that the growth of the overhead is still less than linear.
+previous experiments \mtnote{mostly due to the increased overhead?}\aanote{Well, I guess that's plain.}.
+Despite this, we can notice that the growth of the overhead is consistent with the increment of the number of tasks per node for pilots with 256, 512 and 1024 worker nodes, and much less than linear for the pilot with 2048 worker nodes \aanote{Although that histogram is not trustworty because it is generate with only two runs.}
 
@@
 The last experiments % deals with
-study strong scalability. % Therefore, we
-by running the same amount of tasks for different pilot sizes. We used
+study strong scalability by running the same amount of tasks for different pilot sizes. We used
 % a number of AthenaMP instances equal to
 2048 AthenaMP instances and % tested
-pilots with 256, 512, 1024 and 2048 nodes. Thus, the number of sequential
-AthenaMP instances is equal to 8 for the smallest pilot and is null for the
-largest\mtnote{Not sure I understand this sentence}. Each AthenaMP instance
+pilots with 256, 512, 1024 and 2048 nodes.
+Thus, the number of AthenaMP generations is equal to eight times the size of the smallest pilot and corresponds to the size of the largest pilot.
+As a consequence, the number of consecutive generations of AthenaMP decreases with the pilot size by generating different dynamics within the pilots.
+This experiment aim to show that the pilot overhead is not affected by the concurrency level within the pilot but it depends only on the number of tasks.
+\mtnote{Not sure I understand this sentence}\aanote{My fault.
+It should have been seven and null or 8 and 1.
+It depends on how we consider the word sequential.
+anyway, to avoid confusion I changed the sentence.}.
+Each AthenaMP instance
 simulates sixteen events as in the previous experiment.

### 2017-03-30 overall.tex
Pass on NGE experiments and scattered space savings
Address comments on figure 5

@@
 \begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/titan_backfill_avail.pdf}
+    \includegraphics[clip,width=\columnwidth]{figures/titan_backfill_avail.png}
     %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
@@
 % -----------------------------------------------------------------------------
-\subsection{Efficiency and Scalability of Detector Simulation on Titan}
+\subsection{Efficiency, Scalability, and Reliability of Detector Simulation on Titan}
 \label{ssec:panda_titan}
@@
 
-Initially, all the shared libraries of AthenaMP and the python scripts for the
-configuration stage were stored on the Spider 2 Lustre file system. However, the
-I/O patterns of the initialization and configuration stages degraded the
-performance of the filesystem (Figure~\ref{}\mtnote{We may want to aggregate
-some of the diagrams about lustre's experiments here}). Since Spider 2 is a
-center-wide file system, this resulted in  performance degradation for all OLCF
-resources and users. \mtnote{I am afraid the details about the trace are too
-specific given the space constraints of a SC submission. Please feel free to
-uncomment it if you disagree.}
+\mtnote{TODO: specify that this is work in progress (see Sergey's comment) and
+see how to move it to the data performance subsection}Initially, all the shared
+libraries of AthenaMP and the python scripts for the configuration stage were
+stored on the Spider 2 Lustre file system. However, the I/O patterns of the
+initialization and configuration stages degraded the performance of the
+filesystem (Figure~\ref{}\mtnote{We may want to aggregate some of the diagrams
+about lustre's experiments here}). Since Spider 2 is a center-wide file system,
+this resulted in  performance degradation for all OLCF resources and users.
+\mtnote{I am afraid the details about the trace are too specific given the space
+constraints of a SC submission. Please feel free to uncomment it if you
+disagree.}
 

### 2017-03-31 PandaOverview.tex
Record comments and changes from Danila

@@
 \label{ssec:panda_arch}
+
+\mtnote{DDM: just enacts decisions about registering files into datasets. These
+decisions are made by productions. DDM DOES OT DECIDE which file should be in
+which dataset.}
+\mtnote{No communication between APF and PanDA Server (deleted arrow, adjust
+numbers and related text)}
+\mtnote{No more DDM Service on Edge on grid sites. DDM directly interacts with
+SE (deleted arrows, moved number 8. Update labels and related text)}
 

### 2017-04-01 PandaOverview.tex
Address Danila's and Ruslan's comments

--- 2017-03-29 PandaOverview.tex
@@
 
+\mtnote{From Danila: DDM enacts decisions about registering files into datasets.
+These decisions are made by productions. DDM DOES OT DECIDE which file should be
+in which dataset.}\mtnote{Done.}
+\mtnote{No communication between APF and PanDA Server (deleted arrow, adjust
+numbers and related text)}\mtnote{Done.}
+\mtnote{No more DDM Service on Edge on grid sites. DDM directly interacts with
+SE (deleted arrows, moved number 8. Update labels and related text)}\mtnote{Done.}
+
 The implementation of PanDA consists of several interconnected subsystems, most
@@
     the execution process based on JEDI: From the submission of tasks (1) to the
-    retrieval of their output (20). The monitoring subsystem, the architectural
+    retrieval of their output (19). The monitoring subsystem, the architectural
     details of PanDA Pilot and the communication among subsystem's components
@@
 required users or applications to perform a static conversion between tasks and
-jobs: tasks were described as a set of jobs with a fixed number of events and
-then submitted to the PanDA Server.
+jobs: tasks were described as a set of jobs
+% with a fixed number of events
+and then submitted to the PanDA Server.
 
@@
 resources available for each site. Jobs submitted by JEDI are already bound to
-sites so no further brokerage is needed. Once jobs are bound to sites, the
-Brokerage module communicates to the Data Service module what data sets need to
-be made available on what site (Fig.~\ref{fig:architecture}:6). The Data Service
-communicates these requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7)
-that, when needed, takes care of aggregating files into datasets and containers,
+sites so no further brokerage is needed.
+
+Once jobs are bound to sites, the Brokerage module communicates to the Data
+Service module what data sets need to be made available on what site
+(Fig.~\ref{fig:architecture}:6). The Data Service communicates these
+requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7) that, when needed,
+takes care of % aggregating files into datasets and containers,
 replicating them on the required sites (Fig.~\ref{fig:architecture}:8).
 
-The AutoPyFactory subsystem communicates with the Task Buffer component of the
-PanDA Server, acquiring information about jobs that are ready for execution on
-specific (type of) sites (Fig.~\ref{fig:architecture}:9). On the basis of this
-information, AutoPyFactory defines a set of suitable PanDA Pilots and
-concurrently submits them to a Condor-G broker (Fig.~\ref{fig:architecture}:10).
-This broker submits these pilots wrapped as jobs or VMs to the required sites
-(Fig.~\ref{fig:architecture}:11).
+% AutoPyFactory subsystem communicates with the Task Buffer component of the
+% PanDA Server, acquiring information about jobs that are ready for execution on
+% specific (type of) sites (Fig.~\ref{fig:architecture}:9). On the basis of this
+% information,
+Meanwhile, AutoPyFactory defines a set of suitable PanDA Pilots and concurrently
+submits them to a Condor-G broker (Fig.~\ref{fig:architecture}:9). This broker
+submits these pilots wrapped as jobs or VMs to the required sites
+(Fig.~\ref{fig:architecture}:10).
 
@@
 
-When a PanDA Pilot becomes available on a site, it
-% bootstrap, download their code, collect information about the execution
-% environment and
-calls the Job Dispatcher module of the PanDA Server, requesting for a job to
-execute (Fig.~\ref{fig:architecture}:12). The Job Dispatcher communicates with
-the Task Buffer, requesting for a job that is bound to the site of that pilot
-and ready to be executed. Task Buffer checks the global queue (i.e., the PanDA
-database) and if such a job is available, it passes the job handler to the Job
-Dispatcher. The Job Dispatcher dispatches the job to the PanDA Pilot
-(Fig.~\ref{fig:architecture}:13).
+When a PanDA Pilot becomes available on a site, it calls the Job Dispatcher
+module of the PanDA Server, requesting for a job to execute
+(Fig.~\ref{fig:architecture}:11). The Job Dispatcher interrogates the Task
+Buffer module for a job that is bound to the site of that pilot and ready to be
+executed. Task Buffer checks the global queue (i.e., the PanDA database) and,
+upon availability, returns a suitable job the Job Dispatcher. The Job Dispatcher
+dispatches that job to the PanDA Pilot (Fig.~\ref{fig:architecture}:12).
 
@@
 Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
-subprocess for the execution of the job's payload. The job is setup, input data
-are transferred from the designated staging in location
-(Fig.~\ref{fig:architecture}:14), and the job's payload is executed
-(Fig.~\ref{fig:architecture}:15). Once completed, output is transferred to the
-staging out location (Fig.~\ref{fig:architecture}:16).
+subprocess for the execution of the job's payload. The job is set up, input data
+are transferred from the designated staging-in location
+(Fig.~\ref{fig:architecture}:13), and the job's payload is executed
+(Fig.~\ref{fig:architecture}:14). Once completed, output is transferred to the
+staging-out location (Fig.~\ref{fig:architecture}:15).
 
@@
 The Data Service module of the PanDA Server tracks and collects the output
-generated by each job (Fig.~\ref{fig:architecture}:17), updating jobs'
-attributes via the Task Buffer module (Fig.~\ref{fig:architecture}:18). When the
+generated by each job (Fig.~\ref{fig:architecture}:16), updating jobs'
+attributes via the Task Buffer module (Fig.~\ref{fig:architecture}:17). When the
 output of all the jobs of a task are retrieved, it is made available to the user
-via PanDA Server. When a task is submitted to JEDI, task is marked as done
-(Fig.~\ref{fig:architecture}:19) and the result of its execution is made
-available to the user by JEDI (Fig.~\ref{fig:architecture}:20).
+via PanDA Server. When a task is submitted to JEDI, task is instead marked as
+done (Fig.~\ref{fig:architecture}:18) and the result of its execution is made
+available to the user by JEDI (Fig.~\ref{fig:architecture}:19).
 

### 2017-04-01 PandaEvolution.tex
Ongoing: Addressing Alexei's comments

@@
 Worldwide LHC Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more
-than 100,000 cores at over 100 Grid sites, with a peak performance of 0.3
-petaFLOPS. This capacity will be sufficient for the planned analysis and data
-processing, but it will be insufficient for the Monte Carlo production workflow
-and any extra activity. To alleviate these challenges, ATLAS is engaged in a
-program to expand the current computing model to include additional resources
-such as the opportunistic use of supercomputers as well as commercial and
-academic clouds.
+than
+% 100,000
+300,000 cores at over 100 Grid sites\mtnote{Addresses Alexei's comment.
+Increased Grid capacity to 300K cores. Is the number of sites still 100 or
+should I increase that too?}, with a peak performance of 0.3 petaFLOPS. This
+capacity will be sufficient for the planned analysis and data processing, but it
+will be insufficient for the Monte Carlo production workflow and any extra
+activity. To alleviate these challenges, ATLAS is engaged in a program to expand
+the current computing model to include additional resources such as the
+opportunistic use of supercomputers as well as commercial and academic clouds.
 
@@
 four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
-HERWIG~\cite{corcella2001herwig} and many others. The detector simulator is
-called Geant4~\cite{agostinelli2003geant4} and simulates the interaction of
-these particles with the sensitive material of the ATLAS detector. Each
-interaction creates a so-called hit and all hits are collected and passed on for
-digitalization where hits are further process to mimic the readout of the
+HERWIG~\cite{corcella2001herwig} and many others. \mtnote{Addresses Alexei's comment. Is the following sentence correct?} The detector simulator is
+called Geant4~\cite{agostinelli2003geant4} and simulates both the ATLAS detector and it's interaction physical processes.
+% the interaction of these particles with the sensitive material of the ATLAS
+% detector.
+Each interaction creates a so-called hit and all hits are collected and passed
+on for digitalization where hits are further process to mimic the readout of the
 detector. Finally, reconstruction operates local pattern recognition, creating

### 2017-04-01 nge.tex
Pass on NGE experiments and scattered space savings
minor fixes and response to comment in 5.0.

@@
-\subsection{Rationale} \aanote{I think that the names of the states should be
-removed. They are meaningless without the figures. I left them just in case we
-decide to reintroduce the figures.}\mtnote{Done.}
+% \subsection{Rationale}
+\aanote{I think that the names of the states should be removed. They are
+meaningless without the figures. I left them just in case we decide to
+reintroduce the figures.}\mtnote{Done.}
 
@@
 to Titan's network. Do you have a suggestion for an alternative to ``wide
-area''?}
+area''?} \jhanote{how is the following: "... deployed on the DTNs as Titan's worker
+nodes lack connectivity to the wide area network"}.
 
@@
 efficiency by increasing throughput and by enabling better use of backfill
-utilization. Further, pilots makes easier to support heterogeneous workloads.
+utilization. Further, pilots makes it easier to support heterogeneous
+workloads.
 
@@
 static coupling between MPI scripts submitted to the PBS batch system and
-detector simulations. This makes impossible to schedule multiple generations of
-workload on the same PBS job: Once a number of detector simulations are packaged
-into a PBS job and this job is queued on Titan, no further simulations can be
-added to that job. New simulations have to be packaged into a new PBS job that
-needs to be submitted to Titan on the base of the backfill availability of that
-moment.
+detector simulations. This makes the scheduling of multiple generations of
+workload on the same PBS job impossible: Once a statically defined number of
+detector simulations are packaged into a PBS job and this job is queued on
+Titan, no further simulations can be added to that job. New simulations have
+to be packaged into a new PBS job that needs to be submitted to Titan based 
+upon backfill availability.
 
-The support of  multiple workload generations would enable a more efficient
+The support of  multiple generations of workload would enable a more efficient
 use of the backfill availability of walltime. Currently, when a set of
 simulations ends, the PBS job also ends, independent of whether more wall-time
-would still be available. With a pilot, more simulations could be executed so
-to utilize all the available wall-time, while avoiding further job packaging
-and submission overheads.
+would still be available. With a pilot, additional simulations could be
+executed  to utilize all the available wall-time, while avoiding further job
+packaging and submission overheads.
 
@@
 
-Multiple generations would also help with relaxing two assumptions of the
-current execution model: knowing the number of simulations before submitting the
-MPI script, and having a fixed amount of events per simulation (100 at the
-moment). Pilots would enable the scheduling of simulations independently from
-whether they were available at the moment of submitting the pilot. Further,
-simulations with a varying number of events could be scheduled on a pilot,
-depending on the amount of remaining walltime and the distribution of execution
-time per event, as shown in \S\ref{ssec:panda_titan},
-Figure~\ref{fig:100event-distrib}. This capabilities would be particularly
-useful to increase the PanDA Broker efficiency for availabilities with a wide
-delta between the number of cores and walltime.
+\jhanote{not sure MPI script is optimal description or ambigious.} Multiple
+generations would also relax two assumptions of the current execution model:
+knowing the number of simulations before submitting the MPI script, and having
+a fixed number of events per simulation (100 at the moment). Pilots would
+enable the scheduling of simulations independently from whether they were
+available at the moment of submitting the pilot. Further, simulations with a
+varying number of events could be scheduled on a pilot, depending on the
+amount of remaining walltime and the distribution of execution time per event,
+as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event- distrib}.
+These capabilities would increase the efficiency of the PanDA Broker when
+there is a large difference between the number of cores and walltime.
 
@@
 % -----------------------------------------------------------------------------
-\subsection{Architecture}
+\subsection{Implementation}
 \label{sec:arch}

### 2017-04-01 overall.tex
Fix caption figure 5
Test github with Jack
Ongoing: address part of Sergey's comments, eliminated two figures

@@
 
-% \begin{figure}[htp]
-%     \subfloat[]{
-%         \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
-%         \label{fig:core-hours-utilization}
-%     }
-%
-%     \subfloat[]{
-%         \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
-%         \label{fig:backfill-utilization}
-%     }
-% \caption{(a).(b).}
-% \end{figure}
-
-% \begin{figure}[htp]
-% \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
-% \caption{Caption.}
-% \label{fig:core-hours-utilization}
-% \end{figure}
-
 \begin{figure}[htp]
@@
 comparable difference in total backfill availability.
+
+\mtnote{Elaborate further and put it at the end, after giving several of the
+measurements we would need. ``Currently, this measure of efficiency is being
+refined on the base of the actual minimum amount of walltime that PanDA Broker
+can request when submitting job requests. This is not a fixed value as it
+depends on variable initialization and configuration overheads, simulation time,
+and number of nodes obtained. Part of the undergoing effort is to enable the
+measurements of  the distribution of these parameters to derive an optimization
+function.''}
 
@@
 Not all events take the same time to be simulated:
-Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time of
-1 event from $\approx$2 to $\approx$40 minutes, with a mean of $\approx$14
-minutes. Considering that each worker node process up to 16 events concurrently,
-100 events takes around 2 hours to process.
+% Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time
+% of
+1 event takes from $\approx$2 to $\approx$40 minutes to be simulated, with a
+mean of $\approx$14 minutes. Considering that each worker node process up to 16
+events concurrently, 100 events takes around 2 hours to process.
 
-\begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
-    \caption{Distribution of the time taken by a Geant4 detector simulation to
-    simulate one event. 3000 Events; 30 Titan worker nodes; 16 simulations per
-    node; 100 events per node. Average time per event $\approx$14 min. Broad
-    distribution from $\approx$2 to $\approx$40 minutes.}
-\label{fig:100event-distrib}
-\end{figure}
+% \begin{figure}[htp]
+%     \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
+%     \caption{Distribution of the time taken by a Geant4 detector simulation to
+%     simulate one event. 3000 Events; 30 Titan worker nodes; 16 simulations per
+%     node; 100 events per node. Average time per event $\approx$14 min. Broad
+%     distribution from $\approx$2 to $\approx$40 minutes.}
+% \label{fig:100event-distrib}
+% \end{figure}
 
@@
     work nodes during the experiment window time. Availability was measured
-    62175 times: mean number of work nodes available 680; mean walltime
-    available 680 minutes.}
+    62175 times: mean number of work nodes available 680 (yellow line); mean
+    walltime available 122 minutes (red line).}
 \label{fig:backfill-distrib}
@@
 concurrent simulations computing 50 events is 10.8 minutes, while for 16
-simulations is 14.25 minutes (consistent with
-Figure~\ref{fig:100event-distrib}). Despite an inefficiency of almost 30\%,
-Titan's allocation policy based on number of worker nodes used instead of number
-of cores does not justify the use of 1/2 of the cores available.
+simulations is 14.25 minutes (consistent with the measured distribution of the
+duration of event simulation). Despite an inefficiency of almost 30\%, Titan's
+allocation policy based on number of worker nodes used instead of number of
+cores does not justify the use of 1/2 of the cores available.
 
@@
 of executing the same detector simulation on both Titan or Grid sites.
-Figure~\ref{fig:comparison-titan-grid} compares the efficiencies of Titan to the
-BNL and SIGNET Grid sites, normalized for 8 cores. Effective performance
-per-core at Titan is $\approx$0.57 event per minute, roughly 1/2 of BNL and 1/3
-of SIGNET performances.
+% Figure~\ref{fig:comparison-titan-grid} compares the efficiencies of Titan to
+% the BNL and SIGNET Grid sites, normalized for 8 cores. Effective performance
+% per-core at Titan is $\approx$0.57 event per minute, roughly 1/2 of BNL and
+% 1/3 of SIGNET performances.
+Comparisons of the efficiencies of Titan to  the BNL and SIGNET Grid sites,
+normalized for 8 cores, show that the effective performance per-core at Titan is
+$\approx$0.57 event per minute, roughly 1/2 of BNL and  1/3 of SIGNET
+performances.
 
-\begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/titan_grid_event_rate.pdf}
-    \caption{Comparison of the event rate for finished jobs at Titan (h1), BNL
-    (h2) and GRIDNET (h3). Titan histogram shows 1/2 of the measured values to
-    account for the core difference between Titan (16) and the Grid sites (8).
-    Performance per core: BNL 1.126; SIGNET 0.885; Titan 0.577.}
-\label{fig:comparison-titan-grid}
-\end{figure}
+% \begin{figure}[htp]
+%     \includegraphics[clip,width=\columnwidth]{figures/titan_grid_event_rate.pdf}
+%     \caption{Comparison of the event rate for finished jobs at Titan (h1), BNL
+%     (h2) and GRIDNET (h3). Titan histogram shows 1/2 of the measured values to
+%     account for the core difference between Titan (16) and the Grid sites (8).
+%     Performance per core: BNL 1.126; SIGNET 0.885; Titan 0.577.\mtnote{TO BE DELETED.}}
+% \label{fig:comparison-titan-grid}
+% \end{figure}
 

### 2017-04-01 PandaOverview.tex

@@
 
+\mtnote{Addresses Alexei's comment. I eliminated `event' as abstraction and used
+datum/data/physical process when discussing data management in the whole
+section.}
+
 PanDA's application model assumes tasks, workloads and workflows. Tasks
-represent a set of operations performed on a set of events that are stored in
-one or more input files. Tasks are decomposed into jobs, where each job
-represents the task's set of operations and a partition of the task's events.
-Since 2005, a certain amount of parallelism has been progressively introduced
-for job execution~\cite{crooks2012multi} but, so far, no MPI jobs have been
-considered for production. Jobs are supposed to be relatively self-contained,
-capable of setting up their own execution environment or having a minimal set of
-common dependences.
+represent a set of operations performed on
+% a set of events that are
+data sets stored in one or more input files. Tasks are decomposed into jobs,
+where each job represents the task's set of operations and a partition of the
+task's
+% events
+data set. Since 2005, a certain amount of parallelism has been
+progressively introduced for job execution~\cite{crooks2012multi} but, so far,
+% no MPI jobs have been considered for production
+HEP applications have not required MPI\mtnote{Addresses Alexei's comment. If
+this is still too soft, we can write: ``but HEP applications do not require
+MPI.''}. Jobs are supposed to be relatively self-contained, capable of setting
+up their own execution environment or having a minimal set of common
+dependences.
 
@@
 
-Currently, PanDA's execution model is based on five main abstractions: task,
-job, queue, pilot, and event. Both tasks and jobs are assumed to have attributes
-and states and to be queued into a global queue for execution. Prioritization
-and binding of jobs are assumed to depend on the attributes of each job. Pilot
-is used to indicate the abstraction of resource capabilities. Each job is
-thought to be bound to one pilot and executed on the site where the pilot has
-been instantiated.
+% Currently, PanDA's execution model is based on five main abstractions: task,
+% job, queue, pilot, and event. Both tasks and jobs are assumed to have
+% attributes and states and to be queued into a global queue for execution.
+% Prioritization and binding of jobs are assumed to depend on the attributes of
+% each job. Pilot is used to indicate the abstraction of resource capabilities.
+% Each job is thought to be bound to one pilot and executed on the site where
+% the pilot has been instantiated.
 
-In PanDA's data model, each event refers to the recorded or simulated
-measurement of a physical event. One or more events can be packaged into files
-or other data containers. As with jobs, data have both attributes and states,
-and some of the attributes are shared between events and jobs. Raw,
-reconstruction, and simulation data are assumed to be distributed across
-multiple storage facilities and managed by the ATLAS Distributed Data Management
-(DDM)~\cite{garonne2012atlas}. When necessary, datasets required by each job are
-assumed to be replicated over the network, both for input and output data.
+Currently, PanDA's execution model is based on four main abstractions: task,
+job, queue, and pilot. Both tasks and jobs are assumed to have attributes and
+states and to be queued into a global queue for execution. Prioritization and
+binding of jobs are assumed to depend on the attributes of each job. Pilot is
+used to indicate the abstraction of resource capabilities. Each job is thought
+to be bound to one pilot and executed on the site where the pilot has been
+instantiated.
+
+In PanDA's data model, each
+% event
+datum refers to the recorded or simulated measurement of a physical
+% event
+process.
+% One or more events
+Data can be packaged into files or other containers. As with jobs, data
+have both attributes and states, and some of the attributes are shared between
+events and jobs. Raw, reconstruction, and simulation data are assumed to be
+distributed across multiple storage facilities and managed by the ATLAS
+Distributed Data Management (DDM)~\cite{garonne2012atlas}. When necessary,
+datasets required by each job are assumed to be replicated over the network,
+both for input and output data.
 
@@
 implemented by one or more modules. Databases are used to store eventful
-entities like tasks, jobs and events, and to store information about sites,
+entities and entities' description like tasks, jobs and % events
+input/output data, and to store information about sites,
 resources, logs, and accounting.

### 2017-04-01 overall.tex

--- 2017-03-30 overall.tex
@@
     account for the core difference between Titan (16) and the Grid sites (8).
-    Performance per core: BNL 1.126; SIGNET 0.885; Titan 0.577.}
+    Performance per core: BNL 1.126; SIGNET 0.885; Titan 0.577.\mtnote{TO BE DELATED.}}
 \label{fig:comparison-titan-grid}

### 2017-04-01 nge.tex

@@
 also be MPI-independent, as MPI is used for coordination among multiple worker
-nodes, not by the payload.
+nodes, not by the payload. \jhanote{candidate for compression/contraction into
+earlier paragraph(s)}
 
@@
 submitted to Titan and executed on its worker nodes. The communication between
-PanDA Broker and PanDA Server is abstracted away as not immediately useful to
+PanDA Broker and PanDA Server is abstracted away as it is not immediately useful to
 evaluate the performance and capabilities of a pilot on Titan.
@@
 The UnitManager and the Agent communicate via a database that is instantiated on
-one of Titan's DTN so to be reachable by both modules. A similar approach would
+one of Titan's DTN so as to be reachable by both modules. A similar approach would
 be used to enable PanDA Broker to schedule pilots instead of MPI

### 2017-04-01 Experiments.tex
Minor fixes experiments

@@
 submitting NGE's pilots % on
-to TITAN's batch queue. Because of TITAN's % batch queue
+to Titan's batch queue. Because of Titan's % batch queue
 policies, the turnaround time of each run of our experiments is dominated by
@@
 to use all the 16 cores % present on
-of TITAN's worker nodes.
+of Titan's worker nodes.
 

### 2017-04-01 PandaEvolution.tex

@@
 four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
-HERWIG~\cite{corcella2001herwig} and many others. \mtnote{Addresses Alexei's comment. Is the following sentence correct?} The detector simulator is
-called Geant4~\cite{agostinelli2003geant4} and simulates both the ATLAS detector and it's interaction physical processes.
+HERWIG~\cite{corcella2001herwig} and many others. \mtnote{Addresses Alexei's
+comment. Is the following sentence correct?} The detector simulator is called
+Geant4~\cite{agostinelli2003geant4} and simulates the ATLAS detector and the
+interaction between that detector and particles.
 % the interaction of these particles with the sensitive material of the ATLAS
@@
 Each interaction creates a so-called hit and all hits are collected and passed
-on for digitalization where hits are further process to mimic the readout of the
-detector. Finally, reconstruction operates local pattern recognition, creating
-high-level objects like particles and jets.
+on for digitalization, where hits are further process to mimic the readout of
+the detector. Finally, reconstruction operates local pattern recognition,
+creating high-level objects like particles and jets.
 
@@
 
-\mtnote{Please feel free to add details about backfill as needed.}
-\mtnote{Please note: I know the following departs for the current name given to
-the PanDA subsystem installed on Titan. I think there is a good reason to call
-it a broker instead of a pilot, and I think I explained it in the previous
-paragraphs. Please  take this just as a suggestion, something I would like to
-discuss in our meeting.} \sergeynote{It's not a pilot in conventional sense. I
-call it an agent. It just happened that we were able to use full PanDA pilot's
-code base to serve our purposes on Titan. that's why, by inertia, we still call
-it a pilot.}\mtnote{Thank you. We would have a problem calling it `agent' as we
-use that term to name the pilot of NGE. Would PanDA Broker work?}\sergeynote{OK
-broker it is for this paper.}
+% \mtnote{Please feel free to add details about backfill as needed.}
+% \mtnote{Please note: I know the following departs for the current name given
+% to the PanDA subsystem installed on Titan. I think there is a good reason to
+% call it a broker instead of a pilot, and I think I explained it in the
+% previous paragraphs. Please  take this just as a suggestion, something I would
+% like to discuss in our meeting.} \sergeynote{It's not a pilot in conventional
+% sense. I call it an agent. It just happened that we were able to use full
+% PanDA pilot's code base to serve our purposes on Titan. that's why, by
+% inertia, we still call it a pilot.}\mtnote{Thank you. We would have a problem
+% calling it `agent' as we use that term to name the pilot of NGE. Would PanDA
+% Broker work?}\sergeynote{OK broker it is for this paper.}
 

### 2017-04-01 PandaEvolution.tex
refining subsection titles.

@@
 
-Compared to pilots, backfill has the disadvantage of limiting the amount of work
-nodes that can be requested. Pilots are normal jobs: they can request as many
-worker nodes for as much time as a queue can offer. On the contrary, jobs sized
-on the basis of backfill information availability \jhanote{should be resource
-availability or backfill queue information?}\mtnote{AFAIH, there is no dedicated
-queue to backfill} depend on the number of worker nodes that cannot be given to
-any other job in the Titan's queue.
-% \sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.
+Compared to pilots, backfill has the disadvantage of limiting the amount of
+work nodes that can be requested. Pilots are normal jobs: they can request as
+many worker nodes for as much time as a queue can offer. On the contrary, jobs
+sized on the basis of backfill information availability depend on the number
+of worker nodes that cannot be given to any other job in the Titan's queue. %
+\sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.
 

### 2017-04-01 PandaEvolution.tex
section 3. fixing definition of backfill. more to come in 4.

@@
 
+% Titan’s backfill functionality offers the opportunity to avoid the overhead of
+% queue wait times without using pilot abstraction. Backfill availability is the
+% number of node-hours (or core-hours) that cannot be used for a certain amount
+% of time by any of the jobs already queued on Titan: All queued jobs are either
+% too large or their walltime is too long. At any point in time, Titan’s Moab
+% scheduler can be queried for backfill availability. Based upon the result of
+% this query, a job can be shaped to request no more than the backfill
+% availability. As such, when submitted this job is usually scheduled
+% immediately, spending almost no time in the queue.
+
 Titan’s backfill functionality offers the opportunity to avoid the overhead of
 queue wait times without using pilot abstraction. Backfill availability is the
-number of worker nodes that cannot be used for a certain amount of time by any
-of the jobs already queued on Titan: All queued jobs are either too large or
-their walltime is too long. At any point in time, Titan’s Moab scheduler can be
-queried for backfill availability. Based upon the result of this query, a job
-can be shaped to request no more than the backfill availability. As such, when
-submitted this job is usually scheduled immediately, spending almost no time in
-the queue.
+number of node-hours (or core-hours) that  is not utilized by regular jobs
+submitted to the queues on Titan. Titan’s backfill functionality offers the
+opportunity to avoid the overhead of queue wait times without using pilot
+abstraction.  It is important to note that backfill availability is  computed
+retrospectively and is determined by Titan's operation team. Backfill
+availability represents the theoretical maximum of node-hours that could be
+utilized by ATLAS (or any other project) that submit under backfill mode.
 
@@
 
-Usually, backfill availability is a fraction of the total capacity of the queue
-but the size of Titan mitigates this limitation. Every year, about 10\% of
-Titan's capacity remains unused, corresponding to an average of 30,000 unused
-cores. This equals approximately 270M core hours per year, roughly 30\% of the
-overall capacity of WLCG.
+Backfill availability is typically a small fraction of the total capacity  of
+a resource. Notwithstanding,  given the size of Titan this translates into a
+substantial capacity. Every year, about 10\% of Titan's capacity remains
+unused, corresponding to an average of 30,000 unused cores. This equals
+approximately 270M core hours per year, roughly 30\% of the overall capacity
+of WLCG.
 

### 2017-04-01 overall.tex
Shorter figure
refining subsection titles.
section 3. fixing definition of backfill. more to come in 4.

--- 2017-04-01 overall.tex
@@
-Currently, ATLAS deploys 20 instances of PanDA Broker on 4 Titan's DTNs, 5
+Currently, ATLAS deploys 20 instances of the PanDA Broker on 4 Titan's DTNs, 5
 instances per DTN. Each broker submits and manages the execution of 15 to 300
@@
 We evaluate the efficiency, scalability and reliability of the PanDA Brokers
-along two dimensions: (i) the amount of backfill availability utilized by the
+along two dimensions: (i) backfill utilization by the
 PanDA Brokers; and (ii) the amount of runtime spent performing detector
@@
 % -----------------------------------------------------------------------------
-\subsection{Utilization and Efficiency of PanDA Broker on Titan}
+\subsection{Characterizing the PanDA Broker on Titan}
 \label{ssec:panda_titan}
@@
 % -----------------------------------------------------------------------------
-\subsection{Efficiency, Scalability, and Reliability of Detector Simulation on Titan}
+\subsection{Characterizing the Detector Simulation on Titan}
 \label{ssec:panda_titan}
@@
 % -----------------------------------------------------------------------------
-\subsection{PanDA I/O Impact at OLCF}
+\subsection{Characterizing PanDA I/O}
 

### 2017-04-01 overall.tex
minor fixes in Section 4.

@@
 during the experiment time window.  Efficiency 
-(Figure~\ref{fig:backfill-utilization}, red labels) is defined as the fraction
-of Titan’s cores available via backfill that was utilized by the PanDA Brokers.
-The brokers reached 18\% average efficiency, with a minimum 8.9\% efficiency on
-April 2016 and a maximum 33.5\% efficiency on February 2017. The number of total
-backfill cores available was 38.1M in April 2016, and 33.1M in February 2017.
-This shows that the measured difference in efficiency did not depend on a
-comparable difference in total backfill availability.
+(Figure~\ref{fig:backfill-utilization}, red labels) is defined as the fraction of core-hours utilized by
+the PanDA Brokers (for ATLAS jobs) to the total of Titan’s backfill
+availability. The brokers reached 18\% average efficiency, with a minimum
+8.9\% efficiency on April 2016 and a maximum 33.5\% efficiency on February
+2017. The total backfill available was 38.1M core-hours in April 2016,
+and 33.1M core-hours in February 2017. This shows that the measured difference in
+efficiency did not depend on a comparable difference in total backfill
+availability.
 
@@
     work nodes during the experiment window time. Availability was measured
-    62175 times: mean number of work nodes available 680 (yellow line); mean
+    62175 times: mean number of work nodes available 680 (orange line); mean
     walltime available 122 minutes (red line).}

### 2017-04-01 PandaEvolution.tex
Clean up and reduce 3.0/3.1

--- 2017-04-01 PandaEvolution.tex
@@
 the current computing model to include additional resources such as the
-opportunistic use of supercomputers as well as commercial and academic clouds.
+opportunistic use of supercomputers.
+% as well as commercial and academic clouds.
 
 % -----------------------------------------------------------------------------
-\subsection{Use of Supercomputers with PanDA}
-\label{ssec:panda-supercomputers}
+% \subsection{Use of Supercomputers with PanDA}
+% \label{ssec:panda-supercomputers}
 
@@
 designed to aggregate large amount of computing resources across multiple sites.
-While each site may deploy runtime message-passing capabilities, usually these
-are not used to perform distributed computations.
+While each site may deploy MPI capabilities, usually these are not used to
+perform distributed computations.
 
-There are at least two approaches to enable PanDA WMS to execute ATLAS workloads
-or workflows on supercomputers: (i) using the subset of resources and
-capabilities shared by both supercomputers and WLCG; (ii) reconciling the
-parallel and distributed computing paradigms by means of dedicated abstractions.
-The former is a pragmatic approach that enables the execution of specific
-workloads by prototyping single-point solutions. The latter is a principled
-approach, better suited for a production-grade solution, capable of supporting
-general-purpose workloads and workflows on both supercomputers and Grid
-infrastructures. These two approaches are not mutually exclusive: developing a
-single-point solution gives the opportunity to better understand the problem
-space, supporting the creation of abstractions for the integration of computing
-paradigms.
+We developed and deployed a single-point solution to better understand the
+problem space and study abstractions for the integration of HTC and HPC
+computing paradigms. Specifically, the PanDA team developed a job broker to
+support the execution of part of the ATLAS production Monte Carlo workflow on
+Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
+Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).
 
-This section illustrates the design and architecture of a job broker prototyped
-by the PanDA team. This broker supports execution of part of the ATLAS
-production Monte Carlo workflow on Titan, a leadership-class supercomputer
-managed by the Oak Ridge Leadership Computing Facility (OLCF) at the Oak Ridge
-National Laboratory (ORNL). After an analysis of the results obtained and the
-lessons learned, the following section introduces the design and first
-experimental characterization of a next generation executor (NGE). NGE is
-designed to abstract resources and capabilities, enabling the concurrent
-execution of both parallel and distributed computing on generic high performance
-computing (HPC) machines.
+\mtnote{Move this to the introduction: ``There are at least two approaches to
+enable PanDA WMS to execute ATLAS workloads or workflows on supercomputers: (i)
+using the subset of resources and capabilities shared by both supercomputers and
+WLCG; (ii) reconciling the parallel and distributed computing paradigms by means
+of dedicated abstractions. The former is a pragmatic approach that enables the
+execution of specific workloads by prototyping single-point solutions. The
+latter is a principled approach, better suited for a production-grade solution,
+capable of supporting general-purpose workloads and workflows on both
+supercomputers and Grid infrastructures. \\ This section illustrates the design
+and architecture of a job broker prototyped by the PanDA team. This broker
+supports execution of part of the ATLAS production Monte Carlo workflow on
+Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
+Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).After an
+analysis of the results obtained and the lessons learned, the following section
+introduces the design and first experimental characterization of a next
+generation executor (NGE). NGE is designed to abstract resources and
+capabilities, enabling the concurrent execution of both parallel and distributed
+computing on generic high performance computing (HPC) machines.''}
 

### 2017-04-01 PandaEvolution.tex

@@
 
+% Titan’s backfill functionality offers the opportunity to avoid the overhead of
+% queue wait times without using pilot abstraction. Backfill availability is the
+% number of node-hours (or core-hours) that cannot be used for a certain amount
+% of time by any of the jobs already queued on Titan: All queued jobs are either
+% too large or their walltime is too long. At any point in time, Titan’s Moab
+% scheduler can be queried for backfill availability. Based upon the result of
+% this query, a job can be shaped to request no more than the backfill
+% availability. As such, when submitted this job is usually scheduled
+% immediately, spending almost no time in the queue.
+
 Titan’s backfill functionality offers the opportunity to avoid the overhead of
 queue wait times without using pilot abstraction. Backfill availability is the
-number of worker nodes that cannot be used for a certain amount of time by any
-of the jobs already queued on Titan: All queued jobs are either too large or
-their walltime is too long. At any point in time, Titan’s Moab scheduler can be
-queried for backfill availability. Based upon the result of this query, a job
-can be shaped to request no more than the backfill availability. As such, when
-submitted this job is usually scheduled immediately, spending almost no time in
-the queue.
+number of node-hours (or core-hours) that  is not utilized by regular jobs
+submitted to the queues on Titan. Titan’s backfill functionality offers the
+opportunity to avoid the overhead of queue wait times without using pilot
+abstraction.  It is important to note that backfill availability is  computed
+retrospectively and is determined by Titan's operation team. Backfill
+availability represents the theoretical maximum of node-hours that could be
+utilized by ATLAS (or any other project) that submit under backfill mode.
 
-Compared to pilots, backfill has the disadvantage of limiting the amount of work
-nodes that can be requested. Pilots are normal jobs: they can request as many
-worker nodes for as much time as a queue can offer. On the contrary, jobs sized
-on the basis of backfill information availability \jhanote{should be resource
-availability or backfill queue information?}\mtnote{AFAIH, there is no dedicated
-queue to backfill} depend on the number of worker nodes that cannot be given to
-any other job in the Titan's queue.
-% \sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.
+Compared to pilots, backfill has the disadvantage of limiting the amount of
+work nodes that can be requested. Pilots are normal jobs: they can request as
+many worker nodes for as much time as a queue can offer. On the contrary, jobs
+sized on the basis of backfill information availability depend on the number
+of worker nodes that cannot be given to any other job in the Titan's queue. %
+\sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.
 
-Usually, backfill availability is a fraction of the total capacity of the queue
-but the size of Titan mitigates this limitation. Every year, about 10\% of
-Titan's capacity remains unused, corresponding to an average of 30,000 unused
-cores. This equals approximately 270M core hours per year, roughly 30\% of the
-overall capacity of WLCG.
+Backfill availability is typically a small fraction of the total capacity  of
+a resource. Notwithstanding,  given the size of Titan this translates into a
+substantial capacity. Every year, about 10\% of Titan's capacity remains
+unused, corresponding to an average of 30,000 unused cores. This equals
+approximately 270M core hours per year, roughly 30\% of the overall capacity
+of WLCG.
 

### 2017-04-01 PandaOverview.tex

@@
 progressively introduced for job execution~\cite{crooks2012multi} but, so far,
-% no MPI jobs have been considered for production
 HEP applications have not required MPI\mtnote{Addresses Alexei's comment. If

### 2017-04-01 PandaEvolution.tex
Ongoing: address Alexei's comments

@@
 We developed and deployed a single-point solution to better understand the
-problem space and study abstractions for the integration of HTC and HPC
-computing paradigms. Specifically, the PanDA team developed a job broker to
+problem space of enabling a WMS designed for HTC to execute production workflows
+on resources designed to support HPC. The PanDA team developed a job broker to
 support the execution of part of the ATLAS production Monte Carlo workflow on
@@
 
+\mtnote{Address Alexei's comments: Add a paragraph explaining how this solution has been ported to other HPC machines and reference relevant publications.}
+
 % from the Job Dispatcher module of PanDA Server (as done by a PanDA Pilot);

### 2017-04-01 overall.tex
Run 3 --> 2 fix (minor, local).

@@
 Grid, Cloud, and HPC infrastructures together. These figures confirms the
-relevance of supercomputers' resource contribution to the LHC Run 3, especially
+relevance of supercomputers' resource contribution to the LHC Run 2, especially
 when accounting for the amount of unused backfill availability and the rate of

### 2017-04-02 PandaEvolution.tex
PanDA Broker Architecture and anonymization
Add caption

@@
   \begin{center}
-    \includegraphics[width=\columnwidth]{figures/PanDA_setup_at_OLCF.png}
-    \caption{Schematic view of the PanDA Broker.}
+    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
+    \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates
+    the execution process of a detector simulation job, part of the production
+    ATLAS MD workflow.}
   \end{center}

### 2017-04-02 nge.tex
edits, fixes and comments in 5.0 and 5.1

@@
 
-\jhanote{admittedly i haven't read earlier section in question  carefully, but
-i'm a bit confused about the use of "wide area". wide area is typically used for
-cross data center}\mtnote{Every PanDA Pilot has to communicate with PanDA
-Server. PanDA Server runs on dedicated resources, outside OLCF. Titan's worker
-nodes cannot communicate with PanDA Server because their connectivity is limited
-to Titan's network. Do you have a suggestion for an alternative to ``wide
-area''?} \jhanote{how is the following: "... deployed on the DTNs as Titan's worker
-nodes lack connectivity to the wide area network"}.
+% \jhanote{admittedly i haven't read earlier section in question  carefully, but
+% i'm a bit confused about the use of "wide area". wide area is typically used for
+% cross data center}\mtnote{Every PanDA Pilot has to communicate with PanDA
+% Server. PanDA Server runs on dedicated resources, outside OLCF. Titan's worker
+% nodes cannot communicate with PanDA Server because their connectivity is limited
+% to Titan's network. Do you have a suggestion for an alternative to ``wide
+% area''?} \jhanote{how is the following: "... deployed on the DTNs as Titan's worker
+% nodes lack connectivity to the wide area network"}.
 
-As seen in \S\ref{sec:panda_deployment}, PanDA Broker was deployed on DTNs
-because of the absence of wide area network connectivity on Titan's worker
-nodes. The lack of pilot capabilities impacts both the efficiency and the
-flexibility of PanDA's execution process. Pilots could help to improve
-efficiency by increasing throughput and by enabling better use of backfill
-utilization. Further, pilots makes it easier to support heterogeneous
-workloads.
+As seen in \S\ref{sec:panda_deployment}, PanDA Broker was  deployed on the
+DTNs as Titan's worker nodes lack connectivity to the wide area network The
+lack of pilot capabilities impacts both the efficiency and the flexibility of
+PanDA's execution process. Pilots could improve efficiency by increasing
+throughput and enabling greater backfill utilization. Further, pilots makes it
+easier to support heterogeneous workloads.
 
@@
 
-The support of  multiple generations of workload would enable a more efficient
+The support of  multiple generations of workload would enable more efficient
 use of the backfill availability of walltime. Currently, when a set of
@@
 amount of remaining walltime and the distribution of execution time per event,
-as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event- distrib}.
+as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event-distrib}.
 These capabilities would increase the efficiency of the PanDA Broker when
@@
 
-NGE is a runtime system designed to execute heterogeneous and dynamic workloads
-on diverse resources. Fig.~\ref{fig:arch-overview} illustrates its current
-architecture as deployed on Titan: the two management modules represent a
-simplified version of the PanDA Broker while the agent module is the pilot
-submitted to Titan and executed on its worker nodes. The communication between
-PanDA Broker and PanDA Server is abstracted away as it is not immediately useful to
-evaluate the performance and capabilities of a pilot on Titan.
+\jhanote{Alternative: "NGE is a runtime system to execute heterogeneous and
+dynamically determined tasks that constitute workloads." This definition put
+the focus on tasks and not workloads, which is a level above in the RADICAL
+conceptual layer diagram.} NGE is a runtime system designed to execute
+heterogeneous and dynamic workloads on diverse resources. Fig.~\ref{fig:arch-
+overview} illustrates its current architecture as deployed on Titan: the two
+management modules (Pilot and Unit) represent a simplified version of the PanDA Broker while
+the agent module is the pilot submitted to Titan and executed on its worker
+nodes. The communication between PanDA Broker and PanDA Server is abstracted
+away as it is not immediately useful to evaluate the performance and
+capabilities of a pilot on Titan.
 
@@
 
-NGE exposes an API for the application layer to describe workloads
-(Fig.~\ref{fig:arch-overview}, green squares) and pilots
+NGE exposes an API for the application  \jhanote{workload management?} layer
+to describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
 (Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
-and a UnitManager. The PilotManager submits pilots to Titan's PBS batch system
-via SAGA API (Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker
-to submit MPI scripts. Once scheduled, the pilot's Agent is bootstrapped on
-Titan's MOM node and worker nodes, and  the UnitManager schedules units to the
-Agent's Scheduler (Fig.~\ref{fig:arch-overview}, solid arrow). The Agent's
-Scheduler schedules the units on one or more Agent's Executor for execution. The
-Agent's executors can manage one or more worker nodes, depending on performance
-evaluations.
+and a UnitManager.   The PilotManager submits pilots to Titan's PBS
+\jhanote{for consistency should we say Moab?} batch system via SAGA API
+(Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit
+MPI scripts. Once scheduled, the pilot's Agent is bootstrapped on Titan's MOM
+node and \jhanote{the Agent's Executors on the} worker nodes, and  the
+UnitManager schedules units to the Agent's Scheduler (Fig.~\ref{fig:arch-
+overview}, solid arrow). The Agent's Scheduler schedules the units on one or
+more Agent's Executor for execution. The Agent's executors can manage one or
+more worker nodes, depending on performance evaluations.
 
@@
 be used to enable PanDA Broker to schedule pilots instead of MPI
-scripts.\mtnote{Rewrite this after checking it with Andre.}
+scripts.\mtnote{Rewrite this after checking it with Andre.}\jhanote{In the interest of space and clarity, I would  remove the last sentence of this para. And merge
+the first sentence with the para above.}
 
@@
 
-The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for communication
-and coordination of units execution. ORTE is a spin-off from the Open-MPI
-project and is a critical component of the OpenMPI implementation. It was
-developed to support distributed high-performance computing applications
-operating in a heterogeneous environment. The system transparently provides
-support for interprocess communication, resource discovery and allocation, and
-process launching across a variety of platforms. ORTE provides a mechanism
-similar to the Pilot concept - it allows the user to create a \emph{``dynamic
-virtual machine''} (DVM) that spans multiple nodes. ORTE provides libraries to
-enable the submission, monitoring and managing of tasks, avoiding filesystem
-bottlenecks and race conditions with network sockets. As a consequence, it is
-able to minimize the system overhead while submitting tasks.
+The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
+communication and coordination of the execution of units. ORTE is a critical
+component of the OpenMPI implementation, developed to support distributed
+high-performance computing applications operating in a heterogeneous
+environment~\cite{XXX}. The system \jhanote{which system? ORTE or OpenMPI in
+general?} transparently provides support for interprocess communication,
+resource discovery and allocation, and process launching across a variety of
+platforms. ORTE provides a mechanism similar to the Pilot concept - it allows
+the user to create a \emph{``dynamic virtual machine''} (DVM) that spans
+multiple nodes. ORTE provides libraries to enable the submission, monitoring
+and managing of tasks, avoiding filesystem bottlenecks and race conditions
+with network sockets. As a consequence, it is able to minimize the system
+overhead while submitting tasks.
 

### 2017-04-02 Experiments.tex
5.2 Weak Scalability.
  AA please review/process.

@@
 
-In this section we present experiments % that aim
-to determine the performances of the NGE and to show how it provides a minimal
-overhead while introducing new functionalities.
+We designed experiments to characterize the performance of the NGE on Titan,
+with an emphasis on understanding its overhead and thus the cost of
+introducing new functionalities.  We perform three groups of experiments in
+which we investigate the weak scalability, weak scalability with multiple
+generation, and strong scalability of the NGE.
 
-Experiments % consist in running
-run instances of AthenaMP by using NGE pilots, % where each AthenaMP simulates
-simulating a pre-determined number of events in the ATLAS detector.\mtnote{are
-they 100 events?}\aanote{`Not always.'}
-% taken from ATLAS workload.
-We present three groups of experiments in which we test the NGE for weak
-scalability, weak scalability with multiple generation, and strong scalability.
+Each experiment entails executing multiple instances of AthenaMP using the NGE
+to simulate a pre-determined number of events. All the experiments have been
+performed by  configuring AthenaMP to use all the 16 cores  of Titan's worker
+nodes.
 
-We % collected data about
-measured the execution time of the pilots and of the AthenaMP % that have been
-executed within them, collecting timestamps at % by considering
-all stages of the execution. Experiments have been performed % out of production
-by % using
-submitting NGE's pilots % on
-to Titan's batch queue. Because of Titan's % batch queue
-policies, the turnaround time of each run of our experiments is dominated by
-% the time spent on TITAN's queue
-queue time. Since we are interested only in the performances of the NGE, we
+We  measured the execution time of the pilots and of the AthenaMP  executed
+within them, collecting timestamps at  all stages of the execution.
+Experiments have been performed  by  submitting NGE's pilots  to Titan's batch
+queue.  The turnaround time of an individual run is determined by queue
+waiting times. Since we are interested only in the performances of the NGE, we
 removed queue time from our statistics.
-
-All the experiments have been performed by % setting
-configuring AthenaMP
-% in such a way that it uses
-to use all the 16 cores % present on
-of Titan's worker nodes.
 
@@
 
-% The
-In this experiment % consists in running
-we run as many AthenaMP instances (also referred as tasks from now-on) as the number of nodes controlled by the pilot. Each AthenaMP
-% have been assigned with 100 events whose simulation requires
-simulates 100 events, requiring $\sim 70$ minutes on average.
+In this experiment  we run as many AthenaMP instances (hereafter referred to
+as tasks)  as the number of nodes controlled by the pilot. Each AthenaMP
+simulates 100 events, requiring $\sim 70$ minutes on average. \jhanote{Please
+convert all measurements into seconds!}
 
-Tasks do not % experience
-wait within the NGE Agent's queue % within the pilot
-since % there is
-one node %for
-is available to each AthenaMP instance. % Therefore, d
-Delays in tasks execution are consequence of % only
-three other factors: (i) the bootstrapping of the pilot on the nodes; (ii) the
-UnitManager, as defined in Section~\ref{sec:arch}, that has to dispatch tasks to
-the agent; and (iii) the time that the agent requires to spawn all the tasks on
-the nodes.
+Tasks do not  wait within the NGE Agent's queue since  one node  is available
+to each AthenaMP instance.  Overheads in task execution are consequence
+primarily of the three other factors: (i) the  initial bootstrapping of the
+pilot on the nodes; (ii) the UnitManager's dispatching of  units (tasks) to
+the agent; and (iii) time for the agent to spawn all the tasks on the nodes.
+\jhanote{"spawn" is the right word here. The agent does not spawn tasks. These
+are predefined. I would say "initialize" maybe? Something better?}
 
@@
 %. For all of them, the walltime was
-2 hours walltime. Figure~\ref{fig:weakScal1a} depicts the average pilot
-duration, the average execution time of AthenaMP, and the pilot's overhead as
+2 hours walltime. The time duration is determined by the Titan's walltime policy. Figure~\ref{fig:weakScal1a} depicts the average pilot
+duration, the average execution time of AthenaMP, and the NGE pilot overhead as
 function of the pilot size.
@@
         \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
-    \caption{Weak scalability: average pilot duration, average duration of a
-    single AthenaMP execution, and pilot's overhead for different pilot sizes
-    (250, 500, 1000 and 2000 worker nodes). }
+    \caption{Weak scalability: Average duration of a
+    single AthenaMP execution, average pilot duration and pilot's overhead as a function of different pilot sizes.
+\jhanote{Caption should be in consistent order as the bars in the figure.
+Ideally pilot first then AthenMP.}}
 \label{fig:weakScal1a}
@@
 notice that NGE's overhead does not grow linearly with the number of units.
+\jhanote{It does grow linearly with units! What am I missing??}
 

### 2017-04-02 Experiments.tex
5.2 and 5.3 --  refine and reduce

@@
 
-This experiment is similar to the one presented above but, in this case, we want
-to test also the impact of submitting multiple generations of % new
-AthenaMP to the same pilot.\mtnote{we used `generations' at the beginning of 6,
-I am assuming the reader understands the term by now. Please change if you
-disagree.}\aanote{I agree}
-% instances in place of those that end.
-% The reason behind this experiment is to
-In this way, we stress the pilot's components, % by pushing new tasks while others are ending their execution
-as new tasks are scheduled for execution on the Agent while other tasks are still running.
+The NGE provides an important new capability of submitting multiple
+generations of AthenaMP to the same pilot. In order to investigate the cost of
+doing so, we performed a variant of the weak scalability experiments. This
+stresses the pilot's components, as new tasks are scheduled for execution on
+the Agent while other tasks are still running.
 
-\mtnote{we do
-we need to keep it below 2 hours?}\aanote{You are right. It was misleadind. I Changed. Check if you like it.}
-% decided to send for execution a number of AthenaMP instances equal to
-We performed the experiment by running five AthenaMP instances per node. % and w
-Since we are focusing on the overhead generated by the scheduling and bootstrap of AthenaMP instances, we also reduced  the number of events simulated by each AthenaMP to sixteen in such a way that the running time of each AthenaMP is, on average, $\sim 20$ minutes.
-This choice has been made to avoid the wasting of Titan's core hours but it does not affect the aim of the experiment.  
+\mtnote{we dowe need to keep it below 2 hours?}\aanote{You are right. It was misleadind. I Changed. Check if you like it.}
+
+In these experiments, we run five AthenaMP instances per node.  As these
+experiments are designed to investigate the overhead generated by the
+scheduling and bootstrap of AthenaMP instances, we reduced the number of
+events simulated by each AthenaMP task to sixteen in such a way that the
+running time of each AthenaMP is, on average, $\sim 20$ minutes. This
+experiment design choice does not affect the  objectives or accuracy of the
+experiments, but allows us to scale experiments to large node counts by being
+conservative with allocation.
+
 % \footnote{Note that the number of events has been chosen in such a way
@@
 
-We tested
+We ran
 % different
@@
 previous experiments \mtnote{mostly due to the increased overhead?}\aanote{Well, I guess that's plain.}.
-Despite this, we can notice that the growth of the overhead is consistent with the increment of the number of tasks per node for pilots with 256, 512 and 1024 worker nodes, and much less than linear for the pilot with 2048 worker nodes \aanote{Although that histogram is not trustworty because it is generate with only two runs.}
+Despite this, we notice that the growth of the overhead is consistent with the increment of the number of tasks per node for pilots with 256, 512 and 1024 worker nodes, and much less than linear for the pilot with 2048 worker nodes \aanote{Although that histogram is not trustworty because it is generate with only two runs.}
 
@@
 The last experiments % deals with
-study strong scalability by running the same amount of tasks for different pilot sizes. We used
-% a number of AthenaMP instances equal to
+study strong scalability by running the same number of tasks for different pilot sizes. We used
 2048 AthenaMP instances and % tested
@@
 As a consequence, the number of consecutive generations of AthenaMP decreases with the pilot size by generating different dynamics within the pilots.
-This experiment aim to show that the pilot overhead is not affected by the concurrency level within the pilot but it depends only on the number of tasks.
+These experiments are designed to investigate whether pilot overhead is affected by the degree of concurrency within the pilot and/or the number of tasks.
 \mtnote{Not sure I understand this sentence}\aanote{My fault.

### 2017-04-02 introduction.tex
minor refinement of introduction.
  remove abstract to give to Jack et al.

@@
 The Large Hadron Collider (LHC) was created to explore the fundamental
-properties of matter for the next decades.  Since LHC start-up in 2009,
-multiple experiments  at LHC have collected and distributed hundreds of
-petabytes of data worldwide to hundreds of computer centers. Thousands of
-physicists analyze petascale data volumes daily. The detection of the Higgs
-Boson speaks to the success of the detector and experiment design, as well as
-the sophistication of computing systems devised to analyze the data.
+properties of matter for the next decades.   Multiple experiments at LHC have
+collected and distributed hundreds of petabytes of data worldwide to hundreds
+of computer centers. Thousands of physicists analyze petascale data volumes
+daily. The detection of the Higgs Boson in 2013 speaks to the success of the
+detector and experiment design, as well as the sophistication of computing
+systems devised to analyze the data.
 
-Historically the computing systems consisted of the federation of a hundreds
-to thousands of distributed resources \textemdash{} ranging in scale from
-small to mid-size resource~\cite{foster2003grid}. Although the workloads to be
-executed are independent, the management of the distribution of extremely-
-large workloads across many heterogeneous resources to ensure the effective
+Historically the computing systems used by LHC experiments consisted of the
+federation of a hundreds to thousands of distributed resources \textemdash{}
+ranging in scale from small to mid-size resource~\cite{foster2003grid}.
+Although the workloads to be executed are comprised of tasks that are
+independent of each other, the management of the distribution of 
+workloads across many heterogeneous resources to ensure the effective
 utilization of resources and efficient execution of workloads presents non-
@@
 paradigm of {\it high-throughput computing} viz., the effective execution of
-many independent workloads.
+many independent tasks.
 
-As the LHC prepares for the high-luminosity era (Run 3 in $~\approx$ 2022), it
-is anticipated that the data volumes that will need analyzing will increase by
-10X compared to the current phase (Run 2).   The data will be larger in volume
-but will also require heterogeneous computational processes. In spite of the
-impressive scale of the ATLAS distributed computing system, demand for
-computing systems will significantly outstrip supply (availability).
+As the LHC prepares for the high-luminosity era (Run 4) and Run 3 in
+$~\approx$ 2022, it is anticipated that the data volumes that will need
+analyzing will increase by factors of 10-100  compared to the current phase
+(Run 2). The data will be larger in volume but will also require more
+sophisticated computational processing. In spite of the impressive scale of
+the ATLAS distributed computing system, demand for computing systems will
+significantly outstrip supply (availability).
 
@@
 workload management system with Titan ~\textemdash{} a DOE leadership
-computing facility and scaled to analyze up to \jhanote{XX\%} of the ATLAS
-experiment workload. We present the design of PanDA and show how localized
-substitution to use well established pilot-concepts allow enhanced support for
-heterogeneous workloads (such as molecular dynamics) and advanced execution
-modes.
+computing facility. Consequently, Titan now analyzes up to 5\% of workload  of
+the ATLAS experiment. The case study presents the design and integration of
+PanDA to reach sustained production scales (ca. 100 M core-hours a years). It
+also discusses the investigation of a task execution runtime system on Titan,
+that is based on  the pilot-abstractions and which allows advanced execution
+modes of the ATLAS workload as well as the enhanced support for heterogeneous
+workloads (such as molecular dynamics).
 
-This state-of-practice paper provides multiple contributions.  (i) Documents
-the many design and operational considerations that have been  taken to
-support the sustained, scalable and production usage of Titan for historically
-high-throughput workloads, (ii) Extensions to PanDA to support non-traditional
-heterogeneous workloads and execution modes,  and (iii) As the community looks
-forward to designing the next generation of online analytical platforms
-~\cite{foap} the lessons learnt from our project provide some guidance for how
-current and future experimental and observational systems can be integrated
-with supercomputers in production.
+This state-of-practice paper provides multiple contributions.  It (i)
+documents the many design and operational considerations that have been  taken
+to support the sustained, scalable and production usage of Titan for
+historically high-throughput workloads, (ii) Extensions to PanDA to support
+non-traditional heterogeneous workloads and execution modes,  and (iii) As the
+community looks forward to designing the next generation of online analytical
+platforms~\cite{foap}, this project provides some early lessons and guidance
+for how current and future experimental and observational systems can be
+integrated with production supercomputers in a general and extensible manner.
 

### 2017-04-02 PandaEvolution.tex

@@
 Titan, PanDA pilot was repurposed to serve as a job broker on the DTN nodes of
-Titan. Maintaining the core modules of PanDA Pilot and its stand-alone
-architecture, this prototype called `PanDA Broker' implements functionalities
-to: (i) interrogate Titan about backfill availability; (ii) pull MC jobs and
-events; (iii) wrap the payload of ATLAS jobs into MPI scripts; (iv) submitting
-MPI scripts to Titan's PBS batch system and monitor their execution; and (v)
-staging input/output files. Backfill querying, payload wrapping, and scripts
-submission required a new implementation while pulling ATLAS job and events, and
-file staging were inherited from PanDA Pilot.
+Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of PanDA Pilot
+and its stand-alone architecture, this prototype called `PanDA Broker'
+implements functionalities to: (i) interrogate Titan about backfill
+availability; (ii) pull MC jobs and events; (iii) wrap the payload of ATLAS jobs
+into MPI scripts; (iv) submitting MPI scripts to Titan's PBS batch system and
+monitor their execution; and (v) staging input/output files. Backfill querying,
+payload wrapping, and scripts submission required a new implementation while
+pulling ATLAS job and events, and file staging were inherited from PanDA Pilot.
 
@@
 Once deployed on Titan, every PanDA Broker supports the execution of MC detector
-simulation in 8 steps: (i) PanDA Broker queries Titan's Moab scheduler about
-current backfill availability; (ii) the PanDA Broker queries the Job Dispatcher
-module of the PanDA server for ATLAS jobs that have been bound to Titan by JEDI;
-(iii) Upon receiving the descriptions of those jobs, PanDA Broker pulls their
-input files from the ATLAS DDM service to the DTN; (iv) the PanDA Broker creates
-an MPI script, wrapping enough ATLAS jobs' payload to fit backfill availability;
-(v) the PanDA Broker submits the MPI script to the Titan's PBS batch system via
-RADICAL-SAGA; (vi) upon execution on the worker node(s), the MPI script creates
-configures and executes one AthenaMP for each work node available; (vi) AthenaMP
-spawns 1 event simulation process on each available core (16); (vii) during
-execution, PanDA Broker monitors execution progress and sends PanDA Server
-``heart beats'' for each job.; and (viii) upon completion of each simulation,
-the PanDA Broker locates the output on the shared filesystem, transfer it to
-designated computing centre, and performs cleanup.
+simulation in 9 steps. PanDA Broker queries the Job Dispatcher module of the
+PanDA server for ATLAS jobs that have been bound to Titan by JEDI
+(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA Broker
+pulls jobs' input files from the ATLAS DDM service to the OLCF Lustre file
+system (Fig.~\ref{fig:panda_broker}:3). PanDA Broker queries Titan's Moab
+scheduler about current backfill availability (Fig.~\ref{fig:panda_broker}:2)
+and creates an MPI script, wrapping enough ATLAS jobs' payload to fit backfill
+availability. PanDA Broker submits the MPI script to the Titan's Torque batch
+system via RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).
 
-\mtnote{Address Alexei's comments: Add a paragraph explaining how this solution has been ported to other HPC machines and reference relevant publications.}
+Upon execution on the worker node(s) (Fig.~\ref{fig:panda_broker}:5), the MPI
+script initializes and configures the execution environment
+(Fig.~\ref{fig:panda_broker}:6), and executes one AthenaMP for each available
+work node (Fig.~\ref{fig:panda_broker}:7). AthenaMP retrieves events from Lustre
+(Fig.~\ref{fig:panda_broker}:8) and spawns 1 event simulation process on each of
+the 16 available cores (Fig.~\ref{fig:panda_broker}:9).
+% During execution, PanDA Broker monitors execution progress and sends PanDA
+% Server ``heart beats'' for each job.; and (viii)
+Upon completion of each MPI script, PanDA Broker transfer the jobs' output to
+designated computing centre via the ATLAS DDM(Fig.~\ref{fig:panda_broker}:10),
+and performs cleanup.
+
+\mtnote{Address Alexei's comments: Add a paragraph explaining how this solution
+has been ported to other HPC machines and reference relevant publications.}
+While PanDA Broker implementation is resource specific, it was successfully
+ported also to other supercomputers, including the Ansel (?) at \ldots and
+\ldots at the National Energy Research Scientific Computing Center (NERSC).
 

### 2017-04-02 Experiments.tex
Fixing according Shantenu's comments

@@
 as tasks)  as the number of nodes controlled by the pilot. Each AthenaMP
-simulates 100 events, requiring $\sim 70$ minutes on average. \jhanote{Please
-convert all measurements into seconds!}
+simulates 100 events, requiring $\sim 4200$ seconds on average. \jhanote{Please
+convert all measurements into seconds!}\aanote{Done}
 
@@
 pilot on the nodes; (ii) the UnitManager's dispatching of  units (tasks) to
-the agent; and (iii) time for the agent to spawn all the tasks on the nodes.
+the agent; and (iii) time for the agent to initialize all the tasks on the nodes.
 \jhanote{"spawn" is the right word here. The agent does not spawn tasks. These
-are predefined. I would say "initialize" maybe? Something better?}
+are predefined. I would say "initialize" maybe? Something better?}\aanote{Done. I put initialize, maybe bootstrap would be better?}
 
@@
         \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
-    \caption{Weak scalability: Average duration of a
-    single AthenaMP execution, average pilot duration and pilot's overhead as a function of different pilot sizes.
+    \caption{Weak scalability: average pilot duration, average  duration of a
+    single AthenaMP execution, and pilot's overhead as a function of different pilot sizes.
 \jhanote{Caption should be in consistent order as the bars in the figure.
@@
 We % can
-notice that NGE's overhead does not grow linearly with the number of units.
-\jhanote{It does grow linearly with units! What am I missing??}
+notice that NGE's overhead does grow linearly with the number of units.
+\jhanote{It does grow linearly with units! What am I missing??}\aanote{Corrected. I think that ``not'' came from a previous version of the sentence. }
 
@@
 
-\mtnote{we dowe need to keep it below 2 hours?}\aanote{You are right. It was misleadind. I Changed. Check if you like it.}
+\mtnote{we dowe need to keep it below 2 hours?}\aanote{You are right. It was misleading. I Changed. Check if you like it.}
 
@@
 events simulated by each AthenaMP task to sixteen in such a way that the
-running time of each AthenaMP is, on average, $\sim 20$ minutes. This
+running time of each AthenaMP is, on average, $\sim 1200$ seconds. This
 experiment design choice does not affect the  objectives or accuracy of the
@@
 previous experiments \mtnote{mostly due to the increased overhead?}\aanote{Well, I guess that's plain.}.
-Despite this, we notice that the growth of the overhead is consistent with the increment of the number of tasks per node for pilots with 256, 512 and 1024 worker nodes, and much less than linear for the pilot with 2048 worker nodes \aanote{Although that histogram is not trustworty because it is generate with only two runs.}
+Despite this, we notice that the growth of the overhead is consistent with the increment of the number of tasks per node for pilots with 256, 512 and 1024 worker nodes, and less than linear for the pilot with 2048 worker nodes.
 

### 2017-04-02 data/NGE/script.py

@@ -1,0 +1,55 @@
+import dataAnalysis as da
+import sys
+path = sys.argv[1]
+states = [da.REC,da.SCHED,da.QUEUE,da.EXEC]
+sampling = 0.1
+output = path+".csv"
+outfile = open(output,"w+")
+#256
+folder = "exp1"
+print("Processing "+folder)
+times = da.extractTimes(path+"/"+folder,states,sampling)
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[5],1.0,perc=0.05)
+outfile.write("1 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[4],1.0,perc=0.05)
+outfile.write("1 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[2],1.0,perc=0.05)
+outfile.write("1 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[3],1.0,perc=0.05)
+outfile.write("1 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +"\n")
+#512
+folder = "exp2"
+print("Processing "+folder)
+times = da.extractTimes(path+"/"+folder,states,sampling)
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[5],1.0,perc=0.05)
+outfile.write("2 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[4],1.0,perc=0.05)
+outfile.write("2 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
+(avg,sd,hmin,hmax,count) = da.getAvgAndSDFromList(times[2],1.0,perc=0.05)
+outfile.write("2 " +  str(avg) + " " + str(sd) + " "+ str(hmin) + " " + str(hmax) +" ")
[…]

### 2017-04-02 appendix.tex
Address comments on figure 5

@@
     \item [Section 5.1 -- Figure 4] \url{}
-    \item [Section 5.1 -- Figure 5] \url{}
+    \item [Section 5.1 -- Figure 5] \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5}
     \item [Section 5.1 -- Figure 6] \url{}
@@
 % -----------------------------------------------------------------------------
-\subsection{Section 6.3 - Figure 14}
-\label{apndx:fig14}
-
-Artifacts description.
+\subsection{Section 6.3 - Figure 7}
+\label{apndx:fig9}
 
 \begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
+    \item Raw data: \url{https://github.com/ATLAS-Titan/misc/tree/master/atlas-paper-experiments/data}
+     \item Processed data: \url{ https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/data.csv} 
+    \item Raw data processing: 
+    	\begin{itemize}
+    		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/dataAnalysis.py}
+    		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/script.py}
+    		\item Follow instructions at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-extract-data-from-the-sandbox}
+   	\end{itemize}
+    \item Plotting code: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/plot.gnu}
 \end{itemize}
 
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
+Replicability: Follow instructions at the following link: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-reproduce-experiments-on-TITAN} to prepare the environment on Titan.
+Launch:
+\begin{itemize}
+\item python dispatcher.py 250 250 120 etc/input_100 to replicate experiments with 250 nodes;
+\item python dispatcher.py 500 500 120 etc/input_100 to replicate experiments with 500 nodes;
+\item python dispatcher.py 1000 1000 120 etc/input_100 to replicate experiments with 1000 nodes;
+\item python dispatcher.py 2000 2000 120 etc/input_100 to replicate experiments with 2000 nodes.
+\end{itemize}
 
 % -----------------------------------------------------------------------------
-\subsection{Section 6.3 - Figure 15}
-\label{apndx:fig15}
-
-Artifacts description.
-
+\subsection{Section 6.3 - Figure 8}
+\label{apndx:fig10}
 \begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
+    \item Raw data: \url{https://github.com/ATLAS-Titan/misc/tree/master/atlas-paper-experiments/weak2}
+     \item Processed data: \url{ https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/weak2.csv} 
+    \item Raw data processing: 
+    	\begin{itemize}
+    		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/dataAnalysis.py}
+    		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/script.py}
+    		\item Follow instructions at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-extract-data-from-the-sandbox}
+   	\end{itemize}
+    \item Plotting code: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/plot.gnu}
 \end{itemize}
 
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
+Replicability: Follow instructions at the following link: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-reproduce-experiments-on-TITAN} to prepare the environment on Titan.
+Launch:
+\begin{itemize}
+\item python dispatcher.py 256 1280 256 180 etc/input_16 to replicate experiments with 256 nodes;
+\item python dispatcher.py 512 2560 512 180 etc/input_16 to replicate experiments with 512 nodes;
+\item python dispatcher.py 1024 5120 180 etc/input_16 to replicate experiments with 1024 nodes;
+\item python dispatcher.py 2048 10240 180 etc/input_16 to replicate experiments with 2048 nodes.
+\end{itemize}
 
 % -----------------------------------------------------------------------------
-\subsection{Section 6.3 - Figure 16}
+\subsection{Section 6.3 - Figure 9}
 \label{apndx:fig16}
 
-Artifacts description.
-
+\label{apndx:fig10}
 \begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
+    \item Raw data: \url{https://github.com/ATLAS-Titan/misc/tree/master/atlas-paper-experiments/strong}
+     \item Processed data: \url{ https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/strong.csv} 
+    \item Raw data processing: 
+    	\begin{itemize}
+    		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/dataAnalysis.py}
+    		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/script.py}
+    		\item Follow instructions at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-extract-data-from-the-sandbox}
+   	\end{itemize}
+    \item Plotting code: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/plot.gnu}
 \end{itemize}
 
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
+Replicability: Follow instructions at the following link: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-reproduce-experiments-on-TITAN} to prepare the environment on Titan.
+Launch:
+\begin{itemize}
+\item python dispatcher.py 256  2048 256 180 etc/input_16 to replicate experiments with 256 nodes;
+\item python dispatcher.py 512 2048 512 180 etc/input_16 to replicate experiments with 512 nodes;
+\item python dispatcher.py 1024 2048 180 etc/input_16 to replicate experiments with 1024 nodes;
+\item python dispatcher.py 2048 2048 180 etc/input_16 to replicate experiments with 2048 nodes.
+\end{itemize}
 
 % -----------------------------------------------------------------------------
-\subsection{Section 6.3 - Figure 17}
-\label{apndx:fig17}
-
-Artifacts description.
-
-\begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}

### 2017-04-02 appendix.tex
Solving problem in appendix

@@
 \begin{itemize}
-\item python dispatcher.py 250 250 120 etc/input_100 to replicate experiments with 250 nodes;
-\item python dispatcher.py 500 500 120 etc/input_100 to replicate experiments with 500 nodes;
-\item python dispatcher.py 1000 1000 120 etc/input_100 to replicate experiments with 1000 nodes;
-\item python dispatcher.py 2000 2000 120 etc/input_100 to replicate experiments with 2000 nodes.
+\item python dispatcher.py 250 250 120 etc/input\_100 to replicate experiments with 250 nodes;
+\item python dispatcher.py 500 500 120 etc/input\_100 to replicate experiments with 500 nodes;
+\item python dispatcher.py 1000 1000 120 etc/input\_100 to replicate experiments with 1000 nodes;
+\item python dispatcher.py 2000 2000 120 etc/input\_100 to replicate experiments with 2000 nodes.
 \end{itemize}
@@
 \begin{itemize}
-\item python dispatcher.py 256 1280 256 180 etc/input_16 to replicate experiments with 256 nodes;
-\item python dispatcher.py 512 2560 512 180 etc/input_16 to replicate experiments with 512 nodes;
-\item python dispatcher.py 1024 5120 180 etc/input_16 to replicate experiments with 1024 nodes;
-\item python dispatcher.py 2048 10240 180 etc/input_16 to replicate experiments with 2048 nodes.
+\item python dispatcher.py 256 1280 256 180 etc/input\_16 to replicate experiments with 256 nodes;
+\item python dispatcher.py 512 2560 512 180 etc/input\_16 to replicate experiments with 512 nodes;
+\item python dispatcher.py 1024 5120 180 etc/input\_16 to replicate experiments with 1024 nodes;
+\item python dispatcher.py 2048 10240 180 etc/input\_16 to replicate experiments with 2048 nodes.
 \end{itemize}
@@
 \begin{itemize}
-\item python dispatcher.py 256  2048 256 180 etc/input_16 to replicate experiments with 256 nodes;
-\item python dispatcher.py 512 2048 512 180 etc/input_16 to replicate experiments with 512 nodes;
-\item python dispatcher.py 1024 2048 180 etc/input_16 to replicate experiments with 1024 nodes;
-\item python dispatcher.py 2048 2048 180 etc/input_16 to replicate experiments with 2048 nodes.
+\item python dispatcher.py 256  2048 256 180 etc/input\_16 to replicate experiments with 256 nodes;
+\item python dispatcher.py 512 2048 512 180 etc/input\_16 to replicate experiments with 512 nodes;
+\item python dispatcher.py 1024 2048 180 etc/input\_16 to replicate experiments with 1024 nodes;
+\item python dispatcher.py 2048 2048 180 etc/input\_16 to replicate experiments with 2048 nodes.
 \end{itemize}

### 2017-04-02 nge.tex
Address Shantenu's comments to 6.0/6.1

@@
 
-\jhanote{not sure MPI script is optimal description or ambigious.} Multiple
-generations would also relax two assumptions of the current execution model:
-knowing the number of simulations before submitting the MPI script, and having
-a fixed number of events per simulation (100 at the moment). Pilots would
-enable the scheduling of simulations independently from whether they were
-available at the moment of submitting the pilot. Further, simulations with a
-varying number of events could be scheduled on a pilot, depending on the
-amount of remaining walltime and the distribution of execution time per event,
-as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event-distrib}.
-These capabilities would increase the efficiency of the PanDA Broker when
-there is a large difference between the number of cores and walltime.
+\jhanote{not sure MPI script is optimal description or ambiguous.}\mtnote{Is
+what we used in the previous section(s) to designate what PanDA Broker submits
+to Titan's Torque batch system. Job is already used for what is pulled from
+PanDA server.} Multiple generations would also relax two assumptions of the
+current execution model: knowing the number of simulations before submitting the
+MPI script, and having a fixed number of events per simulation (100 at the
+moment). Pilots would enable the scheduling of simulations independently from
+whether they were available at the moment of submitting the pilot. Further,
+simulations with a varying number of events could be scheduled on a pilot,
+depending on the amount of remaining walltime and the distribution of execution
+time per event, as shown in \S\ref{ssec:panda_titan},
+Figure~\ref{fig:100event-distrib}. These capabilities would increase the
+efficiency of the PanDA Broker when there is a large difference between the
+number of cores and walltime.
 
@@
 \jhanote{Alternative: "NGE is a runtime system to execute heterogeneous and
-dynamically determined tasks that constitute workloads." This definition put
-the focus on tasks and not workloads, which is a level above in the RADICAL
-conceptual layer diagram.} NGE is a runtime system designed to execute
-heterogeneous and dynamic workloads on diverse resources. Fig.~\ref{fig:arch-
-overview} illustrates its current architecture as deployed on Titan: the two
-management modules (Pilot and Unit) represent a simplified version of the PanDA Broker while
-the agent module is the pilot submitted to Titan and executed on its worker
-nodes. The communication between PanDA Broker and PanDA Server is abstracted
-away as it is not immediately useful to evaluate the performance and
-capabilities of a pilot on Titan.
+dynamically determined tasks that constitute workloads." This definition put the
+focus on tasks and not workloads, which is a level above in the RADICAL
+conceptual layer diagram.}\mtnote{Done.} NGE is a runtime system to execute
+heterogeneous and dynamically determined tasks that constitute workloads.
+Fig.~\ref{fig:arch- overview} illustrates its current architecture as deployed
+on Titan: the two management modules (Pilot and Unit) represent a simplified
+version of the PanDA Broker while the agent module is the pilot submitted to
+Titan and executed on its worker nodes. The communication between PanDA Broker
+and PanDA Server is abstracted away as it is not immediately useful to evaluate
+the performance and capabilities of a pilot on Titan.
 
@@
 
-NGE exposes an API for the application  \jhanote{workload management?} layer
-to describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
+NGE exposes an API for the application  \jhanote{workload management?} layer to
+describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
 (Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
-and a UnitManager.   The PilotManager submits pilots to Titan's PBS
-\jhanote{for consistency should we say Moab?} batch system via SAGA API
-(Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit
-MPI scripts. Once scheduled, the pilot's Agent is bootstrapped on Titan's MOM
-node and \jhanote{the Agent's Executors on the} worker nodes, and  the
+and a UnitManager.   The PilotManager submits pilots to Titan's Torque
+\jhanote{for consistency should we say Moab?}\mtnote{Moab = Scheduler; Torque =
+Batch system.} batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash
+arrow), as done by PanDA Broker to submit MPI scripts. Once scheduled, the
+pilot's Agent is bootstrapped on Titan's MOM node and \jhanote{the Agent's
+Executors on the} the Agent's Executors on the worker nodes, and the
 UnitManager schedules units to the Agent's Scheduler (Fig.~\ref{fig:arch-
@@
 more Agent's Executor for execution. The Agent's executors can manage one or
-more worker nodes, depending on performance evaluations.
+more worker nodes, depending on performance evaluations. The UnitManager and the
+Agent communicate via a database that is instantiated on one of Titan's DTN so
+as to be reachable by both modules.
 
-The UnitManager and the Agent communicate via a database that is instantiated on
-one of Titan's DTN so as to be reachable by both modules. A similar approach would
-be used to enable PanDA Broker to schedule pilots instead of MPI
-scripts.\mtnote{Rewrite this after checking it with Andre.}\jhanote{In the interest of space and clarity, I would  remove the last sentence of this para. And merge
-the first sentence with the para above.}
+% A similar approach would be used to enable PanDA Broker to schedule pilots
+% instead of MPI scripts.\mtnote{Rewrite this after checking it with
+% Andre.}
+\jhanote{In the interest of space and clarity, I would  remove the last
+sentence of this para. And merge the first sentence with the para
+above.}\mtnote{Done.}
 
@@
 
-The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
-communication and coordination of the execution of units. ORTE is a critical
-component of the OpenMPI implementation, developed to support distributed
-high-performance computing applications operating in a heterogeneous
-environment~\cite{XXX}. The system \jhanote{which system? ORTE or OpenMPI in
-general?} transparently provides support for interprocess communication,
-resource discovery and allocation, and process launching across a variety of
-platforms. ORTE provides a mechanism similar to the Pilot concept - it allows
-the user to create a \emph{``dynamic virtual machine''} (DVM) that spans
-multiple nodes. ORTE provides libraries to enable the submission, monitoring
-and managing of tasks, avoiding filesystem bottlenecks and race conditions
-with network sockets. As a consequence, it is able to minimize the system
-overhead while submitting tasks.
+The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for communication
+and coordination of the execution of units. ORTE is a critical component of the
+OpenMPI implementation, developed to support distributed high-performance
+computing applications operating in a heterogeneous environment~\cite{XXX}. The
+system \jhanote{which system? ORTE or OpenMPI in general?}\mtnote{I am not sure
+but I think ORTE?} transparently provides support for interprocess
+communication, resource discovery and allocation, and process launching across a
+variety of platforms. ORTE provides a mechanism similar to the Pilot concept -
+it allows the user to create a \emph{``dynamic virtual machine''} (DVM) that
+spans multiple nodes. ORTE provides libraries to enable the submission,
+monitoring and managing of tasks, avoiding filesystem bottlenecks and race
+conditions with network sockets. As a consequence, it is able to minimize the
+system overhead while submitting tasks.
 

### 2017-04-02 PandaEvolution.tex
Address Jack's comments #10

@@
 
-Backfill availability is typically a small fraction of the total capacity  of
-a resource. Notwithstanding,  given the size of Titan this translates into a
+Backfill availability is typically a small fraction of the total capacity  of a
+resource. Notwithstanding,  given the size of Titan this translates into a
 substantial capacity. Every year, about 10\% of Titan's capacity remains
-unused, corresponding to an average of 30,000 unused cores. This equals
-approximately 270M core hours per year, roughly 30\% of the overall capacity
-of WLCG.
+unused~\cite{barker2016us}, corresponding to an average of 30,000 unused cores.
+This equals approximately 270M core hours per year, roughly 10\% of the overall
+capacity of WLCG.
 

### 2017-04-02 PandaOverview.tex
Address Jack's comments #10

@@
 
-Pilot-enabled WMS enable high throughput of tasks execution while supporting
+Pilot-capable WMS enable high throughput of tasks execution while supporting
 interoperability across multiple sites. This is particularly relevant for LHC
@@
 month, analyzing and producing petabytes of data. The design of PanDA WMS
-started in 2005 and its implementation went into production for the LHC Run 1 on
-2009. PanDA was then extended with new subsystems to be deployed on Run 2, on
-2015.
+started in 2005 to support ATLAS, one of the LHC experiments. PanDA went into
+production for the LHC Run 1 on 2009, and was then extended with new subsystems
+to be deployed on Run 2, on 2015.
 

### 2017-04-02 PandaEvolution.tex
Address Jack's comment issue #9

@@
 Titan’s backfill functionality offers the opportunity to avoid the overhead of
-queue wait times without using pilot abstraction. Backfill availability is the
-number of node-hours (or core-hours) that  is not utilized by regular jobs
-submitted to the queues on Titan. Titan’s backfill functionality offers the
-opportunity to avoid the overhead of queue wait times without using pilot
-abstraction.  It is important to note that backfill availability is  computed
-retrospectively and is determined by Titan's operation team. Backfill
-availability represents the theoretical maximum of node-hours that could be
-utilized by ATLAS (or any other project) that submit under backfill mode.
+queue wait times without using pilot abstraction~\cite{maui_backfill_url}.
+Backfill availability is the number of node-hours (or core-hours) that cannot be
+utilized by regular jobs queued on Titan at a specific point in time.  It is
+important to note that backfill availability is computed retrospectively and is
+determined by Titan's operation team. Backfill availability represents the
+theoretical maximum of node-hours that could be utilized by ATLAS (or any other
+project) that submit under backfill mode.
 
-Compared to pilots, backfill has the disadvantage of limiting the amount of
-work nodes that can be requested. Pilots are normal jobs: they can request as
-many worker nodes for as much time as a queue can offer. On the contrary, jobs
-sized on the basis of backfill information availability depend on the number
-of worker nodes that cannot be given to any other job in the Titan's queue. %
-\sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.
+Compared to pilots, backfill has the disadvantage of limiting the amount of work
+nodes that can be requested. Pilots are normal jobs: they can request as many
+worker nodes for as much time as a queue can offer. On the contrary, jobs sized
+on the basis of backfill information availability depend on the number of worker
+nodes that cannot be given to any other job in the Titan's queue at that moment
+in time.
+
+% \sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.
 

### 2017-04-03 PandaOverview.tex
minor corrections and comments in Section 2.

@@
 month, analyzing and producing petabytes of data. The design of PanDA WMS
-started in 2005 to support ATLAS, one of the LHC experiments. PanDA went into
+started in 2005 to support ATLAS. PanDA went into
 production for the LHC Run 1 on 2009, and was then extended with new subsystems
-to be deployed on Run 2, on 2015.
+for Run 2 in 2015.
 
@@
 the PanDA WMS at any point in time, directly or via dedicated application
-frameworks. Consistently, PanDA's security model is based on separation between
+frameworks. PanDA's security model is based on separation between
 authentication, authorization and accounting for both single users and group of
@@
 Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
-Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of ATLAS
-workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}) but, at the
-moment, they are not relevant to understand how PanDA has been ported to
-supercomputers. For a full list of subsystems see Ref.~\cite{panda-wiki_url}.
-Figure~\ref{fig:architecture} shows a diagrammatic representation of PanDA main
-subsystems, highlighting the execution process of tasks while omitting
-monitoring details to improve readability.
+Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of
+ATLAS workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we
+do not discuss them as they are  not relevant to  an understanding of how
+PanDA has been ported to supercomputers. For a full list of subsystems see
+Ref.~\cite{panda-wiki_url}. Figure~\ref{fig:architecture} shows a diagrammatic
+representation of PanDA main subsystems, highlighting the execution process of
+tasks while omitting monitoring details to improve readability.
 
@@
 The relation between tasks and jobs can be one-to-one or one-to-many, and the
-conversion between the two can by static or dynamic. During the LHC Run 1, PanDA
+conversion between the two can by static or dynamic. During LHC Run 1, PanDA
 required users or applications to perform a static conversion between tasks and
@@
 JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users or
-the application layer submits tasks descriptions to JEDI
+the application layer submit task descriptions to JEDI
 (Fig.~\ref{fig:architecture}:1) that stores them into a queue implemented by a
@@
 When a PanDA Pilot becomes available on a site, it calls the Job Dispatcher
-module of the PanDA Server, requesting for a job to execute
+module of the PanDA Server, requesting  a job to execute
 (Fig.~\ref{fig:architecture}:11). The Job Dispatcher interrogates the Task
@@
 executed. Task Buffer checks the global queue (i.e., the PanDA database) and,
-upon availability, returns a suitable job the Job Dispatcher. The Job Dispatcher
+upon availability, returns a suitable job the Job Dispatcher. \jhanote{end of previous sentence needs some attention} The Job Dispatcher
 dispatches that job to the PanDA Pilot (Fig.~\ref{fig:architecture}:12).

### 2017-04-03 PandaEvolution.tex
Iteration 4 and scattered editing

@@
 % -----------------------------------------------------------------------------
-\subsection{Interfacing PanDA with Titan}
+%\subsection{Interfacing PanDA with Titan}
+\subsection{Architectures and Interfaces}
 \label{ssec:panda-titan}
@@
 computational-intensive, requiring less than 2GB of RAM at runtime and with
-small input data requirements. Detector simulation tasks in ATLAS account for
+small input data requirements. Detector simulation tasks in ATLAS are performed
+via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
+GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
 $\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
@@
 % -----------------------------------------------------------------------------
-\subsection{PanDA Broker on Titan}
+% \subsection{PanDA Broker on Titan}
+\subsection{PanDA Broker}
 \label{ssec:panda_titan}
@@
 \begin{figure}
-  \begin{center}
+    \centering
     \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
@@
     ATLAS MD workflow.}
-  \end{center}
 \label{fig:panda_broker}

### 2017-04-03 overall.tex
Iteration 4 and scattered editing

--- 2017-04-01 overall.tex
@@
 
-We evaluate the efficiency, scalability and reliability of the PanDA Brokers
-along two dimensions: (i) backfill utilization by the
-PanDA Brokers; and (ii) the amount of runtime spent performing detector
-simulations. We based these evaluations on measuring the number of cores
-utilized by ATLAS on Titan, the overall backfill availability, the number of
-detector simulations performed on all the resources available to ATLAS and on
-Titan, and the failure rate of these simulations. All the measurements were
-performed between January 2016 and February 2017, hereafter called `experiment
-time window'.
+We evaluate the efficiency, scalability and reliability of the deployment of PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and AthenaMP.
+% two dimensions: (i) backfill utilization by the PanDA Brokers; and (ii) the
+% amount of runtime spent performing detector simulations.
+We based our characterizations on measuring the number of cores utilized by
+ATLAS on Titan, the overall backfill availability, the number of detector
+simulations performed on all the resources available to ATLAS and on Titan, and
+the failure rate of these simulations. We also consider the the time taken to
+setup AthenaMP and to simulate events in the ATLAS detector with it. All the
+measurements were performed between January 2016 and February 2017, hereafter
+called `experiment time window'.
+
+\mtnote{TODO: expand the previous paragraph to include dimensions of AthenaMP characterization.}
 
@@
 
-Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of core-hours
-used by ATLAS on Titan during the experiment time window. ATLAS consumed a total
-of 73.8M core-hours, for an average of $\approx$7M core-hours a month, with a
-minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours in
-February 2017.
+Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of Titan
+core-hours used by ATLAS during the experiment time window. ATLAS consumed a
+total of 73.8M Titan core-hours, for an average of $\approx$7M Titan core-hours
+a month, with a minimum of 3.3M Titan core-hours in April 2016 and a maximum
+14.8M Titan core-hours in February 2017.
 
@@
 (Figure~\ref{fig:backfill-utilization}, red labels) is defined as the fraction
-of Titan’s cores available via backfill that was utilized by the PanDA Brokers.
-The brokers reached 18\% average efficiency, with a minimum 8.9\% efficiency on
-April 2016 and a maximum 33.5\% efficiency on February 2017. The number of total
-backfill cores available was 38.1M in April 2016, and 33.1M in February 2017.
-This shows that the measured difference in efficiency did not depend on a
-comparable difference in total backfill availability.
+of Titan core-hours utilized by the PanDA Brokers (for ATLAS jobs) to the total
+of Titan’s backfill availability. The brokers reached 18\% average efficiency,
+with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\% efficiency on
+February 2017. The total backfill available was 38.1M Titan core-hours in April
+2016, and 33.1M Titan core-hours in February 2017. This shows that the measured
+difference in efficiency did not depend on a comparable difference in total
+backfill availability.
 
@@
 Grid, Cloud, and HPC infrastructures together. These figures confirms the
-relevance of supercomputers' resource contribution to the LHC Run 3, especially
-when accounting for the amount of unused backfill availability and the rate of
-improvement of PanDA efficiency.
+relevance of supercomputers' resource contribution to the LHC Run 2, especially
+when accounting for the amount of unused backfill availability and the
+improvement of PanDA efficiency across the experiment time window.
 
@@
 scheduler at regular intervals during the experiment window time, while
-developing and deploying PanDA Brokers. The mean number of nodes was 680, and
-their mean walltime was 680 minutes. Detector simulations of 100 events, enable
+developing and deploying PanDA Brokers. The mean number of nodes was 691, and
+their mean walltime was 126 minutes. Detector simulations of 100 events, enable
 to use down to 1 node for 1/2 of the mean walltime of backfill availability. As
@@
 
-\begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/titan_backfill_avail.png}
+\begin{figure*}%[htp]
+    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_wnodes_distribution.pdf}
+    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_walltime_distribution.pdf}
+    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
     %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
-    \caption{Backfill availability measures as walltime in minutes vs number of
-    work nodes during the experiment window time. Availability was measured
-    62175 times: mean number of work nodes available 680 (yellow line); mean
-    walltime available 122 minutes (red line).}
+    \caption{Backfill availability: distribution of number of work nodes (left); distribution of walltime in minutes (center); and their correlation (right).
+62175 measures: mean number of work nodes available 691; mean
+    walltime available 126 minutes.}
 \label{fig:backfill-distrib}
-\end{figure}
+\end{figure*}
 
@@
 simulations (3.9\%) on Titan. PanDA Broker could fit the number of events to the
-walltime of the backfill availability on the base of the distribution of
-Figure~\ref{fig:backfill-distrib}. That specific number of event could then be
+walltime of the backfill availability on the base of the distributions of
+Fig.~\ref{fig:backfill-distrib}. That specific number of event could then be
 pulled from the PanDA Event service~\cite{calafiura2015atlas} and given as input
@@
 We use two main parameters to measure the performance of the detector simulation
-jobs submitted to Titan: (i) the time taken to setup
-AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
-GEANT4 simulation toolkit~\cite{agostinelli2003geant4}; and (ii) the
+jobs submitted to Titan: (i) the time taken to setup AthenaMP; and (ii) the
 distribution of the time taken by the Geant4 toolkit to simulate a certain
@@
 and service of AthenaMP is configured by a set of Python scripts. Both these
-operations result in a large number of read operations on the filesystem,
-including those required to access of small python scripts.
+operations result in a large number of read operations on the filesystem shared
+between the worker nodes and the DTNs, including the operations required to
+access of small python scripts.
 
@@
 
-\mtnote{TODO: specify that this is work in progress (see Sergey's comment) and
-see how to move it to the data performance subsection}Initially, all the shared
-libraries of AthenaMP and the python scripts for the configuration stage were
-stored on the Spider 2 Lustre file system. However, the I/O patterns of the
-initialization and configuration stages degraded the performance of the
-filesystem (Figure~\ref{}\mtnote{We may want to aggregate some of the diagrams
-about lustre's experiments here}). Since Spider 2 is a center-wide file system,
-this resulted in  performance degradation for all OLCF resources and users.
-\mtnote{I am afraid the details about the trace are too specific given the space
-constraints of a SC submission. Please feel free to uncomment it if you
-disagree.}
+% \mtnote{TODO: specify that this is work in progress (see Sergey's comment) and
+% see how to move it to the data performance subsection}
 
+Initially, all the shared libraries of AthenaMP and the python scripts for the
+configuration stage were stored on the Spider 2 Lustre file system. However, the
+I/O patterns of the initialization and configuration stages degraded the
+performance of the filesystem.
+% (Figure~\ref{}\mtnote{We may want to aggregate some of the diagrams about
+% lustre's experiments here}).  Since Spider 2 is a center-wide file system,
+% this resulted in  performance degradation for all OLCF resources and users.
+% \mtnote{I am afraid the details about the trace are too specific given the
+% space constraints of a SC  submission. Please feel free to uncomment it if
+% you disagree.}
+%
 % As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
@@
 % off.
-
+%
 % \begin{minipage}{\linewidth}
@@
 % \end{minipage}
-
+%
 % As this problem was identified, the OLCF staff has enabled read-only access to
@@
 % and have it available read-only on a Titan compute node.
+%
+% After careful characterization of the impact of ATLAS jobs on Lustre, the
+% performance degradation
+This was addressed by moving the AthenaMP distribution to a read-only NFS
+directory, shared among DTNs and worker nodes. NFS eliminated the problem of
+metadata contention, improving metadata read performance from $\approx$6,300
+seconds on Lustre to $\approx$1,500 seconds on NFS.
 
-After careful characterization of the impact of ATLAS jobs on Lustre, the
-performance degradation was addressed by moving the AthenaMP distribution to a
-read-only NFS directory, shared among DTNs and worker nodes. This eliminated the
-problem of metadata contention, improving metadata read performance from
-$\approx$6,300 seconds on Lustre to $\approx$1,500 seconds on NFS. Further, at
-configuration time, the input files of each detector simulation job were stored
-to a ramdisk on the work nodes. This offered a marked improvement of reading
-time: from 1,320 seconds on Lustre to 40 seconds on NFS.
-
-%Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
-%improvement on Titan. The circled region illustrates the switch from Lustre to
-%NFS-exported directory for hosting the ATLAS release.
+% Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
+% improvement on Titan. The circled region illustrates the switch from Lustre to
+% NFS-exported directory for hosting the ATLAS release.
 
@@
 %\end{figure}
+
+Once set up, AthenaMP is used to execute 16 concurrent Geant4 simulators on each
+of Titan's worker node. Geant4 requires to read events descriptions from a
+filesystem and simulate them as they would happen within the ATLAS detector. We
+characterized both the compute performance of the simulation and the impact of
+acquiring event descriptions on the filesystem.
 
@@
 % produced with NGE, as discussed in F2F meeting at Rutgers?}
-
+%
 % ----------------------------------------------------------------------------
@@
 % \label{ssec:reliability}
-
-The current design and architecture of the PanDA Broker is proving to be as
-reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
-the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
-while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
-Brokers were responsible for around the 19\% of the failures, compared to the
-29\% of failures produced by the JobDispatcher module of the PanDA Server, and
-the 13\% failures produced by the Geant4 toolkit. The current failure rate of
-the PanDA Brokers confirms the benefits of reusing most of the code base of the
-PanDA Pilot for implementing the PanDA Broker. It also shows that adopting
-third-party libraries like RADICAL-SAGA did not have a measurable adverse effect
-on reliability.
-
+%
 % Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
@@
 % the PanDA Broker failure model.
-
+%
 % \begin{figure}[htp]
@@
 % \end{figure}
+%
+% -----------------------------------------------------------------------------
+% \subsection{Characterizing PanDA I/O}
+%
+% To better understand the I/O impact of ATLAS PanDA project on Titan
+% supercomputing environment
 
-% -----------------------------------------------------------------------------
-\subsection{Characterizing PanDA I/O}
-
-To better understand the I/O impact of ATLAS PanDA project on Titan
-supercomputing environment we analyzed 1,175 jobs ran on the week of 10/25/2016,
-for a total of 174 hours. Table~\ref{panda-olcf-stats} shows the overall
-statistical breakdown of the observed file I/O impact of ATLAS at OLCF\@.
+We studied the impact of acquiring event descriptions on Lustre by analyzing
+1,175 jobs ran on the week of 10/25/2016, for a total of 174 hours.
+Table~\ref{panda-olcf-stats} shows the overall statistical breakdown of the
+observed file I/O.
 % Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written}
@@
 % As can be seen from Table~\ref{panda-olcf-stats},
-
-ATLAS jobs use between 1 and 300 worker nodes, and 35 on average. 75\% of the
-ATLAS jobs consume less than 25 nodes and 92\% less than 100. During the 174
+%
+ATLAS used between 1 and 300 worker nodes, and 35 on average. 75\% of the jobs
+run by ATLAS consumed less than 25 nodes and 92\% less than 100. During the 174
 hours of data collection, 6.75 ATLAS jobs were executed on average per hour,
@@
 \end{tabular}
-\caption{The Statistical breakdown of the I/O impact of 1,175 PanDA jobs executed at OLCF for the week of 10/25/16}
+\caption{The Statistical breakdown of the I/O impact of 1,175 jobs ATLAS executed at OLCF for the week of 10/25/16}
 \label{panda-olcf-stats}
@@
 
-The I/O of the ATLAS job executed on Titan show an interesting pattern.
-Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
-worker node is less than 400 MB on average, while the amount of data written per
-node is less than 170 MB on average. This correlates with our finding that ATLAS
-PanDA jobs are read heavy. However,
+% The I/O of the ATLAS jobs show an interesting pattern.
+% Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
+% worker node is less than 400 MB on average, while the amount of data written
+% per node is less than 170 MB on average. This correlates with our finding
+% that ATLAS PanDA jobs are read heavy. However,
 
-ATLAS PanDA jobs are read heavy: On average, the amount of data read per worker
-node is less than 400 MB, while the amount of data written is less than 170 MB.
-This correlates with our finding that . However,
+ATLAS jobs are read heavy: On average, the amount of data read per worker node
+is less than 400 MB, while the amount of data written is less than 170 MB.
 % as can be seen in Table~\ref{panda-olcf-stats} and figures
 % {fig:atlas-titan-io-read} and {fig:atlas-titan-io-written},
-the distributions of read and written data are quite different: The read
-operation distribution per job shows a long tail, ranging from 12.5 GB to 250
-GB, while the written amount of data has a very narrow distribution.\mtnote{Should we say why?}
+Distributions of read and written data are different: The read operation
+distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
+the written amount of data has a very narrow distribution.
 
@@
 % on average per second per job.
-For the 1,175 ATLAS PanDA jobs observed, the total number of file $open()$
-operations is 172,089,760 and the total number of file $close()$ operations is
-40,132,992. The difference between these two values is still under
-investigation: One possible explanation is that ATLAS PanDA jobs don't call a
-file $close()$ operation per every file $open()$ issued.
+For the 1,175 ATLAS jobs observed, the total number of file $open()$ operations
+is 172,089,760 and the total number of file $close()$ operations is 40,132,992.
+The difference between these two values is still under investigation: One
+possible explanation is that ATLAS jobs don't call a file $close()$ operation
+per every file $open()$ issued.
 
-Overall, based on our experiments with ATLAS jobs, it can be safely concluded
-that the file and metadata I/O load of ATLAS PanDA project on the OLCF Titan
-supercomputing environment and the Spider 2 file system is not detrimental to
-the center operations. At the current scale of the project, the overall impact
-of ATLAS operations on Titan is minimal.
+% based on our experiments with ATLAS jobs, it can be safely concluded
+% that the file and metadata I/O load of ATLAS PanDA project on the OLCF Titan
+% supercomputing environment and the Spider 2 file system is not detrimental to
+% the center operations. At the current scale of the project, the overall impact
+% of ATLAS operations on Titan is minimal.
+
+Overall, the average time taken to read events from input files stored on Lustre is 1,320, comparable to the time taken to read the file required by assembling AthenaMP from NFS.
+Preliminary investigation shows that this time could be reduced to 40 seconds by loading the event descriptions into the RAM disk available on each worker node.
+Events descriptions could be transferred from Lustre to the RAM disk while configuring and initializing AthenaMP, almost halving the time currently required by initiating a Genat4 simulation.
+
+The current design and architecture of the PanDA Broker is proving to be as
+reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
+the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
+while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
+Brokers were responsible for around the 19\% of the failures, compared to the
+29\% of failures produced by the JobDispatcher module of the PanDA Server, and
+the 13\% failures produced by the Geant4 toolkit. The current failure rate of
+the PanDA Brokers confirms the benefits of reusing most of the code base of the
+PanDA Pilot for implementing the PanDA Broker. It also shows that adopting
+third-party libraries like RADICAL-SAGA did not have a measurable adverse effect
+on reliability.

### 2017-04-03 PandaEvolution.tex
checkpoint.

--- 2017-04-02 PandaEvolution.tex
@@
 
-Currently, very few HEP applications could benefit from Titan's GPUs but some
+Currently, very few HEP applications can benefit from Titan's GPUs but some
 computationally-intensive and non memory-intensive tasks of ATLAS workflows can

### 2017-04-03 nge.tex

@@
 what we used in the previous section(s) to designate what PanDA Broker submits
-to Titan's Torque batch system. Job is already used for what is pulled from
+to Titan's PBS batch system. Job is already used for what is pulled from
 PanDA server.} Multiple generations would also relax two assumptions of the
@@
 (Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
-and a UnitManager.   The PilotManager submits pilots to Titan's Torque
-\jhanote{for consistency should we say Moab?}\mtnote{Moab = Scheduler; Torque =
-Batch system.} batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash
-arrow), as done by PanDA Broker to submit MPI scripts. Once scheduled, the
-pilot's Agent is bootstrapped on Titan's MOM node and \jhanote{the Agent's
-Executors on the} the Agent's Executors on the worker nodes, and the
-UnitManager schedules units to the Agent's Scheduler (Fig.~\ref{fig:arch-
-overview}, solid arrow). The Agent's Scheduler schedules the units on one or
-more Agent's Executor for execution. The Agent's executors can manage one or
-more worker nodes, depending on performance evaluations. The UnitManager and the
-Agent communicate via a database that is instantiated on one of Titan's DTN so
-as to be reachable by both modules.
+and a UnitManager.   The PilotManager submits pilots to Titan's PBS \jhanote{for
+consistency should we say Moab?}\mtnote{Moab = Scheduler; Torque = Batch system.
+I used PBS Batch System because it makes easier to speak about both the system
+and the script submitted to it.} batch system via SAGA API
+(Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit
+MPI scripts. Once scheduled, the pilot's Agent is bootstrapped on Titan's MOM
+node and \jhanote{the Agent's Executors on the} the Agent's Executors on the
+worker nodes, and the UnitManager schedules units to the Agent's Scheduler
+(Fig.~\ref{fig:arch- overview}, solid arrow). The Agent's Scheduler schedules
+the units on one or more Agent's Executor for execution. The Agent's executors
+can manage one or more worker nodes, depending on performance evaluations. The
+UnitManager and the Agent communicate via a database that is instantiated on one
+of Titan's DTN so as to be reachable by both modules.
 

### 2017-04-03 PandaEvolution.tex
minor corrections and comments in Section 3.

@@
 
-Titan's architecture, configuration and policies poses several challenges to the
-integration with PanDA. The deployment model of PanDA Pilot is unfeasible on
-Titan: PanDA Pilot requires to contact the Job Dispatcher of the PanDA Server to
-pull jobs to execute but this is not possible on Titan because worker nodes do
-not offer outbound network connectivity. Further, Titan does not support PanDA's
-security model based on certificates and virtual organizations, making the
-PanDA's approach to identity management also unfeasible. While DTNs offer wide
-area network data transfer, an integration with ATLAS DDM is beyond the
-functional and administrative scope of the current prototyping phase. Finally,
-the specific characteristics of the execution environment, especially the
-absence of local storage on the worker nodes and modules tailored to Compute
-Node Linux, require reengineering of ATLAS application frameworks.
+Titan's architecture, configuration and policies poses several challenges to
+the integration with PanDA. The \jhanote{default?} deployment model of PanDA
+Pilot is unfeasible on Titan: PanDA Pilot is required to contact the Job
+Dispatcher of the PanDA Server to pull jobs to execute, but this is not
+possible on Titan because worker nodes do not offer outbound network
+connectivity. Further, Titan does not support PanDA's security model based on
+certificates and virtual organizations, making PanDA's approach to
+identity management also unfeasible. While DTNs offer wide area network data
+transfer, an integration with ATLAS DDM is beyond the functional and
+administrative scope of the current prototyping phase. Finally, the specific
+characteristics of the execution environment, especially the absence of local
+storage on the worker nodes and modules tailored to Compute Node Linux,
+require re-engineering of ATLAS application frameworks.
 
@@
 computationally-intensive and non memory-intensive tasks of ATLAS workflows can
-be offloaded from the Grid to Titan's large amount of cores. Further, when HEP
-tasks can be partitioned to independent jobs, Titan worker nodes can be used to
-execute up to 16 concurrent payloads, one for each available core. Given these
+be off-loaded from the Grid to Titan's large amount of cores. Further, when HEP
+tasks can be partitioned into independent jobs, Titan worker nodes can be used to
+execute up to 16 concurrent payloads, one per each available core. Given these
 constraints and challenges, the type of task most suitable for execution at the
@@
 Each interaction creates a so-called hit and all hits are collected and passed
-on for digitalization, where hits are further process to mimic the readout of
+on for digitalization, where hits are further processed to mimic the readout of
 the detector. Finally, reconstruction operates local pattern recognition,
@@
     the execution process of a detector simulation job, part of the production
-    ATLAS MD workflow.}
+    ATLAS MD workflow. \jhanote{minor error in figure: "ATLAS MC" not "MD"}}
 \label{fig:panda_broker}

### 2017-04-03 PandaOverview.tex
Iteration 4 and scattered editing

@@
 \begin{figure}
-  \begin{center}
+    \centering
     \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
-  \end{center}
-    \caption{PanDA architecture with the subsystems and components
-    relevant to the integration of PanDA with supercomputers. Numbers indicates
-    the execution process based on JEDI: From the submission of tasks (1) to the
+    \caption{PanDA architecture with the subsystems and components relevant to
+    the integration of PanDA with supercomputers. Numbers indicates the
+    execution process based on JEDI: From the submission of tasks (1) to the
     retrieval of their output (19). The monitoring subsystem, the architectural
@@
 executed. Task Buffer checks the global queue (i.e., the PanDA database) and,
-upon availability, returns a suitable job the Job Dispatcher. \jhanote{end of previous sentence needs some attention} The Job Dispatcher
+upon availability, returns a suitable job to the Job Dispatcher. \jhanote{end of
+previous sentence needs some attention}\mtnote{Done.} The Job Dispatcher
 dispatches that job to the PanDA Pilot (Fig.~\ref{fig:architecture}:12).

### 2017-04-03 PandaEvolution.tex
Address Shantenu's comment

@@
 
-Titan's architecture, configuration and policies poses several challenges to
-the integration with PanDA. The \jhanote{default?} deployment model of PanDA
-Pilot is unfeasible on Titan: PanDA Pilot is required to contact the Job
-Dispatcher of the PanDA Server to pull jobs to execute, but this is not
+Titan's architecture, configuration and policies poses several challenges to the
+integration with PanDA. The \jhanote{default?}\mtnote{Done.} default deployment
+model of PanDA Pilot is unfeasible on Titan: PanDA Pilot is required to contact
+the Job Dispatcher of the PanDA Server to pull jobs to execute, but this is not
 possible on Titan because worker nodes do not offer outbound network
 connectivity. Further, Titan does not support PanDA's security model based on
-certificates and virtual organizations, making PanDA's approach to
-identity management also unfeasible. While DTNs offer wide area network data
-transfer, an integration with ATLAS DDM is beyond the functional and
-administrative scope of the current prototyping phase. Finally, the specific
-characteristics of the execution environment, especially the absence of local
-storage on the worker nodes and modules tailored to Compute Node Linux,
-require re-engineering of ATLAS application frameworks.
+certificates and virtual organizations, making PanDA's approach to identity
+management also unfeasible. While DTNs offer wide area network data transfer, an
+integration with ATLAS DDM is beyond the functional and administrative scope of
+the current prototyping phase. Finally, the specific characteristics of the
+execution environment, especially the absence of local storage on the worker
+nodes and modules tailored to Compute Node Linux, require re-engineering of
+ATLAS application frameworks.
 
@@
     the execution process of a detector simulation job, part of the production
-    ATLAS MD workflow. \jhanote{minor error in figure: "ATLAS MC" not "MD"}}
+    ATLAS MD workflow. \jhanote{minor error in figure: "ATLAS MC" not "MD"}\mtnote{Fixed.}}
 \label{fig:panda_broker}

### 2017-04-03 overall.tex
Address Jack's issue #11

@@
 
-Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of Titan
+Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of
 core-hours used by ATLAS during the experiment time window. ATLAS consumed a
-total of 73.8M Titan core-hours, for an average of $\approx$7M Titan core-hours
-a month, with a minimum of 3.3M Titan core-hours in April 2016 and a maximum
-14.8M Titan core-hours in February 2017.
+total of 73.8M core-hours, for an average of $\approx$7M core-hours a month,
+with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours
+in February 2017.
 
@@
     \caption{Gray bars: Number of unused cores of Titan's backfill availability;
-    green bars: Number of cores of backfill availability used by ATLAS via PanDA
-    Brokers on Titan; red labels: Efficiency of PanDA Brokers on Titan defined
-    as percentage of backfill availability used every month of the experiment
-    window.}
+    blue bars: Number of cores of Titan's backfill availability used by ATLAS
+    via PanDA Brokers; red labels: Efficiency of PanDA Brokers defined as
+    percentage of Titan's backfill availability used by ATLAS.}
 \label{fig:backfill-utilization}

### 2017-04-03 overall.tex
Fix some more numbers based on the new measure of number of cores

@@
 
-We evaluate the efficiency, scalability and reliability of the deployment of PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and AthenaMP.
+We evaluate the efficiency, scalability and reliability of the deployment of
+PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and
+AthenaMP. All the measurements were performed between January 2016 and February
+2017, hereafter called `experiment time window'.
+
 % two dimensions: (i) backfill utilization by the PanDA Brokers; and (ii) the
 % amount of runtime spent performing detector simulations.
-We based our characterizations on measuring the number of cores utilized by
-ATLAS on Titan, the overall backfill availability, the number of detector
-simulations performed on all the resources available to ATLAS and on Titan, and
-the failure rate of these simulations. We also consider the the time taken to
-setup AthenaMP and to simulate events in the ATLAS detector with it. All the
-measurements were performed between January 2016 and February 2017, hereafter
-called `experiment time window'.
 
-\mtnote{TODO: expand the previous paragraph to include dimensions of AthenaMP characterization.}
+% We based our characterizations on measuring the number of cores utilized by
+% ATLAS on Titan, the overall backfill availability, the number of detector
+% simulations performed on all the resources available to ATLAS and on Titan,
+% and the failure rate of these simulations. We also consider the the time taken
+% to setup AthenaMP and to simulate events in the ATLAS detector with it. All
+% the measurements were performed between January 2016 and February 2017,
+% hereafter called `experiment time window'.
+
+% \mtnote{TODO: expand the previous paragraph to include dimensions of AthenaMP
+% characterization.}
 
@@
 
-Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of
-core-hours used by ATLAS during the experiment time window. ATLAS consumed a
-total of 73.8M core-hours, for an average of $\approx$7M core-hours a month,
-with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours
-in February 2017.
+% Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of
+% core-hours used by ATLAS during the experiment time window. ATLAS consumed a
+% total of 73.8M core-hours, for an average of $\approx$7M core-hours a month,
+% with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours
+% in February 2017.
+
+Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of core-hours
+used by ATLAS during the experiment time window. ATLAS consumed a total of 39.3M
+core-hours, for an average of $\approx$3.7M core-hours a month, with a minimum
+of 1.7M core-hours in April 2016 and a maximum 7.8M core-hours in February 2017.
 
@@
 
-Figure~\ref{fig:backfill-utilization} (green bars) shows backfill utilization
+Figure~\ref{fig:backfill-utilization} (blue bars) shows backfill utilization
 during the experiment time window. Efficiency
@@
 with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\% efficiency on
-February 2017. The total backfill available was 38.1M Titan core-hours in April
-2016, and 33.1M Titan core-hours in February 2017. This shows that the measured
+February 2017.
+% The total backfill available was 38.1M Titan core-hours in April 2016, and
+% 33.1M Titan core-hours in February 2017. This shows that the measured
+% difference in efficiency did not depend on a comparable difference in total
+% backfill availability.
+The total backfill available was 20.3M Titan core-hours in April
+2016, and 17.6M Titan core-hours in February 2017. This shows that the measured
 difference in efficiency did not depend on a comparable difference in total
@@
 
-\mtnote{Elaborate further and put it at the end, after giving several of the
-measurements we would need. ``Currently, this measure of efficiency is being
-refined on the base of the actual minimum amount of walltime that PanDA Broker
-can request when submitting job requests. This is not a fixed value as it
-depends on variable initialization and configuration overheads, simulation time,
-and number of nodes obtained. Part of the undergoing effort is to enable the
-measurements of  the distribution of these parameters to derive an optimization
-function.''}
+% \mtnote{Elaborate further and put it at the end, after giving several of the
+% measurements we would need. ``Currently, this measure of efficiency is being
+% refined on the base of the actual minimum amount of walltime that PanDA Broker
+% can request when submitting job requests. This is not a fixed value as it
+% depends on variable initialization and configuration overheads, simulation time,
+% and number of nodes obtained. Part of the undergoing effort is to enable the
+% measurements of  the distribution of these parameters to derive an optimization
+% function.''}
 
-\jhanote{I would plot the efficiency on the Y2 axis for both
-[diagrams].}\mtnote{Done but then the first diagram ended up plotting a subset
-of the data of the second diagram. I used only the second diagram.} \jhanote{I
-would suggest similar style of X-axis as previous figure. Difficult to parse
-which is April if we are going to call out a specific month. Also consistency in
-style is generally good to keep cognitive burden low.}\mtnote{Done.}
+% \jhanote{I would plot the efficiency on the Y2 axis for both
+% [diagrams].}\mtnote{Done but then the first diagram ended up plotting a subset
+% of the data of the second diagram. I used only the second diagram.} \jhanote{I
+% would suggest similar style of X-axis as previous figure. Difficult to parse
+% which is April if we are going to call out a specific month. Also consistency
+% in style is generally good to keep cognitive burden low.}\mtnote{Done.}
 
@@
 of time, and 3.5\% of the 6.6B events processed by those jobs. Comparatively,
-Titan contributed 3.9\% of the total of around 200K cores available to ATLAS on
-Grid, Cloud, and HPC infrastructures together. These figures confirms the
-relevance of supercomputers' resource contribution to the LHC Run 2, especially
-when accounting for the amount of unused backfill availability and the
-improvement of PanDA efficiency across the experiment time window.
+Titan contributed 3.9\% of the total of around 200K\mtnote{TODO: Fix this
+number} cores available to ATLAS on Grid, Cloud, and HPC infrastructures
+together. These figures confirms the relevance of supercomputers' resource
+contribution to the LHC Run 2, especially when accounting for the amount of
+unused backfill availability and the improvement of PanDA efficiency across the
+experiment time window.
 

### 2017-04-03 overall.tex
Address Jack's issue #14
Typo
Address Jack's issue #13

@@
 0.9\% of all the 250M detector simulations performed by ATLAS in the same period
-of time, and 3.5\% of the 6.6B events processed by those jobs. Comparatively,
-Titan contributed 3.9\% of the total of around 200K\mtnote{TODO: Fix this
-number} cores available to ATLAS on Grid, Cloud, and HPC infrastructures
-together. These figures confirms the relevance of supercomputers' resource
-contribution to the LHC Run 2, especially when accounting for the amount of
-unused backfill availability and the improvement of PanDA efficiency across the
-experiment time window.
+of time, and 3.5\% of the 6.6B events processed by those jobs.
+% Comparatively, Titan contributed 3.9\% of the total of around
+% 200K\mtnote{TODO: Fix this number} cores available to ATLAS on Grid, Cloud,
+% and HPC infrastructures together.
+These figures confirms the relevance of supercomputers' resource contribution to
+the LHC Run 2, especially when accounting for the amount of unused backfill
+availability and the improvement of PanDA efficiency across the experiment time
+window.
 
@@
 such, further hardware upgrades seem unlikely to improve significantly the
-performance of PanDA Brokers. Nonetheless, the current load suggests that the
-number of brokers per DTN could be increased. This would enable the submission
-of a larger number of concurrent jobs to Titan's PBS queue, allowing for PanDA
-to consume a higher percentage of backfill availability.
+performance of PanDA Brokers. Nonetheless, current DTN load suggests that the
+number of brokers per DTN could be further increased. This would enable the
+submission of a larger number of concurrent jobs to Titan's PBS queue, allowing
+for PanDA to consume a higher percentage of backfill availability.
+
+It should be noted that increasing the number of brokers requires also to
+increase the number of concurrent jobs that can be submitted to Titan's queue.
+Each broker submits 1 job at a time with at least 2 hours walltime. Backfill
+availability is polled every 10 minutes, giving a theoretical maximum of 110
+concurrent jobs submitted to Titan. Titan's policy was already changed in
+September 2016, enabling the submission of 20 concurrent jobs and the deployment
+of 20 PanDA brokers. Further study is required to understand the impact that a
+larger amount of concurrent queued jobs would have on the overall efficiency of
+Titan's scheduler.
 
@@
 mean of $\approx$14 minutes. Considering that each worker node process up to 16
-events concurrently, 100 events takes around 2 hours to process.
+events concurrently, 100 events takes around 2 hours to process. As such, PanDA brokers do not use backfill availability with less than 2 hours walltime.
 

### 2017-04-03 nge.tex
Section 5.0 - refinement.

--- 2017-04-02 nge.tex
@@
 
-\jhanote{not sure MPI script is optimal description or ambiguous.}\mtnote{Is
-what we used in the previous section(s) to designate what PanDA Broker submits
-to Titan's Torque batch system. Job is already used for what is pulled from
-PanDA server.} Multiple generations would also relax two assumptions of the
-current execution model: knowing the number of simulations before submitting the
-MPI script, and having a fixed number of events per simulation (100 at the
-moment). Pilots would enable the scheduling of simulations independently from
-whether they were available at the moment of submitting the pilot. Further,
-simulations with a varying number of events could be scheduled on a pilot,
-depending on the amount of remaining walltime and the distribution of execution
-time per event, as shown in \S\ref{ssec:panda_titan},
-Figure~\ref{fig:100event-distrib}. These capabilities would increase the
-efficiency of the PanDA Broker when there is a large difference between the
-number of cores and walltime.
+% \jhanote{not sure MPI script is optimal description or ambiguous.}\mtnote{Is
+% what we used in the previous section(s) to designate what PanDA Broker submits
+% to Titan's PBS batch system. Job is already used for what is pulled from
+% PanDA server.}
+
+Multiple generations would also relax two assumptions of the current execution
+model: knowing the number of simulations before submitting the MPI script, and
+having a fixed number of events per simulation (100 at the moment). Pilots
+would enable the scheduling of simulations independently from whether they
+were available at the moment of submitting the pilot. Further, simulations
+with a varying number of events could be scheduled on a pilot, depending on
+the amount of remaining walltime and the distribution of execution time per
+event, as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event-
+distrib}. These capabilities would increase the efficiency of the PanDA Broker
+when there is a large difference between the number of cores and walltime.
 
@@
 also be MPI-independent, as MPI is used for coordination among multiple worker
-nodes, not by the payload. \jhanote{candidate for compression/contraction into
-earlier paragraph(s)}
+nodes, not by the payload.
 
@@
 
-\jhanote{Alternative: "NGE is a runtime system to execute heterogeneous and
-dynamically determined tasks that constitute workloads." This definition put the
-focus on tasks and not workloads, which is a level above in the RADICAL
-conceptual layer diagram.}\mtnote{Done.} NGE is a runtime system to execute
-heterogeneous and dynamically determined tasks that constitute workloads.
-Fig.~\ref{fig:arch- overview} illustrates its current architecture as deployed
-on Titan: the two management modules (Pilot and Unit) represent a simplified
-version of the PanDA Broker while the agent module is the pilot submitted to
-Titan and executed on its worker nodes. The communication between PanDA Broker
-and PanDA Server is abstracted away as it is not immediately useful to evaluate
-the performance and capabilities of a pilot on Titan.
+NGE is a runtime system to execute heterogeneous and dynamically determined
+tasks that constitute workloads. Fig.~\ref{fig:arch- overview} illustrates its
+current architecture as deployed on Titan: the two management modules (Pilot
+and Unit) represent a simplified version of the PanDA Broker while the agent
+module is the pilot submitted to Titan and executed on its worker nodes. The
+communication between PanDA Broker and PanDA Server is abstracted away as it
+is not immediately useful to evaluate the performance and capabilities of a
+pilot on Titan.
 
@@
 
-NGE exposes an API for the application  \jhanote{workload management?} layer to
-describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
+NGE exposes an API for the application  \jhanote{workload management?} layer
+to describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
 (Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
-and a UnitManager.   The PilotManager submits pilots to Titan's Torque
-\jhanote{for consistency should we say Moab?}\mtnote{Moab = Scheduler; Torque =
-Batch system.} batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash
-arrow), as done by PanDA Broker to submit MPI scripts. Once scheduled, the
-pilot's Agent is bootstrapped on Titan's MOM node and \jhanote{the Agent's
-Executors on the} the Agent's Executors on the worker nodes, and the
-UnitManager schedules units to the Agent's Scheduler (Fig.~\ref{fig:arch-
-overview}, solid arrow). The Agent's Scheduler schedules the units on one or
-more Agent's Executor for execution. The Agent's executors can manage one or
-more worker nodes, depending on performance evaluations. The UnitManager and the
-Agent communicate via a database that is instantiated on one of Titan's DTN so
-as to be reachable by both modules.
+and a UnitManager.   The PilotManager submits pilots to Titan's PBS  batch
+system via SAGA API (Fig.~\ref{fig:arch-overview}, dash arrow), as done by
+PanDA Broker to submit MPI scripts. Once scheduled, the pilot's Agent is
+bootstrapped on Titan's MOM node and \jhanote{the Agent's Executors on the}
+the Agent's Executors on the worker nodes, and the UnitManager schedules units
+to the Agent's Scheduler (Fig.~\ref{fig:arch- overview}, solid arrow). The
+Agent's Scheduler schedules the units on one or more Agent's Executor for
+execution. The Agent's executors can manage one or more worker nodes,
+depending on performance evaluations. The UnitManager and the Agent
+communicate via a database that is instantiated on one of Titan's DTN so as to
+be reachable by both modules.
+
+% \jhanote{for consistency should we say Moab?}\mtnote{Moab = Scheduler; Torque
+% = Batch system. I used PBS Batch System because it makes easier to speak about
+% both the system and the script submitted to it.}
 
@@
 % Andre.}
-\jhanote{In the interest of space and clarity, I would  remove the last
-sentence of this para. And merge the first sentence with the para
-above.}\mtnote{Done.}
+% \jhanote{In the interest of space and clarity, I would  remove the last
+% sentence of this para. And merge the first sentence with the para
+% above.}\mtnote{Done.}
 

### 2017-04-03 overall.tex
Address Jack's issue #13

@@
 increase the number of concurrent jobs that can be submitted to Titan's queue.
-Each broker submits 1 job at a time with at least 2 hours walltime. Backfill
-availability is polled every 10 minutes, giving a theoretical maximum of 110
-concurrent jobs submitted to Titan. Titan's policy was already changed in
-September 2016, enabling the submission of 20 concurrent jobs and the deployment
-of 20 PanDA brokers. Further study is required to understand the impact that a
-larger amount of concurrent queued jobs would have on the overall efficiency of
-Titan's scheduler.
+Each broker submits 1 job at a time, asking for no more than 300 worker nodes
+and at least 2 hours walltime. When backfill availability exceeds 300 worker
+nodes, multiple brokers are used to submit multiple jobs. Backfill availability
+is polled every 10 minutes, giving a maximum of 11 concurrent sets of jobs
+submitted to Titan. Considering the variable amount of time spent by each job in
+Titan's queue, a large number of brokers and concurrent queued jobs may be
+required to use all the backfill availability.
+
+Titan's policy was already changed in September 2016, enabling the submission of
+20 concurrent jobs. As shown in Fig.~\ref{fig:backfill-utilization}, this
+improved the efficiency of PanDA brokers but further study is required to
+understand the impact that a larger amount of concurrent queued jobs would have
+on the overall efficiency of Titan's scheduler.
 

### 2017-04-03 PandaEvolution.tex
refine.

--- 2017-04-03 PandaEvolution.tex
@@
 
+\jhanote{Do you mean the top of section 3? Or do you mean the very beginning
+section?}
+
 % -----------------------------------------------------------------------------
@@
 
-Titan's architecture, configuration and policies poses several challenges to
-the integration with PanDA. The \jhanote{default?} deployment model of PanDA
-Pilot is unfeasible on Titan: PanDA Pilot is required to contact the Job
-Dispatcher of the PanDA Server to pull jobs to execute, but this is not
+Titan's architecture, configuration and policies poses several challenges to the
+integration with PanDA. The default deployment
+model of PanDA Pilot is unfeasible on Titan: PanDA Pilot is required to contact
+the Job Dispatcher of the PanDA Server to pull jobs to execute, but this is not
 possible on Titan because worker nodes do not offer outbound network
 connectivity. Further, Titan does not support PanDA's security model based on
-certificates and virtual organizations, making PanDA's approach to
-identity management also unfeasible. While DTNs offer wide area network data
-transfer, an integration with ATLAS DDM is beyond the functional and
-administrative scope of the current prototyping phase. Finally, the specific
-characteristics of the execution environment, especially the absence of local
-storage on the worker nodes and modules tailored to Compute Node Linux,
-require re-engineering of ATLAS application frameworks.
+certificates and virtual organizations, making PanDA's approach to identity
+management also unfeasible. While DTNs offer wide area network data transfer, an
+integration with ATLAS DDM is beyond the functional and administrative scope of
+the current prototyping phase. Finally, the specific characteristics of the
+execution environment, especially the absence of local storage on the worker
+nodes and modules tailored to Compute Node Linux, require re-engineering of
+ATLAS application frameworks.
 
@@
     the execution process of a detector simulation job, part of the production
-    ATLAS MD workflow. \jhanote{minor error in figure: "ATLAS MC" not "MD"}}
+    ATLAS MD workflow.}
 \label{fig:panda_broker}

### 2017-04-03 overall.tex
correlation --> scatter plot.

@@
     %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
-    \caption{Backfill availability: distribution of number of work nodes (left); distribution of walltime in minutes (center); and their correlation (right).
+    \caption{Backfill availability: distribution of number of work nodes (left); distribution of walltime in minutes (center); and a scatter plot of the two variables (right).
 62175 measures: mean number of work nodes available 691; mean

### 2017-04-03 Experiments.tex
removing addressed comments in 5.2

@@
 as tasks)  as the number of nodes controlled by the pilot. Each AthenaMP
-simulates 100 events, requiring $\sim 4200$ seconds on average. \jhanote{Please
-convert all measurements into seconds!}\aanote{Done}
+simulates 100 events, requiring $\sim 4200$ seconds on average.
 
@@
 pilot on the nodes; (ii) the UnitManager's dispatching of  units (tasks) to
-the agent; and (iii) time for the agent to initialize all the tasks on the nodes.
-\jhanote{"spawn" is the right word here. The agent does not spawn tasks. These
-are predefined. I would say "initialize" maybe? Something better?}\aanote{Done. I put initialize, maybe bootstrap would be better?}
+the agent; and (iii) time for the agent to bootstrap all the tasks on the nodes.
 
@@
 notice that NGE's overhead does grow linearly with the number of units.
-\jhanote{It does grow linearly with units! What am I missing??}\aanote{Corrected. I think that ``not'' came from a previous version of the sentence. }
 

### 2017-04-03 overall.tex
Address Jack's issue #13

--- 2017-04-03 overall.tex
@@
 performance of PanDA Brokers. Nonetheless, current DTN load suggests that the
-number of brokers per DTN could be further increased. This would enable the
-submission of a larger number of concurrent jobs to Titan's PBS queue, allowing
-for PanDA to consume a higher percentage of backfill availability.
-
-It should be noted that increasing the number of brokers requires also to
-increase the number of concurrent jobs that can be submitted to Titan's queue.
-Each broker submits 1 job at a time with at least 2 hours walltime. Backfill
-availability is polled every 10 minutes, giving a theoretical maximum of 110
-concurrent jobs submitted to Titan. Titan's policy was already changed in
-September 2016, enabling the submission of 20 concurrent jobs and the deployment
-of 20 PanDA brokers. Further study is required to understand the impact that a
-larger amount of concurrent queued jobs would have on the overall efficiency of
-Titan's scheduler.
+number of brokers per DTN could be increased. In September 2016, increasing the
+number of concurrent PanDA Broker from 4 to 20 markedly improved efficiency (see
+Fig.~\ref{fig:backfill-utilization}) but further research is needed to
+understand whether an even larger number brokers would allow PanDA to consume a
+higher percentage of backfill availability.
 

### 2017-04-03 nge.tex
minor fix.

@@
 
-NGE exposes an API for the application  \jhanote{workload management?} layer
-to describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
-(Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
-and a UnitManager.   The PilotManager submits pilots to Titan's PBS  batch
-system via SAGA API (Fig.~\ref{fig:arch-overview}, dash arrow), as done by
-PanDA Broker to submit MPI scripts. Once scheduled, the pilot's Agent is
-bootstrapped on Titan's MOM node and \jhanote{the Agent's Executors on the}
-the Agent's Executors on the worker nodes, and the UnitManager schedules units
-to the Agent's Scheduler (Fig.~\ref{fig:arch- overview}, solid arrow). The
-Agent's Scheduler schedules the units on one or more Agent's Executor for
-execution. The Agent's executors can manage one or more worker nodes,
-depending on performance evaluations. The UnitManager and the Agent
-communicate via a database that is instantiated on one of Titan's DTN so as to
-be reachable by both modules.
+NGE exposes an API to describe workloads (Fig.~\ref{fig:arch-overview}, green
+squares) and pilots (Fig.~\ref{fig:arch-overview}, red circles), and to
+instantiate a PilotManager and a UnitManager.   The PilotManager submits
+pilots to Titan's PBS  batch system via SAGA API (Fig.~\ref{fig:arch-
+overview}, dash arrow), as done by PanDA Broker to submit MPI scripts. Once
+scheduled, the pilot's Agent is bootstrapped on Titan's MOM node and
+\jhanote{the Agent's Executors on the} the Agent's Executors on the worker
+nodes, and the UnitManager schedules units to the Agent's Scheduler
+(Fig.~\ref{fig:arch- overview}, solid arrow). The Agent's Scheduler schedules
+the units on one or more Agent's Executor for execution. The Agent's executors
+can manage one or more worker nodes, depending on performance evaluations. The
+UnitManager and the Agent communicate via a database that is instantiated on
+one of Titan's DTN so as to be reachable by both modules.
 

### 2017-04-03 nge.tex
Comment section 5 addressed, ORTE citations added

@@
 NGE is a runtime system to execute heterogeneous and dynamically determined
-tasks that constitute workloads. Fig.~\ref{fig:arch- overview} illustrates its
+tasks that constitute workloads. Fig.~\ref{fig:arch-overview} illustrates its
 current architecture as deployed on Titan: the two management modules (Pilot
@@
 instantiate a PilotManager and a UnitManager.   The PilotManager submits
-pilots to Titan's PBS  batch system via SAGA API (Fig.~\ref{fig:arch-
-overview}, dash arrow), as done by PanDA Broker to submit MPI scripts. Once
+pilots to Titan's PBS  batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit MPI scripts. Once
 scheduled, the pilot's Agent is bootstrapped on Titan's MOM node and
-\jhanote{the Agent's Executors on the} the Agent's Executors on the worker
+the Agent's Executors on the on the worker
 nodes, and the UnitManager schedules units to the Agent's Scheduler
-(Fig.~\ref{fig:arch- overview}, solid arrow). The Agent's Scheduler schedules
+(Fig.~\ref{fig:arch-overview}, solid arrow). The Agent's Scheduler schedules
 the units on one or more Agent's Executor for execution. The Agent's executors
@@
 OpenMPI implementation, developed to support distributed high-performance
-computing applications operating in a heterogeneous environment~\cite{XXX}. The
-system \jhanote{which system? ORTE or OpenMPI in general?}\mtnote{I am not sure
-but I think ORTE?} transparently provides support for interprocess
-communication, resource discovery and allocation, and process launching across a
-variety of platforms. ORTE provides a mechanism similar to the Pilot concept -
+computing applications operating in a heterogeneous environment~\cite{castain05:_open_rte, cug-2016}.  
+ ORTE transparently provides support for interprocess communication, resource discovery and allocation, and process launching across a variety of platforms.
+ORTE provides a mechanism similar to the Pilot concept -
 it allows the user to create a \emph{``dynamic virtual machine''} (DVM) that

### 2017-04-03 conclusion.tex
added three references.

@@
 high-performance computing resources available to basic research.
+
+Impact of WMS on HEP foreshadows the importance of WMS for other experiments and
+their computational campaign.
+
+Titan has upwards of 93\%. Compare that with Blue Waters 80\% (XDMoD).
+
+Role of abstractions. Important/critical. As platforms available to science
+changes, both in terms of type of infrastructure (Summit) as well as Clouds.
+
+This work is a starting point and represents our initial experience. However, 
+this initial experience allows us to
+
+Sets the basis for more optimized and studied operation points.

### 2017-04-03 Experiments.tex
Comment section 5 addressed, ORTE citations added
Added references to Section 5.3 Figures in the appendix
Minor fix

@@
 
-\subsection{Experiments}
+\subsection{Experiments}\label{sec:ngeExp}
 
@@
     \caption{Weak scalability: average pilot duration, average  duration of a
-    single AthenaMP execution, and pilot's overhead as a function of different pilot sizes.
+    single AthenaMP execution, and pilot's overhead as a function of different pilot sizes (200, 500, 1000 and 2000 nodes).
 \jhanote{Caption should be in consistent order as the bars in the figure.
-Ideally pilot first then AthenMP.}}
+Ideally pilot first then AthenMP.}\aanote{DONE}}
 \label{fig:weakScal1a}

### 2017-04-03 overall.tex
Address backfill definition, Jack's tickets #9, #13, #14

@@
 
-Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of core-hours
-used by ATLAS during the experiment time window. ATLAS consumed a total of 39.3M
-core-hours, for an average of $\approx$3.7M core-hours a month, with a minimum
-of 1.7M core-hours in April 2016 and a maximum 7.8M core-hours in February 2017.
+We calculate the total amount of backfill availability over a period of time by:
+(i) polling backfill availability at regular intervals during that time window;
+(ii) converting the number of worker nodes available and their walltime into
+core-hours; (iii) summing the number of core-hours. We call this number of
+core-hours `total backfill availability'.
+
+Figure~\ref{fig:backfill-utilization} shows the total backfill availability on
+Titan (gray bars) and the number of core-hours of that availability used by
+ATLAS (blue bars) during the experiment time window. ATLAS consumed a total of
+51.4M core-hours, for an average of $\approx$3.7M core-hours a month, with a
+minimum of 1.7M core-hours in April 2016 and a maximum 7.9M core-hours in
+February 2017.
 
@@
     \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.pdf}
-    \caption{Gray bars: Number of unused cores of Titan's backfill availability;
-    blue bars: Number of cores of Titan's backfill availability used by ATLAS
-    via PanDA Brokers; red labels: Efficiency of PanDA Brokers defined as
-    percentage of Titan's backfill availability used by ATLAS.}
+
+    \caption{Titan's total backfill availability: CPU core-hours (gray) and CPU
+    core-hours used by ATLAS (blue). GPU core-hours unaccounted for as they
+    cannot be used by ATLAS. Efficiency of PanDA Brokers defined as percentage
+    of total Titan's backfill availability used by ATLAS (Red labels).}
+
 \label{fig:backfill-utilization}
@@
 
-Figure~\ref{fig:backfill-utilization} (blue bars) shows backfill utilization
-during the experiment time window. Efficiency
-(Figure~\ref{fig:backfill-utilization}, red labels) is defined as the fraction
-of Titan core-hours utilized by the PanDA Brokers (for ATLAS jobs) to the total
-of Titan’s backfill availability. The brokers reached 18\% average efficiency,
-with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\% efficiency on
-February 2017.
+PanDA Brokers' efficiency (Figure~\ref{fig:backfill-utilization}, red labels) is
+defined as the fraction of core-hours utilized by the PanDA Brokers to  Titan’s
+total backfill availability during the experiment time window. The brokers
+reached 18\% average efficiency, with a minimum 7.8\% efficiency on May 2016 and
+a maximum 30.9\% efficiency on February 2017.
 % The total backfill available was 38.1M Titan core-hours in April 2016, and
@@
 % backfill availability.
-The total backfill available was 20.3M Titan core-hours in April
-2016, and 17.6M Titan core-hours in February 2017. This shows that the measured
-difference in efficiency did not depend on a comparable difference in total
-backfill availability.
+The total backfill availability was 20.3M in April 2016, and 17.6M in February
+2017. This shows that the measured difference in efficiency did not depend on a
+comparable difference in the total backfill availability.
 
@@
 % can request when submitting job requests. This is not a fixed value as it
-% depends on variable initialization and configuration overheads, simulation time,
-% and number of nodes obtained. Part of the undergoing effort is to enable the
-% measurements of  the distribution of these parameters to derive an optimization
-% function.''}
+% depends on variable initialization and configuration overheads, simulation
+% time, and number of nodes obtained. Part of the undergoing effort is to
+% enable the measurements of  the distribution of these parameters to derive an
+% optimization function.''}
 
@@
 These figures confirms the relevance of supercomputers' resource contribution to
-the LHC Run 2, especially when accounting for the amount of unused backfill
-availability and the improvement of PanDA efficiency across the experiment time
-window.
+the LHC Run 2, especially when accounting for the amount of unused total
+backfill availability and the improvement of PanDA efficiency across the
+experiment time window.
 
@@
 
-On February 2017, PanDA Brokers used almost twice as much backfill availability
-than in any other month. No relevant code update was made during that period and
-logs indicated that the brokers were able to respond more promptly to backfill
-availability. This is likely due to hardware upgrades on the DTNs. The absence
-of continuous monitoring of those nodes does not allow to quantify bottlenecks
-but spot measurements of their load indicate that a faster CPU and better
-networking were likely responsible for the improved performance.
-
+On February 2017, PanDA Brokers used almost twice as much total backfill
+availability than in any other month (preliminary results for March 2017
+displayed in Fig.~\ref{fig:backfill-utilization} confirm this trend). No
+relevant code update was made during that period and logs indicated that the
+brokers were able to perform faster. This is likely due to hardware upgrades on
+the DTNs. The absence of continuous monitoring of those nodes does not allow to
+quantify bottlenecks but spot measurements of their load indicate that a faster
+CPU and better networking were likely responsible for the improved performance.
 Investigations showed an average CPU load of 3.6\% on the upgraded DTNs. As
 such, further hardware upgrades seem unlikely to improve significantly the
-performance of PanDA Brokers. Nonetheless, current DTN load suggests that the
-number of brokers per DTN could be increased. In September 2016, increasing the
-number of concurrent PanDA Broker from 4 to 20 markedly improved efficiency (see
-Fig.~\ref{fig:backfill-utilization}) but further research is needed to
-understand whether an even larger number brokers would allow PanDA to consume a
-higher percentage of backfill availability.
+performance of PanDA Brokers.
 
@@
 duration of backfill availability. The duration of a detector simulation is a
-function of the number of events: the more events, the longer the simulation.
-Not all events take the same time to be simulated:
+function of the number of events simulated but not all events take the same time
+to be simulated.
 % Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time
 % of
-1 event takes from $\approx$2 to $\approx$40 minutes to be simulated, with a
-mean of $\approx$14 minutes. Considering that each worker node process up to 16
-events concurrently, 100 events takes around 2 hours to process. As such, PanDA brokers do not use backfill availability with less than 2 hours walltime.
+1 event simulation takes from $\approx$2 to $\approx$40 minutes, with a mean of
+$\approx$14 minutes. Considering that each worker node process up to 16 events
+concurrently, 100 events takes an average of 105 minutes to process. As such,
+PanDA brokers do not use backfill availability with less than 105 minutes
+walltime.
 
@@
     %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
-    \caption{Backfill availability: distribution of number of work nodes (left); distribution of walltime in minutes (center); and a scatter plot of the two variables (right).
-62175 measures: mean number of work nodes available 691; mean
-    walltime available 126 minutes.}
+    \caption{Backfill availability: distribution of number of work nodes (left);
+    distribution of walltime in minutes (center); and a scatter plot of the two
+    variables (right). 62175 measures: mean number of work nodes available 691;
+    mean walltime available 126 minutes.}
 \label{fig:backfill-distrib}
@@
 of events than those on Titan. This explains the difference between the
-percentage of detector simulations (9\%) and of events computed by those
-simulations (3.9\%) on Titan. PanDA Broker could fit the number of events to the
-walltime of the backfill availability on the base of the distributions of
+percentage of detector simulations and of events computed by those simulations
+on Titan. PanDA Broker could fit the number of events to the walltime of the
+backfill availability on the base of the distributions of
 Fig.~\ref{fig:backfill-distrib}. That specific number of event could then be
@@
 process is spawn at a time to execute the payload. As a consequence, the
-utilization of a larger portion of Titan's backfill availability depends on the
-the number of concurrent PanDA Brokers instantiated on the DTNs: When all the 20
-PanDA Brokers have submitted a MPI PBS Job, further backfill availability cannot
-be used.
+utilization of a larger portion of Titan's total backfill availability depends
+on the the number of concurrent PanDA Brokers instantiated on the DTNs: When all
+the 20 PanDA Brokers have submitted a MPI PBS Job, further backfill availability
+cannot be used.
 
-Based on the current data, increasing PanDA Brokers efficiency on Titan would
-require around 60 concurrent brokers. Further testing and better monitoring
-capabilities are required to establish whether the DTN capabilities can support
-this load, especially when considering the volume of files staged in and out the
-DTNs. Alternatively, a design of a PanDA Broker capable of managing multiple MPI
-scripts at a time is being evaluated.
+In September 2016, increasing the number of concurrent PanDA Broker from 4 to 20
+markedly improved efficiency (see Fig.~\ref{fig:backfill-utilization}) but
+further research is undergoing to understand whether an even larger number
+brokers would yield similar results. This research focuses on evaluating the the
+overheads of input/output files staging, including its impact on DTNs, and on an
+alternative design of PanDA Broker to enable the concurrent submission of
+multiple MPI scripts~\cite{barreiro2016panda}.
 

### 2017-04-03 nge.tex

@@
 monitoring and managing of tasks, avoiding filesystem bottlenecks and race
-conditions with network sockets. As a consequence, it is able to minimize the
+conditions with network sockets. 
+\aanote{Requires validation: Alternative version of the three sentences above (too many ORTE in my opinion): 
+ORTE provides a mechanism similar to the Pilot concept - it allows the user to create a \emph{``dynamic virtual machine''} (DVM) that spans multiple nodes.
+The DVM transparently provides support for interprocess communication, resource discovery and allocation, and process launching across a variety of platforms.
+Libraries allow the interaction between users and ORTE to enable the submission,
+monitoring and managing of tasks, avoiding filesystem bottlenecks and race
+conditions with network sockets. 
+}
+
+As a consequence, ORTE is able to minimize the
 system overhead while submitting tasks.

### 2017-04-03 conclusion.tex
klunky first pass at conclusions.

--- 2017-03-11 conclusion.tex
@@
 high-performance computing resources available to basic research.
+
+HEP was amongst the first, if not the very first experimental community to
+realize the importance of using WMS to manage their computational campaign(s).
+As computing become increasingly critical for a range of other experiments,
+the impact of WMS on HEP foreshadows the importance of WMS for other
+experiments and their computational campaigns. Other experiments such as SKA,
+LSST etc.., will have their own workload characteristics, resources types and
+federation constraints, as well metrics of performance.  The design and
+experience captured in this paper should proof invaluable for computational
+campaigns that need to evaluate the relative merits of different approaches.
+
+Thanks to the efforts captured in this paper, the time-averaged utilization of
+Titan has consistently been upwards of 90\%. Compare this to the time averaged
+utilization of other similar leadership machines such as NSF's flagship Blue
+Waters, where the average utilization fluctuates between 60-80\% (XDMoD)
+machine~\cite{bw-sucks}. These impressive if not unprecedented operation gains
+aside, this work is just a starting point towards more effective operational
+models for future leadership and online analytical platforms\cite{foap-url},
+which will have to support ever increasing complex workloads with varying
+models for dynamic resource federation. Recent efforts with Google Cloud for
+the CMS experiment -- HEPCloud~\cite{hepcloud,google-hep} represent similar
+yet different approaches to dynamic resource federation.
+
+Our experience demonstrated the advantages arising from a  systems engineering
+and abstractions based approach, in particular the pilot abstraction. As the
+type, heterogeneity and complexity of future platforms  (e.g., Exascale,
+Commercial Clouds) for science increases, there will be a greater need and
+emphasis on abstractions to ensure WMS can both utilize as well as support
+platforms.

### 2017-04-03 introduction.tex
klunky first pass at conclusions.

@@
 workload management system with Titan ~\textemdash{} a DOE leadership
-computing facility. Consequently, Titan now analyzes up to 5\% of workload  of
-the ATLAS experiment. The case study presents the design and integration of
-PanDA to reach sustained production scales (ca. 100 M core-hours a years). It
-also discusses the investigation of a task execution runtime system on Titan,
-that is based on  the pilot-abstractions and which allows advanced execution
-modes of the ATLAS workload as well as the enhanced support for heterogeneous
-workloads (such as molecular dynamics).
+computing facility. Consequently, Titan now analyzes up to 5\% of the
+simulation workload  of the ATLAS experiment. The case study presents the
+design and integration of PanDA to reach sustained production scales (ca. 100
+M core-hours a years). It also discusses the investigation of a task execution
+runtime system on Titan, that is based on  the pilot-abstractions and which
+allows advanced execution modes of the ATLAS workload as well as the enhanced
+support for heterogeneous workloads (such as molecular dynamics).
 
@@
 community looks forward to designing the next generation of online analytical
-platforms~\cite{foap}, this project provides some early lessons and guidance
+platforms~\cite{foap-url}, this project provides some early lessons and guidance
 for how current and future experimental and observational systems can be

### 2017-04-03 PandaEvolution.tex

@@
 % of time by any of the jobs already queued on Titan: All queued jobs are either
-% too large or their walltime is too long. At any point in time, Titan’s Moab
-% scheduler can be queried for backfill availability. Based upon the result of
-% this query, a job can be shaped to request no more than the backfill
-% availability. As such, when submitted this job is usually scheduled
-% immediately, spending almost no time in the queue.
+% too large or their walltime is too long.
 
-Titan’s backfill functionality offers the opportunity to avoid the overhead of
-queue wait times without using pilot abstraction~\cite{maui_backfill_url}.
-Backfill availability is the number of node-hours (or core-hours) that cannot be
-utilized by regular jobs queued on Titan at a specific point in time.  It is
-important to note that backfill availability is computed retrospectively and is
-determined by Titan's operation team. Backfill availability represents the
-theoretical maximum of node-hours that could be utilized by ATLAS (or any other
-project) that submit under backfill mode.
+The backfill optimization of Titan's Moab scheduler offers the opportunity to
+avoid the overhead of queue wait times without using pilot
+abstraction~\cite{maui_backfill_url}. This optimization enables Moab to start
+low-priority jobs as long as they do not delay higher priority jobs,
+independently on whether the low-priority jobs were queued after the
+high-priority jobs.
 
-Compared to pilots, backfill has the disadvantage of limiting the amount of work
-nodes that can be requested. Pilots are normal jobs: they can request as many
-worker nodes for as much time as a queue can offer. On the contrary, jobs sized
-on the basis of backfill information availability depend on the number of worker
-nodes that cannot be given to any other job in the Titan's queue at that moment
-in time.
+When the backfill optimization is enabled, users can interrogate Moab about what
+number of worker nodes and walltime would be available to a low-priority job at
+that moment in time. If a job is immediately submitted to Titan with that number
+of worker nodes and walltime, chances are that Moab will immediately scheduled
+it, reducing its queue time to a minimum. In this paper, we call this number of
+worker nodes and walltime `backfill availability'.
+
+% It is important to note that backfill availability is computed retrospectively
+% and is determined by Titan's operation team. Backfill availability represents
+% the theoretical maximum of node-hours that could be utilized by ATLAS (or any
+% other project) that submit under backfill mode.
+
+Compared to pilots, backfill has the disadvantage of limiting the amount of
+worker nodes that can be requested. Pilots are normal jobs: they can request as
+many worker nodes for as much time as a queue can offer. On the contrary, jobs
+sized on the basis of backfill availability depend on the number of worker nodes
+that cannot be given to any other job in the Titan's queue at that moment in
+time.
 
@@
 
-Backfill availability is typically a small fraction of the total capacity  of a
-resource. Notwithstanding,  given the size of Titan this translates into a
-substantial capacity. Every year, about 10\% of Titan's capacity remains
-unused~\cite{barker2016us}, corresponding to an average of 30,000 unused cores.
-This equals approximately 270M core hours per year, roughly 10\% of the overall
-capacity of WLCG.
+At any point in time, backfill availability is typically a small fraction of the
+total capacity of a resource. Notwithstanding, given the size of Titan this
+translates into a substantial capacity. Every year, about 10\% of Titan's
+capacity remains unused~\cite{barker2016us}, corresponding to an average of
+30,000 unused cores (excluding GPU cores). This equals approximately 270M core
+hours per year, roughly 10\% of the overall capacity of WLCG.
 

### 2017-04-03 conclusion.tex
minor corrections and fixes in 5 and 7.

--- 2017-03-11 conclusion.tex
@@
 high-performance computing resources available to basic research.
+
+HEP was amongst the first, if not the very first experimental community to
+realize the importance of using WMS to manage their computational campaign(s).
+As computing become increasingly critical for a range of experiments, the
+impact of WMS on HEP foreshadows the importance of WMS for other experiments
+(such as SKA, LSST etc.) and their computational campaigns. These experiments
+will have their own workload characteristics, resources types and federation
+constraints, as well metrics of performance.  The design and experience
+captured in this paper should prove invaluable for designing computational
+campaigns and will provide a baseline to evaluate the relative merits of
+different approaches.
+
+Thanks to the efforts captured in this paper, the time-averaged utilization of
+Titan has consistently been upwards of 90\%. Compare this to the time averaged
+utilization of other similar leadership machines such as NSF's flagship Blue
+Waters, where the average utilization fluctuates between 60-80\% (XDMoD)
+machine~\cite{bw-sucks}. These impressive if not unprecedented efficiency
+gains aside, this work is just a starting point towards more effective
+operational models for future leadership and online analytical platforms\cite
+{foap-url}. These platforms will have to support ever increasing complex
+workloads with varying models for dynamic resource federation. Recent efforts
+with Google Cloud for the CMS experiment -- HEPCloud~\cite{hepcloud,google-
+hep} represent similar yet different approaches to dynamic resource
+federation.
+
+Our experience demonstrated the advantages arising from a  systems engineering
+and abstractions based approach, in particular the pilot abstraction. As the
+type, heterogeneity and complexity of future platforms  (e.g., Exascale,
+Commercial Clouds) for science increases, there will be a greater need and
+emphasis on abstractions to ensure WMS can both utilize as well as support
+platforms.

### 2017-04-03 bckground.tex
minor tweaks.

@@
-\jhanote{In this section we should talk less about PanDA and more about other
-WMS -- highlighting the fact that (i) there isn't a comprehensive systems
-(design, architecture and execution properties) discussion of other WMS so
-difficult to compare}\mtnote{Better?}
+% \jhanote{In this section we should talk less about PanDA and more about other
+% WMS -- highlighting the fact that (i) there isn't a comprehensive systems
+% (design, architecture and execution properties) discussion of other WMS so
+% difficult to compare}\mtnote{Better?}
+
 \jhanote{(ii) real comparison should be to the Gliden-in WMS ecosystem and use
@@
 over other solutions.}
-\jhanote{Also might this section be moved towards the rear, as this is  a
-practice paper and not a traditional research paper?}\mtnote{Done.}
+
+% \jhanote{Also might this section be moved towards the rear, as this is  a
+% practice paper and not a traditional research paper?}\mtnote{Done.}
 

### 2017-04-03 conclusion.tex
Iteration conclusions

--- 2017-04-03 conclusion.tex
@@
-The PanDA  system was developed to meet the scale and complexity of LHC
-distributed computing for the ATLAS experiment.  In the process,  the old batch
-job paradigm  of computing in HEP was  discarded  in favor of a  far more
-flexible and scalable  model. The success  of PanDA  at the LHC is leading to
-widespread adoption and testing by other experiments. PanDA  is the first
-exascale  workload management system in HEP, already operating at a million
-computing jobs per day, and processing over an exabyte of data in 2013. Next LHC
-run will pose massive computing  challenges. With a  doubling of the beam
-energy  and luminosity as  well as an increased  need for  simulates  data, the
-data volume is expected to increase with a factor 5--6 or more. Storing and
-processing  this amount of data is a  challenge   that cannot be resolved with
-the currently existing  computing  resources in ATLAS\@. To resolve this
-challenge, ATLAS is turning to commercial  as well as academic Cloud services
-and HPCs via the PanDA system. Also the work underway is enabling the use of
-PanDA by new scientific collaborations and communities as a means  of leveraging
-extreme scale computing  resources with a low barrier of entry. The technology
-base provided by the PanDA system will enhance the usage of a variety  of
-high-performance computing resources available to basic research.
+% PanDA WMS was developed to meet the scale and complexity of LHC
+% distributed computing for the ATLAS experiment. In the process, the old batch
+% job paradigm of computing in HEP was discarded in favor of a far more
+% flexible and scalable model. The success of PanDA at the LHC is leading to
+% widespread adoption and testing by other experiments. PanDA is the first
+% exascale workload management system in HEP, already operating at a million
+% computing jobs per day, and processing over an exabyte of data in 2013. Next
+% LHC run will pose massive computing challenges. With a doubling of the beam
+% energy and luminosity as well as an increased need for simulates data, the
+% data volume is expected to increase with a factor 5--6 or more. Storing and
+% processing this amount of data is a challenge that cannot be resolved with the
+% currently existing computing resources in ATLAS\@. To resolve this challenge,
+% ATLAS is turning to commercial as well as academic Cloud services and HPCs via
+% the PanDA system. Also the work underway is enabling the use of PanDA by new
+% scientific collaborations and communities as a means of leveraging extreme
+% scale computing resources with a low barrier of entry. The technology base
+% provided by the PanDA system will enhance the usage of a variety of
+% high-performance computing resources available to basic research.
 
-HEP was amongst the first, if not the very first experimental community to
-realize the importance of using WMS to manage their computational campaign(s).
-As computing become increasingly critical for a range of other experiments,
-the impact of WMS on HEP foreshadows the importance of WMS for other
-experiments and their computational campaigns. Other experiments such as SKA,
-LSST etc.., will have their own workload characteristics, resources types and
-federation constraints, as well metrics of performance.  The design and
-experience captured in this paper should proof invaluable for computational
-campaigns that need to evaluate the relative merits of different approaches.
+The deployment of PanDA Broker on Titan enabled distributed computing on a
+leadership-class HPC machine at unprecedented scale. In the past 13 months,
+PanDA WMS has consumed almost 52M core-hours on Titan, simulating 3.5\% of the
+total number of detector events of the ATLAS production Monte Carlo workflow. We
+described the implementation and execution process of PanDA WMS
+(\S\ref{sec:panda_overview}) and PanDA Broker (\S\ref{sec:panda_titan}), showing
+how they support and enable distributed computing at this scale on Titan, a
+leadership-class HPC machine managed by OCLF at ORNL.
+
+We characterized the state of practice by evaluating the efficiency, scalability
+and reliability of both PanDA Broker and AthenaMP as deployed on Titan
+(\S\ref{sec:panda_titan}). Our characterization highlighted both the strengths
+and limitations of the current design and implementation: PanDA Brokers enable
+the sustained execution of millions of simulations per week but further work is
+required to optimize its efficiency and reliability (\S\ref{ssec:broker_titan}).
+PanDA Brokers support the concurrent execution of multiple AthenaMP instances,
+enabling each AthenaMP to perform the concurrent execution of up to 16 Geant4
+simulators. Nonetheless, our characterization showed how improving I/O
+performance could reduce overheads (\S\ref{ssec:athenamp_titan}), increasing the
+overall utilization of Titan's backfill availability.
+
+More in general, this paper highlighted the fundamental role of WMS for
+experimental science. HEP was amongst the first, if not the very first
+experimental community to realize the importance of using WMS to manage their
+computational campaign(s). As computing become increasingly critical for a range
+of other experiments, the impact of WMS on HEP foreshadows the importance of WMS
+for other communities.
+% experiments and their computational campaigns.
+Other experiments such as SKA or LSST, will have their own workload
+characteristics, resources types and federation constraints, as well metrics of
+performance. The design and experience captured in this paper should prove
+invaluable for computational campaigns that need to evaluate the relative merits
+of different approaches.
 
@@
 Waters, where the average utilization fluctuates between 60-80\% (XDMoD)
-machine~\cite{bw-sucks}. These impressive if not unprecedented operation gains
-aside, this work is just a starting point towards more effective operational
-models for future leadership and online analytical platforms\cite{foap-url},
-which will have to support ever increasing complex workloads with varying
-models for dynamic resource federation. Recent efforts with Google Cloud for
-the CMS experiment -- HEPCloud~\cite{hepcloud,google-hep} represent similar
-yet different approaches to dynamic resource federation.
+machine~\cite{bw-sucks}. These % impressive if not
+unprecedented operation gains aside, this work is just a starting point towards
+more effective operational models for future leadership and online analytical
+platforms\cite{foap-url}, which will have to support ever increasing complex
+workloads with varying models for dynamic resource federation. Recent efforts
+with Google Cloud for the CMS experiment -- HEPCloud~\cite{hepcloud,google-hep}
+represent similar yet different approaches to dynamic resource federation.
 
-Our experience demonstrated the advantages arising from a  systems engineering
-and abstractions based approach, in particular the pilot abstraction. As the
-type, heterogeneity and complexity of future platforms  (e.g., Exascale,
-Commercial Clouds) for science increases, there will be a greater need and
-emphasis on abstractions to ensure WMS can both utilize as well as support
-platforms.
+Our experience demonstrated also the advantages arising from a systems
+engineering and abstractions based approach, in particular the pilot
+abstraction. As the type, heterogeneity and complexity of future platforms
+(e.g., Exascale, Commercial Clouds) for science increases, there will be a
+greater need and emphasis on abstractions to ensure WMS can both utilize as well
+as support platforms.

### 2017-04-03 Experiments.tex

@@
     \caption{Weak scalability: average pilot duration, average  duration of a
-    single AthenaMP execution, and pilot's overhead as a function of different pilot sizes (200, 500, 1000 and 2000 nodes).
-\jhanote{Caption should be in consistent order as the bars in the figure.
-Ideally pilot first then AthenMP.}\aanote{DONE}}
+    single AthenaMP execution, and pilot's overhead as a function of different pilot sizes (200, 500, 1000 and 2000 nodes).}
 \label{fig:weakScal1a}

### 2017-04-03 PandaEvolution.tex

@@
 supercomputer's queue, once scheduled, they can pull and execute jobs
-independently from the system's queue. Jobs can be concurrently executed on
+independent from the system's queue. Jobs can be concurrently executed on
 every core available to the pilot, and multiple generations of concurrent
@@
 low-priority jobs as long as they do not delay higher priority jobs,
-independently on whether the low-priority jobs were queued after the
+independent of whether the low-priority jobs were queued after the
 high-priority jobs.
 
-When the backfill optimization is enabled, users can interrogate Moab about what
-number of worker nodes and walltime would be available to a low-priority job at
+When the backfill optimization is enabled, users can interrogate Moab about the
+number of worker nodes and walltime that would be available to a low-priority job at
 that moment in time. If a job is immediately submitted to Titan with that number
-of worker nodes and walltime, chances are that Moab will immediately scheduled
+of worker nodes and walltime, chances are that Moab will immediately schedule
 it, reducing its queue time to a minimum. In this paper, we call this number of

### 2017-04-03 conclusion.tex

@@
 computational campaign(s). As computing become increasingly critical for a range
-of other experiments, the impact of WMS on HEP foreshadows the importance of WMS
-for other communities.
-% experiments and their computational campaigns.
-Other experiments such as SKA or LSST, will have their own workload
-characteristics, resources types and federation constraints, as well metrics of
-performance. The design and experience captured in this paper should prove
-invaluable for computational campaigns that need to evaluate the relative merits
-of different approaches.
+of experiments, the impact of WMS on HEP foreshadows the importance of WMS for
+other experiments (such as SKA, LSST etc.). % and their computational campaigns.
+These experiments will have their own workload characteristics, resources types
+and federation constraints, as well metrics of performance.  The design and
+experience captured in this paper should prove invaluable for designing
+computational campaigns and will provide a baseline to evaluate the relative
+merits of different approaches.
 
@@
 Waters, where the average utilization fluctuates between 60-80\% (XDMoD)
-machine~\cite{bw-sucks}. These % impressive if not
-unprecedented operation gains aside, this work is just a starting point towards
+machine~\cite{bw-sucks}. These
+% impressive if not
+unprecedented efficiency gains aside, this work is just a starting point towards
 more effective operational models for future leadership and online analytical
-platforms\cite{foap-url}, which will have to support ever increasing complex
-workloads with varying models for dynamic resource federation. Recent efforts
-with Google Cloud for the CMS experiment -- HEPCloud~\cite{hepcloud,google-hep}
-represent similar yet different approaches to dynamic resource federation.
+platforms~\cite{foap-url}. These platforms will have to support ever increasing
+complex workloads with varying models for dynamic resource federation. Recent
+efforts with Google Cloud for the CMS experiment --
+HEPCloud~\cite{hepcloud,google- hep} represent similar yet different approaches
+to dynamic resource federation.
 
-Our experience demonstrated also the advantages arising from a systems
-engineering and abstractions based approach, in particular the pilot
-abstraction. As the type, heterogeneity and complexity of future platforms
-(e.g., Exascale, Commercial Clouds) for science increases, there will be a
-greater need and emphasis on abstractions to ensure WMS can both utilize as well
-as support platforms.
+Our experience demonstrated the advantages arising from a  systems engineering
+and abstractions based approach, in particular the pilot abstraction. As the
+type, heterogeneity and complexity of future platforms  (e.g., Exascale,
+Commercial Clouds) for science increases, there will be a greater need and
+emphasis on abstractions to ensure WMS can both utilize as well as support
+platforms.

### 2017-04-03 PandaEvolution.tex
Iteration on blackfill: slots + total availability

@@
 When the backfill optimization is enabled, users can interrogate Moab about the
-number of worker nodes and walltime that would be available to a low-priority job at
-that moment in time. If a job is immediately submitted to Titan with that number
-of worker nodes and walltime, chances are that Moab will immediately schedule
-it, reducing its queue time to a minimum. In this paper, we call this number of
-worker nodes and walltime `backfill availability'.
+number of worker nodes and walltime that would be available to a low-priority
+job at that moment in time. If a job is immediately submitted to Titan with that
+number of worker nodes and walltime, chances are that Moab will immediately
+schedule it, reducing its queue time to a minimum. In this paper, we call this
+number of worker nodes and walltime an available `backfill slot'.
 
@@
 many worker nodes for as much time as a queue can offer. On the contrary, jobs
-sized on the basis of backfill availability depend on the number of worker nodes
-that cannot be given to any other job in the Titan's queue at that moment in
-time.
+sized on the basis of an available backfill slot depend on the number of worker
+nodes that cannot be given to any other job in the Titan's queue at that moment
+in time.
 
@@
 
-At any point in time, backfill availability is typically a small fraction of the
-total capacity of a resource. Notwithstanding, given the size of Titan this
-translates into a substantial capacity. Every year, about 10\% of Titan's
-capacity remains unused~\cite{barker2016us}, corresponding to an average of
-30,000 unused cores (excluding GPU cores). This equals approximately 270M core
-hours per year, roughly 10\% of the overall capacity of WLCG.
+At any point in time, the size of an available backfill slot is typically a
+small fraction of the total capacity of a resource. Notwithstanding, given the
+size of Titan this translates into a substantial capacity. Every year, about
+10\% of Titan's capacity remains unused~\cite{barker2016us}, corresponding to an
+average of 30,000 unused cores (excluding GPU cores). This equals
+% approximately 270M core-hours per year,
+to roughly 10\% of the overall capacity of WLCG.
 
@@
 system (Fig.~\ref{fig:panda_broker}:3). PanDA Broker queries Titan's Moab
-scheduler about current backfill availability (Fig.~\ref{fig:panda_broker}:2)
-and creates an MPI script, wrapping enough ATLAS jobs' payload to fit backfill
-availability. PanDA Broker submits the MPI script to the Titan's Torque batch
-system via RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).
+scheduler about the current available backfill slot
+(Fig.~\ref{fig:panda_broker}:2) and creates an MPI script, wrapping enough ATLAS
+jobs' payload to fit the backfill slot. PanDA Broker submits the MPI script to
+the Titan's Torque batch system via RADICAL-SAGA
+(Fig.~\ref{fig:panda_broker}:4).
 

### 2017-04-03 overall.tex
Space saving figures
Iteration conclusions
Iteration on blackfill: slots + total availability

@@
 \subsection{Characterizing the PanDA Broker on Titan}
-\label{ssec:panda_titan}
+\label{ssec:broker_titan}
 
@@
 We calculate the total amount of backfill availability over a period of time by:
-(i) polling backfill availability at regular intervals during that time window;
-(ii) converting the number of worker nodes available and their walltime into
-core-hours; (iii) summing the number of core-hours. We call this number of
+(i) polling the available backfill slots at regular intervals during that time
+window; (ii) converting the number of worker nodes available and their walltime
+into core-hours; (iii) summing the number of core-hours. We call this number of
 core-hours `total backfill availability'.
@@
 \subsection{Characterizing the Detector Simulation on Titan}
-\label{ssec:panda_titan}
+\label{ssec:athenamp_titan}
 
@@
 \begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison.pdf}
+    \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison_vsquashed.pdf}
     \caption{Comparison between distributions of the time taken by a Geant4

### 2017-04-04 conclusion.tex
fixing error.

@@
 efforts with Google Cloud for the CMS experiment --
-HEPCloud~\cite{hepcloud,google- hep} represent similar yet different approaches
+HEPCloud~\cite{hepcloud,googlehep} represent similar yet different approaches
 to dynamic resource federation.

### 2017-04-04 PandaEvolution.tex
Notes clean-up

@@
 ported also to other supercomputers, including the Ansel (?) at \ldots and
-\ldots at the National Energy Research Scientific Computing Center (NERSC).
+Cori at the National Energy Research Scientific Computing Center (NERSC).
 

### 2017-04-04 bckground.tex
Notes clean-up

@@
 
-\jhanote{(ii) real comparison should be to the Gliden-in WMS ecosystem and use
-this as opportunity to narrate the reasons for the divergence.}\mtnote{From our
-meetings with Sergey: Before Run 1, ATLAS management become uncomfortable with
-the pace at which other WMS were being developed, loosing trust in their ability
-to deliver production-grade capabilities. ATLAS USA decided to go solo and
-developed their own solution. PanDA proven to be able to deliver the
-usual pilot-based WMS capabilities at the right scale. For that, it was chosen
-over other solutions.}
+% \jhanote{(ii) real comparison should be to the Gliden-in WMS ecosystem and use
+% this as opportunity to narrate the reasons for the divergence.}\mtnote{From our
+% meetings with Sergey: Before Run 1, ATLAS management become uncomfortable with
+% the pace at which other WMS were being developed, loosing trust in their ability
+% to deliver production-grade capabilities. ATLAS USA decided to go solo and
+% developed their own solution. PanDA proven to be able to deliver the
+% usual pilot-based WMS capabilities at the right scale. For that, it was chosen
+% over other solutions.}
 

### 2017-04-04 PandaOverview.tex
remove addressed comment.
Notes clean-up

@@
 
-\mtnote{Addresses Alexei's comment. I eliminated `event' as abstraction and used
-datum/data/physical process when discussing data management in the whole
-section.}
+% \mtnote{Addresses Alexei's comment. I eliminated `event' as abstraction and used
+% datum/data/physical process when discussing data management in the whole
+% section.}
+
+% \mtnote{Addresses Alexei's comment. If this is still too soft, we can write:
+% ``but HEP applications do not require MPI.''}
 
 PanDA's application model assumes tasks, workloads and workflows. Tasks
-represent a set of operations performed on
-% a set of events that are
-data sets stored in one or more input files. Tasks are decomposed into jobs,
-where each job represents the task's set of operations and a partition of the
-task's
-% events
-data set. Since 2005, a certain amount of parallelism has been
-progressively introduced for job execution~\cite{crooks2012multi} but, so far,
-HEP applications have not required MPI\mtnote{Addresses Alexei's comment. If
-this is still too soft, we can write: ``but HEP applications do not require
-MPI.''}. Jobs are supposed to be relatively self-contained, capable of setting
-up their own execution environment or having a minimal set of common
-dependences.
+represent a set of operations performed on data sets stored in one or more input
+files. Tasks are decomposed into jobs, where each job represents the task's set
+of operations and a partition of the task's data set. Since 2005, a certain
+amount of parallelism has been progressively introduced for job
+execution~\cite{crooks2012multi} but, so far, HEP applications have not required
+MPI. Jobs are supposed to be relatively self-contained, capable of setting up
+their own execution environment or having a minimal set of common dependences.
 
@@
 
-In PanDA's data model, each
-% event
-datum refers to the recorded or simulated measurement of a physical
-% event
-process.
-% One or more events
-Data can be packaged into files or other containers. As with jobs, data
-have both attributes and states, and some of the attributes are shared between
-events and jobs. Raw, reconstruction, and simulation data are assumed to be
-distributed across multiple storage facilities and managed by the ATLAS
-Distributed Data Management (DDM)~\cite{garonne2012atlas}. When necessary,
-datasets required by each job are assumed to be replicated over the network,
-both for input and output data.
+In PanDA's data model, each datum refers to the recorded or simulated
+measurement of a physical process. Data can be packaged into files or other
+containers. As with jobs, data have both attributes and states, and some of the
+attributes are shared between events and jobs. Raw, reconstruction, and
+simulation data are assumed to be distributed across multiple storage facilities
+and managed by the ATLAS Distributed Data Management
+(DDM)~\cite{garonne2012atlas}. When necessary, datasets required by each job are
+assumed to be replicated over the network, both for input and output data.
 
@@
 
-\mtnote{From Danila: DDM enacts decisions about registering files into datasets.
-These decisions are made by productions. DDM DOES OT DECIDE which file should be
-in which dataset.}\mtnote{Done.}
-\mtnote{No communication between APF and PanDA Server (deleted arrow, adjust
-numbers and related text)}\mtnote{Done.}
-\mtnote{No more DDM Service on Edge on grid sites. DDM directly interacts with
-SE (deleted arrows, moved number 8. Update labels and related text)}\mtnote{Done.}
+% \mtnote{From Danila: DDM enacts decisions about registering files into datasets.
+% These decisions are made by productions. DDM DOES OT DECIDE which file should be
+% in which dataset.}\mtnote{Done.}
+% \mtnote{No communication between APF and PanDA Server (deleted arrow, adjust
+% numbers and related text)}\mtnote{Done.}
+% \mtnote{No more DDM Service on Edge on grid sites. DDM directly interacts with
+% SE (deleted arrows, moved number 8. Update labels and related text)}\mtnote{Done.}
 
@@
 (Fig.~\ref{fig:architecture}:11). The Job Dispatcher interrogates the Task
-Buffer module for a job that is bound to the site of that pilot and ready to be
-executed. Task Buffer checks the global queue (i.e., the PanDA database) and,
-upon availability, returns a suitable job to the Job Dispatcher. \jhanote{end of
-previous sentence needs some attention}\mtnote{Done.} The Job Dispatcher
-dispatches that job to the PanDA Pilot (Fig.~\ref{fig:architecture}:12).
+Buffer module for a job that is bound to the site of that pilot and ready to
+be executed. Task Buffer checks the global queue (i.e., the PanDA database)
+and, upon availability, returns a suitable job to the Job Dispatcher. The Job
+Dispatcher dispatches that job to the PanDA Pilot
+(Fig.~\ref{fig:architecture}:12).
 

### 2017-04-04 PandaEvolution.tex
Notes clean-up

@@
 Worldwide LHC Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more
-than
-% 100,000
-300,000 cores at over 100 Grid sites\mtnote{Addresses Alexei's comment.
-Increased Grid capacity to 300K cores. Is the number of sites still 100 or
-should I increase that too?}, with a peak performance of 0.3 petaFLOPS. This
-capacity will be sufficient for the planned analysis and data processing, but it
-will be insufficient for the Monte Carlo production workflow and any extra
-activity. To alleviate these challenges, ATLAS is engaged in a program to expand
-the current computing model to include additional resources such as the
-opportunistic use of supercomputers.
+than 300,000 cores at over 100 Grid sites, with a peak performance of 0.3
+petaFLOPS. This capacity will be sufficient for the planned analysis and data
+processing, but it will be insufficient for the Monte Carlo production workflow
+and any extra activity. To alleviate these challenges, ATLAS is engaged in a
+program to expand the current computing model to include additional resources
+such as the opportunistic use of supercomputers.
+
 % as well as commercial and academic clouds.
+% \mtnote{Addresses Alexei's comment. Increased Grid capacity to 300K cores. Is
+% the number of sites still 100 or should I increase that too?}
 
@@
 
-\mtnote{Move this to the introduction: ``There are at least two approaches to
-enable PanDA WMS to execute ATLAS workloads or workflows on supercomputers: (i)
-using the subset of resources and capabilities shared by both supercomputers and
-WLCG; (ii) reconciling the parallel and distributed computing paradigms by means
-of dedicated abstractions. The former is a pragmatic approach that enables the
-execution of specific workloads by prototyping single-point solutions. The
-latter is a principled approach, better suited for a production-grade solution,
-capable of supporting general-purpose workloads and workflows on both
-supercomputers and Grid infrastructures. \\ This section illustrates the design
-and architecture of a job broker prototyped by the PanDA team. This broker
-supports execution of part of the ATLAS production Monte Carlo workflow on
-Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
-Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).After an
-analysis of the results obtained and the lessons learned, the following section
-introduces the design and first experimental characterization of a next
-generation executor (NGE). NGE is designed to abstract resources and
-capabilities, enabling the concurrent execution of both parallel and distributed
-computing on generic high performance computing (HPC) machines.''}
-
-\jhanote{Do you mean the top of section 3? Or do you mean the very beginning
-section?}
+% \mtnote{Move this to the introduction: ``There are at least two approaches to
+% enable PanDA WMS to execute ATLAS workloads or workflows on supercomputers: (i)
+% using the subset of resources and capabilities shared by both supercomputers and
+% WLCG; (ii) reconciling the parallel and distributed computing paradigms by means
+% of dedicated abstractions. The former is a pragmatic approach that enables the
+% execution of specific workloads by prototyping single-point solutions. The
+% latter is a principled approach, better suited for a production-grade solution,
+% capable of supporting general-purpose workloads and workflows on both
+% supercomputers and Grid infrastructures. \\ This section illustrates the design
+% and architecture of a job broker prototyped by the PanDA team. This broker
+% supports execution of part of the ATLAS production Monte Carlo workflow on
+% Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
+% Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).After an
+% analysis of the results obtained and the lessons learned, the following section
+% introduces the design and first experimental characterization of a next
+% generation executor (NGE). NGE is designed to abstract resources and
+% capabilities, enabling the concurrent execution of both parallel and distributed
+% computing on generic high performance computing (HPC) machines.''}
+%
+% \jhanote{Do you mean the top of section 3? Or do you mean the very beginning
+% section?}
 
@@
 
+% \mtnote{Addresses Alexei's comment. Is the following sentence correct?}
+
 Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
@@
 four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
-HERWIG~\cite{corcella2001herwig} and many others. \mtnote{Addresses Alexei's
-comment. Is the following sentence correct?} The detector simulator is called
-Geant4~\cite{agostinelli2003geant4} and simulates the ATLAS detector and the
-interaction between that detector and particles.
-% the interaction of these particles with the sensitive material of the ATLAS
-% detector.
-Each interaction creates a so-called hit and all hits are collected and passed
-on for digitalization, where hits are further processed to mimic the readout of
-the detector. Finally, reconstruction operates local pattern recognition,
-creating high-level objects like particles and jets.
+HERWIG~\cite{corcella2001herwig} and many others. The detector simulator is
+called Geant4~\cite{agostinelli2003geant4} and simulates the ATLAS detector and
+the interaction between that detector and particles. Each interaction creates a
+so-called hit and all hits are collected and passed on for digitalization, where
+hits are further processed to mimic the readout of the detector. Finally,
+reconstruction operates local pattern recognition, creating high-level objects
+like particles and jets.
 
@@
 cleanup on exit. Upon submission, a copy of the MPI script runs on every
-available worker node,
+available worker node, starting the execution of the ATLAS job's payload in a
+subprocess and waits until its completion.
+
 % Each script knows its MPI rank (an index that runs from zero to a maximum
@@
 % script
-starting the execution of the ATLAS job's payload in a subprocess and waits
-until its completion.
 
@@
 
-\mtnote{Address Alexei's comments: Add a paragraph explaining how this solution
-has been ported to other HPC machines and reference relevant publications.}
+% \mtnote{Address Alexei's comments: Add a paragraph explaining how this solution
+% has been ported to other HPC machines and reference relevant publications.}
 While PanDA Broker implementation is resource specific, it was successfully

### 2017-04-04 nge.tex
Address comments and Notes clean-up

--- 2017-04-03 nge.tex
@@
 % \subsection{Rationale}
-\aanote{I think that the names of the states should be removed. They are
-meaningless without the figures. I left them just in case we decide to
-reintroduce the figures.}\mtnote{Done.}
+% \aanote{I think that the names of the states should be removed. They are
+% meaningless without the figures. I left them just in case we decide to
+% reintroduce the figures.}\mtnote{Done.}
 
@@
 
-As seen in \S\ref{sec:panda_deployment}, PanDA Broker was  deployed on the
-DTNs as Titan's worker nodes lack connectivity to the wide area network The
-lack of pilot capabilities impacts both the efficiency and the flexibility of
-PanDA's execution process. Pilots could improve efficiency by increasing
-throughput and enabling greater backfill utilization. Further, pilots makes it
-easier to support heterogeneous workloads.
+As seen in \S\ref{sec:panda_deployment}, PanDA Broker was  deployed on the DTNs
+as Titan's worker nodes lack connectivity to the wide area network The lack of
+pilot capabilities impacts both the efficiency and the flexibility of PanDA's
+execution process. Pilots could improve efficiency by increasing throughput and
+enabling greater backfill utilization. Further, pilots makes it easier to
+support heterogeneous workloads.
 
@@
 detector simulations are packaged into a PBS job and this job is queued on
-Titan, no further simulations can be added to that job. New simulations have
-to be packaged into a new PBS job that needs to be submitted to Titan based
-upon backfill availability.
+Titan, no further simulations can be added to that job. New simulations have to
+be packaged into a new PBS job that needs to be submitted to Titan based upon
+backfill availability.
 
-The support of  multiple generations of workload would enable more efficient
-use of the backfill availability of walltime. Currently, when a set of
-simulations ends, the PBS job also ends, independent of whether more wall-time
-would still be available. With a pilot, additional simulations could be
-executed  to utilize all the available wall-time, while avoiding further job
-packaging and submission overheads.
+The support of  multiple generations of workload would enable more efficient use
+of the backfill availability of walltime. Currently, when a set of simulations
+ends, the PBS job also ends, independent of whether more wall-time would still
+be available. With a pilot, additional simulations could be executed  to utilize
+all the available wall-time, while avoiding further job packaging and submission
+overheads.
 
@@
 model: knowing the number of simulations before submitting the MPI script, and
-having a fixed number of events per simulation (100 at the moment). Pilots
-would enable the scheduling of simulations independently from whether they
-were available at the moment of submitting the pilot. Further, simulations
-with a varying number of events could be scheduled on a pilot, depending on
-the amount of remaining walltime and the distribution of execution time per
-event, as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event-
-distrib}. These capabilities would increase the efficiency of the PanDA Broker
-when there is a large difference between the number of cores and walltime.
+having a fixed number of events per simulation (100 at the moment). Pilots would
+enable the scheduling of simulations independently from whether they were
+available at the moment of submitting the pilot. Further, simulations with a
+varying number of events could be scheduled on a pilot, depending on the amount
+of remaining walltime and the distribution of execution time per event, as shown
+in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event- distrib}. These
+capabilities would increase the efficiency of the PanDA Broker when there is a
+large difference between the number of cores and walltime.
 
@@
 
-\mtnote{NOTE: I would speak about about backfill/not backfill in the discussion
-of NGE, when speaking about its generality towards resources, i.e., unified
-submission and scheduling process across different resources and multiple
-resources. Part of this generality is being able to submit to whatever batch
-system is supported by SAGA (including Titan's PBS and all its queues) and, in
-case, to multiple resources at the same time.}
-
-\mtnote{NOTE: NGE is not able to use backfill on Titan as we do not have
-specific functionalities to interrogate the Moab scheduler about backfill
-availability. This is why NGE should be presented as the pilot for PanDA Broker
-and not as an alternative to PanDA Broker. Further, NGE alone would not be able
-to speak directly to PanDA Server.}
+% \mtnote{NOTE: I would speak about about backfill/not backfill in the discussion
+% of NGE, when speaking about its generality towards resources, i.e., unified
+% submission and scheduling process across different resources and multiple
+% resources. Part of this generality is being able to submit to whatever batch
+% system is supported by SAGA (including Titan's PBS and all its queues) and, in
+% case, to multiple resources at the same time.}
+%
+% \mtnote{NOTE: NGE is not able to use backfill on Titan as we do not have
+% specific functionalities to interrogate the Moab scheduler about backfill
+% availability. This is why NGE should be presented as the pilot for PanDA Broker
+% and not as an alternative to PanDA Broker. Further, NGE alone would not be able
+% to speak directly to PanDA Server.}
 
@@
 The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for communication
-and coordination of the execution of units. ORTE is a critical component of the
-OpenMPI implementation, developed to support distributed high-performance
-computing applications operating in a heterogeneous environment~\cite{castain05:_open_rte, cug-2016}.  
- ORTE transparently provides support for interprocess communication, resource discovery and allocation, and process launching across a variety of platforms.
-ORTE provides a mechanism similar to the Pilot concept -
-it allows the user to create a \emph{``dynamic virtual machine''} (DVM) that
-spans multiple nodes. ORTE provides libraries to enable the submission,
-monitoring and managing of tasks, avoiding filesystem bottlenecks and race
-conditions with network sockets. As a consequence, it is able to minimize the
-system overhead while submitting tasks.
+and coordination of the execution of units. This environment is a critical
+component of the OpenMPI implementation, developed to support distributed
+high-performance computing applications operating in a heterogeneous
+environment~\cite{castain05:_open_rte, cug-2016}.
+
+ORTE provides a mechanism to create a \emph{``dynamic virtual machine''} (DVM)
+that spans multiple nodes. The DVM transparently provides support for
+interprocess communication, resource discovery and allocation, and process
+launching across a variety of platforms. Libraries enable the interaction
+between users and ORTE to enable the submission, monitoring and managing of
+tasks, avoiding filesystem bottlenecks and race conditions with network sockets.
+As a consequence, ORTE is able to minimize the system overhead while submitting
+tasks.
+
+% ORTE transparently provides support for interprocess communication, resource
+% discovery and allocation, and process launching across a variety of platforms.
+% ORTE provides a mechanism similar to the Pilot concept - it allows the user to
+% create a \emph{``dynamic virtual machine''} (DVM) that spans multiple nodes.
+% ORTE provides libraries to enable the submission, monitoring and managing of
+% tasks, avoiding filesystem bottlenecks and race conditions with network sockets.
+% \aanote{Requires validation: Alternative version of the three sentences above
+% (too many ORTE in my opinion):}
 

### 2017-04-04 Experiments.tex
Notes clean-up

@@
 We designed experiments to characterize the performance of the NGE on Titan,
-with an emphasis on understanding its overhead and thus the cost of
-introducing new functionalities.  We perform three groups of experiments in
-which we investigate the weak scalability, weak scalability with multiple
-generation, and strong scalability of the NGE.
+with an emphasis on understanding its overhead and thus the cost of introducing
+new functionalities.  We perform three groups of experiments in which we
+investigate the weak scalability, weak scalability with multiple generation, and
+strong scalability of the NGE.
 
@@
 We  measured the execution time of the pilots and of the AthenaMP  executed
-within them, collecting timestamps at  all stages of the execution.
-Experiments have been performed  by  submitting NGE's pilots  to Titan's batch
-queue.  The turnaround time of an individual run is determined by queue
-waiting times. Since we are interested only in the performances of the NGE, we
-removed queue time from our statistics.
+within them, collecting timestamps at  all stages of the execution. Experiments
+have been performed  by  submitting NGE's pilots  to Titan's batch queue.  The
+turnaround time of an individual run is determined by queue waiting times. Since
+we are interested only in the performances of the NGE, we removed queue time
+from our statistics.
 
@@
 
-In this experiment  we run as many AthenaMP instances (hereafter referred to
-as tasks)  as the number of nodes controlled by the pilot. Each AthenaMP
-simulates 100 events, requiring $\sim 4200$ seconds on average.
+In this experiment  we run as many AthenaMP instances (hereafter referred to as
+tasks)  as the number of nodes controlled by the pilot. Each AthenaMP simulates
+100 events, requiring $\sim 4200$ seconds on average.
 
-Tasks do not  wait within the NGE Agent's queue since  one node  is available
-to each AthenaMP instance.  Overheads in task execution are consequence
-primarily of the three other factors: (i) the  initial bootstrapping of the
-pilot on the nodes; (ii) the UnitManager's dispatching of  units (tasks) to
-the agent; and (iii) time for the agent to bootstrap all the tasks on the nodes.
+Tasks do not  wait within the NGE Agent's queue since  one node  is available to
+each AthenaMP instance.  Overheads in task execution are consequence primarily
+of the three other factors: (i) the  initial bootstrapping of the pilot on the
+nodes; (ii) the UnitManager's dispatching of  units (tasks) to the agent; and
+(iii) time for the agent to bootstrap all the tasks on the nodes.
 
-We tested % different
-pilots % sizes, i.e. :
-with 250, 500, 1000 and 2000 worker nodes and
-%. For all of them, the walltime was
-2 hours walltime. The time duration is determined by the Titan's walltime policy. Figure~\ref{fig:weakScal1a} depicts the average pilot
-duration, the average execution time of AthenaMP, and the NGE pilot overhead as
-function of the pilot size.
+We tested  pilots  with 250, 500, 1000 and 2000 worker nodes and 2 hours
+walltime. The time duration is determined by the Titan's walltime policy.
+Figure~\ref{fig:weakScal1a} depicts the average pilot duration, the average
+execution time of AthenaMP, and the NGE pilot overhead as function of the pilot
+size.
 
@@
 
-% Firstly, w
-We % can
-observe that, despite some fluctuations due to external factors
-% \footnote{such as the
-(e.g., Titan's shared filesystem and the shared database used by the NGE), the
-average execution time of AthenaMP % oscillates
-ranges between 4200 and 4800 seconds. % Secondly, w
-We % can
-also observe that in all the cases the gap between AthenaMP execution times and
-the pilot durations is minimal, although it slightly increases with the pilot
-size. % By observing NGE's overhead, w
-We % can
-notice that NGE's overhead does grow linearly with the number of units.
+We observe that, despite some fluctuations due to external factors (e.g.,
+Titan's shared filesystem and the shared database used by the NGE), the average
+execution time of AthenaMP  ranges between 4200 and 4800 seconds.  We  also
+observe that in all the cases the gap between AthenaMP execution times and the
+pilot durations is minimal, although it slightly increases with the pilot size.
+We  notice that NGE's overhead does grow linearly with the number of units.
 
@@
 
-The NGE provides an important new capability of submitting multiple
-generations of AthenaMP to the same pilot. In order to investigate the cost of
-doing so, we performed a variant of the weak scalability experiments. This
-stresses the pilot's components, as new tasks are scheduled for execution on
-the Agent while other tasks are still running.
+The NGE provides an important new capability of submitting multiple generations
+of AthenaMP to the same pilot. In order to investigate the cost of doing so, we
+performed a variant of the weak scalability experiments. This stresses the
+pilot's components, as new tasks are scheduled for execution on the Agent while
+other tasks are still running.
 
-\mtnote{we dowe need to keep it below 2 hours?}\aanote{You are right. It was misleading. I Changed. Check if you like it.}
+% \mtnote{we dowe need to keep it below 2 hours?}\aanote{You are right. It was misleading. I Changed. Check if you like it.}
 
 In these experiments, we run five AthenaMP instances per node.  As these
-experiments are designed to investigate the overhead generated by the
-scheduling and bootstrap of AthenaMP instances, we reduced the number of
-events simulated by each AthenaMP task to sixteen in such a way that the
-running time of each AthenaMP is, on average, $\sim 1200$ seconds. This
-experiment design choice does not affect the  objectives or accuracy of the
-experiments, but allows us to scale experiments to large node counts by being
-conservative with allocation.
+experiments are designed to investigate the overhead generated by the scheduling
+and bootstrap of AthenaMP instances, we reduced the number of events simulated
+by each AthenaMP task to sixteen in such a way that the running time of each
+AthenaMP is, on average, $\sim 1200$ seconds. This experiment design choice does
+not affect the  objectives or accuracy of the experiments, but allows us to
+scale experiments to large node counts by being conservative with allocation.
 
@@
 
-We ran
-% different
-pilots
-% sizes, i.e. :
-with 256, 512, 1024 and 2048 worker nodes and
-%. For all of them, the walltime was
-3\mtnote{2?} hours walltime. Figure~\ref{fig:weakScal2a} depicts the average
-pilot duration, the average execution time of five sequential
-generations\mtnote{generations?}\aanote{Changed} of AthenaMP, and the corresponding overhead. We
-% can
-observe that the difference between the two durations is more marked than in the
-previous experiments \mtnote{mostly due to the increased overhead?}\aanote{Well, I guess that's plain.}.
-Despite this, we notice that the growth of the overhead is consistent with the increment of the number of tasks per node for pilots with 256, 512 and 1024 worker nodes, and less than linear for the pilot with 2048 worker nodes.
+We ran pilots with 256, 512, 1024 and 2048 worker nodes and 3\mtnote{2?} hours
+walltime. Figure~\ref{fig:weakScal2a} depicts the average pilot duration, the
+average execution time of five sequential generations of AthenaMP, and the
+corresponding overhead. We observe that the difference between the two durations
+is more marked than in the previous experiments. Despite this, we notice that
+the growth of the overhead is consistent with the increment of the number of
+tasks per node for pilots with 256, 512 and 1024 worker nodes, and less than
+linear for the pilot with 2048 worker nodes.
 
@@
 
-The last experiments % deals with
-study strong scalability by running the same number of tasks for different pilot sizes. We used
-2048 AthenaMP instances and % tested
-pilots with 256, 512, 1024 and 2048 nodes.
-Thus, the number of AthenaMP generations is equal to eight times the size of the smallest pilot and corresponds to the size of the largest pilot.
-As a consequence, the number of consecutive generations of AthenaMP decreases with the pilot size by generating different dynamics within the pilots.
-These experiments are designed to investigate whether pilot overhead is affected by the degree of concurrency within the pilot and/or the number of tasks.
-\mtnote{Not sure I understand this sentence}\aanote{My fault.
-It should have been seven and null or 8 and 1.
-It depends on how we consider the word sequential.
-anyway, to avoid confusion I changed the sentence.}.
-Each AthenaMP instance
-simulates sixteen events as in the previous experiment.
+The last experiments  study strong scalability by running the same number of
+tasks for different pilot sizes. We used 2048 AthenaMP instances and  pilots
+with 256, 512, 1024 and 2048 nodes. Thus, the number of AthenaMP generations is
+equal to eight times the size of the smallest pilot and corresponds to the size
+of the largest pilot. As a consequence, the number of consecutive generations of
+AthenaMP decreases with the pilot size by generating different dynamics within
+the pilots. These experiments are designed to investigate whether pilot overhead
+is affected by the degree of concurrency within the pilot and/or the number of
+tasks. Each AthenaMP instance simulates sixteen events as in the previous
+experiment.
 
-Figure \ref{fig:strongScala} % depicts
-shows the average pilot duration and the average execution time of possibly
-sequential AthenaMP instances.  We % can
-notice that the difference between the pilot duration and the AthenaMP execution
-times is almost constant for all the pilot sizes, although the overall duration
-of the pilot decreases linearly with the pilot size.
+% \mtnote{Not sure I understand this sentence}\aanote{My fault. It should
+% have been seven and null or 8 and 1. It depends on how we consider the word
+% sequential. anyway, to avoid confusion I changed the sentence.}.
+
+Figure \ref{fig:strongScala}  shows the average pilot duration and the average
+execution time of possibly sequential AthenaMP instances.  We  notice that the
+difference between the pilot duration and the AthenaMP execution times is almost
+constant for all the pilot sizes, although the overall duration of the pilot
+decreases linearly with the pilot size.
 

### 2017-04-04 PandaOverview.tex
Eliminate all comments and shift figure around to save space

@@
 
-% We summarize PanDA's design, architecture, and execution process, focussing on
-% those subsystems and components that are most relevant to the the integration
-% of PanDA with supercomputers.
-
-% \mtnote{TODO: replace `resource' with `site'; `http' to `HTTP'?}
-
 % -----------------------------------------------------------------------------
@@
 \label{ssec:panda_design}
-
-% \mtnote{Addresses Alexei's comment. I eliminated `event' as abstraction and used
-% datum/data/physical process when discussing data management in the whole
-% section.}
-
-% \mtnote{Addresses Alexei's comment. If this is still too soft, we can write:
-% ``but HEP applications do not require MPI.''}
 
@@
 and the X.509 standard and on the virtual organization (VO) abstraction.
-
-% Currently, PanDA's execution model is based on five main abstractions: task,
-% job, queue, pilot, and event. Both tasks and jobs are assumed to have
-% attributes and states and to be queued into a global queue for execution.
-% Prioritization and binding of jobs are assumed to depend on the attributes of
-% each job. Pilot is used to indicate the abstraction of resource capabilities.
-% Each job is thought to be bound to one pilot and executed on the site where
-% the pilot has been instantiated.
 
@@
 
-% \begin{itemize}
-%   \item No MPI tasks?
-%   \item Pilot multitenancy?
-%   \item PanDA Server Brokerage component and Event service interaction?
-%   \item PanDA Server Data Service component and Event Service interaction?
-% \end{itemize}
-
-% Different types of job payload (i.e., executable) are assumed, both in terms
-% of user-defined batch scripts or application frameworks for group of users.
-
-% Consistently, pilots are considered to be multitenant as jobs belonging to
-% multiple workflows and/or users can be bound and executed on them.
-% PanDA' data model mostly uses events as the unit of data
-% \mtnote{Is event an abstraction specific to the MD workflow?}.
-
-% \sergeynote{Raw data from the ATLAS detector is captured at CERN and goes
-% through reconstruction step to produce data in secondary formats, smaller in
-% volume and with more and more physics content instead of raw detector hits.
-% Consequently a second, custodial, copy of raw data is being distributed in
-% predetermined volumes to National T1 centers around the world (like BNL  T1
-% center). This insures low raw data redundancy and safety. Multiple copies of
-% chunks of secondary data sets (derived from real raw data) are also
-% distributed to ~100 centers around the word for ease of access and improved
-% availability for analysis. Then raw detector data is stored on Tapes at CERN,
-% just  in case. Most users, who do physics analysis, rarely work with raw
-% data, only a small group of experts needs to work with raw data for
-% specialized tasks like detector calibrations and such. There may be a need to
-% redo a reconstruction step, when errors in previous reconstruction campaign
-% are uncovered or improved detector understanding will lead for a better
-% quality physics. Then data stored on tape is recovered and re-processed
-% again. Simulation data is produced in many places anyway and is not
-% centralized to begin with. Here's the paper that describes current DDM state
-% in ATLAS, data replication strategies and such.
-% http://iopscience.iop.org/article/10.1088/1742-6596/396/3/032045/pdf DDM is a
-% complex and evolving topic in ATLAS.}\mtnote{Thank you. As we are describing
-% the design of PanDA I thought we should not enter into a too detailed
-% description of the DDM design. I eliminated the mistake I had made about
-% `centralized data' but left DDM just as a reference. Any better?}
-
 % -----------------------------------------------------------------------------
@@
 \label{ssec:panda_arch}
-
-% \mtnote{From Danila: DDM enacts decisions about registering files into datasets.
-% These decisions are made by productions. DDM DOES OT DECIDE which file should be
-% in which dataset.}\mtnote{Done.}
-% \mtnote{No communication between APF and PanDA Server (deleted arrow, adjust
-% numbers and related text)}\mtnote{Done.}
-% \mtnote{No more DDM Service on Edge on grid sites. DDM directly interacts with
-% SE (deleted arrows, moved number 8. Update labels and related text)}\mtnote{Done.}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
-    \caption{PanDA architecture with the subsystems and components relevant to
-    the integration of PanDA with supercomputers. Numbers indicates the
-    execution process based on JEDI: From the submission of tasks (1) to the
-    retrieval of their output (19). The monitoring subsystem, the architectural
-    details of PanDA Pilot and the communication among subsystem's components
-    are abstracted to improve clarity.}
-\label{fig:architecture}
-\end{figure}
-
-% -----------------------------------------------------------------------------
-% PANDA SERVER
-% -----------------------------------------------------------------------------
-% \textbf{PanDA Server} is the central hub service of PanDA WMS and it operates
-% as a web service~\cite{maeno2011overview}. It consists of four main components
-% implemented in Python: Task Buffer, Brokerage, Job Dispatcher, and Data
-% Service. Together, these components implement a global queue where to store
-% jobs, share information about jobs and their dataset properties with the other
-% subsystems, and update their states at execution time.
-% -----------------------------------------------------------------------------
-
-% It runs on Apache web server, interacting with the back-end database running
-% on separate machines. PanDA Server implements the Apache worker model, with
-% many independent processes handling client requests in parallel. Since all
-% state is maintained in the central database, the PanDA Server instance itself
-% is stateless. Currently, the production version of PanDA utilizes an Oracle
-% database as backend, but PanDA Server can also work with the MySQL family of
-% databases.
-
-% . It provides a task queue
-% managing all job information centrally. The PanDA server receives jobs through
-% the client interface into the task queue, upon which a brokerage module
-% operates to prioritize and assign work on the basis of job type, priority,
-% input data and its locality, and available CPU resources. The PanDA server
-%
-% and communicating via Python APIs
-
-% Task Buffer manages PanDA's global queue by: (i) storing jobs into the global
-% queue; (ii) providing information about and updating the state of each stored
-% job. The Brokerage component: (i) matches job attributes with site and pilot
-% attributes; (ii) manages the dispatch of input data to processing sites; and
-% (iii) derives PanDA's data pre-placement requirement. The Job Dispatcher: (i)
-% receives requests from active pilots for new jobs to process; (ii) retrieves
-% from Task Buffer the jobs with highest priority and that can be executed on
-% the site of the requesting pilot; (iii) creates a wrapper script for the
-% selected job, tailored to the site of the requesting pilot; and (iv) monitors
-% the state of the job execution on the pilot, updating the Task Buffer
-% component. Finally, the Data Service component: (i) retrieves information from
-% Task Buffer and Brokerage about the data requirements of each job and the site
-% to which each job has been assigned; (ii) interacts with the ATLAS DDM to make
-% input files available to the jobs and manage their output data.
-
-% -----------------------------------------------------------------------------
-% PANDA PILOT
-% -----------------------------------------------------------------------------
-% \textbf{PanDA Pilot} implements the execution environment for PanDA's
-% jobs\cite{nilsson2011atlas}. Implemented in Python, a PanDA Pilot is launched
-% by a wrapper script submitted to a site. Wrapper scripts are tailored to the
-% middleware exposed by each site so to maintain PanDA Pilot site-independent.
-% PanDA Pilot pulls a job for execution, sets up the execution environment
-% including the required input files, spawns the jobs's executable, monitors its
-% execution, and supports the output's staging.
-% -----------------------------------------------------------------------------
-
-% t: (i) prepares the execution environment for PanDA's jobs, collecting
-% information about the compute, data, and network capabilities of the site;
-% (ii) requests a job for execution to the Job Dispatcher component of PanDA
-% Server; (iii) initiates and monitors the execution of pulled jobs, including
-% setting up the job, transferring its input when required, executing the
-% payload, transferring output, and checking log files and local space; and (iv)
-% cleans up when the payload has finished to execute.
-
-% PanDA is a pilot based workload management system.
-% In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload
-% processing on a worker node) are submitted to sites. When these pilot jobs
-% start on a worker node they contact a central server to retrieve a real
-% payload (i.e., an end-user job) and execute it.
-
-% -----------------------------------------------------------------------------
-% PILOT DESCRIPTION: COMMENTED OUT BECAUSE WE DESCRIBE IT LATER ON.
-% -----------------------------------------------------------------------------
-% PanDA Pilot implements the pilot paradigm~\cite{turilli2015comprehensive},
-% decoupling resource acquisition from job execution via multi-stage scheduling.
-% This has several benefits, including increasing throughput by lowering the
-% time taken to queue and schedule jobs; enabling both concurrent and
-% sequential job execution depending on the number of cores available to the
-% pilot and the time for which it remains available; and exposing a unified and
-% consistent interface for the scheduling of jobs, independent from the
-% middleware of the site on which the pilot has been instantiated.
-% -----------------------------------------------------------------------------
-
-% The specific implementation of PanDA Pilot promotes reliability by
-% monitoring the execution of the job's payload, and a certain degree of
-% fault-tolerance by recovering information and, possibly, output files from
-% jobs that failed, even before instantiating the PanDA Pilot.
-
-% Using these pilot-based workflows helps to improve job reliability, optimize
-% resource utilization, allows for opportunistic resources usage, and mitigates
-% many of the problems associated with the inhomogeneities found on the Grid.
-
-% Upon bootstrapping, PanDA Pilot downloads major parts of its runtime code via
-% HTTP from a central Subversion repository. The repository works with an Apache
-% web server, configured with a memory-based web proxy (Squid). The purpose of
-% the cache is to reduce the request load on the back-end Subversion server.
-% That improves performance since only occasional queries trigger a full lookup
-% on the back-end Subversion system, and most external queries are pulled from
-% memory on the front-end web server. The benefits of this system include a
-% high-performance code download service, combined with code updates still being
-% immediately available as soon as they are committed to source code control.
-
-% -----------------------------------------------------------------------------
-% AUTOPYFACTORY
-% -----------------------------------------------------------------------------
-% \textbf{AutoPyFactory} implements a factory to submit pilots locally and to
-% both grid and cloud sites~\cite{caballero2012autopyfactory}. Implemented to be
-% executed as a single demonized process, AutoPyFactory concurrently spawns
-% multiple processes called APFQueue that, on the base of information about jobs
-% queued on the PanDA server and on the state and capabilities of resources,
-% submit pilots to multiple sites with Condor-G~\cite{frey2002condor}.
-% -----------------------------------------------------------------------------
-
-% Each APFQueue communicates with PanDA Server and the Brokerage module to
-% collect information about the state of the jobs it manages. On the basis of
-% this information, an APFQueue submits new pilots to one site. AutoPyFactory
-% implements functionalities to evaluate the number of pilots already
-% instantiated on each site, and to apply a consistent and adequate pressure on
-% their batch queues. Submission is performed via the Condor-G batch
-% system~\cite{frey2002condor}, enabling submission to multiple type of sites.
-% AutoPyFactory exposes a dedicated monitoring service, feeding information to
-% the PanDA Monitoring subsystem. A proxy manager enables certificate-based
-% authentication and authorization for each site.
-
-% As a pilot-based system, PanDA requires some way to get the initial PanDA
-% pilot onto worker nodes at sites. This is done with the help of a component
-% called AutoPyFactory (APF). APF runs in a single daemonized process,
-% launching a separate thread for each internal workflow. Each one of these
-% internal workflows typically serves a single job queue as defined in WMS, and
-% delivers pilots to a single batch queue, either local or remote. The behavior
-% of these APF workflows is determined by the combination of a set of plugins,
-% invoked in a fixed order, in a loop, each one in charge of the performance of
-% a well defined action.
-
-% -----------------------------------------------------------------------------
-% PANDA MONITORING - SPACE SAVING
-% -----------------------------------------------------------------------------
-% \paragraph{\textbf{PanDA Monitoring}} this subsystem is implemented as a
-% web-based, dashboard-style graphical application~\cite{klimentov2011atlas}. It
-% runs on Apache and interacts with the back-end database of PanDA Server to
-% enable persistence. PanDA Monitoring enables users and site administrators to
-% gather information about the status of current and past jobs, data movement,
-% and pilot factory. PanDA Monitoring allows users access to all log files of
-% each job, greatly simplifying code debugging and failure analysis in a
-% distributed computing environment.
-% -----------------------------------------------------------------------------
-
-% -----------------------------------------------------------------------------
-% JEDI
-% -----------------------------------------------------------------------------
-% \textbf{JEDI} exposes a task abstraction for workload and workflows
-% specification and interfaces with PanDA Server to manage their execution on
-% remote sites~\cite{borodin2015scaling}. Implemented as a standalone service,
-% JEDI partitions tasks to jobs, and binds these jobs to sites via dedicated
-% queues. JEDI collects information about sites' resource capability,
-% availability, and performance. Based on this information, JEDI associates
-% sites with similar capabilities to a queue. JEDI replaces and extends the
-% Brokerage module of the PanDA Server.
-% -----------------------------------------------------------------------------
-
-% Each job is assigned to a queue by matching the job's requirements to the
-% capability, availability and performance of the sites of that queue. In this
-% way,
-
-% This module still handles communication between the Task
-% Buffer and the Data Service modules.
-
-% -----------------------------------------------------------------------------
-% \subsection{Execution}
-% \label{ssec:panda_exec}
-
-% The tasks of workloads and workflows submitted to PanDA are partitioned
-% to one or more jobs. Each job is then executed on one of the available
-% PanDA Pilots, chosen among those that the PanDA Server manages across multiple
-% sites. Tasks processes a certain amount of events that are stored into files,
-% and when a task is partitioned to multiple jobs, subsets of events need to be
-% assigned to each job and made available to the pilot on which that job will be
-% executed. Finally, the output of each job need to be retrieved and aggregated
-% into the output of the original task (Fig.~\ref{fig:architecture}).
-
 The relation between tasks and jobs can be one-to-one or one-to-many, and the
@@
 required users or applications to perform a static conversion between tasks and
-jobs: tasks were described as a set of jobs
-% with a fixed number of events
-and then submitted to the PanDA Server.
+jobs: tasks were described as a set of jobs and then submitted to the PanDA
+Server.
 
@@
 requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7) that, when needed,
-takes care of % aggregating files into datasets and containers,
-replicating them on the required sites (Fig.~\ref{fig:architecture}:8).
+takes care of  replicating them on the required sites
+(Fig.~\ref{fig:architecture}:8).
 
-% AutoPyFactory subsystem communicates with the Task Buffer component of the
-% PanDA Server, acquiring information about jobs that are ready for execution on
-% specific (type of) sites (Fig.~\ref{fig:architecture}:9). On the basis of this
-% information,
 Meanwhile, AutoPyFactory defines a set of suitable PanDA Pilots and concurrently
@@
 (Fig.~\ref{fig:architecture}:10).
-
-% AutoPyFactory submits pilots only when similar pilots are not yet available
-% and maintains a predefined amount of pressure on the queue of each site. In
-% this way, resource availability and utilizations is optimized.
 
@@
 
-% \mtnote{Can a pilot request multiple jobs to execute
-% concurrently?}\sergeynote{In principle it should be able to do that, but
-% that's not how it is operated. One pilot - one job is a cleaner model in an
-% environment where resources are ``owned'' by the organization that runs
-% pilots. Simplifies many things. At least as far as I know.}
-
 Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
@@
 staging-out location (Fig.~\ref{fig:architecture}:15).
-
-% During the job's payload execution, the monitoring process of PanDA Pilot
-% checks the status of the job execution and its environment.
 
@@
 
-% \mtnote{Think whether a comparison with other WMS based on the review we did
-% makes sense in this context.}\sergeynote{You provided a nice description of
-% the PanDA system here.}
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
+    \caption{PanDA WMS architecture. Numbers indicates the execution process
+    based on JEDI. Several subsystems and components, the architectural
+    details of PanDA Pilot and the communication among subsystem's components
+    are abstracted to improve clarity.}
+\label{fig:architecture}
+\end{figure}

### 2017-04-04 PandaOverview.tex
Clean up

@@
 Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
-Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of
-ATLAS workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we
-do not discuss them as they are  not relevant to  an understanding of how
-PanDA has been ported to supercomputers. For a full list of subsystems see
-Ref.~\cite{panda-wiki_url}. Figure~\ref{fig:architecture} shows a diagrammatic
+Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of ATLAS
+workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we do not
+discuss them as they are  not relevant to  an understanding of how PanDA has
+been ported to supercomputers. For a full list of subsystems see
+Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture} shows a diagrammatic
 representation of PanDA main subsystems, highlighting the execution process of

### 2017-04-04 nge.tex
Clean up

@@
-% \subsection{Rationale}
-% \aanote{I think that the names of the states should be removed. They are
-% meaningless without the figures. I left them just in case we decide to
-% reintroduce the figures.}\mtnote{Done.}
-
-% \jhanote{admittedly i haven't read earlier section in question  carefully, but
-% i'm a bit confused about the use of "wide area". wide area is typically used for
-% cross data center}\mtnote{Every PanDA Pilot has to communicate with PanDA
-% Server. PanDA Server runs on dedicated resources, outside OLCF. Titan's worker
-% nodes cannot communicate with PanDA Server because their connectivity is limited
-% to Titan's network. Do you have a suggestion for an alternative to ``wide
-% area''?} \jhanote{how is the following: "... deployed on the DTNs as Titan's worker
-% nodes lack connectivity to the wide area network"}.
-
 As seen in \S\ref{sec:panda_deployment}, PanDA Broker was  deployed on the DTNs
@@
 
-% \mtnote{Are there more reasons for wanting multiple generations?}
-
-% \jhanote{not sure MPI script is optimal description or ambiguous.}\mtnote{Is
-% what we used in the previous section(s) to designate what PanDA Broker submits
-% to Titan's PBS batch system. Job is already used for what is pulled from
-% PanDA server.}
-
 Multiple generations would also relax two assumptions of the current execution
@@
 of remaining walltime and the distribution of execution time per event, as shown
-in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event- distrib}. These
+in \S\ref{ssec:panda_titan}, Fig.~\ref{fig:100event- distrib}. These
 capabilities would increase the efficiency of the PanDA Broker when there is a
 large difference between the number of cores and walltime.
-
-% to one in which the number of events is decided on the base of the available
-% walltime. Pilots would enable to package simulations with a heterogeneous
-% number of events, especially when a small amount of cores is available for a
-% long period of time and \textit{vice versa}. In these situations, a PBS job
-% would require all simulations to have a small number of events that would not
-% use all the available walltime.
-
-% a pilot would instead support multiple generations of simulations with a
-% different number of events and therefore duration.
 
@@
 nodes, not by the payload.
-
-% \mtnote{I tried to argue that pilots facilitates the submission to a normal
-% queue but I was not able to find a convincing argumentation. When we consider
-% only Titan, a PBS job is the only way to submit a job to its queue(s). With %
-% NGE, we translate a pilot description to a PBS job and then submit it to a
-% queue via a PBS command. PanDA Brokers do something very similar: they
-% package detector simulations into a PBS job and they submit it via a PBS
-% command. The fact that PanDA Brokers define the number of cores and walltime
-% of a PBS job on the base of backfill information, does not make the packaging
-% or the submission processes any different: PanDA Brokers are already
-% submitting to the nomal queue. If I got it right, PanDA Brokers could already
-% submit a job with, say, 12000 cores and 2 hours walltime. They don't do it
-% because they have no allocation and they would wait a long time in the queue
-% (maybe, let's see what our experiments say about that). I do not see how NGE
-% could make all this any different.}
-
-% \mtnote{NOTE: I would speak about about backfill/not backfill in the discussion
-% of NGE, when speaking about its generality towards resources, i.e., unified
-% submission and scheduling process across different resources and multiple
-% resources. Part of this generality is being able to submit to whatever batch
-% system is supported by SAGA (including Titan's PBS and all its queues) and, in
-% case, to multiple resources at the same time.}
-%
-% \mtnote{NOTE: NGE is not able to use backfill on Titan as we do not have
-% specific functionalities to interrogate the Moab scheduler about backfill
-% availability. This is why NGE should be presented as the pilot for PanDA Broker
-% and not as an alternative to PanDA Broker. Further, NGE alone would not be able
-% to speak directly to PanDA Server.}
 
@@
 
-% We developed and we are currently integrating a prototype of Next Generation
-% Executor (NGE) to add pilot capabilities to the PanDA Broker.
-
 NGE is a runtime system to execute heterogeneous and dynamically determined
@@
 
-% Thus, workloads and pilots are described via the Pilot API and passed to the
-% NGE runtime system. NGE launches the pilots and executes the tasks of the
-% workload on them.
-
 \begin{figure}
@@
    \includegraphics[width=\columnwidth]{figures/rp_architecture_compact_atlaswms_paper.pdf}
-  % \caption{RP overview. An application uses the Pilot API to describe pilots
-  %   (green squares) and units (red circles). The PilotManager (purple box)
-  %   instantiates (dash arrow) pilots, the UnitManager (purple box)
-  %   instantiates (solid arrow) units. Both managers are executed on the user
-  %   workstation. Pilots are launched (dashed arrows) on resources A and B
-  %   via SAGA API, an Agent (purple box) is bootstrapped for each pilot
-  %   (green box), units (solid red circles) are scheduled to the Agents via
-  %   MongoDB and executed by the Agent's Executer (yellow box).
   \caption{NGE Architecture as deployed on Titan. The PilotManager and the
@@
 \end{figure}
-
-% Internally, NGE represents pilots as aggregates of resources independent from
-% the architecture and topology of the target machines, and workloads as a set
-% of units to be executed on the resources of the pilot. Both pilots and units
-% are stateful entities, each with a well-defined state model and life cycle.
-% Their states and state transitions are managed via the three modules of the
-% NGE architecture: PilotManager, UnitManager, and Agent
-% (Fig.~\ref{fig:arch-overview}, purple boxes).
 
@@
 one of Titan's DTN so as to be reachable by both modules.
-
-% \jhanote{for consistency should we say Moab?}\mtnote{Moab = Scheduler; Torque
-% = Batch system. I used PBS Batch System because it makes easier to speak about
-% both the system and the script submitted to it.}
-
-% A similar approach would be used to enable PanDA Broker to schedule pilots
-% instead of MPI scripts.\mtnote{Rewrite this after checking it with
-% Andre.}
-% \jhanote{In the interest of space and clarity, I would  remove the last
-% sentence of this para. And merge the first sentence with the para
-% above.}\mtnote{Done.}
-
-% The lifespan of pilots has four states distributed among the PilotManager,
-% resource, and pilot instance.% (Fig.~\ref{fig:pilot-state-model}).
-% Pilots are instantiated in the state \T{NEW} by the PilotManager, wait in a
-% queue to be launched, and transition to \T{PM\_LAUNCH} when submitted to a
-% Resource Manager (RM) via the SAGA API\@. Pilots wait in the queue of the RM
-% and, once scheduled, become \T{P\_ACTIVE}. They remain in this state until the
-% end of their lifetime, when they transition to \T{DONE}.
-
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=\columnwidth]{figures/pilot_state_model_compact_rp_paper.pdf}
-%  \caption{Pilot State Model. Instantiated in state \T{NEW}, each pilot is
-%    launched (\T{PM\_LAUNCH}) via SAGA API on a resource manager. After
-%    becoming \T{P\_ACTIVE}, it exhausts its duration ending in state
-%    \T{DONE}. The transition between states is sequential and each transition
-%    can be canceled or fail ending in the states \T{CANCELED} or \T{FAILED} (not
-%    depicted to improve the diagram clarity). Boxes color coding as per
-%    Fig.~\ref{fig:arch-overview}, with: orange for queues and blue for states.
-%\label{fig:pilot-state-model}}
-%\end{figure}
-%
-
-% The unit state model has nine states distributed across the UnitManager,
-% MongoDB instance, and Agent.% (Fig.~\ref{fig:unit-state-model}).
-% Instantiated in the state \T{NEW} by the UnitManager, every unit is scheduled
-% on an Agent (\T{UM\_SCHEDULING}) via a queue on a MongoDB instance. The unit
-% is then scheduled on the required number of cores held by the Agent's pilot
-% (\T{A\_SCHEDULING}), and finally executed (\T{A\_EXECUTING}).
-
-% The unit state model pertains also to the input and output data of the units.
-% When required, the input data of a unit are either pushed to the Agent
-% (\T{U\_STAGING\_IN}) or pulled from the Agent (\T{A\_STAGING\_IN}), depending
-% on data locality and sharing requirements. Similarly, the output data of the
-% unit are staged out by the Agent and UnitManager (\T{A\_STAGING\_OUT},
-% \T{U\_STAGING\_OUT}) to a specified destination, e.g., the user workstation.
-% Both input and output staging are optional, depending on the requirements of
-% the units. The actual file transfers are enacted via SAGA, and support
-% (gsi)-scp, (gsi)-sftp, and Globus Online.
-
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=\columnwidth]{figures/unit_state_model_compact_rp_paper.pdf}
-%  \caption{Unit State Model. Instantiated in state \T{NEW}, each unit is
-%    scheduled to an Agent (\T{UM\_SCHEDULING}) via MongoDB, and then scheduled
-%    to an Agent's Executer (\T{A\_SCHEDULING}). When required, unit input data
-%    are staged to the Agent's Pilot (\T{UM\_STAGING\_IN} or
-%    \T{A\_STAGING\_IN}). Each unit is executed (\T{A\_EXECUTING}), unit output
-%    data are staged out (\T{A\_STAGING\_OUT} and \T{UM\_STAGING\_OUT}), and
-%    each unit ends in state (\T{DONE}). Color coding and state omissions
-%    as per Fig.~\ref{fig:pilot-state-model}.
-%\label{fig:unit-state-model}}
-%\end{figure}
-
-%The state transitions represented in Figures~\ref{fig:pilot-state-model}
-%and~\ref{fig:unit-state-model} are sequential and every transition can fail or
-%be canceled by the PilotManager or UnitManager.
-
-% All state transitions are managed by the PilotManager, UnitManager, and Agent
-% components. The only special case is the transition of the pilots to the state
-% \T{P\_ACTIVE} which is dictated by the resource's RM, but managed by the
-% PilotManager.
-
-% Above all NGE's components, the Agent is the one that is worth to mention the
-% most.
 
@@
 tasks.
-
-% ORTE transparently provides support for interprocess communication, resource
-% discovery and allocation, and process launching across a variety of platforms.
-% ORTE provides a mechanism similar to the Pilot concept - it allows the user to
-% create a \emph{``dynamic virtual machine''} (DVM) that spans multiple nodes.
-% ORTE provides libraries to enable the submission, monitoring and managing of
-% tasks, avoiding filesystem bottlenecks and race conditions with network sockets.
-% \aanote{Requires validation: Alternative version of the three sentences above
-% (too many ORTE in my opinion):}
-
-%Each component of RP has a independent semantic scope. This enables modularity
-%isolating implementation complexity and supporting diverse use cases and
-%environments.  For example, unit scheduling can be implemented by exchangeable
-%Scheduler components, suitable for applications of diverse scales, with
-%different coordination patterns, and executed on Beowulf clusters or Cray
-%machines.
-%
-%Components are also designed to be stateless and instantiated concurrently. In
-%this way, RP can manage multiple pilots and units at the same time, resulting
-%in scalable throughput and tolerance to failing components. Concurrent
-%components are coordinated via a dedicated communication mesh which incurs
-%infrastructure and runtime overhead, offset by the lower component complexity
-%and improved overall scalability of the system.
-
-% \begin{enumerate}
-%     \item Rationale and design (why)
-%     \item Architecture (how)
-%     \item Integration (Not present)
-%     \item Characterization (experiments)
-% \end{enumerate}
-
- % \ldots

### 2017-04-04 overall.tex
Notes clean-up
Clean up

@@
 
-% two dimensions: (i) backfill utilization by the PanDA Brokers; and (ii) the
-% amount of runtime spent performing detector simulations.
-
-% We based our characterizations on measuring the number of cores utilized by
-% ATLAS on Titan, the overall backfill availability, the number of detector
-% simulations performed on all the resources available to ATLAS and on Titan,
-% and the failure rate of these simulations. We also consider the the time taken
-% to setup AthenaMP and to simulate events in the ATLAS detector with it. All
-% the measurements were performed between January 2016 and February 2017,
-% hereafter called `experiment time window'.
-
-% \mtnote{TODO: expand the previous paragraph to include dimensions of AthenaMP
-% characterization.}
-
 % -----------------------------------------------------------------------------
@@
 \label{ssec:broker_titan}
-
-% Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of
-% core-hours used by ATLAS during the experiment time window. ATLAS consumed a
-% total of 73.8M core-hours, for an average of $\approx$7M core-hours a month,
-% with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours
-% in February 2017.
 
@@
 
-Figure~\ref{fig:backfill-utilization} shows the total backfill availability on
+Fig.~\ref{fig:backfill-utilization} shows the total backfill availability on
 Titan (gray bars) and the number of core-hours of that availability used by
@@
     \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.pdf}
-
     \caption{Titan's total backfill availability: CPU core-hours (gray) and CPU
@@
     of total Titan's backfill availability used by ATLAS (Red labels).}
-
 \label{fig:backfill-utilization}
@@
 
-PanDA Brokers' efficiency (Figure~\ref{fig:backfill-utilization}, red labels) is
+PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels) is
 defined as the fraction of core-hours utilized by the PanDA Brokers to  Titan’s
@@
 reached 18\% average efficiency, with a minimum 7.8\% efficiency on May 2016 and
-a maximum 30.9\% efficiency on February 2017.
-% The total backfill available was 38.1M Titan core-hours in April 2016, and
-% 33.1M Titan core-hours in February 2017. This shows that the measured
-% difference in efficiency did not depend on a comparable difference in total
-% backfill availability.
-The total backfill availability was 20.3M in April 2016, and 17.6M in February
-2017. This shows that the measured difference in efficiency did not depend on a
-comparable difference in the total backfill availability.
-
-% \mtnote{Elaborate further and put it at the end, after giving several of the
-% measurements we would need. ``Currently, this measure of efficiency is being
-% refined on the base of the actual minimum amount of walltime that PanDA Broker
-% can request when submitting job requests. This is not a fixed value as it
-% depends on variable initialization and configuration overheads, simulation
-% time, and number of nodes obtained. Part of the undergoing effort is to
-% enable the measurements of  the distribution of these parameters to derive an
-% optimization function.''}
-
-% \jhanote{I would plot the efficiency on the Y2 axis for both
-% [diagrams].}\mtnote{Done but then the first diagram ended up plotting a subset
-% of the data of the second diagram. I used only the second diagram.} \jhanote{I
-% would suggest similar style of X-axis as previous figure. Difficult to parse
-% which is April if we are going to call out a specific month. Also consistency
-% in style is generally good to keep cognitive burden low.}\mtnote{Done.}
-
-% Figure~\ref{fig:hpc-workload-utilization} shows that d
+a maximum 30.9\% efficiency on February 2017. The total backfill availability
+was 20.3M in April 2016, and 17.6M in February 2017. This shows that the
+measured difference in efficiency did not depend on a comparable difference in
+the total backfill availability.
 
@@
 0.9\% of all the 250M detector simulations performed by ATLAS in the same period
-of time, and 3.5\% of the 6.6B events processed by those jobs.
-% Comparatively, Titan contributed 3.9\% of the total of around
-% 200K\mtnote{TODO: Fix this number} cores available to ATLAS on Grid, Cloud,
-% and HPC infrastructures together.
-These figures confirms the relevance of supercomputers' resource contribution to
-the LHC Run 2, especially when accounting for the amount of unused total
-backfill availability and the improvement of PanDA efficiency across the
-experiment time window.
-
-% \begin{figure}[htp]
-%     \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
-% \caption{\mtnote{Placeholder for the diagram/data I asked to Danila and
-% Sergey. Consider a table instead of a figure.}}
-% \label{fig:hpc-workload-utilization}
-% \end{figure}
+of time, and 3.5\% of the 6.6B events processed by those jobs. These figures
+confirms the relevance of supercomputers' resource contribution to the LHC Run
+2, especially when accounting for the amount of unused total backfill
+availability and the improvement of PanDA efficiency across the experiment time
+window.
 
@@
 function of the number of events simulated but not all events take the same time
-to be simulated.
-% Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time
-% of
-1 event simulation takes from $\approx$2 to $\approx$40 minutes, with a mean of
-$\approx$14 minutes. Considering that each worker node process up to 16 events
-concurrently, 100 events takes an average of 105 minutes to process. As such,
-PanDA brokers do not use backfill availability with less than 105 minutes
-walltime.
+to be simulated. 1 event simulation takes from $\approx$2 to $\approx$40
+minutes, with a mean of $\approx$14 minutes. Considering that each worker node
+process up to 16 events concurrently, 100 events takes an average of 105 minutes
+to process. As such, PanDA brokers do not use backfill availability with less
+than 105 minutes walltime.
 
-% \begin{figure}[htp]
-%     \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
-%     \caption{Distribution of the time taken by a Geant4 detector simulation to
-%     simulate one event. 3000 Events; 30 Titan worker nodes; 16 simulations per
-%     node; 100 events per node. Average time per event $\approx$14 min. Broad
-%     distribution from $\approx$2 to $\approx$40 minutes.}
-% \label{fig:100event-distrib}
-% \end{figure}
-
-Figure~\ref{fig:backfill-distrib} shows the distribution of backfill
+Fig.~\ref{fig:backfill-distrib} shows the distribution of backfill
 availability on Titan as a function of number of nodes and the time of their
@@
     \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
-    %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
     \caption{Backfill availability: distribution of number of work nodes (left);
@@
 
-% The number of concurrent detector simulation jobs that PanDA Broker can manage
-% depends on the DTNs' resources. Each PanDA Broker can execute 1 MPI PBS job on
-% up to 300 worker nodes for a total of 4,800 cores. Scaling above this
-% threshold requires instantiating multiple PanDA Brokers on the same DTN and,
-% once the DTN resources are saturated, on multiple DTNs.
-
 % -----------------------------------------------------------------------------
@@
 \label{ssec:athenamp_titan}
-
-% \begin{itemize}
-%     \item problem with lustre addressed with ramdisk (done)
-%     \item Competition for cache in AMD processors. (done)
-%     \item Comparison hardware performance Titan/Grid (data acquired)
-%     \item Summary overall figure of performance analogous to those produced
-%     with NGE. Diagram as discussed in F2F meeting at Rutgers. (waiting)
-% \end{itemize}
 
@@
 
-% AthenaMP is a multipurpose framework that needs to be configured depending on
-% the type of payalod that needs to be executed. For Geant4, this configuration
-% process links up to 200 libraries, all requiring filesystem read and write
-% operations.
-
-% The execution of Geant4 is mostly compute-intensive requiring to write around
-% XXKB per 100 events on disk and no more than 2GB of memory.
-
-% -----------------------------------------------------------------------------
-% \subsection{PanDA Shared Library I/O Performance Impact at OLCF}
-
-% Athena, the ATLAS framework has a configuration and initialization stage. At
-% this stage, the running job is assembled on the fly from a large number of
-% shared libraries. Also, at this stage, every algorithm and service is being
-% configured, at run time, by a corresponding set of Python scripts, which
-% results in a large number of read operations accesses to small python
-% scripts, with many includes and imports Python calls.
-
-% \mtnote{TODO: specify that this is work in progress (see Sergey's comment) and
-% see how to move it to the data performance subsection}
-
 Initially, all the shared libraries of AthenaMP and the python scripts for the
@@
 I/O patterns of the initialization and configuration stages degraded the
-performance of the filesystem.
-% (Figure~\ref{}\mtnote{We may want to aggregate some of the diagrams about
-% lustre's experiments here}).  Since Spider 2 is a center-wide file system,
-% this resulted in  performance degradation for all OLCF resources and users.
-% \mtnote{I am afraid the details about the trace are too specific given the
-% space constraints of a SC  submission. Please feel free to uncomment it if
-% you disagree.}
-%
-% As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
-% exhibits a spike corresponding to the beginning of the runs before tapering
-% off.
-%
-% \begin{minipage}{\linewidth}
-% \begin{lstlisting}[language=bash,frame=single,basicstyle=\ttfamily\tiny,caption=ATLAS metadata trace,label=mdstrace]
-% XK7 Application 9205593
-%       39012 RPCs from 300 of 300 nodes
-%         ~69291.96 per sec
-%           37851 LDLM_ENQUEUE RPCs    ~67229.83 per sec
-%                 pmin 13us pavg 42us pmax 4983us
-%           611 LDLM_CANCEL RPCs    ~1085.24 per sec
-%                 pmin 10us pavg 16us pmax 32us
-%           277 MDS_CLOSE RPCs    ~492.00 per sec
-%                 pmin 15us pavg 20us pmax 38us
-%           154 MDS_READPAGE RPCs    ~273.53 per sec
-%                 pmin 170us pavg 292us pmax 671us
-%           86 MDS_GETXATTR RPCs    ~152.75 per sec
-%                 pmin 15us pavg 21us pmax 65us
-%           30 MDS_GETATTR RPCs    ~53.29 per sec
-%                 pmin 16us pavg 20us pmax 27us
-%           3 MDS_REINT RPCs    ~5.33 per sec
-%                 pmin 103us pavg 136us pmax 196us
-%           Overall times
-%                 pmin 10us pavg 42us pmax 4983us
-%
-% XK7 Application 9205355
-%       8698 RPCs from 62 of 62 nodes
-%         ~15449.13 per sec
-%           8445 LDLM_ENQUEUE RPCs    ~14999.76 per sec
-%                 pmin 16us pavg 41us pmax 780us
-%           92 MDS_CLOSE RPCs    ~163.41 per sec
-%                 pmin 15us pavg 23us pmax 52us
-%           55 MDS_READPAGE RPCs    ~97.69 per sec
-%                 pmin 189us pavg 291us pmax 534us
-%           52 LDLM_CANCEL RPCs    ~92.36 per sec
-%                 pmin 12us pavg 16us pmax 25us
-%           41 MDS_GETXATTR RPCs    ~72.82 per sec
-%                 pmin 16us pavg 20us pmax 27us
-%           13 MDS_GETATTR RPCs    ~23.09 per sec
-%                 pmin 16us pavg 21us pmax 38us
-%           Overall times
-%                 pmin 12us pavg 42us pmax 780us
-%
-% \end{lstlisting}
-% \end{minipage}
-%
-% As this problem was identified, the OLCF staff has enabled read-only access to
-% certain NFS-exported directories from Titan compute nodes. This in turn
-% allowed the OLCF staff to install a software package from a Titan login node
-% and have it available read-only on a Titan compute node.
-%
-% After careful characterization of the impact of ATLAS jobs on Lustre, the
-% performance degradation
-This was addressed by moving the AthenaMP distribution to a read-only NFS
-directory, shared among DTNs and worker nodes. NFS eliminated the problem of
-metadata contention, improving metadata read performance from $\approx$6,300
-seconds on Lustre to $\approx$1,500 seconds on NFS.
-
-% Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
-% improvement on Titan. The circled region illustrates the switch from Lustre to
-% NFS-exported directory for hosting the ATLAS release.
-
-%\begin{figure}[!htb]
-%    \centering
-%    \begin{tabular}{cc}
-%        {\includegraphics[width=0.48\textwidth]{figures/panda-completed-jobs-sw-move.pdf}}\\
-%    \end{tabular}
-%    \caption{ATLAS performance improvement on Titan. The circled region shows the switch from Lustre to NFS-exported directory for hosting the ATLAS release.}
-%\label{fig:atlas-perf-improvement}
-%\end{figure}
+performance of the filesystem. This was addressed by moving the AthenaMP
+distribution to a read-only NFS directory, shared among DTNs and worker nodes.
+NFS eliminated the problem of metadata contention, improving metadata read
+performance from $\approx$6,300 seconds on Lustre to $\approx$1,500 seconds on
+NFS.
 
@@
 competes for a single FP scheduler. This creates the overhead shown in
-Figure~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
+Fig.~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
 concurrent simulations computing 50 events is 10.8 minutes, while for 16
@@
 of executing the same detector simulation on both Titan or Grid sites.
-% Figure~\ref{fig:comparison-titan-grid} compares the efficiencies of Titan to
-% the BNL and SIGNET Grid sites, normalized for 8 cores. Effective performance
-% per-core at Titan is $\approx$0.57 event per minute, roughly 1/2 of BNL and
-% 1/3 of SIGNET performances.
 Comparisons of the efficiencies of Titan to  the BNL and SIGNET Grid sites,
@@
 performances.
-
-% \begin{figure}[htp]
-%     \includegraphics[clip,width=\columnwidth]{figures/titan_grid_event_rate.pdf}
-%     \caption{Comparison of the event rate for finished jobs at Titan (h1), BNL
-%     (h2) and GRIDNET (h3). Titan histogram shows 1/2 of the measured values to
-%     account for the core difference between Titan (16) and the Grid sites (8).
-%     Performance per core: BNL 1.126; SIGNET 0.885; Titan 0.577.\mtnote{TO BE DELETED.}}
-% \label{fig:comparison-titan-grid}
-% \end{figure}
 
@@
 
-% \mtnote{Figure summarizing job performance on Titan analogous to those
-% produced with NGE, as discussed in F2F meeting at Rutgers?}
-%
-% ----------------------------------------------------------------------------
-% \subsection{Reliability of PanDA Broker and Detector Simulations on Titan}
-% \label{ssec:reliability}
-%
-% Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
-% recorded by the PanDA Brokers between January 2016 and February 2017. The
-% initial preponderance of unknown failures was progressively addressed by a
-% more accurate failure model, specifically tailored to the PanDA Broker and
-% Titan. Figure~\ref{fig:failures-titan} also shows the progressive improvement
-% of the broker's reliability: In Februrary 2017, errors related to input files
-% and initialization were almost eliminated and execution errors were comparable
-% to those recorded in January 2017 even if almost twice as many simulations
-% were performed (Figure~\ref{fig:backfill-utilization}). The increased amount
-% of unknown failure in February indicates that furter work is needed to improve
-% the PanDA Broker failure model.
-%
-% \begin{figure}[htp]
-%     \includegraphics[clip,width=\columnwidth]{figures/failures_titan.pdf}
-    % \caption{PanDA Broker failures during the experiment time window. Failures
-    % were recorded with 52 exit codes, here aggregated into 6 categories. The
-    % number of failures should be compared to the number of simulations
-    % executed in each month as plotted in
-    % Figure~\ref{fig:backfill-utilization}. Improvements in the failure model
-    % of PanDA Broker shows a decrease in unknown failure and a the progressive
-    % addressing of other categories of failure.}
-% \label{fig:failures-titan}
-% \end{figure}
-%
-% -----------------------------------------------------------------------------
-% \subsection{Characterizing PanDA I/O}
-%
-% To better understand the I/O impact of ATLAS PanDA project on Titan
-% supercomputing environment
-
 We studied the impact of acquiring event descriptions on Lustre by analyzing
@@
 Table~\ref{panda-olcf-stats} shows the overall statistical breakdown of the
-observed file I/O.
-% Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written}
-% show the file read and write I/O histograms for these 1,175 jobs,
-% respectively. Figures~\ref{fig:atlas-titan-file-open}
-% and~\ref{fig:atlas-titan-file-close} show the file $open()$ and $close()$
-% metadata load histograms of the same  1,175 ATLAS jobs, respectively.
-%
-% As can be seen from Table~\ref{panda-olcf-stats},
-%
-ATLAS used between 1 and 300 worker nodes, and 35 on average. 75\% of the jobs
-run by ATLAS consumed less than 25 nodes and 92\% less than 100. During the 174
-hours of data collection, 6.75 ATLAS jobs were executed on average per hour,
-each job running for an average of 1.74 hours.
-% ATLAS jobs issues a large number of file read operations, as can be seen in
-% Table~\ref{panda-olcf-stats}.
-Every job read less than 250 GB and wrote less than 75 GB of data and, on
-average, each job read 20 GB and wrote 6 GB of data.
+observed file I/O. ATLAS used between 1 and 300 worker nodes, and 35 on average.
+75\% of the jobs run by ATLAS consumed less than 25 nodes and 92\% less than
+100. During the 174 hours of data collection, 6.75 ATLAS jobs were executed on
+average per hour, each job running for an average of 1.74 hours. Every job read
+less than 250 GB and wrote less than 75 GB of data and, on average, each job
+read 20 GB and wrote 6 GB of data.
 
@@
 
-% \begin{figure}[!htb]
-%     \centering
-%     \begin{tabular}{cc}
-%         {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_finer_hist.pdf}}\\
-%     \end{tabular}
-%     \caption{ATLAS file read operation histogram on Titan for week of 10/25/16.}
-% \label{fig:atlas-titan-io-read}
-% \end{figure}
-%
-% \begin{figure}[!htb]
-%     \centering
-%     \begin{tabular}{cc}
-%         {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_finer_hist.pdf}}\\
-%     \end{tabular}
-%     \caption{ATLAS file write I/O histogram on Titan for week of 10/25/16.}
-% \label{fig:atlas-titan-io-written}
-% \end{figure}
-
-% The I/O of the ATLAS jobs show an interesting pattern.
-% Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
-% worker node is less than 400 MB on average, while the amount of data written
-% per node is less than 170 MB on average. This correlates with our finding
-% that ATLAS PanDA jobs are read heavy. However,
-
 ATLAS jobs are read heavy: On average, the amount of data read per worker node
 is less than 400 MB, while the amount of data written is less than 170 MB.
-% as can be seen in Table~\ref{panda-olcf-stats} and figures
-% {fig:atlas-titan-io-read} and {fig:atlas-titan-io-written},
 Distributions of read and written data are different: The read operation
@@
 
-% \begin{figure}[!htb]
-%     \centering
-%     \begin{tabular}{cc}
-%         {\includegraphics[width=0.48\textwidth]{figures/panda_file_open_hist.pdf}}\\
-%     \end{tabular}
-%     \caption{ATLAS file $open()$ histogram on Titan for week of 10/25/16.}
-% \label{fig:atlas-titan-file-open}
-% \end{figure}
-%
-% \begin{figure}[!htb]
-%     \centering
-%     \begin{tabular}{cc}
-%         {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
-%     \end{tabular}
-%     \caption{ATLAS file $close()$ histogram on Titan for week of 10/25/16.}
-% \label{fig:atlas-titan-file-close}
-% \end{figure}
+The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
+operations per second (not including file $stat()$ operations) and 5 file
+$close()$ operations per second, with similar distributions. On average, the
+maximum number of file $open()$ operations per job is $\approx$170/s and the
+maximum number of file $close()$ operations is $\approx$39/s. For the 1,175
+ATLAS jobs observed, the total number of file $open()$ operations is 172,089,760
+and the total number of file $close()$ operations is 40,132,992. The difference
+between these two values is still under investigation: One possible explanation
+is that ATLAS jobs don't call a file $close()$ operation per every file $open()$
+issued.
 
-The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
-operations per second (not including file $stat()$ operations) and 5 file $close()$ operations per second, with similar distributions.
-% As can be seen from figures~\ref{fig:atlas-titan-file-open}
-% and~\ref{fig:atlas-titan-file-close}, they .
-On average, the maximum number of file $open()$ operations per job is
-$\approx$170/s and the maximum number of file $close()$ operations is
-$\approx$39/s.
-% on average per second per job.
-For the 1,175 ATLAS jobs observed, the total number of file $open()$ operations
-is 172,089,760 and the total number of file $close()$ operations is 40,132,992.
-The difference between these two values is still under investigation: One
-possible explanation is that ATLAS jobs don't call a file $close()$ operation
-per every file $open()$ issued.
-
-% based on our experiments with ATLAS jobs, it can be safely concluded
-% that the file and metadata I/O load of ATLAS PanDA project on the OLCF Titan
-% supercomputing environment and the Spider 2 file system is not detrimental to
-% the center operations. At the current scale of the project, the overall impact
-% of ATLAS operations on Titan is minimal.
-
-Overall, the average time taken to read events from input files stored on Lustre is 1,320, comparable to the time taken to read the file required by assembling AthenaMP from NFS.
-Preliminary investigation shows that this time could be reduced to 40 seconds by loading the event descriptions into the RAM disk available on each worker node.
-Events descriptions could be transferred from Lustre to the RAM disk while configuring and initializing AthenaMP, almost halving the time currently required by initiating a Genat4 simulation.
+Overall, the average time taken to read events from input files stored on Lustre
+is 1,320, comparable to the time taken to read the file required by assembling
+AthenaMP from NFS. Preliminary investigation shows that this time could be
+reduced to 40 seconds by loading the event descriptions into the RAM disk
+available on each worker node. Events descriptions could be transferred from
+Lustre to the RAM disk while configuring and initializing AthenaMP, almost
+halving the time currently required by initiating a Genat4 simulation.
 

### 2017-04-04 titan.tex
Clean up

--- 2017-03-05 titan.tex
@@
 compute clients and other OLCF resources.
-
-The topology and architecture details of Spider II infrastructure are
-illustrated in Figure~\ref{fig:xk7-compute-node} and
-described as follows:
-
-\begin{figure}[!htb]
-    \centering
-    \begin{tabular}{cc}
-        {\includegraphics[width=0.48\textwidth]{figures/spider2arch.pdf}}\\
-    \end{tabular}
-    \caption{Infrastructure and I/O path between Titan (Cray XK7) and its backend storage}
-    \label{fig:xk7-compute-node}
-\end{figure}
-
-Spider II is based on the Lustre technology~\cite{lustre}.  It is configured
-and deployed as two independent, non-overlapping file systems, each with 144
-Lustre Object Storage Servers (OSSs) and 1,008 Lustre Object Storage Targets
-(OSTs). Spider II offers ~30 PB usable file system capacity to OLCF systems and
-users and has over 1 TB/s aggregate write I/O performance.

### 2017-04-04 workflow.tex
Clean up

@@
 
-\paragraph{Figure~\ref{fig:atlas_workflow} (1)} Multiple end-users (i.e.,
+\paragraph{Fig.~\ref{fig:atlas_workflow} (1)} Multiple end-users (i.e.,
 physicists) submit descriptions of ATLAS Monte Carlo Workflows to the JEDI
@@
 
-\paragraph{Figure~\ref{fig:atlas_workflow} (2)} JEDI submits/hands over events
+\paragraph{Fig.~\ref{fig:atlas_workflow} (2)} JEDI submits/hands over events
 to a PANDA Server. This server manages the execution of all the ATLAS workflows
 and their stages, including the ATLAS Monte Carlo Workflow. In
-Figure~\ref{fig:atlas_workflow} we depicted only how the events of the third
+Fig.~\ref{fig:atlas_workflow} we depicted only how the events of the third
 stage are organized within PANDA Server. Tasks are container of jobs and each
@@
 
-\paragraph{Figure~\ref{fig:atlas_workflow} (3)} Sets of events belonging to
+\paragraph{Fig.~\ref{fig:atlas_workflow} (3)} Sets of events belonging to
 multiple tasks and jobs are pulled from the PANDA Server by one or more PANDA
@@
 
-\paragraph{Figure~\ref{fig:atlas_workflow} (4)} AthenaMP is used to perform the
+\paragraph{Fig.~\ref{fig:atlas_workflow} (4)} AthenaMP is used to perform the
 simulation of the events within the ATLAS detector, one AthenaMP thread for each
@@
 node, each thread being able to simulate an average of 1 event every n??
-minutes?/second? (see Figure~\ref{distribution_amp_titan}). So far, no
+minutes?/second? (see Fig.~\ref{distribution_amp_titan}). So far, no
 description has been given of the flow of data required to simulate each event.
-
-% Questions for Danila/Sergey:
-% \begin{itemize}
-%     \item Does each job of a PANDA Server (and therefore all the tasks of that specific job) belong to the same `user'/VO\@?
-%     \item Does a PANDA Pilot pull jobs/tasks from a PANDA Server or does a PANDA server push jobs/tasks to a PANDA Pilot?
-%     \item Does a PANDA Pilot pull jobs or tasks from a PANDA Server? Are the jobs within a PANDA Pilot the same jobs (i.e., the same container of the same tasks) in the PANDA Server?
-%     \item Do we have 16 PANDA Pilots on a Titan DSL (is this the right acronym?) node? How many DSL nodes?
-%     \item Can the events packaged into a PBS job belong to different submissions, i.e., different jobs, i.e., different users?
-%     \item Do we have the distribution of the time taken on titan to process 1 event of the Monte Carlo workflow?
-% \end{itemize}
-
-% Notes - answers
-%
-% - Does each job of a PANDA Server (and therefore all the tasks of that
-%   specific job) belong to the same ‘user’/VO? • Does a PANDA Pilot pull
-%   jobs/tasks from a PANDA Server or does a PANDA server push jobs/tasks to a
-%   PANDA Pilot?
-%
-% Same user task->job on the server.
-%
-% - Does a PANDA Pilot pull jobs or tasks from a PANDA Server? Are the jobs
-%   within a PANDA Pilot the same jobs (i.e., the same container of the same
-%   tasks) in the PANDA Server?
-%
-% Panda pilot is a edge service on titan, event-driven by backfill
-% availability. When available, it DOES THE PULL from the panda server.
-%
-% - Do we have 16 PANDA Pilots on a Titan DTN (is this the right acronym?)
-%   node? How many DTN nodes?
-%
-% Currently 20, thinking to go higher, on 4 DTN (Data Transfer Node), 5 pilots
-% for DTN, DTN that enable interactive data transfer. New DTN farm is being
-% rolled on. Our set up will change. More node, more pilots per node. This is
-% bound by DTN performance, overloading DTN (currently absence of data on why
-% these nodes are overladed. Network/memory should not be a problem, it is
-% probably purely CPU because of tools used to move files, e.g., GridFTP.)
-%
-% - Can the events packaged into a PBS job belong to different submissions,
-%   i.e., different jobs, i.e., different.
-%
-% They do belong to different tasks.
-%
-% - Do we have the distribution of the time taken on titan to process 1 event
-%   of the Monte Carlo workflow?
-%
-% To be added to Section VI about performance. See presentation:
-% https://drive.google.com/drive/u/0/folders/0BxWZQExiTsWNRTdMQXN4a1g3bHc
 

### 2017-04-04 conclusion.tex
fixing BW head-2-head comparision.

@@
 
-Thanks to the efforts captured in this paper, the time-averaged utilization of
-Titan has consistently been upwards of 90\%. Compare this to the time averaged
-utilization of other similar leadership machines such as NSF's flagship Blue
-Waters, where the average utilization fluctuates between 60-80\% (XDMoD)
-machine~\cite{bw-sucks}. These
-% impressive if not
-unprecedented efficiency gains aside, this work is just a starting point towards
-more effective operational models for future leadership and online analytical
-platforms~\cite{foap-url}. These platforms will have to support ever increasing
-complex workloads with varying models for dynamic resource federation. Recent
-efforts with Google Cloud for the CMS experiment --
-HEPCloud~\cite{hepcloud,googlehep} represent similar yet different approaches
-to dynamic resource federation.
+The 52M core hours used by ATLAS, via PanDA, is over 2\% of the total
+utilization on Titan over the same period, bringing the time-averaged
+utilization of Titan to be consistently upwards of 90\%. Given that the
+average average utilization of most other leadership class machines is less
+(e.g., NSF's flagship Blue Waters the average utilization fluctuates between
+60-80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
+approaches elsewhere. These unprecedented efficiency gains aside, this work is
+just a starting point towards more effective operational models for future
+leadership and online analytical platforms~\cite{foap-url}. These platforms
+will have to support ever increasing complex workloads with varying models for
+dynamic resource federation. Recent efforts with Google Cloud for the CMS
+experiment -- HEPCloud~\cite{hepcloud,googlehep} represent similar yet
+different approaches to dynamic resource federation.
 

### 2017-04-04 Experiments.tex

@@
-% In the batch queue,  queue time becomes dominant but, at the same time, we
-% have more freedom to decide the parameters of the slot.
-
-\subsection{Experiments}\label{sec:ngeExp}
+\subsection{Experiments}
+\label{sec:ngeExp}
 
@@
 walltime. The time duration is determined by the Titan's walltime policy.
-Figure~\ref{fig:weakScal1a} depicts the average pilot duration, the average
+Fig.~\ref{fig:weakScal1a} depicts the average pilot duration, the average
 execution time of AthenaMP, and the NGE pilot overhead as function of the pilot
@@
 
-%\begin{figure}[!htb]
-%        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakOver1.pdf}
-%    \caption{Weak scalability: average overhead by running AthenaMP on 200, 500, 1000 and 2000 nodes.}
-%\label{fig:weakScal1b}
-%\end{figure}
-
 \subsubsection{Weak scalability with multiple generation }
@@
 other tasks are still running.
-
-% \mtnote{we dowe need to keep it below 2 hours?}\aanote{You are right. It was misleading. I Changed. Check if you like it.}
 
@@
 
-% \footnote{Note that the number of events has been chosen in such a way
-% that all the cores of a node run one event.}
-%In this way, every core of each worker node was used to simulate one event,
-%. This allows us to complete
-%reducing the running time of each AthenaMP to ~$20$ minutes. % only.
-
 We ran pilots with 256, 512, 1024 and 2048 worker nodes and 3\mtnote{2?} hours
-walltime. Figure~\ref{fig:weakScal2a} depicts the average pilot duration, the
+walltime. Fig.~\ref{fig:weakScal2a} depicts the average pilot duration, the
 average execution time of five sequential generations of AthenaMP, and the
@@
 
-%\begin{figure}[!htb]
-%        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakOver2.pdf}
-%    \caption{Weak scalability with multiple generations: average overhead by running AthenaMP on 256, 512, 1024 and 2048 nodes.}
-%\label{fig:weakScal2b}
-%\end{figure}
-
 \subsubsection{Strong scalability}
@@
 
-% \mtnote{Not sure I understand this sentence}\aanote{My fault. It should
-% have been seven and null or 8 and 1. It depends on how we consider the word
-% sequential. anyway, to avoid confusion I changed the sentence.}.
-
-Figure \ref{fig:strongScala}  shows the average pilot duration and the average
+Fig.~\ref{fig:strongScala}  shows the average pilot duration and the average
 execution time of possibly sequential AthenaMP instances.  We  notice that the
@@
 \end{figure}
-
-%
-%\begin{figure}[!htb]
-%        \includegraphics[width=0.4\textwidth]{./figures/NGE/strongOver.pdf}
-%    \caption{Strong scalability: average overhead running AthenaMP on 256, 512, 1024 and 2048 nodes.}
-%\label{fig:strongScalb}
-%\end{figure}
-
-%\subsubsection{Heterogeneous execution}
-%
-%The last experiment provides a proof of concept about the ability of the NGE to execute heterogeneous workload.
-%In particular, we coupled the execution of AthenaMP with the execution of Gromacs to simulate molecular dynamics.
-%We performed the experiment by executing at first one AthenaMP per node and then, submitting a Gromacs simulation per core. Each Gromacs simulation requires ~20 minutes.
-%We tested the following pilot sizes : 8, 16, 32, 64.
-%
-%\begin{figure}[!htb]
-%        \includegraphics[width=0.5\textwidth]{./figures/NGE/MDET.pdf}
-%    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
-%\label{fig:strongScala}
-%\end{figure}
-%\begin{figure}[!htb]
-%        \includegraphics[width=0.5\textwidth]{./figures/NGE/MDOver.pdf}
-%    \caption{Average overhead running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
-%\label{fig:MDScal}
-%\end{figure}
-
-% For this reason, the second set of the experiments aims to find
-%sub-optimal parameters with which we can minimize the trade-off between the size
-%of a slot and the time spent in queue waiting for that slot to become available.
-%In other words, we aim to minimize the completion time by finding the best
-%trade-off between execution time and queue time.
-%
-%This execution model introduces slot utilization as one of the key factors for
-%high-performances. This happens because, in order to minimize the time spent in
-%queue, we might asks for slots in advance and, then we could not be able to
-%saturate them when they become available. Thus, this strategy requires a new
-%functionality that allows the job to receive and execute new events while it is
-%already running on the resources. In order to do that we perform the experiments
-%by using a new generation executor that implements such functionality.
-%
-%As last observation, it is important to point out that the percentage of
-%utilization of a slot is minor problem with the current implementation because,
-%due to the dynamics of the Backfill queue, PanDA has a high probability to
-%re-acquire a slot immediately after it has released one\aanote{Are we able to
-%quantify this ``immediately''?}.

### 2017-04-04 introduction.tex
Iteration introduction

--- 2017-04-02 introduction.tex
@@
 properties of matter for the next decades.   Multiple experiments at LHC have
-collected and distributed hundreds of petabytes of data worldwide to hundreds
-of computer centers. Thousands of physicists analyze petascale data volumes
-daily. The detection of the Higgs Boson in 2013 speaks to the success of the
-detector and experiment design, as well as the sophistication of computing
-systems devised to analyze the data.
+collected and distributed hundreds of petabytes of data worldwide to hundreds of
+computer centers. Thousands of physicists analyze petascale data volumes daily.
+The detection of the Higgs Boson in 2013 speaks to the success of the detector
+and experiment design, as well as the sophistication of computing systems
+devised to analyze the data.
 
-Historically the computing systems used by LHC experiments consisted of the
-federation of a hundreds to thousands of distributed resources \textemdash{}
-ranging in scale from small to mid-size resource~\cite{foster2003grid}.
-Although the workloads to be executed are comprised of tasks that are
-independent of each other, the management of the distribution of 
-workloads across many heterogeneous resources to ensure the effective
-utilization of resources and efficient execution of workloads presents non-
-trivial challenges.
+Historically, the computing systems used by LHC experiments consisted of the
+federation of a hundreds to thousands of distributed resources, %\textemdash{}
+ranging in scale from small to mid-size resource~\cite{foster2003grid}. Although
+the workloads to be executed are comprised of tasks that are independent of each
+other, the management of the distribution of workloads across many heterogeneous
+resources to ensure the effective utilization of resources and efficient
+execution of workloads presents non-trivial challenges.
 
 Many software solutions have been developed in response to these challenges.
-One of the LHC experiments, the CMS experiment devised a solution based around
-the HTCondor~\cite{XX} software ecosystem. The ATLAS~\cite{Aad:2008}
-experiment, utilizes the Production and Distributed Analysis (PanDA) workload
-management system~\cite{Maeno2011} (WMS) for distributed data processing and
-analysis. The CMS and ATLAS experiments represent arguably the largest
-production grade distributed computing solutions and have symbolized the
-paradigm of {\it high-throughput computing} viz., the effective execution of
-many independent tasks.
+CMS, one of the LHC experiments, devised a solution based around the
+HTCondor~\cite{thain2005distributed} software ecosystem. The
+ATLAS~\cite{Aad:2008} experiment, utilizes the Production and Distributed
+Analysis (PanDA) workload management system~\cite{Maeno2011} (WMS) for
+distributed data processing and analysis. The CMS and ATLAS experiments
+represent arguably the largest production grade distributed computing solutions
+and have symbolized the paradigm of {\it high-throughput computing}, i.e., the
+effective execution of many independent tasks.
 
-As the LHC prepares for the high-luminosity era (Run 4) and Run 3 in
-$~\approx$ 2022, it is anticipated that the data volumes that will need
-analyzing will increase by factors of 10-100  compared to the current phase
-(Run 2). The data will be larger in volume but will also require more
-sophisticated computational processing. In spite of the impressive scale of
-the ATLAS distributed computing system, demand for computing systems will
-significantly outstrip supply (availability).
+As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity era
+(Run 4), it is anticipated that the data volumes that will need analyzing will
+increase by factors of 10-100 compared to the current phase (Run 2). Data will
+be larger in volume but will also require more sophisticated computational
+processing. In spite of the impressive scale of the ATLAS distributed computing
+system, demand for computing systems will significantly outstrip current and projected supply. %(availability).
 
-There are multiple levels at which this problem needs to be addressed
-urgently, e.g., the utilization of emerging parallel architectures (e.g.,
-platforms), algorithmic and advances in analytical methods (e.g., use of
-Machine Learning) and the ability to exploit different platforms (e.g., clouds
-and supercomputers).
+There are multiple levels at which this problem needs to be addressed: the
+utilization of emerging parallel architectures (e.g., platforms); algorithmic
+and advances in analytical methods (e.g., use of Machine Learning); and the
+ability to exploit different platforms (e.g., clouds and supercomputers).
 
-This paper is a case study of how the ATLAS experiment has "broken free" of
+This paper is a case study of how the ATLAS experiment has ``broken free'' of
 the traditional computational approach of high-throughput computing on
 distributed resources to embrace new platforms, in particular high-performance
-computers. Specifically, we discuss the experience of integrating the PanDA
-workload management system with Titan ~\textemdash{} a DOE leadership
-computing facility. Consequently, Titan now analyzes up to 5\% of workload  of
-the ATLAS experiment. The case study presents the design and integration of
-PanDA to reach sustained production scales (ca. 100 M core-hours a years). It
-also discusses the investigation of a task execution runtime system on Titan,
-that is based on  the pilot-abstractions and which allows advanced execution
-modes of the ATLAS workload as well as the enhanced support for heterogeneous
-workloads (such as molecular dynamics).
+computers. Specifically, we discuss the experience of integrating PanDA WMS with
+a DOE leadership computing facility called Titan to reach sustained production
+scales of approximately 51M core-hours a years. We also discusses the
+investigation of a task execution runtime system on Titan, that is based on the
+pilot-abstractions and which allows advanced execution modes of the ATLAS
+workload as well as the enhanced support for heterogeneous workloads such as
+molecular dynamics.
 
-This state-of-practice paper provides multiple contributions.  It (i)
-documents the many design and operational considerations that have been  taken
-to support the sustained, scalable and production usage of Titan for
-historically high-throughput workloads, (ii) Extensions to PanDA to support
-non-traditional heterogeneous workloads and execution modes,  and (iii) As the
-community looks forward to designing the next generation of online analytical
-platforms~\cite{foap}, this project provides some early lessons and guidance
-for how current and future experimental and observational systems can be
-integrated with production supercomputers in a general and extensible manner.
+% . Consequently, Titan now analyzes up to 3.6\% of the simulation workload of
+% the ATLAS experiment. The case study presents the design and integration of
+% PanDA
 
-% \subsubsection*{The following is a candidate for major compression into a single paragraph}
+This state-of-practice paper provides three main contributions:  (i) a critical
+evaluation of the many design and operational considerations that have been
+taken to support the sustained, scalable and production usage of Titan for
+historically high-throughput workloads; (ii) a preliminary characterization of a
+next generation executor (NGE) for PanDA to support non-traditional
+heterogeneous workloads and execution modes;  and (iii) early lessons and
+guidance for how current and future experimental and observational systems can
+be integrated with production supercomputers in a general and extensible manner.
 
-% \mtnote{Used the following in related work. We may want to reduce/remove it from
-% the introduction: ``Today, PanDA serves several thousand users, managing job
-% distribution to hundreds of ATLAS sites with more than 100,000 CPU cores which
-% process more than a million jobs per day.''} % (Figure~\ref{fig:daily}).
-
-% In this paper, we describe how PanDA has been engineered to execute a specific
-% stage of the ATLAS Monte Carlo workflow on Titan, the larger high-performance
-% computing HPC system currently available in the USA\@.\mtnote{Explain the
-% benefits offered by Titan in terms of multithreading per node and possibly large
-% amount of concurrent nodes. Introduce also the notion of backfill.} This extends
-% the scope of PanDA's compute model, integrating both high-throughput and
-% high-performance computing resources and enabling the concurrent execution of
-% both  single and multi-core jobs. The integration of PanDA and Titan went
-% through three main engineering phases: (i) feasibility study and rapid
-% prototyping of an initial solution; (ii) progressing scaling of the  prototype
-% to saturate the available resources; (iii) study of a product-grade architecture
-% for generic HPC resources. Both phase i and ii have been completed enabling the
-% execution of up to eight million jobs a week on Titan. A prototype has been
-% engineered to support phase iii and experimental data are being collected.
-
-% In the next section we introduce \ldots.
-
-% Why Titan?
-% \begin{itemize}
-%     \item A lot of slow (relative the grid) and homogeneous cores together
-%     \item Grid is saturated
-%     \item Enable HPC as a calss of computing resource
-%     \item Enable future DOE experimental and observational capabilities on HPC
-%     \item Why is so important: more data, run 2 and run 3.
-%     \item growth of grid is flat (economic model) and saturated. We need more CPU because we have more data.
-%     \item ATLAS spend most of the time on simulations this is why we want to offload simulations - that happen happen to be performed via AthenaMP.
-% \end{itemize}
-
-% \mtnote{Moved from related work were we use already this terminology. To be
-% iterated/adjusted for consistency once the first draft will be ready.}
-
-% The term ``workflow'' is used in many disciplines with different meaning. In the
-% field of scientific computing, ``workflow'' assumes different meanings depending
-% on the characteristics of the computation, of the software tools used to support
-% this computation, and of the resources on which it is performed. Further, a
-% workflow may indicate a whole application, a description of the computational
-% process of that application or, more commonly, a series of tasks related by data
-% dependences.
-
-% The lack of a consistent and shared definition of ``workflow'' hinders the
-% understanding of its properties and its relations with related concepts. For
-% example, we need to clarify the difference among ``workflow'', ``workload'',
-% ``task'', or ``job'' but also between workflow ``template'' and ``instance'', or
-% ``data-flow'' and ``control-flow''. This is precondition to specify properly the
-% design of software systems that support the execution of scientific
-% applications.
-
-% In this paper, we use the following definitions:
-
-% \begin{description}
-
-%   \item[Task.] A set of operations to be performed on a computing platform,
-%   alongside a description of the properties and dependences of those operations,
-%   and indications on how the dependences should be satisfied and the operations
-%   should be executed.
-
-%   \item[Job.] A unit of  work performed by submitting a script to a resource
-%   management system (LRMS), like  Slurm or PBS, or by requesting a virtual
-%   machine or a container to a site supporting virtualization. One or more jobs
-%   can perform the operations described with a task.
-
-%   \item[Workload.] A set of jobs that can be executed concurrently, possibly
-%   related by a set of relations. For example, jobs of a workload can share one
-%   or more input files or communicate during execution.
-
-%   \item[Workflow.] Set of jobs, related by a set of relations that define the
-%   order in which each task can be executed. Data dependences are the most common
-%   relations among workloads, used to define the precedence among their
-%   executions.
-
-% \end{description}
+% As the community looks forward to designing the next generation of online
+% analytical platforms~\cite{foap-url}, this project provides some

### 2017-04-04 introduction.tex
Fix bibliography

@@
 heterogeneous workloads and execution modes;  and (iii) early lessons and
-guidance for how current and future experimental and observational systems can
-be integrated with production supercomputers in a general and extensible manner.
-
-% As the community looks forward to designing the next generation of online
-% analytical platforms~\cite{foap-url}, this project provides some
+guidance as the community looks forward to designing the next generation of
+online analytical platforms~\cite{foap-url}, for how current and future
+experimental and observational systems can be integrated with production
+supercomputers in a general and extensible manner.

### 2017-04-04 appendix.tex
Anonymize appendix

--- 2017-03-28 appendix.tex
@@
 %  \item {\bf Binary: }
-  \item {\bf Data set: Available at \\ \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/tree/master/data}}
-  \item {\bf Run-time environment: PanDA Workload Management System, RADICAL-Pilot Pilot System (NGE)}
+  \item {\bf Data set: Available at \\
+  \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}}
+  %\url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/tree/master/data}}
+  \item {\bf Run-time environment: PanDA Workload Management System, NGE}
+  % RADICAL-Pilot Pilot System (NGE)}
   \item {\bf Hardware: OLCF Titan Cray XK7}
@@
 %  \item {\bf Execution: }
-  \item {\bf Output: Available at \\ \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/tree/master/data}}
+  \item {\bf Output: Available at \\
+  \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}}
+  %\url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/tree/master/data}}
   \item {\bf Experiment workflow: Raw data acquisition, data wrangling and filtering, plotting, analysis}
@@
 \begin{itemize}
-    \item PanDA Workload Management System: \\ \url{https://github.com/PanDAWMS}.
-    \item RADICAL-Pilot Pilot System (NGE): \\ \url{https://github.com/radical-cybertools}.
+    \item PanDA Workload Management System: \\
+    \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
+    % \url{https://github.com/PanDAWMS}.
+    \item NGE: \\
+    \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
+    % \url{https://github.com/radical-cybertools}.
 \end{itemize}
@@
 \begin{description}[align=left]
-    \item [Section 5.1 -- Figure 3] \url{}
-    \item [Section 5.1 -- Figure 4] \url{}
-    \item [Section 5.1 -- Figure 5] \url{}
-    \item [Section 5.1 -- Figure 6] \url{}
-    \item [Section 5.2 -- Figure 7] \url{}
-    \item [Section 5.2 -- Figure 8] \url{}
-    \item [Section 5.3 -- Figure 9] \url{}
-    \item [Section 6.3 -- Figure 14] \url{}
-    \item [Section 6.3 -- Figure 15] \url{}
-    \item [Section 6.3 -- Figure 16] \url{}
-    \item [Section 6.3 -- Figure 17] \url{}
-    \item [Section 6.3 -- Figure 18] \url{}
-    \item [Section 6.3 -- Figure 19] \url{}
-    \item [Section 6.3 -- Figure 20] \url{}
-    \item [Section 6.3 -- Figure 21] \url{}
+    \item [Section 4.1 -- Figure 3] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
+    \item [Section 4.2 -- Figure 4] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
+    \item [Section 4.2 -- Figure 5] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
+    \item [Section 5.2 -- Figure 7] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
+    \item [Section 5.2 -- Figure 8] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
+    \item [Section 5.2 -- Figure 9] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
 \end{description}
@@
 % =============================================================================
-\section{OLD -- REMOVE AFTER COMPLETING ABOVE}
-Reproducibility initiative appendices: Artifact Description (AD) and Computational Results Analysis (CRA). Description of the organization of the appendix.
-
-% -----------------------------------------------------------------------------
-\subsection{Section 5.1 - Figure 3}
-\label{apndx:fig3}
-
-Artifacts description.
-
-\begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
-
-% -----------------------------------------------------------------------------
-\subsection{Section 5.1 - Figure 4}
-\label{apndx:fig4}
-
-Artifacts description.
-
-\begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
-
-% -----------------------------------------------------------------------------
-\subsection{Section 5.1 - Figure 5}
-\label{apndx:fig5}
-
-Correlation between backfill availability of cores and walltime on Titan during the experiment time window.
-
-\begin{itemize}
-    \item Raw data: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.txt}, as provided by the PanDA Brokers logs on Titan.
-    \item Wrangling: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.csv}
-    \item Analysis and Plotting: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item Get a log file including regular polling of backfill availability from one or more PanDA Brokers on titan.
-    \item Wrangle the ascii file with the find/replace regular expressions listed at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}, cell \#1.
-    \item Load the Jupyter workbook from a python virtual environment with the module Jupyter installed and all the modules listed at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}, cell \#2.
-    \item Execute each cell of the Jupyter notebook to replicate Figure 5.
-\end{enumerate}
-
-% -----------------------------------------------------------------------------
-\subsection{Section 5.2 - Figure 6}
-\label{apndx:fig6}
-
-Artifacts description.
-
-\begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
-
-% -----------------------------------------------------------------------------
-\subsection{Section 5.2 - Figure 7}
-\label{apndx:fig7}
-
-Artifacts description.
-
-\begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
-
-% -----------------------------------------------------------------------------
-\subsection{Section 5.3 - Figure 8}
-\label{apndx:fig8}
-
-Panda failures on Titan by exit code during the experiment window.
-
-\begin{itemize}
-    \item Raw data: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.csv}
-    \item Wrangling: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Pivot Data \& Wrangled Data tabs)
-    \item Analysis: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Aggregated Data tab)
-    \item Plotting: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Plots tab)
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item Download raw data download in csv format from the ATLAS jobs dashboard.
+% \section{OLD -- REMOVE AFTER COMPLETING ABOVE}
+% Reproducibility initiative appendices: Artifact Description (AD) and Computational Results Analysis (CRA). Description of the organization of the appendix.
+%
+% % -----------------------------------------------------------------------------
+% \subsection{Section 5.1 - Figure 3}
+% \label{apndx:fig3}
+%
+% Artifacts description.
+%
+% \begin{itemize}
+%     \item Experiment code: repository address (when applicable)
+%     \item Raw data: repository address
+%     \item Wrangling: repository address
+%     \item Analysis code: repository address
+%     \item Plotting code: repository address
+% \end{itemize}
+%
+% Replicability:
+%
+% \begin{enumerate}
+%     \item
+% \end{enumerate}
+%
+%
+% % -----------------------------------------------------------------------------
+% \subsection{Section 5.1 - Figure 4}
+% \label{apndx:fig4}
+%
+% Artifacts description.
+%
+% \begin{itemize}
+%     \item Experiment code: repository address (when applicable)
+%     \item Raw data: repository address
+%     \item Wrangling: repository address
+%     \item Analysis code: repository address
+%     \item Plotting code: repository address
+% \end{itemize}
+%
+% Replicability:
+%
+% \begin{enumerate}
+%     \item
+% \end{enumerate}
+%
+%
+% % -----------------------------------------------------------------------------
+% \subsection{Section 5.1 - Figure 5}
+% \label{apndx:fig5}
+%
+% Correlation between backfill availability of cores and walltime on Titan during the experiment time window.
+%
+% \begin{itemize}
+%     \item Raw data: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.txt}, as provided by the PanDA Brokers logs on Titan.
+%     \item Wrangling: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.csv}
+%     \item Analysis and Plotting: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}
+% \end{itemize}
+%
+% Replicability:
+%
+% \begin{enumerate}
+%     \item Get a log file including regular polling of backfill availability from one or more PanDA Brokers on titan.
+%     \item Wrangle the ascii file with the find/replace regular expressions listed at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}, cell \#1.
+%     \item Load the Jupyter workbook from a python virtual environment with the module Jupyter installed and all the modules listed at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}, cell \#2.
+%     \item Execute each cell of the Jupyter notebook to replicate Figure 5.
+% \end{enumerate}
+%
+%
+% % -----------------------------------------------------------------------------
+% \subsection{Section 5.2 - Figure 6}
+% \label{apndx:fig6}
+%
+% Artifacts description.
+%
+% \begin{itemize}
+%     \item Experiment code: repository address (when applicable)
+%     \item Raw data: repository address
+%     \item Wrangling: repository address
+%     \item Analysis code: repository address
+%     \item Plotting code: repository address
+% \end{itemize}
+%
+% Replicability:
+%
+% \begin{enumerate}
+%     \item
+% \end{enumerate}
+%
+%
+% % -----------------------------------------------------------------------------
+% \subsection{Section 5.2 - Figure 7}
+% \label{apndx:fig7}
+%
+% Artifacts description.
+%
+% \begin{itemize}
+%     \item Experiment code: repository address (when applicable)
+%     \item Raw data: repository address
+%     \item Wrangling: repository address
+%     \item Analysis code: repository address
+%     \item Plotting code: repository address
+% \end{itemize}
+%
+% Replicability:
+%
+% \begin{enumerate}
+%     \item
+% \end{enumerate}
+%
+%
+% % -----------------------------------------------------------------------------
+% \subsection{Section 5.3 - Figure 8}
+% \label{apndx:fig8}
+%
+% Panda failures on Titan by exit code during the experiment window.
+%
+% \begin{itemize}
+%     \item Raw data: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.csv}
+%     \item Wrangling: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Pivot Data \& Wrangled Data tabs)
+%     \item Analysis: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Aggregated Data tab)
+%     \item Plotting: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Plots tab)
+% \end{itemize}
+%
+% Replicability:
+%
+% \begin{enumerate}
+%     \item Download raw data download in csv format from the ATLAS jobs dashboard.
 Download link: \url{http://dashb-atlas-job.cern.ch/dashboard/request.py/terminatedjobsstatuscsv?sites=All%20T3210&sitesCat=All%20Countries&resourcetype=All&pandares=ORNL_Titan_MCORE&activities=simul&sitesSort=7&sitesCatSort=0&start=2016-01-01&end=2017-02-28&timeRange=daily&sortBy=16&granularity=Monthly&generic=0&series=30&type=abcb}
-    \item Load the csv file into Excel (or any other software/language).
-    \item Create a pivot table with failure types as columns and months as rows.
-    \item Aggregate data of error codes of the same type by merge columns and adding the values of their raws according to this table:
-    \item Plot the resulting table as a stacked area diagram.
-\end{enumerate}
-
-% -----------------------------------------------------------------------------
-\subsection{Section 6.3 - Figure 14}
-\label{apndx:fig14}
-
-Artifacts description.
-
-\begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
-
-% -----------------------------------------------------------------------------
-\subsection{Section 6.3 - Figure 15}
-\label{apndx:fig15}
-
-Artifacts description.
-
-\begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
-
-% -----------------------------------------------------------------------------
-\subsection{Section 6.3 - Figure 16}
-\label{apndx:fig16}
-
-Artifacts description.
-
-\begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
-
-% -----------------------------------------------------------------------------
-\subsection{Section 6.3 - Figure 17}
-\label{apndx:fig17}
-
-Artifacts description.
-
-\begin{itemize}
-    \item Experiment code: repository address (when applicable)
-    \item Raw data: repository address
-    \item Wrangling: repository address
-    \item Analysis code: repository address
-    \item Plotting code: repository address
-\end{itemize}
-
-Replicability:
-
-\begin{enumerate}
-    \item
-\end{enumerate}
+%     \item Load the csv file into Excel (or any other software/language).
+%     \item Create a pivot table with failure types as columns and months as rows.
+%     \item Aggregate data of error codes of the same type by merge columns and adding the values of their raws according to this table:
+%     \item Plot the resulting table as a stacked area diagram.
+% \end{enumerate}
+%
+%
+% % -----------------------------------------------------------------------------
+% \subsection{Section \ref{sec:ngeExp} - Figure \ref{fig:weakScal1a}}
+% \label{apndx:fig9}
+%
+% \begin{itemize}
+%     \item Raw data: \url{https://github.com/ATLAS-Titan/misc/tree/master/atlas-paper-experiments/data}
+%      \item Processed data: \url{ https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/data.csv}
+%     \item Raw data processing:
+%     	\begin{itemize}
+%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/dataAnalysis.py}
+%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/script.py}
+%     		\item Follow instructions at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-extract-data-from-the-sandbox}
+%    	\end{itemize}
+%     \item Plotting code: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/plot.gnu}
+% \end{itemize}
+%
+% Replicability: Follow instructions at the following link: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-reproduce-experiments-on-TITAN} to prepare the environment on Titan.
+% Launch:
+% \begin{itemize}
+% \item python dispatcher.py 250 250 120 etc/input\_100 to replicate experiments with 250 nodes;
+% \item python dispatcher.py 500 500 120 etc/input\_100 to replicate experiments with 500 nodes;
+% \item python dispatcher.py 1000 1000 120 etc/input\_100 to replicate experiments with 1000 nodes;
+% \item python dispatcher.py 2000 2000 120 etc/input\_100 to replicate experiments with 2000 nodes.
+% \end{itemize}
+%
+%
+%
+% % -----------------------------------------------------------------------------
+% \subsection{Section \ref{sec:ngeExp} - Figure \ref{fig:weakScal2a}}
+% \label{apndx:fig10}
+% \begin{itemize}
+%     \item Raw data: \url{https://github.com/ATLAS-Titan/misc/tree/master/atlas-paper-experiments/weak2}
+%      \item Processed data: \url{ https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/weak2.csv}
+%     \item Raw data processing:
+%     	\begin{itemize}
+%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/dataAnalysis.py}
+%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/script.py}
+%     		\item Follow instructions at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-extract-data-from-the-sandbox}
+%    	\end{itemize}
+%     \item Plotting code: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/plot.gnu}
+% \end{itemize}
+%
+% Replicability: Follow instructions at the following link: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-reproduce-experiments-on-TITAN} to prepare the environment on Titan.
+% Launch:
+% \begin{itemize}
+% \item python dispatcher.py 256 1280 256 180 etc/input\_16 to replicate experiments with 256 nodes;
+% \item python dispatcher.py 512 2560 512 180 etc/input\_16 to replicate experiments with 512 nodes;
+% \item python dispatcher.py 1024 5120 180 etc/input\_16 to replicate experiments with 1024 nodes;
+% \item python dispatcher.py 2048 10240 180 etc/input\_16 to replicate experiments with 2048 nodes.
+% \end{itemize}
+%
+%
+% % -----------------------------------------------------------------------------
+% \subsection{Section \ref{sec:ngeExp} - Figure \ref{fig:strongScala}}
+% \label{apndx:fig16}
+%
+% \label{apndx:fig10}
+% \begin{itemize}
+%     \item Raw data: \url{https://github.com/ATLAS-Titan/misc/tree/master/atlas-paper-experiments/strong}
+%      \item Processed data: \url{ https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/strong.csv}
+%     \item Raw data processing:
+%     	\begin{itemize}
+%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/dataAnalysis.py}
+%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/script.py}
+%     		\item Follow instructions at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-extract-data-from-the-sandbox}
+%    	\end{itemize}
+%     \item Plotting code: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/plot.gnu}
+% \end{itemize}
+%
+% Replicability: Follow instructions at the following link: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-reproduce-experiments-on-TITAN} to prepare the environment on Titan.
+% Launch:
+% \begin{itemize}
+% \item python dispatcher.py 256  2048 256 180 etc/input\_16 to replicate experiments with 256 nodes;
+% \item python dispatcher.py 512 2048 512 180 etc/input\_16 to replicate experiments with 512 nodes;
+% \item python dispatcher.py 1024 2048 180 etc/input\_16 to replicate experiments with 1024 nodes;
+% \item python dispatcher.py 2048 2048 180 etc/input\_16 to replicate experiments with 2048 nodes.
+% \end{itemize}
+%
+%
+%
+% % -----------------------------------------------------------------------------
+%

### 2017-04-04 PandaEvolution.tex
Address Danila's and Alexei's comments

@@
 
-% as well as commercial and academic clouds.
-% \mtnote{Addresses Alexei's comment. Increased Grid capacity to 300K cores. Is
-% the number of sites still 100 or should I increase that too?}
-
-% -----------------------------------------------------------------------------
-% \subsection{Use of Supercomputers with PanDA}
-% \label{ssec:panda-supercomputers}
-
 Modern supercomputers have been designed mainly to support parallel computation
@@
 types of specialized coprocessors.
-
-% \sergeynote{specifically "distributed Grid computing"}\mtnote{done.}.
-% \sergeynote{geographically distributed?}\mtnote{Better?}
 
@@
 
-% \mtnote{Move this to the introduction: ``There are at least two approaches to
-% enable PanDA WMS to execute ATLAS workloads or workflows on supercomputers: (i)
-% using the subset of resources and capabilities shared by both supercomputers and
-% WLCG; (ii) reconciling the parallel and distributed computing paradigms by means
-% of dedicated abstractions. The former is a pragmatic approach that enables the
-% execution of specific workloads by prototyping single-point solutions. The
-% latter is a principled approach, better suited for a production-grade solution,
-% capable of supporting general-purpose workloads and workflows on both
-% supercomputers and Grid infrastructures. \\ This section illustrates the design
-% and architecture of a job broker prototyped by the PanDA team. This broker
-% supports execution of part of the ATLAS production Monte Carlo workflow on
-% Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
-% Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).After an
-% analysis of the results obtained and the lessons learned, the following section
-% introduces the design and first experimental characterization of a next
-% generation executor (NGE). NGE is designed to abstract resources and
-% capabilities, enabling the concurrent execution of both parallel and distributed
-% computing on generic high performance computing (HPC) machines.''}
-%
-% \jhanote{Do you mean the top of section 3? Or do you mean the very beginning
-% section?}
-
 % -----------------------------------------------------------------------------
-%\subsection{Interfacing PanDA with Titan}
 \subsection{Architectures and Interfaces}
@@
 
-% Titan, was the first large-scale system to use a hybrid architecture that
-% utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla
-% K20 GPU accelerators. This hybrid design provides improved energy efficiency,
-% as well as an order of magnitude in computational capacity over its
-% predecessor.
-
 Titan's users submit jobs to Titan's PBS scheduler by logging into login or data
@@
 are in place both for the PBS batch system and shared file systems.
-
-% Data staging is supported via the DTNs, from the wide area network and from
-% the OLCF high performance storage system (HPSS). DTNs are shared across all
-% Titan's users that can schedule jobs to a dedicates batch system to automate
-% data staging. Fair-share policies are in place both for the batch system and
- %for the use of DTN resources.
 
@@
 be off-loaded from the Grid to Titan's large amount of cores. Further, when HEP
-tasks can be partitioned into independent jobs, Titan worker nodes can be used to
-execute up to 16 concurrent payloads, one per each available core. Given these
-constraints and challenges, the type of task most suitable for execution at the
-moment on Titan is Monte Carlo detector simulation. This type of task is mostly
-computational-intensive, requiring less than 2GB of RAM at runtime and with
-small input data requirements. Detector simulation tasks in ATLAS are performed
-via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
-GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
-$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
-offloading.
-
-% \mtnote{Addresses Alexei's comment. Is the following sentence correct?}
+tasks can be partitioned into independent jobs, Titan worker nodes can be used
+to execute up to 16 concurrent payloads, one per each available core. Given
+these constraints and challenges, the type of task most suitable for execution
+at the moment on Titan is Monte Carlo detector simulation. This type of task is
+mostly computational-intensive, requiring less than 2GB of RAM at runtime and
+with small input data requirements. Detector simulation tasks in ATLAS are
+performed via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework
+integrating the GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These
+tasks account for $\approx$ 60\% of all the jobs on WLCG, making them a primary
+candidate for offloading.
 
@@
 
-% While this is different from the HEP computing paradigm, where jobs are
-% independent, it still shares common features such as the use of
-% parallelization. It is not a requirement that HPC machines are able to run
-% any possible task, nor is it relevant how many kinds of job types that can be
-% run. What matters is the total number of cycles that can be offloaded from
-% the Grid.
-
-% Standard ATLAS workflow can not be easily ported to supercomputers due to
-% several complications such as specialized worker node setups, no outbound
-% network connections, limited memory per node, custom operating systems, etc. A
-% reorganization of the standard workflow is therefore needed.
-
 % -----------------------------------------------------------------------------
-% \subsection{PanDA Broker on Titan}
 \subsection{PanDA Broker}
@@
 
-% MC event generation and simulation tasks can be executed without pilots but
-% the time spent waiting in the queue would be too large compared to the
-% execution time of each task.
-
-% Titan’s backfill functionality offers the opportunity to avoid the overhead of
-% queue wait times without using pilot abstraction. Backfill availability is the
-% number of node-hours (or core-hours) that cannot be used for a certain amount
-% of time by any of the jobs already queued on Titan: All queued jobs are either
-% too large or their walltime is too long.
-
 The backfill optimization of Titan's Moab scheduler offers the opportunity to
@@
 
-% It is important to note that backfill availability is computed retrospectively
-% and is determined by Titan's operation team. Backfill availability represents
-% the theoretical maximum of node-hours that could be utilized by ATLAS (or any
-% other project) that submit under backfill mode.
-
 Compared to pilots, backfill has the disadvantage of limiting the amount of
@@
 
-% \sergeynote{any other job currently in the Titan's queue?}\mtnote{Better?}.
-
 At any point in time, the size of an available backfill slot is typically a
@@
 10\% of Titan's capacity remains unused~\cite{barker2016us}, corresponding to an
-average of 30,000 unused cores (excluding GPU cores). This equals
-% approximately 270M core-hours per year,
-to roughly 10\% of the overall capacity of WLCG.
-
-% A system that can use those temporarily free worker nodes with smaller
-% scale tasks would be very valuable. For this reason, Titan offers backfill
-% functionalities: users can interrogate Titan's Moab scheduler about how many
-% free nodes are currently available and for how long. If a user submit a
-% requestvfor that number of nodes and walltime, queue time will be minimal.
-
-% \mtnote{Please feel free to add details about backfill as needed.}
-% \mtnote{Please note: I know the following departs for the current name given
-% to the PanDA subsystem installed on Titan. I think there is a good reason to
-% call it a broker instead of a pilot, and I think I explained it in the
-% previous paragraphs. Please  take this just as a suggestion, something I would
-% like to discuss in our meeting.} \sergeynote{It's not a pilot in conventional
-% sense. I call it an agent. It just happened that we were able to use full
-% PanDA pilot's code base to serve our purposes on Titan. that's why, by
-% inertia, we still call it a pilot.}\mtnote{Thank you. We would have a problem
-% calling it `agent' as we use that term to name the pilot of NGE. Would PanDA
-% Broker work?}\sergeynote{OK broker it is for this paper.}
+average of 30,000 unused cores (excluding GPU cores). This equals to roughly
+10\% of the overall capacity of WLCG.
 
@@
 implements functionalities to: (i) interrogate Titan about backfill
-availability; (ii) pull MC jobs and events; (iii) wrap the payload of ATLAS jobs
-into MPI scripts; (iv) submitting MPI scripts to Titan's PBS batch system and
-monitor their execution; and (v) staging input/output files. Backfill querying,
-payload wrapping, and scripts submission required a new implementation while
-pulling ATLAS job and events, and file staging were inherited from PanDA Pilot.
-
-% \sergeynote{Monitor job state and state transitions using
-% SAGA}\mtnote{Better?}
+availability; (ii) pull ATLAS jobs and events from PanDA Server; (iii) wrap the
+payload of ATLAS jobs into MPI scripts; (iv) submitting MPI scripts to Titan's
+PBS batch system and monitor their execution; and (v) staging and preparing
+input/output files. Backfill querying, payload wrapping, and scripts submission
+required a new implementation while pulling ATLAS job and events, and file
+staging were inherited from PanDA Pilot.
 
@@
 
-% Each script knows its MPI rank (an index that runs from zero to a maximum
-% number of nodes or script copies) as well as the total number of ranks in the
-% current submission. When activated on worker nodes, each copy of the MPI
-% script
-
 MPI scripts are submitted to Titan's PBS batch system via
@@
 system that is shared among DTNs and worker nodes. The input files with the
-events of the ATLAS jobs are downloaded on the shared filesystem via the ATLAS
-DDM service. The MPI script setup process includes making the location of these
-files available to the payload of the ATLAS's jobs. The PanDA Broker can locate
-the payload's output files on the shared filesystem and transfer them from Titan
-to any computing center used by ATLAS.
-% the ATLAS Tier 1 computing center at Brookhaven National Lab (BNL) or any
-% other Grid site.
+events of the ATLAS jobs are downloaded on the shared filesystem from the data
+center of Brookhaven National Laboratory (BNL). The MPI script setup process
+includes making the location of these files available to the payload of the
+ATLAS's jobs. The PanDA Broker can locate the payload's output files on the
+shared filesystem and transfer them from Titan BNL.
 
 Once deployed on Titan, every PanDA Broker supports the execution of MC detector
-simulation in 9 steps. PanDA Broker queries the Job Dispatcher module of the
+simulations in 9 steps. PanDA Broker queries the Job Dispatcher module of the
 PanDA server for ATLAS jobs that have been bound to Titan by JEDI
 (Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA Broker
-pulls jobs' input files from the ATLAS DDM service to the OLCF Lustre file
-system (Fig.~\ref{fig:panda_broker}:3). PanDA Broker queries Titan's Moab
-scheduler about the current available backfill slot
-(Fig.~\ref{fig:panda_broker}:2) and creates an MPI script, wrapping enough ATLAS
-jobs' payload to fit the backfill slot. PanDA Broker submits the MPI script to
-the Titan's Torque batch system via RADICAL-SAGA
-(Fig.~\ref{fig:panda_broker}:4).
+pulls jobs' input files from BNL to the OLCF Lustre file system
+(Fig.~\ref{fig:panda_broker}:2). PanDA Broker queries Titan's Moab scheduler
+about the current available backfill slot (Fig.~\ref{fig:panda_broker}:3) and
+creates an MPI script, wrapping enough ATLAS jobs' payload to fit the backfill
+slot. PanDA Broker submits the MPI script to the Titan's Torque batch system via
+RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).
 
@@
 work node (Fig.~\ref{fig:panda_broker}:7). AthenaMP retrieves events from Lustre
-(Fig.~\ref{fig:panda_broker}:8) and spawns 1 event simulation process on each of
-the 16 available cores (Fig.~\ref{fig:panda_broker}:9).
-% During execution, PanDA Broker monitors execution progress and sends PanDA
-% Server ``heart beats'' for each job.; and (viii)
-Upon completion of each MPI script, PanDA Broker transfer the jobs' output to
-designated computing centre via the ATLAS DDM(Fig.~\ref{fig:panda_broker}:10),
-and performs cleanup.
+(Fig.~\ref{fig:panda_broker}:8) and spawns 1 Geant4 event simulation process on
+each of the 16 available cores (Fig.~\ref{fig:panda_broker}:9). Upon completion
+of each MPI script, PanDA Broker transfer the jobs' output to BNL
+(Fig.~\ref{fig:panda_broker}:10), and performs cleanup.
 
-% \mtnote{Address Alexei's comments: Add a paragraph explaining how this solution
-% has been ported to other HPC machines and reference relevant publications.}
 While PanDA Broker implementation is resource specific, it was successfully
-ported also to other supercomputers, including the Ansel (?) at \ldots and
-Cori at the National Energy Research Scientific Computing Center (NERSC).
-
-% from the Job Dispatcher module of PanDA Server (as done by a PanDA Pilot);
-% (iii) downloading events via PanDA Server's Data Service from ATLAS DDM to
-% Titan's DTN; (iv) wrap the Geant4 payload into a PBS job; (v) submit jobs to
-% Titan's PBS batch scheduler;
-
-% Integration with Titan is the current focus for PanDA developers.
-
-% The project aims to integrate Titan with the PanDA system using an updated
-% PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the
-% worker nodes using the local batch system (PBS) via the SAGA (Simple API for
-% Grid Applications) interface~\cite{SAGA}. This solves several complications
-% of running on HPC worker nodes, including the lack of connectivity to outside
-% world. The pilot can communicate with the PanDA server from the front-end
-% machine.
-
-% \paragraph*{HPC Backfill} HPC facilities are geared towards large scale jobs
-% by design. Time allocation on an HPC is competitive and large projects are
-% often preferred. About 10\% of capacity on a typical HPC machine is unused
-% due to mismatches between job sizes and available resources. The worker nodes
-% sit idle because there are not enough of them or they do not have enough time
-% to handle a large scale computing job. On a machine of the scale of Titan,
-% these 10\% correspond to estimated 300M core hours per year. A system that
-% can occupy those temporarily free nodes with smaller scale tasks would be
-% very valuable.
-
-% This offers a great possibility for PanDA to harvest the opportunistic
-% resources on Titan and use a similar mechanism on other HPCs. Functionality
-% has been added to the PanDA Pilot to collect information about available
-% unused worker nodes on Titan in real time.
-
-% First test of the system were quite successful and show great promise in
-% increasing the resource utilization on Titan.
+ported to other supercomputers, including the HPC2 at the National Research
+Center ``Kurchatov Institute'' (NRC-KI)~\cite{belyaev2015integration},
+Edison/Cori at the National Energy Research Scientific Computing Center
+(NERSC)~\cite{barreiro2016panda}, and SuperMUC at the Leibniz Supercomputing
+Centre (LRZ)~\cite{barreiro2016panda}.

### 2017-04-04 Experiments.tex
saved 4-5 lines and minor corrections.

@@
     \caption{Weak scalability: average pilot duration, average  duration of a
-    single AthenaMP execution, and pilot's overhead as a function of different pilot sizes (200, 500, 1000 and 2000 nodes).}
+    single AthenaMP execution, and pilot's overhead as a function of pilot sizes (200, 500, 1000 and 2000 nodes).}
 \label{fig:weakScal1a}
@@
     duration, average duration of five sequential AthenaMP executions, and
-    pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes).}
+    pilot's overhead for pilot sizes (256, 512, 1024 and 2048 nodes).}
 \label{fig:weakScal2a}

### 2017-04-04 PandaOverview.tex
saved 4-5 lines and minor corrections.

@@
     based on JEDI. Several subsystems and components, the architectural
-    details of PanDA Pilot and the communication among subsystem's components
-    are abstracted to improve clarity.}
+    details of PanDA Pilot and the communication among components
+    are abstracted for clarity.}
 \label{fig:architecture}

### 2017-04-04 PandaOverview.tex
Space-saving section 2

@@
-PanDA is a Workload Management System (WMS) designed to support the execution of
-distributed workloads and workflows via pilots. WMS is middleware for
-discovering and selecting resources, submitting tasks of workloads and
-workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
-abstraction that enables multi-level scheduling by decoupling resource
-acquisition from tasks scheduling~\cite{turilli2015comprehensive}. When
-implemented, a pilot is scheduled on a site and, once active, tasks are
-scheduled to the pilot, not to the site's scheduler.
+% PanDA is a Workload Management System (WMS) designed to support the execution of
+% distributed workloads and workflows via pilots. WMS is middleware for
+% discovering and selecting resources, submitting tasks of workloads and
+% workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
+% abstraction that enables multi-level scheduling by decoupling resource
+% acquisition from tasks scheduling~\cite{turilli2015comprehensive}. When
+% implemented, a pilot is scheduled on a site and, once active, tasks are
+% scheduled to the pilot, not to the site's scheduler.
+%
+% Pilot-capable WMS enable high throughput of tasks execution while supporting
+% interoperability across multiple sites. This is particularly relevant for LHC
+% experiments, where millions of tasks are executed across multiple sites every
+% month, analyzing and producing petabytes of data. The design of PanDA WMS
+% started in 2005 to support ATLAS. PanDA went into
+% production for the LHC Run 1 on 2009, and was then extended with new subsystems
+% for Run 2 in 2015.
 
-Pilot-capable WMS enable high throughput of tasks execution while supporting
-interoperability across multiple sites. This is particularly relevant for LHC
-experiments, where millions of tasks are executed across multiple sites every
-month, analyzing and producing petabytes of data. The design of PanDA WMS
-started in 2005 to support ATLAS. PanDA went into
-production for the LHC Run 1 on 2009, and was then extended with new subsystems
-for Run 2 in 2015.
+PanDA is a Workload Management System (WMS)~\cite{marco2009glite} designed to
+support the execution of distributed workloads and workflows via
+pilots~\cite{turilli2015comprehensive}. Pilot-capable WMS enable high throughput
+of tasks execution via multi-level scheduling while supporting interoperability
+across multiple sites. This is particularly relevant for LHC experiments, where
+millions of tasks are executed across multiple sites every month, analyzing and
+producing petabytes of data. The design of PanDA WMS started in 2005 to support
+ATLAS. PanDA went into production for the LHC Run 1 on 2009, and was then
+extended with new subsystems for Run 2 in 2015.
 
@@
 
-PanDA's application model assumes tasks, workloads and workflows. Tasks
-represent a set of operations performed on data sets stored in one or more input
-files. Tasks are decomposed into jobs, where each job represents the task's set
-of operations and a partition of the task's data set. Since 2005, a certain
-amount of parallelism has been progressively introduced for job
+PanDA's application model assumes tasks grouped into workloads and workflows.
+Tasks represent a set of operations performed on data sets stored in one or more
+input files. Tasks are decomposed into jobs, where each job consists of the
+task's operations performed on a partition of the task's data set. Since 2005, a
+certain amount of parallelism has been progressively introduced for job
 execution~\cite{crooks2012multi} but, so far, HEP applications have not required
-MPI. Jobs are supposed to be relatively self-contained, capable of setting up
-their own execution environment or having a minimal set of common dependences.
+MPI.
+% Jobs are supposed to be capable of setting up their own execution
+% environment and have a minimal set of common dependences.
 
-PanDA's usage model is based on multitenancy of resources and the support of at
-least two types of HEP users: individual researchers and groups executing so
-called `production' workflows. Users are free to submit tasks and workflows to
-the PanDA WMS at any point in time, directly or via dedicated application
-frameworks. PanDA's security model is based on separation between
-authentication, authorization and accounting for both single users and group of
-users. Both authentication and authorization are based on digital certificates
-and the X.509 standard and on the virtual organization (VO) abstraction.
+PanDA's usage model assumes multitenancy of resources and at least two types of
+HEP users: individual researchers and groups executing so called `production'
+workflows.
+% Users can submit tasks and workflows to the PanDA WMS at any point in
+% time, directly or via dedicated application frameworks.
+PanDA's security model is based on separation between authentication,
+authorization and accounting for both single users and group of users. Both
+authentication and authorization are based on digital certificates and on the
+virtual organization abstraction~\cite{foster2001anatomy}.
 
@@
 providing information about the stage of the execution in which each job or data
-item is or has been. Some attributes are assumed to be immutable across
-execution and jobs and data items are assumed to be always in one and only one
-state.
+item is or has been.
+% Some attributes are assumed to be immutable across execution and jobs and data
+% items are assumed to be always in one and only one state.
 
@@
 
-The implementation of PanDA consists of several interconnected subsystems, most
-of them built from off-the-shelf and Open Source components. Subsystems
+The implementation of PanDA WMS consists of several interconnected subsystems,
+most of them built from off-the-shelf and Open Source components. Subsystems
 communicate via dedicated API or HTTP messaging, and each subsystem is
 implemented by one or more modules. Databases are used to store eventful
-entities and entities' description like tasks, jobs and % events
-input/output data, and to store information about sites,
+entities % and entities' description
+like tasks, jobs and  input/output data, and to store information about sites,
 resources, logs, and accounting.
@@
 workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we do not
-discuss them as they are  not relevant to  an understanding of how PanDA has
-been ported to supercomputers. For a full list of subsystems see
+discuss them as they are  not relevant to an understanding of how PanDA has been
+ported to supercomputers. For a full list of subsystems see
 Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture} shows a diagrammatic
@@
 
-The relation between tasks and jobs can be one-to-one or one-to-many, and the
-conversion between the two can by static or dynamic. During LHC Run 1, PanDA
-required users or applications to perform a static conversion between tasks and
-jobs: tasks were described as a set of jobs and then submitted to the PanDA
-Server.
+% The relation between tasks and jobs can be one-to-one or one-to-many, and the
+% conversion between the two can by static or dynamic.
 
-This approach introduced inefficiency both with usability and resource
-utilization~\cite{borodin2015unified}. Ideally, users should not have to reason
-in terms of jobs: Users conceive analyses in terms of one or more, possibly
-related tasks; the `job' abstraction is required by the execution middleware,
-i.e. PanDA. Further, a static partitioning of tasks into jobs does not take into
-account the heterogeneity and dynamicity of the resources of the pilots on which
-each job will be executed.
+During LHC Run 1, PanDA required users to perform a static conversion between
+tasks and jobs: tasks were described as a set of jobs and then submitted to the
+PanDA Server. This introduced inefficiency both with usability and resource
+utilization~\cite{borodin2015unified}. Ideally, users should conceive analyses
+in terms of one or more, possibly related tasks, while the execution manager
+(i.e., PanDA) should partition tasks into jobs, depending on execution
+constraints. Further, the static partitioning of tasks into jobs does not take
+into account the heterogeneity and dynamicity of the resources on which each job
+will be executed, introducing inefficiencies.
 
@@
 available for a defined amount of time or could disappear at any point in time
-as with opportunistic models of resource provision.
+as it happens with opportunistic models of resource provision.
 
-JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users or
-the application layer submit task descriptions to JEDI
-(Fig.~\ref{fig:architecture}:1) that stores them into a queue implemented by a
-database (Fig.~\ref{fig:architecture}:2). Tasks are partitioned into jobs of
-different size, depending on both static and dynamic information about available
-resources (Fig.~\ref{fig:architecture}:3). Jobs are bound to sites with
-resources that best match jobs' requirements, and submitted to the PanDA Server
-for execution (Fig.~\ref{fig:architecture}:4).
+JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users
+submit task descriptions to JEDI (Fig.~\ref{fig:architecture}:1) that stores
+them into a queue implemented by a database (Fig.~\ref{fig:architecture}:2).
+Tasks are partitioned into jobs of different size, depending on both static and
+dynamic information about available resources (Fig.~\ref{fig:architecture}:3).
+Jobs are bound to sites with resources that best match jobs' requirements, and
+submitted to the PanDA Server for execution (Fig.~\ref{fig:architecture}:4).
 
@@
 requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7) that, when needed,
-takes care of  replicating them on the required sites
-(Fig.~\ref{fig:architecture}:8).
+replicates data sets on the required sites (Fig.~\ref{fig:architecture}:8).
 
-Meanwhile, AutoPyFactory defines a set of suitable PanDA Pilots and concurrently
-submits them to a Condor-G broker (Fig.~\ref{fig:architecture}:9). This broker
-submits these pilots wrapped as jobs or VMs to the required sites
-(Fig.~\ref{fig:architecture}:10).
+Meanwhile, AutoPyFactory defines PanDA Pilots, submitting them to a Condor-G
+agent (Fig.~\ref{fig:architecture}:9). Condor-G schedules these pilots wrapped
+as jobs or VMs to the required sites (Fig.~\ref{fig:architecture}:10).
 
-When a PanDA Pilot becomes available on a site, it calls the Job Dispatcher
-module of the PanDA Server, requesting  a job to execute
-(Fig.~\ref{fig:architecture}:11). The Job Dispatcher interrogates the Task
-Buffer module for a job that is bound to the site of that pilot and ready to
-be executed. Task Buffer checks the global queue (i.e., the PanDA database)
-and, upon availability, returns a suitable job to the Job Dispatcher. The Job
-Dispatcher dispatches that job to the PanDA Pilot
+When a PanDA Pilot becomes available, it requests the Job Dispatcher module of
+the PanDA Server for a job to execute (Fig.~\ref{fig:architecture}:11). The Job
+Dispatcher interrogates the Task Buffer module for a job that is bound to the
+site of that pilot and ready to be executed. Task Buffer checks the global queue
+(i.e., the PanDA database) and, upon availability, returns a job to the Job
+Dispatcher. The Job Dispatcher dispatches that job to the PanDA Pilot
 (Fig.~\ref{fig:architecture}:12).
@@
 Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
-subprocess for the execution of the job's payload. The job is set up, input data
-are transferred from the designated staging-in location
-(Fig.~\ref{fig:architecture}:13), and the job's payload is executed
-(Fig.~\ref{fig:architecture}:14). Once completed, output is transferred to the
-staging-out location (Fig.~\ref{fig:architecture}:15).
+subprocess for the execution of the job's payload. Input data are transferred
+from the stage-in location (Fig.~\ref{fig:architecture}:13) and the job's
+payload is executed (Fig.~\ref{fig:architecture}:14). Once completed, output is
+transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).
 
@@
     \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
-    \caption{PanDA WMS architecture. Numbers indicates the execution process
-    based on JEDI. Several subsystems and components, the architectural
-    details of PanDA Pilot and the communication among components
-    are abstracted for clarity.}
+    \caption{PanDA WMS architecture. Numbers indicates the JEDI-based execution
+    process. Several subsystems, components, and architectural and
+    communication details are abstracted for clarity.}
 \label{fig:architecture}

### 2017-04-04 PandaEvolution.tex
saved 4-5 lines and minor corrections.
Scattered space saving editings

@@
-The next LHC run for taking data (Run 3) will require more resources than the
-Worldwide LHC Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more
-than 300,000 cores at over 100 Grid sites, with a peak performance of 0.3
-petaFLOPS. This capacity will be sufficient for the planned analysis and data
-processing, but it will be insufficient for the Monte Carlo production workflow
-and any extra activity. To alleviate these challenges, ATLAS is engaged in a
-program to expand the current computing model to include additional resources
-such as the opportunistic use of supercomputers.
+The upcoming LHC Run 3 will require more resources than the Worldwide LHC
+Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than 300,000
+cores at over 100 Grid sites, with a peak performance of 0.3 petaFLOPS. This
+capacity will be sufficient for the planned analysis and data processing, but it
+will be insufficient for the Monte Carlo production workflow and any extra
+activity. To alleviate these challenges, ATLAS is engaged in a program to expand
+the current computing model to include additional resources such as the
+opportunistic use of supercomputers.
 
-Modern supercomputers have been designed mainly to support parallel computation
-that requires runtime communication. Job execution is parallelized across
-multiple cores, each core calculating a small part of the problem and
-communicating with other cores via a message passing interface (MPI).
-Accordingly, supercomputers have large number of worker nodes, connected through
-a high-speed, low-latency dedicated network. Each worker node has multicore
-CPUs, usually augmented with parallel Graphics Processing Units (GPUs) or other
-types of specialized coprocessors.
+Generally, supercomputers are designed to support parallel computation that
+requires runtime communication. Jobs are executed across multiple cores, each
+core calculating a small part of the problem and communicating with other cores
+via MPI. Accordingly, supercomputers have large number of worker nodes,
+connected through a high-speed, low-latency dedicated network. Each worker node
+has multicore CPUs, usually augmented with Graphics Processing Units (GPUs) or
+other specialized coprocessors.
 
@@
 transfer nodes (DTNs). Titan's authentication and authorization model is based
-on two-factor authentication with a RSA SecurID key, generated every 30 seconds.
-Login nodes and DTNs have out/inbound wide area network connectivity while
-worker nodes have only local network access. Fair-share and allocation policies
-are in place both for the PBS batch system and shared file systems.
+on two-factor authentication with a RSA SecurID key. Login nodes and DTNs have
+out/inbound wide area network connectivity while worker nodes have only local
+network access. Fair-share and allocation policies are in place both for the PBS
+batch system and shared file systems.
 
@@
 computationally-intensive and non memory-intensive tasks of ATLAS workflows can
-be off-loaded from the Grid to Titan's large amount of cores. Further, when HEP
-tasks can be partitioned into independent jobs, Titan worker nodes can be used
-to execute up to 16 concurrent payloads, one per each available core. Given
-these constraints and challenges, the type of task most suitable for execution
-at the moment on Titan is Monte Carlo detector simulation. This type of task is
-mostly computational-intensive, requiring less than 2GB of RAM at runtime and
-with small input data requirements. Detector simulation tasks in ATLAS are
-performed via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework
-integrating the GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These
-tasks account for $\approx$ 60\% of all the jobs on WLCG, making them a primary
-candidate for offloading.
+be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
+partitioned into independent jobs, Titan worker nodes can be used to execute up
+to 16 concurrent payloads, one per each available core. Given these constraints
+and challenges, the type of task most suitable for execution at the moment on
+Titan is Monte Carlo detector simulation. This type of task is mostly
+computational-intensive, requiring less than 2GB of RAM at runtime and with
+small input data requirements. Detector simulation tasks in ATLAS are performed
+via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
+GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
+$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
+offloading.
 
-Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
-(also known as MC production
-chain)~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
+Detector simulation is part of the ATLAS production Monte Carlo (MC)
+workflow~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
 consists of four main stages: event generation, detector simulation,
@@
 
-The backfill optimization of Titan's Moab scheduler offers the opportunity to
-avoid the overhead of queue wait times without using pilot
-abstraction~\cite{maui_backfill_url}. This optimization enables Moab to start
-low-priority jobs as long as they do not delay higher priority jobs,
-independent of whether the low-priority jobs were queued after the
-high-priority jobs.
+The backfill optimization of Titan's Moab scheduler allows to avoid the overhead
+of queue wait times without using pilot abstraction~\cite{maui_backfill_url}.
+With this optimization, Moab starts low-priority jobs when they do not delay
+higher priority jobs, independent of whether the low-priority jobs were queued
+after the high-priority jobs.
 
@@
 worker nodes that can be requested. Pilots are normal jobs: they can request as
-many worker nodes for as much time as a queue can offer. On the contrary, jobs
-sized on the basis of an available backfill slot depend on the number of worker
-nodes that cannot be given to any other job in the Titan's queue at that moment
-in time.
+many worker nodes and walltime as a queue can offer. On the contrary, jobs sized
+according to an available backfill slot depend on the number of worker nodes and
+walltime that cannot be given to any other job at that moment in time.
 
@@
 staging were inherited from PanDA Pilot.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
-    \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates
-    the execution process of a detector simulation job, part of the production
-    ATLAS MD workflow.}
-\label{fig:panda_broker}
-\end{figure}
 
@@
 Centre (LRZ)~\cite{barreiro2016panda}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
+    \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates
+    the execution process of a detector simulation job, part of the production
+    ATLAS MC workflow.}
+\label{fig:panda_broker}
+\end{figure}

### 2017-04-04 Experiments.tex

@@
         \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
-    \caption{Weak scalability: average pilot duration, average  duration of a
-    single AthenaMP execution, and pilot's overhead as a function of pilot sizes (200, 500, 1000 and 2000 nodes).}
+    \caption{Weak scalability: average pilot duration, average duration of one AthenaMP execution, and pilot's overhead as a function of pilot sizes (200, 500, 1000 and 2000 nodes).}
 \label{fig:weakScal1a}
@@
     \caption{Weak scalability with multiple generations: average pilot
-    duration, average duration of five sequential AthenaMP executions, and
-    pilot's overhead for pilot sizes (256, 512, 1024 and 2048 nodes).}
+    duration, average duration of sequential AthenaMP executions, and
+    pilot's overhead for pilot with 256, 512, 1024 and 2048 nodes.}
 \label{fig:weakScal2a}
@@
     \caption{Strong scalability:  average pilot duration, average duration of
-    sequential AthenaMP executions, and pilot's overhead for different pilot
-    sizes (256, 512, 1024 and 2048 nodes).}
+    sequential AthenaMP executions, and pilot's overhead for pilots with 256, 512, 1024 and 2048 nodes.}
 \label{fig:strongScala}

### 2017-04-04 PandaOverview.tex

@@
 execution~\cite{crooks2012multi} but, so far, HEP applications have not required
-MPI.
+a message passing interface (MPI).
 % Jobs are supposed to be capable of setting up their own execution
@@
     process. Several subsystems, components, and architectural and
-    communication details are abstracted for clarity.}
+    communication details are abstracted to improve clarity.}
 \label{fig:architecture}

### 2017-04-04 nge.tex
saved 4-5 lines and minor corrections.

@@
    \includegraphics[width=\columnwidth]{figures/rp_architecture_compact_atlaswms_paper.pdf}
-  \caption{NGE Architecture as deployed on Titan. The PilotManager and the
-  UnitManager reside on one of Titan's DTN while the Agent is executed on
-  Titan's worker nodes. Boxes color coding: gray for entities external to NGE,
-  white for APIs, purple for NGE's modules, green for pilots, yellow for
-  module's components.}
+  \caption{NGE Architecture: The PilotManager and  UnitManager reside on a Titan
+  DTN while the Agent is executed on its worker nodes. Color coding: gray for
+  entities external to NGE, white for APIs, purple for NGE's modules, green for
+  pilots, yellow for module's components.}
 \label{fig:arch-overview}

### 2017-04-04 overall.tex

@@
     \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
-    \caption{Backfill availability: distribution of number of work nodes (left);
-    distribution of walltime in minutes (center); and a scatter plot of the two
-    variables (right). 62175 measures: mean number of work nodes available 691;
-    mean walltime available 126 minutes.}
+
+    \caption{Backfill availability: distribution of number of work nodes (left)
+    and of walltime in minutes (center). Scatter plot of the two variables
+    (right). 62175 measures: mean number of work nodes available 691; mean
+    walltime available 126 minutes.}
+
 \label{fig:backfill-distrib}
@@
     \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison_vsquashed.pdf}
-    \caption{Comparison between distributions of the time taken by a Geant4
-    detector simulation to simulate one event when placing 2 simulations (h1) or
-    1 simulation (h2) per CPU. 2 simulation use 16 cores per node, 1 simulation
-    8, 1 per compute unit. 50 Events; 1 Titan worker nodes; 16 work threads per
-    node; 100 events per node.}
+
+    \caption{Distributions of the time taken to simulate one event when placing
+    2 simulations (h1) or 1 simulation (h2) per Titan's CPU. 2 simulation use 16
+    cores per node, 1 simulation 8. 50 Events; 1 Titan worker nodes; 16 work
+    threads per node; 100 events per node.}
+
 \label{fig:comparison-8-16cores}

### 2017-04-04 Experiments.tex
space saving. 5 lines over.

@@
 
-Each experiment entails executing multiple instances of AthenaMP using the NGE
+Each experiment entails executing multiple instances of AthenaMP 
 to simulate a pre-determined number of events. All the experiments have been
@@
 
-We  measured the execution time of the pilots and of the AthenaMP  executed
+We  measured the execution time of the pilots and of AthenaMP 
 within them, collecting timestamps at  all stages of the execution. Experiments
-have been performed  by  submitting NGE's pilots  to Titan's batch queue.  The
+were performed  by  submitting NGE's pilots  to Titan's batch queue.  The
 turnaround time of an individual run is determined by queue waiting times. Since

### 2017-04-04 conclusion.tex
space saving. 5 lines over.

@@
 and reliability of both PanDA Broker and AthenaMP as deployed on Titan
-(\S\ref{sec:panda_titan}). Our characterization highlighted both the strengths
+(\S\ref{sec:panda_titan}). Our characterization highlighted the strengths
 and limitations of the current design and implementation: PanDA Brokers enable

### 2017-04-04 nge.tex
space saving. 5 lines over.

@@
 module is the pilot submitted to Titan and executed on its worker nodes. The
-communication between PanDA Broker and PanDA Server is abstracted away as it
+communication between PanDA Broker and Server is abstracted away as it
 is not immediately useful to evaluate the performance and capabilities of a

### 2017-04-04 nge.tex
down to 10!

@@
 
-As discussed in \S\ref{sec:panda_titan}, the absence of pilots imposes the
-static coupling between MPI scripts submitted to the PBS batch system and
-detector simulations. This makes the scheduling of multiple generations of
-workload on the same PBS job impossible: Once a statically defined number of
-detector simulations are packaged into a PBS job and this job is queued on
-Titan, no further simulations can be added to that job. New simulations have to
-be packaged into a new PBS job that needs to be submitted to Titan based upon
-backfill availability.
+The absence of pilots imposes the static coupling between MPI scripts
+submitted to the PBS batch system and detector simulations
+(\S\ref{sec:panda_titan}). This makes the scheduling of multiple generations
+of workload on the same PBS job impossible: once a statically defined number
+of detector simulations are packaged into a PBS job and this job is queued on
+Titan, no further simulations can be added to that job. New simulations have
+to be packaged into a new PBS job that needs to be submitted to Titan based
+upon backfill availability.
 
@@
 
-The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for communication
-and coordination of the execution of units. This environment is a critical
-component of the OpenMPI implementation, developed to support distributed
-high-performance computing applications operating in a heterogeneous
-environment~\cite{castain05:_open_rte, cug-2016}.
-
-ORTE provides a mechanism to create a \emph{``dynamic virtual machine''} (DVM)
-that spans multiple nodes. The DVM transparently provides support for
-interprocess communication, resource discovery and allocation, and process
-launching across a variety of platforms. Libraries enable the interaction
-between users and ORTE to enable the submission, monitoring and managing of
-tasks, avoiding filesystem bottlenecks and race conditions with network sockets.
-As a consequence, ORTE is able to minimize the system overhead while submitting
+The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
+communication and coordination of the execution of units. This environment is
+a critical component of the OpenMPI implementation, developed to support
+distributed high-performance computing applications operating in a
+heterogeneous environment~\cite{castain05:_open_rte, cug-2016}. ORTE provides
+a mechanism to create a \emph{``dynamic virtual machine''} (DVM) that spans
+multiple nodes. The DVM transparently provides support for interprocess
+communication, resource discovery and allocation, and process launching across
+a variety of platforms. Libraries enable the interaction between users and
+ORTE to enable the submission, monitoring and managing of tasks, avoiding
+filesystem bottlenecks and race conditions with network sockets. As a
+consequence, ORTE is able to minimize the system overhead while submitting
 tasks.

### 2017-04-04 abstract.tex
abstract --- long first pass.

@@
 
-Experiments at the Large Hadron Collider (LHC) face unprecedented computing
-challenges. Thousands of physicists analyze exabytes of data every year, using
-billions of computing hours on hundreds of computing sites worldwide. PanDA
-(Production and Distributed Analysis) is a workload management system (WMS)
-developed to meet the scale and complexity of LHC distributed computing for the
-ATLAS experiment. PanDA is the first exascale workload management system in HEP,
-executing millions of computing jobs per day, and processing over an exabyte of
-data in 2016. In this paper, we introduce the design and implementation of
-PanDA, describing its deployment on Titan, the third biggest supercomputer in
-the world. We analyze scalability, reliability and performance of PanDA on
-Titan, highlighting the challenges addressed by its architecture and
-implementation. We present preliminary results of experiments performed with the
-Next Generation Executer, a prototype we developed to meet new challenges of
-scale and resource heterogeneity.
+% Experiments at the Large Hadron Collider (LHC) face unprecedented computing
+% challenges. Thousands of physicists analyze exabytes of data every year, using
+% billions of computing hours on hundreds of computing sites worldwide. PanDA
+% (Production and Distributed Analysis) is a workload management system (WMS)
+% developed to meet the scale and complexity of LHC distributed computing for the
+% ATLAS experiment. PanDA is the first exascale workload management system in HEP,
+% executing millions of computing jobs per day, and processing over an exabyte of
+% data in 2016. In this paper, we introduce the design and implementation of
+% PanDA, describing its deployment on Titan, the third biggest supercomputer in
+% the world. We analyze scalability, reliability and performance of PanDA on
+% Titan, highlighting the challenges addressed by its architecture and
+% implementation. We present preliminary results of experiments performed with the
+% Next Generation Executer, a prototype we developed to meet new challenges of
+% scale and resource heterogeneity.
+
+The computing systems used by LHC experiments has historically consisted of
+the federation of hundreds to thousands of distributed resources,  ranging in
+scale from small to mid-size resource. As the LHC prepares for Run 3 in
+$~\approx$ 2022 and the high-luminosity era (Run 4), it is anticipated that
+the data volumes that will need analyzing will increase by factors of 10-100
+compared to the current phase (Run 2). Data will be larger in volume but will
+also require more sophisticated computational processing.  In spite of the
+impressive scale of the existing distributed computing solutions, the
+federation of small to mid-size resources will be insufficient to meet
+projected supply. This paper is a case study of how the ATLAS experiment has
+embraced leadership class high-performance computing in conjunction with
+traditional distributed high-throughput computing. Specifically, we discuss
+the experience of integrating PanDA WMS with the DOE leadership computing
+facility (Titan) to reach sustained production scales of approximately 51M
+core-hours a years. This state-of-practice paper provides three main
+contributions:  (i) a critical evaluation of the many design and operational
+considerations that have been taken to support the sustained, scalable and
+production usage of Titan for historically high-throughput workloads; (ii) a
+preliminary characterization of a next generation executor   for PanDA to
+support non- traditional heterogeneous workloads and advanced execution modes;
+and (iii) early lessons and guidance as the community looks forward to
+designing the next generation of online analytical platforms, for how current
+and future experimental and observational systems can be integrated with
+production supercomputers in a general and extensible manner.

### 2017-04-04 overall.tex
Negative spaces fixing
Small iteration 4

@@
 instances per DTN. Each broker submits and manages the execution of 15 to 300
-ATLAS jobs, one job for each worker node, and a theoretical maximum concurrent
-use of 96,000 cores. Since November 2015, PanDA Brokers have operated only in
-backfill mode, without a defined time allocation, and running at the lowest
-priority on Titan.
+ATLAS jobs, one job for each Titan's worker node, and a theoretical maximum
+concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have operated
+only in backfill mode, without a defined time allocation, and running at the
+lowest priority on Titan.
 
@@
 
-\begin{figure}[htp]
+\begin{figure}[!t]
     \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.pdf}
+    \vspace{-0.3in}
     \caption{Titan's total backfill availability: CPU core-hours (gray) and CPU
@@
 reached 18\% average efficiency, with a minimum 7.8\% efficiency on May 2016 and
-a maximum 30.9\% efficiency on February 2017. The total backfill availability
-was 20.3M in April 2016, and 17.6M in February 2017. This shows that the
-measured difference in efficiency did not depend on a comparable difference in
-the total backfill availability.
+a maximum 30.9\% efficiency on February 2017 (excluding the preliminary results
+of March). The total backfill availability was 20.3M in April 2016, and 17.6M in
+February 2017. This shows that the measured difference in efficiency did not
+depend on a comparable difference in the total backfill availability.
 
@@
 
-Fig.~\ref{fig:backfill-distrib} shows the distribution of backfill
-availability on Titan as a function of number of nodes and the time of their
-availability (i.e., walltime). We recorded these data by polling Titan's Moab
-scheduler at regular intervals during the experiment window time, while
-developing and deploying PanDA Brokers. The mean number of nodes was 691, and
-their mean walltime was 126 minutes. Detector simulations of 100 events, enable
-to use down to 1 node for 1/2 of the mean walltime of backfill availability. As
-such, it offers a good compromise for PanDA Broker efficiency.
+Fig.~\ref{fig:backfill-distrib} shows the distribution of backfill availability
+on Titan as a function of number of nodes and the time of their availability
+(i.e., walltime). We recorded these data by polling Titan's Moab scheduler at
+regular intervals during the experiment window time. The mean number of nodes
+was 691, and their mean walltime was 126 minutes. Detector simulations of 100
+events, enable to use down to 5/6 of the mean walltime of backfill availability.
+As such, it offers a good compromise for PanDA Broker efficiency.
 
-\begin{figure*}%[htp]
+\begin{figure*}[!t]
     \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_wnodes_distribution.pdf}
@@
     \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
-
+   \vspace{-0.1in}
     \caption{Backfill availability: distribution of number of work nodes (left)
@@
 
-Usually, detector simulations performed on the Grid process 10 times the number
-of events than those on Titan. This explains the difference between the
-percentage of detector simulations and of events computed by those simulations
-on Titan. PanDA Broker could fit the number of events to the walltime of the
-backfill availability on the base of the distributions of
-Fig.~\ref{fig:backfill-distrib}. That specific number of event could then be
-pulled from the PanDA Event service~\cite{calafiura2015atlas} and given as input
-to one or more simulations. Once packaged into the MPI script submitted to
-titan's PBS batch system, these simulations would better fit backfill
-availability, increasing the efficiency of PanDA Brokers.
+% Usually, detector simulations performed on the Grid process 10 times the
+% number of events than those on Titan. This explains the difference between the
+% percentage of detector simulations and of events computed by those simulations
+% on Titan.
+PanDA Broker could fit the number of events to the walltime of each available
+backfill slot on the base of the distributions of the time taken by one event to
+be simulated. That specific number of event could then be pulled from the PanDA
+Event service~\cite{calafiura2015atlas} and given as input to one or more
+simulations. Once packaged into the MPI script submitted to titan's PBS batch
+system, these simulations would better fit their available backfill slot,
+contributing to increase the efficiency of PanDA Brokers.
 
@@
 
-Currently, each PanDA Broker creates, submits, and monitor a single MPI PBS
+Currently, each PanDA Broker creates, submits, and monitors a single MPI PBS
 script at a time. This design is inherited from PanDA Pilot where a single
@@
 on the the number of concurrent PanDA Brokers instantiated on the DTNs: When all
-the 20 PanDA Brokers have submitted a MPI PBS Job, further backfill availability
-cannot be used.
+the 20 PanDA Brokers have submitted a MPI PBS script, further backfill
+availability cannot be used.
 
-In September 2016, increasing the number of concurrent PanDA Broker from 4 to 20
-markedly improved efficiency (see Fig.~\ref{fig:backfill-utilization}) but
-further research is undergoing to understand whether an even larger number
-brokers would yield similar results. This research focuses on evaluating the the
+In September 2016, increasing the number of concurrent PanDA brokers from 4 to
+20 markedly improved efficiency (see Fig.~\ref{fig:backfill-utilization}) but
+further research is undergoing to understand whether an even larger number of
+brokers would yield similar results. This research focuses on evaluating the
 overheads of input/output files staging, including its impact on DTNs, and on an
-alternative design of PanDA Broker to enable the concurrent submission of
+alternative design of PanDA Broker that enables the concurrent submission of
 multiple MPI scripts~\cite{barreiro2016panda}.
+
+The current design and architecture of the PanDA Broker is proving to be as
+reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
+the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
+while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
+Brokers were responsible for around the 19\% of the failures, compared to the
+29\% of failures produced by the JobDispatcher module of the PanDA Server, and
+the 13\% failures produced by the Geant4 toolkit. The current failure rate of
+the PanDA Brokers confirms the benefits of reusing most of the code base of the
+PanDA Pilot for implementing the PanDA Broker. It also shows that adopting
+third-party libraries like RADICAL-SAGA did not have a measurable adverse effect
+on reliability.
 
@@
 between the worker nodes and the DTNs, including the operations required to
-access of small python scripts.
+access small python scripts.
 
-Initially, all the shared libraries of AthenaMP and the python scripts for the
-configuration stage were stored on the Spider 2 Lustre file system. However, the
-I/O patterns of the initialization and configuration stages degraded the
-performance of the filesystem. This was addressed by moving the AthenaMP
-distribution to a read-only NFS directory, shared among DTNs and worker nodes.
-NFS eliminated the problem of metadata contention, improving metadata read
-performance from $\approx$6,300 seconds on Lustre to $\approx$1,500 seconds on
-NFS.
+Initially, all the shared libraries of AthenaMP and the python scripts for its
+configuration were stored on the Spider 2 Lustre file system. However, the I/O
+patterns of the initialization and configuration stages degraded the performance
+of the filesystem. This was addressed by moving the AthenaMP distribution to a
+read-only NFS directory, shared among DTNs and worker nodes. NFS eliminated the
+problem of metadata contention, improving metadata read performance from
+$\approx$6,300 seconds on Lustre to $\approx$1,500 seconds on NFS.
 
-Once set up, AthenaMP is used to execute 16 concurrent Geant4 simulators on each
-of Titan's worker node. Geant4 requires to read events descriptions from a
-filesystem and simulate them as they would happen within the ATLAS detector. We
-characterized both the compute performance of the simulation and the impact of
-acquiring event descriptions on the filesystem.
+Once initialized and configured, AthenaMP is used to execute 16 concurrent
+Geant4 simulators on a Titan's worker node. Geant4 requires to read events
+descriptions from a filesystem and simulate them as they would happen within the
+ATLAS detector. We characterized both the compute performance of the simulation
+and the impact of acquiring event descriptions on the filesystem.
 
@@
     \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison_vsquashed.pdf}
-
+   \vspace{-0.3in}
     \caption{Distributions of the time taken to simulate one event when placing
@@
 We used the rate of events processes per minute as a measure of the efficiency
-of executing the same detector simulation on both Titan or Grid sites.
-Comparisons of the efficiencies of Titan to  the BNL and SIGNET Grid sites,
-normalized for 8 cores, show that the effective performance per-core at Titan is
-$\approx$0.57 event per minute, roughly 1/2 of BNL and  1/3 of SIGNET
-performances.
+of executing the same detector simulation on Titan or Grid sites. Comparisons of
+the efficiencies of Titan to the BNL and SIGNET Grid sites, normalized for 8
+cores, show that the effective performance per-core at Titan is $\approx$0.57
+event per minute, roughly 1/2 of BNL and  1/3 of SIGNET performances.
 
@@
 average newer than the CPU of Titan. The heterogeneity of the Grid sites' CPUs
-explain the higher performance variance compared to the performance consistency
-measured on Titan.
+explain the higher performance variance we observed compared to the performance
+consistency we measured on Titan.
 
@@
 halving the time currently required by initiating a Genat4 simulation.
-
-The current design and architecture of the PanDA Broker is proving to be as
-reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
-the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
-while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
-Brokers were responsible for around the 19\% of the failures, compared to the
-29\% of failures produced by the JobDispatcher module of the PanDA Server, and
-the 13\% failures produced by the Geant4 toolkit. The current failure rate of
-the PanDA Brokers confirms the benefits of reusing most of the code base of the
-PanDA Pilot for implementing the PanDA Broker. It also shows that adopting
-third-party libraries like RADICAL-SAGA did not have a measurable adverse effect
-on reliability.

### 2017-04-04 introduction.tex

@@
 Historically, the computing systems used by LHC experiments consisted of the
-federation of a hundreds to thousands of distributed resources, %\textemdash{}
+federation of hundreds to thousands of distributed resources, %\textemdash{}
 ranging in scale from small to mid-size resource~\cite{foster2003grid}. Although

### 2017-04-04 abstract.tex
reduction in abstract.

@@
 
+% As the LHC prepares for Run 3 in
+% $~\approx$ 2022 and the high-luminosity era (Run 4), it is anticipated that
+% the data volumes that will need analyzing will increase by factors of 10-100
+% compared to the current phase (Run 2). Data will be larger in volume but will
+% also require more sophisticated computational processing.
+
 The computing systems used by LHC experiments has historically consisted of
 the federation of hundreds to thousands of distributed resources, ranging in
-scale from small to mid-size resource. As the LHC prepares for Run 3 in
-$~\approx$ 2022 and the high-luminosity era (Run 4), it is anticipated that
-the data volumes that will need analyzing will increase by factors of 10-100
-compared to the current phase (Run 2). Data will be larger in volume but will
-also require more sophisticated computational processing.  In spite of the
-impressive scale of the existing distributed computing solutions, the
-federation of small to mid-size resources will be insufficient to meet
-projected supply. This paper is a case study of how the ATLAS experiment has
-embraced leadership class high-performance computing in conjunction with
-traditional distributed high-throughput computing. Specifically, we discuss
-the experience of integrating PanDA WMS with the DOE leadership computing
-facility (Titan) to reach sustained production scales of approximately 51M
-core-hours a years. This state-of-practice paper provides three main
-contributions:  (i) a critical evaluation of the many design and operational
-considerations that have been taken to support the sustained, scalable and
-production usage of Titan for historically high-throughput workloads; (ii) a
-preliminary characterization of a next generation executor   for PanDA to
-support non- traditional heterogeneous workloads and advanced execution modes;
-and (iii) early lessons and guidance as the community looks forward to
-designing the next generation of online analytical platforms, for how current
-and future experimental and observational systems can be integrated with
-production supercomputers in a general and extensible manner.
+scale from small to mid-size resource.  In spite of the impressive scale of
+the existing distributed computing solutions, the federation of small to mid-
+size resources will be insufficient to meet projected demands of Run 3 (and
+4). This paper is a case study of how the ATLAS experiment has embraced Titan
+-- a DOE leadership facility in conjunction with traditional distributed high-
+throughput computing to reach sustained production scales of approximately 51M
+core-hours a years. The three main contributions of this paper are:  (i) a
+critical evaluation of design and operational considerations  to support the
+sustained, scalable and production usage of Titan;  (ii) a preliminary
+characterization of a next generation executor for PanDA to support new
+workloads and  advanced execution modes; and (iii) early lessons as  for how
+current and future experimental and observational systems can be integrated
+with production supercomputers and other platforms in a general and extensible
+manner.

### 2017-04-04 conclusion.tex
conclusions. refinements.

@@
 
-More in general, this paper highlighted the fundamental role of WMS for
+More generally, this paper highlights the fundamental role of WMS for
 experimental science. HEP was amongst the first, if not the very first
 experimental community to realize the importance of using WMS to manage their
-computational campaign(s). As computing become increasingly critical for a range
-of experiments, the impact of WMS on HEP foreshadows the importance of WMS for
-other experiments (such as SKA, LSST etc.). % and their computational campaigns.
-These experiments will have their own workload characteristics, resources types
-and federation constraints, as well metrics of performance.  The design and
-experience captured in this paper should prove invaluable for designing
-computational campaigns and will provide a baseline to evaluate the relative
-merits of different approaches.
+computational campaign(s). As computing becomes increasingly critical for a
+range of experiments, the impact of WMS on HEP foreshadows the importance of
+WMS for other experiments (such as SKA, LSST etc.). % and their computational
+campaigns. These experiments will have their own workload characteristics,
+resources types and federation constraints, as well metrics of performance.
+The design and experience captured in this paper will prove invaluable for
+designing WMS to support computational campaigns and will provide a baseline
+to evaluate the relative merits of different approaches.
 
@@
 experiment -- HEPCloud~\cite{hepcloud,googlehep} represent similar yet
-different approaches to dynamic resource federation.
-
-Our experience demonstrated the advantages arising from a  systems engineering
-and abstractions based approach, in particular the pilot abstraction. As the
-type, heterogeneity and complexity of future platforms  (e.g., Exascale,
-Commercial Clouds) for science increases, there will be a greater need and
-emphasis on abstractions to ensure WMS can both utilize as well as support
-platforms.
+different approaches to dynamic resource federation. As the type,
+heterogeneity and complexity of future platforms  (e.g., Exascale, Commercial
+Clouds) for science increases, there will be a greater need and emphasis on
+abstractions to ensure WMS can both utilize as well as support evolving
+platforms. Our experience demonstrated the advantages arising from an
+abstractions based approach \textendash{} in particular scalable implementations
+of pilot abstraction.

### 2017-04-04 nge.tex
Negative spaces fixing

@@
-As seen in \S\ref{sec:panda_deployment}, PanDA Broker was  deployed on the DTNs
+As seen in \S\ref{ssec:panda_titan}, PanDA Broker was  deployed on the DTNs
 as Titan's worker nodes lack connectivity to the wide area network The lack of
@@
 of remaining walltime and the distribution of execution time per event, as shown
-in \S\ref{ssec:panda_titan}, Fig.~\ref{fig:100event- distrib}. These
+in \S\ref{ssec:panda_titan}, Fig.~\ref{fig:comparison-8-16cores}. These
 capabilities would increase the efficiency of the PanDA Broker when there is a
@@
    \includegraphics[width=\columnwidth]{figures/rp_architecture_compact_atlaswms_paper.pdf}
+   \vspace{-0.3in}
   \caption{NGE Architecture: The PilotManager and  UnitManager reside on a Titan

### 2017-04-04 abstract.tex
minor tweak of abstract.

@@
 The computing systems used by LHC experiments has historically consisted of
-the federation of hundreds to thousands of distributed resources, ranging in
-scale from small to mid-size resource.  In spite of the impressive scale of
-the existing distributed computing solutions, the federation of small to mid-
-size resources will be insufficient to meet projected demands of Run 3 (and
-4). This paper is a case study of how the ATLAS experiment has embraced Titan
--- a DOE leadership facility in conjunction with traditional distributed high-
+the federation of hundreds to thousands of distributed resources, ranging from
+small to mid-size resource.  In spite of the impressive scale of the existing
+distributed computing solutions, the federation of small to mid-size
+resources will be insufficient to meet projected future demands.
+This paper is a case study of how the ATLAS experiment has embraced Titan -- a
+DOE leadership facility in conjunction with traditional distributed high-
 throughput computing to reach sustained production scales of approximately 51M
@@
 characterization of a next generation executor for PanDA to support new
-workloads and  advanced execution modes; and (iii) early lessons as  for how
+workloads and  advanced execution modes; and (iii) early lessons for how
 current and future experimental and observational systems can be integrated

### 2017-04-04 introduction.tex

@@
 a DOE leadership computing facility called Titan to reach sustained production
-scales of approximately 51M core-hours a years. We also discusses the
+scales of approximately 51M core-hours a year. We also discusses the
 investigation of a task execution runtime system on Titan, that is based on the

### 2017-04-04 PandaEvolution.tex
Negative spaces fixing
Attempting the backflip, careful with faceplants...

@@
 will be insufficient for the Monte Carlo production workflow and any extra
-activity. To alleviate these challenges, ATLAS is engaged in a program to expand
-the current computing model to include additional resources such as the
-opportunistic use of supercomputers.
+activity. To alleviate these challenges, ATLAS is expanding the current
+computing model to include additional resources such as the opportunistic use of
+supercomputers.
 
@@
     \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
+    \vspace{-0.3in}
     \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates

### 2017-04-04 overall.tex
Attempting the backflip, careful with faceplants...

@@
 PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and
-AthenaMP. All the measurements were performed between January 2016 and February
-2017, hereafter called `experiment time window'.
+AthenaMP. We discuss challenges and limitations of our approach at multiple
+levels arising from the specifics of workload, middleware and methods. All the
+measurements were performed between January 2016 and February 2017, hereafter
+called ‘experiment time window
 

### 2017-06-20 paper.tex
Iterate and move related work
Pass on NGE experiments and scattered space savings
PanDA Broker Architecture and anonymization
minor refinement of introduction.
  remove abstract to give to Jack et al.
Iteration 4 and scattered editing
Iteration conclusions
Clean up
Iteration introduction
space saving. 5 lines over.
down to 10!
Negative spaces fixing
fixing text over run / S7.
Fix spaces
Convert paper to IEEETran style for eScience conference

--- 2017-03-19 paper.tex
@@
-\documentclass[sigconf, screen]{acmart}
+\documentclass[conference]{IEEEtran}
+
+% ---------------------------------------------------------------------------
+% Authors are invited to submit unpublished, original work, using the IEEE
+% 8.5 × 11 manuscript guidelines: double-column text using single-spaced 10
+% point font on 8.5 × 11 inch pages. Templates are available from http://www.
+% ieee.org/conferences_events/conferences/publishing/templates.html.
+%
+% The conference proceedings will be made available online through the IEEE
+% Digital Library.
+%
+% Three types of papers will be accepted:
+%
+% - Full research papers (up to 10 pages): focus on new research achievements
+%   in eScience, spanning core technologies and applications.
+% - Experience papers (up to 10 pages): focus on practical outcomes through
+%   applying eScience tools and techniques to novel problems.
+% ---------------------------------------------------------------------------
 
@@
 
-% DOI
-% \acmDOI{10.475/123_4}
-
-% ISBN
-% \acmISBN{123-4567-24-567/08/06}
-
-%Conference
-\acmConference[SC2017]{The International Conference for High Performance Computing, Networking, Storage and Analysis}{November 2017}{Denver, Colorado USA}
-\acmYear{2017}
-\copyrightyear{2017}
-
-% \acmPrice{15.00}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-% -----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
 % BEGIN DOCUMENT
-% -----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
 \begin{document}
 
-\title{A Case Study on Abstractions to Converge High-Throughput and High-Performance Computing}
+\title{Converging High-Throughput and High-Performance Computing: A Case Study}
 
-\author{Author 1}
-\affiliation{\institution{Institution 1}}
-\email{email1}
+% for over three affiliations, or if they all won't fit within the width
+% of the page, use this alternative format:
+% 
+% \author{
+%     \IEEEauthorblockN{
+%         Author 1\IEEEauthorrefmark{1},
+%         Author 2\IEEEauthorrefmark{2},
+%         Author 3\IEEEauthorrefmark{3}, 
+%         Author 4\IEEEauthorrefmark{3}, 
+%         Author 5\IEEEauthorrefmark{4} and
+%         Author 6\IEEEauthorrefmark{4}
+%     }
+%     \IEEEauthorblockA{
+%         \IEEEauthorrefmark{1}Department\\Institution, address, email
+%     }
+%     \IEEEauthorblockA{
+%         \IEEEauthorrefmark{2}Department\\Institution, address, email
+%     }
+%     \IEEEauthorblockA{
+%         \IEEEauthorrefmark{3}Department\\Institution, address, email
+%     }
+%     \IEEEauthorblockA{
+%         \IEEEauthorrefmark{4}Department\\Institution, address, email
+%     }
+% }
 
-\author{Author 2}
-\affiliation{\institution{Institution 2}}
-\email{email2}
-
-\author{Author 3}
-\affiliation{\institution{Institution 3}}
-\email{email3}
-
-\author{Author 4}
-\affiliation{\institution{Institution 4}}
-\email{email4}
-
-\author{Author 5}
-\affiliation{\institution{Institution 5}}
-\email{email5}
-
-\author{Author 6}
-\affiliation{\institution{Institution 6}}
-\email{email6}
-
-% The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{Author 1 et al.}
-
-% ----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
 % ABSTRACT
-% ----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
+\maketitle
 \begin{abstract}
@@
 \end{abstract}
+% \keywords{Distributed computing, Workload management system}
 
-%
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below.
-%
-\begin{CCSXML}
-\end{CCSXML}
-
-% \ccsdesc[500]{Computer systems organization~Embedded systems}
-% \ccsdesc[300]{Computer systems organization~Redundancy}
-% \ccsdesc{Computer systems organization~Robotics}
-% \ccsdesc[100]{Networks~Network reliability}
-
-\keywords{ACM proceedings, \LaTeX, text tagging}
-
-\maketitle
-
-% ----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
 % I - INTRODUCTION
-% ----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
 \section{Introduction}
 \label{sec:intro}
-
 \input{introduction}
 
-% ----------------------------------------------------------------------------
-% II - RELATED WORK
-% ----------------------------------------------------------------------------
-\section{Related Work}
-\label{sec:related}
-
-% \input{WMSLiterature}
-% Please do not correct bckground to background.
-\input{bckground}
-
-% ----------------------------------------------------------------------------
-% III - PanDA OVERVIEW
-% ----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
+% II - PanDA OVERVIEW
+% ---------------------------------------------------------------------------
 \section{PanDA Overview}
 \label{sec:panda_overview}
-
 \input{PandaOverview}
 
-% ----------------------------------------------------------------------------
-% IV - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
-% ----------------------------------------------------------------------------
-\section{Deploying PanDA on a Leadership-scale system}
-\label{sec:panda_deployment}
-
+% ---------------------------------------------------------------------------
+% III - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
+% ---------------------------------------------------------------------------
+%\section{Deploying PanDA on a Leadership-scale system}
+\section{Deploying PanDA on Titan}
+\label{sec:panda_titan}
 \input{PandaEvolution}
 
-% ----------------------------------------------------------------------------
-% V - ANALYSIS AND DISCUSSION
-% ----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
+% IV - ANALYSIS AND DISCUSSION
+% ---------------------------------------------------------------------------
 \section{Analysis and Discussion}
 \label{sec:panda_titan}
-
 \input{overall}
 
-% ----------------------------------------------------------------------------
-% VII - THE NEXT GENERATION EXECUTOR
-% ----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
+% VI - THE NEXT GENERATION EXECUTOR
+% ---------------------------------------------------------------------------
 % \section{PANDA: Future/RoadMap}
@@
 \label{sec:panda_roadmap}
-
 \input{nge}
@@
 
-% ----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
+% VII - RELATED WORK
+% ---------------------------------------------------------------------------
+%\vspace{-0.4in}
+\section{Related Work}
+\label{sec:related}
+\input{bckground}
+
+% ---------------------------------------------------------------------------
 % VIII - CONCLUSION
-% ----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
+%\vspace{-0.1in}
 \section{Conclusion}
 \label{sec:conclusion}
-
 \input{conclusion}
 
-% ----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
 % ACKNOWLEDGEMENTS
-% ----------------------------------------------------------------------------
-\section*{Acknowledgements}
-\label{sec:ack}
+% ---------------------------------------------------------------------------
+% \section*{Acknowledgements}
+% \label{sec:ack}
+%
+% \input{ack}
 
-\input{ack}
+% ---------------------------------------------------------------------------
+% REFERENCES
+% ---------------------------------------------------------------------------
+\newpage
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,bibliography}
 
-% ----------------------------------------------------------------------------
-% REFERENCES
-% ----------------------------------------------------------------------------
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{bibliography}
+% ---------------------------------------------------------------------------
+% A - APPENDIX
+% ---------------------------------------------------------------------------
+% \appendix
+% \section{Artifact Description --- Converging High-Throughput and High-Performance Computing: A Case Study}
+% \label{apnx:ac}
+% \input{appendix}
 

### 2017-06-26 Experiments.tex
down to 10!
Negative spaces fixing
refinement, reduction and reviewers..

@@
 
-\begin{figure}[!htb]
+\begin{figure}[!t]
         \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
+   	\vspace{-0.3in}
     \caption{Weak scalability: average pilot duration, average duration of one AthenaMP execution, and pilot's overhead as a function of pilot sizes (200, 500, 1000 and 2000 nodes).}
@@
 average execution time of five sequential generations of AthenaMP, and the
-corresponding overhead. We observe that the difference between the two durations
+corresponding overhead. The difference between the two durations
 is more marked than in the previous experiments. Despite this, we notice that
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!t]
         \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak2.pdf}
+        \vspace{-0.3in}
     \caption{Weak scalability with multiple generations: average pilot
@@
 
+%\vspace{-0.2in}
+
 \subsubsection{Strong scalability}
 
-The last experiments  study strong scalability by running the same number of
+We investigate strong scalability by running the same number of
 tasks for different pilot sizes. We used 2048 AthenaMP instances and  pilots
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!t]
         \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/strong.pdf}
+        \vspace{-0.3in}
     \caption{Strong scalability:  average pilot duration, average duration of
@@
 \end{figure}
+
+\vspace{-0.05in}

### 2017-06-26 conclusion.tex
refinement, reduction and reviewers..

@@
 
-We characterized the state of practice by evaluating the efficiency, scalability
+We characterized the experience by evaluating the efficiency, scalability
 and reliability of both PanDA Broker and AthenaMP as deployed on Titan
@@
 will have to support ever increasing complex workloads with varying models for
-dynamic resource federation. Recent efforts with Google Cloud for the CMS
-experiment -- HEPCloud~\cite{hepcloud,googlehep} represent similar yet
-different approaches to dynamic resource federation. As the type,
-heterogeneity and complexity of future platforms  (e.g., Exascale, Commercial
-Clouds) for science increases, there will be a greater need and emphasis on
-abstractions to ensure WMS can both utilize as well as support evolving
-platforms. Our experience demonstrated the advantages arising from an
-abstractions based approach \textendash{} in particular scalable implementations
-of pilot abstraction.
+dynamic resource federation.
+
+% Recent efforts with Google Cloud for the CMS
+% experiment -- HEPCloud~\cite{hepcloud,googlehep} represent similar yet
+% different approaches to dynamic resource federation. As the type,
+% heterogeneity and complexity of future platforms  (e.g., Exascale, Commercial
+% Clouds) for science increases, there will be a greater need and emphasis on
+% abstractions to ensure WMS can both utilize as well as support evolving
+% platforms. Our experience demonstrated the advantages arising from an
+% abstractions based approach \textendash{} in particular scalable implementations
+% of pilot abstraction.

### 2017-06-26 introduction.tex
refinement, reduction and reviewers..

@@
 The Large Hadron Collider (LHC) was created to explore the fundamental
-properties of matter for the next decades. Multiple experiments at LHC have
+properties of matter. Multiple experiments at LHC have
 collected and distributed hundreds of petabytes of data worldwide to hundreds of
@@
 other, the management of the distribution of workloads across many heterogeneous
-resources to ensure the effective utilization of resources and efficient
+resources, the effective utilization of resources and efficient
 execution of workloads presents non-trivial challenges.
@@
 be larger in volume but will also require more sophisticated computational
-processing. In spite of the impressive scale of the ATLAS distributed computing
-system, demand for computing systems will significantly outstrip current and projected supply. %(availability).
+processing. In spite of the impressive scale of the ATLAS distributed
+computing system, demand for computing systems will significantly outstrip
+current and projected supply.  There are multiple levels at which this problem
+needs to be addressed: the utilization of emerging parallel architectures
+(e.g., platforms); algorithmic and advances in analytical methods (e.g., use
+of Machine Learning); and the ability to exploit different platforms (e.g.,
+clouds and supercomputers).
 
-There are multiple levels at which this problem needs to be addressed: the
-utilization of emerging parallel architectures (e.g., platforms); algorithmic
-and advances in analytical methods (e.g., use of Machine Learning); and the
-ability to exploit different platforms (e.g., clouds and supercomputers).
-
-This paper is a case study of how the ATLAS experiment has ``broken free'' of
-the traditional computational approach of high-throughput computing on
-distributed resources to embrace new platforms, in particular high-performance
-computers. Specifically, we discuss the experience of integrating PanDA WMS with
-a DOE leadership computing facility called Titan to reach sustained production
-scales of approximately 51M core-hours a year. We also discusses the
-investigation of a task execution runtime system on Titan, that is based on the
-pilot-abstractions and which allows advanced execution modes of the ATLAS
-workload as well as the enhanced support for heterogeneous workloads such as
-molecular dynamics.
+This paper represents the experience of how the ATLAS experiment has ``broken
+free'' of the traditional computational approach of high-throughput computing
+on distributed resources to embrace new platforms, in particular  high-
+performance computers. Specifically, we discuss the experience of integrating
+PanDA WMS with a US DOE leadership computing facility called Titan to reach
+sustained production scales of approximately 51M core-hours a year. We also
+discusses the investigation of a pilot-abstraction based task execution
+runtime system on Titan, which allows advanced execution modes of the ATLAS
+workload as well as the enhanced support for heterogeneous workloads, 
+e.g., molecular dynamics.
 
@@
 
-This state-of-practice paper provides three main contributions:  (i) a critical
+This experience paper provides three main contributions:  (i) a critical
 evaluation of the many design and operational considerations that have been

### 2017-06-26 paper.tex
refinement, reduction and reviewers..

@@
 \title{Converging High-Throughput and High-Performance Computing: A Case Study}
+
+\author{
+    \IEEEauthorblockN{
+        Alessio Angius\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},
+        Danila Oleynik\IEEEauthorrefmark{1}\IEEEauthorrefmark{5},
+        Sergey Panitkin\IEEEauthorrefmark{1}\IEEEauthorrefmark{4},
+        Matteo Turilli\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},\\
+        Kaushik De\IEEEauthorrefmark{5},
+        Alexei Klimentov\IEEEauthorrefmark{4},
+        Sarp H. Oral\IEEEauthorrefmark{6},
+        Jack C. Wells\IEEEauthorrefmark{6} and
+        Shantenu Jha\IEEEauthorrefmark{2}
+    }
+    \IEEEauthorblockA{
+        \IEEEauthorrefmark{1}First authors, alphabetical order
+    }
+    \IEEEauthorblockA{
+        \IEEEauthorrefmark{2}Department of Electric and Computer Engineering, Rutgers University, NJ, USA
+    }
+    \IEEEauthorblockA{
+        \IEEEauthorrefmark{4}Physics Department, Brookhaven National Laboratory, NY, USA
+    }
+    \IEEEauthorblockA{
+        \IEEEauthorrefmark{5}Physics Department, University of Texas Arlington, TX, USA
+    }
+    \IEEEauthorblockA{
+        \IEEEauthorrefmark{6}Oak Ridge Leadership Computing Facility, Oak Ridge National Laboratory, TN, USA
+    }
+}
 

### 2017-06-26 conclusion.tex
ongoing refinement. but more in conclusion.

@@
 
-More generally, this paper highlights the fundamental role of WMS for
-experimental science. HEP was amongst the first, if not the very first
-experimental community to realize the importance of using WMS to manage their
-computational campaign(s). As computing becomes increasingly critical for a
-range of experiments, the impact of WMS on HEP foreshadows the importance of
-WMS for other experiments (such as SKA, LSST etc.). % and their computational
-campaigns. These experiments will have their own workload characteristics,
-resources types and federation constraints, as well metrics of performance.
-The design and experience captured in this paper will prove invaluable for
-designing WMS to support computational campaigns and will provide a baseline
-to evaluate the relative merits of different approaches.
+% More generally, this paper highlights the fundamental role of WMS for
+% experimental science.
+
+HEP was amongst the first, if not the first experimental community to realize
+the importance of using WMS to manage their computational campaign(s). As
+computing becomes increasingly critical for a range of experiments, the
+experience foreshadows the importance of WMS for other experiments (such as
+SKA, LSST etc.).  These experiments will have their own workload
+characteristics, resources types and federation constraints, as well metrics
+of performance. The experience captured in this paper will prove
+invaluable for designing WMS for computational campaigns and will provide a
+baseline to evaluate the relative merits of different approaches.
 

### 2017-06-26 nge.tex
Formatting
minor reduction.

@@
 As seen in \S\ref{ssec:panda_titan}, PanDA Broker was  deployed on the DTNs
-as Titan's worker nodes lack connectivity to the wide area network The lack of
-pilot capabilities impacts both the efficiency and the flexibility of PanDA's
-execution process. Pilots could improve efficiency by increasing throughput and
-enabling greater backfill utilization. Further, pilots makes it easier to
-support heterogeneous workloads.
+as Titan's worker nodes lack connectivity to the wide area network The lack
+of pilot capabilities impacts both the efficiency and the flexibility of
+PanDA's execution process. Pilots could improve efficiency by increasing
+throughput and enabling greater backfill utilization. Further, pilots makes
+it easier to support heterogeneous workloads.
 
@@
 
-The support of  multiple generations of workload would enable more efficient use
-of the backfill availability of walltime. Currently, when a set of simulations
-ends, the PBS job also ends, independent of whether more wall-time would still
-be available. With a pilot, additional simulations could be executed  to utilize
-all the available wall-time, while avoiding further job packaging and submission
-overheads.
+The support of  multiple generations of workload would enable more efficient
+use of the backfill availability of walltime. Currently, when a set of
+simulations ends, the PBS job also ends, independent of whether more
+wall-time would still be available. With a pilot, additional simulations
+could be executed  to utilize all the available wall-time, while avoiding
+further job packaging and submission overheads.
 
-Multiple generations would also relax two assumptions of the current execution
-model: knowing the number of simulations before submitting the MPI script, and
-having a fixed number of events per simulation (100 at the moment). Pilots would
-enable the scheduling of simulations independently from whether they were
-available at the moment of submitting the pilot. Further, simulations with a
-varying number of events could be scheduled on a pilot, depending on the amount
-of remaining walltime and the distribution of execution time per event, as shown
-in \S\ref{ssec:panda_titan}, Fig.~\ref{fig:comparison-8-16cores}. These
-capabilities would increase the efficiency of the PanDA Broker when there is a
-large difference between the number of cores and walltime.
+Multiple generations would also relax two assumptions of the current
+execution model: knowing the number of simulations before submitting the MPI
+script, and having a fixed number of events per simulation (100 at the
+moment). Pilots would enable the scheduling of simulations independently from
+whether they were available at the moment of submitting the pilot. Further,
+simulations with a varying number of events could be scheduled on a pilot,
+depending on the amount of remaining walltime and the distribution of
+execution time per event, as shown in \S\ref{ssec:panda_titan},
+Fig.~\ref{fig:comparison-8-16cores}. These capabilities would increase the
+efficiency of the PanDA Broker when there is a large difference between the
+number of cores and walltime.
 
@@
 develop a specific PBS script for each payload. The submission process would
-also be MPI-independent, as MPI is used for coordination among multiple worker
-nodes, not by the payload.
+also be MPI-independent, as MPI is used for coordination among multiple
+worker nodes, not by the payload.
 
@@
 instantiate a PilotManager and a UnitManager.   The PilotManager submits
-pilots to Titan's PBS  batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit MPI scripts. Once
-scheduled, the pilot's Agent is bootstrapped on Titan's MOM node and
-the Agent's Executors on the on the worker
-nodes, and the UnitManager schedules units to the Agent's Scheduler
-(Fig.~\ref{fig:arch-overview}, solid arrow). The Agent's Scheduler schedules
-the units on one or more Agent's Executor for execution. The Agent's executors
-can manage one or more worker nodes, depending on performance evaluations. The
-UnitManager and the Agent communicate via a database that is instantiated on
-one of Titan's DTN so as to be reachable by both modules.
+pilots to Titan's PBS  batch system via SAGA API (Fig.~\ref{fig:arch-
+overview}, dash arrow), as done by PanDA Broker to submit MPI scripts. Once
+scheduled, the pilot's Agent is bootstrapped on Titan's MOM node and the
+Agent's Executors on the on the worker nodes, and the UnitManager schedules
+units to the Agent's Scheduler (Fig.~\ref{fig:arch-overview}, solid arrow).
+The Agent's Scheduler schedules the units on one or more Agent's Executor for
+execution. The Agent's executors can manage one or more worker nodes,
+depending on performance evaluations. The UnitManager and the Agent
+communicate via a database that is instantiated on one of Titan's DTN so as to
+be reachable by both modules.
 
@@
 communication and coordination of the execution of units. This environment is
-a critical component of the OpenMPI implementation, developed to support
-distributed high-performance computing applications operating in a
-heterogeneous environment~\cite{castain05:_open_rte, cug-2016}. ORTE provides
-a mechanism to create a \emph{``dynamic virtual machine''} (DVM) that spans
-multiple nodes. The DVM transparently provides support for interprocess
-communication, resource discovery and allocation, and process launching across
-a variety of platforms. Libraries enable the interaction between users and
-ORTE to enable the submission, monitoring and managing of tasks, avoiding
-filesystem bottlenecks and race conditions with network sockets. As a
-consequence, ORTE is able to minimize the system overhead while submitting
-tasks.
+a critical component of the OpenMPI implementation ~\cite{castain05:_open_rte,
+cug-2016}. ORTE is able to minimize the system overhead while submitting tasks
+by avoiding filesystem bottlenecks and race conditions with network sockets.
+
+% The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
+% communication and coordination of the execution of units. This environment is
+% a critical component of the OpenMPI implementation, developed to support
+% distributed high-performance computing applications operating in a
+% heterogeneous environment~\cite{castain05:_open_rte, cug-2016}. ORTE provides
+% a mechanism to create a \emph{``dynamic virtual machine''} (DVM) that spans
+% multiple nodes. The DVM transparently provides support for interprocess
+% communication, resource discovery and allocation, and process launching across
+% a variety of platforms. Libraries enable the interaction between users and
+% ORTE to enable the submission, monitoring and managing of tasks, avoiding
+% filesystem bottlenecks and race conditions with network sockets. As a
+% consequence, ORTE is able to minimize the system overhead while submitting
+% tasks.

### 2017-06-26 PandaEvolution.tex
Negative space added in PandaEvolution
Fix spaces
List of comments to address in each section

@@
+Comments to address in this section:
+\begin{enumerate}
+	\item REVIEWER 2: The authors spent quite a few paragraphs for the detail
+	design and implementation of the PanDA WMS, but introduced few background
+	of the LHC applications. According to the paper, I understand the
+	applications running on Titan are a large amount of small independent
+	jobs. Thus, it seems less cost-effective to run these jobs on a
+	world-leading supercomputer like Titan which has a fast and expensive
+	interconnection. It will be more interesting if the authors can compare
+	the solution on Titan (proposed in this paper) with the original
+	distributed computing environment and HPECloud from Google mentioned in
+	the related work section.
+	\item REVIEWER 5: In Section 3.2, GPUs are not considered in this work,
+	which would be a waste for using HPC computing. Could this work be
+	extended to include GPUs or other heterogenous resources?
+\end{enumerate}
+
 The upcoming LHC Run 3 will require more resources than the Worldwide LHC
-Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than 300,000
-cores at over 100 Grid sites, with a peak performance of 0.3 petaFLOPS. This
-capacity will be sufficient for the planned analysis and data processing, but it
-will be insufficient for the Monte Carlo production workflow and any extra
-activity. To alleviate these challenges, ATLAS is expanding the current
-computing model to include additional resources such as the opportunistic use of
-supercomputers.
+Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
+300,000 cores at over 100 Grid sites, with a peak performance of 0.3
+petaFLOPS. This capacity will be sufficient for the planned analysis and data
+processing, but it will be insufficient for the Monte Carlo production
+workflow and any extra activity. To alleviate these challenges, ATLAS is
+expanding the current computing model to include additional resources such as
+the opportunistic use of supercomputers.
 
@@
 requires runtime communication. Jobs are executed across multiple cores, each
-core calculating a small part of the problem and communicating with other cores
-via MPI. Accordingly, supercomputers have large number of worker nodes,
-connected through a high-speed, low-latency dedicated network. Each worker node
-has multicore CPUs, usually augmented with Graphics Processing Units (GPUs) or
-other specialized coprocessors.
+core calculating a small part of the problem and communicating with other
+cores via MPI. Accordingly, supercomputers have large number of worker nodes,
+connected through a high-speed, low-latency dedicated network. Each worker
+node has multicore CPUs, usually augmented with Graphics Processing Units
+(GPUs) or other specialized coprocessors.
 
@@
 ATLAS workloads or workflows involves concurrent and/or sequential runs of
-possibly large amount of jobs, each requiring no or minimal parallelization and
-no runtime communication. Thus, computing infrastructure like WLCG have been
-designed to aggregate large amount of computing resources across multiple sites.
-While each site may deploy MPI capabilities, usually these are not used to
-perform distributed computations.
+possibly large amount of jobs, each requiring no or minimal parallelization
+and no runtime communication. Thus, computing infrastructure like WLCG have
+been designed to aggregate large amount of computing resources across
+multiple sites. While each site may deploy MPI capabilities, usually these
+are not used to perform distributed computations.
 
 We developed and deployed a single-point solution to better understand the
-problem space of enabling a WMS designed for HTC to execute production workflows
-on resources designed to support HPC. The PanDA team developed a job broker to
-support the execution of part of the ATLAS production Monte Carlo workflow on
-Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
-Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).
+problem space of enabling a WMS designed for HTC to execute production
+workflows on resources designed to support HPC. The PanDA team developed a
+job broker to support the execution of part of the ATLAS production Monte
+Carlo workflow on Titan, a leadership-class supercomputer managed by the Oak
+Ridge Leadership Computing Facility (OLCF) at the Oak Ridge National
+Laboratory (ORNL).
 
@@
 
-The Titan supercomputer, current number three on the Top 500 list~\cite{top500},
-is a Cray XK7 system with 18,688 worker nodes and a total of 299,008 CPU cores.
-Each worker node has an AMD Opteron  6274 16-core CPU, a Nvidia Tesla K20X GPU,
-32 GB of RAM and no local storage, though a 16 GB RAM disk can be set up. Work
-nodes use Cray’s Gemini interconnect for inter-node MPI messaging. Titan is
-served by the Spider II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of
-disk storage, and by a 29 PB HPSS tape storage system. Titan’s worker nodes run
-Compute Node Linux, a run time environment based on SUSE Linux Enterprise
-Server.
+The Titan supercomputer, current number three on the Top 500
+list~\cite{top500}, is a Cray XK7 system with 18,688 worker nodes and a total
+of 299,008 CPU cores. Each worker node has an AMD Opteron  6274 16-core CPU,
+a Nvidia Tesla K20X GPU, 32 GB of RAM and no local storage, though a 16 GB
+RAM disk can be set up. Work nodes use Cray’s Gemini interconnect for
+inter-node MPI messaging. Titan is served by the Spider
+II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of disk storage, and
+by a 29 PB HPSS tape storage system. Titan’s worker nodes run Compute Node
+Linux, a run time environment based on SUSE Linux Enterprise Server.
 
-Titan's users submit jobs to Titan's PBS scheduler by logging into login or data
-transfer nodes (DTNs). Titan's authentication and authorization model is based
-on two-factor authentication with a RSA SecurID key. Login nodes and DTNs have
-out/inbound wide area network connectivity while worker nodes have only local
-network access. Fair-share and allocation policies are in place both for the PBS
-batch system and shared file systems.
+Titan's users submit jobs to Titan's PBS scheduler by logging into login or
+data transfer nodes (DTNs). Titan's authentication and authorization model is
+based on two-factor authentication with a RSA SecurID key. Login nodes and
+DTNs have out/inbound wide area network connectivity while worker nodes have
+only local network access. Fair-share and allocation policies are in place
+both for the PBS batch system and shared file systems.
 
-Titan's architecture, configuration and policies poses several challenges to the
-integration with PanDA. The default deployment
-model of PanDA Pilot is unfeasible on Titan: PanDA Pilot is required to contact
-the Job Dispatcher of the PanDA Server to pull jobs to execute, but this is not
-possible on Titan because worker nodes do not offer outbound network
-connectivity. Further, Titan does not support PanDA's security model based on
-certificates and virtual organizations, making PanDA's approach to identity
-management also unfeasible. While DTNs offer wide area network data transfer, an
-integration with ATLAS DDM is beyond the functional and administrative scope of
-the current prototyping phase. Finally, the specific characteristics of the
+Titan's architecture, configuration and policies poses several challenges to
+the integration with PanDA. The default deployment model of PanDA Pilot is
+unfeasible on Titan: PanDA Pilot is required to contact the Job Dispatcher of
+the PanDA Server to pull jobs to execute, but this is not possible on Titan
+because worker nodes do not offer outbound network connectivity. Further,
+Titan does not support PanDA's security model based on certificates and
+virtual organizations, making PanDA's approach to identity management also
+unfeasible. While DTNs offer wide area network data transfer, an integration
+with ATLAS DDM is beyond the functional and administrative scope of the
+current prototyping phase. Finally, the specific characteristics of the
 execution environment, especially the absence of local storage on the worker
@@
 Currently, very few HEP applications can benefit from Titan's GPUs but some
-computationally-intensive and non memory-intensive tasks of ATLAS workflows can
-be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
-partitioned into independent jobs, Titan worker nodes can be used to execute up
-to 16 concurrent payloads, one per each available core. Given these constraints
-and challenges, the type of task most suitable for execution at the moment on
-Titan is Monte Carlo detector simulation. This type of task is mostly
-computational-intensive, requiring less than 2GB of RAM at runtime and with
-small input data requirements. Detector simulation tasks in ATLAS are performed
-via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
-GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
-$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
-offloading.
+computationally-intensive and non memory-intensive tasks of ATLAS workflows
+can be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
+partitioned into independent jobs, Titan worker nodes can be used to execute
+up to 16 concurrent payloads, one per each available core. Given these
+constraints and challenges, the type of task most suitable for execution at
+the moment on Titan is Monte Carlo detector simulation. This type of task is
+mostly computational-intensive, requiring less than 2GB of RAM at runtime and
+with small input data requirements. Detector simulation tasks in ATLAS are
+performed via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework
+integrating the GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These
+tasks account for $\approx$ 60\% of all the jobs on WLCG, making them a
+primary candidate for offloading.
 
 Detector simulation is part of the ATLAS production Monte Carlo (MC)
-workflow~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
-consists of four main stages: event generation, detector simulation,
+workflow~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC
+workflow consists of four main stages: event generation, detector simulation,
 digitization, and reconstruction. Event generation creates sets of particle
-four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
-HERWIG~\cite{corcella2001herwig} and many others. The detector simulator is
-called Geant4~\cite{agostinelli2003geant4} and simulates the ATLAS detector and
-the interaction between that detector and particles. Each interaction creates a
-so-called hit and all hits are collected and passed on for digitalization, where
-hits are further processed to mimic the readout of the detector. Finally,
-reconstruction operates local pattern recognition, creating high-level objects
-like particles and jets.
+four-momenta via different generators, e.g.,
+PYTHIA~\cite{sjostrand2006pythia}, HERWIG~\cite{corcella2001herwig} and many
+others. The detector simulator is called Geant4~\cite{agostinelli2003geant4}
+and simulates the ATLAS detector and the interaction between that detector
+and particles. Each interaction creates a so-called hit and all hits are
+collected and passed on for digitalization, where hits are further processed
+to mimic the readout of the detector. Finally, reconstruction operates local
+pattern recognition, creating high-level objects like particles and jets.
 
@@
 
-The lack of wide area network connectivity on Titan's worker nodes is the most
-relevant challenge for integrating PanDA WMS and Titan. Without connectivity,
-Panda Pilots cannot be scheduled on worker nodes because they would not be able
-to communicate with PanDA Server and therefore pull and execute jobs. This makes
-impossible to port PanDA Pilot to Titan while maintaining the defining feature
-of the pilot abstraction: decoupling resource acquisition from workload
-execution via multi-stage scheduling.
+The lack of wide area network connectivity on Titan's worker nodes is the
+most relevant challenge for integrating PanDA WMS and Titan. Without
+connectivity, Panda Pilots cannot be scheduled on worker nodes because they
+would not be able to communicate with PanDA Server and therefore pull and
+execute jobs. This makes impossible to port PanDA Pilot to Titan while
+maintaining the defining feature of the pilot abstraction: decoupling
+resource acquisition from workload execution via multi-stage scheduling.
 
-The unavailability of pilots is a potential drawback when executing distributed
-workloads like MC detector simulation. Pilots are used to increase the
-throughput of distributed workloads: while pilots have to wait in the
-supercomputer's queue, once scheduled, they can pull and execute jobs
+The unavailability of pilots is a potential drawback when executing
+distributed workloads like MC detector simulation. Pilots are used to
+increase the throughput of distributed workloads: while pilots have to wait
+in the supercomputer's queue, once scheduled, they can pull and execute jobs
 independent from the system's queue. Jobs can be concurrently executed on
@@
 particularly relevant for machines like Titan where queue policies privilege
-parallel jobs on the base of the number of worker nodes they request: the higher
-the number of nodes, the shorter the amount of queue time (modulo fair-share and
-allocation policies).
+parallel jobs on the base of the number of worker nodes they request: the
+higher the number of nodes, the shorter the amount of queue time (modulo
+fair-share and allocation policies).
 
-The backfill optimization of Titan's Moab scheduler allows to avoid the overhead
-of queue wait times without using pilot abstraction~\cite{maui_backfill_url}.
-With this optimization, Moab starts low-priority jobs when they do not delay
-higher priority jobs, independent of whether the low-priority jobs were queued
-after the high-priority jobs.
+The backfill optimization of Titan's Moab scheduler allows to avoid the
+overhead of queue wait times without using pilot
+abstraction~\cite{maui_backfill_url}. With this optimization, Moab starts
+low-priority jobs when they do not delay higher priority jobs, independent of
+whether the low-priority jobs were queued after the high-priority jobs.
 
-When the backfill optimization is enabled, users can interrogate Moab about the
-number of worker nodes and walltime that would be available to a low-priority
-job at that moment in time. If a job is immediately submitted to Titan with that
-number of worker nodes and walltime, chances are that Moab will immediately
-schedule it, reducing its queue time to a minimum. In this paper, we call this
-number of worker nodes and walltime an available `backfill slot'.
+When the backfill optimization is enabled, users can interrogate Moab about
+the number of worker nodes and walltime that would be available to a
+low-priority job at that moment in time. If a job is immediately submitted to
+Titan with that number of worker nodes and walltime, chances are that Moab
+will immediately schedule it, reducing its queue time to a minimum. In this
+paper, we call this number of worker nodes and walltime an available
+`backfill slot'.
 
 Compared to pilots, backfill has the disadvantage of limiting the amount of
-worker nodes that can be requested. Pilots are normal jobs: they can request as
-many worker nodes and walltime as a queue can offer. On the contrary, jobs sized
-according to an available backfill slot depend on the number of worker nodes and
-walltime that cannot be given to any other job at that moment in time.
+worker nodes that can be requested. Pilots are normal jobs: they can request
+as many worker nodes and walltime as a queue can offer. On the contrary, jobs
+sized according to an available backfill slot depend on the number of worker
+nodes and walltime that cannot be given to any other job at that moment in
+time.
 
 At any point in time, the size of an available backfill slot is typically a
-small fraction of the total capacity of a resource. Notwithstanding, given the
-size of Titan this translates into a substantial capacity. Every year, about
-10\% of Titan's capacity remains unused~\cite{barker2016us}, corresponding to an
-average of 30,000 unused cores (excluding GPU cores). This equals to roughly
-10\% of the overall capacity of WLCG.
+small fraction of the total capacity of a resource. Notwithstanding, given
+the size of Titan this translates into a substantial capacity. Every year,
+about 10\% of Titan's capacity remains unused~\cite{barker2016us},
+corresponding to an average of 30,000 unused cores (excluding GPU cores).
+This equals to roughly 10\% of the overall capacity of WLCG.
 
-Given the communication requirements of PanDA Pilots and the unused capacity of
-Titan, PanDA pilot was repurposed to serve as a job broker on the DTN nodes of
-Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of PanDA Pilot
-and its stand-alone architecture, this prototype called `PanDA Broker'
-implements functionalities to: (i) interrogate Titan about backfill
-availability; (ii) pull ATLAS jobs and events from PanDA Server; (iii) wrap the
-payload of ATLAS jobs into MPI scripts; (iv) submitting MPI scripts to Titan's
-PBS batch system and monitor their execution; and (v) staging and preparing
-input/output files. Backfill querying, payload wrapping, and scripts submission
-required a new implementation while pulling ATLAS job and events, and file
-staging were inherited from PanDA Pilot.
+Given the communication requirements of PanDA Pilots and the unused capacity
+of Titan, PanDA pilot was repurposed to serve as a job broker on the DTN
+nodes of Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of
+PanDA Pilot and its stand-alone architecture, this prototype called `PanDA
+Broker' implements functionalities to: (i) interrogate Titan about backfill
+availability; (ii) pull ATLAS jobs and events from PanDA Server; (iii) wrap
+the payload of ATLAS jobs into MPI scripts; (iv) submitting MPI scripts to
+Titan's PBS batch system and monitor their execution; and (v) staging and
+preparing input/output files. Backfill querying, payload wrapping, and
+scripts submission required a new implementation while pulling ATLAS job and
+events, and file staging were inherited from PanDA Pilot.
 
@@
 MPI script enables the execution of unmodified Grid-centric, ATLAS jobs on
-Titan. Typically, a MPI script is workload-specific as it sets up the execution
-environment for a specific payload. This involves organization of worker
-directories, data management, optional input parameters modification, and
-cleanup on exit. Upon submission, a copy of the MPI script runs on every
+Titan. Typically, a MPI script is workload-specific as it sets up the
+execution environment for a specific payload. This involves organization of
+worker directories, data management, optional input parameters modification,
+and cleanup on exit. Upon submission, a copy of the MPI script runs on every
 available worker node, starting the execution of the ATLAS job's payload in a
@@
 Applications (SAGA) offers a unified interface to diverse job schedulers and
-file transferring services. In this way, SAGA provides an interoperability layer
-that lowers the complexity of using distributed infrastructures. Behind the API
-façade, RADICAL-SAGA implements a adaptor architecture: each adaptor interface
-the SAGA API with different middleware systems and services, including the PBS
-batch scheduler of Titan.
+file transferring services. In this way, SAGA provides an interoperability
+layer that lowers the complexity of using distributed infrastructures. Behind
+the API façade, RADICAL-SAGA implements a adaptor architecture: each adaptor
+interface the SAGA API with different middleware systems and services,
+including the PBS batch scheduler of Titan.
 
@@
 system that is shared among DTNs and worker nodes. The input files with the
-events of the ATLAS jobs are downloaded on the shared filesystem from the data
-center of Brookhaven National Laboratory (BNL). The MPI script setup process
-includes making the location of these files available to the payload of the
-ATLAS's jobs. The PanDA Broker can locate the payload's output files on the
-shared filesystem and transfer them from Titan BNL.
+events of the ATLAS jobs are downloaded on the shared filesystem from the
+data center of Brookhaven National Laboratory (BNL). The MPI script setup
+process includes making the location of these files available to the payload
+of the ATLAS's jobs. The PanDA Broker can locate the payload's output files
+on the shared filesystem and transfer them from Titan BNL.
 
-Once deployed on Titan, every PanDA Broker supports the execution of MC detector
-simulations in 9 steps. PanDA Broker queries the Job Dispatcher module of the
-PanDA server for ATLAS jobs that have been bound to Titan by JEDI
-(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA Broker
-pulls jobs' input files from BNL to the OLCF Lustre file system
+Once deployed on Titan, every PanDA Broker supports the execution of MC
+detector simulations in 9 steps. PanDA Broker queries the Job Dispatcher
+module of the PanDA server for ATLAS jobs that have been bound to Titan by
+JEDI (Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA
+Broker pulls jobs' input files from BNL to the OLCF Lustre file system
 (Fig.~\ref{fig:panda_broker}:2). PanDA Broker queries Titan's Moab scheduler
 about the current available backfill slot (Fig.~\ref{fig:panda_broker}:3) and
-creates an MPI script, wrapping enough ATLAS jobs' payload to fit the backfill
-slot. PanDA Broker submits the MPI script to the Titan's Torque batch system via
-RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).
+creates an MPI script, wrapping enough ATLAS jobs' payload to fit the
+backfill slot. PanDA Broker submits the MPI script to the Titan's Torque
+batch system via RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).
 
@@
 (Fig.~\ref{fig:panda_broker}:6), and executes one AthenaMP for each available
-work node (Fig.~\ref{fig:panda_broker}:7). AthenaMP retrieves events from Lustre
-(Fig.~\ref{fig:panda_broker}:8) and spawns 1 Geant4 event simulation process on
-each of the 16 available cores (Fig.~\ref{fig:panda_broker}:9). Upon completion
-of each MPI script, PanDA Broker transfer the jobs' output to BNL
-(Fig.~\ref{fig:panda_broker}:10), and performs cleanup.
+work node (Fig.~\ref{fig:panda_broker}:7). AthenaMP retrieves events from
+Lustre (Fig.~\ref{fig:panda_broker}:8) and spawns 1 Geant4 event simulation
+process on each of the 16 available cores (Fig.~\ref{fig:panda_broker}:9).
+Upon completion of each MPI script, PanDA Broker transfer the jobs' output to
+BNL (Fig.~\ref{fig:panda_broker}:10), and performs cleanup.
 
@@
     \vspace{-0.3in}
-    \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates
-    the execution process of a detector simulation job, part of the production
-    ATLAS MC workflow.}
+    \caption{PanDA Broker architecture as deployed on Titan. Numbers
+    indicates the execution process of a detector simulation job, part of the
+    production ATLAS MC workflow.}
 \label{fig:panda_broker}
 \end{figure}
+\vspace{-0.04in}

### 2017-06-26 conclusion.tex
List of comments to address in each section

@@
-% PanDA WMS was developed to meet the scale and complexity of LHC
-% distributed computing for the ATLAS experiment. In the process, the old batch
-% job paradigm of computing in HEP was discarded in favor of a far more
-% flexible and scalable model. The success of PanDA at the LHC is leading to
-% widespread adoption and testing by other experiments. PanDA is the first
-% exascale workload management system in HEP, already operating at a million
-% computing jobs per day, and processing over an exabyte of data in 2013. Next
-% LHC run will pose massive computing challenges. With a doubling of the beam
-% energy and luminosity as well as an increased need for simulates data, the
-% data volume is expected to increase with a factor 5--6 or more. Storing and
-% processing this amount of data is a challenge that cannot be resolved with the
-% currently existing computing resources in ATLAS\@. To resolve this challenge,
-% ATLAS is turning to commercial as well as academic Cloud services and HPCs via
-% the PanDA system. Also the work underway is enabling the use of PanDA by new
-% scientific collaborations and communities as a means of leveraging extreme
-% scale computing resources with a low barrier of entry. The technology base
-% provided by the PanDA system will enhance the usage of a variety of
-% high-performance computing resources available to basic research.
+Comments to address in this section:
+\begin{enumerate}
+	\item REVIEWER 3: How extensible is this? Do you have a sense of backfill
+	availability on other systems? Is this unique to Titan or Cray or more
+	ubiquitous of an option?
+\end{enumerate}
 
@@
 leadership-class HPC machine at unprecedented scale. In the past 13 months,
-PanDA WMS has consumed almost 52M core-hours on Titan, simulating 3.5\% of the
-total number of detector events of the ATLAS production Monte Carlo workflow. We
-described the implementation and execution process of PanDA WMS
-(\S\ref{sec:panda_overview}) and PanDA Broker (\S\ref{sec:panda_titan}), showing
-how they support and enable distributed computing at this scale on Titan, a
-leadership-class HPC machine managed by OCLF at ORNL.
+PanDA WMS has consumed almost 52M core-hours on Titan, simulating 3.5\% of
+the total number of detector events of the ATLAS production Monte Carlo
+workflow. We described the implementation and execution process of PanDA WMS
+(\S\ref{sec:panda_overview}) and PanDA Broker (\S\ref{sec:panda_titan}),
+showing how they support and enable distributed computing at this scale on
+Titan, a leadership-class HPC machine managed by OCLF at ORNL.
 
-We characterized the experience by evaluating the efficiency, scalability
-and reliability of both PanDA Broker and AthenaMP as deployed on Titan
-(\S\ref{sec:panda_titan}). Our characterization highlighted the strengths
-and limitations of the current design and implementation: PanDA Brokers enable
-the sustained execution of millions of simulations per week but further work is
-required to optimize its efficiency and reliability (\S\ref{ssec:broker_titan}).
-PanDA Brokers support the concurrent execution of multiple AthenaMP instances,
-enabling each AthenaMP to perform the concurrent execution of up to 16 Geant4
-simulators. Nonetheless, our characterization showed how improving I/O
-performance could reduce overheads (\S\ref{ssec:athenamp_titan}), increasing the
-overall utilization of Titan's backfill availability.
+We characterized the experience by evaluating the efficiency, scalability and
+reliability of both PanDA Broker and AthenaMP as deployed on Titan
+(\S\ref{sec:panda_titan}). Our characterization highlighted the strengths and
+limitations of the current design and implementation: PanDA Brokers enable
+the sustained execution of millions of simulations per week but further work
+is required to optimize its efficiency and reliability
+(\S\ref{ssec:broker_titan}). PanDA Brokers support the concurrent execution
+of multiple AthenaMP instances, enabling each AthenaMP to perform the
+concurrent execution of up to 16 Geant4 simulators. Nonetheless, our
+characterization showed how improving I/O performance could reduce overheads
+(\S\ref{ssec:athenamp_titan}), increasing the overall utilization of Titan's
+backfill availability.
 
@@
 characteristics, resources types and federation constraints, as well metrics
-of performance. The experience captured in this paper will prove
-invaluable for designing WMS for computational campaigns and will provide a
-baseline to evaluate the relative merits of different approaches.
+of performance. The experience captured in this paper will prove invaluable
+for designing WMS for computational campaigns and will provide a baseline to
+evaluate the relative merits of different approaches.
 
@@
 (e.g., NSF's flagship Blue Waters the average utilization fluctuates between
-60-80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
-approaches elsewhere. These unprecedented efficiency gains aside, this work is
-just a starting point towards more effective operational models for future
+60--80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
+approaches elsewhere. These unprecedented efficiency gains aside, this work
+is just a starting point towards more effective operational models for future
 leadership and online analytical platforms~\cite{foap-url}. These platforms
-will have to support ever increasing complex workloads with varying models for
-dynamic resource federation.
+will have to support ever increasing complex workloads with varying models
+for dynamic resource federation.
 
-% Recent efforts with Google Cloud for the CMS
-% experiment -- HEPCloud~\cite{hepcloud,googlehep} represent similar yet
-% different approaches to dynamic resource federation. As the type,
-% heterogeneity and complexity of future platforms  (e.g., Exascale, Commercial
-% Clouds) for science increases, there will be a greater need and emphasis on
+% Recent efforts with Google Cloud for the CMS experiment --
+% HEPCloud~\cite{hepcloud,googlehep} represent similar yet different
+% approaches to dynamic resource federation. As the type, heterogeneity and
+% complexity of future platforms  (e.g., Exascale, Commercial Clouds) for
+% science increases, there will be a greater need and emphasis on
 % abstractions to ensure WMS can both utilize as well as support evolving
 % platforms. Our experience demonstrated the advantages arising from an
-% abstractions based approach \textendash{} in particular scalable implementations
-% of pilot abstraction.
+% abstractions based approach \textendash{} in particular scalable
+% implementations of pilot abstraction.

### 2017-06-26 introduction.tex
List of comments to address in each section

@@
+Comments to address in this section:
+\begin{enumerate}
+	\item REVIEWER 1: listing the key observations in one place, and in a
+	consistent fashion. I found The 5 points listed in the first part of the
+	rebuttal quite compelling, and strongly urge the authors to put them in
+	the paper in a prominent location.
+	\item REVIEWER 1: A glossary would be welcomed.
+	\item REVIEWER 5: I would have preferred to see a more detailed list of
+	contributions, including the novelty of the proposed approach as compared
+	to prior work.
+	\item I am ambivalent about the use of the term ``converging'' in the
+	paper's title. While the paper does port workloads to high-performance
+	computing resources, it does not deal with compute-oriented optimizations
+	(parallelization), lowering communication, etc., which are often major
+	concerns in HPC programming. In this sense, the paper is closer in spirit
+	to how to port high-throughput apps to high-performance clusters. Agreed?
+\end{enumerate}
+
 The Large Hadron Collider (LHC) was created to explore the fundamental
-properties of matter. Multiple experiments at LHC have
-collected and distributed hundreds of petabytes of data worldwide to hundreds of
-computer centers. Thousands of physicists analyze petascale data volumes daily.
-The detection of the Higgs Boson in 2013 speaks to the success of the detector
+properties of matter. Multiple experiments at LHC have collected and
+distributed hundreds of petabytes of data worldwide to hundreds of computer
+centers. Thousands of physicists analyze petascale data volumes daily. The
+detection of the Higgs Boson in 2013 speaks to the success of the detector
 and experiment design, as well as the sophistication of computing systems
@@
 federation of hundreds to thousands of distributed resources, %\textemdash{}
-ranging in scale from small to mid-size resource~\cite{foster2003grid}. Although
-the workloads to be executed are comprised of tasks that are independent of each
-other, the management of the distribution of workloads across many heterogeneous
-resources, the effective utilization of resources and efficient
-execution of workloads presents non-trivial challenges.
+ranging in scale from small to mid-size resource~\cite{foster2003grid}.
+Although the workloads to be executed are comprised of tasks that are
+independent of each other, the management of the distribution of workloads
+across many heterogeneous resources, the effective utilization of resources
+and efficient execution of workloads presents non-trivial challenges.
 
@@
 distributed data processing and analysis. The CMS and ATLAS experiments
-represent arguably the largest production grade distributed computing solutions
-and have symbolized the paradigm of {\it high-throughput computing}, i.e., the
-effective execution of many independent tasks.
+represent arguably the largest production grade distributed computing
+solutions and have symbolized the paradigm of {\it high-throughput
+computing}, i.e., the effective execution of many independent tasks.
 
 As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity era
-(Run 4), it is anticipated that the data volumes that will need analyzing will
-increase by factors of 10-100 compared to the current phase (Run 2). Data will
-be larger in volume but will also require more sophisticated computational
-processing. In spite of the impressive scale of the ATLAS distributed
-computing system, demand for computing systems will significantly outstrip
-current and projected supply.  There are multiple levels at which this problem
-needs to be addressed: the utilization of emerging parallel architectures
-(e.g., platforms); algorithmic and advances in analytical methods (e.g., use
-of Machine Learning); and the ability to exploit different platforms (e.g.,
-clouds and supercomputers).
+(Run 4), it is anticipated that the data volumes that will need analyzing
+will increase by factors of 10-100 compared to the current phase (Run 2).
+Data will be larger in volume but will also require more sophisticated
+computational processing. In spite of the impressive scale of the ATLAS
+distributed computing system, demand for computing systems will significantly
+outstrip current and projected supply.  There are multiple levels at which
+this problem needs to be addressed: the utilization of emerging parallel
+architectures (e.g., platforms); algorithmic and advances in analytical
+methods (e.g., use of Machine Learning); and the ability to exploit different
+platforms (e.g., clouds and supercomputers).
 
@@
 runtime system on Titan, which allows advanced execution modes of the ATLAS
-workload as well as the enhanced support for heterogeneous workloads, 
-e.g., molecular dynamics.
+workload as well as the enhanced support for heterogeneous workloads, e.g.,
+molecular dynamics.
 
-% . Consequently, Titan now analyzes up to 3.6\% of the simulation workload of
-% the ATLAS experiment. The case study presents the design and integration of
-% PanDA
+% . Consequently, Titan now analyzes up to 3.6\% of the simulation workload
+% of the ATLAS experiment. The case study presents the design and integration
+% of PanDA
 
@@
 taken to support the sustained, scalable and production usage of Titan for
-historically high-throughput workloads; (ii) a preliminary characterization of a
-next generation executor (NGE) for PanDA to support non-traditional
+historically high-throughput workloads; (ii) a preliminary characterization
+of a next generation executor (NGE) for PanDA to support non-traditional
 heterogeneous workloads and execution modes;  and (iii) early lessons and

### 2017-06-26 nge.tex
List of comments to address in each section

@@
+Comments to address in this section:
+\begin{enumerate}
+	\item REVIEWER 2: In the weak scalability evaluations (in Figure 7 and
+	8), the overhead of the Next Generation Executor (NGE) has grown linearly
+	with the number of the nodes. Can you explain why the average time
+	execution time of AthenaMP is decreased even when the overhead is
+	increasing?
+\end{enumerate}
+
 As seen in \S\ref{ssec:panda_titan}, PanDA Broker was  deployed on the DTNs

### 2017-06-26 overall.tex
List of comments to address in each section

@@
-Currently, ATLAS deploys 20 instances of the PanDA Broker on 4 Titan's DTNs, 5
-instances per DTN. Each broker submits and manages the execution of 15 to 300
-ATLAS jobs, one job for each Titan's worker node, and a theoretical maximum
-concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have operated
-only in backfill mode, without a defined time allocation, and running at the
-lowest priority on Titan.
+Comments to address in this section:
+\begin{enumerate}
+    \item REVIEWER 2: In Figure 4, there are some inconsistencies. The figure
+    caption shows there are only 62175 measures, but in the scatter plot
+    (right) there are 62555 data points. In addition, according to the
+    distribution of the number of work nodes (left), it does not look like
+    the mean number of work available is 691 as the figure caption
+    illustrates.
+    \item REVIEWER 3: t’s interesting that Feb/March have such an increase in
+    backfill availability in 2017. The authors attribute this to hardware
+    upgrades for the DTNs. It’s mentioned that the load is only 3.6\% now and
+    therefore further upgrades will likely not cause similar improvement.
+    What was the previous CPU load? Why were particularly low percentages
+    utilized in April/May 2016?
+    \item REVIEWER 4: “Genat4” at the end of Section 4 is a typo.
+    \item REVIEWER 4: Please explain why you used Titan at all, in spite of
+    your statement that it is 1/2 to 1/3 as efficient as other resources. Was
+    it merely that you were given so much time on this resource that the
+    inefficiency doesn’t matter?
+    \item REVIEWER 4: Why didn’t you benchmark the performance of your
+    workload on other HPC resources that might have delivered better
+    performance, prior to reengineering your entire workflow to deal with the
+    challenges posed by Titan, a machine that doesn’t seem to justify the
+    pain?
+    \item REVIEWER 4: The INCITE program operational requirements for the
+    LCFs are 90\% uptime. Using large batches of Geant4 jobs that run on a
+    single node or less to reach this metric sounds like gaming the system
+    rather than an accomplishment.
+    \item REVIEWER 5: In Figure 3, it shows the unused core hours are much
+    higher than those used by ATLAS. What are the reasons for such low usage?
+\end{enumerate}
+
+Currently, ATLAS deploys 20 instances of the PanDA Broker on 4 Titan's DTNs,
+5 instances per DTN\@. Each broker submits and manages the execution of 15 to
+300 ATLAS jobs, one job for each Titan's worker node, and a theoretical
+maximum concurrent use of 96,000 cores. Since November 2015, PanDA Brokers
+have operated only in backfill mode, without a defined time allocation, and
+running at the lowest priority on Titan.
 
@@
 PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and
-AthenaMP. We discuss challenges and limitations of our approach at multiple
-levels arising from the specifics of workload, middleware and methods. All the
-measurements were performed between January 2016 and February 2017, hereafter
-called ‘experiment time window
+AthenaMP\@. We discuss challenges and limitations of our approach at multiple
+levels arising from the specifics of workload, middleware and methods. All
+the measurements were performed between January 2016 and February 2017,
+hereafter called ‘experiment time window
 
-% -----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
 \subsection{Characterizing the PanDA Broker on Titan}
@@
 
-We calculate the total amount of backfill availability over a period of time by:
-(i) polling the available backfill slots at regular intervals during that time
-window; (ii) converting the number of worker nodes available and their walltime
-into core-hours; (iii) summing the number of core-hours. We call this number of
-core-hours `total backfill availability'.
+We calculate the total amount of backfill availability over a period of time
+by: (i) polling the available backfill slots at regular intervals during that
+time window; (ii) converting the number of worker nodes available and their
+walltime into core-hours; (iii) summing the number of core-hours. We call
+this number of core-hours `total backfill availability'.
 
@@
 Titan (gray bars) and the number of core-hours of that availability used by
-ATLAS (blue bars) during the experiment time window. ATLAS consumed a total of
-51.4M core-hours, for an average of $\approx$3.7M core-hours a month, with a
-minimum of 1.7M core-hours in April 2016 and a maximum 7.9M core-hours in
+ATLAS (blue bars) during the experiment time window. ATLAS consumed a total
+of 51.4M core-hours, for an average of $\approx$3.7M core-hours a month, with
+a minimum of 1.7M core-hours in April 2016 and a maximum 7.9M core-hours in
 February 2017.
@@
     \vspace{-0.3in}
-    \caption{Titan's total backfill availability: CPU core-hours (gray) and CPU
-    core-hours used by ATLAS (blue). GPU core-hours unaccounted for as they
-    cannot be used by ATLAS. Efficiency of PanDA Brokers defined as percentage
-    of total Titan's backfill availability used by ATLAS (Red labels).}
+    \caption{Titan's total backfill availability: CPU core-hours (gray) and
+    CPU core-hours used by ATLAS (blue). GPU core-hours unaccounted for as
+    they cannot be used by ATLAS. Efficiency of PanDA Brokers defined as
+    percentage of total Titan's backfill availability used by ATLAS (Red
+    labels).}
 \label{fig:backfill-utilization}
@@
 
-PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels) is
-defined as the fraction of core-hours utilized by the PanDA Brokers to  Titan’s
-total backfill availability during the experiment time window. The brokers
-reached 18\% average efficiency, with a minimum 7.8\% efficiency on May 2016 and
-a maximum 30.9\% efficiency on February 2017 (excluding the preliminary results
-of March). The total backfill availability was 20.3M in April 2016, and 17.6M in
-February 2017. This shows that the measured difference in efficiency did not
-depend on a comparable difference in the total backfill availability.
+PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels)
+is defined as the fraction of core-hours utilized by the PanDA Brokers to
+Titan’s total backfill availability during the experiment time window. The
+brokers reached 18\% average efficiency, with a minimum 7.8\% efficiency on
+May 2016 and a maximum 30.9\% efficiency on February 2017 (excluding the
+preliminary results of March). The total backfill availability was 20.3M in
+April 2016, and 17.6M in February 2017. This shows that the measured
+difference in efficiency did not depend on a comparable difference in the
+total backfill availability.
 
 During the experiment time window, about 2.25M detector simulation jobs were
-completed on Titan, for a total of 225M events processed. This is equivalent to
-0.9\% of all the 250M detector simulations performed by ATLAS in the same period
-of time, and 3.5\% of the 6.6B events processed by those jobs. These figures
-confirms the relevance of supercomputers' resource contribution to the LHC Run
-2, especially when accounting for the amount of unused total backfill
-availability and the improvement of PanDA efficiency across the experiment time
-window.
+completed on Titan, for a total of 225M events processed. This is equivalent
+to 0.9\% of all the 250M detector simulations performed by ATLAS in the same
+period of time, and 3.5\% of the 6.6B events processed by those jobs. These
+figures confirms the relevance of supercomputers' resource contribution to
+the LHC Run 2, especially when accounting for the amount of unused total
+backfill availability and the improvement of PanDA efficiency across the
+experiment time window.
 
@@
 relevant code update was made during that period and logs indicated that the
-brokers were able to perform faster. This is likely due to hardware upgrades on
-the DTNs. The absence of continuous monitoring of those nodes does not allow to
-quantify bottlenecks but spot measurements of their load indicate that a faster
-CPU and better networking were likely responsible for the improved performance.
-Investigations showed an average CPU load of 3.6\% on the upgraded DTNs. As
-such, further hardware upgrades seem unlikely to improve significantly the
-performance of PanDA Brokers.
+brokers were able to perform faster. This is likely due to hardware upgrades
+on the DTNs. The absence of continuous monitoring of those nodes does not
+allow to quantify bottlenecks but spot measurements of their load indicate
+that a faster CPU and better networking were likely responsible for the
+improved performance. Investigations showed an average CPU load of 3.6\% on
+the upgraded DTNs. As such, further hardware upgrades seem unlikely to
+improve significantly the performance of PanDA Brokers.
 
-Every detector simulation executed on Titan process 100 events. This number of
-events is consistent with the physics of the use case and with the average
+Every detector simulation executed on Titan process 100 events. This number
+of events is consistent with the physics of the use case and with the average
 duration of backfill availability. The duration of a detector simulation is a
-function of the number of events simulated but not all events take the same time
-to be simulated. 1 event simulation takes from $\approx$2 to $\approx$40
-minutes, with a mean of $\approx$14 minutes. Considering that each worker node
-process up to 16 events concurrently, 100 events takes an average of 105 minutes
-to process. As such, PanDA brokers do not use backfill availability with less
-than 105 minutes walltime.
+function of the number of events simulated but not all events take the same
+time to be simulated. 1 event simulation takes from $\approx$2 to $\approx$40
+minutes, with a mean of $\approx$14 minutes. Considering that each worker
+node process up to 16 events concurrently, 100 events takes an average of 105
+minutes to process. As such, PanDA brokers do not use backfill availability
+with less than 105 minutes walltime.
 
-Fig.~\ref{fig:backfill-distrib} shows the distribution of backfill availability
-on Titan as a function of number of nodes and the time of their availability
-(i.e., walltime). We recorded these data by polling Titan's Moab scheduler at
-regular intervals during the experiment window time. The mean number of nodes
-was 691, and their mean walltime was 126 minutes. Detector simulations of 100
-events, enable to use down to 5/6 of the mean walltime of backfill availability.
-As such, it offers a good compromise for PanDA Broker efficiency.
+Fig.~\ref{fig:backfill-distrib} shows the distribution of backfill
+availability on Titan as a function of number of nodes and the time of their
+availability (i.e., walltime). We recorded these data by polling Titan's Moab
+scheduler at regular intervals during the experiment window time. The mean
+number of nodes was 691, and their mean walltime was 126 minutes. Detector
+simulations of 100 events, enable to use down to 5/6 of the mean walltime of
+backfill availability. As such, it offers a good compromise for PanDA Broker
+efficiency.
 
@@
    \vspace{-0.1in}
-    \caption{Backfill availability: distribution of number of work nodes (left)
-    and of walltime in minutes (center). Scatter plot of the two variables
-    (right). 62175 measures: mean number of work nodes available 691; mean
-    walltime available 126 minutes.}
+    \caption{Backfill availability: distribution of number of work nodes
+    (left) and of walltime in minutes (center). Scatter plot of the two
+    variables (right). 62175 measures: mean number of work nodes available
+    691; mean walltime available 126 minutes.}
 
@@
 % Usually, detector simulations performed on the Grid process 10 times the
-% number of events than those on Titan. This explains the difference between the
-% percentage of detector simulations and of events computed by those simulations
-% on Titan.
+% number of events than those on Titan. This explains the difference between
+% the percentage of detector simulations and of events computed by those
+% simulations on Titan.
 PanDA Broker could fit the number of events to the walltime of each available
-backfill slot on the base of the distributions of the time taken by one event to
-be simulated. That specific number of event could then be pulled from the PanDA
-Event service~\cite{calafiura2015atlas} and given as input to one or more
-simulations. Once packaged into the MPI script submitted to titan's PBS batch
-system, these simulations would better fit their available backfill slot,
-contributing to increase the efficiency of PanDA Brokers.
+backfill slot on the base of the distributions of the time taken by one event
+to be simulated. That specific number of event could then be pulled from the
+PanDA Event service~\cite{calafiura2015atlas} and given as input to one or
+more simulations. Once packaged into the MPI script submitted to titan's PBS
+batch system, these simulations would better fit their available backfill
+slot, contributing to increase the efficiency of PanDA Brokers.
 
 The transition from a homogeneous to a heterogeneous number of events per
-detector simulation has implications for the application layer. An even number
-of events across simulations makes it easier to partition, track and package
-events across simulations, especially when they are performed on both the Grid
-and Titan. A homogeneous number of events also helps to keep the size and
-duration of other stages of the MC workflow (\S\ref{ssec:panda-titan}) more
-uniform. Further analysis is needed to evaluate the trade offs between increased
-efficiency of resource utilization and the complexity that would be introduced
-at the application layer.
+detector simulation has implications for the application layer. An even
+number of events across simulations makes it easier to partition, track and
+package events across simulations, especially when they are performed on both
+the Grid and Titan. A homogeneous number of events also helps to keep the
+size and duration of other stages of the MC workflow
+(\S\ref{ssec:panda-titan}) more uniform. Further analysis is needed to
+evaluate the trade offs between increased efficiency of resource utilization
+and the complexity that would be introduced at the application layer.
 
@@
 process is spawn at a time to execute the payload. As a consequence, the
-utilization of a larger portion of Titan's total backfill availability depends
-on the the number of concurrent PanDA Brokers instantiated on the DTNs: When all
-the 20 PanDA Brokers have submitted a MPI PBS script, further backfill
-availability cannot be used.
+utilization of a larger portion of Titan's total backfill availability
+depends on the the number of concurrent PanDA Brokers instantiated on the
+DTNs: When all the 20 PanDA Brokers have submitted a MPI PBS script, further
+backfill availability cannot be used.
 
-In September 2016, increasing the number of concurrent PanDA brokers from 4 to
-20 markedly improved efficiency (see Fig.~\ref{fig:backfill-utilization}) but
-further research is undergoing to understand whether an even larger number of
-brokers would yield similar results. This research focuses on evaluating the
-overheads of input/output files staging, including its impact on DTNs, and on an
-alternative design of PanDA Broker that enables the concurrent submission of
-multiple MPI scripts~\cite{barreiro2016panda}.
+In September 2016, increasing the number of concurrent PanDA brokers from 4
+to 20 markedly improved efficiency (see Fig.~\ref{fig:backfill-utilization})
+but further research is undergoing to understand whether an even larger
+number of brokers would yield similar results. This research focuses on
+evaluating the overheads of input/output files staging, including its impact
+on DTNs, and on an alternative design of PanDA Broker that enables the
+concurrent submission of multiple MPI scripts~\cite{barreiro2016panda}.
 
@@
 the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
-while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
-Brokers were responsible for around the 19\% of the failures, compared to the
-29\% of failures produced by the JobDispatcher module of the PanDA Server, and
-the 13\% failures produced by the Geant4 toolkit. The current failure rate of
-the PanDA Brokers confirms the benefits of reusing most of the code base of the
-PanDA Pilot for implementing the PanDA Broker. It also shows that adopting
-third-party libraries like RADICAL-SAGA did not have a measurable adverse effect
-on reliability.
+while the failure rate of jobs submitted to Titan was a comparable 13.6\%.
+PanDA Brokers were responsible for around the 19\% of the failures, compared
+to the 29\% of failures produced by the JobDispatcher module of the PanDA
+Server, and the 13\% failures produced by the Geant4 toolkit. The current
+failure rate of the PanDA Brokers confirms the benefits of reusing most of
+the code base of the PanDA Pilot for implementing the PanDA Broker. It also
+shows that adopting third-party libraries like RADICAL-SAGA did not have a
+measurable adverse effect on reliability.
 
@@
 
-We use two main parameters to measure the performance of the detector simulation
-jobs submitted to Titan: (i) the time taken to setup AthenaMP; and (ii) the
-distribution of the time taken by the Geant4 toolkit to simulate a certain
-number of events.
+We use two main parameters to measure the performance of the detector
+simulation jobs submitted to Titan: (i) the time taken to setup AthenaMP; and
+(ii) the distribution of the time taken by the Geant4 toolkit to simulate a
+certain number of events.
 
-AthenaMP has an initialization and configuration stage. At initialization time,
-AthenaMP is assembled from a large number of shared libraries, depending on the
-type of payload that will have to be computed. Once initialized, every algorithm
-and service of AthenaMP is configured by a set of Python scripts. Both these
-operations result in a large number of read operations on the filesystem shared
-between the worker nodes and the DTNs, including the operations required to
-access small python scripts.
+AthenaMP has an initialization and configuration stage. At initialization
+time, AthenaMP is assembled from a large number of shared libraries,
+depending on the type of payload that will have to be computed. Once
+initialized, every algorithm and service of AthenaMP is configured by a set
+of Python scripts. Both these operations result in a large number of read
+operations on the filesystem shared between the worker nodes and the DTNs,
+including the operations required to access small python scripts.
 
-Initially, all the shared libraries of AthenaMP and the python scripts for its
-configuration were stored on the Spider 2 Lustre file system. However, the I/O
-patterns of the initialization and configuration stages degraded the performance
-of the filesystem. This was addressed by moving the AthenaMP distribution to a
-read-only NFS directory, shared among DTNs and worker nodes. NFS eliminated the
-problem of metadata contention, improving metadata read performance from
-$\approx$6,300 seconds on Lustre to $\approx$1,500 seconds on NFS.
+Initially, all the shared libraries of AthenaMP and the python scripts for
+its configuration were stored on the Spider 2 Lustre file system. However,
+the I/O patterns of the initialization and configuration stages degraded the
+performance of the filesystem. This was addressed by moving the AthenaMP
+distribution to a read-only NFS directory, shared among DTNs and worker
+nodes. NFS eliminated the problem of metadata contention, improving metadata
+read performance from $\approx$6,300 seconds on Lustre to $\approx$1,500
+seconds on NFS.
 
@@
 Geant4 simulators on a Titan's worker node. Geant4 requires to read events
-descriptions from a filesystem and simulate them as they would happen within the
-ATLAS detector. We characterized both the compute performance of the simulation
-and the impact of acquiring event descriptions on the filesystem.
+descriptions from a filesystem and simulate them as they would happen within
+the ATLAS detector. We characterized both the compute performance of the
+simulation and the impact of acquiring event descriptions on the filesystem.
 
 The AMD Opteron 6274 CPU used on Titan has 16 cores, divided into 8 compute
-units. Each compute units has 1 floating point (FP) scheduler shared between 2
-cores. When using 16 cores for FP-intensive calculations, each pair of cores
-competes for a single FP scheduler. This creates the overhead shown in
+units. Each compute units has 1 floating point (FP) scheduler shared between
+2 cores. When using 16 cores for FP-intensive calculations, each pair of
+cores competes for a single FP scheduler. This creates the overhead shown in
 Fig.~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
 concurrent simulations computing 50 events is 10.8 minutes, while for 16
-simulations is 14.25 minutes (consistent with the measured distribution of the
-duration of event simulation). Despite an inefficiency of almost 30\%, Titan's
-allocation policy based on number of worker nodes used instead of number of
-cores does not justify the use of 1/2 of the cores available.
+simulations is 14.25 minutes (consistent with the measured distribution of
+the duration of event simulation). Despite an inefficiency of almost 30\%,
+Titan's allocation policy based on number of worker nodes used instead of
+number of cores does not justify the use of 1/2 of the cores available.
 
@@
    \vspace{-0.3in}
-    \caption{Distributions of the time taken to simulate one event when placing
-    2 simulations (h1) or 1 simulation (h2) per Titan's CPU. 2 simulation use 16
-    cores per node, 1 simulation 8. 50 Events; 1 Titan worker nodes; 16 work
-    threads per node; 100 events per node.}
+    \caption{Distributions of the time taken to simulate one event when
+    placing 2 simulations (h1) or 1 simulation (h2) per Titan's CPU. 2
+    simulation use 16 cores per node, 1 simulation 8. 50 Events; 1 Titan
+    worker nodes; 16 work threads per node; 100 events per node.}
 
@@
 
-The performance analysis of Titan's AMD CPUs for detector simulations helps also
-to compare Titan and Grid site performances. Usually, Grid sites exposes
+The performance analysis of Titan's AMD CPUs for detector simulations helps
+also to compare Titan and Grid site performances. Usually, Grid sites exposes
 resources with heterogeneous CPU architectures and a maximum of 8 (virtual)
-cores per worker node, while Titan's offer an homogeneous 16 cores architecture.
-We used the rate of events processes per minute as a measure of the efficiency
-of executing the same detector simulation on Titan or Grid sites. Comparisons of
-the efficiencies of Titan to the BNL and SIGNET Grid sites, normalized for 8
-cores, show that the effective performance per-core at Titan is $\approx$0.57
-event per minute, roughly 1/2 of BNL and  1/3 of SIGNET performances.
+cores per worker node, while Titan's offer an homogeneous 16 cores
+architecture. We used the rate of events processes per minute as a measure of
+the efficiency of executing the same detector simulation on Titan or Grid
+sites. Comparisons of the efficiencies of Titan to the BNL and SIGNET Grid
+sites, normalized for 8 cores, show that the effective performance per-core
+at Titan is $\approx$0.57 event per minute, roughly 1/2 of BNL and  1/3 of
+SIGNET performances.
 
-The differences in performance between Titan and the BNL and SIGNET Grid sites
-are due to the FP scheduler competition and the availability of newer
-processors. The CPUs at the Grid sites have one FP scheduler per core and are on
-average newer than the CPU of Titan. The heterogeneity of the Grid sites' CPUs
-explain the higher performance variance we observed compared to the performance
-consistency we measured on Titan.
+The differences in performance between Titan and the BNL and SIGNET Grid
+sites are due to the FP scheduler competition and the availability of newer
+processors. The CPUs at the Grid sites have one FP scheduler per core and are
+on average newer than the CPU of Titan. The heterogeneity of the Grid sites'
+CPUs explain the higher performance variance we observed compared to the
+performance consistency we measured on Titan.
 
@@
 Table~\ref{panda-olcf-stats} shows the overall statistical breakdown of the
-observed file I/O. ATLAS used between 1 and 300 worker nodes, and 35 on average.
-75\% of the jobs run by ATLAS consumed less than 25 nodes and 92\% less than
-100. During the 174 hours of data collection, 6.75 ATLAS jobs were executed on
-average per hour, each job running for an average of 1.74 hours. Every job read
-less than 250 GB and wrote less than 75 GB of data and, on average, each job
-read 20 GB and wrote 6 GB of data.
+observed file I/O. ATLAS used between 1 and 300 worker nodes, and 35 on
+average. 75\% of the jobs run by ATLAS consumed less than 25 nodes and 92\%
+less than 100. During the 174 hours of data collection, 6.75 ATLAS jobs were
+executed on average per hour, each job running for an average of 1.74 hours.
+Every job read less than 250 GB and wrote less than 75 GB of data and, on
+average, each job read 20 GB and wrote 6 GB of data.
 
@@
 \centering
-\begin{tabular}{lllllllll}
- & Num. Nodes & Duration (s) & Read (GB) & Written (GB) & GB Read/nodes & GB Written/nodes & $open()$ & $close()$ \\
-Min & 1 & 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\
-Max & 300 & 7,452 & 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\
-Average & 35.66 & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 & 34,155.74 \\
-Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376 & 231,346.55 & 53,799.08
+\begin{tabular}{lllllllll} & Num. Nodes & Duration (s) & Read (GB) & Written
+ (GB) & GB Read/nodes & GB Written/nodes & $open()$ & $close()$ \\ Min & 1 &
+ 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\ Max & 300 & 7,452 &
+ 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\ Average & 35.66
+ & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 & 34,155.74 \\
+ Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376 & 231,346.55
+ & 53,799.08
 \end{tabular}
-\caption{The Statistical breakdown of the I/O impact of 1,175 jobs ATLAS executed at OLCF for the week of 10/25/16}
+\caption{The Statistical breakdown of the I/O impact of 1,175 jobs ATLAS
+executed at OLCF for the week of 10/25/16}
 \label{panda-olcf-stats}
@@
 
-ATLAS jobs are read heavy: On average, the amount of data read per worker node
-is less than 400 MB, while the amount of data written is less than 170 MB.
-Distributions of read and written data are different: The read operation
-distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
-the written amount of data has a very narrow distribution.
+ATLAS jobs are read heavy: On average, the amount of data read per worker
+node is less than 400 MB, while the amount of data written is less than 170
+MB\@. Distributions of read and written data are different: The read
+operation distribution per job shows a long tail, ranging from 12.5 GB to 250
+GB, while the written amount of data has a very narrow distribution.
 
@@
 maximum number of file $close()$ operations is $\approx$39/s. For the 1,175
-ATLAS jobs observed, the total number of file $open()$ operations is 172,089,760
-and the total number of file $close()$ operations is 40,132,992. The difference
-between these two values is still under investigation: One possible explanation
-is that ATLAS jobs don't call a file $close()$ operation per every file $open()$
-issued.
+ATLAS jobs observed, the total number of file $open()$ operations is
+172,089,760 and the total number of file $close()$ operations is 40,132,992.
+The difference between these two values is still under investigation: One
+possible explanation is that ATLAS jobs don't call a file $close()$ operation
+per every file $open()$ issued.
 
-Overall, the average time taken to read events from input files stored on Lustre
-is 1,320, comparable to the time taken to read the file required by assembling
-AthenaMP from NFS. Preliminary investigation shows that this time could be
-reduced to 40 seconds by loading the event descriptions into the RAM disk
-available on each worker node. Events descriptions could be transferred from
-Lustre to the RAM disk while configuring and initializing AthenaMP, almost
-halving the time currently required by initiating a Genat4 simulation.
+Overall, the average time taken to read events from input files stored on
+Lustre is 1,320, comparable to the time taken to read the file required by
+assembling AthenaMP from NFS. Preliminary investigation shows that this time
+could be reduced to 40 seconds by loading the event descriptions into the RAM
+disk available on each worker node. Events descriptions could be transferred
+from Lustre to the RAM disk while configuring and initializing AthenaMP,
+almost halving the time currently required by initiating a Genat4 simulation.

### 2017-06-27 PandaOverview.tex
Negative spaces fixing
Somewhat drastic space reduction

@@
-% PanDA is a Workload Management System (WMS) designed to support the execution of
-% distributed workloads and workflows via pilots. WMS is middleware for
-% discovering and selecting resources, submitting tasks of workloads and
-% workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
-% abstraction that enables multi-level scheduling by decoupling resource
-% acquisition from tasks scheduling~\cite{turilli2015comprehensive}. When
-% implemented, a pilot is scheduled on a site and, once active, tasks are
-% scheduled to the pilot, not to the site's scheduler.
-%
-% Pilot-capable WMS enable high throughput of tasks execution while supporting
-% interoperability across multiple sites. This is particularly relevant for LHC
-% experiments, where millions of tasks are executed across multiple sites every
-% month, analyzing and producing petabytes of data. The design of PanDA WMS
-% started in 2005 to support ATLAS. PanDA went into
-% production for the LHC Run 1 on 2009, and was then extended with new subsystems
-% for Run 2 in 2015.
+PanDA is a Workload Management System (WMS)%~\cite{marco2009glite} 
+designed to support the execution of distributed workloads and workflows via
+pilots~\cite{turilli2015comprehensive}. Pilot-capable WMS enable high
+throughput of tasks execution via multi-level scheduling while supporting
+interoperability across multiple sites. This is particularly relevant for LHC
+experiments, where millions of tasks are executed across multiple sites every
+month, analyzing and producing petabytes of data. 
+% The design of PanDA WMS started in 2005 to support ATLAS\@. PanDA went into
+% production for the LHC Run 1 on 2009, and was then extended with new
+% subsystems for Run 2 in 2015.
 
-PanDA is a Workload Management System (WMS)~\cite{marco2009glite} designed to
-support the execution of distributed workloads and workflows via
-pilots~\cite{turilli2015comprehensive}. Pilot-capable WMS enable high throughput
-of tasks execution via multi-level scheduling while supporting interoperability
-across multiple sites. This is particularly relevant for LHC experiments, where
-millions of tasks are executed across multiple sites every month, analyzing and
-producing petabytes of data. The design of PanDA WMS started in 2005 to support
-ATLAS. PanDA went into production for the LHC Run 1 on 2009, and was then
-extended with new subsystems for Run 2 in 2015.
+% ---------------------------------------------------------------------------
+% \subsection{Design}
+% \label{ssec:panda_design}
 
-% -----------------------------------------------------------------------------
-\subsection{Design}
-\label{ssec:panda_design}
+% PanDA's application model assumes tasks grouped into workloads and
+% workflows. Tasks represent a set of operations performed on data sets
+% stored in one or more input files. Tasks are decomposed into jobs, where
+% each job consists of the task's operations performed on a partition of the
+% task's data set. Since 2005, a certain amount of parallelism has been
+% progressively introduced for job execution~\cite{crooks2012multi} but, so
+% far, HEP applications have not required a message passing interface (MPI).
 
-PanDA's application model assumes tasks grouped into workloads and workflows.
-Tasks represent a set of operations performed on data sets stored in one or more
-input files. Tasks are decomposed into jobs, where each job consists of the
-task's operations performed on a partition of the task's data set. Since 2005, a
-certain amount of parallelism has been progressively introduced for job
-execution~\cite{crooks2012multi} but, so far, HEP applications have not required
-a message passing interface (MPI).
-% Jobs are supposed to be capable of setting up their own execution
-% environment and have a minimal set of common dependences.
+% PanDA's usage model assumes multitenancy of resources and at least two
+% types of HEP users: individual researchers and groups executing so called
+% `production' workflows. PanDA's security model is based on separation
+% between authentication, authorization and accounting for both single users
+% and group of users. Both authentication and authorization are based on
+% digital certificates and on the virtual organization
+% abstraction~\cite{foster2001anatomy}.
 
-PanDA's usage model assumes multitenancy of resources and at least two types of
-HEP users: individual researchers and groups executing so called `production'
-workflows.
-% Users can submit tasks and workflows to the PanDA WMS at any point in
-% time, directly or via dedicated application frameworks.
-PanDA's security model is based on separation between authentication,
-authorization and accounting for both single users and group of users. Both
-authentication and authorization are based on digital certificates and on the
-virtual organization abstraction~\cite{foster2001anatomy}.
+% Currently, PanDA's execution model is based on four main abstractions:
+% task, job, queue, and pilot. Both tasks and jobs are assumed to have
+% attributes and states and to be queued into a global queue for execution.
+% Prioritization and binding of jobs are assumed to depend on the attributes
+% of each job. Pilot is used to indicate the abstraction of resource
+% capabilities. Each job is thought to be bound to one pilot and executed on
+% the site where the pilot has been instantiated.
 
-Currently, PanDA's execution model is based on four main abstractions: task,
-job, queue, and pilot. Both tasks and jobs are assumed to have attributes and
-states and to be queued into a global queue for execution. Prioritization and
-binding of jobs are assumed to depend on the attributes of each job. Pilot is
-used to indicate the abstraction of resource capabilities. Each job is thought
-to be bound to one pilot and executed on the site where the pilot has been
-instantiated.
+% In PanDA's data model, each datum refers to the recorded or simulated
+% measurement of a physical process. Data can be packaged into files or other
+% containers. As with jobs, data have both attributes and states, and some of
+% the attributes are shared between events and jobs. Raw, reconstruction, and
+% simulation data are assumed to be distributed across multiple storage
+% facilities and managed by the ATLAS Distributed Data Management
+% (DDM)%~\cite{garonne2012atlas}. 
+% When necessary, datasets required by each job are assumed to be replicated
+% over the network, both for input and output data.
 
-In PanDA's data model, each datum refers to the recorded or simulated
-measurement of a physical process. Data can be packaged into files or other
-containers. As with jobs, data have both attributes and states, and some of the
-attributes are shared between events and jobs. Raw, reconstruction, and
-simulation data are assumed to be distributed across multiple storage facilities
-and managed by the ATLAS Distributed Data Management
-(DDM)~\cite{garonne2012atlas}. When necessary, datasets required by each job are
-assumed to be replicated over the network, both for input and output data.
+% PanDA's design supports provenance and traceability for both jobs and data.
+% Attributes enable provenance by linking jobs and data items, providing
+% information like ownership or project affiliation. States enable
+% traceability by providing information about the stage of the execution in
+% which each job or data item is or has been.
 
-PanDA's design supports provenance and traceability for both jobs and data.
-Attributes enable provenance by linking jobs and data items, providing
-information like ownership or project affiliation. States enable traceability by
-providing information about the stage of the execution in which each job or data
-item is or has been.
-% Some attributes are assumed to be immutable across execution and jobs and data
-% items are assumed to be always in one and only one state.
+% ---------------------------------------------------------------------------
+% \subsection{Implementation and Execution}
+% \label{ssec:panda_arch}
 
-% -----------------------------------------------------------------------------
-\subsection{Implementation and Execution}
-\label{ssec:panda_arch}
-
-The implementation of PanDA WMS consists of several interconnected subsystems,
-most of them built from off-the-shelf and Open Source components. Subsystems
-communicate via dedicated API or HTTP messaging, and each subsystem is
-implemented by one or more modules. Databases are used to store eventful
-entities % and entities' description
-like tasks, jobs and  input/output data, and to store information about sites,
-resources, logs, and accounting.
+The implementation of PanDA WMS consists of several interconnected
+subsystems,
+% , most of them built from off-the-shelf and Open Source components.
+% Subsystems communicate 
+communicating via dedicated API or HTTP messaging, and implemented by one or
+more modules. Databases are used to store eventful entities like tasks, jobs
+and input/output data, and to store information about sites, resources,
+logs, and accounting.
 
@@
 Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
-Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of ATLAS
-workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we do not
-discuss them as they are  not relevant to an understanding of how PanDA has been
-ported to supercomputers. For a full list of subsystems see
-Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture} shows a diagrammatic
-representation of PanDA main subsystems, highlighting the execution process of
-tasks while omitting monitoring details to improve readability.
+Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of
+ATLAS workflows % (e.g., PanDA Event Service~\cite{calafiura2015atlas}), 
+but we do not discuss them as they are not relevant to an understanding of
+how PanDA has been ported to supercomputers. For a full list of subsystems
+see Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture} shows a
+diagrammatic representation of PanDA main subsystems, highlighting the
+execution process of tasks while omitting monitoring details to improve
+readability.
 
-% The relation between tasks and jobs can be one-to-one or one-to-many, and the
-% conversion between the two can by static or dynamic.
+% During LHC Run 1, PanDA required users to perform a static conversion
+% between tasks and jobs: tasks were described as a set of jobs and then
+% submitted to the PanDA Server. This introduced inefficiency both with
+% usability and resource utilization.%~\cite{borodin2015unified}. Ideally,
+% users should conceive analyses in terms of one or more, possibly related
+% tasks, while the execution manager (i.e., PanDA) should partition tasks
+% into jobs, depending on execution constraints. Further, due to the static
+% partitioning of tasks into jobs, PanDA instantiates pilots on sites without
+% taking into account the heterogeneity and dynamicity of the resources. An
+% optimal sizing of each job should take into account these properties.
 
-During LHC Run 1, PanDA required users to perform a static conversion between
-tasks and jobs: tasks were described as a set of jobs and then submitted to the
-PanDA Server. This introduced inefficiency both with usability and resource
-utilization~\cite{borodin2015unified}. Ideally, users should conceive analyses
-in terms of one or more, possibly related tasks, while the execution manager
-(i.e., PanDA) should partition tasks into jobs, depending on execution
-constraints. Further, the static partitioning of tasks into jobs does not take
-into account the heterogeneity and dynamicity of the resources on which each job
-will be executed, introducing inefficiencies.
+% Another problem of static job sizing is that PanDA instantiates pilots on
+% sites with different type of resources and different models of availability
+% of those resources. An optimal sizing of each job should take into account
+% these properties. For example, sites may offer cores with different speed,
+% networking with different amount of bandwidth, and resources could be
+% guaranteed to be available for a defined amount of time or could disappear
+% at any point in time as it happens with opportunistic models of resource
+% provision.
 
-Another problem of static job sizing is that PanDA instantiates pilots on sites
-with different type of resources and different models of availability of those
-resources. An optimal sizing of each job should take into account these
-properties. For example, sites may offer cores with different speed, networking
-with different amount of bandwidth, and resources could be guaranteed to be
-available for a defined amount of time or could disappear at any point in time
-as it happens with opportunistic models of resource provision.
+% JEDI was deployed for the LHC Run 2 to address these inefficiencies. 
+Users submit task descriptions to JEDI (Fig.~\ref{fig:architecture}:1) that
+stores them into a queue implemented by a database
+(Fig.~\ref{fig:architecture}:2). Tasks are partitioned into jobs of different
+size, depending on both static and dynamic information about available
+resources (Fig.~\ref{fig:architecture}:3). Jobs are bound to sites with
+resources that best match jobs' requirements, and submitted to the PanDA
+Server for execution (Fig.~\ref{fig:architecture}:4).
 
-JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users
-submit task descriptions to JEDI (Fig.~\ref{fig:architecture}:1) that stores
-them into a queue implemented by a database (Fig.~\ref{fig:architecture}:2).
-Tasks are partitioned into jobs of different size, depending on both static and
-dynamic information about available resources (Fig.~\ref{fig:architecture}:3).
-Jobs are bound to sites with resources that best match jobs' requirements, and
-submitted to the PanDA Server for execution (Fig.~\ref{fig:architecture}:4).
-
-Once submitted to the PanDA Server, jobs are stored by the Task Buffer component
-into a global queue implemented as a database (Fig.~\ref{fig:architecture}:5).
-When jobs are submitted directly to the PanDA Server, the Brokerage component is
-used to bind jobs to available sites, depending on static information about the
-resources available for each site. Jobs submitted by JEDI are already bound to
-sites so no further brokerage is needed.
+Once submitted to the PanDA Server, jobs are stored by the Task Buffer
+component into a global queue implemented as a database
+(Fig.~\ref{fig:architecture}:5). When jobs are submitted directly to the
+PanDA Server, the Brokerage component is used to bind jobs to available
+sites, depending on static information about the resources available for each
+site. Jobs submitted by JEDI are already bound to sites so no further
+brokerage is needed.
 
@@
 (Fig.~\ref{fig:architecture}:6). The Data Service communicates these
-requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7) that, when needed,
-replicates data sets on the required sites (Fig.~\ref{fig:architecture}:8).
+requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7) that, when
+needed, replicates data sets on the required sites
+(Fig.~\ref{fig:architecture}:8).
 
 Meanwhile, AutoPyFactory defines PanDA Pilots, submitting them to a Condor-G
-agent (Fig.~\ref{fig:architecture}:9). Condor-G schedules these pilots wrapped
-as jobs or VMs to the required sites (Fig.~\ref{fig:architecture}:10).
+agent (Fig.~\ref{fig:architecture}:9). Condor-G schedules these pilots
+wrapped as jobs or VMs to the required sites
+(Fig.~\ref{fig:architecture}:10).
 
-When a PanDA Pilot becomes available, it requests the Job Dispatcher module of
-the PanDA Server for a job to execute (Fig.~\ref{fig:architecture}:11). The Job
-Dispatcher interrogates the Task Buffer module for a job that is bound to the
-site of that pilot and ready to be executed. Task Buffer checks the global queue
-(i.e., the PanDA database) and, upon availability, returns a job to the Job
-Dispatcher. The Job Dispatcher dispatches that job to the PanDA Pilot
-(Fig.~\ref{fig:architecture}:12).
+When a PanDA Pilot becomes available, it requests the Job Dispatcher module
+of the PanDA Server for a job to execute (Fig.~\ref{fig:architecture}:11).
+The Job Dispatcher interrogates the Task Buffer module for a job that is
+bound to the site of that pilot and ready to be executed. Task Buffer checks
+the global queue (i.e., the PanDA database) and, upon availability, returns a
+job to the Job Dispatcher. The Job Dispatcher dispatches that job to the
+PanDA Pilot (Fig.~\ref{fig:architecture}:12).
 
@@
 from the stage-in location (Fig.~\ref{fig:architecture}:13) and the job's
-payload is executed (Fig.~\ref{fig:architecture}:14). Once completed, output is
-transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).
+payload is executed (Fig.~\ref{fig:architecture}:14). Once completed, output
+is transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).
 
@@
 generated by each job (Fig.~\ref{fig:architecture}:16), updating jobs'
-attributes via the Task Buffer module (Fig.~\ref{fig:architecture}:17). When the
-output of all the jobs of a task are retrieved, it is made available to the user
-via PanDA Server. When a task is submitted to JEDI, task is instead marked as
-done (Fig.~\ref{fig:architecture}:18) and the result of its execution is made
-available to the user by JEDI (Fig.~\ref{fig:architecture}:19).
+attributes via the Task Buffer module (Fig.~\ref{fig:architecture}:17). When
+the output of all the jobs of a task are retrieved, it is made available to
+the user via PanDA Server. When a task is submitted to JEDI, task is instead
+marked as done (Fig.~\ref{fig:architecture}:18) and the result of its
+execution is made available to the user by JEDI
+(Fig.~\ref{fig:architecture}:19).
 
@@
     \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
-    \caption{PanDA WMS architecture. Numbers indicates the JEDI-based execution
-    process. Several subsystems, components, and architectural and
+    \vspace{-0.3in}
+    \caption{PanDA WMS architecture. Numbers indicates the JEDI-based
+    execution process. Several subsystems, components, and architectural and
     communication details are abstracted to improve clarity.}

### 2017-06-29 introduction.tex
introduce  \ifreview
  to toggle between reviewer comments/not, so as to make it easier to get a handle on
  net space used.
reduction in introduction.   WIP.

@@
+\ifreview
 Comments to address in this section:
@@
 \end{enumerate}
+\fi
 
@@
 
-As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity era
-(Run 4), it is anticipated that the data volumes that will need analyzing
-will increase by factors of 10-100 compared to the current phase (Run 2).
-Data will be larger in volume but will also require more sophisticated
-computational processing. In spite of the impressive scale of the ATLAS
-distributed computing system, demand for computing systems will significantly
-outstrip current and projected supply.  There are multiple levels at which
-this problem needs to be addressed: the utilization of emerging parallel
+In spite of the impressive scale of the ATLAS distributed computing system,
+demand for computing systems will soon significantly outstrip current and
+projected supply.   The data volumes that will need analyzing in LHC-Run 3 in
+$~\approx$ 2022 and the high-luminosity era (Run 4) will increase by factors
+of 10-100 compared to the current phase (Run 2). There are multiple levels at
+which this problem needs to be addressed: the utilization of emerging parallel
 architectures (e.g., platforms); algorithmic and advances in analytical
@@
 
+% As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity era
+% (Run 4), it is anticipated that the data volumes that will need analyzing
+% will increase by factors of 10-100 compared to the current phase (Run 2).
+% Data will be larger in volume but will also require more sophisticated
+% computational processing.
+
 This paper represents the experience of how the ATLAS experiment has ``broken
 free'' of the traditional computational approach of high-throughput computing
-on distributed resources to embrace new platforms, in particular  high-
-performance computers. Specifically, we discuss the experience of integrating
+on distributed resources to embrace new platforms, in particular  
+high-performance computers. Specifically, we discuss the experience of integrating
 PanDA WMS with a US DOE leadership computing facility called Titan to reach

### 2017-07-01 PandaEvolution.tex
introduce  \ifreview
  to toggle between reviewer comments/not, so as to make it easier to get a handle on
  net space used.
Address referees' comments and further reductions

--- 2017-04-04 PandaEvolution.tex
@@
+% \ifreview
+% Comments to address in this section:
+% \begin{enumerate}
+% 	\item REVIEWER 2: The authors spent quite a few paragraphs for the detail
+% 	design and implementation of the PanDA WMS, but introduced few background
+% 	of the LHC applications. According to the paper, I understand the
+% 	applications running on Titan are a large amount of small independent
+% 	jobs. Thus, it seems less cost-effective to run these jobs on a
+% 	world-leading supercomputer like Titan which has a fast and expensive
+% 	interconnection. It will be more interesting if the authors can compare
+% 	the solution on Titan (proposed in this paper) with the original
+% 	distributed computing environment and HPECloud from Google mentioned in
+% 	the related work section.
+% 	\item REVIEWER 5: In Section 3.2, GPUs are not considered in this work,
+% 	which would be a waste for using HPC computing. Could this work be
+% 	extended to include GPUs or other heterogenous resources?
+% \end{enumerate}
+% \fi
 The upcoming LHC Run 3 will require more resources than the Worldwide LHC
-Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than 300,000
-cores at over 100 Grid sites, with a peak performance of 0.3 petaFLOPS. This
-capacity will be sufficient for the planned analysis and data processing, but it
-will be insufficient for the Monte Carlo production workflow and any extra
-activity. To alleviate these challenges, ATLAS is expanding the current
-computing model to include additional resources such as the opportunistic use of
-supercomputers.
+Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
+300,000 cores at over 100 Grid sites, with a peak performance of 0.3
+petaFLOPS\@. This capacity will be sufficient for the planned analysis and
+data processing, but it will be insufficient for the Monte Carlo production
+workflow and any extra activity. To alleviate these challenges, ATLAS is
+expanding the current computing model to include additional resources such as
+the opportunistic use of supercomputers.
 
@@
 requires runtime communication. Jobs are executed across multiple cores, each
-core calculating a small part of the problem and communicating with other cores
-via MPI. Accordingly, supercomputers have large number of worker nodes,
-connected through a high-speed, low-latency dedicated network. Each worker node
-has multicore CPUs, usually augmented with Graphics Processing Units (GPUs) or
-other specialized coprocessors.
+core calculating a small part of the problem and communicating with other
+cores via MPI\@. Accordingly, supercomputers have large number of worker
+nodes, connected through a high-speed, low-latency dedicated network. Each
+worker node has multicore CPUs, usually augmented with Graphics Processing
+Units (GPUs) or other specialized coprocessors.
 
@@
 ATLAS workloads or workflows involves concurrent and/or sequential runs of
-possibly large amount of jobs, each requiring no or minimal parallelization and
-no runtime communication. Thus, computing infrastructure like WLCG have been
-designed to aggregate large amount of computing resources across multiple sites.
-While each site may deploy MPI capabilities, usually these are not used to
-perform distributed computations.
+possibly large amount of jobs, each requiring no or minimal parallelization
+and no runtime communication. Thus, computing infrastructure like WLCG have
+been designed to aggregate large amount of computing resources across
+multiple sites. While each site may deploy MPI capabilities, usually these
+are not used to perform distributed computations.
 
-We developed and deployed a single-point solution to better understand the
-problem space of enabling a WMS designed for HTC to execute production workflows
-on resources designed to support HPC. The PanDA team developed a job broker to
+Currently, ATLAS workloads do not require fast interconnects or specilized
+coprocessors but supercomputers tend not to reach 100\% utilization due to
+the scheduling of jobs requiring large amount of resources. This offers the
+possibility to execute ATLAS-like workloads on supercomputers to increase
+utilization and reducing the waste of available resources.
+
+We developed a single-point solution to better understand the problem space
+of enabling a WMS designed for HTC to execute production workflows on
+resources designed to support HPC\@. The PanDA team developed a job broker to
 support the execution of part of the ATLAS production Monte Carlo workflow on
@@
 
-% -----------------------------------------------------------------------------
-\subsection{Architectures and Interfaces}
+% ---------------------------------------------------------------------------
+\subsection{Architectures, Interfaces and Workloads}
 \label{ssec:panda-titan}
 
-The Titan supercomputer, current number three on the Top 500 list~\cite{top500},
-is a Cray XK7 system with 18,688 worker nodes and a total of 299,008 CPU cores.
-Each worker node has an AMD Opteron  6274 16-core CPU, a Nvidia Tesla K20X GPU,
-32 GB of RAM and no local storage, though a 16 GB RAM disk can be set up. Work
-nodes use Cray’s Gemini interconnect for inter-node MPI messaging. Titan is
-served by the Spider II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of
-disk storage, and by a 29 PB HPSS tape storage system. Titan’s worker nodes run
-Compute Node Linux, a run time environment based on SUSE Linux Enterprise
-Server.
+% The Titan supercomputer, current number three on the Top 500
+% list,%~\cite{top500}, 
+% is a Cray XK7 system with 18,688 worker nodes and a total
+% of 299,008 AMD cores and 18,688 Nvidia Tesla GPUs. 
+% Each worker node has an AMD Opteron 6274 16-core CPU,
+% a Nvidia Tesla K20X GPU, 32 GB of RAM and no local storage, though a 16 GB
+% RAM disk can be set up. Work nodes use Cray’s Gemini interconnect for
+% inter-node MPI messaging. Titan is served by the Spider
+% II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of disk storage, and
+% by a 29 PB HPSS tape storage system. Titan’s worker nodes run Compute Node
+% Linux, a run time environment based on SUSE Linux Enterprise Server.
 
-Titan's users submit jobs to Titan's PBS scheduler by logging into login or data
-transfer nodes (DTNs). Titan's authentication and authorization model is based
-on two-factor authentication with a RSA SecurID key. Login nodes and DTNs have
-out/inbound wide area network connectivity while worker nodes have only local
-network access. Fair-share and allocation policies are in place both for the PBS
-batch system and shared file systems.
+% Titan's users submit jobs to Titan's PBS scheduler by logging into login or
+% data transfer nodes (DTNs). Titan's authentication and authorization model
+% is based on two-factor authentication with a RSA SecurID key. Login nodes
+% and DTNs have out/inbound wide area network connectivity while worker nodes
+% have only local network access. Fair-share and allocation policies are in
+% place both for the PBS batch system and shared file systems.
 
-Titan's architecture, configuration and policies poses several challenges to the
-integration with PanDA. The default deployment
-model of PanDA Pilot is unfeasible on Titan: PanDA Pilot is required to contact
-the Job Dispatcher of the PanDA Server to pull jobs to execute, but this is not
-possible on Titan because worker nodes do not offer outbound network
-connectivity. Further, Titan does not support PanDA's security model based on
-certificates and virtual organizations, making PanDA's approach to identity
-management also unfeasible. While DTNs offer wide area network data transfer, an
-integration with ATLAS DDM is beyond the functional and administrative scope of
-the current prototyping phase. Finally, the specific characteristics of the
-execution environment, especially the absence of local storage on the worker
-nodes and modules tailored to Compute Node Linux, require re-engineering of
-ATLAS application frameworks.
+Titan's architecture, configuration and policies poses several challenges to
+the integration with PanDA\@. The default deployment model of PanDA Pilot is
+unfeasible on Titan: PanDA Pilot is required to contact the Job Dispatcher of
+the PanDA Server to pull jobs to execute, but this is not possible on Titan
+because worker nodes do not offer outbound network connectivity. Further,
+Titan does not support PanDA's security model based on certificates and
+virtual organizations, making PanDA's approach to identity management also
+unfeasible. While Titan's data transfer nodes (DTNs) offer wide area network
+data transfer, an integration with ATLAS DDM is beyond the functional and
+administrative scope of the current prototyping phase. Finally, the specific
+characteristics of the execution environment, especially the absence of local
+storage on the worker nodes and modules tailored to Compute Node Linux,
+require re-engineering of ATLAS application frameworks.
 
 Currently, very few HEP applications can benefit from Titan's GPUs but some
-computationally-intensive and non memory-intensive tasks of ATLAS workflows can
-be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
-partitioned into independent jobs, Titan worker nodes can be used to execute up
-to 16 concurrent payloads, one per each available core. Given these constraints
-and challenges, the type of task most suitable for execution at the moment on
-Titan is Monte Carlo detector simulation. This type of task is mostly
-computational-intensive, requiring less than 2GB of RAM at runtime and with
-small input data requirements. Detector simulation tasks in ATLAS are performed
-via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
-GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
-$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
-offloading.
+computationally-intensive and non memory-intensive tasks of ATLAS workflows
+can be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
+partitioned into independent jobs, Titan worker nodes can be used to execute
+up to 16 concurrent payloads, one per each available core. Given these
+constraints and challenges, the type of task most suitable for execution at
+the moment on Titan is Monte Carlo detector simulation. This type of task is
+mostly computational-intensive, requiring less than 2GB of RAM at runtime and
+small input data. Detector simulation tasks in ATLAS are performed via
+AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
+GEANT4 detector simulation toolkit~\cite{agostinelli2003geant4}. These tasks
+account for $\approx$ 60\% of all the jobs on WLCG, making them a primary
+candidate for offloading.
 
 Detector simulation is part of the ATLAS production Monte Carlo (MC)
-workflow~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
-consists of four main stages: event generation, detector simulation,
-digitization, and reconstruction. Event generation creates sets of particle
-four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
-HERWIG~\cite{corcella2001herwig} and many others. The detector simulator is
-called Geant4~\cite{agostinelli2003geant4} and simulates the ATLAS detector and
-the interaction between that detector and particles. Each interaction creates a
-so-called hit and all hits are collected and passed on for digitalization, where
-hits are further processed to mimic the readout of the detector. Finally,
-reconstruction operates local pattern recognition, creating high-level objects
-like particles and jets.
+workflow~\cite{rimoldi2006atlas}.%,de2013delphes,ritsch2014atlas}. 
+The MC workflow consists of four main stages: event generation, detector
+simulation, digitization, and reconstruction. Event generation creates sets
+of particle four-momenta via different generators, e.g.,
+PYTHIA,%~\cite{sjostrand2006pythia}, 
+HERWIG,%~\cite{corcella2001herwig} 
+and many others. Geant4 simulates the ATLAS detector and the interaction
+between that detector and particles. Each interaction creates a so-called hit
+and all hits are collected and passed on for digitalization, where hits are
+further processed to mimic the readout of the detector. Finally,
+reconstruction operates local pattern recognition, creating high-level
+objects like particles and jets.
 
-% -----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
 \subsection{PanDA Broker}
@@
 particularly relevant for machines like Titan where queue policies privilege
-parallel jobs on the base of the number of worker nodes they request: the higher
-the number of nodes, the shorter the amount of queue time (modulo fair-share and
-allocation policies).
+parallel jobs on the base of the number of worker nodes they request: the
+higher the number of nodes, the shorter the amount of queue time (modulo
+fair-share and allocation policies).
 
-The backfill optimization of Titan's Moab scheduler allows to avoid the overhead
-of queue wait times without using pilot abstraction~\cite{maui_backfill_url}.
-With this optimization, Moab starts low-priority jobs when they do not delay
-higher priority jobs, independent of whether the low-priority jobs were queued
-after the high-priority jobs.
+The backfill optimization of Titan's Moab scheduler allows to avoid the
+overhead of queue wait times without using pilot
+abstraction~\cite{maui_backfill_url}. With this optimization, Moab starts
+low-priority jobs when they do not delay higher priority jobs, independent of
+whether the low-priority jobs were queued after the high-priority ones.
 
-When the backfill optimization is enabled, users can interrogate Moab about the
-number of worker nodes and walltime that would be available to a low-priority
-job at that moment in time. If a job is immediately submitted to Titan with that
-number of worker nodes and walltime, chances are that Moab will immediately
-schedule it, reducing its queue time to a minimum. In this paper, we call this
-number of worker nodes and walltime an available `backfill slot'.
+When the backfill optimization is enabled, users can interrogate Moab about
+the number of worker nodes and walltime that would be available to a
+low-priority job at that moment in time. If a job is immediately submitted to
+Titan with that number of worker nodes and walltime, chances are that Moab
+will immediately schedule it, reducing its queue time to a minimum. In this
+paper, we call this number of worker nodes and walltime an available
+`backfill slot'.
 
 Compared to pilots, backfill has the disadvantage of limiting the amount of
-worker nodes that can be requested. Pilots are normal jobs: they can request as
-many worker nodes and walltime as a queue can offer. On the contrary, jobs sized
-according to an available backfill slot depend on the number of worker nodes and
-walltime that cannot be given to any other job at that moment in time.
+worker nodes that can be requested. Pilots are normal jobs: they can request
+as many worker nodes and walltime as a queue can offer. On the contrary, jobs
+sized according to an available backfill slot depend on the number of worker
+nodes and walltime that cannot be given to any other job at that moment in
+time.
 
 At any point in time, the size of an available backfill slot is typically a
-small fraction of the total capacity of a resource. Notwithstanding, given the
-size of Titan this translates into a substantial capacity. Every year, about
-10\% of Titan's capacity remains unused~\cite{barker2016us}, corresponding to an
-average of 30,000 unused cores (excluding GPU cores). This equals to roughly
-10\% of the overall capacity of WLCG.
+small fraction of the total capacity of a resource. Notwithstanding, given
+the size of Titan this translates into a substantial capacity. Every year,
+about 10\% of Titan's capacity remains unused~\cite{barker2016us},
+corresponding to an average of 30,000 unused cores (excluding GPU cores).
+This equals to roughly 10\% of the overall capacity of WLCG\@.
 
-Given the communication requirements of PanDA Pilots and the unused capacity of
-Titan, PanDA pilot was repurposed to serve as a job broker on the DTN nodes of
-Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of PanDA Pilot
-and its stand-alone architecture, this prototype called `PanDA Broker'
-implements functionalities to: (i) interrogate Titan about backfill
-availability; (ii) pull ATLAS jobs and events from PanDA Server; (iii) wrap the
-payload of ATLAS jobs into MPI scripts; (iv) submitting MPI scripts to Titan's
-PBS batch system and monitor their execution; and (v) staging and preparing
-input/output files. Backfill querying, payload wrapping, and scripts submission
-required a new implementation while pulling ATLAS job and events, and file
-staging were inherited from PanDA Pilot.
+Given the communication requirements of PanDA Pilots and the unused capacity
+of Titan, PanDA pilot was repurposed to serve as a job broker on the DTN
+nodes of Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of
+PanDA Pilot and its stand-alone architecture, this prototype called `PanDA
+Broker' 
+% implements functionalities to: (i) interrogate Titan about backfill
+% availability via a dedicated Moab scheduler command; (ii) pull ATLAS jobs
+% and events from PanDA Server; (iii) wrap the payload of ATLAS jobs into MPI
+% scripts; (iv) submitting workload-specific MPI scripts to Titan's PBS batch
+% system and monitor their execution; and (v) staging and preparing
+% input/output files. Backfill querying, payload wrapping, and scripts
+% submission required a new implementation while pulling ATLAS job and
+% events, and file staging were inherited from PanDA Pilot.
+%
+% Backfill querying is performed via a dedicated Moab scheduler command while
+% a tailored Python MPI script is used to execute the payload of ATLAS jobs.
+% This enables the execution of unmodified Grid-centric, ATLAS jobs on Titan.
+% Typically, a MPI script is workload-specific as it sets up the execution
+% environment for a specific payload. This involves organization of worker
+% directories, data management, optional input parameters modification, and
+% cleanup on exit. Upon submission, a copy of the MPI script runs on every
+% available worker node, starting the execution of the ATLAS job's payload in
+% a subprocess and waits until its completion.
+%
+% MPI scripts are submitted to Titan's PBS batch system via
+% RADICAL-SAGA~\cite{radical-saga_url}, a Python module, compliant with the
+% OGF GFD.90 SAGA specification~\cite{goodale2008simple}. The Simple API for
+% Grid Applications (SAGA) offers a unified interface to diverse job
+% schedulers and file transferring services. In this way, SAGA provides an
+% interoperability layer that lowers the complexity of using distributed
+% infrastructures. Behind the API façade, RADICAL-SAGA implements a adaptor
+% architecture: each adaptor interface the SAGA API with different middleware
+% systems and services, including the PBS batch scheduler of Titan.
+%
+% The data staging capabilities of the PanDA Broker are implemented via a
+% file system shared among DTNs and worker nodes. The input files with the
+% events of the ATLAS jobs are downloaded on the shared filesystem from the
+% data center of Brookhaven National Laboratory (BNL). The MPI script setup
+% process includes making the location of these files available to the
+% payload of the ATLAS's jobs. The PanDA Broker can locate the payload's
+% output files on the shared filesystem and transfer them from Titan BNL\@.
+%
+% Once deployed on Titan, every PanDA Broker 
+supports the execution of MC detector simulations in 9 steps.
 
-Backfill querying is performed via a dedicated Moab scheduler command while a
-tailored Python MPI script is used to execute the payload of ATLAS jobs. This
-MPI script enables the execution of unmodified Grid-centric, ATLAS jobs on
-Titan. Typically, a MPI script is workload-specific as it sets up the execution
-environment for a specific payload. This involves organization of worker
-directories, data management, optional input parameters modification, and
-cleanup on exit. Upon submission, a copy of the MPI script runs on every
-available worker node, starting the execution of the ATLAS job's payload in a
-subprocess and waits until its completion.
-
-MPI scripts are submitted to Titan's PBS batch system via
-RADICAL-SAGA~\cite{radical-saga_url}, a Python module, compliant with the OGF
-GFD.90 SAGA specification~\cite{goodale2008simple}. The Simple API for Grid
-Applications (SAGA) offers a unified interface to diverse job schedulers and
-file transferring services. In this way, SAGA provides an interoperability layer
-that lowers the complexity of using distributed infrastructures. Behind the API
-façade, RADICAL-SAGA implements a adaptor architecture: each adaptor interface
-the SAGA API with different middleware systems and services, including the PBS
-batch scheduler of Titan.
-
-The data staging capabilities of the PanDA Broker are implemented via a file
-system that is shared among DTNs and worker nodes. The input files with the
-events of the ATLAS jobs are downloaded on the shared filesystem from the data
-center of Brookhaven National Laboratory (BNL). The MPI script setup process
-includes making the location of these files available to the payload of the
-ATLAS's jobs. The PanDA Broker can locate the payload's output files on the
-shared filesystem and transfer them from Titan BNL.
-
-Once deployed on Titan, every PanDA Broker supports the execution of MC detector
-simulations in 9 steps. PanDA Broker queries the Job Dispatcher module of the
+Each PanDA Broker deployed on Titan queries the Job Dispatcher module of the
 PanDA server for ATLAS jobs that have been bound to Titan by JEDI
-(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA Broker
-pulls jobs' input files from BNL to the OLCF Lustre file system
+(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA
+Broker pulls jobs' input files from BNL to the OLCF Lustre file system
 (Fig.~\ref{fig:panda_broker}:2). PanDA Broker queries Titan's Moab scheduler
 about the current available backfill slot (Fig.~\ref{fig:panda_broker}:3) and
-creates an MPI script, wrapping enough ATLAS jobs' payload to fit the backfill
-slot. PanDA Broker submits the MPI script to the Titan's Torque batch system via
-RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).
+creates an MPI script, wrapping enough ATLAS jobs' payload to fit the
+backfill slot. PanDA Broker submits the MPI script to the Titan's Torque
+batch system via RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).
 

### 2017-07-01 overall.tex
introduce  \ifreview
  to toggle between reviewer comments/not, so as to make it easier to get a handle on
  net space used.
Address referee's comments and marginal space saving

@@
+\ifreview
 Comments to address in this section:
 \begin{enumerate}
-    \item REVIEWER 2: In Figure 4, there are some inconsistencies. The figure
-    caption shows there are only 62175 measures, but in the scatter plot
-    (right) there are 62555 data points. In addition, according to the
-    distribution of the number of work nodes (left), it does not look like
-    the mean number of work available is 691 as the figure caption
-    illustrates.
-    \item REVIEWER 3: t’s interesting that Feb/March have such an increase in
-    backfill availability in 2017. The authors attribute this to hardware
+    % \item REVIEWER 2: In Figure 4, there are some inconsistencies. The
+    % figure caption shows there are only 62175 measures, but in the scatter
+    % plot (right) there are 62555 data points. In addition, according to the
+    % distribution of the number of work nodes (left), it does not look like
+    % the mean number of work available is 691 as the figure caption
+    % illustrates.
+    \item REVIEWER 3: It’s interesting that Feb/March have such an increase
+    in backfill availability in 2017. The authors attribute this to hardware
     upgrades for the DTNs. It’s mentioned that the load is only 3.6\% now and
@@
     utilized in April/May 2016?
-    \item REVIEWER 4: “Genat4” at the end of Section 4 is a typo.
-    \item REVIEWER 4: Please explain why you used Titan at all, in spite of
-    your statement that it is 1/2 to 1/3 as efficient as other resources. Was
-    it merely that you were given so much time on this resource that the
-    inefficiency doesn’t matter?
+    % \item REVIEWER 4: “Genat4” at the end of Section 4 is a typo.
+    % \item REVIEWER 4: Please explain why you used Titan at all, in spite of
+    % your statement that it is 1/2 to 1/3 as efficient as other resources.
+    % Was it merely that you were given so much time on this resource that
+    % the inefficiency doesn’t matter?
     \item REVIEWER 4: Why didn’t you benchmark the performance of your
@@
     rather than an accomplishment.
-    \item REVIEWER 5: In Figure 3, it shows the unused core hours are much
-    higher than those used by ATLAS. What are the reasons for such low usage?
+    % \item REVIEWER 5: In Figure 3, it shows the unused core hours are much
+    % higher than those used by ATLAS\@. What are the reasons for such low
+    % usage?
 \end{enumerate}
+\fi
 
@@
 have operated only in backfill mode, without a defined time allocation, and
-running at the lowest priority on Titan.
+running at the lowest priority on Titan. Therefore, ATLAS contributed to
+increase the Titan's overall utilization without affecting other users.
 
@@
 the measurements were performed between January 2016 and February 2017,
-hereafter called ‘experiment time window
+hereafter called `experiment time window'.
 
@@
 ATLAS (blue bars) during the experiment time window. ATLAS consumed a total
-of 51.4M core-hours, for an average of $\approx$3.7M core-hours a month, with
-a minimum of 1.7M core-hours in April 2016 and a maximum 7.9M core-hours in
-February 2017.
+of 51.4M core-hours, for an average of $\approx$3.7M core-hours a month. 
+% , with a minimum of 1.7M core-hours in April 2016 and a maximum 7.9M
+% core-hours in February 2017.
 
@@
     CPU core-hours used by ATLAS (blue). GPU core-hours unaccounted for as
-    they cannot be used by ATLAS. Efficiency of PanDA Brokers defined as
+    they cannot be used by ATLAS\@. Efficiency of PanDA Brokers defined as
     percentage of total Titan's backfill availability used by ATLAS (Red
@@
 the upgraded DTNs. As such, further hardware upgrades seem unlikely to
-improve significantly the performance of PanDA Brokers.
+improve significantly the performance of PanDA Brokers.\mtnote{Asked Segey
+and Danila for values previous HW upgrade}
 
@@
 function of the number of events simulated but not all events take the same
-time to be simulated. 1 event simulation takes from $\approx$2 to $\approx$40
-minutes, with a mean of $\approx$14 minutes. Considering that each worker
-node process up to 16 events concurrently, 100 events takes an average of 105
-minutes to process. As such, PanDA brokers do not use backfill availability
-with less than 105 minutes walltime.
+time to be simulated. One event simulation takes from $\approx$2 to
+$\approx$40 minutes, with a mean of $\approx$14 minutes. Considering that
+each worker node process up to 16 events concurrently, 100 events takes an
+average of 105 minutes to process. As such, PanDA brokers do not use backfill
+availability with less than 105 minutes walltime.
 
-Fig.~\ref{fig:backfill-distrib} shows the distribution of backfill
-availability on Titan as a function of number of nodes and the time of their
-availability (i.e., walltime). We recorded these data by polling Titan's Moab
-scheduler at regular intervals during the experiment window time. The mean
-number of nodes was 691, and their mean walltime was 126 minutes. Detector
-simulations of 100 events, enable to use down to 5/6 of the mean walltime of
-backfill availability. As such, it offers a good compromise for PanDA Broker
+Fig.~\ref{fig:backfill-distrib} shows backfill availability on Titan as a
+function of number of nodes and the time of their availability (i.e.,
+walltime). We recorded these data by polling Titan's Moab scheduler at
+regular intervals during the experiment window time. The mean number of nodes
+was 691, and their mean walltime was 126 minutes. Detector simulations of 100
+events, enable to use down to 5/6 of the mean walltime of backfill
+availability. As such, it offers a good compromise for PanDA Broker
 efficiency.
 
-\begin{figure*}[!t]
-    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_wnodes_distribution.pdf}
-    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_walltime_distribution.pdf}
-    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
+%\begin{figure*}[!t]
+\begin{figure}[!t]
+    % \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_wnodes_distribution.pdf}
+    % \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_walltime_distribution.pdf}
+    % \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
+    \includegraphics[clip,width=\columnwidth]{figures/titan_backfill_avail.png}
    \vspace{-0.1in}
-    \caption{Backfill availability: distribution of number of work nodes
-    (left) and of walltime in minutes (center). Scatter plot of the two
-    variables (right). 62175 measures: mean number of work nodes available
-    691; mean walltime available 126 minutes.}
-
+    \caption{
+    % Backfill availability: distribution of number of work nodes (left) and
+    % of walltime in minutes (center). Scatter plot of the two variables.
+    % (right).
+    62555 measures of Backfill availability on Titan during the experiment
+    time window. Mean number of work nodes available 691; mean walltime
+    available 126 minutes.}
 \label{fig:backfill-distrib}
-\end{figure*}
+%\end{figure*}
+\end{figure}
 
@@
 on DTNs, and on an alternative design of PanDA Broker that enables the
-concurrent submission of multiple MPI scripts~\cite{barreiro2016panda}.
+concurrent submission of multiple MPI scripts~\cite{barreiro2016panda}. The
+results of these investigation will contribute to improve the efficienty of
+PanDA Brokers beyond the 30\% limit showed in
+Fig.~\ref{fig:backfill-utilization}.
 
 The current design and architecture of the PanDA Broker is proving to be as
-reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
-the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
-while the failure rate of jobs submitted to Titan was a comparable 13.6\%.
-PanDA Brokers were responsible for around the 19\% of the failures, compared
-to the 29\% of failures produced by the JobDispatcher module of the PanDA
-Server, and the 13\% failures produced by the Geant4 toolkit. The current
-failure rate of the PanDA Brokers confirms the benefits of reusing most of
-the code base of the PanDA Pilot for implementing the PanDA Broker. It also
-shows that adopting third-party libraries like RADICAL-SAGA did not have a
-measurable adverse effect on reliability.
+reliable as PanDA Pilot when used on the WLCG\@. Between Jan 2016 and Feb
+2017, the overall failure rate of all the ATLAS detector simulation jobs was
+14\%, while the failure rate of jobs submitted to Titan was a comparable
+13.6\%. PanDA Brokers were responsible for around the 19\% of the failures,
+compared to the 29\% of failures produced by the JobDispatcher module of the
+PanDA Server, and the 13\% failures produced by the Geant4 toolkit. 
+% The current failure rate of the PanDA Brokers confirms the benefits of
+% reusing most of the code base of the PanDA Pilot for implementing the PanDA
+% Broker. It also shows that adopting third-party libraries like RADICAL-SAGA
+% did not have a measurable adverse effect on reliability.
 
-% -----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
 \subsection{Characterizing the Detector Simulation on Titan}
@@
 We use two main parameters to measure the performance of the detector
-simulation jobs submitted to Titan: (i) the time taken to setup AthenaMP; and
-(ii) the distribution of the time taken by the Geant4 toolkit to simulate a
-certain number of events.
+simulation jobs submitted to Titan: (i) the time taken to setup AthenaMP\@;
+and (ii) the distribution of the time taken by the Geant4 toolkit to simulate
+a certain number of events.
 
 AthenaMP has an initialization and configuration stage. At initialization
-time, AthenaMP is assembled from a large number of shared libraries,
-depending on the type of payload that will have to be computed. Once
-initialized, every algorithm and service of AthenaMP is configured by a set
-of Python scripts. Both these operations result in a large number of read
-operations on the filesystem shared between the worker nodes and the DTNs,
-including the operations required to access small python scripts.
+time, AthenaMP is assembled from a set of shared libraries, depending on the
+type of payload that will have to be computed. Once initialized, every
+algorithm and service of AthenaMP is configured via Python scripts. Both
+these operations result in read operations on the filesystem shared between
+the worker nodes and the DTNs, including the operations required to access
+small python scripts.
 
-Initially, all the shared libraries of AthenaMP and the python scripts for
-its configuration were stored on the Spider 2 Lustre file system. However,
-the I/O patterns of the initialization and configuration stages degraded the
-performance of the filesystem. This was addressed by moving the AthenaMP
-distribution to a read-only NFS directory, shared among DTNs and worker
-nodes. NFS eliminated the problem of metadata contention, improving metadata
-read performance from $\approx$6,300 seconds on Lustre to $\approx$1,500
-seconds on NFS.
+Initially, all the shared libraries and the python scripts of AthenaMP were
+stored on the OLCF Spider 2 Lustre file system. However, the I/O patterns of
+the initialization and configuration stages degraded the performance of the
+filesystem. This was addressed by moving the AthenaMP distribution to a
+read-only NFS directory, shared among Titan's DTNs and worker nodes. NFS
+eliminated the problem of metadata contention, improving metadata read
+performance from $\approx$6,300 seconds on Lustre to $\approx$1,500 seconds
+on NFS\@.
 
@@
     \caption{Distributions of the time taken to simulate one event when
-    placing 2 simulations (h1) or 1 simulation (h2) per Titan's CPU. 2
+    placing 2 simulations (h1) or 1 simulation (h2) per Titan's CPU\@. 2
     simulation use 16 cores per node, 1 simulation 8. 50 Events; 1 Titan
@@
 CPUs explain the higher performance variance we observed compared to the
-performance consistency we measured on Titan.
+performance consistency we measured on Titan. This difference in performance
+is compensated by the amount of resources available on Titan (capable of
+executing up to 30\%/year of the ATLAS detector simulations) and by the lack
+of further resources available to WLCG\@. Also, it should be noted that Titan
+is at the end of its life-cycle and that Summit, Titan's successor, will
+offer cutting-edge performances.
 
@@
 Lustre is 1,320, comparable to the time taken to read the file required by
-assembling AthenaMP from NFS. Preliminary investigation shows that this time
-could be reduced to 40 seconds by loading the event descriptions into the RAM
-disk available on each worker node. Events descriptions could be transferred
-from Lustre to the RAM disk while configuring and initializing AthenaMP,
-almost halving the time currently required by initiating a Genat4 simulation.
+assembling AthenaMP from NFS\@. Preliminary investigation shows that this
+time could be reduced to 40 seconds by loading the event descriptions into
+the RAM disk available on each worker node. Events descriptions could be
+transferred from Lustre to the RAM disk while configuring and initializing
+AthenaMP, almost halving the time currently required by initiating a Geant4
+simulation.

### 2017-07-01 Experiments.tex
Down to 10 pages

@@
+Comments to address in this section:
+\begin{enumerate}
+    \item REVIEWER 2: In the weak scalability evaluations (in Figure 7 and
+    8), the overhead of the Next Generation Executor (NGE) has grown linearly
+    with the number of the nodes. Can you explain why the average time
+    execution time of AthenaMP is decreased even when the overhead is
+    increasing?
+\end{enumerate}
+
 \subsection{Experiments}
@@
 We designed experiments to characterize the performance of the NGE on Titan,
-with an emphasis on understanding its overhead and thus the cost of introducing
-new functionalities.  We perform three groups of experiments in which we
-investigate the weak scalability, weak scalability with multiple generation, and
-strong scalability of the NGE.
+with an emphasis on understanding its overhead and thus the cost of
+introducing new functionalities. We perform three groups of experiments in
+which we investigate the weak scalability, weak scalability with multiple
+generation, and strong scalability of the NGE.
 
-Each experiment entails executing multiple instances of AthenaMP 
-to simulate a pre-determined number of events. All the experiments have been
-performed by  configuring AthenaMP to use all the 16 cores  of Titan's worker
-nodes.
+Each experiment entails executing multiple instances of AthenaMP to simulate
+a pre-determined number of events. All the experiments have been performed by
+configuring AthenaMP to use all the 16 cores of Titan's worker nodes.
 
-We  measured the execution time of the pilots and of AthenaMP 
-within them, collecting timestamps at  all stages of the execution. Experiments
-were performed  by  submitting NGE's pilots  to Titan's batch queue.  The
-turnaround time of an individual run is determined by queue waiting times. Since
-we are interested only in the performances of the NGE, we removed queue time
-from our statistics.
+We measured the execution time of the pilots and of AthenaMP within them,
+collecting timestamps at all stages of the execution. Experiments were
+performed by submitting NGE's pilots to Titan's batch queue. The turnaround
+time of an individual run is determined by queue waiting times. Since we are
+interested only in the performances of the NGE, we removed queue time from
+our statistics.
 
@@
 
-In this experiment  we run as many AthenaMP instances (hereafter referred to as
-tasks)  as the number of nodes controlled by the pilot. Each AthenaMP simulates
-100 events, requiring $\sim 4200$ seconds on average.
+In this experiment we run as many AthenaMP instances (hereafter referred to
+as tasks) as the number of nodes controlled by the pilot. Each AthenaMP
+simulates 100 events, requiring $\sim 4200$ seconds on average.
 
@@
 Fig.~\ref{fig:weakScal1a} depicts the average pilot duration, the average
-execution time of AthenaMP, and the NGE pilot overhead as function of the pilot
-size.
+execution time of AthenaMP, and the NGE pilot overhead as function of the
+pilot size.
 
@@
 We observe that, despite some fluctuations due to external factors (e.g.,
-Titan's shared filesystem and the shared database used by the NGE), the average
-execution time of AthenaMP  ranges between 4200 and 4800 seconds.  We  also
-observe that in all the cases the gap between AthenaMP execution times and the
-pilot durations is minimal, although it slightly increases with the pilot size.
-We  notice that NGE's overhead does grow linearly with the number of units.
+Titan's shared filesystem and the shared database used by the NGE), the
+average execution time of AthenaMP ranges between 4200 and 4800 seconds. We
+also observe that in all the cases the gap between AthenaMP execution times
+and the pilot durations is minimal, although it slightly increases with the
+pilot size. We notice that NGE's overhead does grow linearly with the number
+of units.
 
@@
 
-The NGE provides an important new capability of submitting multiple generations
-of AthenaMP to the same pilot. In order to investigate the cost of doing so, we
-performed a variant of the weak scalability experiments. This stresses the
-pilot's components, as new tasks are scheduled for execution on the Agent while
-other tasks are still running.
+The NGE provides an important new capability of submitting multiple
+generations of AthenaMP to the same pilot. In order to investigate the cost
+of doing so, we performed a variant of the weak scalability experiments. This
+stresses the pilot's components, as new tasks are scheduled for execution on
+the Agent while other tasks are still running.
 
 In these experiments, we run five AthenaMP instances per node. As these
-experiments are designed to investigate the overhead generated by the scheduling
-and bootstrap of AthenaMP instances, we reduced the number of events simulated
-by each AthenaMP task to sixteen in such a way that the running time of each
-AthenaMP is, on average, $\sim 1200$ seconds. This experiment design choice does
-not affect the  objectives or accuracy of the experiments, but allows us to
-scale experiments to large node counts by being conservative with allocation.
+experiments are designed to investigate the overhead generated by the
+scheduling and bootstrap of AthenaMP instances, we reduced the number of
+events simulated by each AthenaMP task to sixteen in such a way that the
+running time of each AthenaMP is, on average, $\sim 1200$ seconds. This
+experiment design choice does not affect the objectives or accuracy of the
+experiments, but allows us to scale experiments to large node counts by being
+conservative with allocation.
 
-We ran pilots with 256, 512, 1024 and 2048 worker nodes and 3\mtnote{2?} hours
-walltime. Fig.~\ref{fig:weakScal2a} depicts the average pilot duration, the
-average execution time of five sequential generations of AthenaMP, and the
-corresponding overhead. The difference between the two durations
-is more marked than in the previous experiments. Despite this, we notice that
-the growth of the overhead is consistent with the increment of the number of
+We ran pilots with 256, 512, 1024 and 2048 worker nodes and 3\mtnote{2?}
+hours walltime. Fig.~\ref{fig:weakScal2a} depicts the average pilot duration,
+the average execution time of five sequential generations of AthenaMP, and
+the corresponding overhead. The difference between the two durations is more
+marked than in the previous experiments. Despite this, we notice that the
+growth of the overhead is consistent with the increment of the number of
 tasks per node for pilots with 256, 512 and 1024 worker nodes, and less than
@@
 
-We investigate strong scalability by running the same number of
-tasks for different pilot sizes. We used 2048 AthenaMP instances and  pilots
-with 256, 512, 1024 and 2048 nodes. Thus, the number of AthenaMP generations is
-equal to eight times the size of the smallest pilot and corresponds to the size
-of the largest pilot. As a consequence, the number of consecutive generations of
-AthenaMP decreases with the pilot size by generating different dynamics within
-the pilots. These experiments are designed to investigate whether pilot overhead
-is affected by the degree of concurrency within the pilot and/or the number of
-tasks. Each AthenaMP instance simulates sixteen events as in the previous
-experiment.
+We investigate strong scalability by running the same number of tasks for
+different pilot sizes. We used 2048 AthenaMP instances and pilots with 256,
+512, 1024 and 2048 nodes. Thus, the number of AthenaMP generations is equal
+to eight times the size of the smallest pilot and corresponds to the size of
+the largest pilot. As a consequence, the number of consecutive generations of
+AthenaMP decreases with the pilot size by generating different dynamics
+within the pilots. These experiments are designed to investigate whether
+pilot overhead is affected by the degree of concurrency within the pilot
+and/or the number of tasks. Each AthenaMP instance simulates sixteen events
+as in the previous experiment.
 
@@
 execution time of possibly sequential AthenaMP instances. We notice that the
-difference between the pilot duration and the AthenaMP execution times is almost
-constant for all the pilot sizes, although the overall duration of the pilot
-decreases linearly with the pilot size.
+difference between the pilot duration and the AthenaMP execution times is
+almost constant for all the pilot sizes, although the overall duration of the
+pilot decreases linearly with the pilot size.
 
@@
         \vspace{-0.3in}
-    \caption{Strong scalability:  average pilot duration, average duration of
+    \caption{Strong scalability: Average pilot duration, average duration of
     sequential AthenaMP executions, and pilot's overhead for pilots with 256, 512, 1024 and 2048 nodes.}

### 2017-07-01 nge.tex
Down to 10 pages

--- 2017-04-04 nge.tex
@@
-As seen in \S\ref{ssec:panda_titan}, PanDA Broker was  deployed on the DTNs
-as Titan's worker nodes lack connectivity to the wide area network The lack of
-pilot capabilities impacts both the efficiency and the flexibility of PanDA's
-execution process. Pilots could improve efficiency by increasing throughput and
-enabling greater backfill utilization. Further, pilots makes it easier to
-support heterogeneous workloads.
+% As seen in \S\ref{ssec:panda_titan}, PanDA Broker is deployed on the DTNs
+% as Titan's worker nodes lack connectivity to the wide area network.
+The lack of pilot capabilities in PanDA Broker impacts both the efficiency
+and the flexibility of PanDA's execution process. Pilots could improve
+efficiency by increasing throughput and enabling greater backfill
+utilization. Further, pilots would make it easier to support heterogeneous
+workloads.
 
-The absence of pilots imposes the static coupling between MPI scripts
-submitted to the PBS batch system and detector simulations
-(\S\ref{sec:panda_titan}). This makes the scheduling of multiple generations
-of workload on the same PBS job impossible: once a statically defined number
-of detector simulations are packaged into a PBS job and this job is queued on
-Titan, no further simulations can be added to that job. New simulations have
-to be packaged into a new PBS job that needs to be submitted to Titan based
-upon backfill availability.
+The absence of pilots 
+% imposes the static coupling between MPI scripts submitted to the PBS batch
+% system and detector simulations (\S\ref{sec:panda_titan}). This
+makes the scheduling of multiple generations of workload on the same PBS job
+impossible: once a statically defined number of detector simulations are
+packaged into a PBS job and this job is queued on Titan, no further
+simulations can be added to that job. New simulations have to be packaged
+into a new PBS job that needs to be submitted to Titan based upon backfill
+and PanDA Brokers availability.
 
-The support of  multiple generations of workload would enable more efficient use
-of the backfill availability of walltime. Currently, when a set of simulations
-ends, the PBS job also ends, independent of whether more wall-time would still
-be available. With a pilot, additional simulations could be executed  to utilize
-all the available wall-time, while avoiding further job packaging and submission
-overheads.
+The support of multiple generations of workload would enable more efficient
+use of the backfill availability walltime. Currently, when a set of
+simulations ends, the PBS job also ends, independent of whether more
+wall-time would still be available. With a pilot, additional simulations
+could be executed to utilize all the available wall-time, while avoiding
+further job packaging and submission overheads.
 
-Multiple generations would also relax two assumptions of the current execution
-model: knowing the number of simulations before submitting the MPI script, and
-having a fixed number of events per simulation (100 at the moment). Pilots would
-enable the scheduling of simulations independently from whether they were
-available at the moment of submitting the pilot. Further, simulations with a
-varying number of events could be scheduled on a pilot, depending on the amount
-of remaining walltime and the distribution of execution time per event, as shown
-in \S\ref{ssec:panda_titan}, Fig.~\ref{fig:comparison-8-16cores}. These
-capabilities would increase the efficiency of the PanDA Broker when there is a
-large difference between the number of cores and walltime.
+Multiple generations would also relax two assumptions of the current
+execution model: knowing the number of simulations before submitting the MPI
+script, and having a fixed number of events per simulation (100 at the
+moment). Pilots would enable the scheduling of simulations independently from
+whether they were available at the moment of submitting the pilot. Further,
+simulations with a varying number of events could be scheduled on a pilot,
+depending on the amount of remaining walltime and the distribution of
+execution time per event, as shown in \S\ref{ssec:panda_titan},
+Fig.~\ref{fig:comparison-8-16cores}. These capabilities would increase the
+efficiency of the PanDA Broker when there is a large difference between the
+number of cores and walltime.
 
@@
 develop a specific PBS script for each payload. The submission process would
-also be MPI-independent, as MPI is used for coordination among multiple worker
-nodes, not by the payload.
+also be MPI-independent, as MPI is used for coordination among multiple
+worker nodes, not by the payload.
 
-% -----------------------------------------------------------------------------
+% ---------------------------------------------------------------------------
 \subsection{Implementation}
@@
 
-The implementation of pilot capabilities within the current PanDA Broker require
-quantification of the effective benefits that it could yield and, on the base of
-this analysis, a dedicated engineering effort. We developed a prototype of a
-pilot system capable of executing on Titan to study experimentally the
-quantitative and qualitative benefits that it could bring to PanDA. We called
-this prototype Next Generation Executor (NGE).
+The implementation of pilot capabilities within the PanDA Broker require
+quantification of the effective benefits that it could yield and, on the base
+of this analysis, a dedicated engineering effort. We developed a prototype of
+a pilot system capable of executing on Titan to study experimentally the
+quantitative and qualitative benefits that it could bring to PanDA\@. We
+called this prototype Next Generation Executor (NGE).
 
@@
 module is the pilot submitted to Titan and executed on its worker nodes. The
-communication between PanDA Broker and Server is abstracted away as it
-is not immediately useful to evaluate the performance and capabilities of a
-pilot on Titan.
+communication between PanDA Broker and Server is abstracted away as it is not
+immediately useful to evaluate the performance and capabilities of a pilot on
+Titan.
 
@@
    \vspace{-0.3in}
-  \caption{NGE Architecture: The PilotManager and  UnitManager reside on a Titan
-  DTN while the Agent is executed on its worker nodes. Color coding: gray for
-  entities external to NGE, white for APIs, purple for NGE's modules, green for
-  pilots, yellow for module's components.}
+  \caption{NGE Architecture: The PilotManager and UnitManager reside on a DTN
+  while the Agent is executed on a worker node. Color coding: gray for
+  entities external to NGE\@; white for APIs; purple for NGE's modules; green
+  for pilots; yellow for module's components.}
 \label{fig:arch-overview}
@@
 squares) and pilots (Fig.~\ref{fig:arch-overview}, red circles), and to
-instantiate a PilotManager and a UnitManager.   The PilotManager submits
-pilots to Titan's PBS  batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit MPI scripts. Once
-scheduled, the pilot's Agent is bootstrapped on Titan's MOM node and
-the Agent's Executors on the on the worker
-nodes, and the UnitManager schedules units to the Agent's Scheduler
-(Fig.~\ref{fig:arch-overview}, solid arrow). The Agent's Scheduler schedules
-the units on one or more Agent's Executor for execution. The Agent's executors
-can manage one or more worker nodes, depending on performance evaluations. The
-UnitManager and the Agent communicate via a database that is instantiated on
-one of Titan's DTN so as to be reachable by both modules.
+instantiate a PilotManager and a UnitManager. The PilotManager submits pilots
+to Titan's PBS batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash
+arrow). %, as done by PanDA Broker to submit MPI scripts. 
+Once scheduled, the pilot's Agent is bootstrapped on a MOM node and the
+Agent's Executors on the worker nodes.
 
-The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
-communication and coordination of the execution of units. This environment is
-a critical component of the OpenMPI implementation, developed to support
-distributed high-performance computing applications operating in a
-heterogeneous environment~\cite{castain05:_open_rte, cug-2016}. ORTE provides
-a mechanism to create a \emph{``dynamic virtual machine''} (DVM) that spans
-multiple nodes. The DVM transparently provides support for interprocess
-communication, resource discovery and allocation, and process launching across
-a variety of platforms. Libraries enable the interaction between users and
-ORTE to enable the submission, monitoring and managing of tasks, avoiding
-filesystem bottlenecks and race conditions with network sockets. As a
-consequence, ORTE is able to minimize the system overhead while submitting
-tasks.
+The UnitManager and the Agent communicate via a database instantiated on a
+DTN so as to be reachable by both modules. The UnitManager schedules units to
+the Agent's Scheduler (Fig.~\ref{fig:arch-overview}, solid arrow) and the
+Agent's Scheduler schedules the units on one or more Agent's Executor.
+% for execution. The Agent's executors can manage one or more worker nodes,
+% depending on performance evaluations.
+
+The NGE Agent uses the Open Run-Time Environment (ORTE) for communication and
+coordination of the execution of units. This environment is a critical
+component of the OpenMPI implementation~\cite{castain05:_open_rte}.
+%, cug-2016}.
+ORTE is able to minimize the system overhead while submitting tasks by
+avoiding filesystem bottlenecks and race conditions with network sockets.

### 2017-07-04 Experiments.tex
Small changes in the text

@@
     increasing?
+    I checked the data.
+Figure 8 was simply wrong.
+I think that we have overwritten the original with an older version.
+Figure 7: I used a smaller integration step to compute the integral of the pilot concurrency.
+It gets better but not as much as I would like.
+Our data has something sketchy.
 \end{enumerate}
@@
 Titan's shared filesystem and the shared database used by the NGE), the
-average execution time of AthenaMP ranges between 4200 and 4800 seconds. We
+average execution time of AthenaMP ranges between 4500 and 4800 seconds. We
 also observe that in all the cases the gap between AthenaMP execution times

### 2017-07-04 overall.tex
Address referee's comment based on Sergey's e-mail

@@
     % illustrates.
-    \item REVIEWER 3: It’s interesting that Feb/March have such an increase
-    in backfill availability in 2017. The authors attribute this to hardware
-    upgrades for the DTNs. It’s mentioned that the load is only 3.6\% now and
-    therefore further upgrades will likely not cause similar improvement.
-    What was the previous CPU load? Why were particularly low percentages
-    utilized in April/May 2016?
+    \item 
+    % REVIEWER 3: It’s interesting that Feb/March have such an increase in
+    % backfill availability in 2017. The authors attribute this to hardware
+    % upgrades for the DTNs. It’s mentioned that the load is only 3.6\% now
+    % and therefore further upgrades will likely not cause similar
+    % improvement. What was the previous CPU load? 
+    Why were particularly low percentages utilized in April/May 2016?
     % \item REVIEWER 4: “Genat4” at the end of Section 4 is a typo.
@@
 improved performance. Investigations showed an average CPU load of 3.6\% on
-the upgraded DTNs. As such, further hardware upgrades seem unlikely to
+the upgraded DTNs, as opposed to the ``high'' utilization reported by OLCF
+for the previous DTNs. As such, further hardware upgrades seem unlikely to
 improve significantly the performance of PanDA Brokers.\mtnote{Asked Segey

### 2017-07-04 overall.tex
minor fixes to language, space saving etc.

@@
 to 20 markedly improved efficiency (see Fig.~\ref{fig:backfill-utilization})
-but further research is undergoing to understand whether an even larger
-number of brokers would yield similar results. This research focuses on
+but further research is ongoing to understand whether an even greater
+number of brokers would yield even greater efficiency. This research focuses on
 evaluating the overheads of input/output files staging, including its impact
@@
 concurrent submission of multiple MPI scripts~\cite{barreiro2016panda}. The
-results of these investigation will contribute to improve the efficienty of
+understanding will contribute to improving the efficiency of
 PanDA Brokers beyond the 30\% limit showed in
@@
 simulation jobs submitted to Titan: (i) the time taken to setup AthenaMP\@;
-and (ii) the distribution of the time taken by the Geant4 toolkit to simulate
+and (ii) the distribution of the time taken by Geant4 to simulate
 a certain number of events.

### 2017-07-04 introduction.tex
reworking introduction.

@@
 	the paper in a prominent location.
-	\item REVIEWER 1: A glossary would be welcomed.
 	\item REVIEWER 5: I would have preferred to see a more detailed list of
@@
 	to how to port high-throughput apps to high-performance clusters. Agreed?
+	{\bf  Need to clarify it is not the task that is being converged but the workload and execution model that is being converged}
+	\item REVIEWER 1: A glossary would be welcomed. {\bf Not addressable}
 \end{enumerate}
@@
 centers. Thousands of physicists analyze petascale data volumes daily. The
-detection of the Higgs Boson in 2013 speaks to the success of the detector
-and experiment design, as well as the sophistication of computing systems
-devised to analyze the data.
+detection of the Higgs Boson in 2013 speaks to the success of the detector and
+experiment design, as well as the sophistication of computing systems devised
+to analyze the data, which historically, consisted of the federation of
+hundreds to thousands of distributed resources, ranging in
+scale from small to mid-size resource~\cite{foster2003grid}.
 
-Historically, the computing systems used by LHC experiments consisted of the
-federation of hundreds to thousands of distributed resources, %\textemdash{}
-ranging in scale from small to mid-size resource~\cite{foster2003grid}.
 Although the workloads to be executed are comprised of tasks that are
@@
 across many heterogeneous resources, the effective utilization of resources
-and efficient execution of workloads presents non-trivial challenges.
-
-Many software solutions have been developed in response to these challenges.
-CMS, one of the LHC experiments, devised a solution based around the
+and efficient execution of workloads present non-trivial challenges. Many
+software solutions have been developed in response to these challenges. The
+CMS experiment, devised a solution based around the
 HTCondor~\cite{thain2005distributed} software ecosystem. The
-ATLAS~\cite{Aad:2008} experiment, utilizes the Production and Distributed
+ATLAS~\cite{Aad:2008} experiment utilizes the Production and Distributed
 Analysis (PanDA) workload management system~\cite{Maeno2011} (WMS) for
 distributed data processing and analysis. The CMS and ATLAS experiments
-represent arguably the largest production grade distributed computing
-solutions and have symbolized the paradigm of {\it high-throughput
-computing}, i.e., the effective execution of many independent tasks.
+utilize, arguably the largest production grade distributed computing
+solutions and have symbolized the paradigm of {\it high-throughput computing},
+i.e., the effective execution of many independent tasks.
 
@@
 demand for computing systems will soon significantly outstrip current and
-projected supply.   The data volumes that will need analyzing in LHC-Run 3 in
-$~\approx$ 2022 and the high-luminosity era (Run 4) will increase by factors
+projected supply.   The data volumes that will need analyzing in LHC-Run 3 
+($~\approx$ 2022) and the high-luminosity era (Run 4) will increase by factors
 of 10-100 compared to the current phase (Run 2). There are multiple levels at
@@
 high-performance computers. Specifically, we discuss the experience of integrating
-PanDA WMS with a US DOE leadership computing facility called Titan to reach
-sustained production scales of approximately 51M core-hours a year. We also
-discusses the investigation of a pilot-abstraction based task execution
-runtime system on Titan, which allows advanced execution modes of the ATLAS
-workload as well as the enhanced support for heterogeneous workloads, e.g.,
-molecular dynamics.
+PanDA WMS with a US DOE leadership machine (Titan) to reach
+sustained production scales of approximately 51M core-hours a year.
 
-% . Consequently, Titan now analyzes up to 3.6\% of the simulation workload
-% of the ATLAS experiment. The case study presents the design and integration
-% of PanDA
+We demonstrate how Titan is more efficient utilized by the mixing of small and
+short-lived tasks in backfill with regular payloads. Cycles otherwise unusable
+(or very difficult to use) are used for science thus, increasing the overall
+utilization on Titan without loss of quality-of-service. The conventional mix
+of jobs at OLCF cannot be effectively backfilled because of size, duration,
+and scheduling policies. Our approach is extensible to any HPC with
+"capability scheduling" policies.  We also investigate a pilot-abstraction
+based task execution runtime system which allows advanced execution modes of
+the ATLAS workload, as well as heterogeneous workloads, e.g., molecular
+dynamics. As such, our work is an exemplary use case for general approaches to
+HPC and HTC integration.
 
-This experience paper provides three main contributions:  (i) a critical
-evaluation of the many design and operational considerations that have been
-taken to support the sustained, scalable and production usage of Titan for
-historically high-throughput workloads; (ii) a preliminary characterization
-of a next generation executor (NGE) for PanDA to support non-traditional
-heterogeneous workloads and execution modes;  and (iii) early lessons and
-guidance as the community looks forward to designing the next generation of
-online analytical platforms~\cite{foap-url}, for how current and future
-experimental and observational systems can be integrated with production
-supercomputers in a general and extensible manner.
+This work demonstrates a viable production route to delivering large amounts
+of computing resources to ATLAS and, in the future, to other experimental and
+observational use cases.  This broadens the use of leadership computing while
+demonstrating how distributed workflows can be integrated with leadership
+resources, and thus how HTC and HPC workloads can be effectively accommodated
+simultaneously.
+
+This paper also provides: (i) a critical evaluation of the many design and
+operational considerations that have been taken to support the sustained,
+scalable and production usage of Titan for historically high-throughput
+workloads, and (ii) early lessons and guidance as the community looks forward
+to designing the next generation of online analytical platforms~\cite{foap-
+url}, for how current and future experimental and observational systems can be
+integrated with production supercomputers in a general and extensible manner.

### 2017-07-04 introduction.tex
addressing more reviewer comments.

@@
 
-We demonstrate how Titan is more efficient utilized by the mixing of small and
-short-lived tasks in backfill with regular payloads. Cycles otherwise unusable
-(or very difficult to use) are used for science thus, increasing the overall
-utilization on Titan without loss of quality-of-service. The conventional mix
-of jobs at OLCF cannot be effectively backfilled because of size, duration,
-and scheduling policies. Our approach is extensible to any HPC with
-"capability scheduling" policies.  We also investigate a pilot-abstraction
-based task execution runtime system which allows advanced execution modes of
-the ATLAS workload, as well as heterogeneous workloads, e.g., molecular
-dynamics. As such, our work is an exemplary use case for general approaches to
-HPC and HTC integration.
+We demonstrate how Titan is more efficiently utilized by the mixing of small
+and short-lived tasks in backfill with regular payloads. Cycles otherwise
+unusable (or very difficult to use) are used for science thus, increasing the
+overall utilization on Titan without loss of quality-of-service. The
+conventional mix of jobs at OLCF cannot be effectively backfilled because of
+size, duration, and scheduling policies. Our approach is extensible to any HPC
+with "capability scheduling" policies.  We also investigate the use of a
+pilot-abstraction based task execution runtime system to flexibly execute
+ATLAS and other heterogeneous workloads (molecular dynamics) using regular
+queues. As such, our approach provides a general solution and investigation of
+the convergence of HPC and HTC execution of workloads.
 
@@
 demonstrating how distributed workflows can be integrated with leadership
-resources, and thus how HTC and HPC workloads can be effectively accommodated
-simultaneously.
+resources, and effectively accommodating HTC and HPC workloads simultaneously.
 

### 2017-07-04 Experiments.tex
minor fixes to language, space saving etc.
refinements  - wordsmithing and shortening.

@@
 Tasks do not wait within the NGE Agent's queue since one node is available to
-each AthenaMP instance. Overheads in task execution are consequence primarily
-of the three other factors: (i) the initial bootstrapping of the pilot on the
-nodes; (ii) the UnitManager's dispatching of units (tasks) to the agent; and
-(iii) time for the agent to bootstrap all the tasks on the nodes.
+each AthenaMP instance. Task execution  overheads result primarily from  three
+factors: (i) the initial bootstrapping of the pilot on the nodes; (ii) the
+UnitManager's dispatching of units (tasks) to the agent; and (iii) time for
+the agent to bootstrap all the tasks on the nodes.
 
@@
 In these experiments, we run five AthenaMP instances per node. As these
-experiments are designed to investigate the overhead generated by the
-scheduling and bootstrap of AthenaMP instances, we reduced the number of
-events simulated by each AthenaMP task to sixteen in such a way that the
-running time of each AthenaMP is, on average, $\sim 1200$ seconds. This
+experiments are designed to investigate the overhead of 
+scheduling and bootstraping of AthenaMP instances, the number of
+events simulated by each AthenaMP task was reduced to sixteen such that the
+running time of each AthenaMP was, on average, $\sim 1200$ seconds. This
 experiment design choice does not affect the objectives or accuracy of the

### 2017-07-04 bckground.tex
refinements  - wordsmithing and shortening.

@@
 
-AliEn, DIRAC, GlideinWMS and PanDA also have similar implementations. These WMS
-were initially implemented to use Grid resources, using one or more components
-to the Condor software ecosystem~\cite{thain2005distributed} and, as with
-GlideinWMS, contributing to its development. Accordingly, all LHC WMS
-implemented Grid-like authentication and authorization systems and adopted a
-computational model based on distributing a large amount of single/few-cores
-tasks across hundreds of sites\mtnote{Is this true?}.
+%~\cite{thain2005distributed}
 
-All the experiments at LHC produces and process large amount of data both from
-actual collisions in the accelerator and from their simulations. Dedicated,
-multi-tiered data systems have been built to store, replicate, and distributed
-these data. All LHC WMS interface with these systems to move data to the sites
-where related compute tasks are executed or to schedule compute tasks where
-(large amount of) data are already stored.
+AliEn, DIRAC, GlideinWMS and PanDA also have similar implementations. These
+WMS were initially implemented to use Grid resources using the Condor software
+ecosystem~\cite{thain2005distributed}. Accordingly, all LHC WMS implemented
+Grid-like authentication and authorization systems and adopted a computational
+model based on distributing a large amount of single/few-cores tasks across
+hundreds of sites\mtnote{Is this true?}.
+
+All LHC experiments produce and process large amounts  of data from actual
+collisions in the accelerator and from their simulations. Dedicated, multi-
+tiered data systems have been built to store, replicate, and distributed these
+data. All LHC WMS interface with these systems to move data to the sites where
+related compute tasks are executed or to schedule compute tasks where (large
+amount of) data are already stored.
 

### 2017-07-04 PandaOverview.tex
minor in Section II.

@@
-PanDA is a Workload Management System (WMS)%~\cite{marco2009glite} 
+PanDA is a Workload Management System (WMS) %~\cite{marco2009glite} 
 designed to support the execution of distributed workloads and workflows via
 pilots~\cite{turilli2015comprehensive}. Pilot-capable WMS enable high
-throughput of tasks execution via multi-level scheduling while supporting
+throughput execution of tasks via multi-level scheduling while supporting
 interoperability across multiple sites. This is particularly relevant for LHC
@@
 agent (Fig.~\ref{fig:architecture}:9). Condor-G schedules these pilots
-wrapped as jobs or VMs to the required sites
+wrapped as jobs to the required sites
 (Fig.~\ref{fig:architecture}:10).
@@
 
-Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
-subprocess for the execution of the job's payload. Input data are transferred
-from the stage-in location (Fig.~\ref{fig:architecture}:13) and the job's
-payload is executed (Fig.~\ref{fig:architecture}:14). Once completed, output
+A PanDA Pilot starts a monitoring process on receiving a job and forks a
+subprocess to execute the job's payload. Input data are transferred
+from the stage-in location (Fig.~\ref{fig:architecture}:13), the job's
+payload is executed (Fig.~\ref{fig:architecture}:14) and once completed, output
 is transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).

### 2017-07-04 PandaEvolution.tex
minor Section III-A

@@
 Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
-300,000 cores at over 100 Grid sites, with a peak performance of 0.3
+300,000 cores at more than 100 Grid sites, with a peak performance of 0.3
 petaFLOPS\@. This capacity will be sufficient for the planned analysis and
@@
 
-Generally, supercomputers are designed to support parallel computation that
-requires runtime communication. Jobs are executed across multiple cores, each
-core calculating a small part of the problem and communicating with other
-cores via MPI\@. Accordingly, supercomputers have large number of worker
+Supercomputers are designed to support parallel computation that
+requires runtime communication. Jobs are executed across multiple nodes, each
+node calculating a small part of the problem and communicating with other
+node, typically via MPI\@. Accordingly, supercomputers have large number of worker
 nodes, connected through a high-speed, low-latency dedicated network. Each
 worker node has multicore CPUs, usually augmented with Graphics Processing
-Units (GPUs) or other specialized coprocessors.
+Units (GPUs) or other specialized co-processors.
 
@@
 ATLAS workloads or workflows involves concurrent and/or sequential runs of
-possibly large amount of jobs, each requiring no or minimal parallelization
+possibly large number of jobs, each requiring minimal, if any parallelization
 and no runtime communication. Thus, computing infrastructure like WLCG have
@@
 Currently, ATLAS workloads do not require fast interconnects or specilized
-coprocessors but supercomputers tend not to reach 100\% utilization due to
+co-processors, but supercomputers tend not to reach 100\% utilization due to
 the scheduling of jobs requiring large amount of resources. This offers the
@@
 Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
-Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).
+Computing Facility (OLCF).
 
@@
 Titan's architecture, configuration and policies poses several challenges to
-the integration with PanDA\@. The default deployment model of PanDA Pilot is
+the deployment of PanDA\@. The default deployment model of PanDA Pilot is
 unfeasible on Titan: PanDA Pilot is required to contact the Job Dispatcher of
@@
 Titan does not support PanDA's security model based on certificates and
-virtual organizations, making PanDA's approach to identity management also
+virtual organizations, making PanDA's approach to identity management 
 unfeasible. While Titan's data transfer nodes (DTNs) offer wide area network
@@
 up to 16 concurrent payloads, one per each available core. Given these
-constraints and challenges, the type of task most suitable for execution at
-the moment on Titan is Monte Carlo detector simulation. This type of task is
-mostly computational-intensive, requiring less than 2GB of RAM at runtime and
-small input data. Detector simulation tasks in ATLAS are performed via
+constraints and challenges, the  Monte Carlo detector simulation task is most
+suitable for execution on Titan at the moment. This type of task is mostly
+computational-intensive, requiring less than 2GB of RAM at runtime and small
+input data. Detector simulation tasks in ATLAS are performed via
 AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
@@
 Detector simulation is part of the ATLAS production Monte Carlo (MC)
-workflow~\cite{rimoldi2006atlas}.%,de2013delphes,ritsch2014atlas}. 
+workflow~\cite{rimoldi2006atlas}. %,de2013delphes,ritsch2014atlas}. 
 The MC workflow consists of four main stages: event generation, detector
@@
 of particle four-momenta via different generators, e.g.,
-PYTHIA,%~\cite{sjostrand2006pythia}, 
-HERWIG,%~\cite{corcella2001herwig} 
+PYTHIA, %~\cite{sjostrand2006pythia}, 
+HERWIG, %~\cite{corcella2001herwig} 
 and many others. Geant4 simulates the ATLAS detector and the interaction
-between that detector and particles. Each interaction creates a so-called hit
+between the detector and particles. Each interaction creates a so-called hit
 and all hits are collected and passed on for digitalization, where hits are

### 2017-07-04 conclusion.tex
minor fixes to language, space saving etc.
minor Section III-A

@@
 showing how they support and enable distributed computing at this scale on
-Titan, a leadership-class HPC machine managed by OCLF at ORNL.
+Titan, a leadership-class HPC machine managed by OCLF.
 

### 2017-07-04 overall.tex
minor fixes Section IV.

@@
 
-Currently, ATLAS deploys 20 instances of the PanDA Broker on 4 Titan's DTNs,
-5 instances per DTN\@. Each broker submits and manages the execution of 15 to
-300 ATLAS jobs, one job for each Titan's worker node, and a theoretical
-maximum concurrent use of 96,000 cores. Since November 2015, PanDA Brokers
-have operated only in backfill mode, without a defined time allocation, and
-running at the lowest priority on Titan. Therefore, ATLAS contributed to
-increase the Titan's overall utilization without affecting other users.
+Currently 20 instances of the PanDA Broker are deployed on 4 Titan's DTNs,
+with 5 instances per each DTN\@. Each broker submits and manages the execution
+of 15 to 300 ATLAS jobs, one job for each Titan's worker node, and a
+theoretical maximum concurrent use of 96,000 cores. Since November 2015, PanDA
+Brokers have operated only in backfill mode, without a defined time
+allocation, and running at the lowest priority on Titan. Therefore, ATLAS
+contributed to an increase of Titan's overall utilization without affecting
+other users.
 

### 2017-07-04 Experiments.tex
scattered fixes in Section V.

@@
 and the pilot durations is minimal, although it slightly increases with the
-pilot size. We notice that NGE's overhead does grow linearly with the number
+pilot size. We notice that NGE's overhead grows linearly with the number
 of units.
 
-\subsubsection{Weak scalability with multiple generation }
+\subsubsection{Weak scalability with multiple generation}
 
 The NGE provides an important new capability of submitting multiple
-generations of AthenaMP to the same pilot. In order to investigate the cost
+generations of tasks to the same pilot. In order to investigate the cost
 of doing so, we performed a variant of the weak scalability experiments. This
@@
 In these experiments, we run five AthenaMP instances per node. As these
-experiments are designed to investigate the overhead of 
-scheduling and bootstraping of AthenaMP instances, the number of
-events simulated by each AthenaMP task was reduced to sixteen such that the
-running time of each AthenaMP was, on average, $\sim 1200$ seconds. This
-experiment design choice does not affect the objectives or accuracy of the
-experiments, but allows us to scale experiments to large node counts by being
-conservative with allocation.
+experiments are designed to investigate the overhead of  scheduling and
+bootstraping of AthenaMP instances, the number of events simulated by each
+AthenaMP task was reduced to sixteen such that the running time of each
+AthenaMP was $\sim 1200$ seconds on average. This experiment design choice
+does not affect the objectives or accuracy of the experiments, but allows us
+to scale experiments to large node counts by conserving project allocation.
 
@@
         \vspace{-0.3in}
-    \caption{Weak scalability with multiple generations: average pilot
+    \caption{Weak scalability with multiple generations (where each generation has approximately 1/6th the number of events compared to Fig. 7): average pilot
     duration, average duration of sequential AthenaMP executions, and

### 2017-07-04 nge.tex
scattered fixes in Section V.

@@
 
-Multiple generations would also relax two assumptions of the current
-execution model: knowing the number of simulations before submitting the MPI
-script, and having a fixed number of events per simulation (100 at the
-moment). Pilots would enable the scheduling of simulations independently from
-whether they were available at the moment of submitting the pilot. Further,
-simulations with a varying number of events could be scheduled on a pilot,
-depending on the amount of remaining walltime and the distribution of
-execution time per event, as shown in \S\ref{ssec:panda_titan},
-Fig.~\ref{fig:comparison-8-16cores}. These capabilities would increase the
-efficiency of the PanDA Broker when there is a large difference between the
-number of cores and walltime.
+Multiple generations would also relax two assumptions of the current execution
+model: knowing the number of simulations before submitting the MPI script, and
+having a fixed number of events per simulation (currently 100). Pilots would
+enable the scheduling of simulations independently from whether they were
+available at the moment of submitting the pilot. Further, simulations with a
+varying number of events could be scheduled on a pilot, depending on the
+amount of remaining walltime and the distribution of execution time per event,
+as shown in \S\ref{ssec:panda_titan}, Fig.~\ref{fig:comparison-8-16cores}.
+These capabilities would increase the efficiency of the PanDA Broker when
+there is a large difference between the number of cores and walltime.
 

### 2017-07-05 PandaEvolution.tex
Address Shantenu's comments

@@
 
-Supercomputers are designed to support parallel computation that
-requires runtime communication. Jobs are executed across multiple nodes, each
-node calculating a small part of the problem and communicating with other
-node, typically via MPI\@. Accordingly, supercomputers have large number of worker
-nodes, connected through a high-speed, low-latency dedicated network. Each
-worker node has multicore CPUs, usually augmented with Graphics Processing
-Units (GPUs) or other specialized co-processors.
+% Supercomputers are designed to support parallel computation that requires
+% runtime communication. Jobs are executed across multiple nodes, each node
+% calculating a small part of the problem and communicating with other node,
+% typically via MPI\@. Accordingly, supercomputers have large number of
+% worker nodes, connected through a high-speed, low-latency dedicated
+% network. Each worker node has multicore CPUs, usually augmented with
+% Graphics Processing Units (GPUs) or other specialized co-processors.
 
@@
 % Linux, a run time environment based on SUSE Linux Enterprise Server.
-
+%
 % Titan's users submit jobs to Titan's PBS scheduler by logging into login or
@@
 computationally-intensive and non memory-intensive tasks of ATLAS workflows
-can be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
+can be off-loaded from the Grid to Titan. Further, when HEP tasks can be
 partitioned into independent jobs, Titan worker nodes can be used to execute

### 2017-07-05 introduction.tex
Fix latex and broken reference

@@
 centers. Thousands of physicists analyze petascale data volumes daily. The
-detection of the Higgs Boson in 2013 speaks to the success of the detector and
-experiment design, as well as the sophistication of computing systems devised
-to analyze the data, which historically, consisted of the federation of
-hundreds to thousands of distributed resources, ranging in
-scale from small to mid-size resource~\cite{foster2003grid}.
+detection of the Higgs Boson in 2013 speaks to the success of the detector
+and experiment design, as well as the sophistication of computing systems
+devised to analyze the data, which historically, consisted of the federation
+of hundreds to thousands of distributed resources, ranging in scale from
+small to mid-size resource~\cite{foster2003grid}.
 
@@
 utilize, arguably the largest production grade distributed computing
-solutions and have symbolized the paradigm of {\it high-throughput computing},
-i.e., the effective execution of many independent tasks.
+solutions and have symbolized the paradigm of {\it high-throughput
+computing}, i.e., the effective execution of many independent tasks.
 
@@
 projected supply.   The data volumes that will need analyzing in LHC-Run 3
-($~\approx$ 2022) and the high-luminosity era (Run 4) will increase by factors
-of 10-100 compared to the current phase (Run 2). There are multiple levels at
-which this problem needs to be addressed: the utilization of emerging parallel
-architectures (e.g., platforms); algorithmic and advances in analytical
-methods (e.g., use of Machine Learning); and the ability to exploit different
-platforms (e.g., clouds and supercomputers).
+($~\approx$ 2022) and the high-luminosity era (Run 4) will increase by
+factors of 10--100 compared to the current phase (Run 2). There are multiple
+levels at which this problem needs to be addressed: the utilization of
+emerging parallel architectures (e.g., platforms); algorithmic and advances
+in analytical methods (e.g., use of Machine Learning); and the ability to
+exploit different platforms (e.g., clouds and supercomputers).
 
-% As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity era
-% (Run 4), it is anticipated that the data volumes that will need analyzing
-% will increase by factors of 10-100 compared to the current phase (Run 2).
-% Data will be larger in volume but will also require more sophisticated
-% computational processing.
+% As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity
+% era (Run 4), it is anticipated that the data volumes that will need
+% analyzing will increase by factors of 10-100 compared to the current phase
+% (Run 2). Data will be larger in volume but will also require more
+% sophisticated computational processing.
 
@@
 on distributed resources to embrace new platforms, in particular
-high-performance computers. Specifically, we discuss the experience of integrating
-PanDA WMS with a US DOE leadership machine (Titan) to reach
+high-performance computers. Specifically, we discuss the experience of
+integrating PanDA WMS with a US DOE leadership machine (Titan) to reach
 sustained production scales of approximately 51M core-hours a year.
@@
 conventional mix of jobs at OLCF cannot be effectively backfilled because of
-size, duration, and scheduling policies. Our approach is extensible to any HPC
-with "capability scheduling" policies.  We also investigate the use of a
-pilot-abstraction based task execution runtime system to flexibly execute
+size, duration, and scheduling policies. Our approach is extensible to any
+HPC with ``capability scheduling'' policies.  We also investigate the use of
+a pilot-abstraction based task execution runtime system to flexibly execute
 ATLAS and other heterogeneous workloads (molecular dynamics) using regular
-queues. As such, our approach provides a general solution and investigation of
-the convergence of HPC and HTC execution of workloads.
+queues. As such, our approach provides a general solution and investigation
+of the convergence of HPC and HTC execution of workloads.
 
@@
 workloads, and (ii) early lessons and guidance as the community looks forward
-to designing the next generation of online analytical platforms~\cite{foap-
-url}, for how current and future experimental and observational systems can be
-integrated with production supercomputers in a general and extensible manner.
+to designing the next generation of online analytical
+platforms~\cite{foap-url}, for how current and future experimental and
+observational systems can be integrated with production supercomputers in a
+general and extensible manner.

### 2017-07-05 Experiments.tex
minor fix.
Changed Agent in Pilot Agent and removed NGE's pilot

@@
 collecting timestamps at all stages of the execution. Experiments were
-performed by submitting NGE's pilots to Titan's batch queue. The turnaround
+performed by submitting pilots to Titan's batch queue. The turnaround
 time of an individual run is determined by queue waiting times. Since we are
@@
 
-Tasks do not wait within the NGE Agent's queue since one node is available to
+Tasks do not wait within the Agent's queue since one node is available to
 each AthenaMP instance. Task execution  overheads result primarily from  three
@@
 Fig.~\ref{fig:weakScal1a} depicts the average pilot duration, the average
-execution time of AthenaMP, and the NGE pilot overhead as function of the
+execution time of AthenaMP, and the pilot overhead as function of the
 pilot size.
@@
 different pilot sizes. We used 2048 AthenaMP instances and pilots with 256,
-512, 1024 and 2048 nodes. Thus, the number of AthenaMP generations is equal
-to eight times the size of the smallest pilot and corresponds to the size of
-the largest pilot. As a consequence, the number of consecutive generations of
-AthenaMP decreases with the pilot size by generating different dynamics
-within the pilots. These experiments are designed to investigate whether
-pilot overhead is affected by the degree of concurrency within the pilot
-and/or the number of tasks. Each AthenaMP instance simulates sixteen events
-as in the previous experiment.
+512, 1024 and 2048 nodes. Thus, the number of AthenaMP generations is equal to
+eight times the size of the smallest pilot and corresponds to the size of the
+largest pilot. As a consequence, the number of consecutive generations of
+AthenaMP decreases with the pilot size. These experiments are designed to
+investigate whether pilot overhead is affected by the degree of concurrency
+within the pilot and/or the number of tasks. Each AthenaMP instance simulates
+sixteen events as in the previous experiment.
 

### 2017-07-05 nge.tex
Changed Agent in Pilot Agent and removed NGE's pilot
Forgot to save nge.tex

@@
   \caption{NGE Architecture: The PilotManager and UnitManager reside on a DTN
-  while the Agent is executed on a worker node. Color coding: gray for
+  while the Pilot Agent is executed on a worker node. Color coding: gray for
   entities external to NGE\@; white for APIs; purple for NGE's modules; green
@@
 arrow). %, as done by PanDA Broker to submit MPI scripts. 
-Once scheduled, the pilot's Agent is bootstrapped on a MOM node and the
+Once scheduled, the Pilot Agent is bootstrapped on a MOM node and the
 Agent's Executors on the worker nodes.
 
-The UnitManager and the Agent communicate via a database instantiated on a
+The UnitManager and the Pilot Agent communicate via a database instantiated on a
 DTN so as to be reachable by both modules. The UnitManager schedules units to
@@
 
-The NGE Agent uses the Open Run-Time Environment (ORTE) for communication and
+The Pilot Agent uses the Open Run-Time Environment (ORTE) for communication and
 coordination of the execution of units. This environment is a critical

### 2017-07-07 paper.tex
Somewhat drastic space reduction
Jack's affiliation.

@@
         Sarp H. Oral\IEEEauthorrefmark{6},
-        Jack C. Wells\IEEEauthorrefmark{6} and
+        Jack C. Wells\IEEEauthorrefmark{7} and
         Shantenu Jha\IEEEauthorrefmark{2}
@@
         }
+    \IEEEauthorblockA{
+        \IEEEauthorrefmark{7}National Center for Computational Sciences, Oak Ridge National Laboratory, TN, USA
         }
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-% 
-% \author{
-%     \IEEEauthorblockN{
-%         Author 1\IEEEauthorrefmark{1},
-%         Author 2\IEEEauthorrefmark{2},
-%         Author 3\IEEEauthorrefmark{3}, 
-%         Author 4\IEEEauthorrefmark{3}, 
-%         Author 5\IEEEauthorrefmark{4} and
-%         Author 6\IEEEauthorrefmark{4}
-%     }
-%     \IEEEauthorblockA{
-%         \IEEEauthorrefmark{1}Department\\Institution, address, email
-%     }
-%     \IEEEauthorblockA{
-%         \IEEEauthorrefmark{2}Department\\Institution, address, email
-%     }
-%     \IEEEauthorblockA{
-%         \IEEEauthorrefmark{3}Department\\Institution, address, email
-%     }
-%     \IEEEauthorblockA{
-%         \IEEEauthorrefmark{4}Department\\Institution, address, email
-%     }
-% }
+}
 
@@
 % ---------------------------------------------------------------------------
-\newpage
+% \newpage
 \bibliographystyle{IEEEtran}

### 2017-07-07 introduction.tex
Fixes, refinements in Introduction.

@@
 
-Although the workloads to be executed are comprised of tasks that are
-independent of each other, the management of the distribution of workloads
-across many heterogeneous resources, the effective utilization of resources
-and efficient execution of workloads present non-trivial challenges. Many
-software solutions have been developed in response to these challenges. The
-CMS experiment, devised a solution based around the
-HTCondor~\cite{thain2005distributed} software ecosystem. The
-ATLAS~\cite{Aad:2008} experiment utilizes the Production and Distributed
-Analysis (PanDA) workload management system~\cite{Maeno2011} (WMS) for
-distributed data processing and analysis. The CMS and ATLAS experiments
-utilize, arguably the largest production grade distributed computing
-solutions and have symbolized the paradigm of {\it high-throughput
-computing}, i.e., the effective execution of many independent tasks.
+The LHC workloads are comprised of tasks that are independent of each other,
+however, the management of the distribution of workloads across many
+heterogeneous resources, the effective utilization of resources and efficient
+execution of workloads present non-trivial challenges. Many software solutions
+have been developed in response to these challenges. The CMS experiment,
+devised a solution based around the HTCondor~\cite{thain2005distributed}
+software ecosystem. The ATLAS~\cite{Aad:2008} experiment utilizes the
+Production and Distributed Analysis (PanDA) workload management
+system~\cite{Maeno2011} (WMS) for distributed data processing and analysis.
+The CMS and ATLAS experiments utilize, arguably the largest production grade
+distributed computing solutions, and have symbolized the paradigm of {\it
+high-throughput computing} (HTC), i.e., the effective execution of many
+independent tasks.
 
-In spite of the impressive scale of the ATLAS distributed computing system,
-demand for computing systems will soon significantly outstrip current and
-projected supply.   The data volumes that will need analyzing in LHC-Run 3
-($~\approx$ 2022) and the high-luminosity era (Run 4) will increase by
-factors of 10--100 compared to the current phase (Run 2). There are multiple
-levels at which this problem needs to be addressed: the utilization of
-emerging parallel architectures (e.g., platforms); algorithmic and advances
-in analytical methods (e.g., use of Machine Learning); and the ability to
-exploit different platforms (e.g., clouds and supercomputers).
+In spite of the impressive scale of the ATLAS distributed computing system --
+in the number of tasks executed, the number of core hours utilized, and the
+number of distributed sites utilized,  demand for computing systems will soon
+significantly outstrip current and projected supply.   The data volumes that
+will need analyzing in LHC-Run 3 ($~\approx$ 2022) and the high-luminosity era
+(Run 4) will increase by factors of 10--100 compared to the current phase (Run
+2). There are multiple levels at which this problem needs to be addressed: the
+utilization of emerging parallel architectures (e.g., platforms); algorithmic
+and advances in analytical methods (e.g., use of Machine Learning); and the
+ability to exploit different platforms (e.g., clouds and supercomputers).
 
@@
 free'' of the traditional computational approach of high-throughput computing
-on distributed resources to embrace new platforms, in particular
-high-performance computers. Specifically, we discuss the experience of
+on distributed resources to embrace new platforms, in particular high-
+performance computers (HPC). Specifically, we discuss the experience of
 integrating PanDA WMS with a US DOE leadership machine (Titan) to reach
@@
 
-We demonstrate how Titan is more efficiently utilized by the mixing of small
-and short-lived tasks in backfill with regular payloads. Cycles otherwise
-unusable (or very difficult to use) are used for science thus, increasing the
-overall utilization on Titan without loss of quality-of-service. The
-conventional mix of jobs at OLCF cannot be effectively backfilled because of
-size, duration, and scheduling policies. Our approach is extensible to any
-HPC with ``capability scheduling'' policies.  We also investigate the use of
-a pilot-abstraction based task execution runtime system to flexibly execute
-ATLAS and other heterogeneous workloads (molecular dynamics) using regular
-queues. As such, our approach provides a general solution and investigation
-of the convergence of HPC and HTC execution of workloads.
+In doing so, we demonstrate how Titan is more efficiently utilized by the
+mixing of small and short-lived tasks in backfill with regular payloads.
+Cycles otherwise unusable (or very difficult to use) are used for science,
+thus increasing the overall utilization on Titan without loss of overall
+quality-of-service. The conventional mix of jobs at OLCF cannot be effectively
+backfilled because of size, duration, and scheduling policies. Our approach is
+extensible to any HPC with ``capability scheduling'' policies.  We also
+investigate the use of a pilot-abstraction based task execution runtime system
+to flexibly execute ATLAS and other heterogeneous workloads (molecular
+dynamics) using regular queues. As such, our approach provides a general
+solution and investigation of the convergence of HPC and HTC execution of
+workloads.
 
@@
 scalable and production usage of Titan for historically high-throughput
-workloads, and (ii) early lessons and guidance as the community looks forward
-to designing the next generation of online analytical
-platforms~\cite{foap-url}, for how current and future experimental and
-observational systems can be integrated with production supercomputers in a
-general and extensible manner.
+workloads, and (ii) early lessons and guidance on designing the next
+generation of online analytical platforms~\cite{foap-url},  so that
+experimental and observational systems can be integrated with production
+supercomputers in a general and extensible manner.

### 2017-07-07 overall.tex
ongoing fixes.

--- 2017-07-04 overall.tex
@@
 
-Currently, ATLAS deploys 20 instances of the PanDA Broker on 4 Titan's DTNs,
+Currently 20 instances of the PanDA Broker are deployed on 4 Titan DTNs, with
 5 instances per DTN\@. Each broker submits and manages the execution of 15 to
-300 ATLAS jobs, one job for each Titan's worker node, and a theoretical
-maximum concurrent use of 96,000 cores. Since November 2015, PanDA Brokers
-have operated only in backfill mode, without a defined time allocation, and
-running at the lowest priority on Titan. Therefore, ATLAS contributed to
-increase the Titan's overall utilization without affecting other users.
+300 ATLAS jobs, one job for each Titan worker node, and a theoretical maximum
+concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have
+operated only in backfill mode, without a defined time allocation, and running
+at the lowest priority on Titan.  Therefore, ATLAS contributed to an increase
+of Titan's overall utilization.
+
+% without affecting other users.
 
@@
 
-PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels)
-is defined as the fraction of core-hours utilized by the PanDA Brokers to
-Titan’s total backfill availability during the experiment time window. The
-brokers reached 18\% average efficiency, with a minimum 7.8\% efficiency on
-May 2016 and a maximum 30.9\% efficiency on February 2017 (excluding the
-preliminary results of March). The total backfill availability was 20.3M in
-April 2016, and 17.6M in February 2017. This shows that the measured
-difference in efficiency did not depend on a comparable difference in the
-total backfill availability.
+PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels) is
+defined as the fraction (or percentage) of core-hours utilized by the PanDA
+Brokers to Titan’s total backfill availability during the experiment time
+window. The brokers reached 18\% average efficiency, with a minimum 7.8\%
+efficiency on May 2016 and a maximum 30.9\% efficiency on February 2017
+(excluding the preliminary results of March). The total backfill availability
+was 20.3M in April 2016, and 17.6M in February 2017. This shows that the
+measured difference in efficiency did not depend on a comparable difference in
+the total backfill availability.
 

### 2017-07-07 abstract.tex
Convert paper to IEEETran style for eScience conference
We have both 51 and 52M core hours.
  Changing 51 to 52M.

@@
-% Experiments at the Large Hadron Collider (LHC) face unprecedented computing
-% challenges. Heterogeneous resources are distributed worldwide, thousands of
-% physicists analyzing the data need remote access to hundreds of computing
-% sites, the volume of processed data is beyond the exabyte scale, and data
-% processing requires more than billions of hours of computing usage per year. The
-% PanDA (Production and Distributed Analysis) system was developed to meet the
-% scale and complexity of LHC distributed computing for the ATLAS experiment. In
-% the process, the old batch job paradigm of computing in HEP was discarded in
-% favor of a far more flexible and scalable model. The success of PanDA at the LHC
-% is leading to widespread adoption and testing by other experiments. PanDA is
-% the first exascale workload management system in HEP, already operating at
-% a million computing jobs per day, and processing over an exabyte of data in
-% 2013. We will describe the design and implementation of PanDA, present data on
-% the performance of PanDA at the LHC, and discuss plans for future evolution
-% of the system to meet new challenges of scale, heterogeneity and increasing
-% user base.
-
-% Experiments at the Large Hadron Collider (LHC) face unprecedented computing
-% challenges. Thousands of physicists analyze exabytes of data every year, using
-% billions of computing hours on hundreds of computing sites worldwide. PanDA
-% (Production and Distributed Analysis) is a workload management system (WMS)
-% developed to meet the scale and complexity of LHC distributed computing for the
-% ATLAS experiment. PanDA is the first exascale workload management system in HEP,
-% executing millions of computing jobs per day, and processing over an exabyte of
-% data in 2016. In this paper, we introduce the design and implementation of
-% PanDA, describing its deployment on Titan, the third biggest supercomputer in
-% the world. We analyze scalability, reliability and performance of PanDA on
-% Titan, highlighting the challenges addressed by its architecture and
-% implementation. We present preliminary results of experiments performed with the
-% Next Generation Executer, a prototype we developed to meet new challenges of
-% scale and resource heterogeneity.
-
-% As the LHC prepares for Run 3 in
-% $~\approx$ 2022 and the high-luminosity era (Run 4), it is anticipated that
-% the data volumes that will need analyzing will increase by factors of 10-100
-% compared to the current phase (Run 2). Data will be larger in volume but will
-% also require more sophisticated computational processing.
-
 The computing systems used by LHC experiments has historically consisted of
-the federation of hundreds to thousands of distributed resources, ranging from
-small to mid-size resource.  In spite of the impressive scale of the existing
-distributed computing solutions, the federation of small to mid-size
-resources will be insufficient to meet projected future demands.
-This paper is a case study of how the ATLAS experiment has embraced Titan -- a
-DOE leadership facility in conjunction with traditional distributed high-
-throughput computing to reach sustained production scales of approximately 51M
-core-hours a years. The three main contributions of this paper are:  (i) a
-critical evaluation of design and operational considerations  to support the
-sustained, scalable and production usage of Titan;  (ii) a preliminary
+the federation of hundreds to thousands of distributed resources, ranging
+from small to mid-size resource.  In spite of the impressive scale of the
+existing distributed computing solutions, the federation of small to mid-size
+resources will be insufficient to meet projected future demands. This paper
+is a case study of how the ATLAS experiment has embraced Titan -- a DOE
+leadership facility in conjunction with traditional distributed high-
+throughput computing to reach sustained production scales of approximately
+52M core-hours a years. The three main contributions of this paper are:  (i)
+a critical evaluation of design and operational considerations  to support
+the sustained, scalable and production usage of Titan;  (ii) a preliminary
 characterization of a next generation executor for PanDA to support new
@@
 current and future experimental and observational systems can be integrated
-with production supercomputers and other platforms in a general and extensible
-manner.
+with production supercomputers and other platforms in a general and
+extensible manner.

### 2017-07-08 conclusion.tex
Adding a last sentence about NGE

@@
 utilization of Titan to be consistently upwards of 90\%. Given that the
-average average utilization of most other leadership class machines is less
+ average utilization of most other leadership class machines is less
 (e.g., NSF's flagship Blue Waters the average utilization fluctuates between
@@
 for dynamic resource federation.
+
+The development of NGE is looking forward to deal with these challenges by providing an abstraction layer for HPC resources together with a set of APIs for the coordination and fast deployment of heterogeneous workloads.
 

### 2017-07-08 conclusion.tex
refine conclusion.

@@
 (\S\ref{sec:panda_titan}). Our characterization highlighted the strengths and
-limitations of the current design and implementation: PanDA Brokers enable
-the sustained execution of millions of simulations per week but further work
-is required to optimize its efficiency and reliability
-(\S\ref{ssec:broker_titan}). PanDA Brokers support the concurrent execution
-of multiple AthenaMP instances, enabling each AthenaMP to perform the
-concurrent execution of up to 16 Geant4 simulators. Nonetheless, our
-characterization showed how improving I/O performance could reduce overheads
+limitations of the current design and implementation: PanDA Brokers enable the
+sustained execution of millions of simulations per week but further work is
+required to optimize its efficiency and reliability
+(\S\ref{ssec:broker_titan}). PanDA Brokers support the concurrent execution of
+multiple AthenaMP instances, enabling each AthenaMP to perform the concurrent
+execution of up to 16 Geant4 simulators. Nonetheless, our characterization
+showed how improving I/O performance could reduce overheads
 (\S\ref{ssec:athenamp_titan}), increasing the overall utilization of Titan's
-backfill availability.
+backfill availability. We introduced PanDA's next generation executor for HPC
+systems, characterized its ability to support multi-generation workloads and
+analyzed its scaling behavior. It enables the high-performance execution of
+new workloads and advanced execution modes of traditional workloads.
 
@@
 
-The development of NGE is looking forward to deal with these challenges by providing an abstraction layer for HPC resources together with a set of APIs for the coordination and fast deployment of heterogeneous workloads.
-
 % Recent efforts with Google Cloud for the CMS experiment --

### 2017-07-08 conclusion.tex
Added a piece (Please check that we are still at 10 pages)

@@
 for dynamic resource federation.
+The development of NGE is looking forward to deal with these challenges by providing an abstraction layer for HPC resources together with a set of APIs for the coordination and fast deployment of heterogeneous workloads.
+In this paper, we provided the results of a preliminary testing of the NGE.
+The testing pointed out the robustness of NGE at handling an high number of tasks and worker nodes by mantaining an overhead that grows linearly with the number of tasks.
 

### 2017-07-08 bckground.tex
pending minor fix.
Addendum to Related Work.

@@
 model based on distributing a large amount of single/few-cores tasks across
-hundreds of sites\mtnote{Is this true?}.
+hundreds of sites.
 
 All LHC experiments produce and process large amounts  of data from actual
-collisions in the accelerator and from their simulations. Dedicated, multi-
-tiered data systems have been built to store, replicate, and distributed these
+collisions in the accelerator and from their simulations. Dedicated,  
+multi-tiered data systems have been built to store, replicate, and distributed these
 data. All LHC WMS interface with these systems to move data to the sites where
@@
 amount of) data are already stored.
+
+It is interesting to note that most WMS developed to support LHC experiments are gradually evolving towards HPC resources, though none have reached sustained operational usage at the scales that PanDA has achieved for ATLAS on Titan.
 

### 2017-07-08 bckground.tex

@@
 
-It is interesting to note that most WMS developed to support LHC experiments are gradually evolving towards HPC resources, though none have reached sustained operational usage at the scales that PanDA has achieved for ATLAS on Titan.
+It is interesting to note that most WMS developed to support LHC experiments
+are gradually evolving towards integrating HPC resources, though none have
+reached sustained operational usage at the scales that PanDA has achieved for
+ATLAS on Titan.
 

### 2017-07-08 conclusion.tex
Commented out AA919's text. I added the NGE contribution a few paragraphs before the end of Conclusion.

--- 2017-07-08 conclusion.tex
@@
 systems, characterized its ability to support multi-generation workloads and
-analyzed its scaling behavior. It enables the high-performance execution of
-new workloads and advanced execution modes of traditional workloads.
+analyzed its scaling behavior.  Performance tests on the next generation
+executor demonstrate linear strong and weak scalability over several orders
+of magnitude  of task and node counts.  Qualitatively, it enables the high-
+performance execution of new workloads and advanced execution modes of
+traditional workloads.
 
@@
 
+% The development of NGE is looking forward to deal with these challenges by providing an abstraction layer for HPC resources together with a set of APIs for the coordination and fast deployment of heterogeneous workloads.
+In this paper, we provided the results of a preliminary testing of the NGE.
+The testing pointed out the robustness of NGE at handling an high number of tasks and worker nodes by mantaining an overhead that grows linearly with the number of tasks.
+
 % Recent efforts with Google Cloud for the CMS experiment --

### 2017-08-28 paper.tex

@@
         Alexei Klimentov\IEEEauthorrefmark{4},
-        Sarp H. Oral\IEEEauthorrefmark{6},
+        Sarp Oral\IEEEauthorrefmark{6},
         Jack C. Wells\IEEEauthorrefmark{7} and

### 2017-08-28 Experiments.tex
Comments to address before submission

@@
 \begin{enumerate}
-    \item REVIEWER 2: In the weak scalability evaluations (in Figure 7 and
-    8), the overhead of the Next Generation Executor (NGE) has grown linearly
-    with the number of the nodes. Can you explain why the average time
-    execution time of AthenaMP is decreased even when the overhead is
-    increasing?
-    I checked the data.
-Figure 8 was simply wrong.
-I think that we have overwritten the original with an older version.
-Figure 7: I used a smaller integration step to compute the integral of the pilot concurrency.
-It gets better but not as much as I would like.
-Our data has something sketchy.
+    \color{red} 
+    \item REVIEWER 3: The scalability benchmarks in section "V.B Experiments"
+    should be better explained: what *extacly* is being kept constant in each
+    experiment, and what is being varied and over what range. Also the
+    captions in the figures do not help much.
 \end{enumerate}
@@
 
-\subsection{Experiments}
-\label{sec:ngeExp}
+% ---------------------------------------------------------------------------
+\subsection{Experiments}\label{sec:ngeExp}
 
@@
 
+% ---------------------------------------------------------------------------
 \subsubsection{Weak scalability}
@@
 as tasks) as the number of nodes controlled by the pilot. Each AthenaMP
-simulates 100 events, requiring $\sim 4200$ seconds on average.
+simulates 100 events, requiring \(\sim 4200\) seconds on average.
 
 Tasks do not wait within the Agent's queue since one node is available to
-each AthenaMP instance. Task execution  overheads result primarily from  three
-factors: (i) the initial bootstrapping of the pilot on the nodes; (ii) the
-UnitManager's dispatching of units (tasks) to the agent; and (iii) time for
-the agent to bootstrap all the tasks on the nodes.
+each AthenaMP instance. Task execution  overheads result primarily from
+three factors: (i) the initial bootstrapping of the pilot on the nodes; (ii)
+the UnitManager's dispatching of units (tasks) to the agent; and (iii) time
+for the agent to bootstrap all the tasks on the nodes.
 
@@
 Fig.~\ref{fig:weakScal1a} depicts the average pilot duration, the average
-execution time of AthenaMP, and the pilot overhead as function of the
-pilot size.
+execution time of AthenaMP, and the pilot overhead as function of the pilot
+size.
 
@@
    	\vspace{-0.3in}
-    \caption{Weak scalability: average pilot duration, average duration of one AthenaMP execution, and pilot's overhead as a function of pilot sizes (200, 500, 1000 and 2000 nodes).}
-\label{fig:weakScal1a}
+    \caption{Weak scalability: average pilot duration, average duration of
+    one AthenaMP execution, and pilot's overhead as a function of pilot sizes
+    (200, 500, 1000 and 2000 nodes).}\label{fig:weakScal1a}
 \end{figure}
@@
 
+% ---------------------------------------------------------------------------
 \subsubsection{Weak scalability with multiple generation}
@@
     \vspace{-0.3in}
-    \caption{Weak scalability with multiple generations (where each generation has approximately 1/6th the number of events compared to Fig. 7): average pilot
-    duration, average duration of sequential AthenaMP executions, and
-    pilot's overhead for pilot with 256, 512, 1024 and 2048 nodes.}
-\label{fig:weakScal2a}
+    \caption{Weak scalability with multiple generations (where each
+    generation has approximately 1/6th the number of events compared to Fig.
+    7): average pilot duration, average duration of sequential AthenaMP
+    executions, and pilot's overhead for pilot with 256, 512, 1024 and 2048
+    nodes.}\label{fig:weakScal2a}
 \end{figure}
@@
 
+% ---------------------------------------------------------------------------
 \subsubsection{Strong scalability}
@@
 different pilot sizes. We used 2048 AthenaMP instances and pilots with 256,
-512, 1024 and 2048 nodes. Thus, the number of AthenaMP generations is equal to
-eight times the size of the smallest pilot and corresponds to the size of the
-largest pilot. As a consequence, the number of consecutive generations of
+512, 1024 and 2048 nodes. Thus, the number of AthenaMP generations is equal
+to eight times the size of the smallest pilot and corresponds to the size of
+the largest pilot. As a consequence, the number of consecutive generations of
 AthenaMP decreases with the pilot size. These experiments are designed to
@@
     \caption{Strong scalability: Average pilot duration, average duration of
-    sequential AthenaMP executions, and pilot's overhead for pilots with 256, 512, 1024 and 2048 nodes.}
-\label{fig:strongScala}
+    sequential AthenaMP executions, and pilot's overhead for pilots with 256,
+    512, 1024 and 2048 nodes.}\label{fig:strongScala}
 \end{figure}

### 2017-08-28 PandaEvolution.tex
Comments to address before submission

@@
-% \ifreview
-% Comments to address in this section:
-% \begin{enumerate}
-% 	\item REVIEWER 2: The authors spent quite a few paragraphs for the detail
-% 	design and implementation of the PanDA WMS, but introduced few background
-% 	of the LHC applications. According to the paper, I understand the
-% 	applications running on Titan are a large amount of small independent
-% 	jobs. Thus, it seems less cost-effective to run these jobs on a
-% 	world-leading supercomputer like Titan which has a fast and expensive
-% 	interconnection. It will be more interesting if the authors can compare
-% 	the solution on Titan (proposed in this paper) with the original
-% 	distributed computing environment and HPECloud from Google mentioned in
-% 	the related work section.
-% 	\item REVIEWER 5: In Section 3.2, GPUs are not considered in this work,
-% 	which would be a waste for using HPC computing. Could this work be
-% 	extended to include GPUs or other heterogenous resources?
-% \end{enumerate}
-% \fi
+\ifreview
+Comments to address in this section:
+\begin{enumerate}
+	\color{red} 
+	\item REVIEWER 2: the technical discussion of TITAN and WLCG differences
+	is abruptly stopped by the description of Monte Carlo event generation
+	and detector simulation (3A).
+	\item REVIEWER 2: The single listing (3B) of similar projects within the
+	collaboration does not detail whether this is derived or separate work.
+	\item REVIEWER 2: How does the TITAN environment prevent usage of DTNs as
+	thin proxies? Is connectivity possible from compute node to DTNs via MPI,
+	message queues, sockets or other means? Are the restrictions a
+	shortcomming of TITAN, or the majority of HPC environments? Given the
+	proclaimed importance of HPC resources, a generic evaluation seems
+	needed.
+	\item REVIEWER 2: Why was the PanDA Broker designed orthogonal to the
+	pilot model, performing both resource acquisition and workload execution?
+	How do pilots conflict with backfill slots, when the former are designed
+	to acquire resources independent of workload requirements? Why does each
+	Broker instance query and use backfill separately? Given that this seems
+	to be a major bottleneck, the choice of such a design begs for
+	motivation.
+	\item REVIEWER 2: Some facts about WLCG seem outdated:
+	\begin{itemize}
+		\item The number of 30,000 unused TITAN cores (§IIIB) is equated to
+		roughly 10\% of the WLCG. However, WLCG currently has 591,282 cores
+		[REBUSCPU]. https://wlcg-rebus.cern.ch/apps/capacities/federations/
+		\item WLCG sites are claimed to provide only 8 cores per worker node
+		(§IIIB). WLCG has a significant number of sites providing 16 or more
+		cores per worker node [REBUSCPU]. https://wlcg-rebus.cern.ch/apps/capacities/federations/
+	\end{itemize}
+	\item REVIEWER 2: Schematic figures (Fig 1, Fig 2) are overly complex,
+	while their captions are not sufficient to understand them. The numbering
+	is only partially explained in the text.
+\end{enumerate}
+\fi
+
 The upcoming LHC Run 3 will require more resources than the Worldwide LHC
@@
 % ---------------------------------------------------------------------------
-\subsection{Architectures, Interfaces and Workloads}
-\label{ssec:panda-titan}
+\subsection{Architectures, Interfaces and Workloads}\label{ssec:panda-titan}
 

### 2017-08-28 conclusion.tex
Comments to address before submission

--- 2017-07-08 conclusion.tex
@@
 Comments to address in this section:
-\begin{enumerate}
-	\item REVIEWER 3: How extensible is this? Do you have a sense of backfill
-	availability on other systems? Is this unique to Titan or Cray or more
-	ubiquitous of an option?
-\end{enumerate}
+\begin{itemize}
+	\color{red} 
+	\item REVIEWER 2: How does Summit, the TITAN successor, compare to the
+	conclusions? The end of the TITAN life-cycle basically invalidates the
+	applicability of all conclusions if this generalisation is missing.
+	\item REVIEWER 3: This statement in the "conclusions" section is quite
+	exaggerated: "The experience captured in this paper will prove invaluable
+	for designing WMS for computational campaigns and will provide a baseline
+	to evaluate the relative merits of different approaches." WMS systems for
+	LHC experiments have been developed and run in production for many years;
+	while this paper adds interesting experience and data points, it's
+	unlikely that this will be "invaluable" for new designs and a cornerstone
+	for their evaluation.
+\end{itemize}
 \fi
@@
 (\S\ref{sec:panda_titan}). Our characterization highlighted the strengths and
-limitations of the current design and implementation: PanDA Brokers enable the
-sustained execution of millions of simulations per week but further work is
-required to optimize its efficiency and reliability
-(\S\ref{ssec:broker_titan}). PanDA Brokers support the concurrent execution of
-multiple AthenaMP instances, enabling each AthenaMP to perform the concurrent
-execution of up to 16 Geant4 simulators. Nonetheless, our characterization
-showed how improving I/O performance could reduce overheads
+limitations of the current design and implementation: PanDA Brokers enable
+the sustained execution of millions of simulations per week but further work
+is required to optimize its efficiency and reliability
+(\S\ref{ssec:broker_titan}). PanDA Brokers support the concurrent execution
+of multiple AthenaMP instances, enabling each AthenaMP to perform the
+concurrent execution of up to 16 Geant4 simulators. Nonetheless, our
+characterization showed how improving I/O performance could reduce overheads
 (\S\ref{ssec:athenamp_titan}), increasing the overall utilization of Titan's
@@
 utilization of Titan to be consistently upwards of 90\%. Given that the
- average utilization of most other leadership class machines is less
-(e.g., NSF's flagship Blue Waters the average utilization fluctuates between
+average utilization of most other leadership class machines is less (e.g.,
+NSF's flagship Blue Waters the average utilization fluctuates between
 60--80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
@@
 for dynamic resource federation.
+
+% The development of NGE is looking forward to deal with these challenges by
+% providing an abstraction layer for HPC resources together with a set of
+% APIs for the coordination and fast deployment of heterogeneous workloads.
+% In this paper, we provided the results of a preliminary testing of the NGE.
+% The testing pointed out the robustness of NGE at handling an high number of
+% tasks and worker nodes by mantaining an overhead that grows linearly with
+% the number of tasks.
 

### 2017-08-28 introduction.tex
Comments to address before submission

@@
 \begin{enumerate}
-	\item REVIEWER 1: listing the key observations in one place, and in a
-	consistent fashion. I found The 5 points listed in the first part of the
-	rebuttal quite compelling, and strongly urge the authors to put them in
-	the paper in a prominent location.
-	\item REVIEWER 5: I would have preferred to see a more detailed list of
-	contributions, including the novelty of the proposed approach as compared
-	to prior work.
-	\item I am ambivalent about the use of the term ``converging'' in the
-	paper's title. While the paper does port workloads to high-performance
-	computing resources, it does not deal with compute-oriented optimizations
-	(parallelization), lowering communication, etc., which are often major
-	concerns in HPC programming. In this sense, the paper is closer in spirit
-	to how to port high-throughput apps to high-performance clusters. Agreed?
-	{\bf  Need to clarify it is not the task that is being converged but the workload and execution model that is being converged}
-	\item REVIEWER 1: A glossary would be welcomed. {\bf Not addressable}
+	\color{red} 
+	\item REVIEWER 1: The introduction is overly verbose and prosaic. The
+	praise for LHC collaborations is off-putting, especially as it ignores
+	other work. For example, claiming "CMS and ATLAS experiments utilize
+	arguably the largest production grade distributed computing solutions"
+	disregards the obvious commercial contenders. By blatantly excluding
+	obvious comparisons, it raises the question how much other work is
+	ignored.
 \end{enumerate}

### 2017-08-28 nge.tex
Comments to address before submission

@@
-% As seen in \S\ref{ssec:panda_titan}, PanDA Broker is deployed on the DTNs
-% as Titan's worker nodes lack connectivity to the wide area network.
+\ifreview
+Comments to address in this section:
+\begin{itemize}
+	\color{red} 
+	\item REVIEWER 3: The first paragraphs of section V. are basically a
+	rediscovery of the advantages of pilot jobs for WLCG task scheduling.
+	Since all those pain points were well-known in the WLCG community before
+	the project started, it should be explained why a pilot job approach was
+	not taken in the first place.
+	\item REVIEWER 3: Figure 6. refers to a "User Workstation" as the running
+	place for the Pilot and Unit Managers, which is never mentioned in the
+	text. (I got the impression that the NGE would run on a DTN node; if this
+	is not the case and separate workstations are being used for the testing
+	phase, this should be made clear.)
+\end{itemize}
+\fi
+
 The lack of pilot capabilities in PanDA Broker impacts both the efficiency
@@
 
-The absence of pilots 
-% imposes the static coupling between MPI scripts submitted to the PBS batch
-% system and detector simulations (\S\ref{sec:panda_titan}). This
-makes the scheduling of multiple generations of workload on the same PBS job
-impossible: once a statically defined number of detector simulations are
-packaged into a PBS job and this job is queued on Titan, no further
-simulations can be added to that job. New simulations have to be packaged
-into a new PBS job that needs to be submitted to Titan based upon backfill
-and PanDA Brokers availability.
+The absence of pilots makes the scheduling of multiple generations of
+workload on the same PBS job impossible: once a statically defined number of
+detector simulations are packaged into a PBS job and this job is queued on
+Titan, no further simulations can be added to that job. New simulations have
+to be packaged into a new PBS job that needs to be submitted to Titan based
+upon backfill and PanDA Brokers availability.
 
@@
 
-Multiple generations would also relax two assumptions of the current execution
-model: knowing the number of simulations before submitting the MPI script, and
-having a fixed number of events per simulation (currently 100). Pilots would
-enable the scheduling of simulations independently from whether they were
-available at the moment of submitting the pilot. Further, simulations with a
-varying number of events could be scheduled on a pilot, depending on the
-amount of remaining walltime and the distribution of execution time per event,
-as shown in \S\ref{ssec:panda_titan}, Fig.~\ref{fig:comparison-8-16cores}.
-These capabilities would increase the efficiency of the PanDA Broker when
-there is a large difference between the number of cores and walltime.
+Multiple generations would also relax two assumptions of the current
+execution model: knowing the number of simulations before submitting the MPI
+script, and having a fixed number of events per simulation (currently 100).
+Pilots would enable the scheduling of simulations independently from whether
+they were available at the moment of submitting the pilot. Further,
+simulations with a varying number of events could be scheduled on a pilot,
+depending on the amount of remaining walltime and the distribution of
+execution time per event, as shown in \S\ref{ssec:panda_titan},
+Fig.~\ref{fig:comparison-8-16cores}. These capabilities would increase the
+efficiency of the PanDA Broker when there is a large difference between the
+number of cores and walltime.
 
@@
 	\vspace{-0.3in}
-  \caption{NGE Architecture: The PilotManager and UnitManager reside on a DTN
-  while the Pilot Agent is executed on a worker node. Color coding: gray for
-  entities external to NGE\@; white for APIs; purple for NGE's modules; green
-  for pilots; yellow for module's components.}
+	\caption{NGE Architecture: The PilotManager and UnitManager reside on a
+  	DTN while the Pilot Agent is executed on a worker node. Color coding:
+  	gray for entities external to NGE\@; white for APIs; purple for NGE's
+  	modules; green for pilots; yellow for module's components.}
 \label{fig:arch-overview}

### 2017-08-28 overall.tex
Checkpoint
Comments to address before submission

@@
 \begin{enumerate}
-    % \item REVIEWER 2: In Figure 4, there are some inconsistencies. The
-    % figure caption shows there are only 62175 measures, but in the scatter
-    % plot (right) there are 62555 data points. In addition, according to the
-    % distribution of the number of work nodes (left), it does not look like
-    % the mean number of work available is 691 as the figure caption
-    % illustrates.
-    \item 
-    % REVIEWER 3: It’s interesting that Feb/March have such an increase in
-    % backfill availability in 2017. The authors attribute this to hardware
-    % upgrades for the DTNs. It’s mentioned that the load is only 3.6\% now
-    % and therefore further upgrades will likely not cause similar
-    % improvement. What was the previous CPU load? 
-    Why were particularly low percentages utilized in April/May 2016?
-    % \item REVIEWER 4: “Genat4” at the end of Section 4 is a typo.
-    % \item REVIEWER 4: Please explain why you used Titan at all, in spite of
-    % your statement that it is 1/2 to 1/3 as efficient as other resources.
-    % Was it merely that you were given so much time on this resource that
-    % the inefficiency doesn’t matter?
-    \item REVIEWER 4: Why didn’t you benchmark the performance of your
-    workload on other HPC resources that might have delivered better
-    performance, prior to reengineering your entire workflow to deal with the
-    challenges posed by Titan, a machine that doesn’t seem to justify the
-    pain?
-    \item REVIEWER 4: The INCITE program operational requirements for the
-    LCFs are 90\% uptime. Using large batches of Geant4 jobs that run on a
-    single node or less to reach this metric sounds like gaming the system
-    rather than an accomplishment.
-    % \item REVIEWER 5: In Figure 3, it shows the unused core hours are much
-    % higher than those used by ATLAS\@. What are the reasons for such low
-    % usage?
+    \color{red} 
+    \item REVIEWER 2: What are the reasons for the low efficiency of PanDA
+    Broker? Why are only 30\% of backfill slots useable, if they are
+    basically uncontested? Why does Fig 3 show an increase of absolute and
+    relative efficiency in Aug-16, when the text claims this in Sep-16 due to
+    more PanDA Brokers? Why are only 20 PanDA Brokers deployed if this is a
+    limiting factor? Why is it questionable that more PanDA Brokers would
+    increase efficiency? If only 5/6 of resources below the mean are
+    unusable, this implies an efficiency of almost 60\% being reachable.
+    \item REVIEWER 2: Given an expected increase in resource need of 10-100,
+    how is the capacity of 4\% WLCG significant? How does this relate to the
+    manpower required to develop and maintain a custom solution per HPC site?
+    How many TITAN resources are actually available to HEP, given the use by
+    other domains? The magnitude of capacities presented as available is in
+    major contrast to the requirements sketched in the introduction.
+    \item REVIEWER 2: How does the poor metadata read performance of
+    1500-6300 s (25\%-100\% of minimum walltime!) affect efficiency? Is it
+    factored into the choice of walltime requests? Is it applicable only to
+    TITAN, to HPC in general, or even to regular WLCG sites? Are there
+    solutions at application, e.g. the use of packaged python scripts and
+    libraries? The information presented is insufficient to estimate whether
+    this issue applies to other environments and usecases as well, or is
+    simply due to suboptimal implementation/configuration.
+    \item REVIEWER 3: The HEP-SPEC06 benchmark, which is used as a
+    ``computing power'' measure in LCG is mostly integer-artihmetic based: FP
+    operations accounts for \~10\% of the average workload (see [1], section
+    ``Why do you use `all\_cpp' instead of SPECint or SPECfp?'') While this
+    figure certainly comes from averaging over many different workloads
+    across all experiments, it seems a bit surprising that FP scheduling
+    contention is responsible for 50\% decrease in compute efficiency on
+    Titan when compared to other WLCG sites (paper page 6).
 \end{enumerate}
@@
 concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have
-operated only in backfill mode, without a defined time allocation, and running
-at the lowest priority on Titan.  Therefore, ATLAS contributed to an increase
-of Titan's overall utilization.
+operated only in backfill mode, without a defined time allocation, and
+running at the lowest priority on Titan.  Therefore, ATLAS contributed to an
+increase of Titan's overall utilization.
 
@@
 
-PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels) is
-defined as the fraction (or percentage) of core-hours utilized by the PanDA
-Brokers to Titan’s total backfill availability during the experiment time
-window. The brokers reached 18\% average efficiency, with a minimum 7.8\%
-efficiency on May 2016 and a maximum 30.9\% efficiency on February 2017
+PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels)
+is defined as the fraction (or percentage) of core-hours utilized by the
+PanDA Brokers to Titan’s total backfill availability during the experiment
+time window. The brokers reached 18\% average efficiency, with a minimum
+7.8\% efficiency on May 2016 and a maximum 30.9\% efficiency on February 2017
 (excluding the preliminary results of March). The total backfill availability
 was 20.3M in April 2016, and 17.6M in February 2017. This shows that the
-measured difference in efficiency did not depend on a comparable difference in
-the total backfill availability.
+measured difference in efficiency did not depend on a comparable difference
+in the total backfill availability.
 
@@
 for the previous DTNs. As such, further hardware upgrades seem unlikely to
-improve significantly the performance of PanDA Brokers.\mtnote{Asked Segey
-and Danila for values previous HW upgrade}
+improve significantly the performance of PanDA Brokers.
 

### 2017-08-28 paper.tex
Comments to address before submission

@@
 \maketitle
+
+\ifreview
+General comments to address:
+\begin{itemize}
+    \color{red} 
+    \item REVIEWER 1: There is a problem with capitalization in some
+    reference titles, ie. [2], [3], [4], [7], [8], [9], [10], [11], [12],
+    [13]. Reference [6] is just an author list, year and title. Should there
+    be more here?
+    \item REVIEWER 2:  Key concepts such as pilot jobs, tasks and workloads
+    are not meaningfully defined.
+    \item REVIEWER 2: the information provided is insufficient to asses in
+    how far any of the presented evaluation, characterization or lessons are
+    applicable outside of TITAN and PanDA Broker combined.
+    \item REVIEWER 2: Several references are of poor quality. For example,
+    [19] is a slide collection even though there are many papers for the
+    referenced work.
+    \item REVIEWER 3: Title looks too general for what amounts to the
+    implementation of an adapter component in a system that is very specific
+    to a single HEP experiment. (In general, title, abstract and conclusions
+    seem to over-state the contributions of this paper, considering this is a
+    general eScience conference and not a WLCG-specific one.)
+\end{itemize}
+\fi
+
 \begin{abstract}
@@
 % ---------------------------------------------------------------------------
-\section{Introduction}
-\label{sec:intro}
+\section{Introduction}\label{sec:intro}
 \input{introduction}
@@
 % ---------------------------------------------------------------------------
-\section{PanDA Overview}
-\label{sec:panda_overview}
+\section{PanDA Overview}\label{sec:panda_overview}
 \input{PandaOverview}
@@
 % ---------------------------------------------------------------------------
-%\section{Deploying PanDA on a Leadership-scale system}
-\section{Deploying PanDA on Titan}
-\label{sec:panda_titan}
+\section{Deploying PanDA on Titan}\label{sec:panda_titan}
 \input{PandaEvolution}
@@
 % ---------------------------------------------------------------------------
-\section{Analysis and Discussion}
-\label{sec:panda_titan}
+\section{Analysis and Discussion}\label{sec:analysis}
 \input{overall}
@@
 % ---------------------------------------------------------------------------
-% VI - THE NEXT GENERATION EXECUTOR
+% V - THE NEXT GENERATION EXECUTOR
 % ---------------------------------------------------------------------------
-% \section{PANDA: Future/RoadMap}
-\section{PANDA\@: The Next Generation Executor}
-\label{sec:panda_roadmap}
+\section{PANDA\@: The Next Generation Executor}\label{sec:panda_roadmap}
 \input{nge}
@@
 % ---------------------------------------------------------------------------
-% VII - RELATED WORK
+% VI - RELATED WORK
 % ---------------------------------------------------------------------------
-%\vspace{-0.4in}
-\section{Related Work}
-\label{sec:related}
+\section{Related Work}\label{sec:related}
 \input{bckground}
@@
 % ---------------------------------------------------------------------------
-% VIII - CONCLUSION
+% VII - CONCLUSION
 % ---------------------------------------------------------------------------
-%\vspace{-0.1in}
-\section{Conclusion}
-\label{sec:conclusion}
+\section{Conclusion}\label{sec:conclusion}
 \input{conclusion}
@@
 % ---------------------------------------------------------------------------
-% \section*{Acknowledgements}
-% \label{sec:ack}
-%
+% \section*{Acknowledgements}\label{sec:ack}
 % \input{ack}
@@
 
-% ---------------------------------------------------------------------------
-% A - APPENDIX
-% ---------------------------------------------------------------------------
-% \appendix
-% \section{Artifact Description --- Converging High-Throughput and High-Performance Computing: A Case Study}
-% \label{apnx:ac}
-% \input{appendix}
-
 % that's all folks

### 2017-09-08 Experiments.tex
Address reviewers' comments: please review/iterate

@@
 which we investigate the weak scalability, weak scalability with multiple
-generation, and strong scalability of the NGE.
+generation, and strong scalability of the NGE\@.
 

### 2017-09-08 PandaEvolution.tex
Address reviewers' comments: please review/iterate

@@
 	is abruptly stopped by the description of Monte Carlo event generation
-	and detector simulation (3A).
+	and detector simulation (3A).\mtnote{I read the section twice and I did
+	not see such an abrupt transition.}
 	\item REVIEWER 2: The single listing (3B) of similar projects within the
-	collaboration does not detail whether this is derived or separate work.
+	collaboration does not detail whether this is derived or separate
+	work.\mtnote{Added a sentence to clarify that PanDA Broker operates
+	within the existing ATLAS software infrastructure. Specified that the
+	ATLAS team has ported PanDA Broker to other supercomputers.}
 	\item REVIEWER 2: How does the TITAN environment prevent usage of DTNs as
@@
 	proclaimed importance of HPC resources, a generic evaluation seems
-	needed.
+	needed.\mtnote{Added a sentence explaining that DTNs are not part of
+	Titan's compute nodes.}
 	\item REVIEWER 2: Why was the PanDA Broker designed orthogonal to the
-	pilot model, performing both resource acquisition and workload execution?
-	How do pilots conflict with backfill slots, when the former are designed
-	to acquire resources independent of workload requirements? Why does each
-	Broker instance query and use backfill separately? Given that this seems
-	to be a major bottleneck, the choice of such a design begs for
-	motivation.
+	pilot model, performing both resource acquisition and workload
+	execution?\mtnote{We explain that PanDA Broker is a prototype and now we
+	are developing NGE.} How do pilots conflict with backfill slots, when the
+	former are designed to acquire resources independent of workload
+	requirements?\mtnote{AFAIK, we never state that pilot conflict with
+	backfill. We argue that the latter mitigates the absence of the former.}
+	Why does each Broker instance query and use backfill separately? Given
+	that this seems to be a major bottleneck, the choice of such a design
+	begs for motivation.\mtnote{Extended a paragraph to explain that PanDA
+	Broker supported rapid deployment and incremental optimization while
+	incurring in functional limitations.}
 	\item REVIEWER 2: Some facts about WLCG seem outdated:
@@
 		roughly 10\% of the WLCG. However, WLCG currently has 591,282 cores
-		[REBUSCPU]. https://wlcg-rebus.cern.ch/apps/capacities/federations/
+		[REBUSCPU].
+		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{Done.}
 		\item WLCG sites are claimed to provide only 8 cores per worker node
 		(§IIIB). WLCG has a significant number of sites providing 16 or more
-		cores per worker node [REBUSCPU]. https://wlcg-rebus.cern.ch/apps/capacities/federations/
+		cores per worker node [REBUSCPU].
+		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{This
+		would require fairly deep rewriting on Section IVB. I edited the
+		sentence referring to 8 cores eliminating the reference to compute
+		nodes. Less than ideal but I did not find a better solution. Any
+		suggestion?}
 	\end{itemize}
@@
 	while their captions are not sufficient to understand them. The numbering
-	is only partially explained in the text.
+	is only partially explained in the text.\mtnote{Added a reference to
+	where the numbering of the figures are explained in the text. I
+	double-checked and the numbering of both figure is fully explained in the
+	text.}
 \end{enumerate}
@@
 Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
-300,000 cores at more than 100 Grid sites, with a peak performance of 0.3
-petaFLOPS\@. This capacity will be sufficient for the planned analysis and
-data processing, but it will be insufficient for the Monte Carlo production
-workflow and any extra activity. To alleviate these challenges, ATLAS is
-expanding the current computing model to include additional resources such as
-the opportunistic use of supercomputers.
-
-% Supercomputers are designed to support parallel computation that requires
-% runtime communication. Jobs are executed across multiple nodes, each node
-% calculating a small part of the problem and communicating with other node,
-% typically via MPI\@. Accordingly, supercomputers have large number of
-% worker nodes, connected through a high-speed, low-latency dedicated
-% network. Each worker node has multicore CPUs, usually augmented with
-% Graphics Processing Units (GPUs) or other specialized co-processors.
+600,000 cores at more than 100 Grid sites, with an aggregated performance of
+8 petaFLOPS\@. This capacity will be sufficient for the planned analysis
+and data processing, but it will be insufficient for the Monte Carlo
+production workflow and any extra activity. To alleviate these challenges,
+ATLAS is expanding the current computing model to include additional
+resources such as the opportunistic use of supercomputers.
 
@@
 \subsection{Architectures, Interfaces and Workloads}\label{ssec:panda-titan}
-
-% The Titan supercomputer, current number three on the Top 500
-% list,%~\cite{top500}, 
-% is a Cray XK7 system with 18,688 worker nodes and a total
-% of 299,008 AMD cores and 18,688 Nvidia Tesla GPUs. 
-% Each worker node has an AMD Opteron 6274 16-core CPU,
-% a Nvidia Tesla K20X GPU, 32 GB of RAM and no local storage, though a 16 GB
-% RAM disk can be set up. Work nodes use Cray’s Gemini interconnect for
-% inter-node MPI messaging. Titan is served by the Spider
-% II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of disk storage, and
-% by a 29 PB HPSS tape storage system. Titan’s worker nodes run Compute Node
-% Linux, a run time environment based on SUSE Linux Enterprise Server.
-%
-% Titan's users submit jobs to Titan's PBS scheduler by logging into login or
-% data transfer nodes (DTNs). Titan's authentication and authorization model
-% is based on two-factor authentication with a RSA SecurID key. Login nodes
-% and DTNs have out/inbound wide area network connectivity while worker nodes
-% have only local network access. Fair-share and allocation policies are in
-% place both for the PBS batch system and shared file systems.
 
@@
 GEANT4 detector simulation toolkit~\cite{agostinelli2003geant4}. These tasks
-account for $\approx$ 60\% of all the jobs on WLCG, making them a primary
+account for \(\approx\)60\% of all the jobs on WLCG, making them a primary
 candidate for offloading.
@@
 Detector simulation is part of the ATLAS production Monte Carlo (MC)
-workflow~\cite{rimoldi2006atlas}. %,de2013delphes,ritsch2014atlas}. 
-The MC workflow consists of four main stages: event generation, detector
-simulation, digitization, and reconstruction. Event generation creates sets
-of particle four-momenta via different generators, e.g.,
-PYTHIA, %~\cite{sjostrand2006pythia}, 
-HERWIG, %~\cite{corcella2001herwig} 
-and many others. Geant4 simulates the ATLAS detector and the interaction
-between the detector and particles. Each interaction creates a so-called hit
-and all hits are collected and passed on for digitalization, where hits are
-further processed to mimic the readout of the detector. Finally,
-reconstruction operates local pattern recognition, creating high-level
-objects like particles and jets.
+workflow~\cite{rimoldi2006atlas}. The MC workflow consists of four main
+stages: event generation, detector simulation, digitization, and
+reconstruction. Event generation creates sets of particle four-momenta via
+different generators, e.g., PYTHIA, HERWIG, and many others. Geant4 simulates
+the ATLAS detector and the interaction between the detector and particles.
+Each interaction creates a so-called hit and all hits are collected and
+passed on for digitalization, where hits are further processed to mimic the
+readout of the detector. Finally, reconstruction operates local pattern
+recognition, creating high-level objects like particles and jets.
 
 % ---------------------------------------------------------------------------
-\subsection{PanDA Broker}
-\label{ssec:panda_titan}
+\subsection{PanDA Broker}\label{ssec:panda_titan}
 
@@
 Compared to pilots, backfill has the disadvantage of limiting the amount of
-worker nodes that can be requested. Pilots are normal jobs: they can request
-as many worker nodes and walltime as a queue can offer. On the contrary, jobs
+resources that can be requested. Pilots are normal jobs: they can request as
+many worker nodes and walltime as a queue can offer. On the contrary, jobs
 sized according to an available backfill slot depend on the number of worker
@@
 corresponding to an average of 30,000 unused cores (excluding GPU cores).
-This equals to roughly 10\% of the overall capacity of WLCG\@.
+This equals to roughly 5\% of the overall capacity of WLCG\@.
 
@@
 of Titan, PanDA pilot was repurposed to serve as a job broker on the DTN
-nodes of Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of
-PanDA Pilot and its stand-alone architecture, this prototype called `PanDA
-Broker' 
-% implements functionalities to: (i) interrogate Titan about backfill
-% availability via a dedicated Moab scheduler command; (ii) pull ATLAS jobs
-% and events from PanDA Server; (iii) wrap the payload of ATLAS jobs into MPI
-% scripts; (iv) submitting workload-specific MPI scripts to Titan's PBS batch
-% system and monitor their execution; and (v) staging and preparing
-% input/output files. Backfill querying, payload wrapping, and scripts
-% submission required a new implementation while pulling ATLAS job and
-% events, and file staging were inherited from PanDA Pilot.
-%
-% Backfill querying is performed via a dedicated Moab scheduler command while
-% a tailored Python MPI script is used to execute the payload of ATLAS jobs.
-% This enables the execution of unmodified Grid-centric, ATLAS jobs on Titan.
-% Typically, a MPI script is workload-specific as it sets up the execution
-% environment for a specific payload. This involves organization of worker
-% directories, data management, optional input parameters modification, and
-% cleanup on exit. Upon submission, a copy of the MPI script runs on every
-% available worker node, starting the execution of the ATLAS job's payload in
-% a subprocess and waits until its completion.
-%
-% MPI scripts are submitted to Titan's PBS batch system via
-% RADICAL-SAGA~\cite{radical-saga_url}, a Python module, compliant with the
-% OGF GFD.90 SAGA specification~\cite{goodale2008simple}. The Simple API for
-% Grid Applications (SAGA) offers a unified interface to diverse job
-% schedulers and file transferring services. In this way, SAGA provides an
-% interoperability layer that lowers the complexity of using distributed
-% infrastructures. Behind the API façade, RADICAL-SAGA implements a adaptor
-% architecture: each adaptor interface the SAGA API with different middleware
-% systems and services, including the PBS batch scheduler of Titan.
-%
-% The data staging capabilities of the PanDA Broker are implemented via a
-% file system shared among DTNs and worker nodes. The input files with the
-% events of the ATLAS jobs are downloaded on the shared filesystem from the
-% data center of Brookhaven National Laboratory (BNL). The MPI script setup
-% process includes making the location of these files available to the
-% payload of the ATLAS's jobs. The PanDA Broker can locate the payload's
-% output files on the shared filesystem and transfer them from Titan BNL\@.
-%
-% Once deployed on Titan, every PanDA Broker 
-supports the execution of MC detector simulations in 9 steps.
+nodes of Titan (Fig.~\ref{fig:panda_broker}). This prototype called `PanDA
+Broker' maintains the core modules of PanDA Pilot and its stand-alone
+architecture. This imposes functional trade-offs (e.g., single-threaded
+architecture, single MPI PBS script submission) but allows for rapid adoption
+and iterative optimization. PanDA Brokers are deployed on DTNs because these
+nodes are part of the OLCF infrastructure and can access Titan without RSA
+SecureID authentication. It should be noted that DTNs are not part of Titan's
+worker nodes and, therefore, are not used to execute Titan's jobs.
 
-Each PanDA Broker deployed on Titan queries the Job Dispatcher module of the
-PanDA server for ATLAS jobs that have been bound to Titan by JEDI
+Currently, up to 20 PanDA Brokers operate within the existing ATLAS
+production software infrastructure, each supporting the execution of MC
+detector simulations in 9 steps. Each broker queries the PanDA Server for
+ATLAS jobs that have been bound to Titan by JEDI
 (Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA
-Broker pulls jobs' input files from BNL to the OLCF Lustre file system
-(Fig.~\ref{fig:panda_broker}:2). PanDA Broker queries Titan's Moab scheduler
-about the current available backfill slot (Fig.~\ref{fig:panda_broker}:3) and
-creates an MPI script, wrapping enough ATLAS jobs' payload to fit the
-backfill slot. PanDA Broker submits the MPI script to the Titan's Torque
-batch system via RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).
+Broker pulls jobs' input files from BNL Data Center to the OLCF Lustre file
+system (Fig.~\ref{fig:panda_broker}:2). PanDA Broker queries Titan's Moab
+scheduler about the current available backfill slot
+(Fig.~\ref{fig:panda_broker}:3) and creates an MPI script, wrapping enough
+ATLAS jobs' payload to fit the backfill slot. PanDA Broker submits the MPI
+script to the Titan's PBS batch system via RADICAL-SAGA
+(Fig.~\ref{fig:panda_broker}:4).
 
@@
 
-While PanDA Broker implementation is resource specific, it was successfully
-ported to other supercomputers, including the HPC2 at the National Research
-Center ``Kurchatov Institute'' (NRC-KI)~\cite{belyaev2015integration},
-Edison/Cori at the National Energy Research Scientific Computing Center
-(NERSC)~\cite{barreiro2016panda}, and SuperMUC at the Leibniz Supercomputing
-Centre (LRZ)~\cite{barreiro2016panda}.
+PanDA Broker implementation is resource specific but the ATLAS team has
+ported it to other supercomputers, including the HPC2 at the National
+Research Center ``Kurchatov Institute''
+(NRC-KI)~\cite{belyaev2015integration}, Edison/Cori at the National Energy
+Research Scientific Computing Center (NERSC)~\cite{barreiro2016panda}, and
+SuperMUC at the Leibniz Supercomputing Centre (LRZ)~\cite{barreiro2016panda}.
 
@@
     \caption{PanDA Broker architecture as deployed on Titan. Numbers
-    indicates the execution process of a detector simulation job, part of the
-    production ATLAS MC workflow.}
-\label{fig:panda_broker}
+    indicates the execution process of a detector simulation job described in
+    \S\ref{ssec:panda_titan}.}\label{fig:panda_broker}
 \end{figure}

### 2017-09-08 PandaOverview.tex
Comments to address before submission
Address reviewers' comments: please review/iterate

@@
+\ifreview
+Comments to address in this section:
+\begin{itemize}
+	\color{red} 
+	\item REVIEWER 2: use of wrong wordings, such as "eventful" instead of
+	the common term "stateful".\mtnote{Done.}
+\end{itemize}
+\fi
+
 PanDA is a Workload Management System (WMS) %~\cite{marco2009glite} 
 designed to support the execution of distributed workloads and workflows via
-pilots~\cite{turilli2015comprehensive}. Pilot-capable WMS enable high
+pilots~\cite{turilli2017comprehensive}. Pilot-capable WMS enable high
 throughput execution of tasks via multi-level scheduling while supporting
@@
 month, analyzing and producing petabytes of data.
-% The design of PanDA WMS started in 2005 to support ATLAS\@. PanDA went into
-% production for the LHC Run 1 on 2009, and was then extended with new
-% subsystems for Run 2 in 2015.
-
-% ---------------------------------------------------------------------------
-% \subsection{Design}
-% \label{ssec:panda_design}
-
-% PanDA's application model assumes tasks grouped into workloads and
-% workflows. Tasks represent a set of operations performed on data sets
-% stored in one or more input files. Tasks are decomposed into jobs, where
-% each job consists of the task's operations performed on a partition of the
-% task's data set. Since 2005, a certain amount of parallelism has been
-% progressively introduced for job execution~\cite{crooks2012multi} but, so
-% far, HEP applications have not required a message passing interface (MPI).
-
-% PanDA's usage model assumes multitenancy of resources and at least two
-% types of HEP users: individual researchers and groups executing so called
-% `production' workflows. PanDA's security model is based on separation
-% between authentication, authorization and accounting for both single users
-% and group of users. Both authentication and authorization are based on
-% digital certificates and on the virtual organization
-% abstraction~\cite{foster2001anatomy}.
-
-% Currently, PanDA's execution model is based on four main abstractions:
-% task, job, queue, and pilot. Both tasks and jobs are assumed to have
-% attributes and states and to be queued into a global queue for execution.
-% Prioritization and binding of jobs are assumed to depend on the attributes
-% of each job. Pilot is used to indicate the abstraction of resource
-% capabilities. Each job is thought to be bound to one pilot and executed on
-% the site where the pilot has been instantiated.
-
-% In PanDA's data model, each datum refers to the recorded or simulated
-% measurement of a physical process. Data can be packaged into files or other
-% containers. As with jobs, data have both attributes and states, and some of
-% the attributes are shared between events and jobs. Raw, reconstruction, and
-% simulation data are assumed to be distributed across multiple storage
-% facilities and managed by the ATLAS Distributed Data Management
-% (DDM)%~\cite{garonne2012atlas}. 
-% When necessary, datasets required by each job are assumed to be replicated
-% over the network, both for input and output data.
-
-% PanDA's design supports provenance and traceability for both jobs and data.
-% Attributes enable provenance by linking jobs and data items, providing
-% information like ownership or project affiliation. States enable
-% traceability by providing information about the stage of the execution in
-% which each job or data item is or has been.
-
-% ---------------------------------------------------------------------------
-% \subsection{Implementation and Execution}
-% \label{ssec:panda_arch}
 
 The implementation of PanDA WMS consists of several interconnected
-subsystems,
-% , most of them built from off-the-shelf and Open Source components.
-% Subsystems communicate 
-communicating via dedicated API or HTTP messaging, and implemented by one or
-more modules. Databases are used to store eventful entities like tasks, jobs
-and input/output data, and to store information about sites, resources,
-logs, and accounting.
+subsystems, communicating via dedicated API or HTTP messaging, and
+implemented by one or more modules. Databases are used to store stateful
+entities like tasks, jobs and input/output data, and to store information
+about sites, resources, logs, and accounting.
 
@@
 Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of
-ATLAS workflows % (e.g., PanDA Event Service~\cite{calafiura2015atlas}), 
-but we do not discuss them as they are not relevant to an understanding of
-how PanDA has been ported to supercomputers. For a full list of subsystems
-see Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture} shows a
-diagrammatic representation of PanDA main subsystems, highlighting the
-execution process of tasks while omitting monitoring details to improve
+ATLAS workflows but we do not discuss them as they are not relevant to an
+understanding of how PanDA has been ported to supercomputers. For a full list
+of subsystems see Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture}
+shows a diagrammatic representation of PanDA main subsystems, highlighting
+the execution process of tasks while omitting monitoring details to improve
 readability.
 
-% During LHC Run 1, PanDA required users to perform a static conversion
-% between tasks and jobs: tasks were described as a set of jobs and then
-% submitted to the PanDA Server. This introduced inefficiency both with
-% usability and resource utilization.%~\cite{borodin2015unified}. Ideally,
-% users should conceive analyses in terms of one or more, possibly related
-% tasks, while the execution manager (i.e., PanDA) should partition tasks
-% into jobs, depending on execution constraints. Further, due to the static
-% partitioning of tasks into jobs, PanDA instantiates pilots on sites without
-% taking into account the heterogeneity and dynamicity of the resources. An
-% optimal sizing of each job should take into account these properties.
-
-% Another problem of static job sizing is that PanDA instantiates pilots on
-% sites with different type of resources and different models of availability
-% of those resources. An optimal sizing of each job should take into account
-% these properties. For example, sites may offer cores with different speed,
-% networking with different amount of bandwidth, and resources could be
-% guaranteed to be available for a defined amount of time or could disappear
-% at any point in time as it happens with opportunistic models of resource
-% provision.
-
-% JEDI was deployed for the LHC Run 2 to address these inefficiencies. 
 Users submit task descriptions to JEDI (Fig.~\ref{fig:architecture}:1) that
@@
 bound to the site of that pilot and ready to be executed. Task Buffer checks
-the global queue (i.e., the PanDA database) and, upon availability, returns a
-job to the Job Dispatcher. The Job Dispatcher dispatches that job to the
-PanDA Pilot (Fig.~\ref{fig:architecture}:12).
+the global queue (i.e., the PanDA DB) and, upon availability, returns a job
+to the Job Dispatcher. The Job Dispatcher dispatches that job to the PanDA
+Pilot (Fig.~\ref{fig:architecture}:12).
 
-A PanDA Pilot starts a monitoring process on receiving a job and forks a
-subprocess to execute the job's payload. Input data are transferred
-from the stage-in location (Fig.~\ref{fig:architecture}:13), the job's
-payload is executed (Fig.~\ref{fig:architecture}:14) and once completed, output
-is transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).
+Each PanDA Pilot starts a monitoring process on receiving a job and forks a
+subprocess to execute the job's payload. Input data are transferred from the
+stage-in location (Fig.~\ref{fig:architecture}:13), the job's payload is
+executed (Fig.~\ref{fig:architecture}:14) and once completed, output is
+transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).
 
@@
     \caption{PanDA WMS architecture. Numbers indicates the JEDI-based
-    execution process. Several subsystems, components, and architectural and
-    communication details are abstracted to improve clarity.}
+    execution process described in \S\ref{sec:panda_overview}. Several
+    subsystems, components, and architectural and communication details are
+    abstracted to improve clarity.}
 \label{fig:architecture}

### 2017-09-08 bckground.tex
Comments to address before submission
Address reviewers' comments: please review/iterate

@@
-% \jhanote{In this section we should talk less about PanDA and more about other
-% WMS -- highlighting the fact that (i) there isn't a comprehensive systems
-% (design, architecture and execution properties) discussion of other WMS so
-% difficult to compare}\mtnote{Better?}
-
-% \jhanote{(ii) real comparison should be to the Gliden-in WMS ecosystem and use
-% this as opportunity to narrate the reasons for the divergence.}\mtnote{From our
-% meetings with Sergey: Before Run 1, ATLAS management become uncomfortable with
-% the pace at which other WMS were being developed, loosing trust in their ability
-% to deliver production-grade capabilities. ATLAS USA decided to go solo and
-% developed their own solution. PanDA proven to be able to deliver the
-% usual pilot-based WMS capabilities at the right scale. For that, it was chosen
-% over other solutions.}
-
-% \jhanote{Also might this section be moved towards the rear, as this is  a
-% practice paper and not a traditional research paper?}\mtnote{Done.}
+\ifreview
+Comments to address in this section:
+\begin{itemize}
+	\color{red} 
+	\item REVIEWER 2: the topic of other HPC usage models [OCCAM,FRHPC],
+	elastic simulation jobs [ELASTICMC], opportunistic HPC usage [ATLASHPC]
+	or proxy infrastructure [ACTATLAS,ACTHYDRA] is not discussed.
+\end{itemize}
+\fi
 
 Several pilot-enabled WMS were developed for the LHC experiments:
-AliEn~\cite{Bagnasco2010} for ALICE; DIRAC~\cite{Paterson2010} for LHCb;
-GlideinWMS~\cite{sfiligoi2008glideinwms} for CMS; and
-PanDA~\cite{maeno2014evolution} for ATLAS. These systems implement similar
+AliEn~\cite{Bagnasco2010} for ALICE\@; DIRAC~\cite{Paterson2010} for LHCb;
+GlideinWMS~\cite{sfiligoi2008glideinwms} for CMS\@; and
+PanDA~\cite{maeno2014evolution} for ATLAS\@. These systems implement similar
 design and architectural principles: centralization of task and resource
@@
 
-AliEn, DIRAC, GlideinWMS and PanDA all share a similar design with two types of
-components: the management ones facing the application layer and centralizing
-the capabilities required to acquire tasks' descriptions and matching them to
-resource capabilities; and resource components used to acquire compute and data
-resources and information about their capabilities. Architecturally, the
-management components include one or more queue and a scheduler that coordinates
-with the resource modules via push/pull protocols. All resource components
-include middleware-specific APIs to request for resources, and a pilot capable
-of pulling tasks from the management modules and executing them on its
-resources.
+AliEn, DIRAC, GlideinWMS and PanDA all share a similar design with two types
+of components: the management ones facing the application layer and
+centralizing the capabilities required to acquire tasks' descriptions and
+matching them to resource capabilities; and resource components used to
+acquire compute and data resources and information about their capabilities.
+Architecturally, the management components include one or more queue and a
+scheduler that coordinates with the resource modules via push/pull protocols.
+All resource components include middleware-specific APIs to request for
+resources, and a pilot capable of pulling tasks from the management modules
+and executing them on its resources.
 
@@
 AliEn, DIRAC, GlideinWMS and PanDA also have similar implementations. These
-WMS were initially implemented to use Grid resources using the Condor software
-ecosystem~\cite{thain2005distributed}. Accordingly, all LHC WMS implemented
-Grid-like authentication and authorization systems and adopted a computational
-model based on distributing a large amount of single/few-cores tasks across
-hundreds of sites.
+WMS were initially implemented to use Grid resources using the Condor
+software ecosystem~\cite{thain2005distributed}. Accordingly, all LHC WMS
+implemented Grid-like authentication and authorization systems and adopted a
+computational model based on distributing a large amount of single/few-cores
+tasks across hundreds of sites.
 
@@
 collisions in the accelerator and from their simulations. Dedicated,
-multi-tiered data systems have been built to store, replicate, and distributed these
-data. All LHC WMS interface with these systems to move data to the sites where
-related compute tasks are executed or to schedule compute tasks where (large
-amount of) data are already stored.
+multi-tiered data systems have been built to store, replicate, and
+distributed these data. All LHC WMS interface with these systems to move data
+to the sites where related compute tasks are executed or to schedule compute
+tasks where (large amount of) data are already stored.
 
@@
 ATLAS on Titan.
-
-% -----------------------------------------------------------------------------
-% NOTES ON TASK, JOBS, AND WORKFLOW TERMINOLOGY
-% -----------------------------------------------------------------------------
-% Currently, the implementation of PanDA WMS has at least two distinguishing
-% features: the scale at which it operates distributed computing, and
-% network-aware scheduling of tasks to sites\mtnote{PanDA team: please let me
-% know whether more unique (i.e., not present in other LHC WMS) features of
-% PanDA implementation should be indicated}. PanDA concurrently supports
-% distributed computations on up to 240,000 cores, submitting more than 365
-% million jobs a year across 100 sites, and processing up to XXXXPB of data
-% distributed worldwide on XX Tier 1-3 facilities. As for 2017, PanDA serves
-% several thousand users, including XX production groups. The scale of this
-% operation is unprecedented, making PanDA the WMS that manages the largest
-% computational campaign in the world.
-
-% PanDA was initially implemented to support computations across an unreliable
-% and relatively expensive networking infrastructure. Accordingly, PanDA
-% assumed a rigid model of data placement based on replication and long-term
-% storage of data on the Grid sites. This model proven too wasteful, requiring
-% storage space for unused and obsolete data. PanDA implementation was evolved
-% to benefit from the development of the internet and of the networking
-% infrastructure for scientific research. Today, PanDA constantly monitors
-% network throughput and latency among its management components, Grid sites and
-% data facilities. PanDA uses this information to schedule tasks on sites,
-% depending on the estimated time required to download and/or replicate input
-% data and to stage out output data.
-
-% The dynamic management of data replication and download is an example of a
-% generalized paradigm shift that distinguishes the implementation of PanDA from
-% other WMS. New components are being implemented to support: (i) dynamic sizing
-% of input dataset based on resource capabilities; (ii) dynamic sizing of the
-% resources held by pilots both in terms of number of cores and duration of the
-% pilot; and (iii) further abstraction of different types of resources to
-% reconcile high-throughput and high-performance computing paradigms.
-% -----------------------------------------------------------------------------
-
-% -----------------------------------------------------------------------------
-% NOTES ON TASK, JOBS, AND WORKFLOW TERMINOLOGY
-% -----------------------------------------------------------------------------
-% Each task may have an arbitrary number of properties like number of cores,
-% executables, or input/output files. Tasks may have precedence interrelations,
-% depending on their data dependences or any other type of dependency mandated
-% by the application algorithms. Tasks with precedence relations have to be
-% executed serially; otherwise, tasks can be executed concurrently. A workload
-% is defined as a set of tasks that can be executed concurrently. As such, a
-% workflow can be composed of a set of workloads.
-
-% The terms ``task'' and ``job'' are also used inconsistently across communities
-% that perform scientific computing. In this paper, ``job'' refers to the unit
-% of work that is submitted to a local resource management system (LRMS), like
-% the Slurm or PBS batch system of a cluster. As such, a task can become a job
-% when is scheduled on a resource that exposes a LRMS but can also become a
-% virtual machine or a container when bootstrapped on an infrastructure
-% supporting virtualization. Tasks can be statically or dynamically grouped
-% into jobs, depending on the resource capabilities and the task requirements.
-
-% Usually, workflows are represented as graphs in which tasks are vertices and
-% relations are edges~\cite{}. Often, graphs are supposed to be acyclic but
-% graphs with cycles have been used to represent workflows of
-% workflows~\cite{}. In this paper, a workflow template is a type of graph
-% while a workflow instance is a workflow template with specific vertices and
-% edges. Further, a workflow instance is ``abstract'' when no resource
-% properties are available for all vertices, ``concrete'' otherwise~\cite{}.
-% When not qualified, the term ``workflow'' indicates a abstract workflow
-% instance.
-
-% Notes for refactoring:
-% \begin{itemize}
-%     \item Naming and nomenclature is a contribution
-%     \item 1, 1.5 pages.
-% \end{itemize}
-% -----------------------------------------------------------------------------

### 2017-09-08 nge.tex
Address reviewers' comments: please review/iterate

@@
 	the project started, it should be explained why a pilot job approach was
-	not taken in the first place.
+	not taken in the first place.\mtnote{Added an opening paragraph recalling
+	the explanation offered in \S\ref{sec:panda_titan}}.
 	\item REVIEWER 3: Figure 6. refers to a "User Workstation" as the running
@@
 	is not the case and separate workstations are being used for the testing
-	phase, this should be made clear.)
+	phase, this should be made clear.)\mtnote{Fixed.}
 \end{itemize}
 \fi
+
+As explained in \S\ref{sec:panda_titan}, PanDA Broker was designed to
+maximize code reutilization of PanDA Pilot. This allowed for rapid adoption
+and incremental optimization while enabling the parallel development of a
+more general solution for executing ATALS workloads on HPC resources.
 
@@
 to Titan's PBS batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash
-arrow). %, as done by PanDA Broker to submit MPI scripts. 
-Once scheduled, the Pilot Agent is bootstrapped on a MOM node and the
+arrow). Once scheduled, the Pilot Agent is bootstrapped on a MOM node and the
 Agent's Executors on the worker nodes.
 
-The UnitManager and the Pilot Agent communicate via a database instantiated on a
-DTN so as to be reachable by both modules. The UnitManager schedules units to
-the Agent's Scheduler (Fig.~\ref{fig:arch-overview}, solid arrow) and the
-Agent's Scheduler schedules the units on one or more Agent's Executor.
-% for execution. The Agent's executors can manage one or more worker nodes,
-% depending on performance evaluations.
+The UnitManager and the Pilot Agent communicate via a database instantiated
+on a DTN so as to be reachable by both modules. The UnitManager schedules
+units to the Agent's Scheduler (Fig.~\ref{fig:arch-overview}, solid arrow)
+and the Agent's Scheduler schedules the units on one or more Agent's
+Executor.
 
-The Pilot Agent uses the Open Run-Time Environment (ORTE) for communication and
-coordination of the execution of units. This environment is a critical
-component of the OpenMPI implementation~\cite{castain05:_open_rte}.
-%, cug-2016}.
-ORTE is able to minimize the system overhead while submitting tasks by
-avoiding filesystem bottlenecks and race conditions with network sockets.
+The Pilot Agent uses the Open Run-Time Environment (ORTE) for communication
+and coordination of the execution of units. This environment is a critical
+component of the OpenMPI implementation~\cite{castain05:_open_rte}. ORTE is
+able to minimize the system overhead while submitting tasks by avoiding
+filesystem bottlenecks and race conditions with network sockets.

### 2017-09-08 overall.tex
Address reviewers' comments: please review/iterate

@@
     \color{red}
+
     \item REVIEWER 2: What are the reasons for the low efficiency of PanDA
     Broker? Why are only 30\% of backfill slots useable, if they are
-    basically uncontested? Why does Fig 3 show an increase of absolute and
+    basically uncontested?
+
+    \item REVIEWER 2: Why does Fig 3 show an increase of absolute and
     relative efficiency in Aug-16, when the text claims this in Sep-16 due to
-    more PanDA Brokers? Why are only 20 PanDA Brokers deployed if this is a
+    more PanDA Brokers?\mtnote{Fixed.}
+
+    \item REVIEWER 2: Why are only 20 PanDA Brokers deployed if this is a
     limiting factor? Why is it questionable that more PanDA Brokers would
     increase efficiency? If only 5/6 of resources below the mean are
-    unusable, this implies an efficiency of almost 60\% being reachable.
+    unusable, this implies an efficiency of almost 60\% being
+    reachable.\mtnote{From our text: ``further research is ongoing to
+    understand whether an even greater number of brokers would yield even
+    greater efficiency. This research focuses on evaluating the overheads of
+    input/output files staging, including its impact on DTNs, and on an
+    alternative design of PanDA Broker that enables the concurrent submission
+    of multiple MPI scripts''. I believe this is an answer good enough.}
+
     \item REVIEWER 2: Given an expected increase in resource need of 10-100,
@@
     major contrast to the requirements sketched in the introduction.
+
     \item REVIEWER 2: How does the poor metadata read performance of
-    1500-6300 s (25\%-100\% of minimum walltime!) affect efficiency? Is it
-    factored into the choice of walltime requests? Is it applicable only to
-    TITAN, to HPC in general, or even to regular WLCG sites? Are there
+    1500-6300s (25\%-100\% of minimum walltime!) affect efficiency? Is it
+    factored into the choice of walltime requests?\mtnote{Asked Sergey and
+    Danila to confirm these numbers and walltime request policy.} Is it
+    applicable only to TITAN, to HPC in general, or even to regular WLCG
+    sites?\mtnote{This is clearly out of scope for this paper.} Are there
     solutions at application, e.g. the use of packaged python scripts and
-    libraries? The information presented is insufficient to estimate whether
-    this issue applies to other environments and usecases as well, or is
-    simply due to suboptimal implementation/configuration.
+    libraries?\mtnote{Added the following: ``While these figures are specific
+    to OLCF and TItan, reengineering on AthenaMP could further improve
+    startup time on every platform.''} The information presented is
+    insufficient to estimate whether this issue applies to other environments
+    and usecases as well, or is simply due to suboptimal
+    implementation/configuration.
+
     \item REVIEWER 3: The HEP-SPEC06 benchmark, which is used as a
@@
     contention is responsible for 50\% decrease in compute efficiency on
-    Titan when compared to other WLCG sites (paper page 6).
+    Titan when compared to other WLCG sites (paper page 6).\mtnote{Wrote an email to Sergey asking for a clarification.}
 \end{enumerate}
@@
 
-% without affecting other users.
-
 We evaluate the efficiency, scalability and reliability of the deployment of
@@
 % ---------------------------------------------------------------------------
-\subsection{Characterizing the PanDA Broker on Titan}
-\label{ssec:broker_titan}
+\subsection{Characterizing the PanDA Broker on Titan}\label{ssec:broker_titan}
 
@@
 ATLAS (blue bars) during the experiment time window. ATLAS consumed a total
-of 51.4M core-hours, for an average of $\approx$3.7M core-hours a month. 
-% , with a minimum of 1.7M core-hours in April 2016 and a maximum 7.9M
-% core-hours in February 2017.
+of 51.4M core-hours, for an average of \(\approx\)3.7M core-hours a month.
 
@@
     percentage of total Titan's backfill availability used by ATLAS (Red
-    labels).}
-\label{fig:backfill-utilization}
+    labels).}\label{fig:backfill-utilization}
 \end{figure}
@@
 function of the number of events simulated but not all events take the same
-time to be simulated. One event simulation takes from $\approx$2 to
-$\approx$40 minutes, with a mean of $\approx$14 minutes. Considering that
+time to be simulated. One event simulation takes from \(\approx\)2 to
+\(\approx\)40 minutes, with a mean of \(\approx\)14 minutes. Considering that
 each worker node process up to 16 events concurrently, 100 events takes an
@@
 
-%\begin{figure*}[!t]
 \begin{figure}[!t]
-    % \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_wnodes_distribution.pdf}
-    % \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_walltime_distribution.pdf}
-    % \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
     \includegraphics[clip,width=\columnwidth]{figures/titan_backfill_avail.png}
     \vspace{-0.1in}
-    \caption{
-    % Backfill availability: distribution of number of work nodes (left) and
-    % of walltime in minutes (center). Scatter plot of the two variables.
-    % (right).
-    62555 measures of Backfill availability on Titan during the experiment
-    time window. Mean number of work nodes available 691; mean walltime
-    available 126 minutes.}
-\label{fig:backfill-distrib}
-%\end{figure*}
+    \caption{62555 measures of Backfill availability on Titan during the
+    experiment time window. Mean number of work nodes available 691; mean
+    walltime available 126 minutes.}\label{fig:backfill-distrib}
 \end{figure}
 
-% Usually, detector simulations performed on the Grid process 10 times the
-% number of events than those on Titan. This explains the difference between
-% the percentage of detector simulations and of events computed by those
-% simulations on Titan.
 PanDA Broker could fit the number of events to the walltime of each available
@@
 
-In September 2016, increasing the number of concurrent PanDA brokers from 4
-to 20 markedly improved efficiency (see Fig.~\ref{fig:backfill-utilization})
-but further research is ongoing to understand whether an even greater
-number of brokers would yield even greater efficiency. This research focuses on
+In August 2016, increasing the number of concurrent PanDA brokers from 4 to
+20 markedly improved efficiency (see Fig.~\ref{fig:backfill-utilization}) but
+further research is ongoing to understand whether an even greater number of
+brokers would yield even greater efficiency. This research focuses on
 evaluating the overheads of input/output files staging, including its impact
@@
 concurrent submission of multiple MPI scripts~\cite{barreiro2016panda}. The
-understanding will contribute to improving the efficiency of
-PanDA Brokers beyond the 30\% limit showed in
-Fig.~\ref{fig:backfill-utilization}.
+understanding will contribute to improving the efficiency of PanDA Brokers
+beyond the 30\% limit showed in Fig.~\ref{fig:backfill-utilization}.
 
@@
 PanDA Server, and the 13\% failures produced by the Geant4 toolkit.
-% The current failure rate of the PanDA Brokers confirms the benefits of
-% reusing most of the code base of the PanDA Pilot for implementing the PanDA
-% Broker. It also shows that adopting third-party libraries like RADICAL-SAGA
-% did not have a measurable adverse effect on reliability.
 
 % ---------------------------------------------------------------------------
-\subsection{Characterizing the Detector Simulation on Titan}
-\label{ssec:athenamp_titan}
+\subsection{Characterizing the Detector Simulation on
+Titan}\label{ssec:athenamp_titan}
 
@@
 simulation jobs submitted to Titan: (i) the time taken to setup AthenaMP\@;
-and (ii) the distribution of the time taken by Geant4 to simulate
-a certain number of events.
+and (ii) the distribution of the time taken by Geant4 to simulate a certain
+number of events.
 
 AthenaMP has an initialization and configuration stage. At initialization
-time, AthenaMP is assembled from a set of shared libraries, depending on the
-type of payload that will have to be computed. Once initialized, every
-algorithm and service of AthenaMP is configured via Python scripts. Both
-these operations result in read operations on the filesystem shared between
-the worker nodes and the DTNs, including the operations required to access
-small python scripts.
+time, AthenaMP is assembled from a very large set of shared libraries,
+depending on the type of payload that will have to be computed. Once
+initialized, every algorithm and service of AthenaMP is configured via Python
+scripts. Both these operations result in read operations on the filesystem
+shared between the worker nodes and the DTNs, including the operations
+required to access small python scripts.
 
@@
 eliminated the problem of metadata contention, improving metadata read
-performance from $\approx$6,300 seconds on Lustre to $\approx$1,500 seconds
-on NFS\@.
+performance from \(\approx\)6,300 seconds on Lustre to \(\approx\)1,500
+seconds on NFS\@. While these figures are specific to OLCF and TItan, reengineering on AthenaMP could further improve startup time on every platform.
 
@@
     simulation use 16 cores per node, 1 simulation 8. 50 Events; 1 Titan
-    worker nodes; 16 work threads per node; 100 events per node.}
-
-\label{fig:comparison-8-16cores}
+    worker nodes; 16 work threads per node; 100 events per
+    node.}\label{fig:comparison-8-16cores}
 \end{figure}
@@
 also to compare Titan and Grid site performances. Usually, Grid sites exposes
-resources with heterogeneous CPU architectures and a maximum of 8 (virtual)
-cores per worker node, while Titan's offer an homogeneous 16 cores
-architecture. We used the rate of events processes per minute as a measure of
-the efficiency of executing the same detector simulation on Titan or Grid
-sites. Comparisons of the efficiencies of Titan to the BNL and SIGNET Grid
-sites, normalized for 8 cores, show that the effective performance per-core
-at Titan is $\approx$0.57 event per minute, roughly 1/2 of BNL and  1/3 of
-SIGNET performances.
+resources with heterogeneous CPU architectures and 8 (virtual) cores, while
+Titan's offer an homogeneous 16 cores architecture. We used the rate of
+events processes per minute as a measure of the efficiency of executing the
+same detector simulation on Titan or Grid sites. Comparisons of the
+efficiencies of Titan to the BNL and SIGNET Grid sites, normalized for 8
+cores, show that the effective performance per-core at Titan is
+\(\approx\)0.57 event per minute, roughly 1/2 of BNL and  1/3 of SIGNET
+performances.
 
@@
 \begin{tabular}{lllllllll} & Num. Nodes & Duration (s) & Read (GB) & Written
- (GB) & GB Read/nodes & GB Written/nodes & $open()$ & $close()$ \\ Min & 1 &
+ (GB) & GB Read/nodes & GB Written/nodes & \(open()\) & \(close()\) \\ Min & 1 &
  1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\ Max & 300 & 7,452 &
@@
 \caption{The Statistical breakdown of the I/O impact of 1,175 jobs ATLAS
-executed at OLCF for the week of 10/25/16}
-\label{panda-olcf-stats}
+executed at OLCF for the week of 10/25/16}\label{panda-olcf-stats}
 \end{table*}
@@
 
-The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
-operations per second (not including file $stat()$ operations) and 5 file
-$close()$ operations per second, with similar distributions. On average, the
-maximum number of file $open()$ operations per job is $\approx$170/s and the
-maximum number of file $close()$ operations is $\approx$39/s. For the 1,175
-ATLAS jobs observed, the total number of file $open()$ operations is
-172,089,760 and the total number of file $close()$ operations is 40,132,992.
-The difference between these two values is still under investigation: One
-possible explanation is that ATLAS jobs don't call a file $close()$ operation
-per every file $open()$ issued.
+The metadata I/O breakdown shows that ATLAS jobs yield 23 file \(open()\)
+operations per second (not including file \(stat()\) operations) and 5 file
+\(close()\) operations per second, with similar distributions. On average,
+the maximum number of file \(open()\) operations per job is \(\approx\)170/s
+and the maximum number of file \(close()\) operations is \(\approx\)39/s. For
+the 1,175 ATLAS jobs observed, the total number of file \(open()\) operations
+is 172,089,760 and the total number of file \(close()\) operations is
+40,132,992. The difference between these two values is still under
+investigation: One possible explanation is that ATLAS jobs don't call a file
+\(close()\) operation per every file \(open()\) issued.
 

### 2017-09-08 bckground.tex
Related work review and comment

@@
 	elastic simulation jobs [ELASTICMC], opportunistic HPC usage [ATLASHPC]
-	or proxy infrastructure [ACTATLAS,ACTHYDRA] is not discussed.
+	or proxy infrastructure [ACTATLAS,ACTHYDRA] is not discussed.\mtnote{I
+	looked for these models and found few references to them. All in all, I
+	am not seeing how these fit within the background of pilot-enabled
+	workload managers.}
 \end{itemize}

### 2017-09-09 overall.tex
Iteration on the base of Sergey's comments

@@
     1500-6300s (25\%-100\% of minimum walltime!) affect efficiency? Is it
-    factored into the choice of walltime requests?\mtnote{Asked Sergey and
-    Danila to confirm these numbers and walltime request policy.} Is it
-    applicable only to TITAN, to HPC in general, or even to regular WLCG
-    sites?\mtnote{This is clearly out of scope for this paper.} Are there
-    solutions at application, e.g. the use of packaged python scripts and
+    factored into the choice of walltime requests?\mtnote{Thanks to Sergey's
+    input I corrected the figure previously reported.} Is it applicable only
+    to TITAN, to HPC in general, or even to regular WLCG sites?\mtnote{This
+    is clearly out of scope for this paper.} Are there solutions at
+    application, e.g. the use of packaged python scripts and
     libraries?\mtnote{Added the following: ``While these figures are specific
@@
     contention is responsible for 50\% decrease in compute efficiency on
-    Titan when compared to other WLCG sites (paper page 6).\mtnote{Wrote an email to Sergey asking for a clarification.}
+    Titan when compared to other WLCG sites (paper page 6).\mtnote{Thanks to
+    Sergey's input I downplayed the role played by FP and up-played the role
+    played by new CPUs.}
 \end{enumerate}
@@
 eliminated the problem of metadata contention, improving metadata read
-performance from \(\approx\)6,300 seconds on Lustre to \(\approx\)1,500
-seconds on NFS\@. While these figures are specific to OLCF and TItan, reengineering on AthenaMP could further improve startup time on every platform.
+performance from a worse case scenario of 6,300s on Lustre to \(\approx\)225s
+on NFS\@. While these figures are specific to OLCF and TItan, reengineering
+of AthenaMP could further improve startup time on every platform.
 
@@
 The differences in performance between Titan and the BNL and SIGNET Grid
-sites are due to the FP scheduler competition and the availability of newer
-processors. The CPUs at the Grid sites have one FP scheduler per core and are
-on average newer than the CPU of Titan. The heterogeneity of the Grid sites'
-CPUs explain the higher performance variance we observed compared to the
-performance consistency we measured on Titan. This difference in performance
-is compensated by the amount of resources available on Titan (capable of
-executing up to 30\%/year of the ATLAS detector simulations) and by the lack
-of further resources available to WLCG\@. Also, it should be noted that Titan
-is at the end of its life-cycle and that Summit, Titan's successor, will
-offer cutting-edge performances.
+sites are due to the FP scheduler competition but mainly to the availability
+of newer processors. 
+% The CPUs at the Grid sites have one FP scheduler per core and are on
+% average newer than the CPU of Titan.
+The heterogeneity of the Grid sites' CPUs explain the higher performance
+variance we observed compared to the performance consistency we measured on
+Titan. This difference in performance is compensated by the amount of
+resources available on Titan (capable of executing up to 30\%/year of the
+ATLAS detector simulations) and by the lack of further resources available to
+WLCG\@. Also, it should be noted that Titan is at the end of its life-cycle
+and that Summit, Titan's successor, will offer cutting-edge performances.
 

### 2017-09-09 introduction.tex
Address reviewers' comments: please review/iterate
Introduction - addressing a reviewer comment.
  "ain't no cure for reviewers with a vendetta"..

@@
 	obvious comparisons, it raises the question how much other work is
-	ignored.
+	ignored. \jhanote{Added "academic" after arguably. Any more?}
 \end{enumerate}
@@
 system~\cite{Maeno2011} (WMS) for distributed data processing and analysis.
-The CMS and ATLAS experiments utilize, arguably the largest production grade
-distributed computing solutions, and have symbolized the paradigm of {\it
-high-throughput computing} (HTC), i.e., the effective execution of many
-independent tasks.
+The CMS and ATLAS experiments utilize, arguably the largest academic
+production grade distributed computing solutions, and have symbolized the
+paradigm of {\it high-throughput computing} (HTC), i.e., the effective
+execution of many independent tasks.
 
@@
 significantly outstrip current and projected supply.   The data volumes that
-will need analyzing in LHC-Run 3 ($~\approx$ 2022) and the high-luminosity era
-(Run 4) will increase by factors of 10--100 compared to the current phase (Run
-2). There are multiple levels at which this problem needs to be addressed: the
-utilization of emerging parallel architectures (e.g., platforms); algorithmic
-and advances in analytical methods (e.g., use of Machine Learning); and the
-ability to exploit different platforms (e.g., clouds and supercomputers).
+will need analyzing in LHC-Run 3 ($\approx$ 2022) and the high-luminosity
+era (Run 4) will increase by factors of 10--100 compared to the current phase
+(Run 2). There are multiple levels at which this problem needs to be
+addressed: the utilization of emerging parallel architectures (e.g.,
+platforms); algorithmic and advances in analytical methods (e.g., use of
+Machine Learning); and the ability to exploit different platforms (e.g.,
+clouds and supercomputers).
 
@@
 thus increasing the overall utilization on Titan without loss of overall
-quality-of-service. The conventional mix of jobs at OLCF cannot be effectively
-backfilled because of size, duration, and scheduling policies. Our approach is
-extensible to any HPC with ``capability scheduling'' policies.  We also
-investigate the use of a pilot-abstraction based task execution runtime system
-to flexibly execute ATLAS and other heterogeneous workloads (molecular
-dynamics) using regular queues. As such, our approach provides a general
-solution and investigation of the convergence of HPC and HTC execution of
-workloads.
+quality-of-service. The conventional mix of jobs at OLCF cannot be
+effectively backfilled because of size, duration, and scheduling policies.
+Our approach is extensible to any HPC with ``capability scheduling''
+policies.  We also investigate the use of a pilot-abstraction based task
+execution runtime system to flexibly execute ATLAS and other heterogeneous
+workloads (molecular dynamics) using regular queues. As such, our approach
+provides a general solution and investigation of the convergence of HPC and
+HTC execution of workloads.
 
@@
 demonstrating how distributed workflows can be integrated with leadership
-resources, and effectively accommodating HTC and HPC workloads simultaneously.
+resources, and effectively accommodating HTC and HPC workloads
+simultaneously.
 

### 2017-09-09 paper.tex
Address reviewers' comments: please review/iterate
Introduction - addressing a reviewer comment.
  "ain't no cure for reviewers with a vendetta"..

@@
     \color{red} 
-    \item REVIEWER 1: There is a problem with capitalization in some
-    reference titles, ie. [2], [3], [4], [7], [8], [9], [10], [11], [12],
-    [13]. Reference [6] is just an author list, year and title. Should there
-    be more here?
+    % \item REVIEWER 1: There is a problem with capitalization in some
+    % reference titles, ie. [2], [3], [4], [7], [8], [9], [10], [11], [12],
+    % [13]. Reference [6] is just an author list, year and title. Should there
+    % be more here?\mtnote{Fixed.}
     \item REVIEWER 2:  Key concepts such as pilot jobs, tasks and workloads
@@
     [19] is a slide collection even though there are many papers for the
-    referenced work.
+    referenced work.\mtnote{Reference [19] is a technical paper.}
     \item REVIEWER 3: Title looks too general for what amounts to the
@@
 % ---------------------------------------------------------------------------
-% ACKNOWLEDGEMENTS
-% ---------------------------------------------------------------------------
-% \section*{Acknowledgements}\label{sec:ack}
-% \input{ack}
-
-% ---------------------------------------------------------------------------
 % REFERENCES
 % ---------------------------------------------------------------------------
-% \newpage
 \bibliographystyle{IEEEtran}
-\bibliography{IEEEabrv,bibliography}
+\bibliography{bibliography}
 

### 2017-09-09 bckground.tex
Acknowledgements.

@@
 	\item REVIEWER 2: the topic of other HPC usage models [OCCAM,FRHPC],
-	elastic simulation jobs [ELASTICMC], opportunistic HPC usage [ATLASHPC]
-	or proxy infrastructure [ACTATLAS,ACTHYDRA] is not discussed.\mtnote{I
-	looked for these models and found few references to them. All in all, I
-	am not seeing how these fit within the background of pilot-enabled
-	workload managers.}
+	elastic simulation jobs [ELASTICMC], opportunistic HPC usage [ATLASHPC] or
+	proxy infrastructure [ACTATLAS,ACTHYDRA] is not discussed.\mtnote{I looked
+	for these models and found few references to them. All in all, I am not
+	seeing how these fit within the background of pilot-enabled workload
+	managers.}\jhanote{As you've exercised diligence in checking, we should
+	ignore these suggestions (which are at best advisory, not mandatory)}
 \end{itemize}

### 2017-09-10 introduction.tex
Typo

@@
 heterogeneous resources, the effective utilization of resources and efficient
-execution of workloads present non-trivial challenges. Many software solutions
-have been developed in response to these challenges. The CMS experiment,
-devised a solution based around the HTCondor~\cite{thain2005distributed}
-software ecosystem. The ATLAS~\cite{Aad:2008} experiment utilizes the
-Production and Distributed Analysis (PanDA) workload management
-system~\cite{Maeno2011} (WMS) for distributed data processing and analysis.
-The CMS and ATLAS experiments utilize, arguably the largest academic
-production grade distributed computing solutions, and have symbolized the
-paradigm of {\it high-throughput computing} (HTC), i.e., the effective
-execution of many independent tasks.
+execution of workloads present non-trivial challenges. Many software
+solutions have been developed in response to these challenges. The CMS
+experiment, devised a solution based around the
+HTCondor~\cite{thain2005distributed} software ecosystem. The
+ATLAS~\cite{Aad:2008} experiment utilizes the Production and Distributed
+Analysis (PanDA) workload management system~\cite{Maeno2011} (WMS) for
+distributed data processing and analysis. The CMS and ATLAS experiments
+utilize, arguably the largest academic production grade distributed computing
+solutions, and have symbolized the paradigm of {\it high-throughput
+computing} (HTC), i.e., the effective execution of many independent tasks.
 
@@
 free'' of the traditional computational approach of high-throughput computing
-on distributed resources to embrace new platforms, in particular high-
-performance computers (HPC). Specifically, we discuss the experience of
+on distributed resources to embrace new platforms, in particular
+high-performance computers (HPC). Specifically, we discuss the experience of
 integrating PanDA WMS with a US DOE leadership machine (Titan) to reach

### 2017-09-10 paper.tex
fixing  title, comments at start.

@@
 
-\title{Converging High-Throughput and High-Performance Computing: A Case Study}
+\title{High-Throughput Computing on High-Performance Platforms: A Case Study}
+
+%\title{Converging High-Throughput and High-Performance Computing: A Case Study}
 
@@
     are not meaningfully defined.
-    \item REVIEWER 2: the information provided is insufficient to asses in
-    how far any of the presented evaluation, characterization or lessons are
-    applicable outside of TITAN and PanDA Broker combined.
-    \item REVIEWER 2: Several references are of poor quality. For example,
-    [19] is a slide collection even though there are many papers for the
-    referenced work.\mtnote{Reference [19] is a technical paper.}
-    \item REVIEWER 3: Title looks too general for what amounts to the
-    implementation of an adapter component in a system that is very specific
-    to a single HEP experiment. (In general, title, abstract and conclusions
-    seem to over-state the contributions of this paper, considering this is a
-    general eScience conference and not a WLCG-specific one.)
+    % \item REVIEWER 2: the information provided is insufficient to asses in
+    % how far any of the presented evaluation, characterization or lessons are
+    % applicable outside of TITAN and PanDA Broker combined.
+    % \item REVIEWER 2: Several references are of poor quality. For example,
+    % [19] is a slide collection even though there are many papers for the
+    % referenced work.\mtnote{Reference [19] is a technical paper.}
+    % \item REVIEWER 3: Title looks too general for what amounts to the
+    % implementation of an adapter component in a system that is very specific
+    % to a single HEP experiment. (In general, title, abstract and conclusions
+    % seem to over-state the contributions of this paper, considering this is a
+    % general eScience conference and not a WLCG-specific one.)
 \end{itemize}

### 2017-09-10 introduction.tex
Typo

@@
 significantly outstrip current and projected supply.   The data volumes that
-will need analyzing in LHC-Run 3 ($\approx$ 2022) and the high-luminosity
+will need analyzing in LHC-Run 3 ($\approx$2022) and the high-luminosity
 era (Run 4) will increase by factors of 10--100 compared to the current phase

### 2017-09-10 introduction.tex
pending.

--- 2017-09-10 introduction.tex
@@
-\ifreview
-Comments to address in this section:
-\begin{enumerate}
-	\color{red} 
-	\item REVIEWER 1: The introduction is overly verbose and prosaic. The
-	praise for LHC collaborations is off-putting, especially as it ignores
-	other work. For example, claiming "CMS and ATLAS experiments utilize
-	arguably the largest production grade distributed computing solutions"
-	disregards the obvious commercial contenders. By blatantly excluding
-	obvious comparisons, it raises the question how much other work is
-	ignored. \jhanote{Added "academic" after arguably. Any more?}
-\end{enumerate}
-\fi
+% \ifreview
+% Comments to address in this section:
+% \begin{enumerate}
+% 	\color{red} 
+% 	\item REVIEWER 1: The introduction is overly verbose and prosaic. The
+% 	praise for LHC collaborations is off-putting, especially as it ignores
+% 	other work. For example, claiming "CMS and ATLAS experiments utilize
+% 	arguably the largest production grade distributed computing solutions"
+% 	disregards the obvious commercial contenders. By blatantly excluding
+% 	obvious comparisons, it raises the question how much other work is
+% 	ignored. \jhanote{Added "academic" after arguably. Any more?}
+% \end{enumerate}
+% \fi
 

### 2017-09-10 PandaEvolution.tex
ongoing cleanup.

@@
-\ifreview
-Comments to address in this section:
-\begin{enumerate}
-	\color{red} 
-	\item REVIEWER 2: the technical discussion of TITAN and WLCG differences
-	is abruptly stopped by the description of Monte Carlo event generation
-	and detector simulation (3A).\mtnote{I read the section twice and I did
-	not see such an abrupt transition.}
-	\item REVIEWER 2: The single listing (3B) of similar projects within the
-	collaboration does not detail whether this is derived or separate
-	work.\mtnote{Added a sentence to clarify that PanDA Broker operates
-	within the existing ATLAS software infrastructure. Specified that the
-	ATLAS team has ported PanDA Broker to other supercomputers.}
-	\item REVIEWER 2: How does the TITAN environment prevent usage of DTNs as
-	thin proxies? Is connectivity possible from compute node to DTNs via MPI,
-	message queues, sockets or other means? Are the restrictions a
-	shortcomming of TITAN, or the majority of HPC environments? Given the
-	proclaimed importance of HPC resources, a generic evaluation seems
-	needed.\mtnote{Added a sentence explaining that DTNs are not part of
-	Titan's compute nodes.}
-	\item REVIEWER 2: Why was the PanDA Broker designed orthogonal to the
-	pilot model, performing both resource acquisition and workload
-	execution?\mtnote{We explain that PanDA Broker is a prototype and now we
-	are developing NGE.} How do pilots conflict with backfill slots, when the
-	former are designed to acquire resources independent of workload
-	requirements?\mtnote{AFAIK, we never state that pilot conflict with
-	backfill. We argue that the latter mitigates the absence of the former.}
-	Why does each Broker instance query and use backfill separately? Given
-	that this seems to be a major bottleneck, the choice of such a design
-	begs for motivation.\mtnote{Extended a paragraph to explain that PanDA
-	Broker supported rapid deployment and incremental optimization while
-	incurring in functional limitations.}
-	\item REVIEWER 2: Some facts about WLCG seem outdated:
-	\begin{itemize}
-		\item The number of 30,000 unused TITAN cores (§IIIB) is equated to
-		roughly 10\% of the WLCG. However, WLCG currently has 591,282 cores
-		[REBUSCPU].
-		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{Done.}
-		\item WLCG sites are claimed to provide only 8 cores per worker node
-		(§IIIB). WLCG has a significant number of sites providing 16 or more
-		cores per worker node [REBUSCPU].
-		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{This
-		would require fairly deep rewriting on Section IVB. I edited the
-		sentence referring to 8 cores eliminating the reference to compute
-		nodes. Less than ideal but I did not find a better solution. Any
-		suggestion?}
-	\end{itemize}
-	\item REVIEWER 2: Schematic figures (Fig 1, Fig 2) are overly complex,
-	while their captions are not sufficient to understand them. The numbering
-	is only partially explained in the text.\mtnote{Added a reference to
-	where the numbering of the figures are explained in the text. I
-	double-checked and the numbering of both figure is fully explained in the
-	text.}
-\end{enumerate}
-\fi
+% \ifreview
+% Comments to address in this section:
+% \begin{enumerate}
+% 	\color{red} 
+% 	\item REVIEWER 2: the technical discussion of TITAN and WLCG differences
+% 	is abruptly stopped by the description of Monte Carlo event generation
+% 	and detector simulation (3A).\mtnote{I read the section twice and I did
+% 	not see such an abrupt transition.}
+% 	\item REVIEWER 2: The single listing (3B) of similar projects within the
+% 	collaboration does not detail whether this is derived or separate
+% 	work.\mtnote{Added a sentence to clarify that PanDA Broker operates
+% 	within the existing ATLAS software infrastructure. Specified that the
+% 	ATLAS team has ported PanDA Broker to other supercomputers.}
+% 	\item REVIEWER 2: How does the TITAN environment prevent usage of DTNs as
+% 	thin proxies? Is connectivity possible from compute node to DTNs via MPI,
+% 	message queues, sockets or other means? Are the restrictions a
+% 	shortcomming of TITAN, or the majority of HPC environments? Given the
+% 	proclaimed importance of HPC resources, a generic evaluation seems
+% 	needed.\mtnote{Added a sentence explaining that DTNs are not part of
+% 	Titan's compute nodes.}
+% 	\item REVIEWER 2: Why was the PanDA Broker designed orthogonal to the
+% 	pilot model, performing both resource acquisition and workload
+% 	execution?\mtnote{We explain that PanDA Broker is a prototype and now we
+% 	are developing NGE.} How do pilots conflict with backfill slots, when the
+% 	former are designed to acquire resources independent of workload
+% 	requirements?\mtnote{AFAIK, we never state that pilot conflict with
+% 	backfill. We argue that the latter mitigates the absence of the former.}
+% 	Why does each Broker instance query and use backfill separately? Given
+% 	that this seems to be a major bottleneck, the choice of such a design
+% 	begs for motivation.\mtnote{Extended a paragraph to explain that PanDA
+% 	Broker supported rapid deployment and incremental optimization while
+% 	incurring in functional limitations.}
+% 	\item REVIEWER 2: Some facts about WLCG seem outdated:
+% 	\begin{itemize}
+% 		\item The number of 30,000 unused TITAN cores (§IIIB) is equated to
+% 		roughly 10\% of the WLCG. However, WLCG currently has 591,282 cores
+% 		[REBUSCPU].
+% 		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{Done.}
+% 		\item WLCG sites are claimed to provide only 8 cores per worker node
+% 		(§IIIB). WLCG has a significant number of sites providing 16 or more
+% 		cores per worker node [REBUSCPU].
+% 		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{This
+% 		would require fairly deep rewriting on Section IVB. I edited the
+% 		sentence referring to 8 cores eliminating the reference to compute
+% 		nodes. Less than ideal but I did not find a better solution. Any
+% 		suggestion?}
+% 	\end{itemize}
+% 	\item REVIEWER 2: Schematic figures (Fig 1, Fig 2) are overly complex,
+% 	while their captions are not sufficient to understand them. The numbering
+% 	is only partially explained in the text.\mtnote{Added a reference to
+% 	where the numbering of the figures are explained in the text. I
+% 	double-checked and the numbering of both figure is fully explained in the
+% 	text.}
+% \end{enumerate}
+% \fi
 

### 2017-09-10 overall.tex
ongoing cleanup.

@@
 
-    \item REVIEWER 2: Why does Fig 3 show an increase of absolute and
-    relative efficiency in Aug-16, when the text claims this in Sep-16 due to
-    more PanDA Brokers?\mtnote{Fixed.}
+    % \item REVIEWER 2: Why does Fig 3 show an increase of absolute and
+    % relative efficiency in Aug-16, when the text claims this in Sep-16 due to
+    % more PanDA Brokers?\mtnote{Fixed.}
 
-    \item REVIEWER 2: Why are only 20 PanDA Brokers deployed if this is a
-    limiting factor? Why is it questionable that more PanDA Brokers would
-    increase efficiency? If only 5/6 of resources below the mean are
-    unusable, this implies an efficiency of almost 60\% being
-    reachable.\mtnote{From our text: ``further research is ongoing to
-    understand whether an even greater number of brokers would yield even
-    greater efficiency. This research focuses on evaluating the overheads of
-    input/output files staging, including its impact on DTNs, and on an
-    alternative design of PanDA Broker that enables the concurrent submission
-    of multiple MPI scripts''. I believe this is an answer good enough.}
+    % \item REVIEWER 2: Why are only 20 PanDA Brokers deployed if this is a
+    % limiting factor? Why is it questionable that more PanDA Brokers would
+    % increase efficiency? If only 5/6 of resources below the mean are
+    % unusable, this implies an efficiency of almost 60\% being
+    % reachable.\mtnote{From our text: ``further research is ongoing to
+    % understand whether an even greater number of brokers would yield even
+    % greater efficiency. This research focuses on evaluating the overheads of
+    % input/output files staging, including its impact on DTNs, and on an
+    % alternative design of PanDA Broker that enables the concurrent submission
+    % of multiple MPI scripts''. I believe this is an answer good enough.}
 
@@
     other domains? The magnitude of capacities presented as available is in
-    major contrast to the requirements sketched in the introduction.
+    major contrast to the requirements sketched in the introduction.\jhanote{MT to address this via a few sentences in Section IV.}
 
-    \item REVIEWER 2: How does the poor metadata read performance of
-    1500-6300s (25\%-100\% of minimum walltime!) affect efficiency? Is it
-    factored into the choice of walltime requests?\mtnote{Thanks to Sergey's
-    input I corrected the figure previously reported.} Is it applicable only
-    to TITAN, to HPC in general, or even to regular WLCG sites?\mtnote{This
-    is clearly out of scope for this paper.} Are there solutions at
-    application, e.g. the use of packaged python scripts and
-    libraries?\mtnote{Added the following: ``While these figures are specific
-    to OLCF and TItan, reengineering on AthenaMP could further improve
-    startup time on every platform.''} The information presented is
-    insufficient to estimate whether this issue applies to other environments
-    and usecases as well, or is simply due to suboptimal
-    implementation/configuration.
+    % \item REVIEWER 2: How does the poor metadata read performance of
+    % 1500-6300s (25\%-100\% of minimum walltime!) affect efficiency? Is it
+    % factored into the choice of walltime requests?\mtnote{Thanks to Sergey's
+    % input I corrected the figure previously reported.} Is it applicable only
+    % to TITAN, to HPC in general, or even to regular WLCG sites?\mtnote{This
+    % is clearly out of scope for this paper.} Are there solutions at
+    % application, e.g. the use of packaged python scripts and
+    % libraries?\mtnote{Added the following: ``While these figures are specific
+    % to OLCF and TItan, reengineering on AthenaMP could further improve
+    % startup time on every platform.''} The information presented is
+    % insufficient to estimate whether this issue applies to other environments
+    % and usecases as well, or is simply due to suboptimal
+    % implementation/configuration.
 
-    \item REVIEWER 3: The HEP-SPEC06 benchmark, which is used as a
-    ``computing power'' measure in LCG is mostly integer-artihmetic based: FP
-    operations accounts for \~10\% of the average workload (see [1], section
-    ``Why do you use `all\_cpp' instead of SPECint or SPECfp?'') While this
-    figure certainly comes from averaging over many different workloads
-    across all experiments, it seems a bit surprising that FP scheduling
-    contention is responsible for 50\% decrease in compute efficiency on
-    Titan when compared to other WLCG sites (paper page 6).\mtnote{Thanks to
-    Sergey's input I downplayed the role played by FP and up-played the role
-    played by new CPUs.}
+    % \item REVIEWER 3: The HEP-SPEC06 benchmark, which is used as a
+    % ``computing power'' measure in LCG is mostly integer-artihmetic based: FP
+    % operations accounts for \~10\% of the average workload (see [1], section
+    % ``Why do you use `all\_cpp' instead of SPECint or SPECfp?'') While this
+    % figure certainly comes from averaging over many different workloads
+    % across all experiments, it seems a bit surprising that FP scheduling
+    % contention is responsible for 50\% decrease in compute efficiency on
+    % Titan when compared to other WLCG sites (paper page 6).\mtnote{Thanks to
+    % Sergey's input I downplayed the role played by FP and up-played the role
+    % played by new CPUs.}
 \end{enumerate}

### 2017-09-10 overall.tex
fix a minor placement error.

@@
     \item REVIEWER 2: What are the reasons for the low efficiency of PanDA
-    Broker? Why are only 30\% of backfill slots useable, if they are
-    basically uncontested?
+    Broker? Why are only 30\% of backfill slots useable, if they are basically
+    uncontested? \jhanote{MT to address this via a few sentences in Section
+    IV.}
 
@@
     other domains? The magnitude of capacities presented as available is in
-    major contrast to the requirements sketched in the introduction.\jhanote{MT to address this via a few sentences in Section IV.}
+    major contrast to the requirements sketched in the introduction.
 

### 2017-09-10 overall.tex
Note about possible new paragraph

--- 2017-09-09 overall.tex
@@
 for the previous DTNs. As such, further hardware upgrades seem unlikely to
-improve significantly the performance of PanDA Brokers.
+improve significantly the performance of PanDA Brokers.\mtnote{Possibly, add
+a small paragraph summarizing why at the moment we are consuming up to 30\%
+of the available backfill.}
 

### 2017-09-10 nge.tex
more comments fixe.

@@
 	the explanation offered in \S\ref{sec:panda_titan}}.
-	\item REVIEWER 3: Figure 6. refers to a "User Workstation" as the running
-	place for the Pilot and Unit Managers, which is never mentioned in the
-	text. (I got the impression that the NGE would run on a DTN node; if this
-	is not the case and separate workstations are being used for the testing
-	phase, this should be made clear.)\mtnote{Fixed.}
+	% \item REVIEWER 3: Figure 6. refers to a "User Workstation" as the running
+	% place for the Pilot and Unit Managers, which is never mentioned in the
+	% text. (I got the impression that the NGE would run on a DTN node; if this
+	% is not the case and separate workstations are being used for the testing
+	% phase, this should be made clear.)\mtnote{Fixed.}
 \end{itemize}

### 2017-09-10 Experiments.tex
removing a query.
Typos

@@
 AthenaMP task was reduced to sixteen such that the running time of each
-AthenaMP was $\sim 1200$ seconds on average. This experiment design choice
+AthenaMP was \(\approx\)1,200 seconds on average. This experiment design choice
 does not affect the objectives or accuracy of the experiments, but allows us

### 2017-09-10 conclusion.tex
Acknowledgements.
another comment non-addressed.

@@
-\ifreview
-Comments to address in this section:
-\begin{itemize}
-	\color{red} 
-	\item REVIEWER 2: How does Summit, the TITAN successor, compare to the
-	conclusions? The end of the TITAN life-cycle basically invalidates the
-	applicability of all conclusions if this generalisation is missing.
-	\item REVIEWER 3: This statement in the "conclusions" section is quite
-	exaggerated: "The experience captured in this paper will prove invaluable
-	for designing WMS for computational campaigns and will provide a baseline
-	to evaluate the relative merits of different approaches." WMS systems for
-	LHC experiments have been developed and run in production for many years;
-	while this paper adds interesting experience and data points, it's
-	unlikely that this will be "invaluable" for new designs and a cornerstone
-	for their evaluation.
-\end{itemize}
-\fi
+% \ifreview
+% Comments to address in this section:
+% \begin{itemize}
+% 	\color{red} 
+% 	% \item REVIEWER 2: How does Summit, the TITAN successor, compare to the
+% 	% conclusions? The end of the TITAN life-cycle basically invalidates the
+% 	% applicability of all conclusions if this generalisation is missing.
+% 	% \item REVIEWER 3: This statement in the "conclusions" section is quite
+% 	% exaggerated: "The experience captured in this paper will prove invaluable
+% 	% for designing WMS for computational campaigns and will provide a baseline
+% 	% to evaluate the relative merits of different approaches." WMS systems for
+% 	% LHC experiments have been developed and run in production for many years;
+% 	% while this paper adds interesting experience and data points, it's
+% 	% unlikely that this will be "invaluable" for new designs and a cornerstone
+% 	% for their evaluation. \jhanote{I think this comment can also be ignored.}
+% \end{itemize}
+% \fi
 
@@
 characteristics, resources types and federation constraints, as well metrics
-of performance. The experience captured in this paper will prove invaluable
+of performance. The experience captured in this paper should be useful
 for designing WMS for computational campaigns and will provide a baseline to
@@
 for dynamic resource federation.
+
+{\bf \footnotesize This work is funded by DOE XXXXXX \par}
 

### 2017-09-10 nge.tex

--- 2017-09-08 nge.tex
@@
-\ifreview
-Comments to address in this section:
-\begin{itemize}
-	\color{red} 
-	\item REVIEWER 3: The first paragraphs of section V. are basically a
-	rediscovery of the advantages of pilot jobs for WLCG task scheduling.
-	Since all those pain points were well-known in the WLCG community before
-	the project started, it should be explained why a pilot job approach was
-	not taken in the first place.\mtnote{Added an opening paragraph recalling
-	the explanation offered in \S\ref{sec:panda_titan}}.
-	\item REVIEWER 3: Figure 6. refers to a "User Workstation" as the running
-	place for the Pilot and Unit Managers, which is never mentioned in the
-	text. (I got the impression that the NGE would run on a DTN node; if this
-	is not the case and separate workstations are being used for the testing
-	phase, this should be made clear.)\mtnote{Fixed.}
-\end{itemize}
-\fi
+% \ifreview
+% Comments to address in this section:
+% \begin{itemize}
+% 	\color{red} 
+% 	% \item REVIEWER 3: The first paragraphs of section V. are basically a
+% 	% rediscovery of the advantages of pilot jobs for WLCG task scheduling.
+% 	% Since all those pain points were well-known in the WLCG community before
+% 	% the project started, it should be explained why a pilot job approach was
+% 	% not taken in the first place.\mtnote{Added an opening paragraph recalling
+% 	% the explanation offered in \S\ref{sec:panda_titan}}.
+% 	% \item REVIEWER 3: Figure 6. refers to a "User Workstation" as the running
+% 	% place for the Pilot and Unit Managers, which is never mentioned in the
+% 	% text. (I got the impression that the NGE would run on a DTN node; if this
+% 	% is not the case and separate workstations are being used for the testing
+% 	% phase, this should be made clear.)\mtnote{Fixed.}
+% \end{itemize}
+% \fi
 

### 2017-09-10 overall.tex

@@
 
-    \item REVIEWER 2: Given an expected increase in resource need of 10-100,
-    how is the capacity of 4\% WLCG significant? How does this relate to the
-    manpower required to develop and maintain a custom solution per HPC site?
-    How many TITAN resources are actually available to HEP, given the use by
-    other domains? The magnitude of capacities presented as available is in
-    major contrast to the requirements sketched in the introduction.
+    % \item REVIEWER 2: Given an expected increase in resource need of 10-100,
+    % how is the capacity of 4\% WLCG significant? How does this relate to the
+    % manpower required to develop and maintain a custom solution per HPC site?
+    % How many TITAN resources are actually available to HEP, given the use by
+    % other domains? The magnitude of capacities presented as available is in
+    % major contrast to the requirements sketched in the introduction.
 

### 2017-09-10 Experiments.tex
fixing and clarifying description of experiments with NGE

@@
 the UnitManager's dispatching of units (tasks) to the agent; and (iii) time
-for the agent to bootstrap all the tasks on the nodes.
+for the Agent to bootstrap all the tasks on the nodes.
 
@@
 
-We ran pilots with 256, 512, 1024 and 2048 worker nodes and 3\mtnote{2?}
-hours walltime. Fig.~\ref{fig:weakScal2a} depicts the average pilot duration,
-the average execution time of five sequential generations of AthenaMP, and
-the corresponding overhead. The difference between the two durations is more
+We ran pilots with 256, 512, 1024 and 2048 worker nodes and 3 hours walltime.
+Fig.~\ref{fig:weakScal2a} depicts the average pilot duration, the average
+execution time of five generations of AthenaMP, and the
+corresponding overhead. The difference between the two durations is more
 marked than in the previous experiments. Despite this, we notice that the
-growth of the overhead is consistent with the increment of the number of
-tasks per node for pilots with 256, 512 and 1024 worker nodes, and less than
-linear for the pilot with 2048 worker nodes.
+growth of the overhead is consistent with the increment of the number of tasks
+per node for pilots with 256, 512 and 1024 worker nodes, and less than linear
+for the pilot with 2048 worker nodes.
 
@@
 different pilot sizes. We used 2048 AthenaMP instances and pilots with 256,
-512, 1024 and 2048 nodes. Thus, the number of AthenaMP generations is equal
-to eight times the size of the smallest pilot and corresponds to the size of
-the largest pilot. As a consequence, the number of consecutive generations of
-AthenaMP decreases with the pilot size. These experiments are designed to
-investigate whether pilot overhead is affected by the degree of concurrency
-within the pilot and/or the number of tasks. Each AthenaMP instance simulates
-sixteen events as in the previous experiment.
+512, 1024 and 2048 nodes. Thus, the number of AthenaMP instances per node
+(i.e., generations) is eight for the smallest pilot (256 nodes), one for the
+largest pilot (2048 nodes), and with the number of generations of AthenaMP
+decreasing with the pilot size. These experiments are designed to investigate
+whether pilot overhead is affected by the degree of concurrency within the
+pilot and/or the number of tasks. Each AthenaMP instance simulates sixteen
+events as in the previous experiment.
 

### 2017-09-10 Experiments.tex
remove comment.

@@
-\ifreview
-Comments to address in this section:
-\begin{enumerate}
-    \color{red} 
-    \item REVIEWER 3: The scalability benchmarks in section "V.B Experiments"
-    should be better explained: what *extacly* is being kept constant in each
-    experiment, and what is being varied and over what range. Also the
-    captions in the figures do not help much.
-\end{enumerate}
-\fi
+% \ifreview
+% Comments to address in this section:
+% \begin{enumerate}
+%     \color{red} 
+%     \item REVIEWER 3: The scalability benchmarks in section "V.B Experiments"
+%     should be better explained: what *extacly* is being kept constant in each
+%     experiment, and what is being varied and over what range. Also the
+%     captions in the figures do not help much.
+% \end{enumerate}
+% \fi
 

### 2017-09-10 conclusion.tex
wordsmithing to prevent single worded sentences.

@@
 
-We characterized the experience by evaluating the efficiency, scalability and
-reliability of both PanDA Broker and AthenaMP as deployed on Titan
-(\S\ref{sec:panda_titan}). Our characterization highlighted the strengths and
-limitations of the current design and implementation: PanDA Brokers enable
-the sustained execution of millions of simulations per week but further work
-is required to optimize its efficiency and reliability
-(\S\ref{ssec:broker_titan}). PanDA Brokers support the concurrent execution
-of multiple AthenaMP instances, enabling each AthenaMP to perform the
-concurrent execution of up to 16 Geant4 simulators. Nonetheless, our
-characterization showed how improving I/O performance could reduce overheads
-(\S\ref{ssec:athenamp_titan}), increasing the overall utilization of Titan's
-backfill availability. We introduced PanDA's next generation executor for HPC
-systems, characterized its ability to support multi-generation workloads and
-analyzed its scaling behavior.  Performance tests on the next generation
-executor demonstrate linear strong and weak scalability over several orders
-of magnitude  of task and node counts.  Qualitatively, it enables the high-
-performance execution of new workloads and advanced execution modes of
-traditional workloads.
+We characterized the efficiency, scalability and reliability of both PanDA
+Broker and AthenaMP as deployed on Titan (\S\ref{sec:panda_titan}). Our
+characterization highlighted the strengths and limitations of the current
+design and implementation: PanDA Brokers enable the sustained execution of
+millions of simulations per week but further work is required to optimize its
+efficiency and reliability (\S\ref{ssec:broker_titan}). PanDA Brokers support
+the concurrent execution of multiple AthenaMP instances, enabling each
+AthenaMP to perform the concurrent execution of up to 16 Geant4 simulators.
+Nonetheless, our characterization showed how improving I/O performance could
+reduce overheads (\S\ref{ssec:athenamp_titan}), increasing the overall
+utilization of Titan's backfill availability. We introduced PanDA's next
+generation executor for HPC systems, characterized its ability to support
+multi-generation workloads and analyzed its scaling behavior.  Performance
+tests on the next generation executor demonstrate linear strong and weak
+scalability over several orders of magnitude  of task and node counts.
+Qualitatively, it enables the high-performance execution of new workloads and
+advanced execution modes of traditional workloads.
 

### 2017-09-10 overall.tex
wordsmithing to prevent single worded sentences.

@@
 is 172,089,760 and the total number of file \(close()\) operations is
-40,132,992. The difference between these two values is still under
-investigation: One possible explanation is that ATLAS jobs don't call a file
-\(close()\) operation per every file \(open()\) issued.
+40,132,992. The difference between these two values is under
+investigation: a possible explanation is that ATLAS jobs don't call a file
+\(close()\) operation for every file \(open()\) issued.
 

### 2017-09-10 overall.tex
suggested tweak in overall.tex.
  Matteo: please review.

@@
 
-PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels)
-is defined as the fraction (or percentage) of core-hours utilized by the
-PanDA Brokers to Titan’s total backfill availability during the experiment
-time window. The brokers reached 18\% average efficiency, with a minimum
-7.8\% efficiency on May 2016 and a maximum 30.9\% efficiency on February 2017
-(excluding the preliminary results of March). The total backfill availability
-was 20.3M in April 2016, and 17.6M in February 2017. This shows that the
-measured difference in efficiency did not depend on a comparable difference
-in the total backfill availability.
+PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels) is
+defined as the fraction (or percentage) of core-hours utilized by the PanDA
+Brokers to Titan’s total backfill availability during the experiment time
+window. The brokers reached 18\% average efficiency, with a minimum efficiency
+of 7.8\% (May 2016) and a maximum efficiency of 30.9\% (Feb. 2017, excluding
+the preliminary results of March). The total backfill availability was 20.3M
+in April 2016, and 17.6M in February 2017. This suggests that the differences
+in efficiency did not depend on a comparable difference in the total backfill
+availability. \jhanote{change last sentence to: ... efficiency does not depend on total backfill availability." ??}
+\jhanote{shouldn't we be consistent in the months used to quote efficiency and
+utilization? April 2016 and Feb 2017 vs May 2016 and Feb 2017??}
 

### 2017-09-10 conclusion.tex
ORNL acknowledgment.

@@
 
-{\bf \footnotesize This work is funded by DOE XXXXXX \par}
+{\bf \footnotesize Acknowledgements: This work is funded by DOE XXXXXX This
+research used resources of the Oak Ridge Leadership Computing Facility at the
+Oak Ridge National Laboratory, which is supported by the Office of Science of
+the U.S.Department of Energy under Contract No. DE-AC05-00OR22725. \par}
 

### 2017-09-10 overall.tex
Fixing description of backfill utilization and efficiency.

--- 2017-09-10 overall.tex
@@
 
-PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels)
-is defined as the fraction (or percentage) of core-hours utilized by the
-PanDA Brokers to Titan’s total backfill availability during the experiment
-time window. The brokers reached 18\% average efficiency, with a minimum
-7.8\% efficiency on May 2016 and a maximum 30.9\% efficiency on February 2017
-(excluding the preliminary results of March). The total backfill availability
-was 20.3M in April 2016, and 17.6M in February 2017. This shows that the
-measured difference in efficiency did not depend on a comparable difference
-in the total backfill availability.
+PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels) is
+defined as the fraction (or percentage) of core-hours utilized by the PanDA
+Brokers of Titan’s total backfill availability during the experiment time
+window. The  average efficiency was 18\%, with a minimum efficiency
+of 7.8\% (May 2016) and a maximum efficiency of 30.9\% (Feb. 2017, excluding
+the preliminary results of March). The total backfill availability was
+$\approx$ 21.5M in April 2016, and 17.6M in February 2017. This suggests that
+the efficiency is invariant of total backfill availability.
+
+% This suggests that the differences
+% in efficiency did not depend on a comparable difference in the total backfill
+% availability.
+
+% \jhanote{change last sentence to: ... efficiency does not depend on total
+% backfill availability." ??}
+% \jhanote{shouldn't we be consistent in the months used to quote efficiency and
+% utilization? April 2016 and Feb 2017 vs May 2016 and Feb 2017??}
 

### 2017-09-10 overall.tex
minor fixes and line savings.

@@
 performance from a worse case scenario of 6,300s on Lustre to \(\approx\)225s
-on NFS\@. While these figures are specific to OLCF and TItan, reengineering
-of AthenaMP could further improve startup time on every platform.
+on NFS\@. While these figures are specific to OLCF and Titan, reengineering
+of AthenaMP could improve startup time on every platform.
 
@@
 
-The performance analysis of Titan's AMD CPUs for detector simulations helps
-also to compare Titan and Grid site performances. Usually, Grid sites exposes
+A performance analysis of Titan's AMD CPUs for detector simulations also helps
+to compare Titan and Grid site performance. Usually, Grid sites expose
 resources with heterogeneous CPU architectures and 8 (virtual) cores, while
-Titan's offer an homogeneous 16 cores architecture. We used the rate of
-events processes per minute as a measure of the efficiency of executing the
-same detector simulation on Titan or Grid sites. Comparisons of the
-efficiencies of Titan to the BNL and SIGNET Grid sites, normalized for 8
-cores, show that the effective performance per-core at Titan is
-\(\approx\)0.57 event per minute, roughly 1/2 of BNL and  1/3 of SIGNET
-performances.
+Titan's offer an homogeneous 16 cores architecture. We used the rate of events
+processes per minute as a measure of the efficiency of executing the same
+detector simulation on Titan or Grid sites. Comparisons of the efficiencies of
+Titan to the BNL and SIGNET Grid sites, normalized for 8 cores, show that the
+effective performance per-core at Titan is
+\(\approx\)0.57 event per minute, roughly 1/2 of BNL and  1/3 of SIGNET.
+
+%performances.
+
+% The CPUs at the Grid sites have one FP scheduler per core and are on
+% average newer than the CPU of Titan.
 
@@
 of newer processors. 
-% The CPUs at the Grid sites have one FP scheduler per core and are on
-% average newer than the CPU of Titan.
 The heterogeneity of the Grid sites' CPUs explain the higher performance

### 2017-09-10 overall.tex
minor fixes to save 3-4 lines.

@@
 
-Currently 20 instances of the PanDA Broker are deployed on 4 Titan DTNs, with
+Currently 20 instances of the PanDA Broker are deployed on 4 DTNs, with
 5 instances per DTN\@. Each broker submits and manages the execution of 15 to
-300 ATLAS jobs, one job for each Titan worker node, and a theoretical maximum
+300 jobs, one job for each Titan worker node, and a theoretical maximum
 concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have
@@
 running at the lowest priority on Titan. Therefore, ATLAS contributed to an
-increase of Titan's overall utilization.
+increase of Titan's utilization.
 

### 2017-09-10 PandaOverview.tex
A bit more space saving

@@
-\ifreview
-Comments to address in this section:
-\begin{itemize}
-	\color{red} 
-	\item REVIEWER 2: use of wrong wordings, such as "eventful" instead of
-	the common term "stateful".\mtnote{Done.}
-\end{itemize}
-\fi
+% \ifreview
+% Comments to address in this section:
+% \begin{itemize}
+% 	\color{red} 
+% 	\item REVIEWER 2: use of wrong wordings, such as "eventful" instead of
+% 	the common term "stateful".\mtnote{Done.}
+% \end{itemize}
+% \fi
 
@@
     subsystems, components, and architectural and communication details are
-    abstracted to improve clarity.}
-\label{fig:architecture}
+    abstracted to improve clarity.}\label{fig:architecture}
 \end{figure}

### 2017-09-11 conclusion.tex
Award Number.

--- 2017-09-10 conclusion.tex
@@
 
-{\bf \footnotesize This work is funded by DOE XXXXXX \par}
+{\bf \footnotesize Acknowledgements: This work is funded by Award number DE-
+SC0016280 from the Office of Advanced Scientific Computing Research within the
+Department of Energy. This research used resources of the Oak Ridge Leadership
+Computing Facility at the Oak Ridge National Laboratory, which is supported by
+the Office of Science of the U.S.Department of Energy under Contract No. DE-
+AC05-00OR22725. \par}
 

### 2017-09-11 paper.tex
added "\\"
A bit more space saving
Author list refine.

@@
 
-\title{High-Throughput Computing on High-Performance Platforms: A Case Study}
+\title{High-Throughput Computing on \\ High-Performance Platforms: A Case Study}
 
@@
     \IEEEauthorblockN{
-        Alessio Angius\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},
         Danila Oleynik\IEEEauthorrefmark{1}\IEEEauthorrefmark{5},
@@
         Matteo Turilli\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},\\
+        Alessio Angius\IEEEauthorrefmark{2},
+        Sarp Oral\IEEEauthorrefmark{6},
         Kaushik De\IEEEauthorrefmark{5},
         Alexei Klimentov\IEEEauthorrefmark{4},
-        Sarp Oral\IEEEauthorrefmark{6},
         Jack C. Wells\IEEEauthorrefmark{7} and
-        Shantenu Jha\IEEEauthorrefmark{2}
+        Shantenu Jha\IEEEauthorrefmark{2}\IEEEauthorrefmark{3}
     }
@@
         \IEEEauthorrefmark{2}Department of Electric and Computer Engineering, Rutgers University, NJ, USA
+    }
+    \IEEEauthorblockA{
+        \IEEEauthorrefmark{3}Computational Science Initiative, Brookhaven National Laboratory, NY, USA
     }
@@
     % reference titles, ie. [2], [3], [4], [7], [8], [9], [10], [11], [12],
-    % [13]. Reference [6] is just an author list, year and title. Should there
-    % be more here?\mtnote{Fixed.}
+    % [13]. Reference [6] is just an author list, year and title. Should
+    % there be more here?\mtnote{Fixed.}
     \item REVIEWER 2:  Key concepts such as pilot jobs, tasks and workloads
-    are not meaningfully defined.
-    % \item REVIEWER 2: the information provided is insufficient to asses in
-    % how far any of the presented evaluation, characterization or lessons are
-    % applicable outside of TITAN and PanDA Broker combined.
+    are not meaningfully defined.    % \item REVIEWER 2: the information provided is insufficient to asses in
+    % how far any of the presented evaluation, characterization or lessons
+    % are applicable outside of TITAN and PanDA Broker combined.
     % \item REVIEWER 2: Several references are of poor quality. For example,
@@
     % \item REVIEWER 3: Title looks too general for what amounts to the
-    % implementation of an adapter component in a system that is very specific
-    % to a single HEP experiment. (In general, title, abstract and conclusions
-    % seem to over-state the contributions of this paper, considering this is a
-    % general eScience conference and not a WLCG-specific one.)
+    % implementation of an adapter component in a system that is very
+    % specific to a single HEP experiment. (In general, title, abstract and
+    % conclusions seem to over-state the contributions of this paper,
+    % considering this is a general eScience conference and not a
+    % WLCG-specific one.)
 \end{itemize}

### 2017-10-28 arxiv/v1/paper.tex
Update arxiv version

\documentclass[conference]{IEEEtran}

\input{head}
\input{include}

% -----------------------------------------------------------------------------
% BEGIN DOCUMENT
% -----------------------------------------------------------------------------
\begin{document}

\title{Converging High-Throughput and High-Performance Computing: A Case Study}

\author{
    \IEEEauthorblockN{
        Alessio Angius\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},
        Danila Oleynik\IEEEauthorrefmark{1}\IEEEauthorrefmark{5},
        Sergey Panitkin\IEEEauthorrefmark{1}\IEEEauthorrefmark{4},
        Matteo Turilli\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},\\
        Kaushik De\IEEEauthorrefmark{5},
        Alexei Klimentov\IEEEauthorrefmark{4},
        Sarp H. Oral\IEEEauthorrefmark{6},
        Jack C. Wells\IEEEauthorrefmark{6} and
        Shantenu Jha\IEEEauthorrefmark{2}
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{1}First authors, alphabetical order
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{2}Department of Electric and Computer Engineering, Rutgers University, NJ, USA
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{4}Physics Department, Brookhaven National Laboratory, NY, USA
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{5}Physics Department, University of Texas Arlington, TX, USA
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{6}Oak Ridge Leadership Computing Facility, Oak Ridge National Laboratory, TN, USA
    }
}

\maketitle

% ----------------------------------------------------------------------------
% ABSTRACT
% ----------------------------------------------------------------------------
\begin{abstract}
The computing systems used by LHC experiments has historically consisted of
the federation of hundreds to thousands of distributed resources, ranging from
small to mid-size resource.  In spite of the impressive scale of the existing
distributed computing solutions, the federation of small to mid-size
resources will be insufficient to meet projected future demands.
This paper is a case study of how the ATLAS experiment has embraced Titan -- a
DOE leadership facility in conjunction with traditional distributed high-
throughput computing to reach sustained production scales of approximately 51M
core-hours a years. The three main contributions of this paper are:  (i) a
critical evaluation of design and operational considerations  to support the
sustained, scalable and production usage of Titan;  (ii) a preliminary
characterization of a next generation executor for PanDA to support new
workloads and  advanced execution modes; and (iii) early lessons for how
current and future experimental and observational systems can be integrated
with production supercomputers and other platforms in a general and extensible
manner.
\end{abstract}

\begin{IEEEkeywords}
Distributed Computing, Workload Management System.
\end{IEEEkeywords}

% ----------------------------------------------------------------------------
% I - INTRODUCTION
% ----------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}

The Large Hadron Collider (LHC) was created to explore the fundamental
properties of matter for the next decades. Multiple experiments at LHC have
collected and distributed hundreds of petabytes of data worldwide to hundreds of
computer centers. Thousands of physicists analyze petascale data volumes daily.
The detection of the Higgs Boson in 2013 speaks to the success of the detector
and experiment design, as well as the sophistication of computing systems
devised to analyze the data.

Historically, the computing systems used by LHC experiments consisted of the
federation of hundreds to thousands of distributed resources, %\textemdash{}
ranging in scale from small to mid-size resource~\cite{foster2003grid}. Although
the workloads to be executed are comprised of tasks that are independent of each
other, the management of the distribution of workloads across many heterogeneous
resources to ensure the effective utilization of resources and efficient
execution of workloads presents non-trivial challenges.

Many software solutions have been developed in response to these challenges.
CMS, one of the LHC experiments, devised a solution based around the
HTCondor~\cite{thain2005distributed} software ecosystem. The
ATLAS~\cite{Aad:2008} experiment, utilizes the Production and Distributed
Analysis (PanDA) workload management system~\cite{Maeno2011} (WMS) for
distributed data processing and analysis. The CMS and ATLAS experiments
represent arguably the largest production grade distributed computing solutions
and have symbolized the paradigm of {\it high-throughput computing}, i.e., the
effective execution of many independent tasks.

As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity era
(Run 4), it is anticipated that the data volumes that will need analyzing will
increase by factors of 10-100 compared to the current phase (Run 2). Data will
be larger in volume but will also require more sophisticated computational
processing. In spite of the impressive scale of the ATLAS distributed computing
system, demand for computing systems will significantly outstrip current and projected supply.

There are multiple levels at which this problem needs to be addressed: the
utilization of emerging parallel architectures (e.g., platforms); algorithmic
and advances in analytical methods (e.g., use of Machine Learning); and the
ability to exploit different platforms (e.g., clouds and supercomputers).

This paper is a case study of how the ATLAS experiment has ``broken free'' of
the traditional computational approach of high-throughput computing on
distributed resources to embrace new platforms, in particular high-performance
computers. Specifically, we discuss the experience of integrating PanDA WMS with
a DOE leadership computing facility called Titan to reach sustained production
scales of approximately 51M core-hours a year. We also discusses the
investigation of a task execution runtime system on Titan, that is based on the
pilot-abstractions and which allows advanced execution modes of the ATLAS
workload as well as the enhanced support for heterogeneous workloads such as
molecular dynamics.

This state-of-practice paper provides three main contributions:  (i) a critical
evaluation of the many design and operational considerations that have been
taken to support the sustained, scalable and production usage of Titan for
historically high-throughput workloads; (ii) a preliminary characterization of a
next generation executor (NGE) for PanDA to support non-traditional
heterogeneous workloads and execution modes;  and (iii) early lessons and
guidance as the community looks forward to designing the next generation of
online analytical platforms~\cite{foap-url}, for how current and future
experimental and observational systems can be integrated with production
supercomputers in a general and extensible manner.

% ----------------------------------------------------------------------------
% II - PanDA OVERVIEW
% ----------------------------------------------------------------------------
\section{PanDA Overview}
\label{sec:panda_overview}

PanDA is a Workload Management System (WMS)~\cite{marco2009glite} designed to
support the execution of distributed workloads and workflows via
pilots~\cite{turilli2015comprehensive}. Pilot-capable WMS enable high throughput
of tasks execution via multi-level scheduling while supporting interoperability
across multiple sites. This is particularly relevant for LHC experiments, where
millions of tasks are executed across multiple sites every month, analyzing and
producing petabytes of data. The design of PanDA WMS started in 2005 to support
ATLAS. PanDA went into production for the LHC Run 1 on 2009, and was then
extended with new subsystems for Run 2 in 2015.

% -----------------------------------------------------------------------------
\subsection{Design}
\label{ssec:panda_design}

PanDA's application model assumes tasks grouped into workloads and workflows.
Tasks represent a set of operations performed on data sets stored in one or more
input files. Tasks are decomposed into jobs, where each job consists of the
task's operations performed on a partition of the task's data set. Since 2005, a
certain amount of parallelism has been progressively introduced for job
execution~\cite{crooks2012multi} but, so far, HEP applications have not required
a message passing interface (MPI).
% Jobs are supposed to be capable of setting up their own execution
% environment and have a minimal set of common dependences.

PanDA's usage model assumes multitenancy of resources and at least two types of
HEP users: individual researchers and groups executing so called `production'
workflows.
% Users can submit tasks and workflows to the PanDA WMS at any point in
% time, directly or via dedicated application frameworks.
PanDA's security model is based on separation between authentication,
authorization and accounting for both single users and group of users. Both
authentication and authorization are based on digital certificates and on the
virtual organization abstraction~\cite{foster2001anatomy}.

Currently, PanDA's execution model is based on four main abstractions: task,
job, queue, and pilot. Both tasks and jobs are assumed to have attributes and
states and to be queued into a global queue for execution. Prioritization and
binding of jobs are assumed to depend on the attributes of each job. Pilot is
used to indicate the abstraction of resource capabilities. Each job is thought
to be bound to one pilot and executed on the site where the pilot has been
instantiated.

In PanDA's data model, each datum refers to the recorded or simulated
measurement of a physical process. Data can be packaged into files or other
containers. As with jobs, data have both attributes and states, and some of the
attributes are shared between events and jobs. Raw, reconstruction, and
simulation data are assumed to be distributed across multiple storage facilities
and managed by the ATLAS Distributed Data Management
(DDM)~\cite{garonne2012atlas}. When necessary, datasets required by each job are
assumed to be replicated over the network, both for input and output data.

PanDA's design supports provenance and traceability for both jobs and data.
Attributes enable provenance by linking jobs and data items, providing
information like ownership or project affiliation. States enable traceability by
providing information about the stage of the execution in which each job or data
item is or has been.
% Some attributes are assumed to be immutable across execution and jobs and data
% items are assumed to be always in one and only one state.

% -----------------------------------------------------------------------------
\subsection{Implementation and Execution}
\label{ssec:panda_arch}

The implementation of PanDA WMS consists of several interconnected subsystems,
most of them built from off-the-shelf and Open Source components. Subsystems
communicate via dedicated API or HTTP messaging, and each subsystem is
implemented by one or more modules. Databases are used to store eventful
entities  like tasks, jobs and  input/output data, and to store information
about sites, resources, logs, and accounting.

Currently, PanDA's architecture has five main subsystems: PanDA
Server~\cite{maeno2011overview},
AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of ATLAS
workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we do not
discuss them as they are  not relevant to an understanding of how PanDA has been
ported to supercomputers. For a full list of subsystems see
Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture} shows a diagrammatic
representation of PanDA main subsystems, highlighting the execution process of
tasks while omitting monitoring details to improve readability.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{panda_architecture.pdf}
    \caption{PanDA WMS architecture. Numbers indicates the JEDI-based execution
    process. Several subsystems, components, and architectural and
    communication details are abstracted to improve clarity.}
\label{fig:architecture}
\end{figure}

During LHC Run 1, PanDA required users to perform a static conversion between
tasks and jobs: tasks were described as a set of jobs and then submitted to the
PanDA Server. This introduced inefficiency both with usability and resource
utilization~\cite{borodin2015unified}. Ideally, users should conceive analyses
in terms of one or more, possibly related tasks, while the execution manager
(i.e., PanDA) should partition tasks into jobs, depending on execution
constraints. Further, the static partitioning of tasks into jobs does not take
into account the heterogeneity and dynamicity of the resources on which each job
will be executed, introducing inefficiencies.

Another problem of static job sizing is that PanDA instantiates pilots on sites
with different type of resources and different models of availability of those
resources. An optimal sizing of each job should take into account these
properties. For example, sites may offer cores with different speed, networking
with different amount of bandwidth, and resources could be guaranteed to be
available for a defined amount of time or could disappear at any point in time
as it happens with opportunistic models of resource provision.

JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users
submit task descriptions to JEDI (Fig.~\ref{fig:architecture}:1) that stores
them into a queue implemented by a database (Fig.~\ref{fig:architecture}:2).
Tasks are partitioned into jobs of different size, depending on both static and
dynamic information about available resources (Fig.~\ref{fig:architecture}:3).
Jobs are bound to sites with resources that best match jobs' requirements, and
submitted to the PanDA Server for execution (Fig.~\ref{fig:architecture}:4).

Once submitted to the PanDA Server, jobs are stored by the Task Buffer component
into a global queue implemented as a database (Fig.~\ref{fig:architecture}:5).
When jobs are submitted directly to the PanDA Server, the Brokerage component is
used to bind jobs to available sites, depending on static information about the
resources available for each site. Jobs submitted by JEDI are already bound to
sites so no further brokerage is needed.

Once jobs are bound to sites, the Brokerage module communicates to the Data
Service module what data sets need to be made available on what site
(Fig.~\ref{fig:architecture}:6). The Data Service communicates these
requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7) that, when needed,
replicates data sets on the required sites (Fig.~\ref{fig:architecture}:8).

Meanwhile, AutoPyFactory defines PanDA Pilots, submitting them to a Condor-G
agent (Fig.~\ref{fig:architecture}:9). Condor-G schedules these pilots wrapped
as jobs or VMs to the required sites (Fig.~\ref{fig:architecture}:10).

When a PanDA Pilot becomes available, it requests the Job Dispatcher module of
the PanDA Server for a job to execute (Fig.~\ref{fig:architecture}:11). The Job
Dispatcher interrogates the Task Buffer module for a job that is bound to the
site of that pilot and ready to be executed. Task Buffer checks the global queue
(i.e., the PanDA database) and, upon availability, returns a job to the Job
Dispatcher. The Job Dispatcher dispatches that job to the PanDA Pilot
(Fig.~\ref{fig:architecture}:12).

Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
subprocess for the execution of the job's payload. Input data are transferred
from the stage-in location (Fig.~\ref{fig:architecture}:13) and the job's
payload is executed (Fig.~\ref{fig:architecture}:14). Once completed, output is
transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).

The Data Service module of the PanDA Server tracks and collects the output
generated by each job (Fig.~\ref{fig:architecture}:16), updating jobs'
attributes via the Task Buffer module (Fig.~\ref{fig:architecture}:17). When the
output of all the jobs of a task are retrieved, it is made available to the user
via PanDA Server. When a task is submitted to JEDI, task is instead marked as
done (Fig.~\ref{fig:architecture}:18) and the result of its execution is made
available to the user by JEDI (Fig.~\ref{fig:architecture}:19).

% ----------------------------------------------------------------------------
% III - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
% ----------------------------------------------------------------------------
%\section{Deploying PanDA on a Leadership-scale system}
\section{Deploying PanDA on Titan}
\label{sec:panda_titan}

The upcoming LHC Run 3 will require more resources than the Worldwide LHC
Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than 300,000
cores at over 100 Grid sites, with a peak performance of 0.3 petaFLOPS. This
capacity will be sufficient for the planned analysis and data processing, but it
will be insufficient for the Monte Carlo production workflow and any extra
activity. To alleviate these challenges, ATLAS is expanding the current
computing model to include additional resources such as the opportunistic use of
supercomputers.

Generally, supercomputers are designed to support parallel computation that
requires runtime communication. Jobs are executed across multiple cores, each
core calculating a small part of the problem and communicating with other cores
via MPI. Accordingly, supercomputers have large number of worker nodes,
connected through a high-speed, low-latency dedicated network. Each worker node
has multicore CPUs, usually augmented with Graphics Processing Units (GPUs) or
other specialized coprocessors.

PanDA WMS has been designed to support distributed Grid computing. Executing
ATLAS workloads or workflows involves concurrent and/or sequential runs of
possibly large amount of jobs, each requiring no or minimal parallelization and
no runtime communication. Thus, computing infrastructure like WLCG have been
designed to aggregate large amount of computing resources across multiple sites.
While each site may deploy MPI capabilities, usually these are not used to
perform distributed computations.

We developed and deployed a single-point solution to better understand the
problem space of enabling a WMS designed for HTC to execute production workflows
on resources designed to support HPC. The PanDA team developed a job broker to
support the execution of part of the ATLAS production Monte Carlo workflow on
Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).

% -----------------------------------------------------------------------------
\subsection{Architectures and Interfaces}
\label{ssec:panda-titan}

The Titan supercomputer, current number three on the Top 500 list~\cite{top500},
is a Cray XK7 system with 18,688 worker nodes and a total of 299,008 CPU cores.
Each worker node has an AMD Opteron  6274 16-core CPU, a Nvidia Tesla K20X GPU,
32 GB of RAM and no local storage, though a 16 GB RAM disk can be set up. Work
nodes use Cray’s Gemini interconnect for inter-node MPI messaging. Titan is
served by the Spider II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of
disk storage, and by a 29 PB HPSS tape storage system. Titan’s worker nodes run
Compute Node Linux, a run time environment based on SUSE Linux Enterprise
Server.

Titan's users submit jobs to Titan's PBS scheduler by logging into login or data
transfer nodes (DTNs). Titan's authentication and authorization model is based
on two-factor authentication with a RSA SecurID key. Login nodes and DTNs have
out/inbound wide area network connectivity while worker nodes have only local
network access. Fair-share and allocation policies are in place both for the PBS
batch system and shared file systems.

Titan's architecture, configuration and policies poses several challenges to the
integration with PanDA. The default deployment
model of PanDA Pilot is unfeasible on Titan: PanDA Pilot is required to contact
the Job Dispatcher of the PanDA Server to pull jobs to execute, but this is not
possible on Titan because worker nodes do not offer outbound network
connectivity. Further, Titan does not support PanDA's security model based on
certificates and virtual organizations, making PanDA's approach to identity
management also unfeasible. While DTNs offer wide area network data transfer, an
integration with ATLAS DDM is beyond the functional and administrative scope of
the current prototyping phase. Finally, the specific characteristics of the
execution environment, especially the absence of local storage on the worker
nodes and modules tailored to Compute Node Linux, require re-engineering of
ATLAS application frameworks.

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows can
be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
partitioned into independent jobs, Titan worker nodes can be used to execute up
to 16 concurrent payloads, one per each available core. Given these constraints
and challenges, the type of task most suitable for execution at the moment on
Titan is Monte Carlo detector simulation. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and with
small input data requirements. Detector simulation tasks in ATLAS are performed
via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
offloading.

Detector simulation is part of the ATLAS production Monte Carlo (MC)
workflow~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
consists of four main stages: event generation, detector simulation,
digitization, and reconstruction. Event generation creates sets of particle
four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
HERWIG~\cite{corcella2001herwig} and many others. The detector simulator is
called Geant4~\cite{agostinelli2003geant4} and simulates the ATLAS detector and
the interaction between that detector and particles. Each interaction creates a
so-called hit and all hits are collected and passed on for digitalization, where
hits are further processed to mimic the readout of the detector. Finally,
reconstruction operates local pattern recognition, creating high-level objects
like particles and jets.

% -----------------------------------------------------------------------------
\subsection{PanDA Broker}
\label{ssec:panda_titan}

The lack of wide area network connectivity on Titan's worker nodes is the most
relevant challenge for integrating PanDA WMS and Titan. Without connectivity,
Panda Pilots cannot be scheduled on worker nodes because they would not be able
to communicate with PanDA Server and therefore pull and execute jobs. This makes
impossible to port PanDA Pilot to Titan while maintaining the defining feature
of the pilot abstraction: decoupling resource acquisition from workload
execution via multi-stage scheduling.

The unavailability of pilots is a potential drawback when executing distributed
workloads like MC detector simulation. Pilots are used to increase the
throughput of distributed workloads: while pilots have to wait in the
supercomputer's queue, once scheduled, they can pull and execute jobs
independent from the system's queue. Jobs can be concurrently executed on
every core available to the pilot, and multiple generations of concurrent
executions can be performed until the pilot's walltime is exhausted. This is
particularly relevant for machines like Titan where queue policies privilege
parallel jobs on the base of the number of worker nodes they request: the higher
the number of nodes, the shorter the amount of queue time (modulo fair-share and
allocation policies).

The backfill optimization of Titan's Moab scheduler allows to avoid the overhead
of queue wait times without using pilot abstraction~\cite{maui_backfill_url}.
With this optimization, Moab starts low-priority jobs when they do not delay
higher priority jobs, independent of whether the low-priority jobs were queued
after the high-priority jobs.

When the backfill optimization is enabled, users can interrogate Moab about the
number of worker nodes and walltime that would be available to a low-priority
job at that moment in time. If a job is immediately submitted to Titan with that
number of worker nodes and walltime, chances are that Moab will immediately
schedule it, reducing its queue time to a minimum. In this paper, we call this
number of worker nodes and walltime an available `backfill slot'.

Compared to pilots, backfill has the disadvantage of limiting the amount of
worker nodes that can be requested. Pilots are normal jobs: they can request as
many worker nodes and walltime as a queue can offer. On the contrary, jobs sized
according to an available backfill slot depend on the number of worker nodes and
walltime that cannot be given to any other job at that moment in time.

At any point in time, the size of an available backfill slot is typically a
small fraction of the total capacity of a resource. Notwithstanding, given the
size of Titan this translates into a substantial capacity. Every year, about
10\% of Titan's capacity remains unused~\cite{barker2016us}, corresponding to an
average of 30,000 unused cores (excluding GPU cores). This equals to roughly
10\% of the overall capacity of WLCG.

Given the communication requirements of PanDA Pilots and the unused capacity of
Titan, PanDA pilot was repurposed to serve as a job broker on the DTN nodes of
Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of PanDA Pilot
and its stand-alone architecture, this prototype called `PanDA Broker'
implements functionalities to: (i) interrogate Titan about backfill
availability; (ii) pull ATLAS jobs and events from PanDA Server; (iii) wrap the
payload of ATLAS jobs into MPI scripts; (iv) submitting MPI scripts to Titan's
PBS batch system and monitor their execution; and (v) staging and preparing
input/output files. Backfill querying, payload wrapping, and scripts submission
required a new implementation while pulling ATLAS job and events, and file
staging were inherited from PanDA Pilot.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{panda_broker_architecture.pdf}
    \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates
    the execution process of a detector simulation job, part of the production
    ATLAS MC workflow.}
\label{fig:panda_broker}
\end{figure}

Backfill querying is performed via a dedicated Moab scheduler command while a
tailored Python MPI script is used to execute the payload of ATLAS jobs. This
MPI script enables the execution of unmodified Grid-centric, ATLAS jobs on
Titan. Typically, a MPI script is workload-specific as it sets up the execution
environment for a specific payload. This involves organization of worker
directories, data management, optional input parameters modification, and
cleanup on exit. Upon submission, a copy of the MPI script runs on every
available worker node, starting the execution of the ATLAS job's payload in a
subprocess and waits until its completion.

MPI scripts are submitted to Titan's PBS batch system via
RADICAL-SAGA~\cite{radical-saga_url}, a Python module, compliant with the OGF
GFD.90 SAGA specification~\cite{goodale2008simple}. The Simple API for Grid
Applications (SAGA) offers a unified interface to diverse job schedulers and
file transferring services. In this way, SAGA provides an interoperability layer
that lowers the complexity of using distributed infrastructures. Behind the API
façade, RADICAL-SAGA implements a adaptor architecture: each adaptor interface
the SAGA API with different middleware systems and services, including the PBS
batch scheduler of Titan.

The data staging capabilities of the PanDA Broker are implemented via a file
system that is shared among DTNs and worker nodes. The input files with the
events of the ATLAS jobs are downloaded on the shared filesystem from the data
center of Brookhaven National Laboratory (BNL). The MPI script setup process
includes making the location of these files available to the payload of the
ATLAS's jobs. The PanDA Broker can locate the payload's output files on the
shared filesystem and transfer them from Titan BNL.

Once deployed on Titan, every PanDA Broker supports the execution of MC detector
simulations in 9 steps. PanDA Broker queries the Job Dispatcher module of the
PanDA server for ATLAS jobs that have been bound to Titan by JEDI
(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA Broker
pulls jobs' input files from BNL to the OLCF Lustre file system
(Fig.~\ref{fig:panda_broker}:2). PanDA Broker queries Titan's Moab scheduler
about the current available backfill slot (Fig.~\ref{fig:panda_broker}:3) and
creates an MPI script, wrapping enough ATLAS jobs' payload to fit the backfill
slot. PanDA Broker submits the MPI script to the Titan's Torque batch system via
RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).

Upon execution on the worker node(s) (Fig.~\ref{fig:panda_broker}:5), the MPI
script initializes and configures the execution environment
(Fig.~\ref{fig:panda_broker}:6), and executes one AthenaMP for each available
work node (Fig.~\ref{fig:panda_broker}:7). AthenaMP retrieves events from Lustre
(Fig.~\ref{fig:panda_broker}:8) and spawns 1 Geant4 event simulation process on
each of the 16 available cores (Fig.~\ref{fig:panda_broker}:9). Upon completion
of each MPI script, PanDA Broker transfer the jobs' output to BNL
(Fig.~\ref{fig:panda_broker}:10), and performs cleanup.

While PanDA Broker implementation is resource specific, it was successfully
ported to other supercomputers, including the HPC2 at the National Research
Center ``Kurchatov Institute'' (NRC-KI)~\cite{belyaev2015integration},
Edison/Cori at the National Energy Research Scientific Computing Center
(NERSC)~\cite{barreiro2016panda}, and SuperMUC at the Leibniz Supercomputing
Centre (LRZ)~\cite{barreiro2016panda}.

% ----------------------------------------------------------------------------
% IV - ANALYSIS AND DISCUSSION
% ----------------------------------------------------------------------------
\section{Analysis and Discussion}
\label{sec:panda_titan}

Currently, ATLAS deploys 20 instances of the PanDA Broker on 4 Titan's DTNs, 5
instances per DTN. Each broker submits and manages the execution of 15 to 300
ATLAS jobs, one job for each Titan's worker node, and a theoretical maximum
concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have operated
only in backfill mode, without a defined time allocation, and running at the
lowest priority on Titan.

We evaluate the efficiency, scalability and reliability of the deployment of
PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and
AthenaMP. We discuss challenges and limitations of our approach at multiple
levels arising from the specifics of workload, middleware and methods. All the
measurements were performed between January 2016 and February 2017, hereafter
called ‘experiment time window

% -----------------------------------------------------------------------------
\subsection{Characterizing the PanDA Broker on Titan}
\label{ssec:broker_titan}

We calculate the total amount of backfill availability over a period of time by:
(i) polling the available backfill slots at regular intervals during that time
window; (ii) converting the number of worker nodes available and their walltime
into core-hours; (iii) summing the number of core-hours. We call this number of
core-hours `total backfill availability'.

Fig.~\ref{fig:backfill-utilization} shows the total backfill availability on
Titan (gray bars) and the number of core-hours of that availability used by
ATLAS (blue bars) during the experiment time window. ATLAS consumed a total of
51.4M core-hours, for an average of $\approx$3.7M core-hours a month, with a
minimum of 1.7M core-hours in April 2016 and a maximum 7.9M core-hours in
February 2017.

\begin{figure}[!t]
    \includegraphics[clip,width=\columnwidth]{backfill_consumption.pdf}
    \caption{Titan's total backfill availability: CPU core-hours (gray) and CPU
    core-hours used by ATLAS (blue). GPU core-hours unaccounted for as they
    cannot be used by ATLAS. Efficiency of PanDA Brokers defined as percentage
    of total Titan's backfill availability used by ATLAS (Red labels).}
\label{fig:backfill-utilization}
\end{figure}

PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels) is
defined as the fraction of core-hours utilized by the PanDA Brokers to  Titan’s
total backfill availability during the experiment time window. The brokers
reached 18\% average efficiency, with a minimum 7.8\% efficiency on May 2016 and
a maximum 30.9\% efficiency on February 2017 (excluding the preliminary results
of March). The total backfill availability was 20.3M in April 2016, and 17.6M in
February 2017. This shows that the measured difference in efficiency did not
depend on a comparable difference in the total backfill availability.

During the experiment time window, about 2.25M detector simulation jobs were
completed on Titan, for a total of 225M events processed. This is equivalent to
0.9\% of all the 250M detector simulations performed by ATLAS in the same period
of time, and 3.5\% of the 6.6B events processed by those jobs. These figures
confirms the relevance of supercomputers' resource contribution to the LHC Run
2, especially when accounting for the amount of unused total backfill
availability and the improvement of PanDA efficiency across the experiment time
window.

On February 2017, PanDA Brokers used almost twice as much total backfill
availability than in any other month (preliminary results for March 2017
displayed in Fig.~\ref{fig:backfill-utilization} confirm this trend). No
relevant code update was made during that period and logs indicated that the
brokers were able to perform faster. This is likely due to hardware upgrades on
the DTNs. The absence of continuous monitoring of those nodes does not allow to
quantify bottlenecks but spot measurements of their load indicate that a faster
CPU and better networking were likely responsible for the improved performance.
Investigations showed an average CPU load of 3.6\% on the upgraded DTNs. As
such, further hardware upgrades seem unlikely to improve significantly the
performance of PanDA Brokers.

Every detector simulation executed on Titan process 100 events. This number of
events is consistent with the physics of the use case and with the average
duration of backfill availability. The duration of a detector simulation is a
function of the number of events simulated but not all events take the same time
to be simulated. 1 event simulation takes from $\approx$2 to $\approx$40
minutes, with a mean of $\approx$14 minutes. Considering that each worker node
process up to 16 events concurrently, 100 events takes an average of 105 minutes
to process. As such, PanDA brokers do not use backfill availability with less
than 105 minutes walltime.

Fig.~\ref{fig:backfill-distrib} shows the distribution of backfill availability
on Titan as a function of number of nodes and the time of their availability
(i.e., walltime). We recorded these data by polling Titan's Moab scheduler at
regular intervals during the experiment window time. The mean number of nodes
was 691, and their mean walltime was 126 minutes. Detector simulations of 100
events, enable to use down to 5/6 of the mean walltime of backfill availability.
As such, it offers a good compromise for PanDA Broker efficiency.

\begin{figure*}[!t]
    \includegraphics[clip,width=0.32\textwidth]{titan_backfill_wnodes_distribution.pdf}
    \includegraphics[clip,width=0.32\textwidth]{titan_backfill_walltime_distribution.pdf}
    \includegraphics[clip,width=0.32\textwidth]{titan_backfill_avail.png}
    \caption{Backfill availability: distribution of number of work nodes (left)
    and of walltime in minutes (center). Scatter plot of the two variables
    (right). 62175 measures: mean number of work nodes available 691; mean
    walltime available 126 minutes.}
\label{fig:backfill-distrib}
\end{figure*}

PanDA Broker could fit the number of events to the walltime of each available
backfill slot on the base of the distributions of the time taken by one event to
be simulated. That specific number of event could then be pulled from the PanDA
Event service~\cite{calafiura2015atlas} and given as input to one or more
simulations. Once packaged into the MPI script submitted to titan's PBS batch
system, these simulations would better fit their available backfill slot,
contributing to increase the efficiency of PanDA Brokers.

The transition from a homogeneous to a heterogeneous number of events per
detector simulation has implications for the application layer. An even number
of events across simulations makes it easier to partition, track and package
events across simulations, especially when they are performed on both the Grid
and Titan. A homogeneous number of events also helps to keep the size and
duration of other stages of the MC workflow (\S\ref{ssec:panda-titan}) more
uniform. Further analysis is needed to evaluate the trade offs between increased
efficiency of resource utilization and the complexity that would be introduced
at the application layer.

Currently, each PanDA Broker creates, submits, and monitors a single MPI PBS
script at a time. This design is inherited from PanDA Pilot where a single
process is spawn at a time to execute the payload. As a consequence, the
utilization of a larger portion of Titan's total backfill availability depends
on the the number of concurrent PanDA Brokers instantiated on the DTNs: When all
the 20 PanDA Brokers have submitted a MPI PBS script, further backfill
availability cannot be used.

In September 2016, increasing the number of concurrent PanDA brokers from 4 to
20 markedly improved efficiency (see Fig.~\ref{fig:backfill-utilization}) but
further research is undergoing to understand whether an even larger number of
brokers would yield similar results. This research focuses on evaluating the
overheads of input/output files staging, including its impact on DTNs, and on an
alternative design of PanDA Broker that enables the concurrent submission of
multiple MPI scripts~\cite{barreiro2016panda}.

The current design and architecture of the PanDA Broker is proving to be as
reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
Brokers were responsible for around the 19\% of the failures, compared to the
29\% of failures produced by the JobDispatcher module of the PanDA Server, and
the 13\% failures produced by the Geant4 toolkit. The current failure rate of
the PanDA Brokers confirms the benefits of reusing most of the code base of the
PanDA Pilot for implementing the PanDA Broker. It also shows that adopting
third-party libraries like RADICAL-SAGA did not have a measurable adverse effect
on reliability.

% -----------------------------------------------------------------------------
\subsection{Characterizing the Detector Simulation on Titan}
\label{ssec:athenamp_titan}

We use two main parameters to measure the performance of the detector simulation
jobs submitted to Titan: (i) the time taken to setup AthenaMP; and (ii) the
distribution of the time taken by the Geant4 toolkit to simulate a certain
number of events.

AthenaMP has an initialization and configuration stage. At initialization time,
AthenaMP is assembled from a large number of shared libraries, depending on the
type of payload that will have to be computed. Once initialized, every algorithm
and service of AthenaMP is configured by a set of Python scripts. Both these
operations result in a large number of read operations on the filesystem shared
between the worker nodes and the DTNs, including the operations required to
access small python scripts.

Initially, all the shared libraries of AthenaMP and the python scripts for its
configuration were stored on the Spider 2 Lustre file system. However, the I/O
patterns of the initialization and configuration stages degraded the performance
of the filesystem. This was addressed by moving the AthenaMP distribution to a
read-only NFS directory, shared among DTNs and worker nodes. NFS eliminated the
problem of metadata contention, improving metadata read performance from
$\approx$6,300 seconds on Lustre to $\approx$1,500 seconds on NFS.

Once initialized and configured, AthenaMP is used to execute 16 concurrent
Geant4 simulators on a Titan's worker node. Geant4 requires to read events
descriptions from a filesystem and simulate them as they would happen within the
ATLAS detector. We characterized both the compute performance of the simulation
and the impact of acquiring event descriptions on the filesystem.

The AMD Opteron 6274 CPU used on Titan has 16 cores, divided into 8 compute
units. Each compute units has 1 floating point (FP) scheduler shared between 2
cores. When using 16 cores for FP-intensive calculations, each pair of cores
competes for a single FP scheduler. This creates the overhead shown in
Fig.~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
concurrent simulations computing 50 events is 10.8 minutes, while for 16
simulations is 14.25 minutes (consistent with the measured distribution of the
duration of event simulation). Despite an inefficiency of almost 30\%, Titan's
allocation policy based on number of worker nodes used instead of number of
cores does not justify the use of 1/2 of the cores available.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{tx8_tx16_comparison_vsquashed.pdf}
    \caption{Distributions of the time taken to simulate one event when placing
    2 simulations (h1) or 1 simulation (h2) per Titan's CPU. 2 simulation use 16
    cores per node, 1 simulation 8. 50 Events; 1 Titan worker nodes; 16 work
    threads per node; 100 events per node.}
\label{fig:comparison-8-16cores}
\end{figure}

The performance analysis of Titan's AMD CPUs for detector simulations helps also
to compare Titan and Grid site performances. Usually, Grid sites exposes
resources with heterogeneous CPU architectures and a maximum of 8 (virtual)
cores per worker node, while Titan's offer an homogeneous 16 cores architecture.
We used the rate of events processes per minute as a measure of the efficiency
of executing the same detector simulation on Titan or Grid sites. Comparisons of
the efficiencies of Titan to the BNL and SIGNET Grid sites, normalized for 8
cores, show that the effective performance per-core at Titan is $\approx$0.57
event per minute, roughly 1/2 of BNL and  1/3 of SIGNET performances.

The differences in performance between Titan and the BNL and SIGNET Grid sites
are due to the FP scheduler competition and the availability of newer
processors. The CPUs at the Grid sites have one FP scheduler per core and are on
average newer than the CPU of Titan. The heterogeneity of the Grid sites' CPUs
explain the higher performance variance we observed compared to the performance
consistency we measured on Titan.

We studied the impact of acquiring event descriptions on Lustre by analyzing
1,175 jobs ran on the week of 10/25/2016, for a total of 174 hours.
Table~\ref{panda-olcf-stats} shows the overall statistical breakdown of the
observed file I/O. ATLAS used between 1 and 300 worker nodes, and 35 on average.
75\% of the jobs run by ATLAS consumed less than 25 nodes and 92\% less than
100. During the 174 hours of data collection, 6.75 ATLAS jobs were executed on
average per hour, each job running for an average of 1.74 hours. Every job read
less than 250 GB and wrote less than 75 GB of data and, on average, each job
read 20 GB and wrote 6 GB of data.

\begin{table*}[t]
\centering
\begin{tabular}{lllllllll}
 & Num. Nodes & Duration (s) & Read (GB) & Written (GB) & GB Read/nodes & GB Written/nodes & $open()$ & $close()$ \\
Min & 1 & 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\
Max & 300 & 7,452 & 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\
Average & 35.66 & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 & 34,155.74 \\
Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376 & 231,346.55 & 53,799.08
\end{tabular}
\caption{The Statistical breakdown of the I/O impact of 1,175 jobs ATLAS executed at OLCF for the week of 10/25/16}
\label{panda-olcf-stats}
\end{table*}

ATLAS jobs are read heavy: On average, the amount of data read per worker node
is less than 400 MB, while the amount of data written is less than 170 MB.
Distributions of read and written data are different: The read operation
distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
the written amount of data has a very narrow distribution.

The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
operations per second (not including file $stat()$ operations) and 5 file
$close()$ operations per second, with similar distributions. On average, the
maximum number of file $open()$ operations per job is $\approx$170/s and the
maximum number of file $close()$ operations is $\approx$39/s. For the 1,175
ATLAS jobs observed, the total number of file $open()$ operations is 172,089,760
and the total number of file $close()$ operations is 40,132,992. The difference
between these two values is still under investigation: One possible explanation
is that ATLAS jobs don't call a file $close()$ operation per every file $open()$
issued.

Overall, the average time taken to read events from input files stored on Lustre
is 1,320, comparable to the time taken to read the file required by assembling
AthenaMP from NFS. Preliminary investigation shows that this time could be
reduced to 40 seconds by loading the event descriptions into the RAM disk
available on each worker node. Events descriptions could be transferred from
Lustre to the RAM disk while configuring and initializing AthenaMP, almost
halving the time currently required by initiating a Genat4 simulation.

% ----------------------------------------------------------------------------
% VI - THE NEXT GENERATION EXECUTOR
% ----------------------------------------------------------------------------
% \section{PANDA: Future/RoadMap}
\section{PANDA\@: The Next Generation Executor}
\label{sec:panda_roadmap}

As seen in \S\ref{ssec:panda_titan}, PanDA Broker was  deployed on the DTNs
as Titan's worker nodes lack connectivity to the wide area network The lack of
pilot capabilities impacts both the efficiency and the flexibility of PanDA's
execution process. Pilots could improve efficiency by increasing throughput and
enabling greater backfill utilization. Further, pilots makes it easier to
support heterogeneous workloads.

The absence of pilots imposes the static coupling between MPI scripts
submitted to the PBS batch system and detector simulations
(\S\ref{sec:panda_titan}). This makes the scheduling of multiple generations
of workload on the same PBS job impossible: once a statically defined number
of detector simulations are packaged into a PBS job and this job is queued on
Titan, no further simulations can be added to that job. New simulations have
to be packaged into a new PBS job that needs to be submitted to Titan based
upon backfill availability.

The support of  multiple generations of workload would enable more efficient use
of the backfill availability of walltime. Currently, when a set of simulations
ends, the PBS job also ends, independent of whether more wall-time would still
be available. With a pilot, additional simulations could be executed  to utilize
all the available wall-time, while avoiding further job packaging and submission
overheads.

Multiple generations would also relax two assumptions of the current execution
model: knowing the number of simulations before submitting the MPI script, and
having a fixed number of events per simulation (100 at the moment). Pilots would
enable the scheduling of simulations independently from whether they were
available at the moment of submitting the pilot. Further, simulations with a
varying number of events could be scheduled on a pilot, depending on the amount
of remaining walltime and the distribution of execution time per event, as shown
in \S\ref{ssec:panda_titan}, Fig.~\ref{fig:comparison-8-16cores}. These
capabilities would increase the efficiency of the PanDA Broker when there is a
large difference between the number of cores and walltime.

Pilots can offer a payload-independent scheduling interface while hiding the
mechanics of coordination and communication among multiple worker nodes. This
could eliminate the need for packaging payload into MPI scripts within the
broker, greatly simplifying the submission process. This simplification would
also enable the submission of different types of payload, without having to
develop a specific PBS script for each payload. The submission process would
also be MPI-independent, as MPI is used for coordination among multiple worker
nodes, not by the payload.

% -----------------------------------------------------------------------------
\subsection{Implementation}
\label{sec:arch}

The implementation of pilot capabilities within the current PanDA Broker require
quantification of the effective benefits that it could yield and, on the base of
this analysis, a dedicated engineering effort. We developed a prototype of a
pilot system capable of executing on Titan to study experimentally the
quantitative and qualitative benefits that it could bring to PanDA. We called
this prototype Next Generation Executor (NGE).

NGE is a runtime system to execute heterogeneous and dynamically determined
tasks that constitute workloads. Fig.~\ref{fig:arch-overview} illustrates its
current architecture as deployed on Titan: the two management modules (Pilot
and Unit) represent a simplified version of the PanDA Broker while the agent
module is the pilot submitted to Titan and executed on its worker nodes. The
communication between PanDA Broker and Server is abstracted away as it
is not immediately useful to evaluate the performance and capabilities of a
pilot on Titan.

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{rp_architecture_compact_atlaswms_paper.pdf}
  \caption{NGE Architecture: The PilotManager and  UnitManager reside on a Titan
  DTN while the Agent is executed on its worker nodes. Color coding: gray for
  entities external to NGE, white for APIs, purple for NGE's modules, green for
  pilots, yellow for module's components.}
\label{fig:arch-overview}
\end{figure}

NGE exposes an API to describe workloads (Fig.~\ref{fig:arch-overview}, green
squares) and pilots (Fig.~\ref{fig:arch-overview}, red circles), and to
instantiate a PilotManager and a UnitManager.   The PilotManager submits
pilots to Titan's PBS  batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit MPI scripts. Once
scheduled, the pilot's Agent is bootstrapped on Titan's MOM node and
the Agent's Executors on the on the worker
nodes, and the UnitManager schedules units to the Agent's Scheduler
(Fig.~\ref{fig:arch-overview}, solid arrow). The Agent's Scheduler schedules
the units on one or more Agent's Executor for execution. The Agent's executors
can manage one or more worker nodes, depending on performance evaluations. The
UnitManager and the Agent communicate via a database that is instantiated on
one of Titan's DTN so as to be reachable by both modules.

The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
communication and coordination of the execution of units. This environment is
a critical component of the OpenMPI implementation, developed to support
distributed high-performance computing applications operating in a
heterogeneous environment~\cite{castain05:_open_rte, cug-2016}. ORTE provides
a mechanism to create a \emph{``dynamic virtual machine''} (DVM) that spans
multiple nodes. The DVM transparently provides support for interprocess
communication, resource discovery and allocation, and process launching across
a variety of platforms. Libraries enable the interaction between users and
ORTE to enable the submission, monitoring and managing of tasks, avoiding
filesystem bottlenecks and race conditions with network sockets. As a
consequence, ORTE is able to minimize the system overhead while submitting
tasks.

\subsection{Experiments}
\label{sec:ngeExp}

We designed experiments to characterize the performance of the NGE on Titan,
with an emphasis on understanding its overhead and thus the cost of introducing
new functionalities.  We perform three groups of experiments in which we
investigate the weak scalability, weak scalability with multiple generation, and
strong scalability of the NGE.

Each experiment entails executing multiple instances of AthenaMP
to simulate a pre-determined number of events. All the experiments have been
performed by  configuring AthenaMP to use all the 16 cores  of Titan's worker
nodes.

We  measured the execution time of the pilots and of AthenaMP
within them, collecting timestamps at  all stages of the execution. Experiments
were performed  by  submitting NGE's pilots  to Titan's batch queue.  The
turnaround time of an individual run is determined by queue waiting times. Since
we are interested only in the performances of the NGE, we removed queue time
from our statistics.

\subsubsection{Weak scalability}

In this experiment  we run as many AthenaMP instances (hereafter referred to as
tasks)  as the number of nodes controlled by the pilot. Each AthenaMP simulates
100 events, requiring $\sim 4200$ seconds on average.

Tasks do not  wait within the NGE Agent's queue since  one node  is available to
each AthenaMP instance.  Overheads in task execution are consequence primarily
of the three other factors: (i) the  initial bootstrapping of the pilot on the
nodes; (ii) the UnitManager's dispatching of  units (tasks) to the agent; and
(iii) time for the agent to bootstrap all the tasks on the nodes.

We tested  pilots  with 250, 500, 1000 and 2000 worker nodes and 2 hours
walltime. The time duration is determined by the Titan's walltime policy.
Fig.~\ref{fig:weakScal1a} depicts the average pilot duration, the average
execution time of AthenaMP, and the NGE pilot overhead as function of the pilot
size.

\begin{figure}[!t]
        \includegraphics[height=4.5cm,width=\columnwidth]{weak1.pdf}
    \caption{Weak scalability: average pilot duration, average duration of one AthenaMP execution, and pilot's overhead as a function of pilot sizes (200, 500, 1000 and 2000 nodes).}
\label{fig:weakScal1a}
\end{figure}

We observe that, despite some fluctuations due to external factors (e.g.,
Titan's shared filesystem and the shared database used by the NGE), the average
execution time of AthenaMP  ranges between 4200 and 4800 seconds.  We  also
observe that in all the cases the gap between AthenaMP execution times and the
pilot durations is minimal, although it slightly increases with the pilot size.
We  notice that NGE's overhead does grow linearly with the number of units.

\subsubsection{Weak scalability with multiple generation }

The NGE provides an important new capability of submitting multiple generations
of AthenaMP to the same pilot. In order to investigate the cost of doing so, we
performed a variant of the weak scalability experiments. This stresses the
pilot's components, as new tasks are scheduled for execution on the Agent while
other tasks are still running.

In these experiments, we run five AthenaMP instances per node.  As these
experiments are designed to investigate the overhead generated by the scheduling
and bootstrap of AthenaMP instances, we reduced the number of events simulated
by each AthenaMP task to sixteen in such a way that the running time of each
AthenaMP is, on average, $\sim 1200$ seconds. This experiment design choice does
not affect the  objectives or accuracy of the experiments, but allows us to
scale experiments to large node counts by being conservative with allocation.

We ran pilots with 256, 512, 1024 and 2048 worker nodes and 3\mtnote{2?} hours
walltime. Fig.~\ref{fig:weakScal2a} depicts the average pilot duration, the
average execution time of five sequential generations of AthenaMP, and the
corresponding overhead. We observe that the difference between the two durations
is more marked than in the previous experiments. Despite this, we notice that
the growth of the overhead is consistent with the increment of the number of
tasks per node for pilots with 256, 512 and 1024 worker nodes, and less than
linear for the pilot with 2048 worker nodes.

\begin{figure}[!t]
        \includegraphics[height=4.5cm,width=\columnwidth]{weak2.pdf}
    \caption{Weak scalability with multiple generations: average pilot
    duration, average duration of sequential AthenaMP executions, and
    pilot's overhead for pilot with 256, 512, 1024 and 2048 nodes.}
\label{fig:weakScal2a}
\end{figure}

%\vspace{-0.2in}

\subsubsection{Strong scalability}

The last experiments  study strong scalability by running the same number of
tasks for different pilot sizes. We used 2048 AthenaMP instances and  pilots
with 256, 512, 1024 and 2048 nodes. Thus, the number of AthenaMP generations is
equal to eight times the size of the smallest pilot and corresponds to the size
of the largest pilot. As a consequence, the number of consecutive generations of
AthenaMP decreases with the pilot size by generating different dynamics within
the pilots. These experiments are designed to investigate whether pilot overhead
is affected by the degree of concurrency within the pilot and/or the number of
tasks. Each AthenaMP instance simulates sixteen events as in the previous
experiment.

Fig.~\ref{fig:strongScala}  shows the average pilot duration and the average
execution time of possibly sequential AthenaMP instances.  We  notice that the
difference between the pilot duration and the AthenaMP execution times is almost
constant for all the pilot sizes, although the overall duration of the pilot
decreases linearly with the pilot size.

\begin{figure}[!t]
        \includegraphics[height=4.5cm,width=\columnwidth]{strong.pdf}
    \caption{Strong scalability:  average pilot duration, average duration of
    sequential AthenaMP executions, and pilot's overhead for pilots with 256, 512, 1024 and 2048 nodes.}
\label{fig:strongScala}
\end{figure}

% ----------------------------------------------------------------------------
% VII - RELATED WORK
% ----------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}

Several pilot-enabled WMS were developed for the LHC experiments:
AliEn~\cite{Bagnasco2010} for ALICE; DIRAC~\cite{Paterson2010} for LHCb;
GlideinWMS~\cite{sfiligoi2008glideinwms} for CMS; and
PanDA~\cite{maeno2014evolution} for ATLAS. These systems implement similar
design and architectural principles: centralization of task and resource
management, and of monitoring and accounting; distribution of task execution
across multiple sites; unification of the application interface; hiding of
resource heterogeneity; and collection of static and sometimes dynamic
information about resources.

AliEn, DIRAC, GlideinWMS and PanDA all share a similar design with two types of
components: the management ones facing the application layer and centralizing
the capabilities required to acquire tasks' descriptions and matching them to
resource capabilities; and resource components used to acquire compute and data
resources and information about their capabilities. Architecturally, the
management components include one or more queue and a scheduler that coordinates
with the resource modules via push/pull protocols. All resource components
include middleware-specific APIs to request for resources, and a pilot capable
of pulling tasks from the management modules and executing them on its
resources.

AliEn, DIRAC, GlideinWMS and PanDA also have similar implementations. These WMS
were initially implemented to use Grid resources, using one or more components
to the Condor software ecosystem~\cite{thain2005distributed} and, as with
GlideinWMS, contributing to its development. Accordingly, all LHC WMS
implemented Grid-like authentication and authorization systems and adopted a
computational model based on distributing a large amount of single/few-cores
tasks across hundreds of sites\mtnote{Is this true?}.

All the experiments at LHC produces and process large amount of data both from
actual collisions in the accelerator and from their simulations. Dedicated,
multi-tiered data systems have been built to store, replicate, and distributed
these data. All LHC WMS interface with these systems to move data to the sites
where related compute tasks are executed or to schedule compute tasks where
(large amount of) data are already stored.

% ----------------------------------------------------------------------------
% VIII - CONCLUSION
% ----------------------------------------------------------------------------
%\vspace{-0.1in}
\section{Conclusion}
\label{sec:conclusion}

The deployment of PanDA Broker on Titan enabled distributed computing on a
leadership-class HPC machine at unprecedented scale. In the past 13 months,
PanDA WMS has consumed almost 52M core-hours on Titan, simulating 3.5\% of the
total number of detector events of the ATLAS production Monte Carlo workflow. We
described the implementation and execution process of PanDA WMS
(\S\ref{sec:panda_overview}) and PanDA Broker (\S\ref{sec:panda_titan}), showing
how they support and enable distributed computing at this scale on Titan, a
leadership-class HPC machine managed by OCLF at ORNL.

We characterized the state of practice by evaluating the efficiency, scalability
and reliability of both PanDA Broker and AthenaMP as deployed on Titan
(\S\ref{sec:panda_titan}). Our characterization highlighted the strengths
and limitations of the current design and implementation: PanDA Brokers enable
the sustained execution of millions of simulations per week but further work is
required to optimize its efficiency and reliability (\S\ref{ssec:broker_titan}).
PanDA Brokers support the concurrent execution of multiple AthenaMP instances,
enabling each AthenaMP to perform the concurrent execution of up to 16 Geant4
simulators. Nonetheless, our characterization showed how improving I/O
performance could reduce overheads (\S\ref{ssec:athenamp_titan}), increasing the
overall utilization of Titan's backfill availability.

More generally, this paper highlights the fundamental role of WMS for
experimental science. HEP was amongst the first, if not the very first
experimental community to realize the importance of using WMS to manage their
computational campaign(s). As computing becomes increasingly critical for a
range of experiments, the impact of WMS on HEP foreshadows the importance of
WMS for other experiments (such as SKA, LSST etc.). % and their computational
campaigns. These experiments will have their own workload characteristics,
resources types and federation constraints, as well metrics of performance.
The design and experience captured in this paper will prove invaluable for
designing WMS to support computational campaigns and will provide a baseline
to evaluate the relative merits of different approaches.

The 52M core hours used by ATLAS, via PanDA, is over 2\% of the total
utilization on Titan over the same period, bringing the time-averaged
utilization of Titan to be consistently upwards of 90\%. Given that the
average average utilization of most other leadership class machines is less
(e.g., NSF's flagship Blue Waters the average utilization fluctuates between
60-80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
approaches elsewhere. These unprecedented efficiency gains aside, this work is
just a starting point towards more effective operational models for future
leadership and online analytical platforms~\cite{foap-url}. These platforms
will have to support ever increasing complex workloads with varying models for
dynamic resource federation. Recent efforts with Google Cloud for the CMS
experiment -- HEPCloud~\cite{hepcloud,googlehep} represent similar yet
different approaches to dynamic resource federation. As the type,
heterogeneity and complexity of future platforms  (e.g., Exascale, Commercial
Clouds) for science increases, there will be a greater need and emphasis on
abstractions to ensure WMS can both utilize as well as support evolving
platforms. Our experience demonstrated the advantages arising from an
abstractions based approach \textendash{} in particular scalable implementations
of pilot abstraction.

% ----------------------------------------------------------------------------
% ACKNOWLEDGEMENTS
% ----------------------------------------------------------------------------
\section*{Acknowledgements}
\label{sec:ack}

This research used resources of the Oak Ridge Leadership Computing Facility,
supported by the Office of Science of the Department of Energy under Contract
DE-AC05-00OR22725.

% ----------------------------------------------------------------------------
% REFERENCES
% ----------------------------------------------------------------------------
\bibliographystyle{IEEEtran}
\bibliography{bibliography}

% that's all folks
\end{document}

### 2017-10-28 arxiv/v2/Experiments.tex
Update arxiv version

--- 2017-09-10 Experiments.tex
@@
-\ifreview
-Comments to address in this section:
-\begin{enumerate}
-    \color{red} 
-    \item REVIEWER 3: The scalability benchmarks in section "V.B Experiments"
-    should be better explained: what *extacly* is being kept constant in each
-    experiment, and what is being varied and over what range. Also the
-    captions in the figures do not help much.
-\end{enumerate}
-\fi
-
 % ---------------------------------------------------------------------------
@@
 collecting timestamps at all stages of the execution. Experiments were
-performed by submitting pilots to Titan's batch queue. The turnaround
-time of an individual run is determined by queue waiting times. Since we are
+performed by submitting pilots to Titan's batch queue. The turnaround time of
+an individual run is determined by queue waiting times. Since we are
 interested only in the performances of the NGE, we removed queue time from
@@
 Tasks do not wait within the Agent's queue since one node is available to
-each AthenaMP instance. Task execution  overheads result primarily from
-three factors: (i) the initial bootstrapping of the pilot on the nodes; (ii)
-the UnitManager's dispatching of units (tasks) to the agent; and (iii) time
-for the Agent to bootstrap all the tasks on the nodes.
+each AthenaMP instance. Task execution  overheads result primarily from three
+factors: (i) the initial bootstrapping of the pilot on the nodes; (ii) the
+UnitManager's dispatching of units (tasks) to the agent; and (iii) time for
+the Agent to bootstrap all the tasks on the nodes.
 
@@
 \begin{figure}[!t]
-    \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
+    \includegraphics[height=4.5cm,width=\columnwidth]{weak1.pdf}
    	\vspace{-0.3in}
@@
 AthenaMP task was reduced to sixteen such that the running time of each
-AthenaMP was \(\approx\)1,200 seconds on average. This experiment design choice
-does not affect the objectives or accuracy of the experiments, but allows us
-to scale experiments to large node counts by conserving project allocation.
+AthenaMP was \(\approx\)1,200 seconds on average. This experiment design
+choice does not affect the objectives or accuracy of the experiments, but
+allows us to scale experiments to large node counts by conserving project
+allocation.
 
@@
 Fig.~\ref{fig:weakScal2a} depicts the average pilot duration, the average
-execution time of five generations of AthenaMP, and the
-corresponding overhead. The difference between the two durations is more
-marked than in the previous experiments. Despite this, we notice that the
-growth of the overhead is consistent with the increment of the number of tasks
-per node for pilots with 256, 512 and 1024 worker nodes, and less than linear
-for the pilot with 2048 worker nodes.
+execution time of five generations of AthenaMP, and the corresponding
+overhead. The difference between the two durations is more marked than in the
+previous experiments. Despite this, we notice that the growth of the overhead
+is consistent with the increment of the number of tasks per node for pilots
+with 256, 512 and 1024 worker nodes, and less than linear for the pilot with
+2048 worker nodes.
 
 \begin{figure}[!t]
-    \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak2.pdf}
+    \includegraphics[height=4.5cm,width=\columnwidth]{weak2.pdf}
     \vspace{-0.3in}
@@
 \begin{figure}[!t]
-    \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/strong.pdf}
+    \includegraphics[height=4.5cm,width=\columnwidth]{strong.pdf}
     \vspace{-0.3in}

### 2017-10-28 arxiv/v2/PandaEvolution.tex
Update arxiv version

--- 2017-09-08 PandaEvolution.tex
@@
-\ifreview
-Comments to address in this section:
-\begin{enumerate}
-	\color{red} 
-	\item REVIEWER 2: the technical discussion of TITAN and WLCG differences
-	is abruptly stopped by the description of Monte Carlo event generation
-	and detector simulation (3A).\mtnote{I read the section twice and I did
-	not see such an abrupt transition.}
-	\item REVIEWER 2: The single listing (3B) of similar projects within the
-	collaboration does not detail whether this is derived or separate
-	work.\mtnote{Added a sentence to clarify that PanDA Broker operates
-	within the existing ATLAS software infrastructure. Specified that the
-	ATLAS team has ported PanDA Broker to other supercomputers.}
-	\item REVIEWER 2: How does the TITAN environment prevent usage of DTNs as
-	thin proxies? Is connectivity possible from compute node to DTNs via MPI,
-	message queues, sockets or other means? Are the restrictions a
-	shortcomming of TITAN, or the majority of HPC environments? Given the
-	proclaimed importance of HPC resources, a generic evaluation seems
-	needed.\mtnote{Added a sentence explaining that DTNs are not part of
-	Titan's compute nodes.}
-	\item REVIEWER 2: Why was the PanDA Broker designed orthogonal to the
-	pilot model, performing both resource acquisition and workload
-	execution?\mtnote{We explain that PanDA Broker is a prototype and now we
-	are developing NGE.} How do pilots conflict with backfill slots, when the
-	former are designed to acquire resources independent of workload
-	requirements?\mtnote{AFAIK, we never state that pilot conflict with
-	backfill. We argue that the latter mitigates the absence of the former.}
-	Why does each Broker instance query and use backfill separately? Given
-	that this seems to be a major bottleneck, the choice of such a design
-	begs for motivation.\mtnote{Extended a paragraph to explain that PanDA
-	Broker supported rapid deployment and incremental optimization while
-	incurring in functional limitations.}
-	\item REVIEWER 2: Some facts about WLCG seem outdated:
-	\begin{itemize}
-		\item The number of 30,000 unused TITAN cores (§IIIB) is equated to
-		roughly 10\% of the WLCG. However, WLCG currently has 591,282 cores
-		[REBUSCPU].
-		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{Done.}
-		\item WLCG sites are claimed to provide only 8 cores per worker node
-		(§IIIB). WLCG has a significant number of sites providing 16 or more
-		cores per worker node [REBUSCPU].
-		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{This
-		would require fairly deep rewriting on Section IVB. I edited the
-		sentence referring to 8 cores eliminating the reference to compute
-		nodes. Less than ideal but I did not find a better solution. Any
-		suggestion?}
-	\end{itemize}
-	\item REVIEWER 2: Schematic figures (Fig 1, Fig 2) are overly complex,
-	while their captions are not sufficient to understand them. The numbering
-	is only partially explained in the text.\mtnote{Added a reference to
-	where the numbering of the figures are explained in the text. I
-	double-checked and the numbering of both figure is fully explained in the
-	text.}
-\end{enumerate}
-\fi
-
 The upcoming LHC Run 3 will require more resources than the Worldwide LHC
@@
 600,000 cores at more than 100 Grid sites, with an aggregated performance of
-8 petaFLOPS\@. This capacity will be sufficient for the planned analysis
-and data processing, but it will be insufficient for the Monte Carlo
-production workflow and any extra activity. To alleviate these challenges,
-ATLAS is expanding the current computing model to include additional
-resources such as the opportunistic use of supercomputers.
+8 petaFLOPS\@. This capacity will be sufficient for the planned analysis and
+data processing, but it will be insufficient for the Monte Carlo production
+workflow and any extra activity. To alleviate these challenges, ATLAS is
+expanding the current computing model to include additional resources such as
+the opportunistic use of supercomputers.
 
@@
 nodes are part of the OLCF infrastructure and can access Titan without RSA
-SecureID authentication. It should be noted that DTNs are not part of Titan's
-worker nodes and, therefore, are not used to execute Titan's jobs.
+SecureID authentication. DTNs are not part of Titan's worker nodes and,
+therefore, are not used to execute Titan's jobs.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
+    \includegraphics[width=\columnwidth]{panda_broker_architecture.pdf}
     \vspace{-0.3in}

### 2017-10-28 arxiv/v2/abstract.tex
Update arxiv version

--- 2017-07-07 abstract.tex
@@
 resources will be insufficient to meet projected future demands. This paper
-is a case study of how the ATLAS experiment has embraced Titan -- a DOE
+is a case study of how the ATLAS experiment has embraced Titan---a DOE
 leadership facility in conjunction with traditional distributed high-

### 2017-10-28 arxiv/v2/conclusion.tex
Update arxiv version

--- 2017-09-10 conclusion.tex
@@
-% \ifreview
-% Comments to address in this section:
-% \begin{itemize}
-% 	\color{red} 
-% 	% \item REVIEWER 2: How does Summit, the TITAN successor, compare to the
-% 	% conclusions? The end of the TITAN life-cycle basically invalidates the
-% 	% applicability of all conclusions if this generalisation is missing.
-% 	% \item REVIEWER 3: This statement in the "conclusions" section is quite
-% 	% exaggerated: "The experience captured in this paper will prove invaluable
-% 	% for designing WMS for computational campaigns and will provide a baseline
-% 	% to evaluate the relative merits of different approaches." WMS systems for
-% 	% LHC experiments have been developed and run in production for many years;
-% 	% while this paper adds interesting experience and data points, it's
-% 	% unlikely that this will be "invaluable" for new designs and a cornerstone
-% 	% for their evaluation. \jhanote{I think this comment can also be ignored.}
-% \end{itemize}
-% \fi
-
 The deployment of PanDA Broker on Titan enabled distributed computing on a
@@
 
-% More generally, this paper highlights the fundamental role of WMS for
-% experimental science.
-
 HEP was amongst the first, if not the first experimental community to realize
@@
 characteristics, resources types and federation constraints, as well metrics
-of performance. The experience captured in this paper should be useful
-for designing WMS for computational campaigns and will provide a baseline to
+of performance. The experience captured in this paper should be useful for
+designing WMS for computational campaigns and will provide a baseline to
 evaluate the relative merits of different approaches.
@@
 
-{\bf \footnotesize This work is funded by DOE XXXXXX \par}
-
-% The development of NGE is looking forward to deal with these challenges by
-% providing an abstraction layer for HPC resources together with a set of
-% APIs for the coordination and fast deployment of heterogeneous workloads.
-% In this paper, we provided the results of a preliminary testing of the NGE.
-% The testing pointed out the robustness of NGE at handling an high number of
-% tasks and worker nodes by mantaining an overhead that grows linearly with
-% the number of tasks.
-
-% Recent efforts with Google Cloud for the CMS experiment --
-% HEPCloud~\cite{hepcloud,googlehep} represent similar yet different
-% approaches to dynamic resource federation. As the type, heterogeneity and
-% complexity of future platforms  (e.g., Exascale, Commercial Clouds) for
-% science increases, there will be a greater need and emphasis on
-% abstractions to ensure WMS can both utilize as well as support evolving
-% platforms. Our experience demonstrated the advantages arising from an
-% abstractions based approach \textendash{} in particular scalable
-% implementations of pilot abstraction.
+{\bf \footnotesize Acknowledgements: This work is funded by Award number
+DE-SC0016280 from the Office of Advanced Scientific Computing Research within
+the Department of Energy. This research used resources of the Oak Ridge
+Leadership Computing Facility at the Oak Ridge National Laboratory, which is
+supported by the Office of Science of the U.S. Department of Energy under
+Contract No. DE-AC05-00OR22725. \par}

### 2017-10-28 arxiv/v2/introduction.tex
Update arxiv version

--- 2017-09-10 introduction.tex
@@
-\ifreview
-Comments to address in this section:
-\begin{enumerate}
-	\color{red} 
-	\item REVIEWER 1: The introduction is overly verbose and prosaic. The
-	praise for LHC collaborations is off-putting, especially as it ignores
-	other work. For example, claiming "CMS and ATLAS experiments utilize
-	arguably the largest production grade distributed computing solutions"
-	disregards the obvious commercial contenders. By blatantly excluding
-	obvious comparisons, it raises the question how much other work is
-	ignored. \jhanote{Added "academic" after arguably. Any more?}
-\end{enumerate}
-\fi
-
 The Large Hadron Collider (LHC) was created to explore the fundamental
@@
 utilize, arguably the largest academic production grade distributed computing
-solutions, and have symbolized the paradigm of {\it high-throughput
-computing} (HTC), i.e., the effective execution of many independent tasks.
+solutions, and have symbolized the paradigm of high-throughput computing
+(HTC), i.e., the effective execution of many independent tasks.
 
-In spite of the impressive scale of the ATLAS distributed computing system --
-in the number of tasks executed, the number of core hours utilized, and the
-number of distributed sites utilized,  demand for computing systems will soon
-significantly outstrip current and projected supply.   The data volumes that
-will need analyzing in LHC-Run 3 ($\approx$2022) and the high-luminosity
-era (Run 4) will increase by factors of 10--100 compared to the current phase
-(Run 2). There are multiple levels at which this problem needs to be
-addressed: the utilization of emerging parallel architectures (e.g.,
-platforms); algorithmic and advances in analytical methods (e.g., use of
-Machine Learning); and the ability to exploit different platforms (e.g.,
+In spite of the impressive scale of the ATLAS distributed computing
+system---in the number of tasks executed, the number of core hours utilized,
+and the number of distributed sites utilized,  demand for computing systems
+will soon significantly outstrip current and projected supply.   The data
+volumes that will need analyzing in LHC-Run 3 (\(\approx\)2022) and the
+high-luminosity era (Run 4) will increase by factors of 10--100 compared to
+the current phase (Run 2). There are multiple levels at which this problem
+needs to be addressed: the utilization of emerging parallel architectures
+(e.g., platforms); algorithmic and advances in analytical methods (e.g., use
+of Machine Learning); and the ability to exploit different platforms (e.g.,
 clouds and supercomputers).
-
-% As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity
-% era (Run 4), it is anticipated that the data volumes that will need
-% analyzing will increase by factors of 10-100 compared to the current phase
-% (Run 2). Data will be larger in volume but will also require more
-% sophisticated computational processing.
 

### 2017-10-28 arxiv/v2/nge.tex
Update arxiv version

--- 2017-09-08 nge.tex
@@
-\ifreview
-Comments to address in this section:
-\begin{itemize}
-	\color{red} 
-	\item REVIEWER 3: The first paragraphs of section V. are basically a
-	rediscovery of the advantages of pilot jobs for WLCG task scheduling.
-	Since all those pain points were well-known in the WLCG community before
-	the project started, it should be explained why a pilot job approach was
-	not taken in the first place.\mtnote{Added an opening paragraph recalling
-	the explanation offered in \S\ref{sec:panda_titan}}.
-	\item REVIEWER 3: Figure 6. refers to a "User Workstation" as the running
-	place for the Pilot and Unit Managers, which is never mentioned in the
-	text. (I got the impression that the NGE would run on a DTN node; if this
-	is not the case and separate workstations are being used for the testing
-	phase, this should be made clear.)\mtnote{Fixed.}
-\end{itemize}
-\fi
-
 As explained in \S\ref{sec:panda_titan}, PanDA Broker was designed to
@@
 % ---------------------------------------------------------------------------
-\subsection{Implementation}
-\label{sec:arch}
+\subsection{Implementation}\label{sec:arch}
 
@@
   \centering
-	\includegraphics[width=\columnwidth]{figures/rp_architecture_compact_atlaswms_paper.pdf}
+	\includegraphics[width=\columnwidth]{rp_architecture_compact_atlaswms_paper.pdf}
 	\vspace{-0.3in}
@@
 arrow). Once scheduled, the Pilot Agent is bootstrapped on a MOM node and the
-Agent's Executors on the worker nodes.
+Agent's Executors on worker nodes.
 

### 2017-10-28 arxiv/v2/overall.tex
Update arxiv version

--- 2017-09-10 overall.tex
@@
-\ifreview
-Comments to address in this section:
-\begin{enumerate}
-    \color{red}
-
-    \item REVIEWER 2: What are the reasons for the low efficiency of PanDA
-    Broker? Why are only 30\% of backfill slots useable, if they are basically
-    uncontested? \jhanote{MT to address this via a few sentences in Section
-    IV.}
-
-    % \item REVIEWER 2: Why does Fig 3 show an increase of absolute and
-    % relative efficiency in Aug-16, when the text claims this in Sep-16 due to
-    % more PanDA Brokers?\mtnote{Fixed.}
-
-    % \item REVIEWER 2: Why are only 20 PanDA Brokers deployed if this is a
-    % limiting factor? Why is it questionable that more PanDA Brokers would
-    % increase efficiency? If only 5/6 of resources below the mean are
-    % unusable, this implies an efficiency of almost 60\% being
-    % reachable.\mtnote{From our text: ``further research is ongoing to
-    % understand whether an even greater number of brokers would yield even
-    % greater efficiency. This research focuses on evaluating the overheads of
-    % input/output files staging, including its impact on DTNs, and on an
-    % alternative design of PanDA Broker that enables the concurrent submission
-    % of multiple MPI scripts''. I believe this is an answer good enough.}
-
-    % \item REVIEWER 2: Given an expected increase in resource need of 10-100,
-    % how is the capacity of 4\% WLCG significant? How does this relate to the
-    % manpower required to develop and maintain a custom solution per HPC site?
-    % How many TITAN resources are actually available to HEP, given the use by
-    % other domains? The magnitude of capacities presented as available is in
-    % major contrast to the requirements sketched in the introduction.
-
-    % \item REVIEWER 2: How does the poor metadata read performance of
-    % 1500-6300s (25\%-100\% of minimum walltime!) affect efficiency? Is it
-    % factored into the choice of walltime requests?\mtnote{Thanks to Sergey's
-    % input I corrected the figure previously reported.} Is it applicable only
-    % to TITAN, to HPC in general, or even to regular WLCG sites?\mtnote{This
-    % is clearly out of scope for this paper.} Are there solutions at
-    % application, e.g. the use of packaged python scripts and
-    % libraries?\mtnote{Added the following: ``While these figures are specific
-    % to OLCF and TItan, reengineering on AthenaMP could further improve
-    % startup time on every platform.''} The information presented is
-    % insufficient to estimate whether this issue applies to other environments
-    % and usecases as well, or is simply due to suboptimal
-    % implementation/configuration.
-
-    % \item REVIEWER 3: The HEP-SPEC06 benchmark, which is used as a
-    % ``computing power'' measure in LCG is mostly integer-artihmetic based: FP
-    % operations accounts for \~10\% of the average workload (see [1], section
-    % ``Why do you use `all\_cpp' instead of SPECint or SPECfp?'') While this
-    % figure certainly comes from averaging over many different workloads
-    % across all experiments, it seems a bit surprising that FP scheduling
-    % contention is responsible for 50\% decrease in compute efficiency on
-    % Titan when compared to other WLCG sites (paper page 6).\mtnote{Thanks to
-    % Sergey's input I downplayed the role played by FP and up-played the role
-    % played by new CPUs.}
-\end{enumerate}
-\fi
-
-Currently 20 instances of the PanDA Broker are deployed on 4 DTNs, with
-5 instances per DTN\@. Each broker submits and manages the execution of 15 to
+Currently 20 instances of the PanDA Broker are deployed on 4 DTNs, with 5
+instances per DTN\@. Each broker submits and manages the execution of 15 to
 300 jobs, one job for each Titan worker node, and a theoretical maximum
@@
 % ---------------------------------------------------------------------------
-\subsection{Characterizing the PanDA Broker on Titan}\label{ssec:broker_titan}
+\subsection{Characterizing the PanDA Broker on
+Titan}\label{ssec:broker_titan}
 
@@
 \begin{figure}[!t]
-    \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.pdf}
+    \includegraphics[clip,width=\columnwidth]{backfill_consumption.pdf}
     \vspace{-0.3in}
@@
 
-PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels) is
-defined as the fraction (or percentage) of core-hours utilized by the PanDA
-Brokers of Titan’s total backfill availability during the experiment time
-window. The  average efficiency was 18\%, with a minimum efficiency
-of 7.8\% (May 2016) and a maximum efficiency of 30.9\% (Feb. 2017, excluding
-the preliminary results of March). The total backfill availability was
-$\approx$ 21.5M in April 2016, and 17.6M in February 2017. This suggests that
-the efficiency is invariant of total backfill availability.
-
-% This suggests that the differences
-% in efficiency did not depend on a comparable difference in the total backfill
-% availability.
-
-% \jhanote{change last sentence to: ... efficiency does not depend on total
-% backfill availability." ??}
-% \jhanote{shouldn't we be consistent in the months used to quote efficiency and
-% utilization? April 2016 and Feb 2017 vs May 2016 and Feb 2017??}
+PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels)
+is defined as the fraction (or percentage) of core-hours utilized by the
+PanDA Brokers of Titan’s total backfill availability during the experiment
+time window. The  average efficiency was 18\%, with a minimum efficiency of
+7.8\% (May 2016) and a maximum efficiency of 30.9\% (Feb. 2017, excluding the
+preliminary results of March). The total backfill availability was
+\(\approx\)21.5M in April 2016, and 17.6M in February 2017. This suggests
+that the efficiency is invariant of total backfill availability.
 
@@
 \begin{figure}[!t]
-    \includegraphics[clip,width=\columnwidth]{figures/titan_backfill_avail.png}
-    \vspace{-0.1in}
+    \includegraphics[clip,width=\columnwidth]{titan_backfill_avail.png}
+    \vspace{-0.3in}
     \caption{62555 measures of Backfill availability on Titan during the
@@
 \begin{figure}[htp]
-    \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison_vsquashed.pdf}
+    \includegraphics[clip,width=\columnwidth]{tx8_tx16_comparison_vsquashed.pdf}
     \vspace{-0.3in}
@@
 
-A performance analysis of Titan's AMD CPUs for detector simulations also helps
-to compare Titan and Grid site performance. Usually, Grid sites expose
+A performance analysis of Titan's AMD CPUs for detector simulations also
+helps to compare Titan and Grid site performance. Usually, Grid sites expose
 resources with heterogeneous CPU architectures and 8 (virtual) cores, while
-Titan's offer an homogeneous 16 cores architecture. We used the rate of events
-processes per minute as a measure of the efficiency of executing the same
-detector simulation on Titan or Grid sites. Comparisons of the efficiencies of
-Titan to the BNL and SIGNET Grid sites, normalized for 8 cores, show that the
-effective performance per-core at Titan is
-\(\approx\)0.57 event per minute, roughly 1/2 of BNL and  1/3 of SIGNET.
-
-%performances.
-
-% The CPUs at the Grid sites have one FP scheduler per core and are on
-% average newer than the CPU of Titan.
+Titan's offer an homogeneous 16 cores architecture. We used the rate of
+events processes per minute as a measure of the efficiency of executing the
+same detector simulation on Titan or Grid sites. Comparisons of the
+efficiencies of Titan to the BNL and SIGNET Grid sites, normalized for 8
+cores, show that the effective performance per-core at Titan is
+\(\approx\)0.57 event per minute, roughly 1/2 of BNL and  1/3 of SIGNET\@.
 
@@
 sites are due to the FP scheduler competition but mainly to the availability
-of newer processors. 
-The heterogeneity of the Grid sites' CPUs explain the higher performance
-variance we observed compared to the performance consistency we measured on
-Titan. This difference in performance is compensated by the amount of
-resources available on Titan (capable of executing up to 30\%/year of the
-ATLAS detector simulations) and by the lack of further resources available to
-WLCG\@. Also, it should be noted that Titan is at the end of its life-cycle
-and that Summit, Titan's successor, will offer cutting-edge performances.
+of newer processors. The heterogeneity of the Grid sites' CPUs explain the
+higher performance variance we observed compared to the performance
+consistency we measured on Titan. This difference in performance is
+compensated by the amount of resources available on Titan (capable of
+executing up to 30\%/year of the ATLAS detector simulations) and by the lack
+of further resources available to WLCG\@. Also, it should be noted that Titan
+is at the end of its life-cycle and that Summit, Titan's successor, will
+offer cutting-edge performances.
 
@@
 \begin{tabular}{lllllllll} & Num. Nodes & Duration (s) & Read (GB) & Written
- (GB) & GB Read/nodes & GB Written/nodes & \(open()\) & \(close()\) \\ Min & 1 &
- 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\ Max & 300 & 7,452 &
- 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\ Average & 35.66
- & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 & 34,155.74 \\
- Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376 & 231,346.55
- & 53,799.08
+ (GB) & GB Read/nodes & GB Written/nodes & \(open()\) & \(close()\) \\ Min &
+ 1 & 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\ Max & 300 &
+ 7,452 & 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\ Average
+ & 35.66 & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 &
+ 34,155.74 \\ Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376
+ & 231,346.55 & 53,799.08
 \end{tabular}
@@
 is 172,089,760 and the total number of file \(close()\) operations is
-40,132,992. The difference between these two values is under
-investigation: a possible explanation is that ATLAS jobs don't call a file
+40,132,992. The difference between these two values is under investigation: a
+possible explanation is that ATLAS jobs don't call a file
 \(close()\) operation for every file \(open()\) issued.
